### 2012-12-05 doc/description.tex
added description in latex

\documentclass[11pt]{article}
\usepackage{amssymb}

\begin{document}

\section{Application model}
The application is a set of instances which cooperates together. Each instance is defined according to a class, it means the class defines the structure, relationships of the instance\footnote{we use notation $instance \lhd class$ to describe the fact instance is defined according a class}.

The application model consists of model of classes -structural model and of instances - data model. The whole application is described as:

\begin{verbatim}
Application = < Class*, Instance*>
\end{verbatim}
The application is considered to be consistent if all instances are created according a class existing in the application.

$\forall i \in instances(Application), \exists c \in classes(Application): i \lhd c $

\subsection{Application structure}
The structure of the application is define by five concepts:
\begin{itemize}
	\item Class : Class represents a basic structural concept in the application model. It has a unique name, one or more properties and PrimitiveCollection and a class can be associated to other classes in the application.  
	\item Property : Property represents a feature of  a class which is can be represented as a primitive type. The property can be mandatory, can have a default value
	\item PrimitiveCollection : the PrimitiveCollection sort represents any collection (set, bag, etc.) of primitive types connected to a class. The collection has its default value and it can be mandatory.
	\item Association : the association represents a connection between two classes. It has a unique name and it contains a name of the class which is referenced by the association. The class which owns the association is consider to be a starting class of an association, referenced class is consider to be an ending class of an association. The cardinalities defines the multiplicity of the association.
	\item App-Type : represents primitive types in the application. There are usually defined types such as String, Integer, Boolean etc. in contrast there is only one type in our model, because we focus on structural and data changes and type casting operations are not important for us. The only App-Type type is called APP-STRING.
\end{itemize}

\begin{verbatim}
Class = <label, property*, primitiveCollection*, association*>
Property = <label, App-Type, defaultValue, mandatory, unique>
PrimitiveCollection = <label, App-Type, defaultValue, mandatory>
Association = <label, referencedClass, startCardinality, 
              endCardinality>
App-Type = APP-STRING 
unique = Bool
mandatory = Bool
referencesClass = label
startCardinality = NzNat
endCardinality = NzNat
Bool = true | false
NzNat = [0 - 9]*
\end{verbatim}

\end{document}

### 2012-12-05 doc/description.tex
description updated

@@
 Class = <label, property*, primitiveCollection*, association*>
-Property = <label, App-Type, defaultValue, mandatory, unique>
+Property = <label, App-Type, defaultValue, mandatory>
 PrimitiveCollection = <label, App-Type, defaultValue, mandatory>
@@
 App-Type = APP-STRING 
-unique = Bool
 mandatory = Bool

### 2012-12-05 doc/description.tex
property: mandatory changed to cardinality

@@
 Class = <label, property*, primitiveCollection*, association*>
-Property = <label, App-Type, defaultValue, mandatory>
+Property = <label, App-Type, defaultValue, cardinality>
 PrimitiveCollection = <label, App-Type, defaultValue, mandatory>
@@
 App-Type = APP-STRING 
-mandatory = Bool
+cardinality = NzNat
 referencesClass = label

### 2012-12-07 doc/description.tex
description of ORM operations started

--- 2012-12-05 doc/description.tex
@@
 \begin{itemize}
-	\item Class : Class represents a basic structural concept in the application model.
-It has a unique name, one or more properties and PrimitiveCollection and a class can be associated to other classes in the application.
+	\item Class : Class represents a basic structural concept in the application model. It has a unique name, one or more properties and a class can be associated to other classes in the application.  
 	\item Property : Property represents a feature of  a class which is can be represented as a primitive type. The property can be mandatory, can have a default value
-	\item PrimitiveCollection : the PrimitiveCollection sort represents any collection (set, bag, etc.) of primitive types connected to a class.
-The collection has its default value and it can be mandatory.
 	\item Association : the association represents a connection between two classes.
@@
 \begin{verbatim}
-Class = <label, property*, primitiveCollection*, association*>
-Property = <label, App-Type, defaultValue, mandatory>
-PrimitiveCollection = <label, App-Type, defaultValue, mandatory>
+Class = <label, property*, association*>
+Property = <label, App-Type, defaultValue, cardinality>
 Association = <label, referencedClass, startCardinality, 
@@
 
+\newpage
+\appendix
+\section{Object-Relational Mapping}
+There are defined functions for object-relational mapping used in the model. The purpose of a function is indicated in its name and should be obvious from its definition.
+
+\subsection{Mapping of Types}
+The mapping of types assumes a bijection between application types and database types, otherwise these has to be additional information for type mapping.
+We focus on structural and data change therefore we simplify types and its mapping.
+\begin{verbatim}
+op type2type : App-Type -> Db-Type .
+eq type2type(APP-STRING) = DB-STRING .
+\end{verbatim}
+
+\subsection{Class Mapping}
+We assume the primary key column is created automatically by database, therefore the primary key is always created with name Id and od NzNat type and properties with cardinality 0 or 1 are mapped into columns.
+The associations are ignored int this mapping as it is one of the partial mapping function used in eORM.
+
+\begin{verbatim}
+op class2table : Class -> Table .
+eq class2table(C) = table(name(C), primaryKey('Id), 
+                    props2cols(singlePropertyList(C)), []) .
+\end{verbatim}
+
+\subsection{Mapping of Properties}
+There are two types of properties - the first type called single properties has cardinality 0 or 1, the second one (called collections) has cardinality greater than 1.
+Each type has its own rules of mapping.
+
+\subsubsection{Mapping of Single Properties}
+\begin{verbatim}
+op props2cols : List{Property} -> List{Column} .
+	eq props2cols([]) = ([]).List{Column} .
+	eq props2cols(Props) = prop2col(head(Props)) props2cols(tail(Props)) .
+\end{verbatim}
+
 \end{document}

### 2012-12-07 doc/description.tex
uodates in description;

@@
 op type2type : App-Type -> Db-Type .
+
 eq type2type(APP-STRING) = DB-STRING .
@@
 op class2table : Class -> Table .
+
 eq class2table(C) = table(name(C), primaryKey('Id), 
@@
 
-\subsubsection{Mapping of Single Properties}
+\subsubsection{Mapping of Single Property}
+Each single property is mapped to a column, when the cardinality is equal to 1 the NOTNULL constraint is added to a column. 
 \begin{verbatim}
 op props2cols : List{Property} -> List{Column} .
+op props2cols : Property -> Column .
+
 eq props2cols([]) = ([]).List{Column} .
-	eq props2cols(Props) = prop2col(head(Props)) props2cols(tail(Props)) .
+ceq props2cols(Prop) = [] if cardinality(Prop) > 2 .
+ceq props2cols(Prop) = column(name(Prop), type2type(type(Prop)), NOTNULL)
+    if cardinality(Prop) == 1 .
+ceq props2cols(Prop) = column(name(Prop), type2type(type(Prop)), []) 
+   if cardinality(Prop) == 0 .
+eq props2cols(Props) = props2cols(head(Props)) props2cols(tail(Props)) .
+\end{verbatim}
+A property with cardinality greater than one has to be mapped int a table not a column, thus we have a special mapping function for collections.
+
+\subsubsection{Mapping of Collections}
+A property collection is mapped into a table which contains a foreign key. The key references the table representing the owning class.
+\begin{verbatim}
+op collections2tables : Qid List{Property} -> List{Table} .
+op collections2tables : Qid Property -> Table .
+
+eq collections2tables(Q, []) = ([]).List{Table} .
+eq collections2tables(Q, Prop) = table(name(Prop), primaryKey('Id), 
+   column('value, type2type(type(Prop)), [] ), foreignKey(Q, Q, NOTNULL)) .
+eq collections2tables(Q, Props) = collections2tables(Q, head(Props)) 
+   collections2tables(Q, tail(Props)) .
+\end{verbatim}
+
+\subsection{Mapping of Associations}
+\begin{verbatim}
+
 \end{verbatim}

### 2012-12-07 doc/description.tex
orm exmplained

@@
 \begin{verbatim}
-Application = < Class*, Instance*>
+op application : List{Class} List{Instance} -> Application [ctor] .
 \end{verbatim}
@@
 
+\section{Database model}
+The relational database consists of database schema which defines the structure of the database and data. The database is a pair od sets:
+\begin{verbatim}
+op database : List{Table} List{Row} -> Database [ctor] .
+\end{verbatim}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \newpage
@@
 \subsection{Mapping of Associations}
+A association is mapped into database as a foreign key in table representing associating class or an association can be mapped as a coupling table in case the association cardinality is M:N or 1:N . 
 \begin{verbatim}
+op associations2db : Qid List{Association} Database -> Database .
+op associations2db : Qid Association Database -> Database .
 
+eq associations2db(Q, [], D) = D .
+eq associations2db(Q, Assoc, EDb) = EDb .
+ceq associations2db(Q, Assoc, D) = addFK(foreignKey(name(Assoc), Q, NOTNULL), 
+	   referenceA(Assoc), D) if startCardinality(Assoc) < 2 . 
+eq associations2db(Q, Assoc, D) = addTable(table(name(Assoc), 
+   primaryKey('Id'), [], 
+   foreignKey(Q, Q, NOTNULL) 
+   foreignKey(name(Assoc), name(Assoc), NOTNULL)), D) .
+	eq associations2db(Q, Assoc, D) = associations2db(Q, tail(Assoc), 
+   associations2db(Q, head(Assoc), D)) .
 \end{verbatim}

### 2012-12-07 doc/description.tex
database modules refactoring

@@
 \end{verbatim}
+
+\subsection{Database Schema Model}
+The database schema model is defined by following concepts:
+\begin{itemize}
+	\item Table : table represents a basic concept of database schema.
+It has a unique name, one or more column and it can be related to other tables in the schema by foreign keys.
+Rows in the table represents stored data.
+	\item Column : columns define possible data values and types which can be part of a table record.
+	\item Foreign key :
+	\item Primary key :
+	\item Db-Type : represents primitive types in the database. There are usually defined types such as Char, Integer, Boolean etc. There is only one type defined in the model which is called DB-STRING.
+	\item Constraint : there are two types of constraints defined in the model.
+Both constraints are column constraints - first constraint defines unique records in a column, second constraint defines non-empty columns.
+\end{itemize}
 

### 2012-12-07 doc/description.tex
renaming in the model, extended description

@@
 \usepackage{amssymb}
+%\usepackage{semantic}
 
@@
 \subsection{Application structure}
-The structure of the application is define by five concepts:
+The structure of the application is defined by five concepts. Each concept is described and its constructor is introduced.
+
 \begin{itemize}
 	\item Class : Class represents a basic structural concept in the application model. It has a unique name, one or more properties and a class can be associated to other classes in the application.  
-	\item Property : Property represents a feature of  a class which is can be represented as a primitive type. The property can be mandatory, can have a default value
+	\begin{verbatim}
+op class : Qid List{Property} List{Association} -> Class [ctor] .
+\end{verbatim}
+	\item Property : Property represents a feature of  a class which is can be represented as a primitive type. The property can be mandatory, can have a default value.
+	\begin{verbatim}
+op property : Qid App-Type String Nat -> Property [ctor] .
+\end{verbatim}
 	\item Association : the association represents a connection between two classes.
@@
 The cardinalities defines the multiplicity of the association.
+	\begin{verbatim}
+op association : Qid Qid Nat Nat -> Association [ctor] .
+\end{verbatim}
 	\item App-Type : represents primitive types in the application.
@@
 The only App-Type type is called APP-STRING.
+	\begin{verbatim}
+op APP-STRING : -> App-Type [ctor] .
+	\end{verbatim}
 \end{itemize}
-
+An summary of application model:
 \begin{verbatim}
@@
               endCardinality>
+
 App-Type = APP-STRING 
@@
 NzNat = [0 - 9]*
+\end{verbatim}
+
+\subsection{Application Structure evolution}
+\subsubsection{Add Class}
+\begin{verbatim}
+op addClass : Class Application -> Application .
+ceq addClass(C, A) = errApplication("inconsistent", A) 
+   if containsName(classes(A), name(C))  
+   or not referencesExist(associations(C), classes(A)) .
+eq addClass(C, A) = application(classes(A) C, instances(A)) .
 \end{verbatim}
@@
 \begin{itemize}
-	\item Table : table represents a basic concept of database schema.
+	\item Table: table represents a basic concept of database schema.
 It has a unique name, one or more column and it can be related to other tables in the schema by foreign keys.
 Rows in the table represents stored data.
+	\begin{verbatim}
+op table : Qid PrimaryKey List{Column} List{ForeignKey} -> Table [ctor].
+	\end{verbatim}
 	\item Column : columns define possible data values and types which can be part of a table record.
-	\item Foreign key :
-	\item Primary key :
-	\item Db-Type : represents primitive types in the database. There are usually defined types such as Char, Integer, Boolean etc. There is only one type defined in the model which is called DB-STRING.
-	\item Constraint : there are two types of constraints defined in the model.
+	\begin{verbatim}
+op column : Qid Db-Type List{Constraint} -> Column [ctor] .
+	\end{verbatim}
+
+	\item Foreign key:
+	\begin{verbatim}
+op foreignKey : Qid Qid List{Constraint} -> ForeignKey [ctor] .
+	\end{verbatim}
+
+	\item Primary key:
+	\begin{verbatim}
+op primaryKey : Qid -> PrimaryKey [ctor] . 	
+	\end{verbatim}
+
+	\item Db-Type: represents primitive types in the database. There are usually defined types such as Char, Integer, Boolean etc. There is only one type defined in the model which is called DB-STRING.
+	\begin{verbatim}
+op DB-STRING : -> Db-Type [ctor] .
+	\end{verbatim}
+
+	\item Constraint: there are two types of constraints defined in the model.
 Both constraints are column constraints - first constraint defines unique records in a column, second constraint defines non-empty columns.
+	\begin{verbatim}
+op NOTNULL : -> Constraint [ctor] .
+op UNIQUE : -> Constraint [ctor] .
+	\end{verbatim}
 \end{itemize}
+A summary of database model:
+\begin{verbatim}
+Table = < name, PK, column*, foreignKey* >	
+Column = < name, type, constraint* >
+PrimaryKey = name
+ForeignKey = < name, referencedTableName, constraint* >
+Constraint = NOTNULL | UNIQUE 
+Db-Type = DB-STRING
+\end{verbatim}
+
+\section{Data Evolution in ORM System}
+The Object-Relational system consist of an application and a database. The system has to be in consistent state. The extended object-relational mapping helps to assure consistency of the system.
+\begin{verbatim}
+op system : Application Database -> OrmSystem [ctor] .
+\end{verbatim}
+%TODO define consistency
+Each eORM function applied on an ORM system produces a new generation of the system. If the new generation is consistent it can be used and evolved, otherwise the system cannot be evolved.
+
+\subsection{Creating Evolution}
+Creational operations does not have impact on stored data they change the structure of the application and database by adding new objects into model.
+
+\subsubsection{Add Class}
+The function addClass adds new class into an application and its equivalent into database. 
+\begin{verbatim}
+op addClass : Class OrmSystem -> OrmSystem .
+eq addClass(C, system(A, D)) = system(addClass(C, A), 
+   addClass2Db(C, D)) .
+
+op addClass2Db : Class Database -> Database .
+eq addClass2Db(C, D) = associations2db(name(C), 
+   associations(C), addTables(collections2tables(name(C), 
+   collectionPropertyList(C)), addTable(class2table(C), D))) .
+\end{verbatim}
+
+There is limitation connected with the implementation of platform. If there is a collection with the same name already existing table the evolution fails => rename a collection table
+
+\subsection{Removing Evolution}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \newpage
@@
 \section{Object-Relational Mapping}
-There are defined functions for object-relational mapping used in the model. The purpose of a function is indicated in its name and should be obvious from its definition.
+There are defined functions for object-relational mapping used in the model.
+The purpose of a function is indicated in its name and should be obvious from its definition.
+These functions do not serve for full object-relational mapping, their purpose is to help with data evolution and its propagation from application to database.
 

### 2012-12-07 doc/description.tex
addTables renamed to addTable

@@
 \usepackage{amssymb}
-%\usepackage{semantic}
+\usepackage{semantic}
 
@@
 
-\subsection{Application Structure evolution}
+\subsection{Application Structure Evolution}
+The evolution of the application persistent layer is called refactoring. In this section we provide an overview of refactorings allowed in our ORM system.
+
 \subsubsection{Add Class}
+$
+\inference[AddClass(C, A)=]{\forall c \in classes(A) : c.name != C.name}
+{application(classes(A) \cup C, instances(A))}
+$
+
 \begin{verbatim}
@@
 
-There is limitation connected with the implementation of platform. If there is a collection with the same name already existing table the evolution fails => rename a collection table
+There is limitation connected with the implementation of platform. If there is a collection with the same name already existing table the evolution fails.
 

### 2012-12-07 doc/description.tex
updated description addTable

@@
 
+\subsection{Database Schema Evolution}
+\subsubsection{AddTable}
+
+$
+\inference[AddTable(T, D)=]{\forall t \in tables(D) : t.name != T.name\\
+	\forall f \in T.foreignKeys, \exists t in T.tables : f.reference == t.name
+}{database(T \cup tables(D), data(D))}
+$
+
+\begin{verbatim}
+op addTable : Table Database -> Database .
+ceq addTable(T, D) = errDatabase("inconsistent", D) 
+   if containsName(tables(D), name(T)) 
+   or (foreignKeys(T) =/= [] 
+   and not referencesExist(foreignKeys(T), tables(D))) .
+eq addTable(T, D) = database(tables(D) T, data(D)) .
+\end{verbatim}
+
 \section{Data Evolution in ORM System}

### 2012-12-13 doc/description.tex
add/remove class, add property

--- 2012-12-07 doc/description.tex
@@
 \usepackage{amssymb}
-%\usepackage{semantic}
+\usepackage{amsmath}
+\usepackage{semantic}
+\usepackage{listings}
+\usepackage{url}
 
 \begin{document}
+\tableofcontents
 
-\section{Application model}
-The application is a set of instances which cooperates together.
-Each instance is defined according to a class, it means the class defines the structure, relationships of the instance\footnote{we use notation $instance \lhd class$ to describe the fact instance is defined according a class}.
+\newpage
+\section{Evolution of ORM System}
+The ORM system consist of application and database.
+The application and database has to be in consistent state according to a concrete ORM algorithm, otherwise the application cannot work properly.
+It means that during the evolution we have to respect a concrete ORM to produce consistent result.
+$$
+OrmSystem = (Application, Database)
+$$
+%TODO define consistency
 
-The application model consists of model of classes -structural model and of instances - data model. The whole application is described as:
+There are several way how the evolution of ORM system can be implemented.
+The way when the application is evolved, database scheme re-genareted by an ORM framework and data migrated manually is the most common nowadays.
+The goal of our project is to minimize the manual work in the process of data evolution.
+We have following options:
+\begin{itemize}
+	\item Backward round-trip engineering can be used.
+It means the application and database is evolved and then the difference in structure is used to define data migration.
+This approach is inappropriate for situations where there is stored data as a change in database schema can be connected with more possible changes of data.
+Hence an additional information from user will be needed to migrate data correctly.
+    \item Forward engineering can be used in two ways.
+First approach evolves application, re-generate a database schema and generate a data migration according to definition of application evolution.
+Second approach uses the definition of application evolution as a source for evolution of database schema and data migration.
+In contrast this approach preserves the original database - it only alters its structure, whereas the first approach creates a new database.
+\end{itemize}
+All approaches can be used and all of them have their pros and cons.
+We decide to use forward engineering approach as it is a natural way of evolution.
+Next we decide to generate both migrations (for database schema and data) from definition of application evolution.
+This allows us to keep history of all layers and to verify the evolution on multiple levels.
+The algorithm is illustrated in Figure \ref{fig:algorithm}.
 
-\begin{verbatim}
-op application : List{Class} List{Instance} -> Application [ctor] .
-\end{verbatim}
+The figure shows all important components of the domain.
+There are three static components (static because they do not define behavior, but state of the system): an application, a database schema and stored data.
+Then there are four dynamic components (they define transformations between components and transition between system states): an object-relational mapping is a transformation between application and database schema, then are two components for evolution of application and of database.
+(To simplify the model we will consider the database schema and data are evolved by one component).
+Last component is the most important one - it is component which maps the evolution from application level to relational database level.
+We call this component evolution object-relational mapping (eORM).
+
+There is a description of each component following. The description defines set of models of static components and dynamic components are defined as functions.
+%TODO: prerequisities - lables (if they are unique it is not necessary to check the name)
+%%%%%%%%%%%%%%%%%%%
+
+\section{Application Component}
+The application component represent an application created using object oriented language.
+It means the application is a set of instances which cooperates together.
+Each instance is defined according to a class - the class defines the structure, relationships of the instance.
+We use notation $instance \lhd class$ to describe the fact instance is defined according a class.
+
+The application model consists of model of classes - structural model and of instances - data model. The whole application is described as:
+
+$$
+Application = (Class*, Instance*)
+$$
+%TODO: are instances necessary on this level?
+
 The application is considered to be consistent if all instances are created according a class existing in the application. 
+$$
+\forall i \in instances(Application), \exists c \in classes(Application): i \lhd c 
+$$
+%TODO define consistency
 
-$\forall i \in instances(Application), \exists c \in classes(Application): i \lhd c $
-
-\subsection{Application structure}
+\subsection{Application Structure Definition}
 The structure of the application is defined by five concepts. Each concept is described and its constructor is introduced.
@@
 	\item Class : Class represents a basic structural concept in the application model. It has a unique name, one or more properties and a class can be associated to other classes in the application.  
-	\begin{verbatim}
-op class : Qid List{Property} List{Association} -> Class [ctor] .
-\end{verbatim}
+$$
+Class = (label, Property*, Association*)
+$$
 	\item Property : Property represents a feature of  a class which is can be represented as a primitive type. The property can be mandatory, can have a default value.
-	\begin{verbatim}
-op property : Qid App-Type String Nat -> Property [ctor] .
-\end{verbatim}
+$$
+Property = (label, App-Type, DefaultValue, Cardinality)
+$$
+%TODO owning class
 	\item Association : the association represents a connection between two classes.
@@
 The cardinalities defines the multiplicity of the association.
-	\begin{verbatim}
-op association : Qid Qid Nat Nat -> Association [ctor] .
-\end{verbatim}
+$$
+Association = (label, classRef, startCardinality, endCardinality)
+$$
 	\item App-Type : represents primitive types in the application.
@@
 The only App-Type type is called APP-STRING.
-	\begin{verbatim}
-op APP-STRING : -> App-Type [ctor] .
-	\end{verbatim}
+$$
+App-Type = APPSTRING
+$$
 \end{itemize}
 An summary of application model:
-\begin{verbatim}
-Class = <label, property*, association*>
-Property = <label, App-Type, defaultValue, cardinality>
-Association = <label, referencedClass, startCardinality, 
-              endCardinality>
+$$
+cardinality = NzNat
+$$
+$$
+referencesClass = label
+$$
+$$
+startCardinality = NzNat
+$$
+$$
+endCardinality = NzNat
+$$
+$$
+NzNat = [0 - 9]*
+$$
 
-App-Type = APP-STRING 
-cardinality = NzNat
-referencesClass = label
-startCardinality = NzNat
-endCardinality = NzNat
-Bool = true | false
-NzNat = [0 - 9]*
-\end{verbatim}
+\subsection{Application Evolution}
+\label{sec:appEvolution}
+\subsubsection{Application Creation}
+$$AddClass: Class \times Application \rightarrow Application $$
 
-\subsection{Application Structure evolution}
-\subsubsection{Add Class}
-\begin{verbatim}
-op addClass : Class Application -> Application .
-ceq addClass(C, A) = errApplication("inconsistent", A) 
-   if containsName(classes(A), name(C))  
-   or not referencesExist(associations(C), classes(A)) .
-eq addClass(C, A) = application(classes(A) C, instances(A)) .
-\end{verbatim}
+$
+AddClass(c, a) = \inference{\forall s \in classes(A) : name(s) \neq name(c)}
+{application(classes(a) \cup c, instances(a))}
+$
+
+$$AddProperty : Label \times Property \times Application \rightarrow Application $$
+
+$
+AddProperty(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \not \exist q \in properties(p) : name(q) \neq name(p)}{
+application(classes(a)) \wedge properties(c) \cup p }
+$
+
+\subsubsection{Application Deconstruction}
+$$RemoveClass: Label \times Application \rightarrow Application $$
+$
+RemoveClass(q, a) = \inference{\exists c \in classes(a) : name(c) = q \\
+\wedge not$ $instantiated(c) = \{ \} \wedge not$ $associated(c)}{applicattion(classes(a) \setminus c, instances(a)) }
+$
+
+$$RemoveProperty: Label \times Label \times Application \rightarrow Application $$
+
+$
+RemoveProperty(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{}
+$
+
+%%%%%%%%%%%%%%%%%%%
 
@@
 The relational database consists of database schema which defines the structure of the database and data. The database is a pair od sets:
-\begin{verbatim}
-op database : List{Table} List{Row} -> Database [ctor] .
-\end{verbatim}
+$$
+Database = (table*, data*)
+$$
 
@@
 Rows in the table represents stored data.
-	\begin{verbatim}
-op table : Qid PrimaryKey List{Column} List{ForeignKey} -> Table [ctor].
-	\end{verbatim}
+$$
+Table = (label, primaryKey, Column*, ForeignKey*)
+$$
 	\item Column : columns define possible data values and types which can be part of a table record.
-	\begin{verbatim}
-op column : Qid Db-Type List{Constraint} -> Column [ctor] .
-	\end{verbatim}
+$$
+Column = (label, Db-Type, Constraint*)
+$$
 
 	\item Foreign key:
-	\begin{verbatim}
-op foreignKey : Qid Qid List{Constraint} -> ForeignKey [ctor] .
-	\end{verbatim}
+$$
+ForeignKey = (label, tableRef, Constraint*)
+$$
 
 	\item Primary key:
-	\begin{verbatim}
-op primaryKey : Qid -> PrimaryKey [ctor] . 	
-	\end{verbatim}
+$$
+PrimaryKey =  (label) 	
+$$
 
 	\item Db-Type: represents primitive types in the database. There are usually defined types such as Char, Integer, Boolean etc. There is only one type defined in the model which is called DB-STRING.
-	\begin{verbatim}
-op DB-STRING : -> Db-Type [ctor] .
-	\end{verbatim}
+$$
+Db-Type = DBSTRING
+$$
 
@@
 Both constraints are column constraints - first constraint defines unique records in a column, second constraint defines non-empty columns.
-	\begin{verbatim}
-op NOTNULL : -> Constraint [ctor] .
-op UNIQUE : -> Constraint [ctor] .
-	\end{verbatim}
+$$
+Constraint = NOTNULL | UNIQUE 
+$$
 \end{itemize}
-A summary of database model:
-\begin{verbatim}
-Table = < name, PK, column*, foreignKey* >	
-Column = < name, type, constraint* >
-PrimaryKey = name
-ForeignKey = < name, referencedTableName, constraint* >
-Constraint = NOTNULL | UNIQUE 
-Db-Type = DB-STRING
-\end{verbatim}
 
-\section{Data Evolution in ORM System}
-The Object-Relational system consist of an application and a database. The system has to be in consistent state. The extended object-relational mapping helps to assure consistency of the system.
-\begin{verbatim}
-op system : Application Database -> OrmSystem [ctor] .
-\end{verbatim}
-%TODO define consistency
-Each eORM function applied on an ORM system produces a new generation of the system. If the new generation is consistent it can be used and evolved, otherwise the system cannot be evolved.
-
-\subsection{Creating Evolution}
-Creational operations does not have impact on stored data they change the structure of the application and database by adding new objects into model.
-
-\subsubsection{Add Class}
-The function addClass adds new class into an application and its equivalent into database. 
-\begin{verbatim}
-op addClass : Class OrmSystem -> OrmSystem .
-eq addClass(C, system(A, D)) = system(addClass(C, A), 
-   addClass2Db(C, D)) .
-
-op addClass2Db : Class Database -> Database .
-eq addClass2Db(C, D) = associations2db(name(C), 
-   associations(C), addTables(collections2tables(name(C), 
-   collectionPropertyList(C)), addTable(class2table(C), D))) .
-\end{verbatim}
-
-There is limitation connected with the implementation of platform. If there is a collection with the same name already existing table the evolution fails => rename a collection table
-
-\subsection{Removing Evolution}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newpage
-\appendix
 \section{Object-Relational Mapping}
@@
 
+%TODO exist \rho to execute mapping and \rho' (?) just returning a name of mapped object
+
 \subsection{Mapping of Types}
@@
 We focus on structural and data change therefore we simplify types and its mapping.
-\begin{verbatim}
-op type2type : App-Type -> Db-Type .
+$$
+\rho : AppType \rightarrow DbType
+$$
 
-eq type2type(APP-STRING) = DB-STRING .
-\end{verbatim}
+$$
+\rho(APPSTRING) = DBSTRING
+$$
 
-\subsection{Class Mapping}
+\subsection{Mapping of Classes}
 We assume the primary key column is created automatically by database, therefore the primary key is always created with name Id and od NzNat type and properties with cardinality 0 or 1 are mapped into columns.
@@
 
-\begin{verbatim}
-op class2table : Class -> Table .
+$$
+\rho : Class \rightarrow Table
+$$
 
-eq class2table(C) = table(name(C), primaryKey('Id), 
-                    props2cols(singlePropertyList(C)), []) .
-\end{verbatim}
+$$
+\rho(c) = table(name(c), primaryKey("Id"), \rho(properties(c), \{ \})
+$$
 
@@
 \subsubsection{Mapping of Single Property}
-Each single property is mapped to a column, when the cardinality is equal to 1 the NOTNULL constraint is added to a column. 
-\begin{verbatim}
-op props2cols : List{Property} -> List{Column} .
-op props2cols : Property -> Column .
+Each single property is mapped to a column, when the cardinality is equal to 1 the $NOTNULL$ constraint is added to a column.
+A property with cardinality greater than one has to be mapped int a table not a column, thus we have a special mapping case for collections.
+A collection is mapped into a table which contains a foreign key.
+The key references the table representing the owning class.
 
-eq props2cols([]) = ([]).List{Column} .
-ceq props2cols(Prop) = [] if cardinality(Prop) > 2 .
-ceq props2cols(Prop) = column(name(Prop), type2type(type(Prop)), NOTNULL)
-    if cardinality(Prop) == 1 .
-ceq props2cols(Prop) = column(name(Prop), type2type(type(Prop)), []) 
-   if cardinality(Prop) == 0 .
-eq props2cols(Props) = props2cols(head(Props)) props2cols(tail(Props)) .
-\end{verbatim}
-A property with cardinality greater than one has to be mapped int a table not a column, thus we have a special mapping function for collections.
+%TODO split \rho function?
 
-\subsubsection{Mapping of Collections}
-A property collection is mapped into a table which contains a foreign key. The key references the table representing the owning class.
-\begin{verbatim}
-op collections2tables : Qid List{Property} -> List{Table} .
-op collections2tables : Qid Property -> Table .
+$
+\rho : Property, Database \rightarrow Column 
+$
 
-eq collections2tables(Q, []) = ([]).List{Table} .
-eq collections2tables(Q, Prop) = table(name(Prop), primaryKey('Id), 
-   column('value, type2type(type(Prop)), [] ), foreignKey(Q, Q, NOTNULL)) .
-eq collections2tables(Q, Props) = collections2tables(Q, head(Props)) 
-   collections2tables(Q, tail(Props)) .
-\end{verbatim}
+$\rho(p, d) = \begin{cases}
+ \inference{cardinality(p) = 0 \\ \wedge \rho(owningClass(p) \in classes(d)}{columns(owninngClass(p)) \cup column(name(p), \rho(type(p)), \{ \})} \\ \\
+
+ \inference{cardinality(p) = 1 \\ \wedge \rho(owningClass(p) \in classes(d)}{columns(owninngClass(p)) \cup col(name(p), \rho(type(p)), NOTNULL)} \\ \\
+
+ \inference{cardinality(p) > 2 \\ \wedge \rho(owningClass(p) \in classes(d)}{\begin{gathered} tables(d) \cup table(name(p), pk("Id"), col("value", DBSTRING), \\ fk(name(p), \rho(owningClass(p))) \end{gathered}}
+ \end{cases}$
 
 \subsection{Mapping of Associations}
-A association is mapped into database as a foreign key in table representing associating class or an association can be mapped as a coupling table in case the association cardinality is M:N or 1:N . 
-\begin{verbatim}
-op associations2db : Qid List{Association} Database -> Database .
-op associations2db : Qid Association Database -> Database .
+A association is mapped into database as a foreign key in table representing associating class or an association can be mapped as a coupling table in case the association cardinality is M:N or 1:N.
 
-eq associations2db(Q, [], D) = D .
-eq associations2db(Q, Assoc, EDb) = EDb .
-ceq associations2db(Q, Assoc, D) = addFK(foreignKey(name(Assoc), Q, NOTNULL), 
-	   referenceA(Assoc), D) if startCardinality(Assoc) < 2 . 
-eq associations2db(Q, Assoc, D) = addTable(table(name(Assoc), 
-   primaryKey('Id'), [], 
-   foreignKey(Q, Q, NOTNULL) 
-   foreignKey(name(Assoc), name(Assoc), NOTNULL)), D) .
-	eq associations2db(Q, Assoc, D) = associations2db(Q, tail(Assoc), 
-   associations2db(Q, head(Assoc), D)) .
-\end{verbatim}
+$
+\rho(a,d) = \begin{cases}
+ \inference{startCardinality(a) = 0 \\ \wedge \rho(owningClass(p) \in classes(d)  \wedge \rho(reference(a)) \in tables(d)}{foreignKeys(\rho(assoc(a)) \cup foreignKey(name(a), owningClass(a),  \{ \}) } \\ \\
+
+ \inference{startCardinality(a) = 1 \\ \wedge \rho(owningClass(p) \in tables(d)  \wedge \rho(reference(a)) \in tables(d)}{\begin{gathered}
+ foreignKeys(\rho(assoc(a)) \cup foreignKey(name(a), \\ owningClass(a),  NOTNULL)
+\end{gathered}
+} \\ \\
+
+  \inference{startCardinality > 1 \\ \wedge \rho(owningClass(p)) \in tables(d) \wedge \rho(reference(a)) \in tables(d)}{
+  \begin{gathered}
+  tables(d) \cup table(name(a), pk("Id"), \{ \}, \\ foreignKey(owningClass(a),owningClass(a), NOTNULL) \\ foreignkey(reference(a), reference(a), NOTNULL)) 
+  \end{gathered}}
+
+ \end{cases}
+$
+
+%TODO FK nemusí mít jméno - odkazuje jen na tabulku a id
+
+\section{Data Evolution in ORM System}
+The Object-Relational system consist of an application and a database.
+The system has to be in consistent state, therefore the evolution of application has to be propagated correctly into database.
+Each evolution operation consist of two parts - applying evolution operation on application and on database.
+In general an evolutionary operation $X : OrmSystem \rightarrow OrmSystem $ can be defined as:
+
+$$
+\inference[X(s) = ]{}{system(X(application(s)), \rho_{X}(database(s))) }
+$$
+where the $\rho_X$ is the mapping of the evolution from application to database.
+The mapping of the evolution change is called extended Object-Relational Mapping (eORM).
+The eORM mapping helps to assure consistency of the system by transforming application evolution into database schema and data.
+Each eORM function applied on an ORM system produces a new generation of the system.
+If the function cannot be applied on a system the system cannot be evolved by the function.
+
+This section defines eORM for each trivial operation evolving application from \ref{sec:appEvolution}, moreover there are described more complex cases of evolution.
+
+\subsection{Creating of a System}
+Creational operations does not have impact on stored data they change the structure of the application and database by adding new objects into model.
+
+\subsubsection{Add Class}
+
+$
+\rho_{addClass}(c, d) = \inference{\forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
+	\forall a \in \rho(associations(c)) : reference(a) \in tables(d) 
+}{ %TODO podmínky?
+\begin{gathered}
+database(tables(d) \cup \rho(c) \cup \rho(properties(c)) \\ \cup \rho(associations(c)), instances(d))
+\end{gathered}
+}
+$
+\subsubsection{Add Property}
+$
+\rho_{addProperty}(l, p, d) =  \begin{cases}
+\inference{\exists t \in tables(d) : name(t) = l \\ \wedge (cardinality(p) = 0 \vee cardinality(p) = 1)}{database(tables(d) \wedge columns(t) \cup \rho(p), instances(d))} 
+\\ \\
+\inference{\exists t \in tables(d) : name(t) = l \\ \wedge cardinality(p) > 1}{database(tables(d) \cup \rho(p), instances(d))}  
+  \end{cases}
+$
+
+\subsection{Deconstruction of a System}
+\subsubsection{Remove Class}
+$
+\rho_{removeClass}(c,d) = \inference{\rho(c) \in tables(d) \\ \wedge \not \exists r \in data(d) : owningTable(r) \neq \rho(c) \\ \not \exists j \in tables(d), f \in foreignKeys(j) : reference(f) = \rho(c)  }{tables(d) \setminus \rho(c)}
+$
+
+\subsubsection{Remove Property}
 

### 2012-12-13 doc/description.tex
added definition of the set Label

@@
 
-There is a description of each component following. The description defines set of models of static components and dynamic components are defined as functions.
-%TODO: prerequisities - lables (if they are unique it is not necessary to check the name)
+There is a description of each component following.
+The description defines set of models of static components and dynamic components are defined as functions.
+There are several common parts of the model we can define in advance.
+The model uses a set $Labels$ which contains all possible identifiers for the model.
+Each model component which has a label ($label \in Labels$) is unambiguously identified in the model.
+
 %%%%%%%%%%%%%%%%%%%
@@
 It means the application is a set of instances which cooperates together.
-Each instance is defined according to a class - the class defines the structure, relationships of the instance.
-We use notation $instance \lhd class$ to describe the fact instance is defined according a class.
-
-The application model consists of model of classes - structural model and of instances - data model. The whole application is described as:
+Each instance is defined according to a class - the class defines the structure and relationships of the instance.
+The application component model consists of  classes defining the structure of application:
 
 $$
-Application = (Class*, Instance*)
+Application = (Class*)
 $$
-%TODO: are instances necessary on this level?
-
-The application is considered to be consistent if all instances are created according a class existing in the application. 
-$$
-\forall i \in instances(Application), \exists c \in classes(Application): i \lhd c 
-$$
-%TODO define consistency
 
@@
 
-%TODO exist \rho to execute mapping and \rho' (?) just returning a name of mapped object
-
 \subsection{Mapping of Types}

### 2012-12-13 doc/description.tex
text about operations, cleaned definitions

@@
 \label{sec:appEvolution}
+
+The evolution is defined as a set of operations which change structure of an application.
+The definition of an operation consists of two parts - a conditions of operation feasibility and operation impact on the application structure.
+We decide to use following transcription:
+
+$$
+operation = \inference{conditions$ $of$ $feasibility}{impact$ $on$ $application$ $structure}
+$$
+If the conditions of feasibility are not met, the operation cannot impact the application structure.
+%TODO monade
+
 \subsubsection{Application Creation}
@@
 AddClass(c, a) = \inference{\forall s \in classes(A) : name(s) \neq name(c)}
-{application(classes(a) \cup c, instances(a))}
+{classes(a) \cup c}
 $
@@
 AddProperty(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \not \exist q \in properties(p) : name(q) \neq name(p)}{
-application(classes(a)) \wedge properties(c) \cup p }
+ properties(c) \cup p }
 $
@@
 RemoveClass(q, a) = \inference{\exists c \in classes(a) : name(c) = q \\
-\wedge not$ $instantiated(c) = \{ \} \wedge not$ $associated(c)}{applicattion(classes(a) \setminus c, instances(a)) }
+\wedge not$ $instantiated(c) = \{ \} \wedge not$ $associated(c)}{classes(a) \setminus c }
 $
@@
 $
-RemoveProperty(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{}
+RemoveProperty(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{properties(c) \setminus p }
 $
@@
 $
-\rho : Property, Database \rightarrow Column 
+\rho : Property \times Database \rightarrow Column 
 $
@@
 \begin{gathered}
-database(tables(d) \cup \rho(c) \cup \rho(properties(c)) \\ \cup \rho(associations(c)), instances(d))
+tables(d) \cup \rho(c) \cup \rho(properties(c)) \cup \rho(associations(c))
 \end{gathered}
@@
 $
-\rho_{addProperty}(l, p, d) =  \begin{cases}
-\inference{\exists t \in tables(d) : name(t) = l \\ \wedge (cardinality(p) = 0 \vee cardinality(p) = 1)}{database(tables(d) \wedge columns(t) \cup \rho(p), instances(d))} 
-\\ \\
-\inference{\exists t \in tables(d) : name(t) = l \\ \wedge cardinality(p) > 1}{database(tables(d) \cup \rho(p), instances(d))}  
-  \end{cases}
+\rho_{addProperty}(l, p, d) = 
+\inference{\exists t \in tables(d) : name(t) = l}{\rho(p,d)}
+
 $
@@
 $
+%TODO remove by label
 
 \subsubsection{Remove Property}
+$
+\rho_{removeProperty}(l_c, l_p, d) = \inference{\exists t \in tables(d) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_p }{
+columns(t) \setminus c
+}
+$
 

### 2012-12-14 doc/description.tex
add association
remove assoc from app
url changed tu hyperref + added helper functions for application

@@
 \usepackage{listings}
-\usepackage{url}
+\usepackage{hyperref}
 
@@
 $$
+$classes : Application \rightarrow Class*$ \\
+$classes((classes)) = classes$
 
 \subsection{Application Structure Definition}
-The structure of the application is defined by five concepts. Each concept is described and its constructor is introduced.
+The structure of the application is defined by five concepts. Each concept is defined and a set of manipulating functions is define to manipulate the application model easily.
 
@@
 	\item Class : Class represents a basic structural concept in the application model. It has a unique name, one or more properties and a class can be associated to other classes in the application.
-$$
-Class = (label, Property*, Association*)
-$$
+
+$$Class = (label, Property*, Association*)$$
+$name : Class \rightarrow Label$ \\
+$name((lab, props, assocs)) = lab$ \\ \\
+$properties : Class \rightarrow Property*$ \\
+$properties((lab, props, assocs)) = props $\\ \\
+$primitiveProperties : Class \rightarrow Property*$ \\
+$primitiveProperties((lab, props, assocs)) = props \wedge \forall p \in props : cardinality(p) \leq 1 $\\ \\
+$collectionProperties : Class \rightarrow Property*$ \\
+$collectionProperties((lab, props, assocs)) = props \wedge \forall p \in props : cardinality(p) > 1 $\\ \\
+$associations : Class \rightarrow Association*$ \\
+$associations((lab, props, assocs)) = assocs $
+
 	\item Property : Property represents a feature of  a class which is can be represented as a primitive type. The property can be mandatory, can have a default value.
 $$
-Property = (label, App-Type, DefaultValue, Cardinality)
+Property = (label, AppType, DefaultValue, Cardinality)
 $$
+$name : Property \rightarrow Label$ \\
+$name((lab, t, val, n)) = lab$ \\ \\
+$type : Property \times AppType$ \\
+$type((lab, t, val, n)) = t$ \\ \\
+$cardinality : Property \rightarrow NzNat$ \\
+$cardinality((lab, t, val, n)) = n$
+
 %TODO owning class
@@
 $$
+$name : Association \rightarrow Label$ \\
+$name((lab, ref, n_1, n_2)) = lab$\\ \\
+$reference : Association \rightarrow Label$ \\
+$reference((lab, ref, n_1, n_2)) = ref$\\ \\
+$startCardinality : Association \rightarrow NzNat$ \\
+$startCardinality((lab, ref, n_1, n_2)) = n_1$\\ \\
+$endCardinality : Association \rightarrow NzNat$ \\
+$endCardinality((lab, ref, n_1, n_2)) = n_2$
+
 	\item App-Type : represents primitive types in the application.
@@
 $$
-App-Type = APPSTRING
+AppType = APPSTRING
 $$
 \end{itemize}
-An summary of application model:
+The types used in application model:
 $$
@@
 
-\subsection{Application Evolution}
+\section{Application Evolution}
 \label{sec:appEvolution}
@@
 
-\subsubsection{Application Creation}
+\subsection{Application Creation}
 $$AddClass: Class \times Application \rightarrow Application $$
@@
 $
+$$AddAssociation : Label \times Association \times Application \rightarrow Application $$
+$
+AddAssociation(l, s, a) = \inference{\not \exists f \in associations(a) : name(f) = name(s) \\ \wedge \exists c in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(a) \cup s}
+$
 
-\subsubsection{Application Deconstruction}
+\subsection{Application Deconstruction}
 $$RemoveClass: Label \times Application \rightarrow Application $$
@@
 RemoveProperty(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{properties(c) \setminus p }
+$
+
+$$RemoveAssociation : Label \times Labels \times Application \rightarrow Application $$
+
+$
+RemoveAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists s \in associations(c) : name(s) = l_a }{associations(c) \setminus s}
 $
@@
 \subsection{Mapping of Properties}
-There are two types of properties - the first type called single properties has cardinality 0 or 1, the second one (called collections) has cardinality greater than 1.
-Each type has its own rules of mapping.
-
-\subsubsection{Mapping of Single Property}
-Each single property is mapped to a column, when the cardinality is equal to 1 the $NOTNULL$ constraint is added to a column.
-A property with cardinality greater than one has to be mapped int a table not a column, thus we have a special mapping case for collections.
-A collection is mapped into a table which contains a foreign key.
-The key references the table representing the owning class.
+There are two types of properties - the first type has cardinality 0 or 1 and it is mapped to a column (when the cardinality is equal to 1 the $NOTNULL$ constraint is added to a column).
+A property with cardinality greater than one has to be mapped int a table not a column, thus we have a special mapping case.
 
@@
 \inference{\exists t \in tables(d) : name(t) = l}{\rho(p,d)} 
-
+$
+\subsection{Add Association}
+$
+\rho_{AddAssociation}(l, a, d) = \inference{\exists t_1, t_2 \in tables(d) : name(t_1) = l \\ \wedge name(t_2) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_1) : name(f) = name(a) }{\rho(a,d)}
 $
@@
 $
+%TODO 1:N M:N
 

### 2012-12-14 doc/description.tex
owningClass function

@@
 
-\begin{itemize}
-	\item Class : Class represents a basic structural concept in the application model. It has a unique name, one or more properties and a class can be associated to other classes in the application.
+\paragraph{Class} represents a basic structural concept in the application model. It has a unique name, one or more properties and a class can be associated to other classes in the application.
 
@@
 $primitiveProperties : Class \rightarrow Property*$ \\
-$primitiveProperties((lab, props, assocs)) = props \wedge \forall p \in props : cardinality(p) \leq 1 $\\ \\
+$primitiveProperties((lab, props, assocs)) = props', \forall p \in props' : cardinality(p) \leq 1 $\\ \\
 $collectionProperties : Class \rightarrow Property*$ \\
-$collectionProperties((lab, props, assocs)) = props \wedge \forall p \in props : cardinality(p) > 1 $\\ \\
+$collectionProperties((lab, props, assocs)) = props', \forall p \in props' : cardinality(p) > 1 $\\ \\
 $associations : Class \rightarrow Association*$ \\
@@
 
-	\item Property : Property represents a feature of  a class which is can be represented as a primitive type. The property can be mandatory, can have a default value.
+\paragraph{Property} represents a feature of  a class which is can be represented as a primitive type. The property can be mandatory, can have a default value.
 $$
@@
 $cardinality : Property \rightarrow NzNat$ \\
-$cardinality((lab, t, val, n)) = n$
+$cardinality((lab, t, val, n)) = n$ \\ \\
+$owningClass : Property \times Application \rightarrow Class $ \\
+$owningClass(p, a) = c, c \in classes(a) \wedge p \in properties(c) $
 
-%TODO owning class
-	\item Association : the association represents a connection between two classes.
+\paragraph {Association} represents a connection between two classes.
 It has a unique name and it contains a name of the class which is referenced by the association.
@@
 $endCardinality : Association \rightarrow NzNat$ \\
-$endCardinality((lab, ref, n_1, n_2)) = n_2$
+$endCardinality((lab, ref, n_1, n_2)) = n_2$ \\ \\
+$owningClass : Association \times Application \rightarrow Class $ \\
+$owningClass(as, a) = c, c \in classes(a) \wedge as \in associations(c) $
 
-	\item App-Type : represents primitive types in the application.
+\paragraph{AppType} represents primitive types in the application.
 There are usually defined types such as String, Integer, Boolean etc.
@@
 $$
-\end{itemize}
 The types used in application model:

### 2012-12-14 doc/description.tex
formating improvement, mappings improvement

@@
 $$
-Application = (Class*)
+Application = (Class*, Instance*)
 $$
 $classes : Application \rightarrow Class*$ \\
-$classes((classes)) = classes$
+$classes((classes)) = classes$ \\ \\
+$instantiated : Class \times Application \rightarrow Boolean $
+%TODO: variable types
 
@@
 $associations : Class \rightarrow Association*$ \\
-$associations((lab, props, assocs)) = assocs $
+$associations((lab, props, assocs)) = assocs $ \\ \\
+$associated :  Class \times Application \rightarrow Boolean $
+%TODO: variable types
 
@@
 $owningClass(p, a) = c, c \in classes(a) \wedge p \in properties(c) $
+%TODO: variable types
 
@@
 $owningClass(as, a) = c, c \in classes(a) \wedge as \in associations(c) $
+%TODO: variable types
 
@@
 $
-AddProperty(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \not \exist q \in properties(p) : name(q) \neq name(p)}{
+AddProperty(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in properties(p) : name(q) \neq name(p)}{
  properties(c) \cup p }
@@
 $
-AddAssociation(l, s, a) = \inference{\not \exists f \in associations(a) : name(f) = name(s) \\ \wedge \exists c in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(a) \cup s}
+AddAssociation(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(a) \cup s}
 $
@@
 RemoveClass(q, a) = \inference{\exists c \in classes(a) : name(c) = q \\
-\wedge not$ $instantiated(c) = \{ \} \wedge not$ $associated(c)}{classes(a) \setminus c }
+\wedge not$ $instantiated(c) = \varnothing \wedge not$ $associated(c)}{classes(a) \setminus c }
 $
@@
 
-\section{Database model}
+\section{Database Component Model}
 The relational database consists of database schema which defines the structure of the database and data. The database is a pair od sets:
@@
 The database schema model is defined by following concepts:
-\begin{itemize}
-	\item Table: table represents a basic concept of database schema.
+\paragraph{Table} represents a basic concept of database schema.
 It has a unique name, one or more column and it can be related to other tables in the schema by foreign keys.
@@
 $$
-	\item Column : columns define possible data values and types which can be part of a table record.
+$name : Table \rightarrow Label $ \\
+$name((lab, pk, cols, fks)) = lab$ \\ \\
+$primaryKey : Table \rightarrow PrimaryKey $ \\
+$primaryKey((lab, pk, cols, fks)) = pk$ \\ \\
+$columns : Table \rightarrow Column* $ \\
+$columns((lab, pk, cols, fks)) = cols$ \\ \\
+$foreignKeys : Table \rightarrow ForeignKey* $ \\
+$foreignKeys((lab, pk, cols, fks)) = fks$
+
+\paragraph{Column} columns define possible data values and types which can be part of a table record.
 $$
-Column = (label, Db-Type, Constraint*)
+Column = (label, DbType, DefaultValue, Constraint*)
 $$
+$name : Column \rightarrow Label $ \\
+$name((lab, t, val, cons)) = lab $ \\ \\
+$type : Column \rightarrow DbType $ \\
+$type((lab, t, val, cons)) = t $ \\ \\
+$constraints : Column \rightarrow Constraint* $ \\
+$constraints((lab, t, val, cons)) = cons $ \\ \\
+$owningTable : Column \times Database \rightarrow Table $ \\
+$constraints(c, d) = tab, tab \in tables(d) \wedge c \in columns(d) $
 
-	\item Foreign key:
+\paragraph{Foreign key}
 $$
@@
 $$
+$name : ForeignKey \rightarrow Label $ \\
+$name((lab, tRef, cons)) = lab $ \\ \\
+$reference : ForeignKey \rightarrow Label $ \\
+$reference((lab, tRef, cons)) = tRef $ \\ \\
+$reference : ForeignKey \rightarrow Constraint* $ \\
+$reference((lab, tRef, cons)) = cons $ \\ \\
+$owningTable : ForeignKey \times Database \rightarrow Table $ \\
+$owningTable(fk, d) = tab, tab \in tables(d) \wedge fk \in foreignKeys(d) $
 
-	\item Primary key:
+\paragraph{Primary key}
 $$
@@
 
-	\item Db-Type: represents primitive types in the database. There are usually defined types such as Char, Integer, Boolean etc. There is only one type defined in the model which is called DB-STRING.
+\paragraph{DbType} represents primitive types in the database. There are usually defined types such as Char, Integer, Boolean etc. There is only one type defined in the model which is called DBSTRING.
 $$
-Db-Type = DBSTRING
+DbType = DBSTRING
 $$
 
-	\item Constraint: there are two types of constraints defined in the model.
+\paragraph{Constraint} there are two types of constraints defined in the model.
 Both constraints are column constraints - first constraint defines unique records in a column, second constraint defines non-empty columns.
-$$
-Constraint = NOTNULL | UNIQUE 
-$$
-\end{itemize}
+
+\begin{center}
+$Constraint = NOTNULL$ $|$ $UNIQUE $
+\end{center}
 
@@
 We focus on structural and data change therefore we simplify types and its mapping.
-$$
-\rho : AppType \rightarrow DbType
-$$
 
-$$
-\rho(APPSTRING) = DBSTRING
-$$
+\begin{center}
+$ \rho : AppType \rightarrow DbType $ \\
+$ \rho(APPSTRING) = DBSTRING $
+\end{center}
 
@@
 
-$$
-\rho : Class \rightarrow Table
-$$
-
-$$
-\rho(c) = table(name(c), primaryKey("Id"), \rho(properties(c), \{ \})
-$$
+\begin{center}
+$\rho : Class \rightarrow Table $ \\ 
+$ \rho(c) = table(name(c), primaryKey("Id"), \rho(primitiveProperties(c), \varnothing $\\$ \cup \rho(associations(c)) \cup \rho(collectionProperties)$
+\end{center}
 
@@
 $\rho(p, d) = \begin{cases}
- \inference{cardinality(p) = 0 \\ \wedge \rho(owningClass(p) \in classes(d)}{columns(owninngClass(p)) \cup column(name(p), \rho(type(p)), \{ \})} \\ \\
+ \inference{cardinality(p) = 0 \\ \wedge \rho(owningClass(p)) \in classes(d)}{columns(\rho(owningClass(p))) \cup column(name(p), \rho(type(p)), \varnothing)} \\ \\
 
- \inference{cardinality(p) = 1 \\ \wedge \rho(owningClass(p) \in classes(d)}{columns(owninngClass(p)) \cup col(name(p), \rho(type(p)), NOTNULL)} \\ \\
+ \inference{cardinality(p) = 1 \\ \wedge \rho(owningClass(p) \in classes(d)}{columns(\rho(owningClass(p))) \cup column(name(p), \rho(type(p)), NOTNULL)} \\ \\
 
- \inference{cardinality(p) > 2 \\ \wedge \rho(owningClass(p) \in classes(d)}{\begin{gathered} tables(d) \cup table(name(p), pk("Id"), col("value", DBSTRING), \\ fk(name(p), \rho(owningClass(p))) \end{gathered}}
+ \inference{cardinality(p) > 2 \\ \wedge \rho(owningClass(p) \in classes(d)}{\begin{gathered} table(name(p), primaryKey("Id"), column("value", \rho(type(p))), \\ foreignKey(name(p), name(owningClass(p))) \end{gathered}}
  \end{cases}$
@@
 \rho(a,d) = \begin{cases}
- \inference{startCardinality(a) = 0 \\ \wedge \rho(owningClass(p) \in classes(d)  \wedge \rho(reference(a)) \in tables(d)}{foreignKeys(\rho(assoc(a)) \cup foreignKey(name(a), owningClass(a),  \{ \}) } \\ \\
+ \inference{startCardinality(a) = 0 \\ \wedge \rho(owningClass(p)) \in classes(d)  \wedge \rho(reference(a)) \in tables(d)}{
+ \begin{gathered}
+ foreignKeys(\rho(reference(a))) \\ \cup foreignKey(name(a), owningClass(a),  \varnothing) 
+ \end{gathered}
+ }
+
+  \\ \\
 
  \inference{startCardinality(a) = 1 \\ \wedge \rho(owningClass(p) \in tables(d)  \wedge \rho(reference(a)) \in tables(d)}{\begin{gathered}
- foreignKeys(\rho(assoc(a)) \cup foreignKey(name(a), \\ owningClass(a),  NOTNULL)
+foreignKeys(\rho(reference(a))) \\ \cup  foreignKey(name(a), \\ owningClass(a),  NOTNULL)
 \end{gathered}
@@
   \begin{gathered}
-  tables(d) \cup table(name(a), pk("Id"), \{ \}, \\ foreignKey(owningClass(a),owningClass(a), NOTNULL) \\ foreignkey(reference(a), reference(a), NOTNULL)) 
+ table(name(a), pk("Id"), \varnothing, \\ foreignKey(owningClass(a),owningClass(a), NOTNULL) \\ foreignkey(reference(a), reference(a), NOTNULL)) 
   \end{gathered}}
@@
 Each evolution operation consist of two parts - applying evolution operation on application and on database.
-In general an evolutionary operation $X : OrmSystem \rightarrow OrmSystem $ can be defined as:
-
-$$
-\inference[X(s) = ]{}{system(X(application(s)), \rho_{X}(database(s))) }
-$$
-where the $\rho_X$ is the mapping of the evolution from application to database.
 The mapping of the evolution change is called extended Object-Relational Mapping (eORM).
@@
 	\forall a \in \rho(associations(c)) : reference(a) \in tables(d) 
-}{ %TODO podmínky?
+}{
 \begin{gathered}
-tables(d) \cup \rho(c) \cup \rho(properties(c)) \cup \rho(associations(c))
+tables(d) \cup \rho(c) \cup \rho(collectionProperties(c)) \\ \cup \rho(associations(c))
 \end{gathered}
@@
 $
-\rho_{removeClass}(c,d) = \inference{\rho(c) \in tables(d) \\ \wedge \not \exists r \in data(d) : owningTable(r) \neq \rho(c) \\ \not \exists j \in tables(d), f \in foreignKeys(j) : reference(f) = \rho(c)  }{tables(d) \setminus \rho(c)}
+\rho_{removeClass}(c,d) = \inference{\rho(c) \in tables(d) \\ \wedge \not \exists r \in data(d) : owningTable(r) = \rho(c) \\ \not \exists j \in tables(d), f \in foreignKeys(j) : reference(f) = \rho(c)  }{tables(d) \setminus \rho(c)}
 $
@@
 $
-\rho_{removeProperty}(l_c, l_p, d) = \inference{\exists t \in tables(d) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_p  }{
+\rho_{removeProperty}(l_c, l_p, d) = \begin{cases}
+ \inference{\exists t \in tables(d) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_p  }{
 columns(t) \setminus c
-}
+} \\ \\
+  \inference{}{}
+ \end{cases}
+
 $

### 2012-12-14 doc/description.tex
new helpers updating mappings

@@
 $$
-Application = (Class*, Instance*)
+Application = <Class*, Instance*>
 $$
 $classes : Application \rightarrow Class*$ \\
-$classes((classes)) = classes$ \\ \\
-$instantiated : Class \times Application \rightarrow Boolean $
+$classes(<cs, is>) = cs, cs \in Class* \wedge is \in Instance* $ \\ \\
+$instantiated : Class \times Application \rightarrow Boolean $ \\
+\\
+$class : Label \times Application \rightarrow Class  $ \\ 
+$class(l, a) = c, c \in classes(a) \wedge name(c) = l, l \in Label, a \in Application* $ 
 %TODO: variable types
@@
 \subsection{Application Creation}
-$$AddClass: Class \times Application \rightarrow Application $$
+$$addClass: Class \times Application \rightarrow Application $$
 
 $
-AddClass(c, a) = \inference{\forall s \in classes(A) : name(s) \neq name(c)}
+addClass(c, a) = \inference{\forall s \in classes(A) : name(s) \neq name(c)}
 {classes(a) \cup c}
@@
 $
-AddAssociation(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(a) \cup s}
+AddAssociation(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(c) \cup s}
 $
@@
 $$
+$table : Label \times Database \rightarrow Table $ \\
+$table(l, d) = t, t \in tables(d) \wedge name(t) = l$ \\
 
@@
 \begin{center}
-$ \rho : AppType \rightarrow DbType $ \\
-$ \rho(APPSTRING) = DBSTRING $
+$ \rho_{t} : AppType \rightarrow DbType $ \\
+$ \rho_{t}(APPSTRING) = DBSTRING $
 \end{center}
@@
 \begin{center}
-$\rho : Class \rightarrow Table $ \\ 
-$ \rho(c) = table(name(c), primaryKey("Id"), \rho(primitiveProperties(c), \varnothing $\\$ \cup \rho(associations(c)) \cup \rho(collectionProperties)$
+$\rho_{c}: Class \rightarrow Table $ \\ 
+$\rho_{c}(c) = table(name(c), primaryKey("Id"), \rho_{p}(primitiveProperties(c), \varnothing $\\$ \cup \rho_{a}(associations(c)) \cup \rho_{p}(collectionProperties)$
 \end{center}
@@
 $
-\rho : Property \times Database \rightarrow Column 
+\rho_p : Property \times Database \rightarrow Column 
 $
 
-$\rho(p, d) = \begin{cases}
- \inference{cardinality(p) = 0 \\ \wedge \rho(owningClass(p)) \in classes(d)}{columns(\rho(owningClass(p))) \cup column(name(p), \rho(type(p)), \varnothing)} \\ \\
+$\rho_p(p, d) = \begin{cases}
+ \inference{cardinality(p) = 0 \\ \wedge \rho_c(owningClass(p)) \in tables(d)}{\begin{gathered}
+ columns(table(name(owningClass(p)),d) \\ \cup column(name(p), \rho_t(type(p)), \varnothing)
+\end{gathered}
+} \\ \\
 
- \inference{cardinality(p) = 1 \\ \wedge \rho(owningClass(p) \in classes(d)}{columns(\rho(owningClass(p))) \cup column(name(p), \rho(type(p)), NOTNULL)} \\ \\
+ \inference{cardinality(p) = 1 \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL)} \\ \\
 
- \inference{cardinality(p) > 2 \\ \wedge \rho(owningClass(p) \in classes(d)}{\begin{gathered} table(name(p), primaryKey("Id"), column("value", \rho(type(p))), \\ foreignKey(name(p), name(owningClass(p))) \end{gathered}}
+ \inference{cardinality(p) > 2 \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered} table(name(p), primaryKey("Id"), column("value", \rho_t(type(p))), \\ foreignKey(name(p), name(owningClass(p))) \end{gathered}}
  \end{cases}$
@@
 $
-\rho(a,d) = \begin{cases}
- \inference{startCardinality(a) = 0 \\ \wedge \rho(owningClass(p)) \in classes(d)  \wedge \rho(reference(a)) \in tables(d)}{
+\rho_a(a,d) = \begin{cases}
+ \inference{startCardinality(a) = 0 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{
  \begin{gathered}
- foreignKeys(\rho(reference(a))) \\ \cup foreignKey(name(a), owningClass(a),  \varnothing) 
+ foreignKeys(u) \\ \cup foreignKey(name(a), name(t),  \varnothing) 
  \end{gathered}
  }
-
   \\ \\
-
- \inference{startCardinality(a) = 1 \\ \wedge \rho(owningClass(p) \in tables(d)  \wedge \rho(reference(a)) \in tables(d)}{\begin{gathered}
-foreignKeys(\rho(reference(a))) \\ \cup  foreignKey(name(a), \\ owningClass(a),  NOTNULL)
+ \inference{startCardinality(a) = 1 \\\\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{\begin{gathered}
+foreignKeys(u) \\ \cup  foreignKey(name(a), name(t),  NOTNULL)
 \end{gathered}
@@
 
-%TODO FK nemusí mít jméno - odkazuje jen na tabulku a id
+%TODO handle instances
 
@@
 
+%TODO jak se postupně vyvíjí systém
 \subsection{Creating of a System}
@@
 \rho_{addClass}(c, d) = \inference{\forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
-	\forall a \in \rho(associations(c)) : reference(a) \in tables(d) 
+	\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) 
 }{
@@
 $
-\rho_{addProperty}(l, p, d) = 
-\inference{\exists t \in tables(d) : name(t) = l}{\rho(p,d)} 
+\rho_{addProperty}(p, d) = 
+\inference{\exists t \in tables(d) : name(t) = name(owningClass(p))}{\rho(p,d)} 
 $
@@
 $
-\rho_{AddAssociation}(l, a, d) = \inference{\exists t_1, t_2 \in tables(d) : name(t_1) = l \\ \wedge name(t_2) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_1) : name(f) = name(a) }{\rho(a,d)}
+\rho_{AddAssociation}(a, d) = \inference{\exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) }{\rho(a,d)}
 $
@@
 } \\ \\
+  \inference{\exists t \in tables(d) : name(t) = l_c \\ \wedge \not \exists c \in columns(t) : name(c) = l_p \\ \wedge \exists u \in tables(d) : name(u) = l_p }{tables(d) \setminus u }
+ \end{cases}
+$
+
+\subsubsection{Remove Association}
+$
+\rho_{removeAssociation}(l_c, l_a, d) = \begin{cases}
+ \inference{\exists t \in tables(d) : name(t) = l_c \wedge }{foreignKeys(\rho(reference(a))) \ }
+
  \inference{}{}
  \end{cases}
-
 $
-%TODO 1:N M:N
 

### 2012-12-14 doc/description.tex
adding context in database operations

@@
 $\rho_{c}: Class \rightarrow Table $ \\ 
-$\rho_{c}(c) = table(name(c), primaryKey("Id"), \rho_{p}(primitiveProperties(c), \varnothing $\\$ \cup \rho_{a}(associations(c)) \cup \rho_{p}(collectionProperties)$
+$\rho_{c}(c,d) = d \vdash table(name(c), primaryKey("Id"), \rho_{p}(primitiveProperties(c), \varnothing $\\$ \cup \rho_{a}(associations(c)) \cup \rho_{p}(collectionProperties)$
 \end{center}
@@
 $\rho_p(p, d) = \begin{cases}
- \inference{cardinality(p) = 0 \\ \wedge \rho_c(owningClass(p)) \in tables(d)}{\begin{gathered}
+ \inference{d \vdash cardinality(p) = 0 \\ \wedge \rho_c(owningClass(p)) \in tables(d)}{\begin{gathered} d \vdash 
  columns(table(name(owningClass(p)),d) \\ \cup column(name(p), \rho_t(type(p)), \varnothing)
@@
 
- \inference{cardinality(p) = 1 \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL)} \\ \\
+ \inference{d \vdash cardinality(p) = 1 \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{d \vdash columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL)} \\ \\
 
- \inference{cardinality(p) > 2 \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered} table(name(p), primaryKey("Id"), column("value", \rho_t(type(p))), \\ foreignKey(name(p), name(owningClass(p))) \end{gathered}}
+ \inference{d \vdash cardinality(p) > 2 \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}d \vdash  table(name(p), primaryKey("Id"), column("value", \rho_t(type(p))), \\ foreignKey(name(p), name(owningClass(p))) \end{gathered}}
  \end{cases}$
@@
 \rho_a(a,d) = \begin{cases}
- \inference{startCardinality(a) = 0 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{
+ \inference{d \vdash startCardinality(a) = 0 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{d \vdash 
  \begin{gathered}
@@
   \\ \\
- \inference{startCardinality(a) = 1 \\\\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{\begin{gathered}
+ \inference{d \vdash startCardinality(a) = 1 \\\\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{\begin{gathered} d \vdash 
 foreignKeys(u) \\ \cup  foreignKey(name(a), name(t),  NOTNULL)
@@
 
-  \inference{startCardinality > 1 \\ \wedge \rho(owningClass(p)) \in tables(d) \wedge \rho(reference(a)) \in tables(d)}{
-  \begin{gathered}
+  \inference{d \vdash  startCardinality > 1 \\ \wedge \rho(owningClass(p)) \in tables(d) \wedge \rho(reference(a)) \in tables(d)}{
+  \begin{gathered} d \vdash 
  table(name(a), pk("Id"), \varnothing, \\ foreignKey(owningClass(a),owningClass(a), NOTNULL) \\ foreignkey(reference(a), reference(a), NOTNULL)) 
@@
 $
-\rho_{addClass}(c, d) = \inference{\forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
+\Phi_{addClass}(c, d) = \inference{d \vdash \forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
 	\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) 
@@
 \begin{gathered}
-tables(d) \cup \rho(c) \cup \rho(collectionProperties(c)) \\ \cup \rho(associations(c))
+d \vdash  tables(d) \cup \rho(c) \cup \rho(collectionProperties(c)) \\ \cup \rho(associations(c))
 \end{gathered}
@@
 $
-\rho_{addProperty}(p, d) = 
-\inference{\exists t \in tables(d) : name(t) = name(owningClass(p))}{\rho(p,d)} 
+\Phi_{addProperty}(p, d) = 
+\inference{d \vdash \exists t \in tables(d) : name(t) = name(owningClass(p))}{d \vdash \rho(p,d)} 
 $
@@
 $
-\rho_{AddAssociation}(a, d) = \inference{\exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) }{\rho(a,d)}
+\Phi_{AddAssociation}(a, d) = \inference{d \vdash \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) }{d \vdash \rho(a,d)}
 $
@@
 $
-\rho_{removeClass}(c,d) = \inference{\rho(c) \in tables(d) \\ \wedge \not \exists r \in data(d) : owningTable(r) = \rho(c) \\ \not \exists j \in tables(d), f \in foreignKeys(j) : reference(f) = \rho(c)  }{tables(d) \setminus \rho(c)}
+\Phi_{removeClass}(c,d) = \inference{d \vdash \rho(c) \in tables(d) \\ \wedge \not \exists r \in data(d) : owningTable(r) = \rho(c) \\ \not \exists j \in tables(d), f \in foreignKeys(j) : reference(f) = \rho(c)  }{d \vdash tables(d) \setminus \rho(c)}
 $
@@
 $
-\rho_{removeProperty}(l_c, l_p, d) = \begin{cases}
- \inference{\exists t \in tables(d) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_p  }{
+\Phi_{removeProperty}(l_c, l_p, d) = \begin{cases}
+ \inference{d \vdash \exists t \in tables(d) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_p  }{d \vdash
 columns(t) \setminus c
 } \\ \\
-  \inference{\exists t \in tables(d) : name(t) = l_c \\ \wedge \not \exists c \in columns(t) : name(c) = l_p \\ \wedge \exists u \in tables(d) : name(u) = l_p }{tables(d) \setminus u }
+  \inference{d \vdash \exists t \in tables(d) : name(t) = l_c \\ \wedge \not \exists c \in columns(t) : name(c) = l_p \\ \wedge \exists u \in tables(d) : name(u) = l_p }{d \vdash tables(d) \setminus u }
  \end{cases}
@@
 $
-\rho_{removeAssociation}(l_c, l_a, d) = \begin{cases}
- \inference{\exists t \in tables(d) : name(t) = l_c \wedge }{foreignKeys(\rho(reference(a))) \ }
+\Phi_{removeAssociation}(l_c, l_a, d) = \begin{cases}
+ \inference{d \vdash \exists t \in tables(d) : name(t) = l_c \wedge }{d \vdash foreignKeys(\rho(reference(a))) \ }
 

### 2012-12-14 doc/description.tex
start working on including data into model

@@
 
- \inference{d \vdash cardinality(p) = 1 \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{d \vdash columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL)} \\ \\
+ \inference{ d \vdash cardinality(p) = 1 
+ \\ \wedge \exists t : t = table(name(owningClass(p)), d) 
+ \\ \wedge \not instanciated(owningClass(p))
+ }{ d \vdash columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL)} \\ \\
 
@@
  \end{cases}$
+
+The new foreign key can cause inconsistency of the system when added to a table which contains data.
+This case is not handled by the mapping function above.
+We have to extend the function definition of a feature that allows the mapping between the data in both tables.
+The mapping is a function:
+$$f_m : Property \rightarrow Instance \times Instance $$
 

### 2012-12-16 doc/chunk-possible ebolutions.tex
created images, added mandatory attribute to Property, changed ORM

--- 2012-12-14 doc/description.tex
@@
-\documentclass[11pt]{article}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{semantic}
-\usepackage{listings}
-\usepackage{hyperref}
-
-\begin{document}
-\tableofcontents
-
-\newpage
-\section{Evolution of ORM System}
-The ORM system consist of application and database.
-The application and database has to be in consistent state according to a concrete ORM algorithm, otherwise the application cannot work properly.
-It means that during the evolution we have to respect a concrete ORM to produce consistent result.
-$$
-OrmSystem = (Application, Database)
-$$
-%TODO define consistency
-
 There are several way how the evolution of ORM system can be implemented.
@@
 This allows us to keep history of all layers and to verify the evolution on multiple levels.
-The algorithm is illustrated in Figure \ref{fig:algorithm}.
-
-The figure shows all important components of the domain.
-There are three static components (static because they do not define behavior, but state of the system): an application, a database schema and stored data.
-Then there are four dynamic components (they define transformations between components and transition between system states): an object-relational mapping is a transformation between application and database schema, then are two components for evolution of application and of database.
-(To simplify the model we will consider the database schema and data are evolved by one component).
-Last component is the most important one - it is component which maps the evolution from application level to relational database level.
-We call this component evolution object-relational mapping (eORM).
-
-There is a description of each component following.
-The description defines set of models of static components and dynamic components are defined as functions.
-There are several common parts of the model we can define in advance.
-The model uses a set $Labels$ which contains all possible identifiers for the model.
-Each model component which has a label ($label \in Labels$) is unambiguously identified in the model.
-
-%%%%%%%%%%%%%%%%%%%
-
-\section{Application Component}
-The application component represent an application created using object oriented language.
-It means the application is a set of instances which cooperates together.
-Each instance is defined according to a class - the class defines the structure and relationships of the instance.
-The application component model consists of  classes defining the structure of application:
-
-$$
-Application = <Class*, Instance*>
-$$
-$classes : Application \rightarrow Class*$ \\
-$classes(<cs, is>) = cs, cs \in Class* \wedge is \in Instance* $ \\ \\
-$instantiated : Class \times Application \rightarrow Boolean $ \\
-\\
-$class : Label \times Application \rightarrow Class  $ \\ 
-$class(l, a) = c, c \in classes(a) \wedge name(c) = l, l \in Label, a \in Application* $ 
-%TODO: variable types
-
-\subsection{Application Structure Definition}
-The structure of the application is defined by five concepts. Each concept is defined and a set of manipulating functions is define to manipulate the application model easily.
-
-\paragraph{Class} represents a basic structural concept in the application model. It has a unique name, one or more properties and a class can be associated to other classes in the application.
-
-$$Class = (label, Property*, Association*)$$
-$name : Class \rightarrow Label$ \\
-$name((lab, props, assocs)) = lab$ \\ \\
-$properties : Class \rightarrow Property*$ \\
-$properties((lab, props, assocs)) = props $\\ \\
-$primitiveProperties : Class \rightarrow Property*$ \\
-$primitiveProperties((lab, props, assocs)) = props', \forall p \in props' : cardinality(p) \leq 1 $\\ \\
-$collectionProperties : Class \rightarrow Property*$ \\
-$collectionProperties((lab, props, assocs)) = props', \forall p \in props' : cardinality(p) > 1 $\\ \\
-$associations : Class \rightarrow Association*$ \\
-$associations((lab, props, assocs)) = assocs $ \\ \\
-$associated :  Class \times Application \rightarrow Boolean $
-%TODO: variable types
-
-\paragraph{Property} represents a feature of  a class which is can be represented as a primitive type. The property can be mandatory, can have a default value.
-$$
-Property = (label, AppType, DefaultValue, Cardinality)
-$$
-$name : Property \rightarrow Label$ \\
-$name((lab, t, val, n)) = lab$ \\ \\
-$type : Property \times AppType$ \\
-$type((lab, t, val, n)) = t$ \\ \\
-$cardinality : Property \rightarrow NzNat$ \\
-$cardinality((lab, t, val, n)) = n$ \\ \\
-$owningClass : Property \times Application \rightarrow Class $ \\
-$owningClass(p, a) = c, c \in classes(a) \wedge p \in properties(c) $
-%TODO: variable types
-
-\paragraph {Association} represents a connection between two classes.
-It has a unique name and it contains a name of the class which is referenced by the association.
-The class which owns the association is consider to be a starting class of an association, referenced class is consider to be an ending class of an association.
-The cardinalities defines the multiplicity of the association.
-$$
-Association = (label, classRef, startCardinality, endCardinality)
-$$
-$name : Association \rightarrow Label$ \\
-$name((lab, ref, n_1, n_2)) = lab$\\ \\
-$reference : Association \rightarrow Label$ \\
-$reference((lab, ref, n_1, n_2)) = ref$\\ \\
-$startCardinality : Association \rightarrow NzNat$ \\
-$startCardinality((lab, ref, n_1, n_2)) = n_1$\\ \\
-$endCardinality : Association \rightarrow NzNat$ \\
-$endCardinality((lab, ref, n_1, n_2)) = n_2$ \\ \\
-$owningClass : Association \times Application \rightarrow Class $ \\
-$owningClass(as, a) = c, c \in classes(a) \wedge as \in associations(c) $
-%TODO: variable types
-
-\paragraph{AppType} represents primitive types in the application.
-There are usually defined types such as String, Integer, Boolean etc.
-in contrast there is only one type in our model, because we focus on structural and data changes and type casting operations are not important for us.
-The only App-Type type is called APP-STRING.
-$$
-AppType = APPSTRING
-$$
-The types used in application model:
-$$
-cardinality = NzNat
-$$
-$$
-referencesClass = label
-$$
-$$
-startCardinality = NzNat
-$$
-$$
-endCardinality = NzNat
-$$
-$$
-NzNat = [0 - 9]*
-$$
-
-\section{Application Evolution}
-\label{sec:appEvolution}
-
-The evolution is defined as a set of operations which change structure of an application.
-The definition of an operation consists of two parts - a conditions of operation feasibility and operation impact on the application structure.
-We decide to use following transcription:
-
-$$
-operation = \inference{conditions$ $of$ $feasibility}{impact$ $on$ $application$ $structure}
-$$
-If the conditions of feasibility are not met, the operation cannot impact the application structure.
-%TODO monade
-
-\subsection{Application Creation}
-$$addClass: Class \times Application \rightarrow Application $$
-
-$
-addClass(c, a) = \inference{\forall s \in classes(A) : name(s) \neq name(c)}
-{classes(a) \cup c}
-$
-
-$$AddProperty : Label \times Property \times Application \rightarrow Application $$
-
-$
-AddProperty(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in properties(p) : name(q) \neq name(p)}{
- properties(c) \cup p }
-$
-$$AddAssociation : Label \times Association \times Application \rightarrow Application $$
-$
-AddAssociation(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(c) \cup s}
-$
-
-\subsection{Application Deconstruction}
-$$RemoveClass: Label \times Application \rightarrow Application $$
-$
-RemoveClass(q, a) = \inference{\exists c \in classes(a) : name(c) = q \\
-\wedge not$ $instantiated(c) = \varnothing \wedge not$ $associated(c)}{classes(a) \setminus c }
-$
-
-$$RemoveProperty: Label \times Label \times Application \rightarrow Application $$
-
-$
-RemoveProperty(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{properties(c) \setminus p }
-$
-
-$$RemoveAssociation : Label \times Labels \times Application \rightarrow Application $$
-
-$
-RemoveAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists s \in associations(c) : name(s) = l_a }{associations(c) \setminus s}
-$
-
-%%%%%%%%%%%%%%%%%%%
-
-\section{Database Component Model}
-The relational database consists of database schema which defines the structure of the database and data. The database is a pair od sets:
-$$
-Database = (table*, data*)
-$$
-$table : Label \times Database \rightarrow Table $ \\
-$table(l, d) = t, t \in tables(d) \wedge name(t) = l$ \\
-
-\subsection{Database Schema Model}
-The database schema model is defined by following concepts:
-\paragraph{Table} represents a basic concept of database schema.
-It has a unique name, one or more column and it can be related to other tables in the schema by foreign keys.
-Rows in the table represents stored data.
-$$
-Table = (label, primaryKey, Column*, ForeignKey*)
-$$
-$name : Table \rightarrow Label $ \\
-$name((lab, pk, cols, fks)) = lab$ \\ \\
-$primaryKey : Table \rightarrow PrimaryKey $ \\
-$primaryKey((lab, pk, cols, fks)) = pk$ \\ \\
-$columns : Table \rightarrow Column* $ \\
-$columns((lab, pk, cols, fks)) = cols$ \\ \\
-$foreignKeys : Table \rightarrow ForeignKey* $ \\
-$foreignKeys((lab, pk, cols, fks)) = fks$
-
-\paragraph{Column} columns define possible data values and types which can be part of a table record.
-$$
-Column = (label, DbType, DefaultValue, Constraint*)
-$$
-$name : Column \rightarrow Label $ \\
-$name((lab, t, val, cons)) = lab $ \\ \\
-$type : Column \rightarrow DbType $ \\
-$type((lab, t, val, cons)) = t $ \\ \\
-$constraints : Column \rightarrow Constraint* $ \\
-$constraints((lab, t, val, cons)) = cons $ \\ \\
-$owningTable : Column \times Database \rightarrow Table $ \\
-$constraints(c, d) = tab, tab \in tables(d) \wedge c \in columns(d) $
-
-\paragraph{Foreign key}
-$$
-ForeignKey = (label, tableRef, Constraint*)
-$$
-$name : ForeignKey \rightarrow Label $ \\
-$name((lab, tRef, cons)) = lab $ \\ \\
-$reference : ForeignKey \rightarrow Label $ \\
-$reference((lab, tRef, cons)) = tRef $ \\ \\
-$reference : ForeignKey \rightarrow Constraint* $ \\
-$reference((lab, tRef, cons)) = cons $ \\ \\
-$owningTable : ForeignKey \times Database \rightarrow Table $ \\
-$owningTable(fk, d) = tab, tab \in tables(d) \wedge fk \in foreignKeys(d) $
-
-\paragraph{Primary key}
-$$
-PrimaryKey =  (label) 	
-$$
-
-\paragraph{DbType} represents primitive types in the database. There are usually defined types such as Char, Integer, Boolean etc. There is only one type defined in the model which is called DBSTRING.
-$$
-DbType = DBSTRING
-$$
-
-\paragraph{Constraint} there are two types of constraints defined in the model.
-Both constraints are column constraints - first constraint defines unique records in a column, second constraint defines non-empty columns.
-
-\begin{center}
-$Constraint = NOTNULL$ $|$ $UNIQUE $
-\end{center}
-
-\section{Object-Relational Mapping}
-There are defined functions for object-relational mapping used in the model.
-The purpose of a function is indicated in its name and should be obvious from its definition.
-These functions do not serve for full object-relational mapping, their purpose is to help with data evolution and its propagation from application to database.
-
-\subsection{Mapping of Types}
-The mapping of types assumes a bijection between application types and database types, otherwise these has to be additional information for type mapping.
-We focus on structural and data change therefore we simplify types and its mapping.
-
-\begin{center}
-$ \rho_{t} : AppType \rightarrow DbType $ \\
-$ \rho_{t}(APPSTRING) = DBSTRING $
-\end{center}
-
-\subsection{Mapping of Classes}
-We assume the primary key column is created automatically by database, therefore the primary key is always created with name Id and od NzNat type and properties with cardinality 0 or 1 are mapped into columns.
-The associations are ignored int this mapping as it is one of the partial mapping function used in eORM.
-
-\begin{center}
-$\rho_{c}: Class \rightarrow Table $ \\ 
-$\rho_{c}(c) = table(name(c), primaryKey("Id"), \rho_{p}(primitiveProperties(c), \varnothing $\\$ \cup \rho_{a}(associations(c)) \cup \rho_{p}(collectionProperties)$
-\end{center}
-
-\subsection{Mapping of Properties}
-There are two types of properties - the first type has cardinality 0 or 1 and it is mapped to a column (when the cardinality is equal to 1 the $NOTNULL$ constraint is added to a column).
-A property with cardinality greater than one has to be mapped int a table not a column, thus we have a special mapping case.
-
-%TODO split \rho function?
-
-$
-\rho_p : Property \times Database \rightarrow Column 
-$
-
-$\rho_p(p, d) = \begin{cases}
- \inference{cardinality(p) = 0 \\ \wedge \rho_c(owningClass(p)) \in tables(d)}{\begin{gathered}
- columns(table(name(owningClass(p)),d) \\ \cup column(name(p), \rho_t(type(p)), \varnothing)
-\end{gathered}
-} \\ \\
-
- \inference{cardinality(p) = 1 \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL)} \\ \\
-
- \inference{cardinality(p) > 2 \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered} table(name(p), primaryKey("Id"), column("value", \rho_t(type(p))), \\ foreignKey(name(p), name(owningClass(p))) \end{gathered}}
- \end{cases}$
-
-\subsection{Mapping of Associations}
-A association is mapped into database as a foreign key in table representing associating class or an association can be mapped as a coupling table in case the association cardinality is M:N or 1:N.
-
-$
-\rho_a(a,d) = \begin{cases}
- \inference{startCardinality(a) = 0 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{
- \begin{gathered}
- foreignKeys(u) \\ \cup foreignKey(name(a), name(t),  \varnothing) 
- \end{gathered}
- }
-  \\ \\
- \inference{startCardinality(a) = 1 \\\\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{\begin{gathered}
-foreignKeys(u) \\ \cup  foreignKey(name(a), name(t),  NOTNULL)
-\end{gathered}
-} \\ \\
-
-  \inference{startCardinality > 1 \\ \wedge \rho(owningClass(p)) \in tables(d) \wedge \rho(reference(a)) \in tables(d)}{
-  \begin{gathered}
- table(name(a), pk("Id"), \varnothing, \\ foreignKey(owningClass(a),owningClass(a), NOTNULL) \\ foreignkey(reference(a), reference(a), NOTNULL)) 
-  \end{gathered}}
-
- \end{cases}
-$
-
-%TODO handle instances
-
-\section{Data Evolution in ORM System}
-The Object-Relational system consist of an application and a database.
-The system has to be in consistent state, therefore the evolution of application has to be propagated correctly into database.
-Each evolution operation consist of two parts - applying evolution operation on application and on database.
-The mapping of the evolution change is called extended Object-Relational Mapping (eORM).
-The eORM mapping helps to assure consistency of the system by transforming application evolution into database schema and data.
-Each eORM function applied on an ORM system produces a new generation of the system.
-If the function cannot be applied on a system the system cannot be evolved by the function.
-
-This section defines eORM for each trivial operation evolving application from \ref{sec:appEvolution}, moreover there are described more complex cases of evolution.
-
-%TODO jak se postupně vyvíjí systém
-\subsection{Creating of a System}
-Creational operations does not have impact on stored data they change the structure of the application and database by adding new objects into model.
-
-\subsubsection{Add Class}
-
-$
-\rho_{addClass}(c, d) = \inference{\forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
-	\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) 
-}{
-\begin{gathered}
-tables(d) \cup \rho(c) \cup \rho(collectionProperties(c)) \\ \cup \rho(associations(c))
-\end{gathered}
-}
-$
-\subsubsection{Add Property}
-$
-\rho_{addProperty}(p, d) = 
-\inference{\exists t \in tables(d) : name(t) = name(owningClass(p))}{\rho(p,d)} 
-$
-\subsection{Add Association}
-$
-\rho_{AddAssociation}(a, d) = \inference{\exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) }{\rho(a,d)}
-$
-
-\subsection{Deconstruction of a System}
-\subsubsection{Remove Class}
-$
-\rho_{removeClass}(c,d) = \inference{\rho(c) \in tables(d) \\ \wedge \not \exists r \in data(d) : owningTable(r) = \rho(c) \\ \not \exists j \in tables(d), f \in foreignKeys(j) : reference(f) = \rho(c)  }{tables(d) \setminus \rho(c)}
-$
-%TODO remove by label
-
-\subsubsection{Remove Property}
-$
-\rho_{removeProperty}(l_c, l_p, d) = \begin{cases}
- \inference{\exists t \in tables(d) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_p  }{
-columns(t) \setminus c
-} \\ \\
-  \inference{\exists t \in tables(d) : name(t) = l_c \\ \wedge \not \exists c \in columns(t) : name(c) = l_p \\ \wedge \exists u \in tables(d) : name(u) = l_p }{tables(d) \setminus u }
- \end{cases}
-$
-
-\subsubsection{Remove Association}
-$
-\rho_{removeAssociation}(l_c, l_a, d) = \begin{cases}
- \inference{\exists t \in tables(d) : name(t) = l_c \wedge }{foreignKeys(\rho(reference(a))) \ }
-
- \inference{}{}
- \end{cases}
-$
-
-\end{document}

### 2012-12-16 doc/description.tex
created images, added mandatory attribute to Property, changed ORM

@@
 \usepackage{hyperref}
+\usepackage{graphicx}
 
@@
 \newpage
-\section{Evolution of ORM System}
-The ORM system consist of application and database.
+\section{ORM System}
+The ORM system represents an application created using object oriented language and relational database.
+The components of such an application and theirs relations are in Figure \ref{fig:appStructure}.
+We define the application in our model as:
+
+$$system = < Application, Database, \rho >$$
+$instantiated : Class \times Application \rightarrow Boolean $ \\
+
 The application and database has to be in consistent state according to a concrete ORM algorithm, otherwise the application cannot work properly.
-It means that during the evolution we have to respect a concrete ORM to produce consistent result.
-$$
-OrmSystem = (Application, Database)
-$$
-%TODO define consistency
+It means that during the evolution we have to respect a concrete ORM to produce consistent result too.
 
-There are several way how the evolution of ORM system can be implemented.
-The way when the application is evolved, database scheme re-genareted by an ORM framework and data migrated manually is the most common nowadays.
-The goal of our project is to minimize the manual work in the process of data evolution.
-We have following options:
+\begin{figure}
+\begin{center}
+	\includegraphics[scale=0.3]{./images/system}
+	\caption{The model of one generation of an ORM system consists of application persistent layer and relational database for storing instances.}
+\end{center}
+	\label{fig:appStructure}
+\end{figure}
+
 \begin{itemize}
-	\item Backward round-trip engineering can be used.
-It means the application and database is evolved and then the difference in structure is used to define data migration.
-This approach is inappropriate for situations where there is stored data as a change in database schema can be connected with more possible changes of data.
-Hence an additional information from user will be needed to migrate data correctly.
-    \item Forward engineering can be used in two ways.
-First approach evolves application, re-generate a database schema and generate a data migration according to definition of application evolution.
-Second approach uses the definition of application evolution as a source for evolution of database schema and data migration.
-In contrast this approach preserves the original database - it only alters its structure, whereas the first approach creates a new database.
+	\item \textbf{TODO define consistency}
+	\item \textbf{TODO: $\rho(Application) = Database$ show how database is build from Application}
 \end{itemize}
-All approaches can be used and all of them have their pros and cons.
-We decide to use forward engineering approach as it is a natural way of evolution.
-Next we decide to generate both migrations (for database schema and data) from definition of application evolution.
-This allows us to keep history of all layers and to verify the evolution on multiple levels.
-The algorithm is illustrated in Figure \ref{fig:algorithm}.
 
-The figure shows all important components of the domain.
+All important components of the domain are shown in Figure \ref{fig:appStructure}.
 There are three static components (static because they do not define behavior, but state of the system): an application, a database schema and stored data.
@@
 
-\section{Application Component}
-The application component represent an application created using object oriented language.
-It means the application is a set of instances which cooperates together.
+\section{Application Model}
+The model represents a persistent layer of an application created using object oriented language.
+It means the layer is a set of instances which cooperates together.
 Each instance is defined according to a class - the class defines the structure and relationships of the instance.
-The application component model consists of  classes defining the structure of application:
+The instances are stored in the relational database in our case, therefore we define only the structure of classes in our model:
 
 $$
-Application = <Class*, Instance*>
+Application = <Class*>
 $$
@@
 $classes(<cs, is>) = cs, cs \in Class* \wedge is \in Instance* $ \\ \\
-$instantiated : Class \times Application \rightarrow Boolean $ \\
-\\
 $class : Label \times Application \rightarrow Class  $ \\ 
-$class(l, a) = c, c \in classes(a) \wedge name(c) = l, l \in Label, a \in Application* $ 
+$class(l, a) = c, c \in classes(a) \wedge name(c) = l, l \in Label, a \in Application* $ \\
 %TODO: variable types
 
-\subsection{Application Structure Definition}
-The structure of the application is defined by five concepts. Each concept is defined and a set of manipulating functions is define to manipulate the application model easily.
+The structure consist not only of classes, the structure of the application is defined by five concepts.
+Each concept is defined and a set of manipulating functions is define to manipulate the application model easily.
 
@@
 
-$$Class = (label, Property*, Association*)$$
+$$Class = < label, Property*, Association* >$$
 $name : Class \rightarrow Label$ \\
@@
 $$
-Property = (label, AppType, DefaultValue, Cardinality)
+Property = < label, AppType, DefaultValue, Cardinality, Mandatory >
 $$
@@
 $cardinality((lab, t, val, n)) = n$ \\ \\
+$mandatory : Property \rightarrow Boolean $
+$mandatory(<lab, t, val, n, m>) = m $ \\ \\
 $owningClass : Property \times Application \rightarrow Class $ \\
@@
 $$
-$$
-NzNat = [0 - 9]*
-$$
-
-\section{Application Evolution}
-\label{sec:appEvolution}
-
-The evolution is defined as a set of operations which change structure of an application.
-The definition of an operation consists of two parts - a conditions of operation feasibility and operation impact on the application structure.
-We decide to use following transcription:
-
-$$
-operation = \inference{conditions$ $of$ $feasibility}{impact$ $on$ $application$ $structure}
-$$
-If the conditions of feasibility are not met, the operation cannot impact the application structure.
-%TODO monade
-
-\subsection{Application Creation}
-$$addClass: Class \times Application \rightarrow Application $$
-
-$
-addClass(c, a) = \inference{\forall s \in classes(A) : name(s) \neq name(c)}
-{classes(a) \cup c}
-$
-
-$$AddProperty : Label \times Property \times Application \rightarrow Application $$
-
-$
-AddProperty(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in properties(p) : name(q) \neq name(p)}{
- properties(c) \cup p }
-$
-$$AddAssociation : Label \times Association \times Application \rightarrow Application $$
-$
-AddAssociation(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(c) \cup s}
-$
-
-\subsection{Application Deconstruction}
-$$RemoveClass: Label \times Application \rightarrow Application $$
-$
-RemoveClass(q, a) = \inference{\exists c \in classes(a) : name(c) = q \\
-\wedge not$ $instantiated(c) = \varnothing \wedge not$ $associated(c)}{classes(a) \setminus c }
-$
-
-$$RemoveProperty: Label \times Label \times Application \rightarrow Application $$
-
-$
-RemoveProperty(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{properties(c) \setminus p }
-$
-
-$$RemoveAssociation : Label \times Labels \times Application \rightarrow Application $$
-
-$
-RemoveAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists s \in associations(c) : name(s) = l_a }{associations(c) \setminus s}
-$
 
@@
 
-\section{Database Component Model}
-The relational database consists of database schema which defines the structure of the database and data. The database is a pair od sets:
+\section{Database Model}
+The relational database consists of database schema which defines the structure of the database and data, which in our ORM system represents instances. The database is a pair:
 $$
-Database = (table*, data*)
+Database = <table*, data*>
 $$
@@
 
+\subsection{Stored Data Model}
+
 \section{Object-Relational Mapping}
@@
 \subsection{Mapping of Properties}
-There are two types of properties - the first type has cardinality 0 or 1 and it is mapped to a column (when the cardinality is equal to 1 the $NOTNULL$ constraint is added to a column).
+There are two types of properties - the first type has cardinality 1  and it is mapped to a column.
 A property with cardinality greater than one has to be mapped int a table not a column, thus we have a special mapping case.
-
-%TODO split \rho function?
+When the property is mandatory the $NOTNULL$ constraint is added to the column with property value.
+This constraint affect already stored instances a new mandatory value has to be added to each instance.
 
@@
 $\rho_p(p, d) = \begin{cases}
- \inference{d \vdash cardinality(p) = 0 \\ \wedge \rho_c(owningClass(p)) \in tables(d)}{\begin{gathered} d \vdash 
- columns(table(name(owningClass(p)),d) \\ \cup column(name(p), \rho_t(type(p)), \varnothing)
-\end{gathered}
-} \\ \\
-
- \inference{ d \vdash cardinality(p) = 1 
+  \inference{ d \vdash cardinality(p) = 1 \wedge ! mandatory(p)
  \\ \wedge \exists t : t = table(name(owningClass(p)), d) 
  \\ \wedge \not instanciated(owningClass(p))
- }{ d \vdash columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL)} \\ \\
+ }{ d \vdash columns(t) \cup column(name(p), \rho_t(type(p)), \varnothing )} \\ \\
 
- \inference{d \vdash cardinality(p) > 2 \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}d \vdash  table(name(p), primaryKey("Id"), column("value", \rho_t(type(p))), \\ foreignKey(name(p), name(owningClass(p))) \end{gathered}}
+  \inference{ d \vdash cardinality(p) = 1 \wedge mandatory(p)
+ \\ \wedge \exists t : t = table(name(owningClass(p)), d) 
+ \\ \wedge ! instanciated(owningClass(p))
+ }{ d \vdash columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL )} 
+\\ \\
+
+ \inference{d \vdash cardinality(p) > 1 \wedge !
+mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}d \vdash  table(name(p), primaryKey("Id"), column("value", \rho_t(type(p))), \\ foreignKey(name(p), name(owningClass(p)), \varnothing) \end{gathered}}
+\\ \\
+
+ \inference{d \vdash cardinality(p) > 1 \wedge mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge !
+instanciated(owningClass(p))}{\begin{gathered}d \vdash  table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), NOTNULL), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL) \end{gathered}}
+\\
  \end{cases}$
-
-The new foreign key can cause inconsistency of the system when added to a table which contains data.
-This case is not handled by the mapping function above.
-We have to extend the function definition of a feature that allows the mapping between the data in both tables.
-The mapping is a function:
-$$f_m : Property \rightarrow Instance \times Instance $$
+\textbf{TODO: add value into instances in NOTNULL columns}
 
@@
   \end{gathered}}  
-
  \end{cases}
@@
 
+The new foreign key can cause inconsistency of the system when added to a table which contains data.
+This case is not handled by the mapping function above.
+We have to extend the function definition of a feature that allows the mapping between the data in both tables.
+The mapping is a function:
+$$f_m : Property \rightarrow Instance \times Instance $$
 %TODO handle instances
@@
 
+The state of a system is called generation and transformation from one state to another is called evolution.
+The state of the system defines a context for each evolution and it determines if the evolution is successful or not.
+
+\begin{figure}
+\begin{center}
+	\includegraphics[scale=0.4]{./images/evolution}
+	\caption{The evolution of data changes the system on all levels}
+\end{center}
+	\label{fig:evolution}
+\end{figure}
+
 This section defines eORM for each trivial operation evolving application from \ref{sec:appEvolution}, moreover there are described more complex cases of evolution.
 
-%TODO jak se postupně vyvíjí systém
-\subsection{Creating of a System}
+\begin{itemize}
+	\item \textbf{TODO show how a whole system is evolved}
+	\item \textbf{TODO state monade}
+\end{itemize}
+
+\subsection{Evolution of Application}
+\label{sec:appEvolution}
+
+The evolution is defined as a set of operations which change structure of an application.
+The definition of an operation consists of two parts - a conditions of operation feasibility and operation impact on the application structure.
+We decide to use following transcription:
+
+$$
+operation = \inference{conditions$ $of$ $feasibility}{impact$ $on$ $application$ $structure}
+$$
+If the conditions of feasibility are not met, the operation cannot impact the application structure.
+%TODO monade
+
+\subsubsection{Application Creation}
+$$addClass: Class \times Application \rightarrow Application $$
+
+$
+addClass(c, a) = \inference{\forall s \in classes(A) : name(s) \neq name(c)}
+{classes(a) \cup c}
+$
+
+$$AddProperty : Label \times Property \times Application \rightarrow Application $$
+
+$
+AddProperty(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in properties(p) : name(q) \neq name(p)}{
+ properties(c) \cup p }
+$
+$$AddAssociation : Label \times Association \times Application \rightarrow Application $$
+$
+AddAssociation(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(c) \cup s}
+$
+
+\subsubsection{Application Deconstruction}
+$$RemoveClass: Label \times Application \rightarrow Application $$
+$
+RemoveClass(q, a) = \inference{\exists c \in classes(a) : name(c) = q \\
+\wedge not$ $instantiated(c) = \varnothing \wedge not$ $associated(c)}{classes(a) \setminus c }
+$
+
+$$RemoveProperty: Label \times Label \times Application \rightarrow Application $$
+
+$
+RemoveProperty(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{properties(c) \setminus p }
+$
+
+$$RemoveAssociation : Label \times Labels \times Application \rightarrow Application $$
+
+$
+RemoveAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists s \in associations(c) : name(s) = l_a }{associations(c) \setminus s}
+$
+
+\subsubsection{Application Alternation}
+\paragraph{Rename}
+\paragraph{Change Cardinality}
+\paragraph{Copy Property}
+
+\subsection{ORM System Evolution}
 Creational operations does not have impact on stored data they change the structure of the application and database by adding new objects into model.
 
-\subsubsection{Add Class}
+\subsubsection{Creation of a System}
+\paragraph{Add Class}
 
@@
 $
-\subsubsection{Add Property}
+\paragraph{Add Property}
 $
@@
 $
-\subsection{Add Association}
+\paragraph{Add Association}
 $
@@
 
-\subsection{Deconstruction of a System}
-\subsubsection{Remove Class}
+\subsubsection{Deconstruction of a System}
+\paragraph{Remove Class}
 $
@@
 
-\subsubsection{Remove Property}
+\paragraph{Remove Property}
 $
@@
 
-\subsubsection{Remove Association}
+\paragraph{Remove Association}
 $
-\Phi_{removeAssociation}(l_c, l_a, d) = \begin{cases}
- \inference{d \vdash \exists t \in tables(d) : name(t) = l_c \wedge }{d \vdash foreignKeys(\rho(reference(a))) \ }
-
- \inference{}{}
+\Phi_{removeAssociation}(l_c, l_a, d) = d \vdash \begin{cases}
+ \inference{\exists t \in tables(d) : name(t) = l_c 
+ \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_a 
+ \\ \wedge not containsData(t)}{d \vdash foreignKeys(t) \setminus f }
+ \\ \\
+ \inference{\exists t \in tables(d) : name(t) = l_a 
+ \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_c
+ \\ \wedge not containsData(t)}{d \vdash tables(d) \setminus t}
  \end{cases}
@@
 
+\subsubsection{Altering of a System}
+
 \end{document}

### 2012-12-19 doc/description.tex
a lot of fixes,  addProperty with data

@@
 $classes : Application \rightarrow Class*$ \\
-$classes(<cs, is>) = cs, cs \in Class* \wedge is \in Instance* $ \\ \\
+$classes(<cs, is >) = = cs, cs \in Class* \wedge is \in Instance* $ \\ \\
 $class : Label \times Application \rightarrow Class  $ \\ 
@@
 $name : Class \rightarrow Label$ \\
-$name((lab, props, assocs)) = lab$ \\ \\
+$name(< lab, props, assocs  >) = lab$ \\ \\
 $properties : Class \rightarrow Property*$ \\
-$properties((lab, props, assocs)) = props $\\ \\
+$properties(< lab, props, assocs  >) = props $\\ \\
 $primitiveProperties : Class \rightarrow Property*$ \\
-$primitiveProperties((lab, props, assocs)) = props', \forall p \in props' : cardinality(p) \leq 1 $\\ \\
+$primitiveProperties(< lab, props, assocs  >) = props', \forall p \in props' : cardinality(p) \leq 1 $\\ \\
 $collectionProperties : Class \rightarrow Property*$ \\
-$collectionProperties((lab, props, assocs)) = props', \forall p \in props' : cardinality(p) > 1 $\\ \\
+$collectionProperties(< lab, props, assocs  >) = props', \forall p \in props' : cardinality(p) > 1 $\\ \\
 $associations : Class \rightarrow Association*$ \\
-$associations((lab, props, assocs)) = assocs $ \\ \\
+$associations(< lab, props, assocs  >) = assocs $ \\ \\
 $associated :  Class \times Application \rightarrow Boolean $
@@
 $name : Property \rightarrow Label$ \\
-$name((lab, t, val, n)) = lab$ \\ \\
+$name(< lab, t, val, n  >) = lab$ \\ \\
 $type : Property \times AppType$ \\
-$type((lab, t, val, n)) = t$ \\ \\
+$type(< lab, t, val, n  >) = t$ \\ \\
 $cardinality : Property \rightarrow NzNat$ \\
-$cardinality((lab, t, val, n)) = n$ \\ \\
+$cardinality(< lab, t, val, n  >) = n$ \\ \\
 $mandatory : Property \rightarrow Boolean $
-$mandatory(<lab, t, val, n, m>) = m $ \\ \\
+$mandatory(<lab, t, val, n, m >) = = m $ \\ \\
 $owningClass : Property \times Application \rightarrow Class $ \\
@@
 $name : Association \rightarrow Label$ \\
-$name((lab, ref, n_1, n_2)) = lab$\\ \\
+$name(< lab, ref, n_1, n_2  >) = lab$\\ \\
 $reference : Association \rightarrow Label$ \\
-$reference((lab, ref, n_1, n_2)) = ref$\\ \\
+$reference(< lab, ref, n_1, n_2  >) = ref$\\ \\
 $startCardinality : Association \rightarrow NzNat$ \\
-$startCardinality((lab, ref, n_1, n_2)) = n_1$\\ \\
+$startCardinality(< lab, ref, n_1, n_2  >) = n_1$\\ \\
 $endCardinality : Association \rightarrow NzNat$ \\
-$endCardinality((lab, ref, n_1, n_2)) = n_2$ \\ \\
+$endCardinality(< lab, ref, n_1, n_2  >) = n_2$ \\ \\
 $owningClass : Association \times Application \rightarrow Class $ \\
@@
 $$
-Database = <table*, data*>
+Database = <table*, row*>
 $$
@@
 $name : Table \rightarrow Label $ \\
-$name((lab, pk, cols, fks)) = lab$ \\ \\
+$name(< lab, pk, cols, fks  >) = lab$ \\ \\
 $primaryKey : Table \rightarrow PrimaryKey $ \\
-$primaryKey((lab, pk, cols, fks)) = pk$ \\ \\
+$primaryKey(< lab, pk, cols, fks  >) = pk$ \\ \\
 $columns : Table \rightarrow Column* $ \\
-$columns((lab, pk, cols, fks)) = cols$ \\ \\
+$columns(< lab, pk, cols, fks  >) = cols$ \\ \\
 $foreignKeys : Table \rightarrow ForeignKey* $ \\
-$foreignKeys((lab, pk, cols, fks)) = fks$
+$foreignKeys(< lab, pk, cols, fks  >) = fks$
 
@@
 $name : Column \rightarrow Label $ \\
-$name((lab, t, val, cons)) = lab $ \\ \\
+$name(< lab, t, val, cons  >) = lab $ \\ \\
 $type : Column \rightarrow DbType $ \\
-$type((lab, t, val, cons)) = t $ \\ \\
+$type(< lab, t, val, cons  >) = t $ \\ \\
 $constraints : Column \rightarrow Constraint* $ \\
-$constraints((lab, t, val, cons)) = cons $ \\ \\
+$constraints(< lab, t, val, cons  >) = cons $ \\ \\
 $owningTable : Column \times Database \rightarrow Table $ \\
@@
 $name : ForeignKey \rightarrow Label $ \\
-$name((lab, tRef, cons)) = lab $ \\ \\
+$name(< lab, tRef, cons  >) = lab $ \\ \\
 $reference : ForeignKey \rightarrow Label $ \\
-$reference((lab, tRef, cons)) = tRef $ \\ \\
+$reference(< lab, tRef, cons  >) = tRef $ \\ \\
 $reference : ForeignKey \rightarrow Constraint* $ \\
-$reference((lab, tRef, cons)) = cons $ \\ \\
+$reference(< lab, tRef, cons  >) = cons $ \\ \\
 $owningTable : ForeignKey \times Database \rightarrow Table $ \\
@@
 $$
-PrimaryKey =  (label) 	
+PrimaryKey =  < label > 	
 $$
@@
 
-\subsection{Stored Data Model}
+\subsection{Model of Stored Data}
+The representation of stored data in the relational database is based on pairs:
+$$pair = < refColumn, value > | < refFK, value>$$
+
+\textbf{TODO: primary key}
+where the $refColumn$ reference a column or foreign key by its name and $value$ represents concrete value stored in the database. A row in a table is a set of pairs with reference to a concrete table:
+
+$$row = < refTable, pair* >$$
+
+\textbf{TODO: define consistency of rows and schema}
 
@@
  \\ \wedge \not instanciated(owningClass(p))
- }{ d \vdash columns(t) \cup column(name(p), \rho_t(type(p)), \varnothing )} \\ \\
+ }{\begin{gathered}
+	 d \vdash columns(t) \cup column(name(p), \rho_t(type(p)), \varnothing )
+\end{gathered}
+ } \\ \\
 
   \inference{ d \vdash cardinality(p) = 1 \wedge mandatory(p)
- \\ \wedge \exists t : t = table(name(owningClass(p)), d) 
- \\ \wedge ! instanciated(owningClass(p))
- }{ d \vdash columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL )} 
+ \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}
+ 	d \vdash columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL ) \\
+
+\end{gathered}
+} 
 \\ \\
@@
  \inference{d \vdash cardinality(p) > 1 \wedge !
-mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}d \vdash  table(name(p), primaryKey("Id"), column("value", \rho_t(type(p))), \\ foreignKey(name(p), name(owningClass(p)), \varnothing) \end{gathered}}
+mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}d \vdash  table(name(p), primaryKey("Id"), column("value", \rho_t(type(p))), \\ foreignKey(name(p), name(owningClass(p)), \varnothing) \\
+
+\end{gathered}}
 \\ \\
 
- \inference{d \vdash cardinality(p) > 1 \wedge mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge !
-instanciated(owningClass(p))}{\begin{gathered}d \vdash  table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), NOTNULL), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL) \end{gathered}}
-\\
+ \inference{d \vdash cardinality(p) > 1 \wedge mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d) }{\begin{gathered}d \vdash  table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), NOTNULL), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL)
+	 \end{gathered}}
  \end{cases}$
-\textbf{TODO: add value into instances in NOTNULL columns}
 
@@
   \\ \\
- \inference{d \vdash startCardinality(a) = 1 \\\\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{\begin{gathered} d \vdash 
-foreignKeys(u) \\ \cup  foreignKey(name(a), name(t),  NOTNULL)
+ \inference{d \vdash startCardinality(a) = 1 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{\begin{gathered} d \vdash 
+foreignKeys(u) \\ \cup  foreignKey(name(a), name(t),  NOTNULL) \\
 \end{gathered}
@@
 $
+\textbf{TODO: check + add cardinalities/images and split(?), data in NOTNULL}
 
@@
 \begin{center}
-	\includegraphics[scale=0.4]{./images/evolution}
+	\includegraphics[scale=0.4]{./images/evolution_simple}
 	\caption{The evolution of data changes the system on all levels}
@@
 $
-AddProperty(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in properties(p) : name(q) \neq name(p)}{
- properties(c) \cup p }
+AddProperty(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(p) : name(q) \neq name(p)}{properties(c) \cup p }
 $
@@
 $
-\Phi_{addProperty}(p, d) = 
-\inference{d \vdash \exists t \in tables(d) : name(t) = name(owningClass(p))}{d \vdash \rho(p,d)} 
+\Phi_{addProperty}(p, d) = \begin{cases}
+\inference{d \vdash \exists t \in tables(d) : name(t) = name(owningClass(p)) \wedge (! mandatory(p) \vee ! instantiated(owningClass(p))}{d \vdash \rho(p,d)} 
+\\ \\ 
+\inference{d \vdash \exists t \in tables(d) : name(t) = name(owningClass(p)) \wedge mandatory(p) \wedge instantiated(owningClass(p)) \wedge cardinality(p) \leq 1}{d \vdash \rho(p,d) \wedge \forall r \in data(d), reference(r) = t : pair(r) \cup pair(name(p), defaultValue(p))}
+\\ \\
+\inference{d \vdash \exists t \in tables(d) : name(t) = name(owningClass(p)) \wedge mandatory(p) \wedge instantiated(owningClass(p)) \wedge cardinality(p) \leq 1}{\begin{gathered}
+ \rho(p,d) \wedge \forall r \in row(owningClass(p)) : data(d)\\ \cup row(name(p), pair("value", defaultValue(p)), \\ pair(name(p), value(primaryKey(r))) 
+\end{gathered}
+}
+
+\end{cases}
 $
+
 \paragraph{Add Association}

### 2012-12-19 doc/description.tex
simplified orm of classes
updated eorm of addClass

@@
 $\rho_{c}: Class \rightarrow Table $ \\ 
-$\rho_{c}(c,d) = d \vdash table(name(c), primaryKey("Id"), \rho_{p}(primitiveProperties(c), \varnothing $\\$ \cup \rho_{a}(associations(c)) \cup \rho_{p}(collectionProperties)$
+$\rho_{c}(c,d) = tables(d) \cup table(name(c), primaryKey("Id"), \varnothing, \varnothing) $
 \end{center}
@@
 $\rho_p(p, d) = \begin{cases}
-  \inference{ d \vdash cardinality(p) = 1 \wedge ! mandatory(p)
+  \inference{cardinality(p) = 1 \wedge ! mandatory(p)
  \\ \wedge \exists t : t = table(name(owningClass(p)), d) 
@@
  }{\begin{gathered}
-	 d \vdash columns(t) \cup column(name(p), \rho_t(type(p)), \varnothing )
+	  columns(t) \cup column(name(p), \rho_t(type(p)), \varnothing )
 \end{gathered}
@@
 
-  \inference{ d \vdash cardinality(p) = 1 \wedge mandatory(p)
+  \inference{cardinality(p) = 1 \wedge mandatory(p)
  \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}
- 	d \vdash columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL ) \\
+ 	 columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL ) \\
 
@@
 
- \inference{d \vdash cardinality(p) > 1 \wedge !
-mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}d \vdash  table(name(p), primaryKey("Id"), column("value", \rho_t(type(p))), \\ foreignKey(name(p), name(owningClass(p)), \varnothing) \\
+ \inference{ cardinality(p) > 1 \wedge !
+mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}tables(d) \cup  table(name(p), primaryKey("Id"), column("value", \rho_t(type(p))), \\ foreignKey(name(p), name(owningClass(p)), \varnothing) \\
 
@@
 
- \inference{d \vdash cardinality(p) > 1 \wedge mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d) }{\begin{gathered}d \vdash  table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), NOTNULL), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL)
+ \inference{ cardinality(p) > 1 \wedge mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d) }{\begin{gathered} tables(d) \cup  table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), NOTNULL), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL)
 	 \end{gathered}}
@@
 \rho_a(a,d) = \begin{cases}
- \inference{d \vdash startCardinality(a) = 0 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{d \vdash 
+ \inference{ startCardinality(a) = 0 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{ 
  \begin{gathered}
@@
   \\ \\
- \inference{d \vdash startCardinality(a) = 1 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{\begin{gathered} d \vdash 
+ \inference{ startCardinality(a) = 1 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{\begin{gathered}  
 foreignKeys(u) \\ \cup  foreignKey(name(a), name(t),  NOTNULL) \\
@@
 
-  \inference{d \vdash  startCardinality > 1 \\ \wedge \rho(owningClass(p)) \in tables(d) \wedge \rho(reference(a)) \in tables(d)}{
-  \begin{gathered} d \vdash 
+  \inference{  startCardinality > 1 \\ \wedge \rho(owningClass(p)) \in tables(d) \wedge \rho(reference(a)) \in tables(d)}{
+  \begin{gathered}  
  table(name(a), pk("Id"), \varnothing, \\ foreignKey(owningClass(a),owningClass(a), NOTNULL) \\ foreignkey(reference(a), reference(a), NOTNULL)) 
@@
 $
-\textbf{TODO: check + add cardinalities/images and split(?), data in NOTNULL}
 
@@
 \subsubsection{Creation of a System}
-\paragraph{Add Class}
+\subsubsection{Add Class}
 
 $
-\Phi_{addClass}(c, d) = \inference{d \vdash \forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
+\Phi_{addClass}(c, d) = \inference{ \forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
 	\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) 
@@
 \begin{gathered}
-d \vdash  tables(d) \cup \rho(c) \cup \rho(collectionProperties(c)) \\ \cup \rho(associations(c))
+  \rho(c) \wedge \rho(singleProperties(c)) \wedge \rho(collectionProperties(c)) \\ \wedge \rho(associations(c))
 \end{gathered}
@@
 $
-\paragraph{Add Property}
+\subsubsection{Add Property}
 $
 \Phi_{addProperty}(p, d) = \begin{cases}
-\inference{d \vdash \exists t \in tables(d) : name(t) = name(owningClass(p)) \wedge (! mandatory(p) \vee ! instantiated(owningClass(p))}{d \vdash \rho(p,d)} 
+\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \wedge (! mandatory(p) \vee ! instantiated(owningClass(p))}{ \rho(p,d)} 
 \\ \\ 
-\inference{d \vdash \exists t \in tables(d) : name(t) = name(owningClass(p)) \wedge mandatory(p) \wedge instantiated(owningClass(p)) \wedge cardinality(p) \leq 1}{d \vdash \rho(p,d) \wedge \forall r \in data(d), reference(r) = t : pair(r) \cup pair(name(p), defaultValue(p))}
+\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \wedge mandatory(p) \wedge instantiated(owningClass(p)) \wedge cardinality(p) \leq 1}{ \rho(p,d) \wedge \forall r \in data(d), reference(r) = t : pair(r) \cup pair(name(p), defaultValue(p))}
 \\ \\
-\inference{d \vdash \exists t \in tables(d) : name(t) = name(owningClass(p)) \wedge mandatory(p) \wedge instantiated(owningClass(p)) \wedge cardinality(p) \leq 1}{\begin{gathered}
+\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \wedge mandatory(p) \wedge instantiated(owningClass(p)) \wedge cardinality(p) \leq 1}{\begin{gathered}
  \rho(p,d) \wedge \forall r \in row(owningClass(p)) : data(d)\\ \cup row(name(p), pair("value", defaultValue(p)), \\ pair(name(p), value(primaryKey(r))) 
@@
 
-\paragraph{Add Association}
+\subsubsection{Add Association}
 $
-\Phi_{AddAssociation}(a, d) = \inference{d \vdash \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) }{d \vdash \rho(a,d)}
+\Phi_{AddAssociation}(a, d) = \inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) }{ \rho(a,d)}
 $
@@
 \subsubsection{Deconstruction of a System}
-\paragraph{Remove Class}
+\subsubsection{Remove Class}
 $
-\Phi_{removeClass}(c,d) = \inference{d \vdash \rho(c) \in tables(d) \\ \wedge \not \exists r \in data(d) : owningTable(r) = \rho(c) \\ \not \exists j \in tables(d), f \in foreignKeys(j) : reference(f) = \rho(c)  }{d \vdash tables(d) \setminus \rho(c)}
+\Phi_{removeClass}(c,d) = \inference{ \rho(c) \in tables(d) \\ \wedge \not \exists r \in data(d) : owningTable(r) = \rho(c) \\ \not \exists j \in tables(d), f \in foreignKeys(j) : reference(f) = \rho(c)  }{ tables(d) \setminus \rho(c)}
 $
@@
 
-\paragraph{Remove Property}
+\subsubsection{Remove Property}
 $
 \Phi_{removeProperty}(l_c, l_p, d) = \begin{cases}
- \inference{d \vdash \exists t \in tables(d) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_p  }{d \vdash
+ \inference{ \exists t \in tables(d) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_p  }{d \vdash
 columns(t) \setminus c
 } \\ \\
-  \inference{d \vdash \exists t \in tables(d) : name(t) = l_c \\ \wedge \not \exists c \in columns(t) : name(c) = l_p \\ \wedge \exists u \in tables(d) : name(u) = l_p }{d \vdash tables(d) \setminus u }
+  \inference{ \exists t \in tables(d) : name(t) = l_c \\ \wedge \not \exists c \in columns(t) : name(c) = l_p \\ \wedge \exists u \in tables(d) : name(u) = l_p }{ tables(d) \setminus u }
  \end{cases}
@@
 
-\paragraph{Remove Association}
+\subsubsection{Remove Association}
 $
-\Phi_{removeAssociation}(l_c, l_a, d) = d \vdash \begin{cases}
+\Phi_{removeAssociation}(l_c, l_a, d) =  \begin{cases}
  \inference{\exists t \in tables(d) : name(t) = l_c 
  \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_a 
- \\ \wedge not containsData(t)}{d \vdash foreignKeys(t) \setminus f }
+ \\ \wedge not containsData(t)}{ foreignKeys(t) \setminus f }
  \\ \\
@@
  \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_c
- \\ \wedge not containsData(t)}{d \vdash tables(d) \setminus t}
+ \\ \wedge not containsData(t)}{ tables(d) \setminus t}
  \end{cases}

### 2012-12-19 doc/description.tex
reformated eorm add property

@@
 \Phi_{addProperty}(p, d) = \begin{cases}
-\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \wedge (! mandatory(p) \vee ! instantiated(owningClass(p))}{ \rho(p,d)} 
+\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge (! mandatory(p) \vee ! instantiated(owningClass(p))}{ \rho(p,d)} 
 \\ \\ 
-\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \wedge mandatory(p) \wedge instantiated(owningClass(p)) \wedge cardinality(p) \leq 1}{ \rho(p,d) \wedge \forall r \in data(d), reference(r) = t : pair(r) \cup pair(name(p), defaultValue(p))}
+\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
+ \rho(p,d) \wedge \forall r \in data(d), reference(r) = t : pair(r) \\ \cup pair(name(p), defaultValue(p))
+\end{gathered}
+} 
 \\ \\
-\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \wedge mandatory(p) \wedge instantiated(owningClass(p)) \wedge cardinality(p) \leq 1}{\begin{gathered}
+\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
  \rho(p,d) \wedge \forall r \in row(owningClass(p)) : data(d)\\ \cup row(name(p), pair("value", defaultValue(p)), \\ pair(name(p), value(primaryKey(r))) 

### 2012-12-19 doc/description.tex
updated addAssociation orm

@@
 $
-\Phi_{AddAssociation}(a, d) = \inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) }{ \rho(a,d)}
+\Phi_{AddAssociation}(a, d) = \begin{cases}
+\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) }{ \rho(a,d) }
+ \end{cases}
 $

### 2012-12-20 doc/description.tex
structural changes in document

@@
 We call this component evolution object-relational mapping (eORM).
+
+\section{Data Evolution in ORM System}
+The Object-Relational system consist of an application and a database.
+The system has to be in consistent state, therefore the evolution of application has to be propagated correctly into database.
+Each evolution operation consist of two parts - applying evolution operation on application and on database.
+The mapping of the evolution change is called extended Object-Relational Mapping (eORM).
+The eORM mapping helps to assure consistency of the system by transforming application evolution into database schema and data.
+Each eORM function applied on an ORM system produces a new generation of the system.
+If the function cannot be applied on a system the system cannot be evolved by the function.
+
+The state of a system is called generation and transformation from one state to another is called evolution.
+The state of the system defines a context for each evolution and it determines if the evolution is successful or not.
+
+\begin{figure}
+\begin{center}
+	\includegraphics[scale=0.4]{./images/evolution_simple}
+	\caption{The evolution of data changes the system on all levels}
+\end{center}
+	\label{fig:evolution}
+\end{figure}
+
+\begin{itemize}
+	\item \textbf{TODO show how a whole system is evolved}
+	\item \textbf{TODO state monade}
+\end{itemize}
 
@@
  \inference{ cardinality(p) > 1 \wedge !
-mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}tables(d) \cup  table(name(p), primaryKey("Id"), column("value", \rho_t(type(p))), \\ foreignKey(name(p), name(owningClass(p)), \varnothing) \\
+mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}tables(d) \cup  table(name(p), primaryKey("Id"), column("value", \rho_t(type(p)), \varnothing), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL) \\
 
@@
 
-\section{Data Evolution in ORM System}
-The Object-Relational system consist of an application and a database.
-The system has to be in consistent state, therefore the evolution of application has to be propagated correctly into database.
-Each evolution operation consist of two parts - applying evolution operation on application and on database.
-The mapping of the evolution change is called extended Object-Relational Mapping (eORM).
-The eORM mapping helps to assure consistency of the system by transforming application evolution into database schema and data.
-Each eORM function applied on an ORM system produces a new generation of the system.
-If the function cannot be applied on a system the system cannot be evolved by the function.
-
-The state of a system is called generation and transformation from one state to another is called evolution.
-The state of the system defines a context for each evolution and it determines if the evolution is successful or not.
-
-\begin{figure}
-\begin{center}
-	\includegraphics[scale=0.4]{./images/evolution_simple}
-	\caption{The evolution of data changes the system on all levels}
-\end{center}
-	\label{fig:evolution}
-\end{figure}
-
-This section defines eORM for each trivial operation evolving application from \ref{sec:appEvolution}, moreover there are described more complex cases of evolution.
-
-\begin{itemize}
-	\item \textbf{TODO show how a whole system is evolved}
-	\item \textbf{TODO state monade}
-\end{itemize}
-
+\section{Evolution}
 \subsection{Evolution of Application}

### 2012-12-20 doc/description.tex
updated addAsociation so it handles data too

@@
 
+\title{Evolution of ORM Systems: Formal Foundations}
+\author{Ondrej Macek}
+
 \begin{document}
-\tableofcontents
+\tableofcontents \newpage
 
-\newpage
+\maketitle
 \section{ORM System}
@@
 $
-\Phi_{AddAssociation}(a, d) = \begin{cases}
-\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) }{ \rho(a,d) }
+\Phi_{AddAssociation}(a, f_m, d) = \begin{cases}
+\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
+\rho(a,d) \wedge \forall r \in f_m(owningClass(a), reference(a)) : \\ data(t_t) \cup row(name(t_s), pair(name(a), \\ value(primaryKey(first(r)))) 
+\end{gathered}
+ }
+ \\ \\
+\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) > 1}{\begin{gathered}
+\rho(a,d) \wedge \forall r \in f_m(owningClass(a), reference(a)),\\ t \in tables(d) : name(t) = name(a) : \\ data(t) \cup row(name(a), pair(name(t_s), \\ value(primaryKey(first(r)))) pair(name(t_t), \\ value(primaryKey(second(r)))))
+\end{gathered}
+ }
+
  \end{cases}
@@
 
-\subsubsection{Altering of a System}
+\subsubsection{Altering a System}
 

### 2012-12-20 doc/description.tex
updates in removes

@@
 \subsubsection{Creation of a System}
-\subsubsection{Add Class}
+\paragraph{Add Class}
 
-$
+$$
 \Phi_{addClass}(c, d) = \inference{ \forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
@@
 }
-$
-\subsubsection{Add Property}
-$
+$$
+\paragraph{Add Property}
+$$
 \Phi_{addProperty}(p, d) = \begin{cases}
@@
 \end{cases}
-$
+$$
 
-\subsubsection{Add Association}
-$
+\paragraph{Add Association}
+$$
 \Phi_{AddAssociation}(a, f_m, d) = \begin{cases}
@@
  \end{cases}
-$
+$$
 
 \subsubsection{Deconstruction of a System}
-\subsubsection{Remove Class}
-$
+\paragraph{Remove Class}
+$$
 \Phi_{removeClass}(c,d) = \inference{ \rho(c) \in tables(d) \\ \wedge \not \exists r \in data(d) : owningTable(r) = \rho(c) \\ \not \exists j \in tables(d), f \in foreignKeys(j) : reference(f) = \rho(c)  }{ tables(d) \setminus \rho(c)}
-$
+$$
 %TODO remove by label
 
-\subsubsection{Remove Property}
-$
-\Phi_{removeProperty}(l_c, l_p, d) = \begin{cases}
- \inference{ \exists t \in tables(d) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_p  }{d \vdash
-columns(t) \setminus c
+\paragraph{Remove Property}
+$$
+\Phi_{removeProperty}(p, d) = \begin{cases}
+ \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \exists c \in columns(t) : name(c) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p)) }{columns(t) \setminus c
 } \\ \\
-  \inference{ \exists t \in tables(d) : name(t) = l_c \\ \wedge \not \exists c \in columns(t) : name(c) = l_p \\ \wedge \exists u \in tables(d) : name(u) = l_p }{ tables(d) \setminus u }
+  \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \not \exists c \in columns(t) : name(c) = name(p) \\ \wedge \exists u \in tables(d) : name(u) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p))}{ tables(d) \setminus u }
  \end{cases}
-$
+$$
 
-\subsubsection{Remove Association}
-$
-\Phi_{removeAssociation}(l_c, l_a, d) =  \begin{cases}
- \inference{\exists t \in tables(d) : name(t) = l_c 
- \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_a 
- \\ \wedge not containsData(t)}{ foreignKeys(t) \setminus f }
+\paragraph{Remove Association}
+$$
+\Phi_{removeAssociation}(a, d) =  \begin{cases}
+ \inference{\exists t \in tables(d) : name(t) = name(owningClass(a))
+ \\ \wedge \exists f \in foreignKeys(t) : name(f) = name(a)
+ \\ \wedge \forall r \in data(d) : owningTable(r) \neq t}{ foreignKeys(t) \setminus f }
  \\ \\
- \inference{\exists t \in tables(d) : name(t) = l_a 
- \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_c
- \\ \wedge not containsData(t)}{ tables(d) \setminus t}
+ \inference{\exists t \in tables(d) : name(t) = name(a)
+ \\ \wedge \exists f \in foreignKeys(t) : name(f) = name(owningClass(a))
+ \\ \wedge \forall r \in data(d) : owningTable(r) \neq t}{ tables(d) \setminus t}
  \end{cases}
-$
+$$
 

### 2012-12-20 doc/description.tex
added texts

@@
 \maketitle
+\begin{abstract}
+The evolution of an application is a common issue during software development.
+The formal description of the evolution operations in context of  software created using an object oriented programming language and a relational database is introduced in this paper.
+Special attention is given to the data stored in databases (instances) and theirs consistency during evolution process.
+\end{abstract}
+
+\textbf{Keywords:} data evolution, data migration
+
 \section{ORM System}
@@
 $$system = < Application, Database, \rho >$$
-$instantiated : Class \times Application \rightarrow Boolean $ \\
 
 The application and database has to be in consistent state according to a concrete ORM algorithm, otherwise the application cannot work properly.
-It means that during the evolution we have to respect a concrete ORM to produce consistent result too.
+It means that during the evolution we have to respect a concrete ORM to produce consistent result.
 
@@
 
-All important components of the domain are shown in Figure \ref{fig:appStructure}.
+an object-relational mapping is a transformation between application and database schema
+
+$instantiated : Class \times Application \rightarrow Boolean $ \\
+
+\section{Data Evolution in ORM System}
+The data evolution is a change of a state of the ORM system.
+The change of a state is described via evolution operations on application level, the change than has to be propagated into the rest of the system - database schema and data.
+The evolution process is in Figure \ref{fig:evolution}.
 There are three static components (static because they do not define behavior, but state of the system): an application, a database schema and stored data.
-Then there are four dynamic components (they define transformations between components and transition between system states): an object-relational mapping is a transformation between application and database schema, then are two components for evolution of application and of database.
-(To simplify the model we will consider the database schema and data are evolved by one component).
+Then there are two dynamic components (they define transition between system states): component for evolution of application and component for database evolution.
+To simplify the model we will consider the database schema and data are evolved by one component.
 Last component is the most important one - it is component which maps the evolution from application level to relational database level.
@@
 
-\section{Data Evolution in ORM System}
-The Object-Relational system consist of an application and a database.
-The system has to be in consistent state, therefore the evolution of application has to be propagated correctly into database.
-Each evolution operation consist of two parts - applying evolution operation on application and on database.
+%Each evolution operation consist of two parts - applying evolution operation on application and on database.
 The mapping of the evolution change is called extended Object-Relational Mapping (eORM).
+The state of a system is called generation and transformation from one state to another is called evolution.
+The state of the system defines a context for each evolution and it determines if the evolution is successful or not.
 The eORM mapping helps to assure consistency of the system by transforming application evolution into database schema and data.
@@
 
-The state of a system is called generation and transformation from one state to another is called evolution.
-The state of the system defines a context for each evolution and it determines if the evolution is successful or not.
-
 \begin{figure}
@@
 	\includegraphics[scale=0.4]{./images/evolution_simple}
-	\caption{The evolution of data changes the system on all levels}
+	\caption{The evolution of data changes the system on all levels. The figure shows all the components of the evolution process.}
 \end{center}
@@
 
-There is a description of each component following.
-The description defines set of models of static components and dynamic components are defined as functions.
+A description of each component follows.
+Models of static components and dynamic components are defined, then ORM and eORM is defined.
 There are several common parts of the model we can define in advance.
@@
 \textbf{TODO: define consistency of rows and schema}
+\textbf{TODO: define operations such as instantiated}
 
 \section{Object-Relational Mapping}
-There are defined functions for object-relational mapping used in the model.
+There are defined functions for object-relational mapping used in the model in this section.
 The purpose of a function is indicated in its name and should be obvious from its definition.
-These functions do not serve for full object-relational mapping, their purpose is to help with data evolution and its propagation from application to database.
+These functions do not serve for full object-relational mapping, their purpose is to help with data evolution (see Section \ref{sec:eorm}) and its propagation from application to database.
 
@@
  \inference{ cardinality(p) > 1 \wedge !
-mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}tables(d) \cup  table(name(p), primaryKey("Id"), column("value", \rho_t(type(p)), \varnothing), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL) \\
+mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}tables(d) \cup  table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), \varnothing), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL) \\
 
@@
 
-The new foreign key can cause inconsistency of the system when added to a table which contains data.
-This case is not handled by the mapping function above.
-We have to extend the function definition of a feature that allows the mapping between the data in both tables.
-The mapping is a function:
-$$f_m : Property \rightarrow Instance \times Instance $$
-%TODO handle instances
+\section{Evolution}
+\label{sec:eorm}
+There are introduced operations for evolution of an application first as they are a source for the evolution of the whole ORM system.
+The mapping of the application evolution into a database schema and stored data are introduced later in this section too.
 
-\section{Evolution}
 \subsection{Evolution of Application}
@@
 
-The evolution is defined as a set of operations which change structure of an application.
+The evolution of an application is defined as a set of operations which change a structure of the application.
 The definition of an operation consists of two parts - a conditions of operation feasibility and operation impact on the application structure.
@@
 If the conditions of feasibility are not met, the operation cannot impact the application structure.
-%TODO monade
 
@@
 
-\subsection{ORM System Evolution}
-Creational operations does not have impact on stored data they change the structure of the application and database by adding new objects into model.
+\subsection{Extended Object-Relation Mapping}
+The extended object-relational mapping maps the evolution of the application level into the database schema and stored data.
+This section introduces the mapping for evolution operations introduced earlier.
 
 \subsubsection{Creation of a System}
-\paragraph{Add Class}
-
+\paragraph{Add Class} The $AddClass$ operation changes the structure of the database schema only.
 $$
-\Phi_{addClass}(c, d) = \inference{ \forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
+\Phi_{AC}(c, d) = \inference{ \forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
 	\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) 
@@
 $$
-\paragraph{Add Property}
+
+\paragraph{Add Property} The function $AddProperty$ on level of database does not change the structure only, but it has an impact on stored data too.
+This is because a mandatory property can be added into a class with already existing instances, then a default value has to be added to all existing instances.
 $$
-\Phi_{addProperty}(p, d) = \begin{cases}
+\Phi_{AP}(p, d) = \begin{cases}
 \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge (! mandatory(p) \vee ! instantiated(owningClass(p))}{ \rho(p,d)} 
@@
 
-\paragraph{Add Association}
+\paragraph{Add Association} Adding an association into a database is complicated because of already existing instances.
+When an association between classes is added the instances of both classes has to be paired.
+This pairing is provided by the function $f_m$:
+$$f_m : Property \rightarrow Instance \times Instance $$
+
+\textbf{TODO: define the features of the $f_m$}
+
+The mapping of the $AddAssociation$ function is defined as:
 $$
-\Phi_{AddAssociation}(a, f_m, d) = \begin{cases}
+\Phi_{AA}(a, f_m, d) = \begin{cases}
 \inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
@@
 $$
-\Phi_{removeClass}(c,d) = \inference{ \rho(c) \in tables(d) \\ \wedge \not \exists r \in data(d) : owningTable(r) = \rho(c) \\ \not \exists j \in tables(d), f \in foreignKeys(j) : reference(f) = \rho(c)  }{ tables(d) \setminus \rho(c)}
+\Phi_{RC}(c,d) = \inference{ \rho(c) \in tables(d) \\ \wedge \not \exists r \in data(d) : owningTable(r) = \rho(c) \\ \not \exists j \in tables(d), f \in foreignKeys(j) : reference(f) = \rho(c)  }{ tables(d) \setminus \rho(c)}
 $$
@@
 $$
-\Phi_{removeProperty}(p, d) = \begin{cases}
+\Phi_{RP}(p, d) = \begin{cases}
  \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \exists c \in columns(t) : name(c) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p)) }{columns(t) \setminus c
@@
 $$
-\Phi_{removeAssociation}(a, d) =  \begin{cases}
+\Phi_{RA}(a, d) =  \begin{cases}
  \inference{\exists t \in tables(d) : name(t) = name(owningClass(a))

### 2012-12-21 doc/description.tex
added texts

@@
 \begin{abstract}
-The evolution of an application is a common issue during software development.
-The formal description of the evolution operations in context of  software created using an object oriented programming language and a relational database is introduced in this paper.
+The formal definition of the evolution in context of  software created using an object oriented programming language and a relational database is introduced in this paper.
 Special attention is given to the data stored in databases (instances) and theirs consistency during evolution process.
@@
 \textbf{Keywords:} data evolution, data migration
+\section{Introduction}
+The evolution of an application is a common issue during software development.
+The evolution can occur from many reasons in different parts of application lifecycle.
+The evolution of a software  created using an object oriented programming language and a relational database is usually processed as code refactoring and database migration is processed as a set of SQL scripts.
+The code refactoring is supported by many IDEs and the database migration is supported by many frameworks and tools, nevertheless these tools are usually not capable to solve complicated migration cases or preserve stored data.
+Another disadvantage is the evolution has to be define twice - as a code refactoring and as a database migration.
 
-\section{ORM System}
-The ORM system represents an application created using object oriented language and relational database.
-The components of such an application and theirs relations are in Figure \ref{fig:appStructure}.
-We define the application in our model as:
+This paper provides a formal description of an evolutionary framework, which evolves a database including stored data according to a evolution in application.
+There are defined models representing an application and database in the paper, together with functions which are able to perform an evolution of the whole software.
 
-$$system = < Application, Database, \rho >$$
+The paper is organized as follows: in Section \ref{sec:evolution} the concept of ORM software and its evolution is discussed, then models of application and database are introduced in Section \ref{sec:appModel} and in Section \ref{sec:dbModel}, the ORM is defined in Section \ref{sec:orm} and the evolutionary transformations are defined in Section \ref{sec:eorm}.
 
-The application and database has to be in consistent state according to a concrete ORM algorithm, otherwise the application cannot work properly.
-It means that during the evolution we have to respect a concrete ORM to produce consistent result.
+\section{Data Evolution in ORM Software}
+\label{sec:evolution}
+The ORM software consist of an application created using object oriented language and relational database.
+The application and database are in consistent state according to a concrete ORM function, otherwise the software cannot work properly.
+An ORM software is illustrated in Figure \ref{fig:appStructure}.
+The software can be defined as a triple:
+$$software = < Application, Database, \rho >$$
+Models of $Application$ and $Database$ will be introduced in detail in Section \ref{sec:appModel} and in Section \ref{sec:dbModel}, the ORM function $\rho$ is defined in Section \ref{sec:orm}.
+The consistent state of application and database is defined as:
+$$\rho(Application) = Database$$
 
@@
 
-\begin{itemize}
-	\item \textbf{TODO define consistency}
-	\item \textbf{TODO: $\rho(Application) = Database$ show how database is build from Application}
-\end{itemize}
+The evolution of an software is a transformation of one consistent state to another consistent state: $$\epsilon : system \rightarrow system $$ We will call consistent states of one software \emph{generations} of the software.
+The evolution is quite easy if we consider only structure of an application and database and ignore instances (stored data).
+The evolution can be described in such a case as:
+\begin{equation*}
+\epsilon(software(App, Db)) = \{a \rightarrow \epsilon_{App}(App)  | software(a, \rho(a), \rho)\}
+\end{equation*}
+It means only the structure of the application is evolved and the new database is obtained as a result od ORM.
+Stored data are usually useless in many cases as their structure does not correspond to the new database schema.
+We believe the data preservation and consistency is a crucial issue during the software evolution.
 
-an object-relational mapping is a transformation between application and database schema
-
-$instantiated : Class \times Application \rightarrow Boolean $ \\
-
-\section{Data Evolution in ORM System}
-The data evolution is a change of a state of the ORM system.
-The change of a state is described via evolution operations on application level, the change than has to be propagated into the rest of the system - database schema and data.
-The evolution process is in Figure \ref{fig:evolution}.
-There are three static components (static because they do not define behavior, but state of the system): an application, a database schema and stored data.
-Then there are two dynamic components (they define transition between system states): component for evolution of application and component for database evolution.
-To simplify the model we will consider the database schema and data are evolved by one component.
-Last component is the most important one - it is component which maps the evolution from application level to relational database level.
-We call this component evolution object-relational mapping (eORM).
-
-%Each evolution operation consist of two parts - applying evolution operation on application and on database.
-The mapping of the evolution change is called extended Object-Relational Mapping (eORM).
-The state of a system is called generation and transformation from one state to another is called evolution.
-The state of the system defines a context for each evolution and it determines if the evolution is successful or not.
-The eORM mapping helps to assure consistency of the system by transforming application evolution into database schema and data.
-Each eORM function applied on an ORM system produces a new generation of the system.
-If the function cannot be applied on a system the system cannot be evolved by the function.
+The evolution process we propose is in Figure \ref{fig:evolution}.
+The transition from the first generation to the second one is based on application evolution which is interpreted on the database level.
+The interpretation of application evolution is called evolutionary object-relation mapping (eORM) and its defined as a function:
+\begin{equation*}
+	\Phi : \epsilon_{app} \rightarrow \epsilon_{db}
+\end{equation*}
+The evolution of the software can be than defined as:
+\begin{equation*}
+\epsilon(software(App, Db)) = \{software(\epsilon_{app}(App), \Phi((\epsilon_{app}, Db), \rho)\}
+\end{equation*}
 
@@
 \begin{itemize}
-	\item \textbf{TODO show how a whole system is evolved}
 	\item \textbf{TODO state monade}
@@
 
-A description of each component follows.
-Models of static components and dynamic components are defined, then ORM and eORM is defined.
-There are several common parts of the model we can define in advance.
-The model uses a set $Labels$ which contains all possible identifiers for the model.
-Each model component which has a label ($label \in Labels$) is unambiguously identified in the model.
-
-%%%%%%%%%%%%%%%%%%%
-
 \section{Application Model}
+\label{sec:appModel}
 The model represents a persistent layer of an application created using object oriented language.
@@
 Each instance is defined according to a class - the class defines the structure and relationships of the instance.
-The instances are stored in the relational database in our case, therefore we define only the structure of classes in our model:
+The instances are stored in the relational database in our case, therefore we define only the structure of classes in our model.
+
+The model uses a set $Labels$ which contains all possible identifiers for the model. Each model component which has a label ($label \in Labels$) is unambiguously identified in the model.
 
 $$
-Application = <Class*>
+Application = < Class* > 
 $$
 $classes : Application \rightarrow Class*$ \\
-$classes(<cs, is >) = = cs, cs \in Class* \wedge is \in Instance* $ \\ \\
+$classes(<cs, is >) = cs, cs \in Class* \wedge is \in Instance* $ \\ \\
 $class : Label \times Application \rightarrow Class  $ \\ 
-$class(l, a) = c, c \in classes(a) \wedge name(c) = l, l \in Label, a \in Application* $ \\
-%TODO: variable types
-
-The structure consist not only of classes, the structure of the application is defined by five concepts.
-Each concept is defined and a set of manipulating functions is define to manipulate the application model easily.
+$class(l, a) = c, c \in classes(a) \wedge name(c) = l, l \in Label, a = Application $ \\ \\
+The structure of an application consists not only of classes, it is defined by five other concepts.
+Besides the structure definition, there is defined a set of functions, which allows to manipulate the application model easily.
 
@@
 \section{Database Model}
-The relational database consists of database schema which defines the structure of the database and data, which in our ORM system represents instances. The database is a pair:
+\label{sec:dbModel}
+The relational database consists of database schema which defines the structure of the database and data, which in our ORM system represents instances. The database is:
 $$
@@
 \subsection{Model of Stored Data}
-The representation of stored data in the relational database is based on pairs:
-$$pair = < refColumn, value > | < refFK, value>$$
 
-\textbf{TODO: primary key}
-where the $refColumn$ reference a column or foreign key by its name and $value$ represents concrete value stored in the database. A row in a table is a set of pairs with reference to a concrete table:
+A row in a table is tuple consisting of reference to a concrete table and a set of value pairs, which represents concrete values of concrete columns.
+$$row = < refTable, pair* >$$
+$$pair = < reference, value >$$
+The $reference $ is a name of a column, a foreign key or its primary key and $value$ represents concrete value stored in the database.
+Each row in a database has to refer to a table in database schema, otherwise the database is inconsistent.
+The consistency of database is then:
+\begin{equation*}
+\begin{gathered}
+\forall r \in data(d), \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) \exists u \in columns(t) \cup foreignKeys(t) : reference(p) = name(u) 
+\end{gathered}
+\end{equation*}
 
-$$row = < refTable, pair* >$$
+\begin{itemize}
+	\item \textbf{TODO: define operations such as instantiated}
+\end{itemize}
 
-\textbf{TODO: define consistency of rows and schema}
-\textbf{TODO: define operations such as instantiated}
+$instantiated : Class \times Application \rightarrow Boolean $ \\
 
 \section{Object-Relational Mapping}
+\label{sec:orm}
 There are defined functions for object-relational mapping used in the model in this section.
@@
 These functions do not serve for full object-relational mapping, their purpose is to help with data evolution (see Section \ref{sec:eorm}) and its propagation from application to database.
+
+\begin{itemize}
+	\item \textbf{TODO: $\rho(Application) = Database$ show how database is build from Application}
+\end{itemize}
 

### 2012-12-22 doc/description.tex
orm algorithm

@@
 \usepackage{graphicx}
+\usepackage{algpseudocode}
 
@@
 The software can be defined as a triple:
-$$software = < Application, Database, \rho >$$
+\begin{equation*}
+	software = < Application, Database, \rho >
+\end{equation*}
 Models of $Application$ and $Database$ will be introduced in detail in Section \ref{sec:appModel} and in Section \ref{sec:dbModel}, the ORM function $\rho$ is defined in Section \ref{sec:orm}.
@@
 \end{equation*}
+There are two functions connected with stored data - first of them verify if the table contains any data:
+$$contains : Table \times Database \rightarrow Boolean $$
+\begin{equation*}
+\begin{gathered}
+	contains(t, d) = \begin{cases}
+ \inference{\exists r \in data(d) : refTable(r) = name(t) }{true} \\ \\
+  \inference{\forall r \in data(d) : refTable(r) \neq name(t)}{false}
+ \end{cases} \\ \\
+ t \in tables(d), d \in Database
+\end{gathered}
+\end{equation*}
 
-\begin{itemize}
-	\item \textbf{TODO: define operations such as instantiated}
-\end{itemize}
-
-$instantiated : Class \times Application \rightarrow Boolean $ \\
+second of them verify if there are instances of a class:
+$$instantiated : Class \times Software \rightarrow Boolean $$ 
+\begin{equation*}
+\begin{gathered}
+	instantiated(c, s) = \begin{cases}
+ \inference{\exists t \in tables(database(s)) : name(t) = name(c) \\ \wedge contains(t, database(s))}{true} \\ \\
+  \inference{\exists t \in tables(database(s)) : name(t) = name(c) \\ \wedge ! contains(t, database(s))}{false} \\ \\
+ \end{cases} \\ \\
+ c \in classes(application(s)), s \in Software 
+\end{gathered}
+\end{equation*}
 
@@
 There are defined functions for object-relational mapping used in the model in this section.
-The purpose of a function is indicated in its name and should be obvious from its definition.
+The purpose of a function is indicated by its name and should be obvious from its definition.
 These functions do not serve for full object-relational mapping, their purpose is to help with data evolution (see Section \ref{sec:eorm}) and its propagation from application to database.
+However the database can be obtained from an application model by using following algorithm:
 
-\begin{itemize}
-	\item \textbf{TODO: $\rho(Application) = Database$ show how database is build from Application}
-\end{itemize}
+\begin{algorithmic}[1]
+	\Require $a \in Application$
+	\State d = Database
+	\ForAll{$c \in classes(a)$} 
+		\State $\rho_c(c), d$
+		\ForAll{$p \in properties(c)$} 
+			\State $\rho_p(p, d)$
+		\EndFor
+	 \EndFor
+	\ForAll{$c \in classes(a)$}
+		\ForAll{$as \in associations(c)$} 
+			\State $\rho_{a}(as, d)$
+		\EndFor
+	\EndFor
+\end{algorithmic}
 
@@
 
-\begin{center}
-$ \rho_{t} : AppType \rightarrow DbType $ \\
-$ \rho_{t}(APPSTRING) = DBSTRING $
-\end{center}
+\begin{equation*}
+\begin{gathered}
+	\rho_{t} : AppType \rightarrow DbType  \\
+ 	\rho_{t}(APPSTRING) = DBSTRING 
+\end{gathered}
+\end{equation*}
 
@@
 \subsubsection{Application Alternation}
-\paragraph{Rename}
+\paragraph{Rename Class}
+$$\rho_{RenC} : Label \times Label \times Application \rightarrow Class $$
+\begin{equation*}
+	\rho_{RenC}(l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \forall e \in classes(a) : name(e) \neq l_n}{\begin{gathered}
+c.name = l_n \wedge \forall f \in associations(e), \\ e \in classes(a), reference(f) = l_c : f.reference = l_n 
+\end{gathered}
+}
+\end{equation*}
+
+\paragraph{Rename Property}
+$$\rho_{RenP} : Property \times Label \times Application \rightarrow Property $$
+\paragraph{Rename Association}
+$$\rho_{RenA} : Association \times Label \times Application \rightarrow Association $$
 \paragraph{Change Cardinality}
+$$\rho_{CcP} : Property \times NzNat \times Application \rightarrow Property $$
+$$\rho_{cardA} : Association \times NzNat \times Application \rightarrow Association $$
 \paragraph{Copy Property}
+$$\rho_{copyP} : Property \times Class \times Class \times Application \rightarrow Class \times Class $$
 
@@
 
-\subsubsection{Creation of a System}
+\subsubsection{Creation of a Software}
 \paragraph{Add Class} The $AddClass$ operation changes the structure of the database schema only.
@@
 
-\subsubsection{Deconstruction of a System}
+\subsubsection{Deconstruction of a Software}
 \paragraph{Remove Class}
@@
 
-\subsubsection{Altering a System}
+\subsubsection{Altering a Software}
 

### 2012-12-29 doc/description.tex
added state monade and copy property

@@
 
-\section{Data Evolution in ORM Software}
+\section{ORM Software}
 \label{sec:evolution}
@@
 \end{figure}
-
-\begin{itemize}
-	\item \textbf{TODO state monade}
-\end{itemize}
 
@@
 The definition of an operation consists of two parts - a conditions of operation feasibility and operation impact on the application structure.
-We decide to use following transcription:
+If the conditions of feasibility are not met, the operation cannot change the application structure - the software cannot evolve.
+This is important in case we have an evolution consisting of several evolution operations.
+These operations has to be proceed in concrete order and all of them has to change the software otherwise the software can become inconsistent from user point of view (the structure or stored data will not correspond with the users' reality).
+To assure the right behavior of evolution operations and to propagate operations fail to the user we extend the definition of a software to make it a monad.
+The extension consists in defining a software state and operations $ret$ and $evolve$, which is used as a concatenation of operations as well.
+$$state = Consistent s | Inconsistent s, s \in Softwares$$
 
-$$
-operation = \inference{conditions$ $of$ $feasibility}{impact$ $on$ $application$ $structure}
-$$
-If the conditions of feasibility are not met, the operation cannot impact the application structure.
+$$ret : Software \rightarrow State $$
+\begin{equation*}
+\begin{gathered}
+	ret(s) = \begin{cases}
+ Consistent s, consistent(s)
+ \\
+ Inconsistent s, ! consistent(s)
+ \end{cases}
+\end{gathered}
+\end{equation*}
+
+Each evolution operation has a definition: $Software \rightarrow State$ and all evolution operations creates a set called $Operations$, then a concatenation of evolution operations is defined as:
+
+$$evolve : State \times Operations \rightarrow State  $$
+\begin{equation*}
+\begin{gathered}
+ evolve(t, o) = \begin{cases}
+ t, \; t = Inconsistent \; s \\
+ o(s), \; t = Consistent \; s 
+\end{cases}\\ 
+t \in States, o \in Operatons, s \in Softwares
+\end{gathered}
+\end{equation*}
+
+We decide to use following transcription to cover all aspects  of operation definition:
+
+\begin{equation*}
+\begin{gathered}
+operation(parameters, software) = \begin{cases}
+  \inference{conditions \; of \; feasibility}{impact \; on \; software \; structure}\\ \\
+  \inference{otherwise}{Inconsistent \; software}
+ \end{cases}
+\end{gathered}
+\end{equation*}
+
+The second branch of the operation definition is an implicit assumption - if an operation is not feasible in a software, its state become inconsistent.
+We will not state this implicit assumption in the operations definitions to increase theirs readability.
 
@@
 \paragraph{Rename Class}
-$$\rho_{RenC} : Label \times Label \times Application \rightarrow Class $$
+$$\rho_{RenC} : Label \times Label \times Application \rightarrow Application $$
 \begin{equation*}
@@
 \paragraph{Rename Property}
-$$\rho_{RenP} : Property \times Label \times Application \rightarrow Property $$
+$$\rho_{RenP} : Label \times Label \times Label \times Application \rightarrow Application $$
+\begin{equation*}
+	\rho_{RenP}(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_o \\ \wedge \forall e \in properties(c) : name(e) \neq l_n}{\begin{gathered}
+p.name = l_n 
+\end{gathered}
+}
+\end{equation*}
+
 \paragraph{Rename Association}
-$$\rho_{RenA} : Association \times Label \times Application \rightarrow Association $$
+$$\rho_{RenA} : Label \times Label \times Label \times Application \rightarrow Application $$
+\begin{equation*}
+	\rho_{RenP}(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists a \in associations(c) : name(a) = l_o \\ \wedge \forall e \in associations(c) : name(e) \neq l_n}{\begin{gathered}
+a.name = l_n 
+\end{gathered}
+}
+\end{equation*}
+
+\paragraph{Copy Property}
+All transformations described earlier does change only the structure of a application.
+These operation can be used in case there are no instances in the application.
+If instances are present in the application, the decomposition transformations could cause a data lost.
+Therefore we define  the transformation for copying a property from one class to another.
+The copy property is the basic way to copy data from instance of a class to instance of another class.
+
+The copy property can be implemented as a transformation composed by operations for adding and removing property, nevertheless this composition would not work with instances, thus we define the copy property operation as a basic and atomic operation which will help us to handle instances.
+
+$$\rho_{copyP} : Label \times Label \times Label \times (Instance \times Instance) \times Application \rightarrow Application $$
+\begin{equation*}
+	\rho_{copy}(l_s, l_t, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p \\ \wedge \exists e \in classes(a) : name(e) = l_t  \\ \wedge \forall q \in properties(c) : name(q) \neq l_p}{\begin{gathered}
+properites(e) \cup p 
+\end{gathered}
+}
+\end{equation*}
+
+With a copy property transformation defined we are able to define advanced evolution scenarios.
+
+\subsection{Composing of Evolution Operations}
+
 \paragraph{Change Cardinality}
@@
 $$\rho_{cardA} : Association \times NzNat \times Application \rightarrow Association $$
-\paragraph{Copy Property}
-$$\rho_{copyP} : Property \times Class \times Class \times Application \rightarrow Class \times Class $$
 
@@
 \subsubsection{Altering a Software}	
+\paragraph{Copy Column}
+$$\Phi_{copyC} : Label \times Label \times Label \times f_m \times Database \rightarrow Database $$
+
+\begin{equation*}
+	\begin{gathered}
+		\Phi_{copyC}(l_c, l_s, l_t, f_m, d) = \begin{cases}
+ 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{columns(e) \cup c \wedge \forall r \in rows(t) : f_m(r) \cup getPair(l_c,r)}
+	\end{cases}
+	\end{gathered}
+\end{equation*}
 

### 2012-12-30 doc/description.tex
text about transformations, format updates in app model

@@
 The model uses a set $Labels$ which contains all possible identifiers for the model. Each model component which has a label ($label \in Labels$) is unambiguously identified in the model.
-
-$$
+\begin{gather*}
   Application = < Class* >
-$$
-$classes : Application \rightarrow Class*$ \\
-$classes(<cs, is >) = cs, cs \in Class* \wedge is \in Instance* $ \\ \\
-$class : Label \times Application \rightarrow Class  $ \\ 
-$class(l, a) = c, c \in classes(a) \wedge name(c) = l, l \in Label, a = Application $ \\ \\
+  \\ 
+  app : Class* \rightarrow  Application \\ \\
+  classes : Application \rightarrow Class* \\
+  classes(app(cs)) = cs  \\\\
+  class : Label \times Application \rightarrow Class   \\ 
+  class(l, a) = c, c \in classes(a) \wedge name(c) = l \\ \\
+  l \in Label, a = Application, cs \in Class* 
+  \\ \\
+  associated :  Label \times Application \rightarrow Boolean \\
+  owningClass : Property \times Application \rightarrow Class  \\ owningClass(p, a) = c, c \in classes(a) \wedge p \in properties(c) \\
+  owningClass : Association \times Application \rightarrow Class  \\
+owningClass(as, a) = c, c \in classes(a) \wedge as \in associations(c)
+\end{gather*}
 The structure of an application consists not only of classes, it is defined by five other concepts.
@@
 
-$$Class = < label, Property*, Association* >$$
-$name : Class \rightarrow Label$ \\
-$name(< lab, props, assocs  >) = lab$ \\ \\
-$properties : Class \rightarrow Property*$ \\
-$properties(< lab, props, assocs  >) = props $\\ \\
-$primitiveProperties : Class \rightarrow Property*$ \\
-$primitiveProperties(< lab, props, assocs  >) = props', \forall p \in props' : cardinality(p) \leq 1 $\\ \\
-$collectionProperties : Class \rightarrow Property*$ \\
-$collectionProperties(< lab, props, assocs  >) = props', \forall p \in props' : cardinality(p) > 1 $\\ \\
-$associations : Class \rightarrow Association*$ \\
-$associations(< lab, props, assocs  >) = assocs $ \\ \\
-$associated :  Class \times Application \rightarrow Boolean $
-%TODO: variable types
+$$Class = <label, Property*, Association*>$$
+\begin{equation*}
+\begin{alignat}{2}
+cl : Label \times Property* \times Association* \rightarrow Class &\quad&\text{constructor}\\ \\
+% TODO association has to be only to self (all associated has to exists)
+name : Class \rightarrow Label &\quad&\text{name of the class}\\
+name(cl(l, props, assocs)) = l \\ \\
+properties : Class \rightarrow Property* &\quad&\text{all properties}\\
+properties(cl(l, props, assocs)) = props \\ \\
+primitiveProperties : Class \rightarrow Property* &\quad&\text{all primitive properties}\\
+primitiveProperties(cl(l, props, assocs)) = props', \\ props' \subset props,  \forall p \in props' : cardinality(p) \leq 1 \\ \\
+collectionProperties : Class \rightarrow Property* &\quad&\text{all collections}\\
+collectionProperties(cl(l, props, assocs)) = props', \\ props' \subset props, \forall p \in props' : cardinality(p) > 1 \\ \\
+associations : Class \rightarrow Association* &\quad&\text{all associations}\\
+associations(cl(l, props, assocs)) = assocs
+\end{alignat}
+\end{equation*}
+$$l \in Label, props \in Properties, assocs \in Associations$$
 
 \paragraph{Property} represents a feature of  a class which is can be represented as a primitive type. The property can be mandatory, can have a default value.
-$$
-Property = < label, AppType, DefaultValue, Cardinality, Mandatory >
-$$
-$name : Property \rightarrow Label$ \\
-$name(< lab, t, val, n  >) = lab$ \\ \\
-$type : Property \times AppType$ \\
-$type(< lab, t, val, n  >) = t$ \\ \\
-$cardinality : Property \rightarrow NzNat$ \\
-$cardinality(< lab, t, val, n  >) = n$ \\ \\
-$mandatory : Property \rightarrow Boolean $
-$mandatory(<lab, t, val, n, m >) = = m $ \\ \\
-$owningClass : Property \times Application \rightarrow Class $ \\
-$owningClass(p, a) = c, c \in classes(a) \wedge p \in properties(c) $
+\begin{gather*}
+Property = <label, AppType, DefaultValue, Cardinality, Mandatory> \\ \\
+prop : Label \times AppType \times String \times \mathbb{N^{*}} \times Boolean \rightarrow Property \\ \\
+name : Property \rightarrow Label \\
+name(prop(l, t, val, n, m)) = l \\ \\
+type : Property \times AppType \\
+type(prop(l, t, val, n, m)) = t \\ \\
+cardinality : Property \rightarrow \mathbb{N^{*}} \\
+cardinality(prop(l, t, val, n, m)) = n \\ \\
+mandatory : Property \rightarrow Boolean \\
+mandatory(prop(l, t, val, n, m)) = m  \\ \\
+l \in Label, t \in AppType, val \in String*, n \in \mathbb{N^{*}}, m \in Boolean
+\end{gather*}
+
 %TODO: variable types
@@
 The cardinalities defines the multiplicity of the association.
-$$
-Association = (label, classRef, startCardinality, endCardinality)
-$$
-$name : Association \rightarrow Label$ \\
-$name(< lab, ref, n_1, n_2  >) = lab$\\ \\
-$reference : Association \rightarrow Label$ \\
-$reference(< lab, ref, n_1, n_2  >) = ref$\\ \\
-$startCardinality : Association \rightarrow NzNat$ \\
-$startCardinality(< lab, ref, n_1, n_2  >) = n_1$\\ \\
-$endCardinality : Association \rightarrow NzNat$ \\
-$endCardinality(< lab, ref, n_1, n_2  >) = n_2$ \\ \\
-$owningClass : Association \times Application \rightarrow Class $ \\
-$owningClass(as, a) = c, c \in classes(a) \wedge as \in associations(c) $
+\begin{gather*}
+Association = <label, classRef, startCardinality, endCardinality> \\ \\
+assoc : Label \times Label \times \mathbb{N} \times \mathbb{N} \rightarrow Association
+name : Association \rightarrow Label \\
+name(assoc(l, ref, n_1, n_2)) = l\\ \\
+reference : Association \rightarrow Label \\
+reference(assoc(l, ref, n_1, n_2)) = ref\\ \\
+startCardinality : Association \rightarrow \mathbb{N} \\
+startCardinality(assoc(l, ref, n_1, n_2)) = n_1\\ \\
+endCardinality : Association \rightarrow \mathbb{N} \\
+endCardinality(assoc(l, ref, n_1, n_2)) = n_2 \\ \\
+l, ref \in Label,  n_1, n_2 \in \mathbb{N}
+\end{gather*}
+
 %TODO: variable types
@@
 AppType = APPSTRING
-$$
-The types used in application model:
-$$
-cardinality = NzNat
-$$
-$$
-referencesClass = label
-$$
-$$
-startCardinality = NzNat
-$$
-$$
-endCardinality = NzNat
 $$
@@
 $name : Table \rightarrow Label $ \\
-$name(< lab, pk, cols, fks  >) = lab$ \\ \\
+$name(< lab, pk, cols, fks)) = lab$ \\ \\
 $primaryKey : Table \rightarrow PrimaryKey $ \\
-$primaryKey(< lab, pk, cols, fks  >) = pk$ \\ \\
+$primaryKey(< lab, pk, cols, fks)) = pk$ \\ \\
 $columns : Table \rightarrow Column* $ \\
-$columns(< lab, pk, cols, fks  >) = cols$ \\ \\
+$columns(< lab, pk, cols, fks)) = cols$ \\ \\
 $foreignKeys : Table \rightarrow ForeignKey* $ \\
-$foreignKeys(< lab, pk, cols, fks  >) = fks$
+$foreignKeys(< lab, pk, cols, fks)) = fks$
 
@@
 $name : Column \rightarrow Label $ \\
-$name(< lab, t, val, cons  >) = lab $ \\ \\
+$name(< lab, t, val, cons)) = lab $ \\ \\
 $type : Column \rightarrow DbType $ \\
-$type(< lab, t, val, cons  >) = t $ \\ \\
+$type(< lab, t, val, cons)) = t $ \\ \\
 $constraints : Column \rightarrow Constraint* $ \\
-$constraints(< lab, t, val, cons  >) = cons $ \\ \\
+$constraints(< lab, t, val, cons)) = cons $ \\ \\
 $owningTable : Column \times Database \rightarrow Table $ \\
@@
 $name : ForeignKey \rightarrow Label $ \\
-$name(< lab, tRef, cons  >) = lab $ \\ \\
+$name(< lab, tRef, cons)) = lab $ \\ \\
 $reference : ForeignKey \rightarrow Label $ \\
-$reference(< lab, tRef, cons  >) = tRef $ \\ \\
+$reference(< lab, tRef, cons)) = tRef $ \\ \\
 $reference : ForeignKey \rightarrow Constraint* $ \\
-$reference(< lab, tRef, cons  >) = cons $ \\ \\
+$reference(< lab, tRef, cons)) = cons $ \\ \\
 $owningTable : ForeignKey \times Database \rightarrow Table $ \\
@@
 \subsection{Mapping of Classes}
-We assume the primary key column is created automatically by database, therefore the primary key is always created with name Id and od NzNat type and properties with cardinality 0 or 1 are mapped into columns.
+We assume the primary key column is created automatically by database, therefore the primary key is always created with name Id and of the $\mathbb{N^{*}}$ type and properties with cardinality 0 or 1 are mapped into columns.
 The associations are ignored int this mapping as it is one of the partial mapping function used in eORM.
@@
 
-The evolution of an application is defined as a set of operations which change a structure of the application.
+The evolution of a software is defined as a set of transformations which change a structure of the software.
 The definition of an operation consists of two parts - a conditions of operation feasibility and operation impact on the application structure.
@@
 To assure the right behavior of evolution operations and to propagate operations fail to the user we extend the definition of a software to make it a monad.
-The extension consists in defining a software state and operations $ret$ and $evolve$, which is used as a concatenation of operations as well.
-$$state = Consistent s | Inconsistent s, s \in Softwares$$
+The extension consists in defining a software state and operations $ret$ and $\oplus$ (called evolve), which is used as a concatenation of operations as well.
+\begin{gather*}
+state = Consistent \; s \: | \: Inconsistent \; s, v \in Softwares 
+\end{gather*}
 
-$$ret : Software \rightarrow State $$
 \begin{equation*}
 \begin{gathered}
+ret : Software \rightarrow State \\ \\
 	ret(s) = \begin{cases}
- Consistent s, consistent(s)
+ Consistent \; s, consistent(s)
  \\
- Inconsistent s, ! consistent(s)
+ Inconsistent\;  s, ! consistent(s)
  \end{cases}
@@
 \end{equation*}
-
-Each evolution operation has a definition: $Software \rightarrow State$ and all evolution operations creates a set called $Operations$, then a concatenation of evolution operations is defined as:
-
-$$evolve : State \times Operations \rightarrow State  $$
+Each evolution transformation has a definition: $Software \rightarrow State$ and all evolution transformations creates a set called $Transformations$, then a concatenation of evolution transformation is defined as:
 \begin{equation*}
 \begin{gathered}
- evolve(t, o) = \begin{cases}
- t, \; t = Inconsistent \; s \\
- o(s), \; t = Consistent \; s 
+ \oplus : State \times Transformations \rightarrow State  \\ \\
+ u \oplus t = \begin{cases}
+ u, \; u = Inconsistent \; s \\
+ t(u), \; u = Consistent \; s 
 \end{cases}\\ 
-t \in States, o \in Operatons, s \in Softwares
+u \in States, t \in Transformations, s \in Softwares
 \end{gathered}
 \end{equation*}
+Each transformation consist in changes of the application and the database.
+To describe these two aspects od software evolution we decide to define transformations for each layer separately.
+The separation of transformation contexts help us to define transformation specific for application or database evolution, which can ease both - understanding and defining the transformation.
+The transformations specific for a context are called operations and they creates two sets- first one for application evolution: $AppOps$  and second one for database evolution $DbOps$.
+Applying of a transformation is defined:
+\begin{gather*}
+u \oplus t = ret(<application(u) \oplus o_{app}, database(u) \oplus o_{db}>) \\
+u \in Softwares, t \in Transformations, o_{app} \in AppOps, o_{db} \in DbOps
+\end{gather*}
+The function $\Phi$ provides a mapping between the $AppOps$ and $DbOps$ sets.
+The $\Phi$ function makes possible to use a description of an application evolution as a source for the evolution of the whole software.
+\begin{gather*}
+\Phi : AppOps \rightarrow DbOps
+\end{gather*}
 
+The definition of an operation consists of two parts - a conditions of operation feasibility and operation impact on the context.
 We decide to use following transcription to cover all aspects  of operation definition:
@@
 \begin{gathered}
-operation(parameters, software) = \begin{cases}
-  \inference{conditions \; of \; feasibility}{impact \; on \; software \; structure}\\ \\
-  \inference{otherwise}{Inconsistent \; software}
+operation(parameters, context) = \begin{cases}
+  \inference{conditions \; of \; feasibility}{impact \; on \; context}\\ \\
+  \inference{otherwise}{Inconsistent \; context}
  \end{cases}
@@
 \paragraph{Change Cardinality}
-$$\rho_{CcP} : Property \times NzNat \times Application \rightarrow Property $$
-$$\rho_{cardA} : Association \times NzNat \times Application \rightarrow Association $$
+$$\rho_{CcP} : Property \times \mathbb{N^{*}} \times Application \rightarrow Property $$
+$$\rho_{cardA} : Association \times \mathbb{N^{*}} \times Application \rightarrow Association $$
 

### 2012-12-30 doc/description.tex
updated application operations typesetting

@@
   owningClass : Association \times Application \rightarrow Class  \\
-owningClass(as, a) = c, c \in classes(a) \wedge as \in associations(c)
+owningClass(as, a) = c, c \in classes(a) \wedge as \in associations(c) \\ \\
+owningTable : Column \times Database \rightarrow Table  \\
+owningTable(c, d) = tab, tab \in tables(d) \wedge c \in columns(d) \\ \\
+owningTable : ForeignKey \times Database \rightarrow Table  \\
+owningTable(fk, d) = tab, tab \in tables(d) \wedge fk \in foreignKeys(d)\\ \\ 
 \end{gather*}
@@
 The relational database consists of database schema which defines the structure of the database and data, which in our ORM system represents instances. The database is:
-$$
-Database = <table*, row*>
-$$
-$table : Label \times Database \rightarrow Table $ \\
-$table(l, d) = t, t \in tables(d) \wedge name(t) = l$ \\
+\begin{gather*}
+Database = <table*, row*> \\ \\
+db : Tables \times Rows \rightarrow Database \\ \\
+tables : Database \rightarrow Tables \\
+tables(db(tabs, rows)) = tabs \\ \\
+data : Database \rightarrow Rows \\
+data(db(tabs, rows)) = rows \\ \\
+table : Label \times Database \rightarrow Table  \\
+table(l, d) = t, t \in tables(d) \wedge name(t) = l \\ \\
+tabs \in Tables, rows \in Rows, l \in Labels, d \in Databases
+\end{gather*}
 
@@
 Rows in the table represents stored data.
-$$
-Table = (label, primaryKey, Column*, ForeignKey*)
-$$
-$name : Table \rightarrow Label $ \\
-$name(< lab, pk, cols, fks)) = lab$ \\ \\
-$primaryKey : Table \rightarrow PrimaryKey $ \\
-$primaryKey(< lab, pk, cols, fks)) = pk$ \\ \\
-$columns : Table \rightarrow Column* $ \\
-$columns(< lab, pk, cols, fks)) = cols$ \\ \\
-$foreignKeys : Table \rightarrow ForeignKey* $ \\
-$foreignKeys(< lab, pk, cols, fks)) = fks$
+\begin{gather*}
+Table = <label, primaryKey, Column*, ForeignKey*>\\ \\
+tab : Label \times PrimaryKeys \times Columns \times ForeignKeys \rightarrow Table \\ \\
+name : Table \rightarrow Label  \\
+name(tab(l, pk, cols, fks)) = l \\ \\
+primaryKey : Table \rightarrow PrimaryKeys  \\
+primaryKey(tab(l, pk, cols, fks)) = pk \\ \\
+columns : Table \rightarrow Columns  \\
+columns(tab(l, pk, cols, fks)) = cols \\ \\
+foreignKeys : Table \rightarrow ForeignKeys  \\
+foreignKeys(tab(l, pk, cols, fks)) = fks \\ \\
+l \in Label, pk \in PrimaryKey, cols \in Columns, fks \in ForeignKeys
+\end{gather*}
 
 \paragraph{Column} columns define possible data values and types which can be part of a table record.
-$$
-Column = (label, DbType, DefaultValue, Constraint*)
-$$
-$name : Column \rightarrow Label $ \\
-$name(< lab, t, val, cons)) = lab $ \\ \\
-$type : Column \rightarrow DbType $ \\
-$type(< lab, t, val, cons)) = t $ \\ \\
-$constraints : Column \rightarrow Constraint* $ \\
-$constraints(< lab, t, val, cons)) = cons $ \\ \\
-$owningTable : Column \times Database \rightarrow Table $ \\
-$constraints(c, d) = tab, tab \in tables(d) \wedge c \in columns(d) $
+\begin{gather*}
+Column = <label, DbType, DefaultValue, Constraint*> \\ \\
+col : Label \times DbType \times String \times Constraints \rightarrow Column \\ \\
+name : Column \rightarrow Label  \\
+name(col(l, t, val, cons)) = l  \\ \\
+type : Column \rightarrow DbType  \\
+type(col(l, t, val, cons)) = t  \\ \\
+constraints : Column \rightarrow Constraints  \\
+constraints(col(l, t, val, cons)) = cons  \\ \\
+l \in Labels, t \in DbType, val \in Strings, cons \in Constraints
+\end{gather*}
 
 \paragraph{Foreign key}
-$$
-ForeignKey = (label, tableRef, Constraint*)
-$$
-$name : ForeignKey \rightarrow Label $ \\
-$name(< lab, tRef, cons)) = lab $ \\ \\
-$reference : ForeignKey \rightarrow Label $ \\
-$reference(< lab, tRef, cons)) = tRef $ \\ \\
-$reference : ForeignKey \rightarrow Constraint* $ \\
-$reference(< lab, tRef, cons)) = cons $ \\ \\
-$owningTable : ForeignKey \times Database \rightarrow Table $ \\
-$owningTable(fk, d) = tab, tab \in tables(d) \wedge fk \in foreignKeys(d) $
+\begin{gather*}
+ForeignKey = <label, tableRef, Constraint*> \\ \\
+fk : Label \times Label \times Constraints \rightarrow ForeignKey \\ \\
+name : ForeignKey \rightarrow Label \\
+name(fk(l, tRef, cons)) = lab  \\ \\
+reference : ForeignKey \rightarrow Label  \\
+reference(fk(l, tRef, cons)) = tRef  \\ \\
+reference : ForeignKey \rightarrow Constraint*  \\
+reference(fk(l, tRef, cons)) = cons  \\ \\
+l, tRef \in Label, cons \in Constraints
+\end{gather*}
 
 \paragraph{Primary key}
-$$
-PrimaryKey =  < label > 	
-$$
+\begin{gather*}
+PrimaryKey =  < label > 	\\
+pk : Label \rightarrow PrimaryKey
+\end{gather*}
 
@@
 Both constraints are column constraints - first constraint defines unique records in a column, second constraint defines non-empty columns.
+\begin{gather*}
+Constraint = NOTNULL \; | \; UNIQUE 
+\end{gather*}
 
-\begin{center}
-$Constraint = NOTNULL$ $|$ $UNIQUE $
-\end{center}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \subsection{Model of Stored Data}
+A database consists not only from schema but also from data which are represented as rows in a table.
+A table row in our model is a tuple consisting of reference to a concrete table and a set of value pairs, which represents concrete values of concrete columns.
+\begin{gather*}
+row = < refTable, pair* > \\
+row : Label \times Pairs \rightarrow Row \\ \\
+refTable : Row \rightarrow Label \\
+refTable(row(l, pairs) = l \\ \\
+contains : Table \times Database \rightarrow Boolean \\
+contains(t, d) = \begin{cases}
+ \inference{\exists r \in data(d) : refTable(r) = name(t) }{true} \\ \\
+  \inference{\forall r \in data(d) : refTable(r) \neq name(t)}{false}
+ \end{cases} \\ \\
+ l \in Labels, pairs \in Pairs, t \in tables(d), d \in Database
+\end{gather*}
+There functions $contains$ verify if the table contains any data.
 
-A row in a table is tuple consisting of reference to a concrete table and a set of value pairs, which represents concrete values of concrete columns.
-$$row = < refTable, pair* >$$
-$$pair = < reference, value >$$
+\begin{gather*}
+pair = < refCol, value > \\
+[Label \times Strings] \rightarrow Pair
+\end{gather*}
 The $reference $ is a name of a column, a foreign key or its primary key and $value$ represents concrete value stored in the database.
@@
 \forall r \in data(d), \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) \exists u \in columns(t) \cup foreignKeys(t) : reference(p) = name(u) 
-\end{gathered}
-\end{equation*}
-There are two functions connected with stored data - first of them verify if the table contains any data:
-$$contains : Table \times Database \rightarrow Boolean $$
-\begin{equation*}
-\begin{gathered}
-	contains(t, d) = \begin{cases}
- \inference{\exists r \in data(d) : refTable(r) = name(t) }{true} \\ \\
-  \inference{\forall r \in data(d) : refTable(r) \neq name(t)}{false}
- \end{cases} \\ \\
- t \in tables(d), d \in Database
 \end{gathered}
@@
  	 columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL ) \\ 
-
-\end{gathered}
-} 
+\end{gathered}} 
 \\ \\
@@
 mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}tables(d) \cup  table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), \varnothing), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL) \\
-
 \end{gathered}}
@@
 \subsubsection{Application Creation}
-$$addClass: Class \times Application \rightarrow Application $$
-
-$
-addClass(c, a) = \inference{\forall s \in classes(A) : name(s) \neq name(c)}
-{classes(a) \cup c}
-$
-
-$$AddProperty : Label \times Property \times Application \rightarrow Application $$
-
-$
-AddProperty(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(p) : name(q) \neq name(p)}{properties(c) \cup p }
-$
-$$AddAssociation : Label \times Association \times Application \rightarrow Application $$
-$
-AddAssociation(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(c) \cup s}
-$
+\begin{gather*}
+addClass: Class \times Application \rightarrow Application \\ 
+addClass(c, a) = \inference{\forall s \in classes(a) : name(s) \neq name(c)}{classes(a) \cup c} \\ \\
+c \in Classes, a \in Applications
+\end{gather*}
+\begin{gather*}
+AddProperty : Label \times Property \times Application \rightarrow Application \\
+AddProperty(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(p) : name(q) \neq name(p)}{properties(c) \cup p } \\
+l \in Label, p \in Properties, a \in Applications
+\end{gather*}
+\begin{gather*}
+AddAssociation : Label \times Association \times Application \rightarrow Application \\
+AddAssociation(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(c) \cup s}\\
+l \in Labels, s \in Associations, a \in Applications
+\end{gather*}
 
 \subsubsection{Application Deconstruction}
-$$RemoveClass: Label \times Application \rightarrow Application $$
-$
+\begin{gather*}
+RemoveClass: Label \times Application \rightarrow Application  \\
 RemoveClass(q, a) = \inference{\exists c \in classes(a) : name(c) = q \\
-\wedge not$ $instantiated(c) = \varnothing \wedge not$ $associated(c)}{classes(a) \setminus c }
-$
-
-$$RemoveProperty: Label \times Label \times Application \rightarrow Application $$
-
-$
-RemoveProperty(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{properties(c) \setminus p }
-$
-
-$$RemoveAssociation : Label \times Labels \times Application \rightarrow Application $$
-
-$
-RemoveAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists s \in associations(c) : name(s) = l_a }{associations(c) \setminus s}
-$
+\wedge !\: instantiated(c) = \varnothing \wedge ! \: associated(c)} {classes(a) \setminus c } \\
+q \in Label, a \in Applications
+\end{gather*}
+\begin{gather*}
+RemoveProperty: Label \times Label \times Application \rightarrow Application \\
+RemoveProperty(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{properties(c) \setminus p }\\
+l_c, l_p \in Label, a \in Applications 
+\end{gather*}
+\begin{gather*}
+RemoveAssociation : Label \times Labels \times Application \rightarrow Application \\ 
+RemoveAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists s \in associations(c) : name(s) = l_a }{associations(c) \setminus s} \\
+l_c, l_a \in Labels, a \in Applications
+\end{gather*}
 
@@
 \paragraph{Rename Class}
-$$\rho_{RenC} : Label \times Label \times Application \rightarrow Application $$
-\begin{equation*}
+\begin{gather*}
+\rho_{RenC} : Label \times Label \times Application \rightarrow Application \\ 
 \rho_{RenC}(l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \forall e \in classes(a) : name(e) \neq l_n}{\begin{gathered}
 c.name = l_n \wedge \forall f \in associations(e), \\ e \in classes(a), reference(f) = l_c : f.reference = l_n 
-\end{gathered}
-}
-\end{equation*}
+\end{gathered}}\\
+l_o, l_n \in Label, a \in Applications
+\end{gather*}
 
 \paragraph{Rename Property}
-$$\rho_{RenP} : Label \times Label \times Label \times Application \rightarrow Application $$
-\begin{equation*}
+\begin{gather*}
+\rho_{RenP} : Label \times Label \times Label \times Application \rightarrow Application \\ \\
 	\rho_{RenP}(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_o \\ \wedge \forall e \in properties(c) : name(e) \neq l_n}{\begin{gathered}
 p.name = l_n 
-\end{gathered}
-}
-\end{equation*}
+\end{gathered}}\\
+l_c, l_o, l_n \in Label, a \in Applications
+\end{gather*}
 
 \paragraph{Rename Association}
-$$\rho_{RenA} : Label \times Label \times Label \times Application \rightarrow Application $$
-\begin{equation*}
+\begin{gather*}
+\rho_{RenA} : Label \times Label \times Label \times Application \rightarrow Application \\
 	\rho_{RenP}(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists a \in associations(c) : name(a) = l_o \\ \wedge \forall e \in associations(c) : name(e) \neq l_n}{\begin{gathered}
 a.name = l_n 
-\end{gathered}
-}
-\end{equation*}
+\end{gathered}}\\
+l_c, l_o, l_n \in Label, a \in Applications
+\end{gather*}
 

### 2012-12-30 doc/description.tex
restructurization of the text

@@
 \begin{document}
-\tableofcontents \newpage
 
@@
 \begin{abstract}
-The formal definition of the evolution in context of  software created using an object oriented programming language and a relational database is introduced in this paper.
+The evolution of a software is a common issue during software lifecycle.
+The evolution od code and database are usually separated tasks.
+We discuss the evolution of software in context of  software created using an object oriented programming language and a relational database.
+The formal definition of the evolutionary framework is introduced in this paper.
 Special attention is given to the data stored in databases (instances) and theirs consistency during evolution process.
@@
 \section{Introduction}
-The evolution of an application is a common issue during software development.
-The evolution can occur from many reasons in different parts of application lifecycle.
-The evolution of a software  created using an object oriented programming language and a relational database is usually processed as code refactoring and database migration is processed as a set of SQL scripts.
-The code refactoring is supported by many IDEs and the database migration is supported by many frameworks and tools, nevertheless these tools are usually not capable to solve complicated migration cases or preserve stored data.
-Another disadvantage is the evolution has to be define twice - as a code refactoring and as a database migration.
+The evolution of a software is a common issue during software development.
+The evolution can occur from many reasons in different stages of software lifecycle.
+The evolution of a software created using an object oriented programming language and a relational database is usually processed in two steps - as a code refactoring and as a database migration, which is processed as an SQL script.
+The code refactoring is supported by many IDEs, whereas the database migration is usually processed manually.
+There are frameworks and tools capable of database evolution, nevertheless these tools are usually not capable to solve complicated migration cases or preserve stored data.
+Another disadvantage of such an approach to data evolution is the evolution has to be define twice - as a code refactoring and as a database migration.
 
-This paper provides a formal description of an evolutionary framework, which evolves a database including stored data according to a evolution in application.
+This paper provides a formal definition of an evolutionary framework, which evolves a database including stored data according to a evolution in application.
 There are defined models representing an application and database in the paper, together with functions which are able to perform an evolution of the whole software.
 
-The paper is organized as follows: in Section \ref{sec:evolution} the concept of ORM software and its evolution is discussed, then models of application and database are introduced in Section \ref{sec:appModel} and in Section \ref{sec:dbModel}, the ORM is defined in Section \ref{sec:orm} and the evolutionary transformations are defined in Section \ref{sec:eorm}.
+The paper is organized as follows: in Section \ref{sec:evoIntro} the concept of ORM software and its evolution is discussed, then models of application and database are introduced in Section \ref{sec:appModel} and in Section \ref{sec:dbModel}, the ORM is defined in Section \ref{sec:orm} and the evolutionary transformations are defined in Section \ref{sec:eorm}.
 
 \section{ORM Software}
-\label{sec:evolution}
-The ORM software consist of an application created using object oriented language and relational database.
-The application and database are in consistent state according to a concrete ORM function, otherwise the software cannot work properly.
-An ORM software is illustrated in Figure \ref{fig:appStructure}.
-The software can be defined as a triple:
-\begin{equation*}
-	software = < Application, Database, \rho >
-\end{equation*}
-Models of $Application$ and $Database$ will be introduced in detail in Section \ref{sec:appModel} and in Section \ref{sec:dbModel}, the ORM function $\rho$ is defined in Section \ref{sec:orm}.
-The consistent state of application and database is defined as:
-$$\rho(Application) = Database$$
-
+\label{sec:evoIntro}
 \begin{figure}
-\begin{center}
+\centering
 	\includegraphics[scale=0.3]{./images/system}
 	\caption{The model of one generation of an ORM system consists of application persistent layer and relational database for storing instances.}
-\end{center}
 \label{fig:appStructure}
 \end{figure}
-
-The evolution of an software is a transformation of one consistent state to another consistent state: $$\epsilon : system \rightarrow system $$ We will call consistent states of one software \emph{generations} of the software.
-The evolution is quite easy if we consider only structure of an application and database and ignore instances (stored data).
-The evolution can be described in such a case as:
-\begin{equation*}
-\epsilon(software(App, Db)) = \{a \rightarrow \epsilon_{App}(App)  | software(a, \rho(a), \rho)\}
-\end{equation*}
-It means only the structure of the application is evolved and the new database is obtained as a result od ORM.
-Stored data are usually useless in many cases as their structure does not correspond to the new database schema.
-We believe the data preservation and consistency is a crucial issue during the software evolution.
-
-The evolution process we propose is in Figure \ref{fig:evolution}.
-The transition from the first generation to the second one is based on application evolution which is interpreted on the database level.
-The interpretation of application evolution is called evolutionary object-relation mapping (eORM) and its defined as a function:
-\begin{equation*}
-	\Phi : \epsilon_{app} \rightarrow \epsilon_{db}
-\end{equation*}
-The evolution of the software can be than defined as:
-\begin{equation*}
-\epsilon(software(App, Db)) = \{software(\epsilon_{app}(App), \Phi((\epsilon_{app}, Db), \rho)\}
-\end{equation*}
-
+A software implemented in an object-oriented language, which uses relational database as a data storage is context of this paper.
+The situation is illustrated in the Figure \ref{fig:appStructure}.
+There are three important components of such a software: application (its persistent layer concretely), database (consisting of database schema and stored data) and an ORM (we will use $\rho$ as a symbol for ORM), therefore we define a software a triple:
+\begin{gather*}
+	software = < Application, Database, \rho > \\
+	soft : Applications \times Databases \times ORM \rightarrow Software \\ \\
+	application : Software \rightarrow Application \\
+	application(soft(a, d, \rho)) = a \\ \\
+	database : Software \rightarrow Database \\
+	database(soft(a, d, \rho)) = d \\
+	a \in Applications, d \in Databases
+\end{gather*}
+The software has to be in consistent state so it is beneficial for its user. The state of the software is defined to communicate the sate of the software to its user:
+\begin{gather*}
+	state = Consistent\; s \: | \: Inconsistent \; s, \\
+	s \in Softwares \\ \\
+	state : Software \rightarrow State \\
+	state(s) = \begin{cases}
+ 		\inference{consistent(application(s)) \wedge \\ consistent(database(s)) \wedge \rho(application(s)) = database(s)}{Consistent \; s}\\ \\
+ 		\inference{otherwise}{Inconsistent \; s}
+ 	\end{cases}
+\end{gather*}
+\subsection{Software Evolution}
 \begin{figure}
-\begin{center}
+\centering
 	\includegraphics[scale=0.4]{./images/evolution_simple}
 	\caption{The evolution of data changes the system on all levels. The figure shows all the components of the evolution process.}
-\end{center}
 	\label{fig:evolution}
 \end{figure}
+The evolution of a software is a transformation from one consistent state to another one.
+These states are called generations of the software.
+The evolution consists of change in application and in database as shown in the Figure \ref{fig:evolution}.
+We focus on situations when the evolution is initiated on the application level (as code refactoring), however the evolution has to be processed on the database level too.
+It means there is a set of application transformations called $AppOps$ and a set of database transformations called $DbOps$ and a mapping between these two sets.
+We call this mapping $\Phi$:
+\begin{gather*}
+\Phi : AppOps \rightarrow DbOps
+\end{gather*}
 
-\section{Application Model}
+The evolution of the software is defined as the $\oplus$ operator:
+$$\oplus : State \times AppOps \rightarrow State $$
+and similarly the $\oplus_{app}$ and the $\oplus_{db}$ operators for evolving application and database:
+\begin{gather*}
+\oplus_{app} : Application \times AppOps \rightarrow Application \\
+ \oplus_{db} : Database \times DbOps \rightarrow Database
+\end{gather*}
+The definition of the evolving operator is inspired by the monads (as used e.g.
+in Haskell) so the state of the evolution can be passed from one evolution step to another, which helps us to concatenate evolution transformations.
+\begin{gather*}
+ u \oplus t = \begin{cases}
+ 	\inference{u = Inconsistent \; s}{u} \\ \\
+ 	\inference{u = Consistent \; s}{state(soft(application(u) \oplus_{app} t, database(u) \oplus_{db} \Phi(t))} 
+\end{cases}\\ 
+u \in States, t \in AppOps, s \in Softwares
+\end{gather*}
+\begin{gather*}
+ a \oplus o = \begin{cases}
+ 	\inference{!\;consistent(a)}{a} \\ \\
+ 	\inference{consistent(a)}{a \oplus_{app} o} 
+\end{cases}\\ 
+a \in Applications, o \in AppOps
+\end{gather*}
+\begin{gather*}
+ d \oplus o = \begin{cases}
+ 	\inference{!\;consistent(d)}{d} \\ \\
+ 	\inference{consistent(d)}{d \oplus_{app} o} 
+\end{cases}\\ 
+d \in Applications, o \in DbOps
+\end{gather*}
+
+\section{Software Model}
+The model of an ORM software consists of three parts - an application model, a database model and a ORM.
+
+\subsection{Application Model}
 \label{sec:appModel}
@@
 
-%%%%%%%%%%%%%%%%%%%
-
-\section{Database Model}
+\subsection{Database Model}
 \label{sec:dbModel}
@@
 
-\subsection{Database Schema Model}
+\subsubsection{Database Schema Model}
 The database schema model is defined by following concepts:
@@
 
-\subsection{Model of Stored Data}
+\subsubsection{Model of Stored Data}
 A database consists not only from schema but also from data which are represented as rows in a table.
@@
 
-\section{Object-Relational Mapping}
+\subsection{Object-Relational Mapping}
 \label{sec:orm}
@@
 
-\subsection{Mapping of Types}
+\subsubsection{Mapping of Types}
 The mapping of types assumes a bijection between application types and database types, otherwise these has to be additional information for type mapping.
@@
 
-\subsection{Mapping of Properties}
+\subsubsection{Mapping of Properties}
 There are two types of properties - the first type has cardinality 1  and it is mapped to a column.
@@
 
-\subsection{Mapping of Associations}
+\subsubsection{Mapping of Associations}
 A association is mapped into database as a foreign key in table representing associating class or an association can be mapped as a coupling table in case the association cardinality is M:N or 1:N.
@@
 
-\section{Evolution}
-\label{sec:eorm}
-There are introduced operations for evolution of an application first as they are a source for the evolution of the whole ORM system.
-The mapping of the application evolution into a database schema and stored data are introduced later in this section too.
+\section{Evolutionaty Transformations}
+Each evolutionary transformation consist in changes of the application (a transformation from the $AppOps$ set) and the database (a transformation from the  $DbOps$ set).
+The function $\Phi$ provides a mapping between the $AppOps$ and $DbOps$ sets, which makes the $AppOps$ set a source for the transformation of the whole software.
 
-\subsection{Evolution of Application}
-\label{sec:appEvolution}
-
-The evolution of a software is defined as a set of transformations which change a structure of the software.
-The definition of an operation consists of two parts - a conditions of operation feasibility and operation impact on the application structure.
-If the conditions of feasibility are not met, the operation cannot change the application structure - the software cannot evolve.
-This is important in case we have an evolution consisting of several evolution operations.
-These operations has to be proceed in concrete order and all of them has to change the software otherwise the software can become inconsistent from user point of view (the structure or stored data will not correspond with the users' reality).
-To assure the right behavior of evolution operations and to propagate operations fail to the user we extend the definition of a software to make it a monad.
-The extension consists in defining a software state and operations $ret$ and $\oplus$ (called evolve), which is used as a concatenation of operations as well.
+The definition of a transformation consists of two parts - a conditions of transformation feasibility and its impact on the context.
+If the conditions of feasibility are not met, the transformation cannot change the context - the software cannot evolve.
+We decide to use following transcription to cover all aspects  of transformation definition:
 \begin{gather*}
-state = Consistent \; s \: | \: Inconsistent \; s, v \in Softwares 
-\end{gather*}
-
-\begin{equation*}
-\begin{gathered}
-ret : Software \rightarrow State \\ \\
-	ret(s) = \begin{cases}
- Consistent \; s, consistent(s)
- \\
- Inconsistent\;  s, ! consistent(s)
- \end{cases}
-\end{gathered}
-\end{equation*}
-Each evolution transformation has a definition: $Software \rightarrow State$ and all evolution transformations creates a set called $Transformations$, then a concatenation of evolution transformation is defined as:
-\begin{equation*}
-\begin{gathered}
- \oplus : State \times Transformations \rightarrow State  \\ \\
- u \oplus t = \begin{cases}
- u, \; u = Inconsistent \; s \\
- t(u), \; u = Consistent \; s 
-\end{cases}\\ 
-u \in States, t \in Transformations, s \in Softwares
-\end{gathered}
-\end{equation*}
-Each transformation consist in changes of the application and the database.
-To describe these two aspects od software evolution we decide to define transformations for each layer separately.
-The separation of transformation contexts help us to define transformation specific for application or database evolution, which can ease both - understanding and defining the transformation.
-The transformations specific for a context are called operations and they creates two sets- first one for application evolution: $AppOps$  and second one for database evolution $DbOps$.
-Applying of a transformation is defined:
-\begin{gather*}
-u \oplus t = ret(<application(u) \oplus o_{app}, database(u) \oplus o_{db}>) \\
-u \in Softwares, t \in Transformations, o_{app} \in AppOps, o_{db} \in DbOps
-\end{gather*}
-The function $\Phi$ provides a mapping between the $AppOps$ and $DbOps$ sets.
-The $\Phi$ function makes possible to use a description of an application evolution as a source for the evolution of the whole software.
-\begin{gather*}
-\Phi : AppOps \rightarrow DbOps
-\end{gather*}
-
-The definition of an operation consists of two parts - a conditions of operation feasibility and operation impact on the context.
-We decide to use following transcription to cover all aspects  of operation definition:
-
-\begin{equation*}
-\begin{gathered}
 operation(parameters, context) = \begin{cases}
   \inference{conditions \; of \; feasibility}{impact \; on \; context}\\ \\
-  \inference{otherwise}{Inconsistent \; context}
+  \inference{otherwise}{context}
  \end{cases}
-\end{gathered}
-\end{equation*}
+\end{gather*}
 
@@
 
+\subsection{Application Transformation}
 \subsubsection{Application Creation}
@@
 
-\paragraph{Copy Property}
-All transformations described earlier does change only the structure of a application.
-These operation can be used in case there are no instances in the application.
-If instances are present in the application, the decomposition transformations could cause a data lost.
-Therefore we define  the transformation for copying a property from one class to another.
-The copy property is the basic way to copy data from instance of a class to instance of another class.
-
-The copy property can be implemented as a transformation composed by operations for adding and removing property, nevertheless this composition would not work with instances, thus we define the copy property operation as a basic and atomic operation which will help us to handle instances.
-
-$$\rho_{copyP} : Label \times Label \times Label \times (Instance \times Instance) \times Application \rightarrow Application $$
-\begin{equation*}
-	\rho_{copy}(l_s, l_t, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p \\ \wedge \exists e \in classes(a) : name(e) = l_t  \\ \wedge \forall q \in properties(c) : name(q) \neq l_p}{\begin{gathered}
-properites(e) \cup p 
-\end{gathered}
-}
-\end{equation*}
-
-With a copy property transformation defined we are able to define advanced evolution scenarios.
-
-\subsection{Composing of Evolution Operations}
-
-\paragraph{Change Cardinality}
-$$\rho_{CcP} : Property \times \mathbb{N^{*}} \times Application \rightarrow Property $$
-$$\rho_{cardA} : Association \times \mathbb{N^{*}} \times Application \rightarrow Association $$
-
-\subsection{Extended Object-Relation Mapping}
-The extended object-relational mapping maps the evolution of the application level into the database schema and stored data.
-This section introduces the mapping for evolution operations introduced earlier.
-
-\subsubsection{Creation of a Software}
+\subsection{Database Transformation}
 \paragraph{Add Class} The $AddClass$ operation changes the structure of the database schema only.
@@
 $$
-
 \paragraph{Add Property} The function $AddProperty$ on level of database does not change the structure only, but it has an impact on stored data too.
@@
 
+\subsection{Composing of Evolution Operations}
+\paragraph{Copy Property}
+All transformations described earlier does change only the structure of a application.
+These operation can be used in case there are no instances in the application.
+If instances are present in the application, the decomposition transformations could cause a data lost.
+Therefore we define  the transformation for copying a property from one class to another.
+The copy property is the basic way to copy data from instance of a class to instance of another class.
+
+The copy property can be implemented as a transformation composed by operations for adding and removing property, nevertheless this composition would not work with instances, thus we define the copy property operation as a basic and atomic operation which will help us to handle instances.
+
+$$\rho_{copyP} : Label \times Label \times Label \times (Instance \times Instance) \times Application \rightarrow Application $$
+\begin{equation*}
+	\rho_{copy}(l_s, l_t, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p \\ \wedge \exists e \in classes(a) : name(e) = l_t  \\ \wedge \forall q \in properties(c) : name(q) \neq l_p}{\begin{gathered}
+properites(e) \cup p 
+\end{gathered}
+}
+\end{equation*}
+
+With a copy property transformation defined we are able to define advanced evolution scenarios.
+
+\paragraph{Change Cardinality}
+$$\rho_{CcP} : Property \times \mathbb{N^{*}} \times Application \rightarrow Property $$
+$$\rho_{cardA} : Association \times \mathbb{N^{*}} \times Application \rightarrow Association $$
+
+\section{Summary}
+
 \end{document}
+% This is important in case we have an evolution consisting of several evolution operations.
+These operations has to be proceed in concrete order and all of them has to change the software otherwise the software can become inconsistent from user point of view (the structure or stored data will not correspond with the users' reality).
+To assure the right behavior of evolution operations and to propagate operations fail to the user we extend the definition of a software to make it a monad.
+The extension consists in defining a software state and operations $ret$ and $\oplus$ (called evolve), which is used as a concatenation of operations as well.
+%
+%Each evolution transformation has a definition: $Software \rightarrow State$ and all evolution transformations creates a set called $Transformations$, then a concatenation of evolution transformation is defined as:
+%\begin{equation*}
+%\begin{gathered}
+% \oplus : State \times Transformations \rightarrow State  \\ \\
+% u \oplus t = \begin{cases}
+% u, \; u = Inconsistent \; s \\
+% t(u), \; u = Consistent \; s 
+%\end{cases}\\ 
+%u \in States, t \in Transformations, s \in Softwares
+%\end{gathered}
+%\end{equation*}
+%
+% To describe these two aspects od software evolution we decide to define transformations for each layer separately.
+The separation of transformation contexts help us to define transformation specific for application or database evolution, which can ease both - understanding and defining the transformation.
+The transformations specific for a context are called operations and they creates two sets- first one for application evolution: $AppOps$  and second one for database evolution $DbOps$.
+Applying of a transformation is defined:
+%\begin{gather*}
+%u \oplus t = ret(<application(u) \oplus o_{app}, database(u) \oplus o_{db}>) \\
+%u \in Softwares, t \in Transformations, o_{app} \in AppOps, o_{db} \in DbOps
+%\end{gather*}

### 2012-12-31 doc/description.tex
set of sw Evolution defined

@@
 
-This paper provides a formal definition of an evolutionary framework, which evolves a database including stored data according to a evolution in application.
-There are defined models representing an application and database in the paper, together with functions which are able to perform an evolution of the whole software.
+There are three main goals of our evolutionary framework: 1) provide one source of evolution definition for both: code and database 2) provide a tool, which will be able to preserve stored data 3) minimize the manual work during data evolution in a software.
+
+This paper provides a formal definition of such a framework, which is based on model driven architecture.
+There are defined models representing an application and database in this paper, together with transformations which are able to perform an evolution of the whole software and preserve data.
 
@@
 We focus on situations when the evolution is initiated on the application level (as code refactoring), however the evolution has to be processed on the database level too.
-It means there is a set of application transformations called $AppOps$ and a set of database transformations called $DbOps$ and a mapping between these two sets.
-We call this mapping $\Phi$:
+It means there is a set of application transformations called $AppOps$ and a set of database transformations called $DbOps$.
+Because there are needed different parameters for application and database evolution the set of transformation called $E$ which evolves the whole software and a mapping between $E$ and $AppOps$, respectively $DbOps$ is defined.
+We call this mappings $\Psi$ and $\Phi$:
 \begin{gather*}
-\Phi : AppOps \rightarrow DbOps
+\Psi : E \rightarrow AppOps \\
+\Phi : E \rightarrow DbOps
 \end{gather*}
+A subset of $E$ represents an evolution transformation of a software, which mets the first goal of our framework - a single source of data evolution for the whole software.
 
-The evolution of the software is defined as the $\oplus$ operator:
-$$\oplus : State \times AppOps \rightarrow State $$
-and similarly the $\oplus_{app}$ and the $\oplus_{db}$ operators for evolving application and database:
+The application of a transformation on a software state (evolution) is marked by the $\oplus$ operator:
+$$\oplus : State \times E \rightarrow State $$
+and similarly the $\circ$ and the $\bullet$ operators for evolving application and database:
 \begin{gather*}
-\oplus_{app} : Application \times AppOps \rightarrow Application \\
- \oplus_{db} : Database \times DbOps \rightarrow Database
+\circ : Application \times AppOps \rightarrow Application \\
+ \bullet : Database \times DbOps \rightarrow Database
 \end{gather*}
 The definition of the evolving operator is inspired by the monads (as used e.g.
-in Haskell) so the state of the evolution can be passed from one evolution step to another, which helps us to concatenate evolution transformations.
+in Haskell programming language) so the state of the software can be passed from one evolution step to another, which helps us to concatenate evolution transformations.
+
 \begin{gather*}
@@
  	\inference{u = Inconsistent \; s}{u} \\ \\
- 	\inference{u = Consistent \; s}{state(soft(application(u) \oplus_{app} t, database(u) \oplus_{db} \Phi(t))} 
+ 	\inference{u = Consistent \; s}{state(soft(application(u) \circ \Psi(t), database(u) \bullet \Phi(t))} 
 \end{cases}\\ 
-u \in States, t \in AppOps, s \in Softwares
+u \in States, t \in E, s \in Softwares
 \end{gather*}
 \begin{gather*}
- a \oplus o = \begin{cases}
+ a \circ o = \begin{cases}
  	\inference{!\;consistent(a)}{a} \\ \\
- 	\inference{consistent(a)}{a \oplus_{app} o} 
+ 	\inference{consistent(a)}{a \circ o} 
 \end{cases}\\ 
@@
 \begin{gather*}
- d \oplus o = \begin{cases}
+ d \bullet o = \begin{cases}
  	\inference{!\;consistent(d)}{d} \\ \\
- 	\inference{consistent(d)}{d \oplus_{app} o} 
+ 	\inference{consistent(d)}{d \circ o} 
 \end{cases}\\ 
@@
 \section{Evolutionaty Transformations}
-Each evolutionary transformation consist in changes of the application (a transformation from the $AppOps$ set) and the database (a transformation from the  $DbOps$ set).
+Each evolutionary transformation of a software consist in changes of the application (a transformation from the $AppOps$ set) and the database (a transformation from the  $DbOps$ set).
 The function $\Phi$ provides a mapping between the $AppOps$ and $DbOps$ sets, which makes the $AppOps$ set a source for the transformation of the whole software.

### 2012-12-31 doc/description.tex
changed the structure of the chapter about Transformations

@@
 	software = < Application, Database, \rho > \\
-	soft : Applications \times Databases \times ORM \rightarrow Software \\ \\
+	sw : Applications \times Databases \times ORM \rightarrow Software \\ \\
 	application : Software \rightarrow Application \\
-	application(soft(a, d, \rho)) = a \\ \\
+	application(sw(a, d, \rho)) = a \\ \\
 	database : Software \rightarrow Database \\
-	database(soft(a, d, \rho)) = d \\
+	database(sw(a, d, \rho)) = d \\
 	a \in Applications, d \in Databases
@@
  	\inference{u = Inconsistent \; s}{u} \\ \\
- 	\inference{u = Consistent \; s}{state(soft(application(u) \circ \Psi(t), database(u) \bullet \Phi(t))} 
+ 	\inference{u = Consistent \; s}{state(sw(\Psi(u, t), \Phi(u,t))} 
 \end{cases}\\ 
@@
 The model of an ORM software consists of three parts - an application model, a database model and a ORM.
+The definitions of all parts assume a set $Label$ exist which contains all possible identifiers and each modeled entity is identified unambiguously identified in the model.
+Next there is a set $Values$ containing all possible values, which can be used in the model.
 
@@
 It means the layer is a set of instances which cooperates together.
-Each instance is defined according to a class - the class defines the structure and relationships of the instance.
+Each instance is defined according to a class - the class defines the structure of its instances and relationships between them an the rest of the application.
 The instances are stored in the relational database in our case, therefore we define only the structure of classes in our model.
+Besides the structure definition, there is defined a set of functions, which allows to manipulate the application model easily.
 
-The model uses a set $Labels$ which contains all possible identifiers for the model. Each model component which has a label ($label \in Labels$) is unambiguously identified in the model.
 \begin{gather*}
-  Application = < Class* >
-  \\ 
+  Application = < Class* > \\ 
   app : Class* \rightarrow  Application \\ \\
@@
   associated :  Label \times Application \rightarrow Boolean \\
-  owningClass : Property \times Application \rightarrow Class  \\ owningClass(p, a) = c, c \in classes(a) \wedge p \in properties(c) \\
-  owningClass : Association \times Application \rightarrow Class  \\
-owningClass(as, a) = c, c \in classes(a) \wedge as \in associations(c) \\ \\
-owningTable : Column \times Database \rightarrow Table  \\
-owningTable(c, d) = tab, tab \in tables(d) \wedge c \in columns(d) \\ \\
-owningTable : ForeignKey \times Database \rightarrow Table  \\
-owningTable(fk, d) = tab, tab \in tables(d) \wedge fk \in foreignKeys(d)\\ \\ 
  \end{gather*}
-The structure of an application consists not only of classes, it is defined by five other concepts.
-Besides the structure definition, there is defined a set of functions, which allows to manipulate the application model easily.
 
@@
 
-\paragraph{Property} represents a feature of  a class which is can be represented as a primitive type. The property can be mandatory, can have a default value.
+\paragraph{Property} represents a feature of  a class which is represented as a primitive type.
+The property can be mandatory, can have a default value and can represent a single value or a collection of values.
 \begin{gather*}
@@
 It has a unique name and it contains a name of the class which is referenced by the association.
-The class which owns the association is consider to be a starting class of an association, referenced class is consider to be an ending class of an association.
-The cardinalities defines the multiplicity of the association.
+The class which owns the association is consider to be the starting class of an association, referenced class is consider to be the ending class of an association.
+The cardinalities defines the multiplicity of the association ends.
 \begin{gather*}
@@
 
-\paragraph{DbType} represents primitive types in the database. There are usually defined types such as Char, Integer, Boolean etc. There is only one type defined in the model which is called DBSTRING.
+\paragraph{DbType} represents primitive types in the database.
+There are usually defined types such as Varchar, Integer, Boolean etc.
+There are only two types defined in the model $DBSTRING$ as a default type of values in columns and $DBINT$ as a default type for keys.
 $$
-DbType = DBSTRING
+DbType = DBSTRING \; | \; DBINT 
 $$
@@
  \inference{\exists r \in data(d) : refTable(r) = name(t) }{true} \\ \\
-  \inference{\forall r \in data(d) : refTable(r) \neq name(t)}{false}
+  \inference{otherwise}{false}
  \end{cases} \\ \\
@@
 \end{gather*}
-There functions $contains$ verify if the table contains any data.
+There functions $contains$ verify if the table contains data in context of concrete database.
 
@@
 pair = < refCol, value > \\
-[Label \times Strings] \rightarrow Pair
+[Label \times Value] \rightarrow Pair\\ \\
+\mathbf{reference} : Pair \rightarrow Label \\
+reference([l,val]) = l \\ \\
+\mathbf{value} : Pair \rightarrow Label \\ 
+value([l,val]) = val \\
+l \in Label, val \in Value
 \end{gather*}
-The $reference $ is a name of a column, a foreign key or its primary key and $value$ represents concrete value stored in the database.
-Each row in a database has to refer to a table in database schema, otherwise the database is inconsistent.
+The $refCol$ is a name of a column, a foreign key or its primary key and $value$ represents concrete value stored in the database.
+Each row in a database has to refer to a table in database schema and has a structure corresponding to the one defined by a table, otherwise the database is inconsistent.
 The consistency of database is then:
-\begin{equation*}
-\begin{gathered}
-\forall r \in data(d), \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) \exists u \in columns(t) \cup foreignKeys(t) : reference(p) = name(u) 
-\end{gathered}
-\end{equation*}
+\begin{gather*}
+\mathbf{consistent} : Database \rightarrow Boolean \\
+consistent(d) = \begin{cases}
+ 	\inference{\forall r \in data(d), \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) :\\ \exists u \in columns(t) \cup foreignKeys(t) : reference(p) = name(u)}{true}
+ 	\\ \\
+ 	\inference{otherwise}{false}
+ \end{cases}
+\end{gather*}
 
-second of them verify if there are instances of a class:
+\subsection{Extension of Software Definition}
+
+\begin{gather*}
+ owningClass : Property \times Application \rightarrow Class  \\ owningClass(p, a) = c, c \in classes(a) \wedge p \in properties(c) \\
+  owningClass : Association \times Application \rightarrow Class  \\
+owningClass(as, a) = c, c \in classes(a) \wedge as \in associations(c) \\ \\
+owningTable : Column \times Database \rightarrow Table  \\
+owningTable(c, d) = tab, tab \in tables(d) \wedge c \in columns(d) \\ \\
+owningTable : ForeignKey \times Database \rightarrow Table  \\
+owningTable(fk, d) = tab, tab \in tables(d) \wedge fk \in foreignKeys(d)\\ \\ 
+\end{gather*}
+
+The database stores instances of the application, thus the $instantiated$ function has to be defined:
+
 $$instantiated : Class \times Software \rightarrow Boolean $$ 
@@
  \inference{\exists t \in tables(database(s)) : name(t) = name(c) \\ \wedge contains(t, database(s))}{true} \\ \\
-  \inference{\exists t \in tables(database(s)) : name(t) = name(c) \\ \wedge ! contains(t, database(s))}{false} \\ \\
+  \inference{otherwise}{false} \\ \\
  \end{cases} \\ \\
@@
 \begin{center}
-$\rho_{c}: Class \rightarrow Table $ \\ 
+$\rho_{c}: Class \times Database \rightarrow Database $ \\ 
 $\rho_{c}(c,d) = tables(d) \cup table(name(c), primaryKey("Id"), \varnothing, \varnothing) $
@@
 $
-\rho_p : Property \times Database \rightarrow Column 
+\rho_p : Property \times Database \rightarrow Database 
 $
@@
 $\rho_p(p, d) = \begin{cases}
-  \inference{cardinality(p) = 1 \wedge ! mandatory(p)
+  \inference{cardinality(p) = 1 \wedge !\;mandatory(p)
  \\ \wedge \exists t : t = table(name(owningClass(p)), d) 
@@
 
- \inference{ cardinality(p) > 1 \wedge !
-mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}tables(d) \cup  table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), \varnothing), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL) \\
+ \inference{ cardinality(p) > 1 \wedge !\;mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}tables(d) \cup  table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), \varnothing), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL) \\
 \end{gathered}}
@@
 \section{Evolutionaty Transformations}
-Each evolutionary transformation of a software consist in changes of the application (a transformation from the $AppOps$ set) and the database (a transformation from the  $DbOps$ set).
-The function $\Phi$ provides a mapping between the $AppOps$ and $DbOps$ sets, which makes the $AppOps$ set a source for the transformation of the whole software.
+The set of all possible data evolution cases $E$ was defined in Section \ref{sec:evoIntro} together with sets $AppOps$ and $DbOps$ for evolving various layers of the software.
+The semantic of transformations in the set $E$ is defined by their interpretation on individual software layers.
+This section provides an overview of basic transformations (evolution cases) and theirs interpretation on the application and the database.
+The section starts with simple transformations as the complex one can be often implemented as theirs concatenation.
+The transformations are divided into three groups - first group creates the structure of a software, second group removes parts of a software structure and third group is able to alter the structure.
 
-The definition of a transformation consists of two parts - a conditions of transformation feasibility and its impact on the context.
+The definition of a transformation consists of two parts - a conditions of transformation feasibility and its impact on the context (application or database).
 If the conditions of feasibility are not met, the transformation cannot change the context - the software cannot evolve.
@@
 \end{gather*}
-
 The second branch of the operation definition is an implicit assumption - if an operation is not feasible in a software, its state become inconsistent.
@@
 
-\subsection{Application Transformation}
-\subsubsection{Application Creation}
+\subsection{Software Creation}
+There are defined three transformations which are able to create a software: $newClass$, $addProperty$, $addAssociation$.
+These three transformations can be used to create a software from a scratch.
+The first of them has no impact on stored data, whereas the rest of them can in some cases.
+\begin{gather*}
+\mathbf{addProperty} : Label \times Property \times State \rightarrow State \\
+\Psi(s, addProperty(l, p)) = application(s) \circ addProp(l,p) \\
+\Phi(s, addProperty(l, p)) = database(s) \bullet insertProp(l, p) \\
+s \in Softwares, l \in Label, p \in Properties
+\end{gather*}
+Adding a property into an application structure has to verify a name collision only:
+\begin{gather*}
+addProp : Label \times Property \times Application \rightarrow Application \\
+addProp(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(p) : name(q) \neq name(p)}{properties(c) \cup p } \\
+l \in Label, p \in Properties, a \in Applications
+\end{gather*}
+Whereas the $insertProp$ on database level can change stored data too, because a mandatory property can be added into a class with already existing instances, then a default value has to be added to all existing instances.
+\begin{gather*}
+insertProp(p, d) = \begin{cases}
+\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge (!\;mandatory(p) \vee !\;instantiated(owningClass(p))}{ \rho(p,d)} 
+\\ \\ 
+\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
+ \rho(p,d) \wedge \forall r \in data(d), reference(r) = name(t) : pairs(r) \\ \cup [name(p), defaultValue(p)]
+\end{gathered}
+} 
+\\ \\
+\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
+ \rho(p,d) \wedge \forall r \in row(owningClass(p)) : data(d)\\ \cup row(name(p), ["value", defaultValue(p)], \\ [name(p), value(primaryKey(r))]) 
+\end{gathered}
+} 
+\end{cases}\\
+p \in Properties, d \in Dataabses
+\end{gather*}
+
+\begin{gather*}
+\mathbf{addAsociation} : Label \times Association \times (Row \times Row) \times State \rightarrow State \\
+\Psi(s, addAsociation(l, a, f_m)) = application(s) \circ addAssoc(l, a) \\
+\Phi(s, addAsociation(l, a, f_m)) = database(s) \bullet insertAssoc(l, a, f_m)
+\end{gather*}
+Adding an association into an application structure has to verify if there is no name collision and if there is the associated class present in the model.
+\begin{gather*}
+addAssoc : Label \times Association \times Application \rightarrow Application \\
+addAssoc(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(c) \cup s}\\
+l \in Labels, s \in Associations, a \in Applications
+\end{gather*}
+Adding an association into a database is complicated because of already existing instances.
+When an association between classes is added the instances of both classes has to be paired.
+This pairing is provided by the function $f_m$:
+$$f_m : \rightarrow Instance \times Instance $$
+
+\textbf{TODO: define the features of the $f_m$ function}
+
+The $insertAssoc$ function is defined as:
+\begin{gather*}
+insertAssoc : Association \times Database \times (Row \times Row) \\
+insertAssoc(a, d, f_m) = \begin{cases}
+\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
+\rho(a,d) \wedge \forall r \in f_m(owningClass(a), reference(a)) : \\ data(t_t) \cup row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]) 
+\end{gathered}
+ }
+ \\ \\
+\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) > 1}{\begin{gathered}
+\rho(a,d) \wedge \forall r \in f_m(owningClass(a), reference(a)),\\ t \in tables(d) : name(t) = name(a) : \\ data(t) \cup row(name(a), [name(t_s), \\ value(primaryKey(first(r)))] [name(t_t), \\ value(primaryKey(second(r)))])
+\end{gathered}
+ } 
+ \end{cases}\\
+ a \in Associations, d \in Databases
+\end{gather*}
+
+\begin{gather*}
+\mathbf{newClass} : Class \times State \rightarrow State \\
+\Psi(s, newClass(c)) = application(s) \circ addClass(c) \\
+\Phi(s, newClass(c)) = database(s) \bullet insertClass(c) 
+\end{gather*}
 \begin{gather*}
@@
 \end{gather*}
+\paragraph{Add Class} The $insertClass$ operation changes the structure of the database schema only.
 \begin{gather*}
-AddProperty : Label \times Property \times Application \rightarrow Application \\
-AddProperty(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(p) : name(q) \neq name(p)}{properties(c) \cup p } \\
-l \in Label, p \in Properties, a \in Applications
+insertClass : Class \times Database \rightarrow Database \\
+insertClass(c, d) = \inference{ \forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
+	\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) 
+}{
+\begin{gathered}
+  \rho(c) \wedge \rho(singleProperties(c)) \wedge \rho(collectionProperties(c)) \\ \wedge \rho(associations(c))
+\end{gathered}
+}
 \end{gather*}
-\begin{gather*}
-AddAssociation : Label \times Association \times Application \rightarrow Application \\
-AddAssociation(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(c) \cup s}\\
-l \in Labels, s \in Associations, a \in Applications
-\end{gather*}
+
+\subsection{Software Decomposition}
+\subsection{Software Alternation}
 
@@
 RemoveClass(q, a) = \inference{\exists c \in classes(a) : name(c) = q \\
-\wedge !\: instantiated(c) = \varnothing \wedge ! \: associated(c)} {classes(a) \setminus c } \\
+\wedge !\: instantiated(c) = \varnothing \wedge !\: associated(c)} {classes(a) \setminus c } \\
 q \in Label, a \in Applications
@@
 \subsection{Database Transformation}
-\paragraph{Add Class} The $AddClass$ operation changes the structure of the database schema only.
-$$
-\Phi_{AC}(c, d) = \inference{ \forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
-	\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) 
-}{
-\begin{gathered}
-  \rho(c) \wedge \rho(singleProperties(c)) \wedge \rho(collectionProperties(c)) \\ \wedge \rho(associations(c))
-\end{gathered}
-}
-$$
-\paragraph{Add Property} The function $AddProperty$ on level of database does not change the structure only, but it has an impact on stored data too.
-This is because a mandatory property can be added into a class with already existing instances, then a default value has to be added to all existing instances.
-$$
-\Phi_{AP}(p, d) = \begin{cases}
-\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge (! mandatory(p) \vee ! instantiated(owningClass(p))}{ \rho(p,d)} 
-\\ \\ 
-\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
- \rho(p,d) \wedge \forall r \in data(d), reference(r) = t : pair(r) \\ \cup pair(name(p), defaultValue(p))
-\end{gathered}
-} 
-\\ \\
-\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
- \rho(p,d) \wedge \forall r \in row(owningClass(p)) : data(d)\\ \cup row(name(p), pair("value", defaultValue(p)), \\ pair(name(p), value(primaryKey(r))) 
-\end{gathered}
-}
-
-\end{cases}
-$$
-
-\paragraph{Add Association} Adding an association into a database is complicated because of already existing instances.
-When an association between classes is added the instances of both classes has to be paired.
-This pairing is provided by the function $f_m$:
-$$f_m : Property \rightarrow Instance \times Instance $$
-
-\textbf{TODO: define the features of the $f_m$}
-
-The mapping of the $AddAssociation$ function is defined as:
-$$
-\Phi_{AA}(a, f_m, d) = \begin{cases}
-\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
-\rho(a,d) \wedge \forall r \in f_m(owningClass(a), reference(a)) : \\ data(t_t) \cup row(name(t_s), pair(name(a), \\ value(primaryKey(first(r)))) 
-\end{gathered}
- }
- \\ \\
-\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) > 1}{\begin{gathered}
-\rho(a,d) \wedge \forall r \in f_m(owningClass(a), reference(a)),\\ t \in tables(d) : name(t) = name(a) : \\ data(t) \cup row(name(a), pair(name(t_s), \\ value(primaryKey(first(r)))) pair(name(t_t), \\ value(primaryKey(second(r)))))
-\end{gathered}
- }
-
- \end{cases}
-$$
 

### 2013-01-02 doc/description.tex
changes in organization of Transformation chapter
reorganization of transformations

@@
  }{\begin{gathered}
-	  columns(t) \cup column(name(p), \rho_t(type(p)), \varnothing )
+	  column(name(p), \rho_t(type(p)), \varnothing )
 \end{gathered}
@@
  \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}
- 	 columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL ) \\ 
+ 	 column(name(p), \rho_t(type(p)), NOTNULL ) \\ 
 \end{gathered}} 
@@
 
- \inference{ cardinality(p) > 1 \wedge mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d) }{\begin{gathered} tables(d) \cup  table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), NOTNULL), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL)
+ \inference{ cardinality(p) > 1 \wedge mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d) }{\begin{gathered}  table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), NOTNULL), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL)
 	 \end{gathered}}
@@
 insertProp(p, d) = \begin{cases}
-\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge (!\;mandatory(p) \vee !\;instantiated(owningClass(p))}{ \rho(p,d)} 
+\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge (!\;mandatory(p) \vee !\;instantiated(owningClass(p))}{ columns(t) \cup \rho(p,d)} 
 \\ \\ 
-\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
- \rho(p,d) \wedge \forall r \in data(d), reference(r) = name(t) : pairs(r) \\ \cup [name(p), defaultValue(p)]
+\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered} columns(t) \cup 
+ \rho(p,d) \\ \wedge \forall r \in data(d), reference(r) = name(t) : pairs(r) \\ \cup [name(p), defaultValue(p)]
 \end{gathered}
@@
 \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
- \rho(p,d) \wedge \forall r \in row(owningClass(p)) : data(d)\\ \cup row(name(p), ["value", defaultValue(p)], \\ [name(p), value(primaryKey(r))]) 
+tables(d) \cup  \rho(p,d) \wedge \forall r \in row(owningClass(p)) : data(d)\\ \cup row(name(p), ["value", defaultValue(p)], \\ [name(p), value(primaryKey(r))]) 
 \end{gathered}
@@
 \inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
-\rho(a,d) \wedge \forall r \in f_m(owningClass(a), reference(a)) : \\ data(t_t) \cup row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]) 
+columns(t_s) \cup \rho(a,d) \\ \wedge \forall r \in f_m(owningClass(a), reference(a)) : \\ data(t_t) \cup row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]) 
 \end{gathered}
@@
  \\ \\
-\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) > 1}{\begin{gathered}
-\rho(a,d) \wedge \forall r \in f_m(owningClass(a), reference(a)),\\ t \in tables(d) : name(t) = name(a) : \\ data(t) \cup row(name(a), [name(t_s), \\ value(primaryKey(first(r)))] [name(t_t), \\ value(primaryKey(second(r)))])
+\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) > 1}{\begin{gathered} tables(d) \cup
+\rho(a,d) \\ \wedge \forall r \in f_m(owningClass(a), reference(a)),\\ t \in tables(d) : name(t) = name(a) : \\ data(t) \cup row(name(a), [name(t_s), \\ value(primaryKey(first(r)))] [name(t_t), \\ value(primaryKey(second(r)))])
 \end{gathered}
@@
 \end{gather*}
-\paragraph{Add Class} The $insertClass$ operation changes the structure of the database schema only.
+\paragraph{Add Class} The $insertClass$ operation changes the structure of the database schema only, because the $insertAssoc$ transformation needs a mapping function to work properly, which is not part of the $newClass$ definition.
+The situation reflect the modeling praxis when the structure of a class is defined and then it is associations are created.
 \begin{gather*}
@@
 insertClass(c, d) = \inference{ \forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
-	\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) 
+	%\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) \\ 
+	assocaiations(c) = \varnothing
 }{
 \begin{gathered}
-  \rho(c) \wedge \rho(singleProperties(c)) \wedge \rho(collectionProperties(c)) \\ \wedge \rho(associations(c))
+  tables(d) \cup \rho(c) \\ \wedge \rho(singleProperties(c)) \wedge \rho(collectionProperties(c)) \\ 
 \end{gathered}
@@
 \subsection{Software Decomposition}
-\subsection{Software Alternation}
-
-\subsubsection{Application Deconstruction}
 \begin{gather*}
-RemoveClass: Label \times Application \rightarrow Application  \\
-RemoveClass(q, a) = \inference{\exists c \in classes(a) : name(c) = q \\
+RemoveClass: Label \times Software \rightarrow Software  \\
+\Psi RemoveClass(l, s) = applications(s) \circ remClass(l) \\
+\Phi RemoveClass(l, s) = database(s) \bullet dropClass(l) \\
+\\ l \in Label, s \in Software 
+\\ \\
+remClass(l, a) = \inference{\exists c \in classes(a) : name(c) = l \\
 \wedge !\: instantiated(c) = \varnothing \wedge !\: associated(c)} {classes(a) \setminus c } \\
-q \in Label, a \in Applications
 \end{gather*}
 \begin{gather*}
-RemoveProperty: Label \times Label \times Application \rightarrow Application \\
+dropClass(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \not \exists r \in data(d) : owningTable(r) = l \\  \not \exists j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) = l }{ tables(d) \setminus \rho(c)}
+\end{gather*}
+
+\begin{gather*}
+RemoveProperty: Label \times Label \times Software \rightarrow Software \\
+\Psi RemoveProperty(l, l_p, s) =  applications(s) \circ  remProp(l, l_p) \\
+\Phi RemoveProperty(l, l_p, s) =  database(s) \bullet dropProp(l, l_p) \\
 RemoveProperty(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{properties(c) \setminus p }\\
@@
 \begin{gather*}
-RemoveAssociation : Label \times Labels \times Application \rightarrow Application \\ 
-RemoveAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists s \in associations(c) : name(s) = l_a }{associations(c) \setminus s} \\
+dropProp(p, d) = \begin{cases}
+ \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \exists c \in columns(t) : name(c) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p)) }{columns(t) \setminus c
+} \\ \\
+  \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \not \exists c \in columns(t) : name(c) = name(p) \\ \wedge \exists u \in tables(d) : name(u) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p))}{ tables(d) \setminus u }
+ \end{cases}
+\end{gather*}
+
+\begin{gather*}
+RemoveAssociation : Label \times Labels \times Software \rightarrow Software \\ 
+\Psi RemoveAssociation(l, l_a, s) = application(s) \circ remAss(l, l_a) \\
+\Phi RemoveAssociation(l, l_a, s) = database(s) \bullet dropAssoc(l, l_a) \\
+RemoveAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) \setminus f} \\
 l_c, l_a \in Labels, a \in Applications
 \end{gather*}
+\begin{gather*}
+dropAssoc(l, l_a, d) =  \begin{cases}
+ \inference{\exists t \in tables(d) : name(t) = l
+ \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_a
+ \\ \wedge \forall r \in data(d) : owningTable(r) \neq t}{ foreignKeys(t) \setminus f }
+ \\ \\
+ \inference{\exists t \in tables(d) : name(t) = l_a
+ \\ \wedge \exists f \in foreignKeys(t) : name(f) = l
+ \\ \wedge \forall r \in data(d) : reference(r) \neq t}{ tables(d) \setminus t}
+ \end{cases}
+\end{gather*}
 
-\subsubsection{Application Alternation}
+\subsection{Software Alternation}
 \paragraph{Rename Class}
@@
 
-\subsection{Database Transformation}
-
-\subsubsection{Deconstruction of a Software}
-\paragraph{Remove Class}
-$$
-\Phi_{RC}(c,d) = \inference{ \rho(c) \in tables(d) \\ \wedge \not \exists r \in data(d) : owningTable(r) = \rho(c) \\ \not \exists j \in tables(d), f \in foreignKeys(j) : reference(f) = \rho(c)  }{ tables(d) \setminus \rho(c)}
-$$
-%TODO remove by label
-
-\paragraph{Remove Property}
-$$
-\Phi_{RP}(p, d) = \begin{cases}
- \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \exists c \in columns(t) : name(c) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p)) }{columns(t) \setminus c
-} \\ \\
-  \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \not \exists c \in columns(t) : name(c) = name(p) \\ \wedge \exists u \in tables(d) : name(u) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p))}{ tables(d) \setminus u }
- \end{cases}
-$$
-
-\paragraph{Remove Association}
-$$
-\Phi_{RA}(a, d) =  \begin{cases}
- \inference{\exists t \in tables(d) : name(t) = name(owningClass(a))
- \\ \wedge \exists f \in foreignKeys(t) : name(f) = name(a)
- \\ \wedge \forall r \in data(d) : owningTable(r) \neq t}{ foreignKeys(t) \setminus f }
- \\ \\
- \inference{\exists t \in tables(d) : name(t) = name(a)
- \\ \wedge \exists f \in foreignKeys(t) : name(f) = name(owningClass(a))
- \\ \wedge \forall r \in data(d) : owningTable(r) \neq t}{ tables(d) \setminus t}
- \end{cases}
-$$
-
 \subsubsection{Altering a Software}	
-\paragraph{Copy Column}
-$$\Phi_{copyC} : Label \times Label \times Label \times f_m \times Database \rightarrow Database $$
-
-\begin{equation*}
-	\begin{gathered}
-		\Phi_{copyC}(l_c, l_s, l_t, f_m, d) = \begin{cases}
- 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{columns(e) \cup c \wedge \forall r \in rows(t) : f_m(r) \cup getPair(l_c,r)}
-	\end{cases}
-	\end{gathered}
-\end{equation*}
-
-\subsection{Composing of Evolution Operations}
 \paragraph{Copy Property}
@@
 With a copy property transformation defined we are able to define advanced evolution scenarios.
+
+\paragraph{Copy Column}
+$$\Phi_{copyC} : Label \times Label \times Label \times f_m \times Database \rightarrow Database $$
+
+\begin{equation*}
+	\begin{gathered}
+		\Phi_{copyC}(l_c, l_s, l_t, f_m, d) = \begin{cases}
+ 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{columns(e) \cup c \wedge \forall r \in rows(t) : f_m(r) \cup getPair(l_c,r)}
+	\end{cases}
+	\end{gathered}
+\end{equation*}
+
+\subsection{Composing of Evolution Operations}
 

### 2013-01-03 doc/description.tex
finished basic operations

@@
 \begin{gather*}
-RemoveClass: Label \times Software \rightarrow Software  \\
+RemoveClass: Label \times Software \rightarrow State  \\
 \Psi RemoveClass(l, s) = applications(s) \circ remClass(l) \\
@@
 \begin{gather*}
-RemoveProperty: Label \times Label \times Software \rightarrow Software \\
+RemoveProperty: Label \times Label \times State \rightarrow State \\
 \Psi RemoveProperty(l, l_p, s) =  applications(s) \circ  remProp(l, l_p) \\
@@
 \begin{gather*}
-RemoveAssociation : Label \times Labels \times Software \rightarrow Software \\ 
+RemoveAssociation : Label \times Labels \times State \rightarrow State \\ 
 \Psi RemoveAssociation(l, l_a, s) = application(s) \circ remAss(l, l_a) \\
@@
 \begin{gather*}
-\rho_{RenC} : Label \times Label \times Application \rightarrow Application \\ 
-\rho_{RenC}(l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \forall e \in classes(a) : name(e) \neq l_n}{\begin{gathered}
-c.name = l_n \wedge \forall f \in associations(e), \\ e \in classes(a), reference(f) = l_c : f.reference = l_n 
+RenameClass : Label \times Label \times State \rightarrow State \\ 
+\Psi(RenameClass(l_o, l_n), s) = application(s) \circ renClass(l_o, l_n) \\
+\Phi(RenameClass(l_o, l_n), s) = database(s) \bullet alterClass(l_o, l_n) \\ \\
+renClass:  Label \times Label \times Application \rightarrow Application \\
+renClass(l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_o \\ \wedge \forall e \in classes(a) : name(e) \neq l_n}{\begin{gathered}
+c.name = l_n \wedge \forall e \in classes(a) \\ \wedge \forall (f \in associations(e) \wedge reference(f) = l_o ) \\ : f.reference = l_n 
+\end{gathered}}\\ \\
+alterTable:  Label \times Label \times Database \rightarrow Database \\
+alterTable(l_o, l_n, d) = \inference{\exists t \in tables(d) : name(d) = l_o \\ \wedge \forall e \in tables(d) : name(e) \neq l_n}{\begin{gathered}
+t.name = l_n \\ \wedge \forall e \in tables(a) \wedge \forall (f \in fks(e) \wedge reference(f) = l_o ) \\ : f.reference = l_n 
 \end{gathered}}\\
-l_o, l_n \in Label, a \in Applications
+l_o, l_n \in Label, s \in States, a \in Applications, d \in Databases
 \end{gather*}
@@
 \begin{gather*}
-\rho_{RenP} : Label \times Label \times Label \times Application \rightarrow Application \\ \\
-	\rho_{RenP}(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_o \\ \wedge \forall e \in properties(c) : name(e) \neq l_n}{\begin{gathered}
+RenameProperty : Label \times Label \times Label \times State \rightarrow State \\ \\
+\Psi(RenameProperty(l_c, l_o, l_n, s)) = application(s) \circ renProp(l_c, l_o, l_n) \\
+\Phi(RenameProperty(l_c, l_o, l_n, s)) = database(s) \bullet alterColumn(l_c, l_o, l_n) \\\\
+renProp : Label \times Label \times Label \times Application \rightarrow Application \\
+	renProp(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_o \\ \wedge \forall e \in properties(c) : name(e) \neq l_n}{\begin{gathered}
 p.name = l_n 
-\end{gathered}}\\
-l_c, l_o, l_n \in Label, a \in Applications
+\end{gathered}}\\\\
+alterColumn : Label \times Label \times Label \times Database \rightarrow Database \\
+	alterColumn(l_c, l_o, l_n, d) = \begin{cases}
+ \inference{\exists t \in tables(a) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_o \\ \wedge \forall e \in columns(t) : name(e) \neq l_n}{\begin{gathered}
+c.name = l_n 
+\end{gathered}}
+\\ \\
+\inference{\exists t \in tables(a) : name(t) = l_o \\ \wedge \forall e \in tables(t) : name(e) \neq l_n}{\begin{gathered}
+t.name = l_n 
+\end{gathered}}
+ \end{cases}
+ \\
+l_c, l_o, l_n \in Label, a \in Applications, s \in States, d \in Databases
 \end{gather*}
@@
 \begin{gather*}
-\rho_{RenA} : Label \times Label \times Label \times Application \rightarrow Application \\
-	\rho_{RenP}(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists a \in associations(c) : name(a) = l_o \\ \wedge \forall e \in associations(c) : name(e) \neq l_n}{\begin{gathered}
-a.name = l_n 
-\end{gathered}}\\
-l_c, l_o, l_n \in Label, a \in Applications
+RenameAssociation : Label \times Label \times Label \times State \rightarrow State \\
+\Psi(RenameAssocaiation(l_c, l_o, l_n), s) = application(s) \circ renAssoc(l_c, l_o, l_n)\\
+\Phi(RenameAssocaiation(l_c, l_o, l_n), s) = database(s) \bullet alterAsspoc(l_c, l_o, l_n)\\\\
+renAssoc(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists b \in associations(c) : name(b) = l_o \\ \wedge \forall e \in associations(c) : name(e) \neq l_n}{\begin{gathered}
+b.name = l_n 
+\end{gathered}}\\\\
+alterColumn : Label \times Label \times Label \times Database \rightarrow Database \\
+	alterAssoc(l_c, l_o, l_n, d) = \begin{cases}
+?? \inference{\exists t \in tables(a) : name(t) = l_c \\ \wedge \exists f \in fks(t) : name(c) = l_o \\ \wedge \forall e \in fks(t) : name(e) \neq l_n}{\begin{gathered}
+f.name = l_n 
+\end{gathered}}
+\\ \\
+\inference{\exists t \in tables(a) : name(t) = l_o \\ \wedge \forall e \in tables(t) : name(e) \neq l_n}{\begin{gathered}
+t.name = l_n 
+\end{gathered}}
+ \end{cases}
+ \\
+l_c, l_o, l_n \in Label, a \in Applications, s \in Sates, d \in Databases
 \end{gather*}

### 2013-01-03 doc/description.tex
added manipulators for database structure

@@
 
+\subsection{Transformations for Manipulating a Database}
+The database is composed from two separated parts - database schema and data. Several transformation are defined to make manipulating a database easily in this paper.
+
+\begin{gather*}
+addColumn : Label \times Column \times Database \rightarrow Database \\ addColum(l, c, d) = \begin{cases}
+columns(t) \cup c \; \mathbf{if} t \in tables(d) \wedge name(t) = l \\ \;\;\; \wedge \forall f \in columns(t) : name(f) \neq name(c)
+\\ 
+Inconsistent \;d 
+ \end{cases}
+\\ l \in Label, c \in Columns, d \in Databases
+\end{gather*}
+
+\begin{gather*}
+addTable : Table \times Database \rightarrow Database \\ 
+addTable(t, d) = \begin{cases}
+tables(d) \cup c \; \mathbf{if} \forall e \in tables(d) \wedge name(t) \neq name(e)
+\\ 
+Inconsistent \;d 
+ \end{cases}
+\\ t \in Tables,  d \in Databases
+\end{gather*}
+
 \subsection{Extension of Software Definition}
@@
 $\rho_{c}: Class \times Database \rightarrow Database $ \\ 
-$\rho_{c}(c,d) = tables(d) \cup table(name(c), primaryKey("Id"), \varnothing, \varnothing) $
+$\rho_{c}(c,d) = table(name(c), primaryKey("Id"), \varnothing, \varnothing) $
 \end{center}
@@
 insertProp(p, d) = \begin{cases}
-\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge (!\;mandatory(p) \vee !\;instantiated(owningClass(p))}{ columns(t) \cup \rho(p,d)} 
+\inference{!\:mandatory(p) \vee !\:instantiated(owningClass(p))}{addColumn(l, \rho(p,d), d)} 
 \\ \\ 
-\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered} columns(t) \cup 
- \rho(p,d) \\ \wedge \forall r \in data(d), reference(r) = name(t) : pairs(r) \\ \cup [name(p), defaultValue(p)]
+\inference{ \wedge mandatory(p) \vee instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered} addColumn(l, \rho(p,d), d) \\ \bullet \forall r \in data(d), reference(r) = name(t) : pairs(r) \\ \cup [name(p), defaultValue(p)]
 \end{gathered}
@@
 \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
-tables(d) \cup  \rho(p,d) \wedge \forall r \in row(owningClass(p)) : data(d)\\ \cup row(name(p), ["value", defaultValue(p)], \\ [name(p), value(primaryKey(r))]) 
+addTable(\rho(p,d), d) \bullet \forall r \in row(owningClass(p)) : data(d)\\ \cup row(name(p), ["value", defaultValue(p)], \\ [name(p), value(primaryKey(r))]) 
 \end{gathered}
@@
 \inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
-columns(t_s) \cup \rho(a,d) \\ \wedge \forall r \in f_m(owningClass(a), reference(a)) : \\ data(t_t) \cup row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]) 
+addColumn(name(t_s), \rho(a,d), d) \\ \bullet \forall r \in f_m(owningClass(a), reference(a)) : \\ data(t_t) \cup row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]) 
 \end{gathered}
@@
  \\ \\
-\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) > 1}{\begin{gathered} tables(d) \cup
-\rho(a,d) \\ \wedge \forall r \in f_m(owningClass(a), reference(a)),\\ t \in tables(d) : name(t) = name(a) : \\ data(t) \cup row(name(a), [name(t_s), \\ value(primaryKey(first(r)))] [name(t_t), \\ value(primaryKey(second(r)))])
+\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) > 1}{\begin{gathered} addTable(\rho(a,d),d) \\ \bullet \forall r \in f_m(owningClass(a), reference(a)),\\ t \in tables(d) : name(t) = name(a) : \\ data(t) \cup row(name(a), [name(t_s), \\ value(primaryKey(first(r)))] [name(t_t), \\ value(primaryKey(second(r)))])
 \end{gathered}
@@
 \begin{gathered}
-  tables(d) \cup \rho(c) \\ \wedge \rho(singleProperties(c)) \wedge \rho(collectionProperties(c)) \\ 
+ addTable(\rho(c, d),d) \\ \wedge t \rho(singleProperties(c)) \wedge \rho(collectionProperties(c)) \\ 
 \end{gathered}
@@
 
-\subsubsection{Altering a Software}	
 \paragraph{Copy Property}
-All transformations described earlier does change only the structure of a application.
-These operation can be used in case there are no instances in the application.
-If instances are present in the application, the decomposition transformations could cause a data lost.
+All transformations described earlier does change mostly the structure of a application only.
 Therefore we define  the transformation for copying a property from one class to another.
@@
 
-The copy property can be implemented as a transformation composed by operations for adding and removing property, nevertheless this composition would not work with instances, thus we define the copy property operation as a basic and atomic operation which will help us to handle instances.
+The copy property can be implemented as a transformation composed by operations for adding and removing property, nevertheless this composition would not work with instances, where the $copyData$ transformation is used.
 

### 2013-01-05 doc/description.tex
more partial transformations

@@
 [Label \times Value] \rightarrow Pair\\ \\
+\mathbf{pairByRef} : Label \times Row \rightarrow Pair \\
+pairByRef(l,i) = \begin{cases}
+	\inference{\exists p \in pairs(i) : reference(p) = l}{p} \\ \\
+	\varnothing
+ \end{cases}\\\\
 \mathbf{reference} : Pair \rightarrow Label \\
@@
 \begin{gather*}
-addColumn : Label \times Column \times Database \rightarrow Database \\ addColum(l, c, d) = \begin{cases}
-columns(t) \cup c \; \mathbf{if} t \in tables(d) \wedge name(t) = l \\ \;\;\; \wedge \forall f \in columns(t) : name(f) \neq name(c)
-\\ 
+addColumn : Label \times Column* \times Database \rightarrow Database \\ addColum(l, c, d) = \begin{cases}
+\inference{|c| > 1}{addColumn(l, head(c), d) \bullet addColumn(l, tail(c), d)}
+\\ \\
+\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c)}{columns(t) \cup c}
+\\ \\
 Inconsistent \;d 
  \end{cases}
-\\ l \in Label, c \in Columns, d \in Databases
+\\ l \in Label, c = Column*, d = Database
 \end{gather*}
@@
 \begin{gather*}
-addTable : Table \times Database \rightarrow Database \\ 
+	copyColumn : Label \times Label \times Label \times Database \rightarrow Database \\
+	copyColumn(l_s, l_t, l_c, d) = \begin{cases}
+ 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{columns(e) \cup c }
+	\end{cases} \\
+	l_s,l_t,l_c \in Label, d = Database
+\end{gather*}
+
+\begin{gather*}
+addTable : Table* \times Database \rightarrow Database \\ 
 addTable(t, d) = \begin{cases}
-tables(d) \cup c \; \mathbf{if} \forall e \in tables(d) \wedge name(t) \neq name(e)
-\\ 
+	\inference{|t| > 1}{addTable(head(t),d) \bullet addTable(tail(t), d)}
+\\ \\	
+	\inference{|t| = 1 \wedge \forall e \in tables(d) \wedge name(t) \neq name(e)}{tables(d) \cup c}
+\\ \\
 Inconsistent \;d 
  \end{cases}
-\\ t \in Tables,  d \in Databases
+\\ t = Table*,  d = Database
+\end{gather*}
+
+\begin{gather*}
+insertRow : Row* \times Database \rightarrow Database \\
+insertRow(r, d) = \begin{cases}
+	\inference{|r|>1}{insertRow(head(r), d) \bullet insertRow(tail(r), d)} 
+\\ \\
+ 	\inference{|r|=1 \wedge \exists t \in tables(d) : name(t) = reference(r) }{data(d) \cup r}
+\\ \\
+ Inconsistent \; d
+ \end{cases}
+\\ r = Row*, d = Database
+\end{gather*}
+
+\begin{gather*}
+appendValues : Label \times Pair* \times Database \rightarrow Daatabase \\
+appendValues(l, p, d) = \begin{cases}
+ 	\inference{|p| > 1}{appendValues(l, head(p), d)\bullet appendValues(l, tail(p), d)} 
+ \\ \\
+ 	\inference{|p| = 1 \wedge \exists r \in data(d), reference(r) = l}{pairs(r) \cup [name(p), defaultValue(p)]}
+ \\ \\
+ 	 Inconsistent \; d
+ \end{cases} 
+ \\
+l \in Label, p = Pair*, d = Databases
+\end{gather*}
+
+\begin{gather*}
+insertPair : Row \times Pair \times Database \rightarrow Database \\
+insertPair(r_t, p, d) = \begin{cases}
+ \inference{\forall e \in pairs(r_t) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r_t), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{ pairs(r_t) \cup p}
+ \\\\
+ Inconsistent \; d
+ \end{cases}
+\end{gather*}
+
+\begin{gather*}
+copyValue : Label \times Label \times Label \times (Instancece \times Instance) \times Database \times Database \\
+copyValue(l_s, l_t, l_{sc}, l_{tc} f_m, d) = \begin{cases}
+ 	\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc} }{\begin{gathered}
+\forall p \in f_m(t_s, t_t)  : \\ insertPair(second(p), [l_{tc},value(pairByRef(l_{sc}, first(p))),d)
+\end{gathered}}
+ \end{cases}
 \end{gather*}
@@
 
- \inference{ cardinality(p) > 1 \wedge !\;mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}tables(d) \cup  table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), \varnothing), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL) \\
+ \inference{ cardinality(p) > 1 \wedge !\;mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), \varnothing), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL) \\
 \end{gathered}}
@@
 \\ \\ 
-\inference{ \wedge mandatory(p) \vee instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered} addColumn(l, \rho(p,d), d) \\ \bullet \forall r \in data(d), reference(r) = name(t) : pairs(r) \\ \cup [name(p), defaultValue(p)]
+\inference{ \wedge mandatory(p) \vee instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered} addColumn(l, \rho(p,d), d) \\ \bullet \forall r \in data(d), reference(r) = name(t) : \\ appendValue(name(t), [name(p), defaultValue(p)], d)
 \end{gathered}
@@
 \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
-addTable(\rho(p,d), d) \bullet \forall r \in row(owningClass(p)) : data(d)\\ \cup row(name(p), ["value", defaultValue(p)], \\ [name(p), value(primaryKey(r))]) 
+addTable(\rho(p,d), d) \bullet \forall r \in row(owningClass(p)) :\\ insertRow(row(name(p), ["value", defaultValue(p)], \\ [name(p), value(primaryKey(r))]),d) 
 \end{gathered}
@@
 \inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
-addColumn(name(t_s), \rho(a,d), d) \\ \bullet \forall r \in f_m(owningClass(a), reference(a)) : \\ data(t_t) \cup row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]) 
+addForeignKey(name(t_s), \rho(a,d), d) \\ \bullet \forall r \in f_m(owningClass(a), reference(a)) : \\ insertRow(row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]), d)
 \end{gathered}
@@
  \\ \\
-\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) > 1}{\begin{gathered} addTable(\rho(a,d),d) \\ \bullet \forall r \in f_m(owningClass(a), reference(a)),\\ t \in tables(d) : name(t) = name(a) : \\ data(t) \cup row(name(a), [name(t_s), \\ value(primaryKey(first(r)))] [name(t_t), \\ value(primaryKey(second(r)))])
+\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) > 1}{\begin{gathered} addTable(\rho(a,d),d) \\ \bullet \forall r \in f_m(owningClass(a), reference(a)),\\ t \in tables(d) : name(t) = name(a) : \\ insertRow(row(name(a), [name(t_s), \\ value(primaryKey(first(r)))] [name(t_t), \\ value(primaryKey(second(r)))]), d)
 \end{gathered}
@@
 \Psi(s, newClass(c)) = application(s) \circ addClass(c) \\
-\Phi(s, newClass(c)) = database(s) \bullet insertClass(c) 
+\Phi(s, newClass(c)) = database(s)  \bullet addTable(\rho(c,d), d) \\ \bullet addColumn(name(c), \{\forall r \in singleProperties(c) | \rho(r,d)\}, d) \\ \bullet addTable(name(c), \{\forall r \in collectionProperties(c) | \rho(r,d)\}, d)
 \end{gather*}
@@
 The situation reflect the modeling praxis when the structure of a class is defined and then it is associations are created.
-\begin{gather*}
-insertClass : Class \times Database \rightarrow Database \\
-insertClass(c, d) = \inference{ \forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
-	%\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) \\ 
-	assocaiations(c) = \varnothing
-}{
-\begin{gathered}
- addTable(\rho(c, d),d) \\ \wedge t \rho(singleProperties(c)) \wedge \rho(collectionProperties(c)) \\ 
-\end{gathered}
-}
-\end{gather*}
+%\begin{gather*}
+%insertClass : Class \times Database \rightarrow Database \\
+%insertClass(c, d) = \inference{ \forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
+%	%\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) \\ 
+%	assocaiations(c) = \varnothing
+%}{
+%\begin{gathered}
+% addTable(\rho(c, d),d) \\ \wedge t \rho(singleProperties(c)) \wedge \rho(collectionProperties(c)) \\ 
+%\end{gathered}
+%}
+%\end{gather*}
 
@@
 Therefore we define  the transformation for copying a property from one class to another.
-The copy property is the basic way to copy data from instance of a class to instance of another class.
+To copy property is the basic way to move data from instances of a class to instance of another class.
+Note the transformation does not move a single value, but it move the values for all instances in the software.
 
 The copy property can be implemented as a transformation composed by operations for adding and removing property, nevertheless this composition would not work with instances, where the $copyData$ transformation is used.
-
-$$\rho_{copyP} : Label \times Label \times Label \times (Instance \times Instance) \times Application \rightarrow Application $$
-\begin{equation*}
-	\rho_{copy}(l_s, l_t, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p \\ \wedge \exists e \in classes(a) : name(e) = l_t  \\ \wedge \forall q \in properties(c) : name(q) \neq l_p}{\begin{gathered}
-properites(e) \cup p 
-\end{gathered}
-}
-\end{equation*}
+\begin{gather*}
+	CopyProperty : Label \times Label \times Label \times (Instance \times Instance) \times State \rightarrow State 
+\\
+\Psi(CopyProperty(l_s, l_t, l_p, f_m), s) = application(s) \circ copyProp(l_s,l_t, l_p) \\
+\Phi(CopyProperty(l_s, l_t, l_p, f_m), s) = database(s) \bullet copyColumn(l_s, l_t, l_p)\\  \bullet copyValues(l_s, l_t, l_p, f_m) \\ 
+\end{gather*}
 
 With a copy property transformation defined we are able to define advanced evolution scenarios.
-
-\paragraph{Copy Column}
-$$\Phi_{copyC} : Label \times Label \times Label \times f_m \times Database \rightarrow Database $$
-
-\begin{equation*}
-	\begin{gathered}
-		\Phi_{copyC}(l_c, l_s, l_t, f_m, d) = \begin{cases}
- 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{columns(e) \cup c \wedge \forall r \in rows(t) : f_m(r) \cup getPair(l_c,r)}
-	\end{cases}
-	\end{gathered}
-\end{equation*}
 
@@
 
-\section{Summary}
+\section{Conclusion}
 
 \end{document}
+
+%	CopyProperty(l_s, l_t, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p \\ \wedge \exists e \in classes(a) : name(e) = l_t  \\ \wedge \forall q \in properties(c) : name(q) \neq l_p}{\begin{gathered}
+%properites(e) \cup p 
+%\end{gathered}}
+
 % This is important in case we have an evolution consisting of several evolution operations.

### 2013-01-07 doc/description.tex
formating and minor fixes

@@
 There are three important components of such a software: application (its persistent layer concretely), database (consisting of database schema and stored data) and an ORM (we will use $\rho$ as a symbol for ORM), therefore we define a software a triple:
-\begin{gather*}
+\begin{align*}
 	software = < Application, Database, \rho > \\
@@
 	a \in Applications, d \in Databases
-\end{gather*}
+\end{align*}
 The software has to be in consistent state so it is beneficial for its user. The state of the software is defined to communicate the sate of the software to its user:
-\begin{gather*}
+\begin{align*}
 	state = Consistent\; s \: | \: Inconsistent \; s, \\
@@
  	\end{cases}
-\end{gather*}
+\end{align*}
 \subsection{Software Evolution}
@@
 We call this mappings $\Psi$ and $\Phi$:
-\begin{gather*}
+\begin{align*}
 \Psi : E \rightarrow AppOps \\
 \Phi : E \rightarrow DbOps
-\end{gather*}
+\end{align*}
 A subset of $E$ represents an evolution transformation of a software, which mets the first goal of our framework - a single source of data evolution for the whole software.
@@
 and similarly the $\circ$ and the $\bullet$ operators for evolving application and database:
-\begin{gather*}
+\begin{align*}
 \circ : Application \times AppOps \rightarrow Application \\
  \bullet : Database \times DbOps \rightarrow Database
-\end{gather*}
+\end{align*}
 The definition of the evolving operator is inspired by the monads (as used e.g.
@@
 
-\begin{gather*}
+\begin{align*}
  u \oplus t = \begin{cases}
@@
 u \in States, t \in E, s \in Softwares
-\end{gather*}
-\begin{gather*}
+\end{align*}
+\begin{align*}
  a \circ o = \begin{cases}
@@
 a \in Applications, o \in AppOps
-\end{gather*}
-\begin{gather*}
+\end{align*}
+\begin{align*}
  d \bullet o = \begin{cases}
@@
 d \in Applications, o \in DbOps
-\end{gather*}
+\end{align*}
 
@@
 \label{sec:appModel}
-The model represents a persistent layer of an application created using object oriented language.
-It means the layer is a set of instances which cooperates together.
+The application model represents a persistent layer of an application created using object oriented language and transformations which can alter the model. 
+\subsubsection{Application Structure}
 Each instance is defined according to a class - the class defines the structure of its instances and relationships between them an the rest of the application.
@@
 
-\begin{gather*}
-  Application = < Class* > \\ 
-  app : Class* \rightarrow  Application \\ \\
-  classes : Application \rightarrow Class* \\
+\begin{align*}
+  \mathbf{Application} = < Class* > \\ 
+  \mathbf{app} : Class* \rightarrow  Application \\ \\
+  \mathbf{classes} : Application \rightarrow Class* \\
   classes(app(cs)) = cs  \\\\
-  class : Label \times Application \rightarrow Class   \\ 
+  \mathbf{class} : Label \times Application \rightarrow Class   \\ 
   class(l, a) = c, c \in classes(a) \wedge name(c) = l \\ \\
@@
   \\ \\
-  associated :  Label \times Application \rightarrow Boolean \\
- \end{gather*}
+  \mathbf{associated} :  Label \times Application \rightarrow Boolean \\
+\end{align*}
 
 \paragraph{Class} represents a basic structural concept in the application model. It has a unique name, one or more properties and a class can be associated to other classes in the application. 
-
-$$Class = <label, Property*, Association*>$$
-\begin{equation*}
-\begin{alignat}{2}
-cl : Label \times Property* \times Association* \rightarrow Class &\quad&\text{constructor}\\ \\
+\begin{align*}
+	\mathbf{Class} = <label, Property*, Association*> \\
+	\mathbf{cl} : Label \times Property* \times Association* \rightarrow Class \\\\
 % TODO association has to be only to self (all associated has to exists)
-name : Class \rightarrow Label &\quad&\text{name of the class}\\
+	\mathbf{name} : Class \rightarrow Label \\
 	name(cl(l, props, assocs)) = l \\ \\
-properties : Class \rightarrow Property* &\quad&\text{all properties}\\
+	\mathbf{properties} : Class \rightarrow Property* \\
 	properties(cl(l, props, assocs)) = props \\ \\
-primitiveProperties : Class \rightarrow Property* &\quad&\text{all primitive properties}\\
-primitiveProperties(cl(l, props, assocs)) = props', \\ props' \subset props,  \forall p \in props' : cardinality(p) \leq 1 \\ \\
-collectionProperties : Class \rightarrow Property* &\quad&\text{all collections}\\
-collectionProperties(cl(l, props, assocs)) = props', \\ props' \subset props, \forall p \in props' : cardinality(p) > 1 \\ \\
-associations : Class \rightarrow Association* &\quad&\text{all associations}\\
-associations(cl(l, props, assocs)) = assocs
-\end{alignat}
-\end{equation*}
-$$l \in Label, props \in Properties, assocs \in Associations$$
+	\mathbf{primitiveProperties} : Class \rightarrow Property* \\	primitiveProperties(cl(l, props, assocs)) = props', \\ props' \subset props,  \forall p \in props' : cardinality(p) \leq 1 \\ \\
+	\mathbf{collectionProperties} : Class \rightarrow Property* \\	collectionProperties(cl(l, props, assocs)) = props', \\ props' \subset props, \forall p \in props' : cardinality(p) > 1 \\ \\
+	\mathbf{associations} : Class \rightarrow Association* \\	associations(cl(l, props, assocs)) = assocs \\ \\
+l \in Label, props \in Properties, assocs \in Associations
+\end{align*}
 
@@
 The property can be mandatory, can have a default value and can represent a single value or a collection of values.
-\begin{gather*}
-Property = <label, AppType, DefaultValue, Cardinality, Mandatory> \\ \\
-prop : Label \times AppType \times String \times \mathbb{N^{*}} \times Boolean \rightarrow Property \\ \\
-name : Property \rightarrow Label \\
+\begin{align*}
+	\mathbf{Property} = <label, AppType, DefaultValue, Cardinality, Mandatory> \\ \\
+	\mathbf{prop} : Label \times AppType \times String \times \mathbb{N^{*}} \times Boolean \rightarrow Property \\ \\
+	\mathbf{name} : Property \rightarrow Label \\
 	name(prop(l, t, val, n, m)) = l \\ \\
-type : Property \times AppType \\
+	\mathbf{type} : Property \times AppType \\
 	type(prop(l, t, val, n, m)) = t \\ \\
-cardinality : Property \rightarrow \mathbb{N^{*}} \\
+	\mathbf{cardinality} : Property \rightarrow \mathbb{N^{*}} \\
 	cardinality(prop(l, t, val, n, m)) = n \\ \\
-mandatory : Property \rightarrow Boolean \\
+	\mathbf{mandatory} : Property \rightarrow Boolean \\
 	mandatory(prop(l, t, val, n, m)) = m  \\ \\
 l \in Label, t \in AppType, val \in String*, n \in \mathbb{N^{*}}, m \in Boolean
-\end{gather*}
+\end{align*}
 
@@
 The cardinalities defines the multiplicity of the association ends.
-\begin{gather*}
-Association = <label, classRef, startCardinality, endCardinality> \\ \\
-assoc : Label \times Label \times \mathbb{N} \times \mathbb{N} \rightarrow Association
-name : Association \rightarrow Label \\
+\begin{align*}
+	\mathbf{Association} = <label, classRef, startCardinality, endCardinality> \\ \\
+	\mathbf{assoc} : Label \times Label \times \mathbb{N} \times \mathbb{N} \rightarrow Association\\
+	\mathbf{name} : Association \rightarrow Label \\
 	name(assoc(l, ref, n_1, n_2)) = l\\ \\
-reference : Association \rightarrow Label \\
+	\mathbf{reference} : Association \rightarrow Label \\
 	reference(assoc(l, ref, n_1, n_2)) = ref\\ \\
-startCardinality : Association \rightarrow \mathbb{N} \\
+	\mathbf{startCardinality} : Association \rightarrow \mathbb{N} \\
 	startCardinality(assoc(l, ref, n_1, n_2)) = n_1\\ \\
-endCardinality : Association \rightarrow \mathbb{N} \\
+	\mathbf{endCardinality} : Association \rightarrow \mathbb{N} \\
 	endCardinality(assoc(l, ref, n_1, n_2)) = n_2 \\ \\
 l, ref \in Label,  n_1, n_2 \in \mathbb{N}
-\end{gather*}
+\end{align*}
 
@@
 The only App-Type type is called APP-STRING.
-$$
-AppType = APPSTRING
-$$
+\begin{align*}
+\mathbf{AppType} = APPSTRING
+\end{align*}
+
+\subsubsection{Application Transformation}
 
@@
 The relational database consists of database schema which defines the structure of the database and data, which in our ORM system represents instances. The database is:
-\begin{gather*}
-Database = <table*, row*> \\ \\
-db : Tables \times Rows \rightarrow Database \\ \\
-tables : Database \rightarrow Tables \\
-tables(db(tabs, rows)) = tabs \\ \\
-data : Database \rightarrow Rows \\
-data(db(tabs, rows)) = rows \\ \\
-table : Label \times Database \rightarrow Table  \\
+\begin{align*}
+	\mathbf{Database} = <table*, row*> \\ \\
+	\mathbf{db} : Tables \times Rows \rightarrow Database \\ \\
+	\mathbf{tables} : Database \rightarrow Tables \\
+	tables(db(tabs, rs)) = tabs \\ \\
+	\mathbf{rows} : Database \rightarrow Rows \\
+	rows(db(tabs, rs)) = rs \\ \\
+	\mathbf{table} : Label \times Database \rightarrow Table  \\
 	table(l, d) = t, t \in tables(d) \wedge name(t) = l \\ \\
-tabs \in Tables, rows \in Rows, l \in Labels, d \in Databases
-\end{gather*}
+	tabs = Table*, rs = Row*, l \in Label, d = Database
+\end{align*}
 
@@
 Rows in the table represents stored data.
-\begin{gather*}
-Table = <label, primaryKey, Column*, ForeignKey*>\\ \\
-tab : Label \times PrimaryKeys \times Columns \times ForeignKeys \rightarrow Table \\ \\
-name : Table \rightarrow Label  \\
+\begin{align*}
+	\mathbf{Table} = <label, primaryKey, Column*, ForeignKey*>\\ \\
+	\mathbf{tab} 	: Label \times PrimaryKeys \times Columns \times ForeignKeys \rightarrow Table \\ \\
+	\mathbf{name} : Table \rightarrow Label  \\
 	name(tab(l, pk, cols, fks)) = l \\ \\
-primaryKey : Table \rightarrow PrimaryKeys  \\
+	\mathbf{primaryKey} : Table \rightarrow PrimaryKeys  \\
 	primaryKey(tab(l, pk, cols, fks)) = pk \\ \\
-columns : Table \rightarrow Columns  \\
+	\mathbf{columns} : Table \rightarrow Columns  \\
 	columns(tab(l, pk, cols, fks)) = cols \\ \\
-foreignKeys : Table \rightarrow ForeignKeys  \\
+	\mathbf{foreignKeys} : Table \rightarrow ForeignKeys  \\
 	foreignKeys(tab(l, pk, cols, fks)) = fks \\ \\
-l \in Label, pk \in PrimaryKey, cols \in Columns, fks \in ForeignKeys
-\end{gather*}
+l \in Label, pk = PrimaryKey, cols = Column*, fks = ForeignKey*
+\end{align*}
 
 \paragraph{Column} columns define possible data values and types which can be part of a table record.
-\begin{gather*}
-Column = <label, DbType, DefaultValue, Constraint*> \\ \\
-col : Label \times DbType \times String \times Constraints \rightarrow Column \\ \\
-name : Column \rightarrow Label  \\
+\begin{align*}
+	\mathbf{Column} = <label, DbType, DefaultValue, Constraint*> \\ \\
+	\mathbf{col} : Label \times DbType \times String \times Constraints \rightarrow Column \\ \\
+	\mathbf{name} : Column \rightarrow Label  \\
 	name(col(l, t, val, cons)) = l  \\ \\
-type : Column \rightarrow DbType  \\
+	\mathbf{type} : Column \rightarrow DbType  \\
 	type(col(l, t, val, cons)) = t  \\ \\
-constraints : Column \rightarrow Constraints  \\
+	\mathbf{constraints} : Column \rightarrow Constraints  \\
 	constraints(col(l, t, val, cons)) = cons  \\ \\
-l \in Labels, t \in DbType, val \in Strings, cons \in Constraints
-\end{gather*}
+l \in Label, t = DbType, val = Strings, cons = Constraints
+\end{align*}
 
 \paragraph{Foreign key}
-\begin{gather*}
-ForeignKey = <label, tableRef, Constraint*> \\ \\
-fk : Label \times Label \times Constraints \rightarrow ForeignKey \\ \\
-name : ForeignKey \rightarrow Label \\
+\begin{align*}
+	\mathbf{ForeignKey} = <label, tableRef, Constraint*> \\ \\
+	\mathbf{fk} : Label \times Label \times Constraints \rightarrow ForeignKey \\ \\
+	\mathbf{name} : ForeignKey \rightarrow Label \\
 	name(fk(l, tRef, cons)) = lab  \\ \\
-reference : ForeignKey \rightarrow Label  \\
+	\mathbf{reference} : ForeignKey \rightarrow Label  \\
 	reference(fk(l, tRef, cons)) = tRef  \\ \\
-reference : ForeignKey \rightarrow Constraint*  \\
-reference(fk(l, tRef, cons)) = cons  \\ \\
+	\mathbf{constaints} : ForeignKey \rightarrow Constraint*  \\
+	constaints(fk(l, tRef, cons)) = cons  \\ \\
 l, tRef \in Label, cons \in Constraints
-\end{gather*}
+\end{align*}
 
 \paragraph{Primary key}
-\begin{gather*}
-PrimaryKey =  < label > 	\\
-pk : Label \rightarrow PrimaryKey
-\end{gather*}
+\begin{align*}
+	\mathbf{PrimaryKey} =  < label > 	\\
+	\mathbf{pk} : Label \rightarrow PrimaryKey
+\end{align*}
 
@@
 There are only two types defined in the model $DBSTRING$ as a default type of values in columns and $DBINT$ as a default type for keys.
-$$
-DbType = DBSTRING \; | \; DBINT 
-$$
+\begin{align*}
+	\mathbf{DbType} = DBSTRING \; | \; DBINT 
+\end{align*}
 
@@
 Both constraints are column constraints - first constraint defines unique records in a column, second constraint defines non-empty columns.
-\begin{gather*}
-Constraint = NOTNULL \; | \; UNIQUE 
-\end{gather*}
+\begin{align*}
+	\mathbf{Constraint} = NOTNULL \; | \; UNIQUE 
+\end{align*}
 
@@
 A table row in our model is a tuple consisting of reference to a concrete table and a set of value pairs, which represents concrete values of concrete columns.
-\begin{gather*}
-row = < refTable, pair* > \\
-row : Label \times Pairs \rightarrow Row \\ \\
-refTable : Row \rightarrow Label \\
+\begin{align*}
+	\mathbf{Row} = < refTable, pair* > \\
+	\mathbf{row} : Label \times Pairs \rightarrow Row \\ \\
+	\mathbf{refTable} : Row \rightarrow Label \\
 	refTable(row(l, pairs) = l \\ \\
-contains : Table \times Database \rightarrow Boolean \\
+	\mathbf{contains} : Table \times Database \rightarrow Boolean \\
 	contains(t, d) = \begin{cases}
 			\inference{\exists r \in data(d) : refTable(r) = name(t) }{true} \\ \\
-  \inference{otherwise}{false}
+		 	false
  		\end{cases} \\ \\
- l \in Labels, pairs \in Pairs, t \in tables(d), d \in Database
-\end{gather*}
+ l \in Label, ps = Pair*, t \in tables(d), d \in Database
+\end{align*}
 There functions $contains$ verify if the table contains data in context of concrete database.
 
-\begin{gather*}
-pair = < refCol, value > \\
+\begin{align*}
+	\mathbf{Pair} = < refCol, value > \\
 	[Label \times Value] \rightarrow Pair\\ \\
+	\mathbf{reference} : Pair \rightarrow Label \\
+	reference([l,val]) = l \\ \\
+	\mathbf{value} : Pair \rightarrow Label \\ 
+	value([l,val]) = val \\ \\
 	\mathbf{pairByRef} : Label \times Row \rightarrow Pair \\
@@
 			\varnothing
- \end{cases}\\\\
-\mathbf{reference} : Pair \rightarrow Label \\
-reference([l,val]) = l \\ \\
-\mathbf{value} : Pair \rightarrow Label \\ 
-value([l,val]) = val \\
+		\end{cases}\\
 	l \in Label, val \in Value
-\end{gather*}
+\end{align*}
 The $refCol$ is a name of a column, a foreign key or its primary key and $value$ represents concrete value stored in the database.
@@
 The consistency of database is then:
-\begin{gather*}
+\begin{align*}
 	\mathbf{consistent} : Database \rightarrow Boolean \\
 	consistent(d) = \begin{cases}
- 	\inference{\forall r \in data(d), \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) :\\ \exists u \in columns(t) \cup foreignKeys(t) : reference(p) = name(u)}{true}
+ 		\inference{\forall r \in data(d), \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) :\\ \exists u \in columns(t) \cup foreignKeys(t) : \\ reference(p) = name(u)}{true}
  	\\ \\
- 	\inference{otherwise}{false}
+ 	false
  \end{cases}
-\end{gather*}
+\end{align*}
 
@@
 
-\begin{gather*}
-addColumn : Label \times Column* \times Database \rightarrow Database \\ addColum(l, c, d) = \begin{cases}
-\inference{|c| > 1}{addColumn(l, head(c), d) \bullet addColumn(l, tail(c), d)}
-\\ \\
-\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c)}{columns(t) \cup c}
-\\ \\
+\begin{align*}
+	\mathbf{addColumn} : Label \times Column* \times Database \rightarrow Database \\ 	
+	addColum(l, c, d) = \begin{cases}
+		\inference{|c| > 1}{addColumn(l, head(c), d) \bullet addColumn(l, tail(c), d)} \\ \\
+		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c)}{columns(t) \cup c}\\ \\
 		Inconsistent \;d 
- \end{cases}
-\\ l \in Label, c = Column*, d = Database
-\end{gather*}
+	 \end{cases}\\ 
+	 l \in Label, c = Column*, d = Database
+\end{align*}
 
-\begin{gather*}
-	copyColumn : Label \times Label \times Label \times Database \rightarrow Database \\
+\begin{align*}
+	\mathbf{copyColumn} : Label \times Label \times Label \times Database \rightarrow Database \\
 	copyColumn(l_s, l_t, l_c, d) = \begin{cases}
@@
 	l_s,l_t,l_c \in Label, d = Database
-\end{gather*}
+\end{align*}
 
-\begin{gather*}
-addTable : Table* \times Database \rightarrow Database \\ 
+\begin{align*}
+	\mathbf{addTable} : Table* \times Database \rightarrow Database \\ 
 	addTable(t, d) = \begin{cases}
@@
 		Inconsistent \;d 
- \end{cases}
-\\ t = Table*,  d = Database
-\end{gather*}
+	 \end{cases}\\ 
+	t = Table*,  d = Database
+\end{align*}
 
-\begin{gather*}
-insertRow : Row* \times Database \rightarrow Database \\
+\begin{align*}
+	\mathbf{insertRow} : Row* \times Database \rightarrow Database \\
 	insertRow(r, d) = \begin{cases}
-	\inference{|r|>1}{insertRow(head(r), d) \bullet insertRow(tail(r), d)} 
-\\ \\
- 	\inference{|r|=1 \wedge \exists t \in tables(d) : name(t) = reference(r) }{data(d) \cup r}
-\\ \\
+			\inference{|r|>1}{insertRow(head(r), d) \bullet insertRow(tail(r), d)} \\ \\
+		 	\inference{|r|=1 \wedge \exists t \in tables(d) : name(t) = reference(r) }{data(d) \cup r} \\ \\
 			 Inconsistent \; d
- \end{cases}
-\\ r = Row*, d = Database
-\end{gather*}
+		 \end{cases}	\\
+		 r = Row*, d = Database
+\end{align*}
 
-\begin{gather*}
-appendValues : Label \times Pair* \times Database \rightarrow Daatabase \\
+\begin{align*}
+	\mathbf{appendValues} : Label \times Pair* \times Database \rightarrow Daatabase \\
 	appendValues(l, p, d) = \begin{cases}
- 	\inference{|p| > 1}{appendValues(l, head(p), d)\bullet appendValues(l, tail(p), d)} 
- \\ \\
- 	\inference{|p| = 1 \wedge \exists r \in data(d), reference(r) = l}{pairs(r) \cup [name(p), defaultValue(p)]}
- \\ \\
+ 		\inference{|p| > 1}{appendValues(l, head(p), d)\bullet appendValues(l, tail(p), d)}  \\ \\
+ 		\inference{|p| = 1 \wedge \forall e \in pairs(r_t) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r_t), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p} \\ \\
 	 	 Inconsistent \; d
- \end{cases} 
- \\
+ 		\end{cases} \\
 l \in Label, p = Pair*, d = Databases
-\end{gather*}
+\end{align*}
 
-\begin{gather*}
-insertPair : Row \times Pair \times Database \rightarrow Database \\
-insertPair(r_t, p, d) = \begin{cases}
- \inference{\forall e \in pairs(r_t) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r_t), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{ pairs(r_t) \cup p}
- \\\\
- Inconsistent \; d
- \end{cases}
-\end{gather*}
-
-\begin{gather*}
-copyValue : Label \times Label \times Label \times (Instancece \times Instance) \times Database \times Database \\
+\begin{align*}
+	\mathbf{copyValue} : Label \times Label \times Label \times (Instancece \times Instance) \times Database \times Database \\
 	copyValue(l_s, l_t, l_{sc}, l_{tc} f_m, d) = \begin{cases}
  		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc} }{\begin{gathered}
-\forall p \in f_m(t_s, t_t)  : \\ insertPair(second(p), [l_{tc},value(pairByRef(l_{sc}, first(p))),d)
+			\forall p \in f_m(t_s, t_t)  : \\ appendValues(name(second(p)), \\ [l_{tc},value(pairByRef(l_{sc}, first(p))),d)
 		\end{gathered}}
  	\end{cases}
-\end{gather*}
+\end{align*}
 
@@
 
-\begin{gather*}
+\begin{align*}
  owningClass : Property \times Application \rightarrow Class  \\ owningClass(p, a) = c, c \in classes(a) \wedge p \in properties(c) \\
@@
 owningTable(fk, d) = tab, tab \in tables(d) \wedge fk \in foreignKeys(d)\\ \\ 
-\end{gather*}
+\end{align*}
 
@@
 We decide to use following transcription to cover all aspects of transformation definition:
-\begin{gather*}
+\begin{align*}
 operation(parameters, context) = \begin{cases}
@@
  \end{cases}
-\end{gather*}
+\end{align*}
 The second branch of the operation definition is an implicit assumption - if an operation is not feasible in a software, its state become inconsistent.
@@
 The first of them has no impact on stored data, whereas the rest of them can in some cases.
-\begin{gather*}
+\begin{align*}
 \mathbf{addProperty} : Label \times Property \times State \rightarrow State \\
@@
 s \in Softwares, l \in Label, p \in Properties
-\end{gather*}
+\end{align*}
 Adding a property into an application structure has to verify a name collision only:
-\begin{gather*}
+\begin{align*}
 addProp : Label \times Property \times Application \rightarrow Application \\
@@
 l \in Label, p \in Properties, a \in Applications
-\end{gather*}
+\end{align*}
 Whereas the $insertProp$ on database level can change stored data too, because a mandatory property can be added into a class with already existing instances, then a default value has to be added to all existing instances.
-\begin{gather*}
+\begin{align*}
 insertProp(p, d) = \begin{cases}
@@
 p \in Properties, d \in Dataabses
-\end{gather*}
+\end{align*}
 
-\begin{gather*}
+\begin{align*}
 \mathbf{addAsociation} : Label \times Association \times (Row \times Row) \times State \rightarrow State \\
@@
 \Phi(s, addAsociation(l, a, f_m)) = database(s) \bullet insertAssoc(l, a, f_m)
-\end{gather*}
+\end{align*}
 Adding an association into an application structure has to verify if there is no name collision and if there is the associated class present in the model.
-\begin{gather*}
+\begin{align*}
 addAssoc : Label \times Association \times Application \rightarrow Application \\
@@
 l \in Labels, s \in Associations, a \in Applications
-\end{gather*}
+\end{align*}
 Adding an association into a database is complicated because of already existing instances.
@@
 The $insertAssoc$ function is defined as:
-\begin{gather*}
+\begin{align*}
 insertAssoc : Association \times Database \times (Row \times Row) \\
@@
  a \in Associations, d \in Databases
-\end{gather*}
+\end{align*}
 
-\begin{gather*}
+\begin{align*}
 \mathbf{newClass} : Class \times State \rightarrow State \\
@@
 \Phi(s, newClass(c)) = database(s)  \bullet addTable(\rho(c,d), d) \\ \bullet addColumn(name(c), \{\forall r \in singleProperties(c) | \rho(r,d)\}, d) \\ \bullet addTable(name(c), \{\forall r \in collectionProperties(c) | \rho(r,d)\}, d)
-\end{gather*}
-\begin{gather*}
+\end{align*}
+\begin{align*}
 addClass: Class \times Application \rightarrow Application \\ 
@@
 c \in Classes, a \in Applications
-\end{gather*}
+\end{align*}
 \paragraph{Add Class} The $insertClass$ operation changes the structure of the database schema only, because the $insertAssoc$ transformation needs a mapping function to work properly, which is not part of the $newClass$ definition.
 The situation reflect the modeling praxis when the structure of a class is defined and then it is associations are created.
-%\begin{gather*}
+%\begin{align*}
 %insertClass : Class \times Database \rightarrow Database \\
@@
 %}
-%\end{gather*}
+%\end{align*}
 
 \subsection{Software Decomposition}
-\begin{gather*}
+\begin{align*}
 RemoveClass: Label \times Software \rightarrow State  \\
@@
 \wedge !\: instantiated(c) = \varnothing \wedge !\: associated(c)} {classes(a) \setminus c } \\
-\end{gather*}
-\begin{gather*}
+\end{align*}
+\begin{align*}
 dropClass(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \not \exists r \in data(d) : owningTable(r) = l \\  \not \exists j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) = l }{ tables(d) \setminus \rho(c)}
-\end{gather*}
+\end{align*}
 
-\begin{gather*}
+\begin{align*}
 RemoveProperty: Label \times Label \times State \rightarrow State \\
@@
 l_c, l_p \in Label, a \in Applications 
-\end{gather*}
-\begin{gather*}
+\end{align*}
+\begin{align*}
 dropProp(p, d) = \begin{cases}
@@
  \end{cases}
-\end{gather*}
+\end{align*}
 
-\begin{gather*}
+\begin{align*}
 RemoveAssociation : Label \times Labels \times State \rightarrow State \\ 
@@
 l_c, l_a \in Labels, a \in Applications
-\end{gather*}
-\begin{gather*}
+\end{align*}
+\begin{align*}
 dropAssoc(l, l_a, d) =  \begin{cases}
@@
  \end{cases}
-\end{gather*}
+\end{align*}
 
@@
 \paragraph{Rename Class}
-\begin{gather*}
+\begin{align*}
 RenameClass : Label \times Label \times State \rightarrow State \\ 
@@
 l_o, l_n \in Label, s \in States, a \in Applications, d \in Databases
-\end{gather*}
+\end{align*}
 
 \paragraph{Rename Property}
-\begin{gather*}
+\begin{align*}
 RenameProperty : Label \times Label \times Label \times State \rightarrow State \\ \\
@@
 l_c, l_o, l_n \in Label, a \in Applications, s \in States, d \in Databases
-\end{gather*}
+\end{align*}
 
 \paragraph{Rename Association}
-\begin{gather*}
+\begin{align*}
 RenameAssociation : Label \times Label \times Label \times State \rightarrow State \\
 \Psi(RenameAssocaiation(l_c, l_o, l_n), s) = application(s) \circ renAssoc(l_c, l_o, l_n)\\
-\Phi(RenameAssocaiation(l_c, l_o, l_n), s) = database(s) \bullet alterAsspoc(l_c, l_o, l_n)\\\\
+\Phi(RenameAssocaiation(l_c, l_o, l_n), s) = \begin{cases}
+	\inference{\mathbf{TODO}}{database(s) \bullet alterTable(l_c, l_o, l_n)}
+ \\
+  \inference{\mathbf{TODO}}{database(s) \bullet alterColumn(l_c, l_o, l_n)}
+ \end{cases}
+\\\\
 renAssoc(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists b \in associations(c) : name(b) = l_o \\ \wedge \forall e \in associations(c) : name(e) \neq l_n}{\begin{gathered}
@@
 alterColumn : Label \times Label \times Label \times Database \rightarrow Database \\
-	alterAssoc(l_c, l_o, l_n, d) = \begin{cases}
-?? \inference{\exists t \in tables(a) : name(t) = l_c \\ \wedge \exists f \in fks(t) : name(c) = l_o \\ \wedge \forall e \in fks(t) : name(e) \neq l_n}{\begin{gathered}
-f.name = l_n 
-\end{gathered}}
-\\ \\
-\inference{\exists t \in tables(a) : name(t) = l_o \\ \wedge \forall e \in tables(t) : name(e) \neq l_n}{\begin{gathered}
-t.name = l_n 
-\end{gathered}}
- \end{cases}
- \\
+%	alterAssoc(l_c, l_o, l_n, d) = \begin{cases}
+%?? \inference{\exists t \in tables(a) : name(t) = l_c \\ \wedge \exists f \in fks(t) : name(c) = l_o \\ \wedge \forall e \in fks(t) : name(e) \neq l_n}{\begin{gathered}
+%f.name = l_n 
+%\end{gathered}}
+%\\ \\
+%\inference{\exists t \in tables(a) : name(t) = l_o \\ \wedge \forall e \in tables(t) : name(e) \neq l_n}{\begin{gathered}
+%t.name = l_n 
+%\end{gathered}}
+% \end{cases}
+% \\
 l_c, l_o, l_n \in Label, a \in Applications, s \in Sates, d \in Databases
-\end{gather*}
+\end{align*}
 
@@
 The copy property can be implemented as a transformation composed by operations for adding and removing property, nevertheless this composition would not work with instances, where the $copyData$ transformation is used.
-\begin{gather*}
-	CopyProperty : Label \times Label \times Label \times (Instance \times Instance) \times State \rightarrow State 
-\\
+\begin{align*}
+	\mathbf{CopyProperty} : Label \times Label \times Label \times (Instance \times Instance) \times State \rightarrow State \\
 	\Psi(CopyProperty(l_s, l_t, l_p, f_m), s) = application(s) \circ copyProp(l_s,l_t, l_p) \\
-\Phi(CopyProperty(l_s, l_t, l_p, f_m), s) = database(s) \bullet copyColumn(l_s, l_t, l_p)\\  \bullet copyValues(l_s, l_t, l_p, f_m) \\ 
-\end{gather*}
+	\Phi(CopyProperty(l_s, l_t, l_p, f_m), s) = \begin{cases}
+		\inference{\mathbf{TODO}}{\begin{gathered}
+			database(s) \bullet copyColumn(l_s, l_t, l_p) \\ \bullet copyValues(l_s, l_t, l_p, f_m) 
+			\end{gathered}} \\ \\
+		\inference{\mathbf{TODO}}{\begin{gathered}
+			database(s) \bullet copyTable(l_s, l_t, l_p) \\ \bullet copyRows(l_s, l_t, l_p, f_m) 
+			\end{gathered}}
+ \end{cases}
+\end{align*}
 
 With a copy property transformation defined we are able to define advanced evolution scenarios.
-
-\subsection{Composing of Evolution Operations}
 
@@
 Applying of a transformation is defined:
-%\begin{gather*}
+%\begin{align*}
 %u \oplus t = ret(<application(u) \oplus o_{app}, database(u) \oplus o_{db}>) \\
 %u \in Softwares, t \in Transformations, o_{app} \in AppOps, o_{db} \in DbOps
-%\end{gather*}
+%\end{align*}

### 2013-01-08 doc/description.tex
updated orm mappings and app transformation moved into one section

@@
 \end{align*}
-
-\subsubsection{Application Transformation}
 
@@
 We focus on structural and data change therefore we simplify types and its mapping.
-
-\begin{equation*}
-\begin{gathered}
-	\rho_{t} : AppType \rightarrow DbType  \\
+\begin{align*}
+	\mathbf{\rho_{t}} : AppType \rightarrow DbType  \\
  	\rho_{t}(APPSTRING) = DBSTRING 
-\end{gathered}
-\end{equation*}
+\end{align*}
 
@@
 
-\begin{center}
-$\rho_{c}: Class \times Database \rightarrow Database $ \\ 
-$\rho_{c}(c,d) = table(name(c), primaryKey("Id"), \varnothing, \varnothing) $
-\end{center}
+\begin{align*}
+	\mathbf{\rho_{c}} : Class*  \rightarrow Table* \\
+	\rho_{c}(c) = \begin{cases}
+		\inference{c = \varnothing}{\varnothing}\\\\
+		\inference{|c| = 1}{table(name(c), primaryKey("Id"), \varnothing, \varnothing)}  \\ \\
+		\inference{|c| > 1}{\rho_c(head(c)) \cup \rho_c(tail(c))}
+ \end{cases}\\
+ c = Class
+\end{align*}
 
@@
 
-$
-\rho_p : Property \times Database \rightarrow Database 
-$
-
-$\rho_p(p, d) = \begin{cases}
+\begin{align*}
+	\mathbf{\rho_p} : Property* \rightarrow Column* \\ \\
+	\rho_{sp}(p) = \begin{cases}
  		\inference{cardinality(p) = 1 \wedge !\;mandatory(p)
+ 			\\ \wedge \exists t : t = table(name(owningClass(p)), d) 
+		 }{\begin{gathered}
+	  		column(name(p), \rho_t(type(p)), \varnothing )
+		\end{gathered}} \\ \\
+  		\inference{cardinality(p) = 1 \wedge mandatory(p)
 			 \\ \wedge \exists t : t = table(name(owningClass(p)), d)
@@
 		}{\begin{gathered}
-	  column(name(p), \rho_t(type(p)), \varnothing )
-\end{gathered}
- } \\ \\
-
-  \inference{cardinality(p) = 1 \wedge mandatory(p)
- \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}
  	 		column(name(p), \rho_t(type(p)), NOTNULL ) \\ 
 		\end{gathered}}
-\\ \\
-
- \inference{ cardinality(p) > 1 \wedge !\;mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered}table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), \varnothing), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL) \\
+ 	\end{cases}
+\end{align*}
+\begin{align*}
+	\mathbf{\rho_p} : Property* \rightarrow Table* \\ \\
+	\rho_{cp}(p) = \begin{cases}
+ 	\inference{cardinality(p) > 1 \wedge !\;mandatory(p) \\ 
+ 		\wedge 	\exists t : t = table(name(owningClass(p)), d)
+ 		}{\begin{gathered}
+ 			table(name(p), primaryKey("Id"), \\ column("value", 			\rho_t(type(p)), \varnothing), \\ foreignKey(name(p), 			name(owningClass(p)), NOTNULL) \\
+	\end{gathered}}\\ \\
+ 	\inference{ cardinality(p) > 1 \wedge mandatory(p) \\ 
+ 		\wedge \exists t : t = table(name(owningClass(p)), d) 
+ 		}{\begin{gathered}  
+ 			table(name(p), primaryKey("Id"), \\ 
+ 			column("value", \rho_t(type(p)), NOTNULL), \\ 			foreignKey(name(p), name(owningClass(p)), NOTNULL)
 	 \end{gathered}}
-\\ \\
-
- \inference{ cardinality(p) > 1 \wedge mandatory(p) \\ \wedge \exists t : t = table(name(owningClass(p)), d) }{\begin{gathered}  table(name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), NOTNULL), \\ foreignKey(name(p), name(owningClass(p)), NOTNULL)
-	 \end{gathered}}
- \end{cases}$
+ 	\end{cases}
+\end{align*}
 
@@
 
-$
-\rho_a(a,d) = \begin{cases}
- \inference{ startCardinality(a) = 0 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{ 
- \begin{gathered}
- foreignKeys(u) \\ \cup foreignKey(name(a), name(t),  \varnothing) 
- \end{gathered}
+\begin{align*}
+	\mathbf{\rho_a} : Association* \rightarrow ForeignKey* \\
+	\rho_a(a) = \begin{cases}
+		\inference{ startCardinality(a) = 0 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)
+		}{
+			foreignKey(name(a), name(t),  \varnothing) 
 	 	}
   \\ \\
- \inference{ startCardinality(a) = 1 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{\begin{gathered}  
-foreignKeys(u) \\ \cup  foreignKey(name(a), name(t),  NOTNULL) \\
-\end{gathered}
-} \\ \\
-
-  \inference{  startCardinality > 1 \\ \wedge \rho(owningClass(p)) \in tables(d) \wedge \rho(reference(a)) \in tables(d)}{
-  \begin{gathered}  
+ 	 \inference{ startCardinality(a) = 1 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)
+ 	 }{ 
+		foreignKey(name(a), name(t),  NOTNULL)
+	}
+	 \end{cases}
+\end{align*}
+\begin{align*} 
+	\mathbf{\rho_a} : Association* \rightarrow Table* \\
+  	\rho_a(a) = \inference{  startCardinality > 1 \\ \wedge \rho(owningClass(p)) \in tables(d) \wedge \rho(reference(a)) \in tables(d)
+  		}{\begin{gathered}  
 		 table(name(a), pk("Id"), \varnothing, \\ foreignKey(owningClass(a),owningClass(a), NOTNULL) \\ foreignkey(reference(a), reference(a), NOTNULL)) 
   		\end{gathered}}  
- \end{cases}
-$
+\end{align*}
+
+\section{Model Manipulating Transformation}
+\subsection{Application Transformation}
+\paragraph{Add Property}Adding a property into an application structure has to verify a name collision only:
+\begin{align*}
+ 	\mathbf{addProp} : Label \times Property \times Application \rightarrow Application \\
+	addProp(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(p) : name(q) \neq name(p)}{properties(c) \cup p } \\
+	l \in Label, p \in Properties, a \in Applications
+\end{align*}
+\paragraph{Add Association} Adding an association into an application structure has to verify if there is no name collision and if there is the associated class present in the model.
+\begin{align*}
+	\mathbf{addAssoc} : Label \times Association \times Application \rightarrow Application \\
+	addAssoc(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(c) \cup s}\\
+	l \in Labels, s \in Associations, a \in Applications
+\end{align*}
+\paragraph{Add Class} Adding a class is limited to the case when the class has no associations, this is because it simplifies the database migration and because it is a natural way of modeling.
+Next the transformation $addClass$ has to verify there is no name collision
+\begin{align*}
+	\mathbf{addClass} : Class \times Application \rightarrow Application \\ 
+	addClass(c, a) = \inference{\forall s \in classes(a) : name(s) \neq name(c) \\ \wedge associations(c) = \varnothing }{classes(a) \cup c} \\ \\
+	c \in Classes, a \in Applications
+\end{align*}
+\paragraph{Remove Property}
+\begin{align*}
+ 	\mathbf{remProp} : Label \times Label \times Application \rightarrow Application \\
+ 	remProp(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p
+	}{properties(c) \setminus p }\\
+	l_c, l_p \in Label, a \in Applications 
+\end{align*}
+\paragraph{Remove Association}
+\begin{align*}
+	\mathbf{remAssoc} : Label \times Label \times Application \rightarrow Application \\
+	remAssoc(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) \setminus f} \\
+	l_c, l_a \in Labels, a \in Applications
+\end{align*}
+\paragraph{Remove Class} The class can be removed only in case it is not referenced by any association existing in the application.
+\begin{align*}
+	\mathbf{remClass} : Label \times Application \rightarrow Application \\
+	remClass(l, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge !\: associated(c)
+	} {classes(a) \setminus c }\\
+	l \in Label, a = Application
+\end{align*}
+\paragraph{Rename Property}
+\begin{align*}
+	\mathbf{renProp} : Label \times Label \times Label \times Application \rightarrow Application \\
+	renProp(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_o \\ \wedge \forall e \in properties(c) : name(e) \neq l_n
+	}{\begin{gathered}
+		p.name = l_n 
+	\end{gathered}}
+\end{align*}
+\paragraph{Rename Association}
+\begin{align*}
+	\mathbf{renAssoc} : Label \times Label \times Label \times Application \rightarrow Application \\
+	renAssoc(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists b \in associations(c) : name(b) = l_o \\ \wedge \forall e \in associations(c) : name(e) \neq l_n
+	}{\begin{gathered}
+		b.name = l_n 
+	\end{gathered}}
+\end{align*}
+\paragraph{Rename Class}
+\begin{align*}
+	\mathbf{renClass} :  Label \times Label \times Application \rightarrow Application \\
+	renClass(l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_o \\ \wedge \forall e \in classes(a) : name(e) \neq l_n
+	}{\begin{gathered}
+		c.name = l_n \wedge \forall e \in classes(a) \\ \wedge \forall (f \in associations(e) \wedge reference(f) = l_o ) \\ : f.reference = l_n 
+	\end{gathered}}
+\end{align*}
 
@@
 \end{align*}
-Adding a property into an application structure has to verify a name collision only:
-\begin{align*}
-addProp : Label \times Property \times Application \rightarrow Application \\
-addProp(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(p) : name(q) \neq name(p)}{properties(c) \cup p } \\
-l \in Label, p \in Properties, a \in Applications
-\end{align*}
+
 Whereas the $insertProp$ on database level can change stored data too, because a mandatory property can be added into a class with already existing instances, then a default value has to be added to all existing instances.
@@
 \end{align*}
-Adding an association into an application structure has to verify if there is no name collision and if there is the associated class present in the model.
-\begin{align*}
-addAssoc : Label \times Association \times Application \rightarrow Application \\
-addAssoc(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(c) \cup s}\\
-l \in Labels, s \in Associations, a \in Applications
-\end{align*}
+
 Adding an association into a database is complicated because of already existing instances.
@@
 \end{align*}
-\begin{align*}
-addClass: Class \times Application \rightarrow Application \\ 
-addClass(c, a) = \inference{\forall s \in classes(a) : name(s) \neq name(c)}{classes(a) \cup c} \\ \\
-c \in Classes, a \in Applications
-\end{align*}
+
 \paragraph{Add Class} The $insertClass$ operation changes the structure of the database schema only, because the $insertAssoc$ transformation needs a mapping function to work properly, which is not part of the $newClass$ definition.
@@
 \\ l \in Label, s \in Software 
-\\ \\
-remClass(l, a) = \inference{\exists c \in classes(a) : name(c) = l \\
-\wedge !\: instantiated(c) = \varnothing \wedge !\: associated(c)} {classes(a) \setminus c } \\
 \end{align*}
@@
 \Psi RemoveProperty(l, l_p, s) =  applications(s) \circ  remProp(l, l_p) \\
-\Phi RemoveProperty(l, l_p, s) =  database(s) \bullet dropProp(l, l_p) \\
-RemoveProperty(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{properties(c) \setminus p }\\
-l_c, l_p \in Label, a \in Applications 
-\end{align*}
+\Phi RemoveProperty(l, l_p, s) =  database(s) \bullet dropProp(l, l_p) \end{align*}
 \begin{align*}
@@
 \Phi RemoveAssociation(l, l_a, s) = database(s) \bullet dropAssoc(l, l_a) \\
-RemoveAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) \setminus f} \\
 l_c, l_a \in Labels, a \in Applications
@@
 \Phi(RenameClass(l_o, l_n), s) = database(s) \bullet alterClass(l_o, l_n) \\ \\
-renClass:  Label \times Label \times Application \rightarrow Application \\
-renClass(l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_o \\ \wedge \forall e \in classes(a) : name(e) \neq l_n}{\begin{gathered}
-c.name = l_n \wedge \forall e \in classes(a) \\ \wedge \forall (f \in associations(e) \wedge reference(f) = l_o ) \\ : f.reference = l_n 
-\end{gathered}}\\ \\
+\end{align*}
+
+\begin{align*}
 alterTable:  Label \times Label \times Database \rightarrow Database \\
@@
 \Phi(RenameProperty(l_c, l_o, l_n, s)) = database(s) \bullet alterColumn(l_c, l_o, l_n) \\\\
-renProp : Label \times Label \times Label \times Application \rightarrow Application \\
-	renProp(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_o \\ \wedge \forall e \in properties(c) : name(e) \neq l_n}{\begin{gathered}
-p.name = l_n 
-\end{gathered}}\\\\
+\end{align*}
+
+\begin{align*}
 alterColumn : Label \times Label \times Label \times Database \rightarrow Database \\
@@
 \\\\
-renAssoc(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists b \in associations(c) : name(b) = l_o \\ \wedge \forall e \in associations(c) : name(e) \neq l_n}{\begin{gathered}
-b.name = l_n 
-\end{gathered}}\\\\
 alterColumn : Label \times Label \times Label \times Database \rightarrow Database \\
@@
 	\mathbf{CopyProperty} : Label \times Label \times Label \times (Instance \times Instance) \times State \rightarrow State \\
-	\Psi(CopyProperty(l_s, l_t, l_p, f_m), s) = application(s) \circ copyProp(l_s,l_t, l_p) \\
+	\Psi(CopyProperty(l_s, l_t, l_p, f_m), s) = \inference{\exists c \in classes(a) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)} \\
 	\Phi(CopyProperty(l_s, l_t, l_p, f_m), s) = \begin{cases}

### 2013-01-08 doc/description.tex
database transformation moved into section about transformations
appendValues renamed to insertValues

@@
 
-\subsection{Transformations for Manipulating a Database}
-The database is composed from two separated parts - database schema and data. Several transformation are defined to make manipulating a database easily in this paper.
-
-\begin{align*}
-	\mathbf{addColumn} : Label \times Column* \times Database \rightarrow Database \\ 	
-	addColum(l, c, d) = \begin{cases}
-		\inference{|c| > 1}{addColumn(l, head(c), d) \bullet addColumn(l, tail(c), d)} \\ \\
-		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c)}{columns(t) \cup c}\\ \\
-		Inconsistent \;d 
-	 \end{cases}\\ 
-	 l \in Label, c = Column*, d = Database
-\end{align*}
-
-\begin{align*}
-	\mathbf{copyColumn} : Label \times Label \times Label \times Database \rightarrow Database \\
-	copyColumn(l_s, l_t, l_c, d) = \begin{cases}
- 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{columns(e) \cup c }
-	\end{cases} \\
-	l_s,l_t,l_c \in Label, d = Database
-\end{align*}
-
-\begin{align*}
-	\mathbf{addTable} : Table* \times Database \rightarrow Database \\ 
-	addTable(t, d) = \begin{cases}
-		\inference{|t| > 1}{addTable(head(t),d) \bullet addTable(tail(t), d)}
-\\ \\	
-		\inference{|t| = 1 \wedge \forall e \in tables(d) \wedge name(t) \neq name(e)}{tables(d) \cup c}
-\\ \\
-		Inconsistent \;d 
-	 \end{cases}\\ 
-	t = Table*,  d = Database
-\end{align*}
-
-\begin{align*}
-	\mathbf{insertRow} : Row* \times Database \rightarrow Database \\
-	insertRow(r, d) = \begin{cases}
-			\inference{|r|>1}{insertRow(head(r), d) \bullet insertRow(tail(r), d)} \\ \\
-		 	\inference{|r|=1 \wedge \exists t \in tables(d) : name(t) = reference(r) }{data(d) \cup r} \\ \\
-			 Inconsistent \; d
-		 \end{cases}	\\
-		 r = Row*, d = Database
-\end{align*}
-
-\begin{align*}
-	\mathbf{appendValues} : Label \times Pair* \times Database \rightarrow Daatabase \\
-	appendValues(l, p, d) = \begin{cases}
- 		\inference{|p| > 1}{appendValues(l, head(p), d)\bullet appendValues(l, tail(p), d)}  \\ \\
- 		\inference{|p| = 1 \wedge \forall e \in pairs(r_t) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r_t), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p} \\ \\
-	 	 Inconsistent \; d
- 		\end{cases} \\
-l \in Label, p = Pair*, d = Databases
-\end{align*}
-
-\begin{align*}
-	\mathbf{copyValue} : Label \times Label \times Label \times (Instancece \times Instance) \times Database \times Database \\
-	copyValue(l_s, l_t, l_{sc}, l_{tc} f_m, d) = \begin{cases}
- 		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc} }{\begin{gathered}
-			\forall p \in f_m(t_s, t_t)  : \\ appendValues(name(second(p)), \\ [l_{tc},value(pairByRef(l_{sc}, first(p))),d)
-		\end{gathered}}
- 	\end{cases}
-\end{align*}
-
 \subsection{Extension of Software Definition}
@@
 	\end{gathered}}
+\end{align*}
+
+\subsection{Database Transformation}
+The database is composed from two separated parts - database schema and data. Several transformation are defined to make manipulating a database easily in this paper.
+\paragraph{Add Column}
+\begin{align*}
+	\mathbf{addColumn} : Label \times Column* \times Database \rightarrow Database \\ 	
+	addColum(l, c, d) = \begin{cases}
+		\inference{|c| > 1}{addColumn(l, head(c), d) \bullet addColumn(l, tail(c), d)} \\ \\
+		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c)}{columns(t) \cup c}\\ \\
+		Inconsistent \;d 
+	 \end{cases}\\ 
+	 l \in Label, c = Column*, d = Database
+\end{align*}
+\paragraph{COPY COLUMN}
+\begin{align*}
+	\mathbf{copyColumn} : Label \times Label \times Label \times Database \rightarrow Database \\
+	copyColumn(l_s, l_t, l_c, d) = \begin{cases}
+ 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{columns(e) \cup c }
+	\end{cases} \\
+	l_s,l_t,l_c \in Label, d = Database
+\end{align*}
+\paragraph{Add Table}
+\begin{align*}
+	\mathbf{addTable} : Table* \times Database \rightarrow Database \\ 
+	addTable(t, d) = \begin{cases}
+		\inference{|t| > 1}{addTable(head(t),d) \bullet addTable(tail(t), d)}
+\\ \\	
+		\inference{|t| = 1 \wedge \forall e \in tables(d) \wedge name(t) \neq name(e)}{tables(d) \cup c}
+\\ \\
+		Inconsistent \;d 
+	 \end{cases}\\ 
+	t = Table*,  d = Database
+\end{align*}
+\paragraph{Insert row}
+\begin{align*}
+	\mathbf{insertRow} : Row* \times Database \rightarrow Database \\
+	insertRow(r, d) = \begin{cases}
+			\inference{|r|>1}{insertRow(head(r), d) \bullet insertRow(tail(r), d)} \\ \\
+		 	\inference{|r|=1 \wedge \exists t \in tables(d) : name(t) = reference(r) }{data(d) \cup r} \\ \\
+			 Inconsistent \; d
+		 \end{cases}	\\
+		 r = Row*, d = Database
+\end{align*}
+\paragraph{Insert Values}
+\begin{align*}
+	\mathbf{insertValues} : Label \times Pair* \times Database \rightarrow Daatabase \\
+	insertValues(l, p, d) = \begin{cases}
+ 		\inference{|p| > 1}{insertValues(l, head(p), d)\bullet insertValues(l, tail(p), d)}  \\ \\
+ 		\inference{|p| = 1 \wedge \forall e \in pairs(r_t) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r_t), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p} \\ \\
+	 	 Inconsistent \; d
+ 		\end{cases} \\
+l \in Label, p = Pair*, d = Databases
+\end{align*}
+\paragraph{COPY VALUES}
+\begin{align*}
+	\mathbf{copyValue} : Label \times Label \times Label \times (Instancece \times Instance) \times Database \times Database \\
+	copyValue(l_s, l_t, l_{sc}, l_{tc} f_m, d) = \begin{cases}
+ 		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc} }{\begin{gathered}
+			\forall p \in f_m(t_s, t_t)  : \\ insertValues(name(second(p)), \\ [l_{tc},value(pairByRef(l_{sc}, first(p))),d)
+		\end{gathered}}
+ 	\end{cases}
 \end{align*}

### 2013-01-09 doc/description.tex
fixed associated(l,a) function for class

@@
   class(l, a) = c, c \in classes(a) \wedge name(c) = l \\ \\
-  l \in Label, a = Application, cs \in Class* 
-  \\ \\
-  \mathbf{associated} :  Label \times Application \rightarrow Boolean \\
+  l \in Label, a = Application, cs = Class* 
 \end{align*}
@@
 	\mathbf{associations} : Class \rightarrow Association* \\	associations(cl(l, props, assocs)) = assocs \\ \\
-l \in Label, props \in Properties, assocs \in Associations
+	\mathbf{associated} :  Label \times Application \rightarrow Boolean \\
+  	associated(l,a) = \begin{cases}
+  		\inference{\exists c \in classes(a) \wedge \exists e \in associations(c) : reference(e) = l
+  	}{ true } \\
+ 		false
+ 	\end{cases} \\ \\
+	l \in Label, props = Property*, assocs = Association*, a = Application
 \end{align*}
@@
 \end{align*}
-
-%TODO: variable types
 

### 2013-01-09 doc/description.tex
row and pair reformated

@@
 	mandatory(prop(l, t, val, n, m)) = m  \\ \\
+	\mathbf{owningClass} : Property \times Application \rightarrow Class  \\ 	
+	owningClass(p, a) = \inference{\exists c \in classes(a) \wedge p \in properties(c)}{c}  \\ \\
 	l \in Label, t \in AppType, val \in String*, n \in \mathbb{N^{*}}, m \in Boolean 
@@
 	endCardinality(assoc(l, ref, n_1, n_2)) = n_2 \\ \\
-l, ref \in Label,  n_1, n_2 \in \mathbb{N}
+	\mathbf{owningClass} : Association \times Application \rightarrow Class  \\
+	owningClass(as, a) = \inference{\exists c \in classes(a) \wedge as \in associations(c)}{c}  \\ \\	
+	l, ref \in Label,  n_1, n_2 \in \mathbb{N}, as = Association
 \end{align*}
@@
 	constraints(col(l, t, val, cons)) = cons  \\ \\
-l \in Label, t = DbType, val = Strings, cons = Constraints
+	\mathbf{owningTable} : Column \times Database \rightarrow Table  \\
+	owningTable(c, d) = \inference{\exists tab \in tables(d) \wedge c \in columns(tab)}{tab} \\ \\
+l \in Label, t = DbType, val = String,\\ cons = Constraints, c = Column, d = Database
 \end{align*}
@@
 	constaints(fk(l, tRef, cons)) = cons  \\ \\
-l, tRef \in Label, cons \in Constraints
+	\mathbf{owningTable} : ForeignKey \times Database \rightarrow Table  \\
+	owningTable(fk, d) = \inference{\exists tab \in tables(d) \wedge fk \in foreignKeys(tab)}{tab}\\\\
+	l, tRef \in Label, cons = Constraint*, d = Database, fk = ForeignKey
 \end{align*}
@@
 	\mathbf{refTable} : Row \rightarrow Label \\
-	refTable(row(l, pairs) = l \\ \\
+	refTable(row(l, ps) = l \\ \\
+	\mathbf{pairs} : Row \rightarrow Pair* \\
+	pairs(row(l, ps)) = ps \\ \\
+	\mathbf{pairByRef} : Label \times Row \rightarrow Pair \\
+	pairByRef(l,i) = \begin{cases}
+		\inference{\exists p \in pairs(i) : reference(p) = l}{p} \\ \\
+			\varnothing
+		\end{cases}\\
+	 l \in Label, ps = Pair*, i = Row
+\end{align*}
+The value stored in concrete column of the row is represented by pair, where the $refCol$ is a name of the column, a foreign key or its primary key and $value$ represents concrete value .
+\begin{align*}
+	\mathbf{Pair} = < refCol, value > \\
+	[Label \times Value] \rightarrow Pair\\ \\
+	\mathbf{reference} : Pair \rightarrow Label \\
+	reference([l,val]) = l \\ \\
+	\mathbf{value} : Pair \rightarrow Label \\ 
+	value([l,val]) = val \\ \\
+	l \in Label, val \in Value
+\end{align*}
+There functions $contains$ verify if the table contains data in context of concrete database.
+\begin{align*}
 	\mathbf{contains} : Table \times Database \rightarrow Boolean \\
@@
 \end{align*}
-There functions $contains$ verify if the table contains data in context of concrete database.
-
-\begin{align*}
-	\mathbf{Pair} = < refCol, value > \\
-	[Label \times Value] \rightarrow Pair\\ \\
-	\mathbf{reference} : Pair \rightarrow Label \\
-	reference([l,val]) = l \\ \\
-	\mathbf{value} : Pair \rightarrow Label \\ 
-	value([l,val]) = val \\ \\
-	\mathbf{pairByRef} : Label \times Row \rightarrow Pair \\
-	pairByRef(l,i) = \begin{cases}
-			\inference{\exists p \in pairs(i) : reference(p) = l}{p} \\ \\
-			\varnothing
-		\end{cases}\\
-	l \in Label, val \in Value
-\end{align*}
-The $refCol$ is a name of a column, a foreign key or its primary key and $value$ represents concrete value stored in the database.
 Each row in a database has to refer to a table in database schema and has a structure corresponding to the one defined by a table, otherwise the database is inconsistent.
@@
 \subsection{Extension of Software Definition}
-
-\begin{align*}
- owningClass : Property \times Application \rightarrow Class  \\ owningClass(p, a) = c, c \in classes(a) \wedge p \in properties(c) \\
-  owningClass : Association \times Application \rightarrow Class  \\
-owningClass(as, a) = c, c \in classes(a) \wedge as \in associations(c) \\ \\
-owningTable : Column \times Database \rightarrow Table  \\
-owningTable(c, d) = tab, tab \in tables(d) \wedge c \in columns(d) \\ \\
-owningTable : ForeignKey \times Database \rightarrow Table  \\
-owningTable(fk, d) = tab, tab \in tables(d) \wedge fk \in foreignKeys(d)\\ \\ 
-\end{align*}
-
 The database stores instances of the application, thus the $instantiated$ function has to be defined:
@@
 	\rho_{sp}(p) = \begin{cases}
+		\inference{p = \varnothing}{\varnothing}\\
+		\inference{|p| > 1}{\rho_{sp}(head(p)) \cup \rho_{sp}(tail(p))}\\\\
  		\inference{cardinality(p) = 1 \wedge !\;mandatory(p)

### 2013-01-09 doc/description.tex
orm mapping works with collections

@@
 	\rho_{cp}(p) = \begin{cases}
+		\inference{p = \varnothing}{\varnothing}\\
+		\inference{|p| > 1}{\rho_{cp}(head(p)) \cup \rho_{cp}(tail(p))}\\\\
      	\inference{cardinality(p) > 1 \wedge !\;mandatory(p) \\ 
@@
 	\rho_a(a) = \begin{cases}
+		\inference{a = \varnothing}{\varnothing}\\
+		\inference{|a| > 1}{\rho_{a}(head(a)) \cup \rho_{a}(tail(a))}\\\\
 		\inference{ startCardinality(a) = 0 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)
@@
 	\mathbf{\rho_a} : Association* \rightarrow Table* \\
-  	\rho_a(a) = \inference{  startCardinality > 1 \\ \wedge \rho(owningClass(p)) \in tables(d) \wedge \rho(reference(a)) \in tables(d)
+	\rho_a(a) = \begin{cases}
+ 		\inference{a = \varnothing}{\varnothing}\\
+		\inference{|a| > 1}{\rho_{a}(head(a)) \cup \rho_{a}(tail(a))}\\\\
+		\inference{  startCardinality > 1 \\ \wedge \rho(owningClass(p)) \in tables(d) \wedge \rho(reference(a)) \in tables(d)
   		}{\begin{gathered}  
@@
   		\end{gathered}}  
+ 	\end{cases}
 \end{align*}

### 2013-01-09 doc/description.tex
copy property defined

@@
 	\mathbf{CopyProperty} : Label \times Label \times Label \times (Instance \times Instance) \times State \rightarrow State \\
-	\Psi(CopyProperty(l_s, l_t, l_p, f_m), s) = \inference{\exists c \in classes(a) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)} \\
+	\Psi(CopyProperty(l_s, l_t, l_p, f_m), s) = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)} \\\\
 	\Phi(CopyProperty(l_s, l_t, l_p, f_m), s) = \begin{cases}
-		\inference{\mathbf{TODO}}{\begin{gathered}
+		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) \leq 1 }{\begin{gathered}
 			database(s) \bullet copyColumn(l_s, l_t, l_p) \\ \bullet copyValues(l_s, l_t, l_p, f_m) 
 			\end{gathered}} \\ \\
-		\inference{\mathbf{TODO}}{\begin{gathered}
+		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) > 1}{\begin{gathered}
 			database(s) \bullet copyTable(l_s, l_t, l_p) \\ \bullet copyRows(l_s, l_t, l_p, f_m) 
@@
 
-With a copy property transformation defined we are able to define advanced evolution scenarios.
+With the copy property transformation defined we are able to define advanced evolution scenarios.
 

### 2013-01-09 doc/description.tex
has type of: \lhd

@@
 The situation is illustrated in the Figure \ref{fig:appStructure}.
-There are three important components of such a software: application (its persistent layer concretely), database (consisting of database schema and stored data) and an ORM (we will use $\rho$ as a symbol for ORM), therefore we define a software a triple:
+There are three important components of such a software: application (its persistent layer concretely), database (consisting of database schema and stored data) and an ORM (we will use $\rho$ as a symbol for ORM), therefore we define a software as a triple:
 \begin{align*}
-	software = < Application, Database, \rho > \\
+	\mathbf{software} = < Application, Database, \rho > \\
 	sw : Applications \times Databases \times ORM \rightarrow Software \\ \\
-	application : Software \rightarrow Application \\
+	\mathbf{application} : Software \rightarrow Application \\
 	application(sw(a, d, \rho)) = a \\ \\
-	database : Software \rightarrow Database \\
+	\mathbf{database} : Software \rightarrow Database \\
 	database(sw(a, d, \rho)) = d \\
-	a \in Applications, d \in Databases
+	a \lhd Application, d \lhd Database
 \end{align*}
+The software definition above introduce the first type used in this paper. The notation $a \lhd Application$ indicates the variable $a$ has the type $Application$.
+
 The software has to be in consistent state so it is beneficial for its user. The state of the software is defined to communicate the sate of the software to its user:
 \begin{align*}
-	state = Consistent\; s \: | \: Inconsistent \; s, \\
-	s \in Softwares  \\ \\
-	state : Software \rightarrow State \\
+ 	\mathbf{state} = Consistent\; s \: | \: Inconsistent \; s, \\\\
+	\mathbf{state} : Software \rightarrow State \\
 	state(s) = \begin{cases}
- 		\inference{consistent(application(s)) \wedge \\ consistent(database(s)) \wedge \rho(application(s)) = database(s)}{Consistent \; s}\\ \\
+ 		\inference{consistent(application(s)) \wedge \\ consistent(database(s)) \\ \wedge \rho(application(s)) = database(s)}{Consistent \; s}\\ \\
  		\inference{otherwise}{Inconsistent \; s}
- 	\end{cases}
+ 	\end{cases} \\ \\
+ 	s \lhd Software  
 \end{align*}
@@
 \end{cases}\\ 
-u \in States, t \in E, s \in Softwares
+u \lhd State, t \in E, s \lhd Software
 \end{align*}
@@
 \end{cases}\\ 
-a \in Applications, o \in AppOps
+a \lhd Application, o \in AppOps
 \end{align*}
@@
 \end{cases}\\ 
-d \in Applications, o \in DbOps
+d \lhd Application, o \in DbOps
 \end{align*}
@@
 The model of an ORM software consists of three parts - an application model, a database model and a ORM.
+The model of application and database creates a type system for both kinds ORM and evolutionary transformations.
+
 The definitions of all parts assume a set $Label$ exist which contains all possible identifiers and each modeled entity is identified unambiguously identified in the model.
@@
   class(l, a) = c, c \in classes(a) \wedge name(c) = l \\ \\
-  l \in Label, a = Application, cs = Class* 
+  l \in Label,a \lhd Applications, cs = Class* 
 \end{align*}
@@
  	\end{cases} \\ \\
-	l \in Label, props = Property*, assocs = Association*, a = Application
+	l \in Label, props = Property*, assocs = Association*,a \lhd Applications
 \end{align*}
@@
 	owningClass(p, a) = \inference{\exists c \in classes(a) \wedge p \in properties(c)}{c}  \\ \\
-	l \in Label, t \in AppType, val \in String*, n \in \mathbb{N^{*}}, m \in Boolean 
+	l \in Label, t \lhd AppType, val \in Values, n \in \mathbb{N^{*}}, m \in Boolean 
 \end{align*}
@@
 	owningClass(as, a) = \inference{\exists c \in classes(a) \wedge as \in associations(c)}{c}  \\ \\	
-	l, ref \in Label,  n_1, n_2 \in \mathbb{N}, as = Association
+	l, ref \in Label,  n_1, n_2 \in \mathbb{N}, as \lhd Association
 \end{align*}
@@
 	table(l, d) = t, t \in tables(d) \wedge name(t) = l \\ \\
-	tabs = Table*, rs = Row*, l \in Label, d = Database
+	tabs = Table*, rs = Row*, l \in Label,d \lhd Database
 \end{align*}
@@
 	foreignKeys(tab(l, pk, cols, fks)) = fks \\ \\
-l \in Label, pk = PrimaryKey, cols = Column*, fks = ForeignKey*
+l \in Label, pk \lhd PrimaryKey, cols = Column*, fks = ForeignKey*
 \end{align*}
@@
 	owningTable(c, d) = \inference{\exists tab \in tables(d) \wedge c \in columns(tab)}{tab} \\ \\
-l \in Label, t = DbType, val = String,\\ cons = Constraints, c = Column, d = Database
+l \in Label, t \lhd DbType, val \in Values,\\ cons \lhd Constraints, c \lhd Column,d \lhd Database
 \end{align*}
@@
 	owningTable(fk, d) = \inference{\exists tab \in tables(d) \wedge fk \in foreignKeys(tab)}{tab}\\\\
-	l, tRef \in Label, cons = Constraint*, d = Database, fk = ForeignKey
+	l, tRef \in Label, cons \lhd Constraint*,d \lhd Database, fk \lhd ForeignKey
 \end{align*}
@@
 		\end{cases}\\
-	 l \in Label, ps = Pair*, i = Row
+	 l \in Label, ps = Pair*, i \lhd Row
 \end{align*}
@@
  		\end{cases} \\ \\
- l \in Label, ps = Pair*, t \in tables(d), d \in Database
+ l \in Label, ps = Pair*, t \in tables(d), d \lhd Database
 \end{align*}
@@
 	\Require $a \in Application$
-	\State d = Database
+	\State $d \lhd Database$
 	\ForAll{$c \in classes(a)$} 
@@
  \end{cases}\\
- c = Class
+ c \lhd Class
 \end{align*}
@@
     	 \end{gathered}}
- 	\end{cases}
+ 	\end{cases}\\
+ 	p \lhd Property
 \end{align*}
@@
 	}
-	 \end{cases}
+	 \end{cases}\\
+	 a \lhd Association
 \end{align*}
@@
   		\end{gathered}}  
- 	\end{cases}
+ 	\end{cases}\\
+	 a \lhd Association
 \end{align*}
@@
 	addProp(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(p) : name(q) \neq name(p)}{properties(c) \cup p } \\
-	l \in Label, p \in Properties, a \in Applications
+	l \in Label, p \lhd Property, a \lhd Applications
 \end{align*}
@@
 	addAssoc(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(c) \cup s}\\
-	l \in Labels, s \in Associations, a \in Applications
+	l \in Labels, s \in Associations, a \lhd Applications
 \end{align*}
@@
 	addClass(c, a) = \inference{\forall s \in classes(a) : name(s) \neq name(c) \\ \wedge associations(c) = \varnothing }{classes(a) \cup c} \\ \\
-	c \in Classes, a \in Applications
+	c \lhd Class, a \lhd Applications
 \end{align*}
@@
 	}{properties(c) \setminus p }\\
-	l_c, l_p \in Label, a \in Applications 
+	l_c, l_p \in Label, a \lhd Applications 
 \end{align*}
@@
 	remAssoc(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) \setminus f} \\
-	l_c, l_a \in Labels, a \in Applications
+	l_c, l_a \in Labels, a \lhd Applications
 \end{align*}
@@
 	} {classes(a) \setminus c }\\
-	l \in Label, a = Application
+	l \in Label,a \lhd Applications
 \end{align*}
@@
 	 \end{cases}\\ 
-	 l \in Label, c = Column*, d = Database
+	 l \in Label, c \lhd Column*,d \lhd Database
 \end{align*}
@@
 	\end{cases} \\
-	l_s,l_t,l_c \in Label, d = Database
+	l_s,l_t,l_c \in Label,d \lhd Database
 \end{align*}
@@
 	 \end{cases}\\ 
-	t = Table*,  d = Database
+	t = Table*, d \lhd Database
 \end{align*}
@@
 		 \end{cases}	\\
-		 r = Row*, d = Database
+		 r = Row*,d \lhd Database
 \end{align*}
@@
  		\end{cases} \\
-l \in Label, p = Pair*, d = Databases
+l \in Label, p = Pair*,d \lhd Databases
 \end{align*}
@@
 \Phi(s, addProperty(l, p)) = database(s) \bullet insertProp(l, p) \\
-s \in Softwares, l \in Label, p \in Properties
+s \lhd Software, l \in Label, p \in Properties
 \end{align*}
@@
  \end{cases}\\
- a \in Associations, d \in Databases
+ a \in Associations, d \lhd Database
 \end{align*}
@@
 \Phi RemoveAssociation(l, l_a, s) = database(s) \bullet dropAssoc(l, l_a) \\
-l_c, l_a \in Labels, a \in Applications
+l_c, l_a \in Labels, a \lhd Applications
 \end{align*}
@@
 \end{gathered}}\\
-l_o, l_n \in Label, s \in States, a \in Applications, d \in Databases
+l_o, l_n \in Label, s \in States, a \lhd Applications, d \lhd Database
 \end{align*}
@@
  \\
-l_c, l_o, l_n \in Label, a \in Applications, s \in States, d \in Databases
+l_c, l_o, l_n \in Label, a \lhd Applications, s \in States, d \lhd Database
 \end{align*}
@@
 % \\
-l_c, l_o, l_n \in Label, a \in Applications, s \in Sates, d \in Databases
+l_c, l_o, l_n \in Label, a \lhd Applications, s \in Sates, d \lhd Database
 \end{align*}
@@
 %\end{cases}\\ 
-%u \in States, t \in Transformations, s \in Softwares
+%u \in States, t \in Transformations, s \lhd Software
 %\end{gathered}

### 2013-01-09 doc/description.tex
Improved model definitions descriptions

@@
 
-The software has to be in consistent state so it is beneficial for its user. The state of the software is defined to communicate the sate of the software to its user:
+The software has to be in consistent state so its users could have benefit from its useage. The state of the software is defined to communicate the sate of the software to its user:
 \begin{align*}
@@
 \subsubsection{Application Structure}
-Each instance is defined according to a class - the class defines the structure of its instances and relationships between them an the rest of the application.
+Each instance in an application is defined according to some class - the class defines the structure of its instances and relationships between them an the rest of the application.
 The instances are stored in the relational database in our case, therefore we define only the structure of classes in our model.
-Besides the structure definition, there is defined a set of functions, which allows to manipulate the application model easily.
-
 \begin{align*}
@@
 
-\paragraph{Class} represents a basic structural concept in the application model. It has a unique name, one or more properties and a class can be associated to other classes in the application. 
+\paragraph{Class} Class represents a basic structural unit in the application model. It has a unique name, one or more properties and it can be associated to other classes in the application. 
 \begin{align*}
@@
 	\mathbf{cl} : Label \times Property* \times Association* \rightarrow Class \\\\
-% TODO association has to be only to self (all associated has to exists)
 	\mathbf{name} : Class \rightarrow Label \\
@@
 	properties(cl(l, props, assocs)) = props \\ \\
-	\mathbf{primitiveProperties} : Class \rightarrow Property* \\	primitiveProperties(cl(l, props, assocs)) = props', \\ props' \subset props,  \forall p \in props' : cardinality(p) \leq 1 \\ \\
-	\mathbf{collectionProperties} : Class \rightarrow Property* \\	collectionProperties(cl(l, props, assocs)) = props', \\ props' \subset props, \forall p \in props' : cardinality(p) > 1 \\ \\
 	\mathbf{associations} : Class \rightarrow Association* \\	associations(cl(l, props, assocs)) = assocs \\ \\
+\end{align*}
+Three more function are defined to obtain information about a class - $primitives$ returns all properties of the class which represent a single value, whereas $collections$ returns all properties representing collections.
+The last one $associated$ determine if a class is associated by any other class in an application.
+\begin{align*}
+	\mathbf{primitives} : Class \rightarrow Property* \\	primitives(cl(l, props, assocs)) = props', \\ props' \subset props,  \forall p \in props' : cardinality(p) \leq 1 \\ \\
+	\mathbf{collections} : Class \rightarrow Property* \\	collections(cl(l, props, assocs)) = props', \\ props' \subset props, \forall p \in props' : cardinality(p) > 1 \\ \\
 	\mathbf{associated} :  Label \times Application \rightarrow Boolean \\
   	associated(l,a) = \begin{cases}
-  		\inference{\exists c \in classes(a) \wedge \exists e \in associations(c) : reference(e) = l
-  	}{ true } \\
+  		\inference{\exists c \in classes(a) \wedge \exists e \in associations(c) : \\ reference(e) = l
+  	}{ true } \\\\
  		false
@@
 
-\paragraph{Property} represents a feature of  a class which is represented as a primitive type.
-The property can be mandatory, can have a default value and can represent a single value or a collection of values.
+\paragraph{Property} Property represents a feature of a class which is represented as a primitive type.
+The property can be mandatory, can have a default value and according to its cardinality it can represent a single value or a collection of values.
+The $owningClass$ function returns the class in the model which owns the property.
 \begin{align*}
@@
 
-\paragraph {Association} represents a connection between two classes.
-It has a unique name and it contains a name of the class which is referenced by the association.
+\paragraph {Association} Assoctation represents a connection between two classes.
+It has a unique name and the reference is represented by the label of referenced class.
 The class which owns the association is consider to be the starting class of an association, referenced class is consider to be the ending class of an association.
@@
 	\mathbf{Association} = <label, classRef, startCardinality, endCardinality> \\ \\
-	\mathbf{assoc} : Label \times Label \times \mathbb{N} \times \mathbb{N} \rightarrow Association\\
+	\mathbf{assoc} : Label \times Label \times \mathbb{N} \times \mathbb{N} \rightarrow Association\\ \\
 	\mathbf{name} : Association \rightarrow Label \\
@@
 
-\paragraph{AppType} represents primitive types in the application.
+\paragraph{Application type} Application type ($AppType$) represents primitive type in the application.
 There are usually defined types such as String, Integer, Boolean etc.
 in contrast there is only one type in our model, because we focus on structural and data changes and type casting operations are not important for us.
-The only App-Type type is called APP-STRING.
+The only App-Type type is called $APPSTRING$.
 \begin{align*}
@@
 \begin{align*}
-	\mathbf{Database} = <table*, row*> \\ \\
+	\mathbf{Database} = < Table*, Row*> \\ \\
 	\mathbf{db} : Tables \times Rows \rightarrow Database \\ \\
@@
 	rows(db(tabs, rs)) = rs \\ \\
-	\mathbf{table} : Label \times Database \rightarrow Table  \\
-	table(l, d) = t, t \in tables(d) \wedge name(t) = l \\ \\
 	tabs = Table*, rs = Row*, l \in Label,d \lhd Database
 \end{align*}
-
-\subsubsection{Database Schema Model}
-The database schema model is defined by following concepts:
 \paragraph{Table} represents a basic concept of database schema.
@@
 
+The function $findTable$ finds and returns a table in the database according to the provided label.
+\begin{align*}
+	\mathbf{findTable} : Label \times Database \rightarrow Table  \\
+	findTable(l, d) = \inference{\exists t \in tables(d) \wedge name(t) = l}{t}  \\ \\
+	l \in Label, d \lhd Database
+\end{align*}
+
 \paragraph{Column} columns define possible data values and types which can be part of a table record.
@@
 
-\paragraph{Foreign key}
+\paragraph{Foreign key} Foreign key is a reference to another table's primary key, it has an unique name and it can be constrained.
+The value of a foreign key is $\varnothing$ or a non-zero natural number.
 \begin{align*}
@@
 
-\paragraph{Primary key}
+\paragraph{Primary key} Primary key is unambiguous identifier of a record in a table.
+The primary key is always provided by the database as a non-zero natural number.
+Primary key is always defined with constraints $NOTNULL$ and $UNIQUE$.
 \begin{align*}
@@
 
-\paragraph{DbType} represents primitive types in the database.
+\paragraph{Data Types} Database data types $DbType$ represents primitive types in the database.
 There are usually defined types such as Varchar, Integer, Boolean etc.
-There are only two types defined in the model $DBSTRING$ as a default type of values in columns and $DBINT$ as a default type for keys.
+There is only one type defined in the model $DBSTRING$ as a default type of values in columns.
 \begin{align*}
-	\mathbf{DbType} = DBSTRING \; | \; DBINT 
+	\mathbf{DbType} = DBSTRING
 \end{align*}
 
-\paragraph{Constraint} there are two types of constraints defined in the model.
+\paragraph{Constraints} There are two types of constraints defined in the model.
 Both constraints are column constraints - first constraint defines unique records in a column, second constraint defines non-empty columns.
@@
 
-\subsubsection{Model of Stored Data}
-A database consists not only from schema but also from data which are represented as rows in a table.
-A table row in our model is a tuple consisting of reference to a concrete table and a set of value pairs, which represents concrete values of concrete columns.
+\paragraph{Stored Data} A database consists not only from schema but also from data which are represented as rows in a table.
+A table row in our model is a tuple consisting of value pairs, which represents concrete values of concrete columns.
+Each row contains a name of a table it belongs to.
 \begin{align*}
@@
 \end{align*}
-There functions $contains$ verify if the table contains data in context of concrete database.
+The definition of data allows definition of function $contains$ which verify if the table contains data in context of concrete database and $instantiated$, which search for instances of a class with given name and returns true if there are rows for the table in the database.
 \begin{align*}
@@
 \end{align*}
+\begin{align*}
+	\mathbf{instantiated} : Class \times Software \rightarrow Boolean \\
+	instantiated(c, s) = \begin{cases} \inference{\exists t \in tables(database(s)) : name(t) = name(c) \\ \wedge contains(t, database(s))}{true} \\ \\
+  false
+ \end{cases} \\ \\
+	 c \in classes(application(s)), s \lhd Software 
+\end{align*}
+
+\subsubsection{Database consistency}
 Each row in a database has to refer to a table in database schema and has a structure corresponding to the one defined by a table, otherwise the database is inconsistent.
@@
 \end{align*}
-
-\subsection{Extension of Software Definition}
-The database stores instances of the application, thus the $instantiated$ function has to be defined:
-
-$$instantiated : Class \times Software \rightarrow Boolean $$ 
-\begin{equation*}
-\begin{gathered}
-	instantiated(c, s) = \begin{cases}
- \inference{\exists t \in tables(database(s)) : name(t) = name(c) \\ \wedge contains(t, database(s))}{true} \\ \\
-  \inference{otherwise}{false} \\ \\
- \end{cases} \\ \\
- c \in classes(application(s)), s \in Software 
-\end{gathered}
-\end{equation*}
 
@@
 \Psi(s, newClass(c)) = application(s) \circ addClass(c) \\
-\Phi(s, newClass(c)) = database(s)  \bullet addTable(\rho(c,d), d) \\ \bullet addColumn(name(c), \{\forall r \in singleProperties(c) | \rho(r,d)\}, d) \\ \bullet addTable(name(c), \{\forall r \in collectionProperties(c) | \rho(r,d)\}, d)
+\Phi(s, newClass(c)) = database(s)  \bullet addTable(\rho(c,d), d) \\ \bullet addColumn(name(c), \{\forall r \in singleProperties(c) | \rho(r,d)\}, d) \\ \bullet addTable(name(c), \{\forall r \in collections(c) | \rho(r,d)\}, d)
 \end{align*}
@@
 %\begin{gathered}
-% addTable(\rho(c, d),d) \\ \wedge t \rho(singleProperties(c)) \wedge \rho(collectionProperties(c)) \\ 
+% addTable(\rho(c, d),d) \\ \wedge t \rho(singleProperties(c)) \wedge \rho(collections(c)) \\ 
 %\end{gathered}

### 2013-01-09 doc/description.tex
improved description of ORM

@@
 \label{sec:orm}
-There are defined functions for object-relational mapping used in the model in this section.
-The purpose of a function is indicated by its name and should be obvious from its definition.
-These functions do not serve for full object-relational mapping, their purpose is to help with data evolution (see Section \ref{sec:eorm}) and its propagation from application to database.
-However the database can be obtained from an application model by using following algorithm:
+There are defined functions for object-relational mapping used in this paper.
+The mapping function itself are not able to create a database from the application as they are mentioned to be used together with evolutionary transformations introduced in section \ref{sec:transformations}.
 
-\begin{algorithmic}[1]
-	\Require $a \in Application$
-	\State $d \lhd Database$
-	\ForAll{$c \in classes(a)$} 
-		\State $\rho_c(c), d$
-		\ForAll{$p \in properties(c)$} 
-			\State $\rho_p(p, d)$
-		\EndFor
-	 \EndFor
-	\ForAll{$c \in classes(a)$}
-		\ForAll{$as \in associations(c)$} 
-			\State $\rho_{a}(as, d)$
-		\EndFor
-	\EndFor
-\end{algorithmic}
-
-\subsubsection{Mapping of Types}
-The mapping of types assumes a bijection between application types and database types, otherwise these has to be additional information for type mapping.
-We focus on structural and data change therefore we simplify types and its mapping.
+\paragraph{Mapping of Types} The mapping of types assumes a bijection between application types and database types, otherwise these has to be additional information for type mapping.
+We focus on structural and data change therefore there is only one type in both our models.
+The mapping is then streightforward:
 \begin{align*}
@@
 
-\subsection{Mapping of Classes}
-We assume the primary key column is created automatically by database, therefore the primary key is always created with name Id and of the $\mathbb{N^{*}}$ type and properties with cardinality 0 or 1 are mapped into columns.
-The associations are ignored int this mapping as it is one of the partial mapping function used in eORM.
-
+\paragraph{Mapping of Classes} The class is mapped into an table with  automatically created primary key.
+The primary key is always created with name Id and of the $\mathbb{N^{*}}$ type.
+The properties and associations are mapped separately.
 \begin{align*}
@@
 \subsubsection{Mapping of Properties}
-There are two types of properties - the first type has cardinality 1  and it is mapped to a column.
-A property with cardinality greater than one has to be mapped int a table not a column, thus we have a special mapping case.
+There are two kinds of properties - the first kind has cardinality equal to one and it is mapped to a column; second kind - a property with cardinality greater than one has to be mapped to a table, thus there are two mapping cases.
 When the property is mandatory the $NOTNULL$ constraint is added to the column with property value.
-This constraint affect already stored instances a new mandatory value has to be added to each instance.
-
 \begin{align*}
@@
 
-\subsubsection{Mapping of Associations}
-A association is mapped into database as a foreign key in table representing associating class or an association can be mapped as a coupling table in case the association cardinality is M:N or 1:N.
-
+\paragraph{Mapping of Associations} An association is mapped into database as a foreign key in table representing associating class or as a table in case the association cardinality is one to many or many to many.
 \begin{align*}
@@
 
-\section{Model Manipulating Transformation}
+\section{Transformation for Models Manipulation}
+The evolution affects both parts of software and as shown in section \ref{sec:evoIntro} it is mapped to transformation specific for application (set $AppOps$ of transformations) and database (set $DbOps$).
+These sets are defined in this section.
 \subsection{Application Transformation}
+This section provide transformations for manipulating the application layer of a software - the $AppOps$ set. The set contains transformations for model creation, alternation and decomposition.
 \paragraph{Add Property}Adding a property into an application structure has to verify a name collision only:

### 2013-01-09 doc/description.tex
Improved application helpers

--- 2013-01-09 doc/description.tex
@@
 
-\section{Model Manipulating Transformation}
+\section{Transformation for Models Manipulation}
+The evolution affects both parts of software and as shown in section \ref{sec:evoIntro} it is mapped to transformation specific for application (set $AppOps$ of transformations) and database (set $DbOps$).
+These sets are defined in this section.
 \subsection{Application Transformation}
-\paragraph{Add Property}Adding a property into an application structure has to verify a name collision only:
+This section provide transformations for manipulating the application layer of a software - the $AppOps$ set. The set contains transformations for model creation, alternation and decomposition.
+
+\paragraph{Add Property}Before adding a property into an application structure, the name collision only has to be verified:
 \begin{align*}
@@
 \end{align*}
-\paragraph{Add Association} Adding an association into an application structure has to verify if there is no name collision and if there is the associated class present in the model.
+
+\paragraph{Add Association} Adding an association into an application structure has to verify if there is no name collision and if there is the associated class present in the application model.
 \begin{align*}
@@
 \end{align*}
+
 \paragraph{Add Class} Adding a class is limited to the case when the class has no associations, this is because it simplifies the database migration and because it is a natural way of modeling.
@@
 \end{align*}
-\paragraph{Remove Property}
+
+\paragraph{Remove Property or Association} The property or association can be simply removed from a class:
 \begin{align*}
@@
 \end{align*}
-\paragraph{Remove Association}
 \begin{align*}
@@
 \end{align*}
+
 \paragraph{Remove Class} The class can be removed only in case it is not referenced by any association existing in the application.
@@
 \end{align*}
-\paragraph{Rename Property}
+
+\paragraph{Rename Property and Association} Both renaming operations can change a name of the property, association respectively if there are no name collisions:
 \begin{align*}
@@
 \end{align*}
-\paragraph{Rename Association}
 \begin{align*}
@@
 \end{align*}
-\paragraph{Rename Class}
+
+\paragraph{Rename Class} When a class is renamed, the name collisions has to be verified and then all referencing associations has to be updated.
 \begin{align*}
@@
 	}{\begin{gathered}
-		c.name = l_n \wedge \forall e \in classes(a) \\ \wedge \forall (f \in associations(e) \wedge reference(f) = l_o ) \\ : f.reference = l_n 
+		c.name = l_n \wedge \forall e \in classes(a) \\ \wedge \forall (f \in associations(e) \wedge reference(f) = l_o ) :\\ f.reference = l_n 
 	\end{gathered}}

### 2013-01-09 doc/description.tex
Improvements in data transformations

@@
 \begin{align*}
+	\mathbf{tableData} : Label \times Row* \rightarrow Row* \\
+	tableData(l, r) = \begin{cases}
+ 		\inference{r = \varnothing}{\varnothing} \\
+ 		\inference{|r| > 1}{tableData(head(r)) \cup tableData(tail(r))}\\ \\
+ 		\inference{|r| = 1 \wedge refTable(r) = l}{r}\\\\
+ 		\inference{|r| = 1 \wedge refTable(r) \neq l}{\varnothing}
+ \end{cases} \\ \\
+	r \in Rows, l \in Label
+\end{align*}
+\begin{align*}
 	\mathbf{instantiated} : Class \times Software \rightarrow Boolean \\
@@
 \subsection{Database Transformation}
-The database is composed from two separated parts - database schema and data. Several transformation are defined to make manipulating a database easily in this paper.
-\paragraph{Add Column}
+The database is composed from two separated parts - database schema and data. Several transformation are defined to make manipulating a database easily.
+\paragraph{Add Column} A new column can be add into a table if there are no name collisions.
+If the column has the $NOTNULL$ constraint, then the default value has to be added to all rows in the table.
 \begin{align*}
@@
 		\inference{|c| > 1}{addColumn(l, head(c), d) \bullet addColumn(l, tail(c), d)} \\ \\
-		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c)}{columns(t) \cup c}\\ \\
+		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c) \\ \wedge NOTNULL \not \in constraints(c) }{columns(t) \cup c}\\ \\
+		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c) \\ \wedge NOTNULL \in constraints(c) }{columns(t) \cup c}\\ \bullet appendValues(l, [name(c), defVal(c)]) \\ \\
 		Inconsistent \;d 
@@
 \end{align*}
-\paragraph{COPY COLUMN}
+
+\paragraph{Copy Column} The transformation creates a copy of a column in a another table in a database. The transformation changes only the structure of database - stored data are not changed.
 \begin{align*}
@@
 	copyColumn(l_s, l_t, l_c, d) = \begin{cases}
- 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{columns(e) \cup c }
+ 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{columns(e) \cup c}
 	\end{cases} \\
-	l_s,l_t,l_c \in Label,d \lhd Database
+	l_s,l_t,l_c \in Label \wedge l_s \neq l_t,d \lhd Database
 \end{align*}
-\paragraph{Add Table}
+\paragraph{Add Table} The transformation creates a new table in the database. The new table should not contain any foreign keys.
 \begin{align*}
@@
 \\ \\	
-		\inference{|t| = 1 \wedge \forall e \in tables(d) \wedge name(t) \neq name(e)}{tables(d) \cup c}
+		\inference{|t| = 1 \wedge \forall e \in tables(d) \wedge name(t) \neq name(e) \\ \wedge foreignKeys(t) = \varnothing }{tables(d) \cup t}
 \\ \\
@@
 \end{align*}
-\paragraph{Insert row}
+\paragraph{Insert row} The function allow to insert a row(s) into a database.
 \begin{align*}
@@
 			\inference{|r|>1}{insertRow(head(r), d) \bullet insertRow(tail(r), d)} \\ \\
-		 	\inference{|r|=1 \wedge \exists t \in tables(d) : name(t) = reference(r) }{data(d) \cup r} \\ \\
+		 	\inference{|r|=1 \wedge \exists t \in tables(d) : name(t) = reference(r) \\ \wedge \forall e \in tableData(t, d) : pkValue(e) \neq pkValue(r)}{data(d) \cup r} \\ \\
 			 Inconsistent \; d
@@
 \end{align*}
-\paragraph{Insert Values}
+\paragraph{Append Values} The function appends a pair to all rows of a table.
 \begin{align*}
-	\mathbf{insertValues} : Label \times Pair* \times Database \rightarrow Daatabase \\
-	insertValues(l, p, d) = \begin{cases}
- 		\inference{|p| > 1}{insertValues(l, head(p), d)\bullet insertValues(l, tail(p), d)}  \\ \\
- 		\inference{|p| = 1 \wedge \forall e \in pairs(r_t) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r_t), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p} \\ \\
+	\mathbf{appendValues} : Label \times Pair \times Database \rightarrow Database \\
+	appemdValues(l, p, d) = \begin{cases}
+ 		\inference{\exists t \in tables(d) : name(t) = l, \\ r = tableData(l, d)}{insertValue(r, p, d)} \\ \\
+ 		Inconsistent \; d
+ 	\end{cases}\\ \\
+ 	l \in Label, d \lhd Database
+\end{align*}
+\paragraph{Insert Value} The function inserts a new pair into an existing row.
+\begin{align*}
+	\mathbf{insertValues} : Row \times Pair* \times Database \rightarrow Daatabase \\
+	insertValues(r, p, d) = \begin{cases}
+ 		\inference{|p| > 1}{insertValues(r, head(p), d)\bullet insertValues(r, tail(p), d)}  \\ \\
+ 		\inference{|p| = 1 \wedge \forall e \in pairs(r) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p} \\ \\
 	 	 Inconsistent \; d
  		\end{cases} \\
-l \in Label, p = Pair*,d \lhd Databases
+r \lhd Row, p = Pair*,d \lhd Databases
 \end{align*}
@@
 \begin{align*}
-	\mathbf{CopyProperty} : Label \times Label \times Label \times (Instance \times Instance) \times State \rightarrow State \\
+	\mathbf{CopyProperty} : Label \times Label \times Label \times (Row \times Row) \times State \rightarrow State \\
 	\Psi(CopyProperty(l_s, l_t, l_p, f_m), s) = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)} \\\\

### 2013-01-14 doc/description.tex
copyValue, copyRow

@@
 \subsection{Database Transformation}
-The database is composed from two separated parts - database schema and data. Several transformation are defined to make manipulating a database easily.
+The database is composed from two separated parts - database schema and data.
+Transformations are defined to manipulate the database schema and to manipulate stored data.
+These transformation were created with respect to evolutionary transformation, thus they ....
+
 \paragraph{Add Column} A new column can be add into a table if there are no name collisions.
@@
 		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c) \\ \wedge NOTNULL \not \in constraints(c) }{columns(t) \cup c}\\ \\
-		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c) \\ \wedge NOTNULL \in constraints(c) }{columns(t) \cup c}\\ \bullet appendValues(l, [name(c), defVal(c)]) \\ \\
+		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c) \\ \wedge NOTNULL \in constraints(c) }{columns(t) \cup c}\\ \bullet addValueToTable(l, [name(c), defVal(c)]) \\ \\
 		Inconsistent \;d 
@@
 \end{align*}
+
+\paragraph{Alter Column Name}
+\begin{align*}
+\mathbf{alterColumn} : Label \times Label \times Label \times Database \rightarrow Database \\
+	alterColumn(l_c, l_o, l_n, d) =
+ \inference{\exists t \in tables(a) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_o \\ \wedge \forall e \in columns(t) : name(e) \neq l_n}{\begin{gathered}
+c.name = l_n 
+\end{gathered}}
+ \\
+l_c, l_o, l_n \in Label, d \lhd Database
+\end{align*}
+
 \paragraph{Add Table} The transformation creates a new table in the database. The new table should not contain any foreign keys.
@@
 \end{align*}
-\paragraph{Append Values} The function appends a pair to all rows of a table.
+\paragraph{Add Value To Table} The function appends a pair to all rows of a table.
 \begin{align*}
-	\mathbf{appendValues} : Label \times Pair \times Database \rightarrow Database \\
-	appemdValues(l, p, d) = \begin{cases}
- 		\inference{\exists t \in tables(d) : name(t) = l, \\ r = tableData(l, d)}{insertValue(r, p, d)} \\ \\
+	\mathbf{addValueToTable} : Label \times Pair \times Database \rightarrow Database \\
+	addValueToTable(l, p, d) = \begin{cases}
+ 		\inference{\exists t \in tables(d) : name(t) = l}	{\begin{gathered}
+		insertValue(head(tableData(l, d)), p, d) \\ \bullet addValueToTable(tail(tableData(l, d)), p, d)
+	\end{gathered}} \\ \\
  		Inconsistent \; d
  	\end{cases}\\ \\
- 	l \in Label, d \lhd Database
+ 	l \in Label, p \lhd Pair,  d \lhd Database
 \end{align*}
+
 \paragraph{Insert Value} The function inserts a new pair into an existing row.
@@
 \end{align*}
-\paragraph{COPY VALUES}
+\paragraph{Copy Value} The function copy values from one column in a table to another column in the same or in another table - according to mapping function $f_m$. 
 \begin{align*}
-	\mathbf{copyValue} : Label \times Label \times Label \times (Instancece \times Instance) \times Database \times Database \\
+	\mathbf{copyValue} : Label \times Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
 	copyValue(l_s, l_t, l_{sc}, l_{tc} f_m, d) = \begin{cases}
- 		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc} }{\begin{gathered}
-			\forall p \in f_m(t_s, t_t)  : \\ insertValues(name(second(p)), \\ [l_{tc},value(pairByRef(l_{sc}, first(p))),d)
+		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc}}{\begin{gathered}
+			insertValues(second(p), \\ [l_{tc},value(pairByRef(l_{sc}, first(p)))],d)
 		\end{gathered}}
@@
 \end{align*}
+
+\paragraph{Copy rows}
+\begin{align*}
+	\mathbf{copyRows} : Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
+	copyRows(l_s, l_t, f_m, d) = \begin{cases}
+ 	\inference{|f_m| = \varnothing}{d} \\\\
+ 	\inference{|f_m| = 1 \\ \wedge \exists t_s, t_t \in tables(d) : name(t_s) = l_s, name(t_t) = l_t \\ \wedge t_s \cong t_t }{data(d) \cup insertRow(changeRef(l_t, first(f_m)), d)} \\\\
+ 	\inference{|f_m| > 1}{\begin{gathered}
+		 d \bullet copyRows(l_s, l_t, head(f_m)) \bullet copyRows(l_s, l_t, tail(f_m))
+	\end{gathered}}
+  \end{cases}
+\end{align*}
+
+%\begin{align*}
+%	v(l, fm, d) = \begin{cases}
+%	\inference{fm = \varnothing}{d}\\ \\
+%	\inference{|fm| > 1}{v(l, head(fm)) \bullet v(l, tail(fm))}\\ \\
+%	\inference{|fm| = 1 \exists c \in pairs(first(fm)) : reference(c) = l \\ \wedge \exists e \in pairs(second(fm)) : reference(e) = l}{e \cup pairByRef(l,c)} 
+% \end{cases}
+%\end{align*}
 
@@
 \\ \\ 
-\inference{ \wedge mandatory(p) \vee instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered} addColumn(l, \rho(p,d), d) \\ \bullet \forall r \in data(d), reference(r) = name(t) : \\ appendValue(name(t), [name(p), defaultValue(p)], d)
+\inference{ \wedge mandatory(p) \vee instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered} addColumn(l, \rho(p,d), d) \\ \bullet \forall r \in data(d), reference(r) = name(t) : \\ addValueToTable(name(t), [name(p), defaultValue(p)], d)
 \end{gathered}
@@
 \end{align*}
-
 \begin{align*}
@@
 \Psi(RenameProperty(l_c, l_o, l_n, s)) = application(s) \circ renProp(l_c, l_o, l_n) \\
-\Phi(RenameProperty(l_c, l_o, l_n, s)) = database(s) \bullet alterColumn(l_c, l_o, l_n) \\\\
-\end{align*}
-
-\begin{align*}
-alterColumn : Label \times Label \times Label \times Database \rightarrow Database \\
-	alterColumn(l_c, l_o, l_n, d) = \begin{cases}
- \inference{\exists t \in tables(a) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_o \\ \wedge \forall e \in columns(t) : name(e) \neq l_n}{\begin{gathered}
-c.name = l_n 
-\end{gathered}}
-\\ \\
-\inference{\exists t \in tables(a) : name(t) = l_o \\ \wedge \forall e \in tables(t) : name(e) \neq l_n}{\begin{gathered}
-t.name = l_n 
-\end{gathered}}
- \end{cases}
- \\
-l_c, l_o, l_n \in Label, a \lhd Applications, s \in States, d \lhd Database
+\Phi(RenameProperty(l_c, l_o, l_n, s)) = \begin{cases}
+ 	\inference{cardinality(propertyByName(l_c, l_o, application(s))) \leq 1}{database(s) \bullet alterColumn(l_c, l_o, l_n))} \\\\
+ 	\inference{cardinality(propertyByName(l_c, l_o, application(s))) > 1}{alterTable(l_o, l_n, database(s))}
+ \end{cases}\\
+l_c, l_o, l_n \in Label,  s \lhd Software
 \end{align*}
@@
 \Phi(RenameAssocaiation(l_c, l_o, l_n), s) = \begin{cases}
-	\inference{\mathbf{TODO}}{database(s) \bullet alterTable(l_c, l_o, l_n)}
- \\
-  \inference{\mathbf{TODO}}{database(s) \bullet alterColumn(l_c, l_o, l_n)}
+	\inference{\exists a  \in getAssocByname(application(s)) : \\
+  name(a) = l_c \\ \wedge startCardinality(a) > 1 }{database(s) \bullet alterTable(l_c, l_o, l_n)}
+ \\ \\
+  \inference{\exists a  \in getAssocByname(application(s)) : \\
+  name(a) = l_c \\ \wedge startCardinality(a) \leq 1 }{\begin{gathered}
+database(s) \\ \bullet alterForeignKey(endClass(a) ,l_c, l_o, l_n)
+\end{gathered}}
  \end{cases}
 \\\\
-alterColumn : Label \times Label \times Label \times Database \rightarrow Database \\
 %	alterAssoc(l_c, l_o, l_n, d) = \begin{cases}
@@
 
+\section{Refactoring Cases}
+\subsection{Move Property}
+\subsection{Extract and Inline Class}
+
 \section{Conclusion}

### 2013-01-14 doc/description.tex
change reference of a row
getAssocByname, propertyByName

@@
 
+\paragraph{Get Property or Association}
+\begin{align*}
+	\mathbf{propertyByName} : Label \times Label \times Application \rightarrow Property \\
+	propertyByName(l_c, l_p, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{p} \\
+	l_c, l_p \in Label, a \lhd Application
+\end{align*}
+\begin{align*}
+	\mathbf{associationByName} : Label \times Label \times Application \rightarrow Association \\
+	associationByName(l_c, l_a, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists o \in associations(c) : name(o) = l_a}{o} \\
+	l_c, l_a \in Label, a \lhd Application
+\end{align*}
+
 \paragraph{Remove Property or Association} The property or association can be simply removed from a class:
@@
 		\end{gathered}}
- 	\end{cases}
+ 	\end{cases} \\
+ 	l_s, l_t, l_{sc}, l_{tc} \in Label, d \lhd Database
 \end{align*}
@@
 	\end{gathered}}
-  \end{cases}
+  \end{cases} \\
+  l_s, l_t \in Label, d \lhd Database
+\end{align*}
+
+\paragraph{Change ref}
+\begin{align*}
+	\mathbf{changeRef} : Label \times Row \rightarrow Row \\
+	changeRef(l, r) = \inference{}{reference(r) = l } \\
+	l \in Label, r \lhd Row
 \end{align*}
@@
 \Phi(RenameAssocaiation(l_c, l_o, l_n), s) = \begin{cases}
-	\inference{\exists a  \in getAssocByname(application(s)) : \\
+	\inference{\exists a  \in associationByName(application(s)) : \\
   name(a) = l_c \\ \wedge startCardinality(a) > 1 }{database(s) \bullet alterTable(l_c, l_o, l_n)}
  \\ \\
-  \inference{\exists a  \in getAssocByname(application(s)) : \\
+  \inference{\exists a  \in associationByName(application(s)) : \\
   name(a) = l_c \\ \wedge startCardinality(a) \leq 1 }{\begin{gathered}

### 2013-01-14 doc/description.tex
\cong

@@
 l \in Label, pk \lhd PrimaryKey, cols = Column*, fks = ForeignKey*
-\end{align*}
-
-The function $findTable$ finds and returns a table in the database according to the provided label.
-\begin{align*}
-	\mathbf{findTable} : Label \times Database \rightarrow Table  \\
-	findTable(l, d) = \inference{\exists t \in tables(d) \wedge name(t) = l}{t}  \\ \\
-	l \in Label, d \lhd Database
 \end{align*}
@@
 \subsection{Database Transformation}
-The database is composed from two separated parts - database schema and data.
-Transformations are defined to manipulate the database schema and to manipulate stored data.
-These transformation were created with respect to evolutionary transformation, thus they ....
+The database is composed from two separated parts - database schema and data. Transformations are defined to manipulate the database schema and to manipulate stored data.
+
+\paragraph{Find Table} The function $findTable$ finds and returns a table in the database according to the provided label.
+\begin{align*}
+	\mathbf{findTable} : Label \times Database \rightarrow Table  \\
+	findTable(l, d) = \inference{\exists t \in tables(d) \wedge name(t) = l}{t}  \\ \\
+	l \in Label, d \lhd Database
+\end{align*}
+
+\paragraph{Compare Tables} The need to compare two tables during the evolution results into the comparing operator $\cong$. The operator has to be defined for columns and foreign keys too.
+\begin{align*}
+	\mathbf{\cong} : Table \times Table \rightarrow Boolean \\
+	t_1 \cong t_2 = \begin{cases}
+ 		\inference{|columns(t_1)| = |columns(t_2)| \\ \wedge |foreignKeys(t_1)| = |foreignKeys(t_2)| \\ \wedge \forall c \in columns(t_1) \exists e \in columns(t_2) : c = e \\ \wedge \forall f \in foreignKeys(t_1) \exists g \in foreignKeys(t_2) : f = g }{true}\\\\
+ false
+ \end{cases}\\
+ t_1 \lhd Table, t_2 \lhd Table
+\end{align*}
+\begin{align*}
+	\mathbf{\cong} : Column \times Column \rightarrow Boolean \\
+	c_1 \cong c_2 = \begin{cases}
+ 		\inference{name(c_1) = name(c_2) \\ \wedge type(c_1) = type(c_2) \\ \wedge constraints(c_1) = comstraints(c_2)}{true}
+ 		\\ \\
+		false
+ \end{cases}\\
+ c_1 \lhd Column, c_2 \lhd Column
+\end{align*}
+\begin{align*}
+	\mathbf{\cong} : ForeignKey \times ForeignKey \rightarrow Boolean \\
+	f_1 \cong f_2 = \begin{cases}
+ 		\inference{name(f_1) = name(f_2) \\ \wedge reference(f_1) = reference(f_2) \\ \wedge constraints(f_1) = comstraints(f_2}{true}
+ 		\\ \\
+		false
+ \end{cases}
+\end{align*}
 
@@
 
-\paragraph{Change Cardinality}
-$$\rho_{CcP} : Property \times \mathbb{N^{*}} \times Application \rightarrow Property $$
-$$\rho_{cardA} : Association \times \mathbb{N^{*}} \times Application \rightarrow Association $$
+%\paragraph{Change Cardinality}
+%$$\rho_{CcP} : Property \times \mathbb{N^{*}} \times Application \rightarrow Property $$
+%$$\rho_{cardA} : Association \times \mathbb{N^{*}} \times Application \rightarrow Association $$
 
@@
 \end{document}
-
-%	CopyProperty(l_s, l_t, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p \\ \wedge \exists e \in classes(a) : name(e) = l_t  \\ \wedge \forall q \in properties(c) : name(q) \neq l_p}{\begin{gathered}
-%properites(e) \cup p 
-%\end{gathered}}
-
-% This is important in case we have an evolution consisting of several evolution operations.
-These operations has to be proceed in concrete order and all of them has to change the software otherwise the software can become inconsistent from user point of view (the structure or stored data will not correspond with the users' reality).
-To assure the right behavior of evolution operations and to propagate operations fail to the user we extend the definition of a software to make it a monad.
-The extension consists in defining a software state and operations $ret$ and $\oplus$ (called evolve), which is used as a concatenation of operations as well.
-%
-%Each evolution transformation has a definition: $Software \rightarrow State$ and all evolution transformations creates a set called $Transformations$, then a concatenation of evolution transformation is defined as:
-%\begin{equation*}
-%\begin{gathered}
-% \oplus : State \times Transformations \rightarrow State  \\ \\
-% u \oplus t = \begin{cases}
-% u, \; u = Inconsistent \; s \\
-% t(u), \; u = Consistent \; s 
-%\end{cases}\\ 
-%u \in States, t \in Transformations, s \lhd Software
-%\end{gathered}
-%\end{equation*}
-%
-% To describe these two aspects od software evolution we decide to define transformations for each layer separately.
-The separation of transformation contexts help us to define transformation specific for application or database evolution, which can ease both - understanding and defining the transformation.
-The transformations specific for a context are called operations and they creates two sets- first one for application evolution: $AppOps$  and second one for database evolution $DbOps$.
-Applying of a transformation is defined:
-%\begin{align*}
-%u \oplus t = ret(<application(u) \oplus o_{app}, database(u) \oplus o_{db}>) \\
-%u \in Softwares, t \in Transformations, o_{app} \in AppOps, o_{db} \in DbOps
-%\end{align*}

### 2013-01-14 doc/description.tex
added pkValue

@@
 \begin{align*}
-	\mathbf{Row} = < refTable, pair* > \\
+	\mathbf{Row} = < refTable, pkValue, pair* > \\
 	\mathbf{row} : Label \times Pairs \rightarrow Row \\ \\
 	\mathbf{refTable} : Row \rightarrow Label \\
-	refTable(row(l, ps) = l \\ \\
+	refTable(row(l, n, ps) = l \\ \\
 	\mathbf{pairs} : Row \rightarrow Pair* \\
-	pairs(row(l, ps)) = ps \\ \\
-	\mathbf{pairByRef} : Label \times Row \rightarrow Pair \\
-	pairByRef(l,i) = \begin{cases}
-		\inference{\exists p \in pairs(i) : reference(p) = l}{p} \\ \\
-			\varnothing
-		\end{cases}\\
-	 l \in Label, ps = Pair*, i \lhd Row
+	pairs(row(l, n, ps)) = ps \\ \\
+	\mathbf{pkValue} : Row \rightarrow N \\
+	pkValue(row(l, n, ps)) = n \\ \\
+	 l \in Label, ps = Pair*, n \in N
 \end{align*}
-The value stored in concrete column of the row is represented by pair, where the $refCol$ is a name of the column, a foreign key or its primary key and $value$ represents concrete value .
+The value stored in concrete column of the row is represented by the pair, where the $refCol$ is a name of the column, a foreign key or its primary key and $value$ represents concrete value .
 \begin{align*}
@@
 
+\paragraph{Find pair}
+\begin{align*}
+\mathbf{findPair} : Label \times Row \rightarrow Pair \\
+	findPair(l, r) = \begin{cases}
+		\inference{\exists p \in pairs(r) : reference(p) = l}{p} \\ \\
+			\varnothing
+		\end{cases}\\
+		l \in Label, r \lhd Row
+\end{align*}
+
 \paragraph{Add Column} A new column can be add into a table if there are no name collisions.
@@
 		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc}}{\begin{gathered}
-			insertValues(second(p), \\ [l_{tc},value(pairByRef(l_{sc}, first(p)))],d)
+			insertValues(second(p), \\ [l_{tc},value(findPair(l_{sc}, first(p)))],d)
 		\end{gathered}}
@@
 %	\inference{|fm| > 1}{v(l, head(fm)) \bullet v(l, tail(fm))}\\ \\
-%	\inference{|fm| = 1 \exists c \in pairs(first(fm)) : reference(c) = l \\ \wedge \exists e \in pairs(second(fm)) : reference(e) = l}{e \cup pairByRef(l,c)} 
+%	\inference{|fm| = 1 \exists c \in pairs(first(fm)) : reference(c) = l \\ \wedge \exists e \in pairs(second(fm)) : reference(e) = l}{e \cup findPair(l,c)} 
 % \end{cases}

### 2013-01-15 doc/description.tex
added section headers
fixed typos and updated text structure

@@
 \end{figure}
-A software implemented in an object-oriented language, which uses relational database as a data storage is context of this paper.
-The situation is illustrated in the Figure \ref{fig:appStructure}.
+A software implemented using an object-oriented language, which uses relational database as a data storage (as illustrated in the Figure \ref{fig:appStructure}) is object of interest of this paper.
 There are three important components of such a software: application (its persistent layer concretely), database (consisting of database schema and stored data) and an ORM (we will use $\rho$ as a symbol for ORM), therefore we define a software as a triple:
@@
 
-The software has to be in consistent state so its users could have benefit from its useage. The state of the software is defined to communicate the sate of the software to its user:
+The software has to be in consistent state so its users could have benefit from its usage.
+The state of the software is defined to communicate the sate of the software to its user.
+The components of a software (application and database) are transformed likewise, thus the state is defined for them too.
 \begin{align*}
- 	\mathbf{state} = Consistent\; s \: | \: Inconsistent \; s, \\\\
+ 	\mathbf{state} = Consistent\; s \: | \: Inconsistent \; s,  \\	s \lhd Software \vee s \lh Application \vee s \lhd Database
+\end{align*}
+Before the definition of the software consistency is provided, the notation of function transcription has to be introduced.
+Usually the definition of a function consists of two parts - an effect of the function and conditions when the effect occurs We decide to use following transcription, to simplify the notation and to improve understanding of functions:
+\begin{align*}
+f(x) =   \inference{conditions}{effect}
+\end{align*}
+The function querying the state (it means consistency) of a software is defined as:
+\begin{align*}
+ 	\mathbf{consistent} : State \rightarrow Boolean \\
+ 	consistent(s) = \begin{cases}
+ 		\inference{s = Consistent(s)}{true}\\\\
+ 		\inference{s = Inconsistent(s)}{false}\\
+ \end{cases}\\
+ s \lhd State
+\end{align*}
+The consistency of application and database is discussed later in this paper. The definition of software state only is provided:
+\begin{align*}
 	\mathbf{state} : Software \rightarrow State \\
@@
  	\end{cases} \\ \\
- 	s \lhd Software  
 \end{align*}
+
 \subsection{Software Evolution}
@@
 The evolution of a software is a transformation from one consistent state to another one.
-These states are called generations of the software.
+These states are called generations of the software and the function which changed the state of a software are called transformations.
 The evolution consists of change in application and in database as shown in the Figure \ref{fig:evolution}.
-We focus on situations when the evolution is initiated on the application level (as code refactoring), however the evolution has to be processed on the database level too.
-It means there is a set of application transformations called $AppOps$ and a set of database transformations called $DbOps$.
-Because there are needed different parameters for application and database evolution the set of transformation called $E$ which evolves the whole software and a mapping between $E$ and $AppOps$, respectively $DbOps$ is defined.
-We call this mappings $\Psi$ and $\Phi$:
+We focus on situations when the evolution is initiated on the application level (as code refactoring), thus the transformation are designed to conform to the refactoring cases.
+However a case when the evolution begins on database level is also possible (eg.
+as a result of database optimization) it is not considered in this paper.
+
+The evolution of the whole software has to be processed on the both software components - application and database.
+It means there is a set of application transformations called $AppT$ and a set of database transformations called $DbT$.
+Because there are needed different parameters for application and database evolution the evolution of database can not be derived from application evolution only.
+Thus the set of transformation called $E$ is defined to evolve the whole software together with mappings $\Psi$ between $E$ and $AppT$ and $\Phi$ between $E$ and $DbT$ is defined:
 \begin{align*}
-\Psi : E \rightarrow AppOps \\
-\Phi : E \rightarrow DbOps
+\Psi : E \rightarrow AppT \\
+\Phi : E \rightarrow DbT
 \end{align*}
-A subset of $E$ represents an evolution transformation of a software, which mets the first goal of our framework - a single source of data evolution for the whole software.
+The set $E$ containing the evolution transformations of a software mets our first goal - a single source of data evolution for the whole software.
 
@@
 \begin{align*}
-\circ : Application \times AppOps \rightarrow Application \\
- \bullet : Database \times DbOps \rightarrow Database
+\circ : State \times AppT \rightarrow State \\
+ \bullet : State \times DbT \rightarrow State
 \end{align*}
@@
 in Haskell programming language) so the state of the software can be passed from one evolution step to another, which helps us to concatenate evolution transformations.
-
 \begin{align*}
- u \oplus t = \begin{cases}
- 	\inference{u = Inconsistent \; s}{u} \\ \\
- 	\inference{u = Consistent \; s}{state(sw(\Psi(u, t), \Phi(u,t))} 
+ state(s) \oplus t = \begin{cases}
+ 	\inference{! consistent(state(s))}{u} \\ \\
+ 	\inference{consistent(state(s))}{state(sw(\Psi(s, t), \Phi(s,t))} 
 \end{cases}\\ 
-u \lhd State, t \in E, s \lhd Software
+s \lhd Software, t \in E
 \end{align*}
 \begin{align*}
- a \circ o = \begin{cases}
- 	\inference{!\;consistent(a)}{a} \\ \\
- 	\inference{consistent(a)}{a \circ o} 
+ state(a) \circ e = \begin{cases}
+ 	\inference{!\;consistent(state(a))}{a} \\ \\
+ 	\inference{consistent(state(a))}{e(a)} 
 \end{cases}\\ 
-a \lhd Application, o \in AppOps
+a \lhd Application, e \in AppT
 \end{align*}
 \begin{align*}
- d \bullet o = \begin{cases}
- 	\inference{!\;consistent(d)}{d} \\ \\
- 	\inference{consistent(d)}{d \circ o} 
+ state(d) \bullet e = \begin{cases}
+ 	\inference{!\;consistent(state(d))}{d} \\ \\
+ 	\inference{consistent(state(d))}{e(d)} 
 \end{cases}\\ 
-d \lhd Application, o \in DbOps
+d \lhd Application, e \in DbT
 \end{align*}
@@
 \section{Software Model}
-The model of an ORM software consists of three parts - an application model, a database model and a ORM.
-The model of application and database creates a type system for both kinds ORM and evolutionary transformations.
+The model of an ORM software consists of three parts - an application model, a database model and an ORM.
+The model of application and database creates a type system for both ORM and evolutionary transformations.
 
-The definitions of all parts assume a set $Label$ exist which contains all possible identifiers and each modeled entity is identified unambiguously identified in the model.
+The definitions of all parts assume a set $Label$ exists which contains all possible identifiers and each modeled entity is unambiguously identified in the model.
 Next there is a set $Values$ containing all possible values, which can be used in the model.
@@
 \label{sec:appModel}
-The application model represents a persistent layer of an application created using object oriented language and transformations which can alter the model. 
+The application model represents an application persistent layer created using object oriented language and functions which can alter the model. 
 \subsubsection{Application Structure}
@@
 	properties(cl(l, props, assocs)) = props \\ \\
-	\mathbf{associations} : Class \rightarrow Association* \\	associations(cl(l, props, assocs)) = assocs \\ \\
+	\mathbf{associations} : Class \rightarrow Association* \\	associations(cl(l, props, assocs)) = assocs \\ 
+	l \in Label, props \lhd Property*, assocs \lhd Association*
 \end{align*}
-Three more function are defined to obtain information about a class - $primitives$ returns all properties of the class which represent a single value, whereas $collections$ returns all properties representing collections.
-The last one $associated$ determine if a class is associated by any other class in an application.
+Three more function are defined to obtain information about a class - function $primitives$ returns all properties of the class which represent a single value, whereas function $collections$ returns all properties representing collection of values.
+The last function called $associated$ determine if a class is associated by any other class in an application.
 \begin{align*}
@@
 \begin{align*}
-	\mathbf{Property} = <label, AppType, DefaultValue, Cardinality, Mandatory> \\ \\
+	\mathbf{Property} = <label, AppType, DefaultValue, Cardinality, Mandatory> \\
 	\mathbf{prop} : Label \times AppType \times String \times \mathbb{N^{*}} \times Boolean \rightarrow Property \\ \\
@@
 
-\paragraph {Association} Assoctation represents a connection between two classes.
+\paragraph {Association} Associtation represents a connection between two classes.
 It has a unique name and the reference is represented by the label of referenced class.
 The class which owns the association is consider to be the starting class of an association, referenced class is consider to be the ending class of an association.
-The cardinalities defines the multiplicity of the association ends.
+The cardinalities defines the multiplicity of both association ends.
 \begin{align*}
-	\mathbf{Association} = <label, classRef, startCardinality, endCardinality> \\ \\
+	\mathbf{Association} = <label, classRef, startCardinality, endCardinality> \\
 	\mathbf{assoc} : Label \times Label \times \mathbb{N} \times \mathbb{N} \rightarrow Association\\ \\
@@
 in contrast there is only one type in our model, because we focus on structural and data changes and type casting operations are not important for us.
-The only App-Type type is called $APPSTRING$.
+The only $AppType$ type is called $APPSTRING$.
 \begin{align*}
@@
 
+\subsubsection{Application consistency}
+
 \subsection{Database Model}
 \label{sec:dbModel}
-The relational database consists of database schema which defines the structure of the database and data, which in our ORM system represents instances. The database is:
+The relational database consists of database schema which defines the structure of the database and data, which in our ORM system represents stored instances. The database is defined as a tuple:
 \begin{align*}
-	\mathbf{Database} = < Table*, Row*> \\ \\
+	\mathbf{Database} = < Table*, Row*> \\
 	\mathbf{db} : Tables \times Rows \rightarrow Database \\ \\
@@
 \end{align*}
-\paragraph{Table} represents a basic concept of database schema.
-It has a unique name, one or more column and it can be related to other tables in the schema by foreign keys.
+\paragraph{Table} Table represents a basic concept of database schema.
+It has a unique name, one or more columns and it can be related to other tables in the schema by foreign keys.
 Rows in the table represents stored data.
 \begin{align*}
-	\mathbf{Table} = <label, primaryKey, Column*, ForeignKey*>\\ \\
+	\mathbf{Table} = <label, primaryKey, Column*, ForeignKey*>\\
 	\mathbf{tab} 	: Label \times PrimaryKeys \times Columns \times ForeignKeys \rightarrow Table \\ \\
@@
 
-\paragraph{Column} columns define possible data values and types which can be part of a table record.
+\paragraph{Column} Column defines data values and types which can be part of a table record.
 \begin{align*}
-	\mathbf{Column} = <label, DbType, DefaultValue, Constraint*> \\ \\
+	\mathbf{Column} = <label, DbType, DefaultValue, Constraint*> \\
 	\mathbf{col} : Label \times DbType \times String \times Constraints \rightarrow Column \\ \\
@@
 	owningTable(c, d) = \inference{\exists tab \in tables(d) \wedge c \in columns(tab)}{tab} \\ \\
-l \in Label, t \lhd DbType, val \in Values,\\ cons \lhd Constraints, c \lhd Column,d \lhd Database
+l \in Label, t \lhd DbType, val \in Values,\\ cons \lhd Constraint*, c \lhd Column,d \lhd Database
 \end{align*}
@@
 \begin{align*}
-	\mathbf{ForeignKey} = <label, tableRef, Constraint*> \\ \\
+	\mathbf{ForeignKey} = <label, tableRef, Constraint*> \\ 
 	\mathbf{fk} : Label \times Label \times Constraints \rightarrow ForeignKey \\ \\
@@
 \paragraph{Primary key} Primary key is unambiguous identifier of a record in a table.
-The primary key is always provided by the database as a non-zero natural number.
+The primary key is always provided (automatically generated) by the database as a non-zero natural number.
 Primary key is always defined with constraints $NOTNULL$ and $UNIQUE$.
@@
 There are usually defined types such as Varchar, Integer, Boolean etc.
-There is only one type defined in the model $DBSTRING$ as a default type of values in columns.
+There is only one type of column values defined in the model called $DBSTRING$.
+Next the $DBN$ type is defined for the values of keys.
 \begin{align*}
-	\mathbf{DbType} = DBSTRING
+	\mathbf{DbType} = DBSTRING | DBN
 \end{align*}
@@
 \paragraph{Constraints} There are two types of constraints defined in the model.
-Both constraints are column constraints - first constraint defines unique records in a column, second constraint defines non-empty columns.
+Both constraints are column constraints - first constraint defines non-empty columns, second constraint defines there has to be unique records in a column or foreign key.
 \begin{align*}
@@
 \paragraph{Stored Data} A database consists not only from schema but also from data which are represented as rows in a table.
-A table row in our model is a tuple consisting of value pairs, which represents concrete values of concrete columns.
+A table row in our model is a tuple consisting of value pairs, which represents concrete value of concrete column.
 Each row contains a name of a table it belongs to.
@@
 \end{align*}
-The value stored in concrete column of the row is represented by the pair, where the $refCol$ is a name of the column, a foreign key or its primary key and $value$ represents concrete value .
+The value stored in concrete column of the row is represented by a pair, where the $refCol$ is a name of the column, a foreign key or its primary key and $value$ represents concrete value .
 \begin{align*}
@@
 	reference([l,val]) = l \\ \\
-	\mathbf{value} : Pair \rightarrow Label \\ 
+	\mathbf{value} : Pair \rightarrow Value \\ 
 	value([l,val]) = val \\ \\
@@
  \end{cases} \\ \\
-	r \in Rows, l \in Label
+	r \lhd Row*, l \in Label
 \end{align*}
@@
 \subsubsection{Database consistency}
-Each row in a database has to refer to a table in database schema and has a structure corresponding to the one defined by a table, otherwise the database is inconsistent.
+The database consistency is founded on the consistency of structure and consistency between structure and stored values.
+Each row in a database has to refer to a table in database schema and it has to have the structure corresponding to the one defined by a table, otherwise the database is inconsistent.
 The consistency of database is then:
@@
 
-\paragraph{Mapping of Types} The mapping of types assumes a bijection between application types and database types, otherwise these has to be additional information for type mapping.
-We focus on structural and data change therefore there is only one type in both our models.
-The mapping is then streightforward:
+\paragraph{Mapping of Types} The mapping of types assumes a bijection between application types and database types, otherwise there has to be additional information for type mapping.
+We focus on structural and data change therefore there is only one type of values in both our models.
+The mapping is then straightforward:
 \begin{align*}
@@
 
-\paragraph{Mapping of Classes} The class is mapped into an table with  automatically created primary key.
+\paragraph{Mapping of Classes} The class is mapped into a table with  automatically created primary key.
 The primary key is always created with name Id and of the $\mathbb{N^{*}}$ type.
-The properties and associations are mapped separately.
+The properties and associations has to be mapped separately.
 \begin{align*}
@@
 \subsubsection{Mapping of Properties}
-There are two kinds of properties - the first kind has cardinality equal to one and it is mapped to a column; second kind - a property with cardinality greater than one has to be mapped to a table, thus there are two mapping cases.
+There are two kinds of properties - primitive property with cardinality equal to one is mapped to a column; second kind - collection with cardinality greater than one is mapped to table, thus there are two mapping cases.
 When the property is mandatory the $NOTNULL$ constraint is added to the column with property value.
@@
 		\end{gathered}}
- 	\end{cases}
+	\end{cases}\\
+	p \lhd Property
 \end{align*}
@@
 
-\paragraph{Mapping of Associations} An association is mapped into database as a foreign key in table representing associating class or as a table in case the association cardinality is one to many or many to many.
+\paragraph{Mapping of Associations} An association is mapped into database as a foreign key in table representing associated class or as a table in case the association cardinality is one to many or many to many.
+Thus there are two mapping cases as for properties.
 \begin{align*}
@@
 \section{Transformation for Models Manipulation}
-The evolution affects both parts of software and as shown in section \ref{sec:evoIntro} it is mapped to transformation specific for application (set $AppOps$ of transformations) and database (set $DbOps$).
+The evolution affects both parts of software and as shown in section \ref{sec:evoIntro} it is mapped to transformation specific for application (set $AppT$ of transformations) and database (set $DbT$).
 These sets are defined in this section.
 \subsection{Application Transformation}
-This section provide transformations for manipulating the application layer of a software - the $AppOps$ set. The set contains transformations for model creation, alternation and decomposition.
+This section provide transformations for manipulating the application layer of a software - the $AppT$ set.
+The set contains not only transformations for model creation, alternation and decomposition, but some operations to obtain information about a model too.
 
-\paragraph{Add Property}Before adding a property into an application structure, the name collision only has to be verified:
+\paragraph{Get Property or Association}
+\begin{align*}
+	\mathbf{propertyByName} : Label \times Label \times Application \rightarrow Property \\
+	propertyByName(l_c, l_p, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{p} \\
+	l_c, l_p \in Label, a \lhd Application
+\end{align*}
+\begin{align*}
+	\mathbf{associationByName} : Label \times Label \times Application \rightarrow Association \\
+	associationByName(l_c, l_a, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists o \in associations(c) : name(o) = l_a}{o} \\
+	l_c, l_a \in Label, a \lhd Application
+\end{align*}
+
+\paragraph{Add Property} The name collision only has to be verified, before adding a property into an application structure:
 \begin{align*}
@@
 	c \lhd Class, a \lhd Applications
-\end{align*}
-
-\paragraph{Get Property or Association}
-\begin{align*}
-	\mathbf{propertyByName} : Label \times Label \times Application \rightarrow Property \\
-	propertyByName(l_c, l_p, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{p} \\
-	l_c, l_p \in Label, a \lhd Application
-\end{align*}
-\begin{align*}
-	\mathbf{associationByName} : Label \times Label \times Application \rightarrow Association \\
-	associationByName(l_c, l_a, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists o \in associations(c) : name(o) = l_a}{o} \\
-	l_c, l_a \in Label, a \lhd Application
 \end{align*}
@@
 
-\paragraph{Copy Column} The transformation creates a copy of a column in a another table in a database. The transformation changes only the structure of database - stored data are not changed.
-\begin{align*}
-	\mathbf{copyColumn} : Label \times Label \times Label \times Database \rightarrow Database \\
-	copyColumn(l_s, l_t, l_c, d) = \begin{cases}
- 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{columns(e) \cup c}
-	\end{cases} \\
-	l_s,l_t,l_c \in Label \wedge l_s \neq l_t,d \lhd Database
-\end{align*}
-
 \paragraph{Alter Column Name}
@@
 \end{align*}
+
+\paragraph{Copy Column} The transformation creates a copy of a column in a another table in the database. The transformation changes only the structure of database - stored data are not changed.
+\begin{align*}
+	\mathbf{copyColumn} : Label \times Label \times Label \times Database \rightarrow Database \\
+	copyColumn(l_s, l_t, l_c, d) = \begin{cases}
+ 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{state(d) \bullet addColumn(l_t, c)}
+	\end{cases} \\
+	l_s,l_t,l_c \in Label \wedge l_s \neq l_t,d \lhd Database
+\end{align*}
+
 \paragraph{Insert row} The function allow to insert a row(s) into a database.
@@
 \end{align*}
+
+\paragraph{Insert Value} The function inserts a new pair into an existing row.
+\begin{align*}
+	\mathbf{insertValues} : Row \times Pair* \times Database \rightarrow Daatabase \\
+	insertValues(r, p, d) = \begin{cases}
+ 		\inference{|p| > 1}{insertValues(r, head(p), d)\bullet insertValues(r, tail(p), d)}  \\ \\
+ 		\inference{|p| = 1 \wedge \forall e \in pairs(r) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p} \\ \\
+	 	 Inconsistent \; d
+ 		\end{cases} \\
+r \lhd Row, p = Pair*,d \lhd Databases
+\end{align*}
+
 \paragraph{Add Value To Table} The function appends a pair to all rows of a table.
@@
 
-\paragraph{Insert Value} The function inserts a new pair into an existing row.
-\begin{align*}
-	\mathbf{insertValues} : Row \times Pair* \times Database \rightarrow Daatabase \\
-	insertValues(r, p, d) = \begin{cases}
- 		\inference{|p| > 1}{insertValues(r, head(p), d)\bullet insertValues(r, tail(p), d)}  \\ \\
- 		\inference{|p| = 1 \wedge \forall e \in pairs(r) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p} \\ \\
-	 	 Inconsistent \; d
- 		\end{cases} \\
-r \lhd Row, p = Pair*,d \lhd Databases
-\end{align*}
 \paragraph{Copy Value} The function copy values from one column in a table to another column in the same or in another table - according to mapping function $f_m$. 
@@
 
-%\begin{align*}
-%	v(l, fm, d) = \begin{cases}
-%	\inference{fm = \varnothing}{d}\\ \\
-%	\inference{|fm| > 1}{v(l, head(fm)) \bullet v(l, tail(fm))}\\ \\
-%	\inference{|fm| = 1 \exists c \in pairs(first(fm)) : reference(c) = l \\ \wedge \exists e \in pairs(second(fm)) : reference(e) = l}{e \cup findPair(l,c)} 
-% \end{cases}
-%\end{align*}
-
 \section{Evolutionaty Transformations}
-The set of all possible data evolution cases $E$ was defined in Section \ref{sec:evoIntro} together with sets $AppOps$ and $DbOps$ for evolving various layers of the software.
-The semantic of transformations in the set $E$ is defined by their interpretation on individual software layers.
+The set of all possible data evolution cases $E$ was defined in Section \ref{sec:evoIntro}.
+The semantic of transformations in the set $E$ is defined by their interpretation on individual software components.
 This section provides an overview of basic transformations (evolution cases) and theirs interpretation on the application and the database.
@@
 If the conditions of feasibility are not met, the transformation cannot change the context - the software cannot evolve.
-We decide to use following transcription to cover all aspects of transformation definition:
-\begin{align*}
-operation(parameters, context) = \begin{cases}
-  \inference{conditions \; of \; feasibility}{impact \; on \; context}\\ \\
-  \inference{otherwise}{context}
- \end{cases}
-\end{align*}
 The second branch of the operation definition is an implicit assumption - if an operation is not feasible in a software, its state become inconsistent.
-We will not state this implicit assumption in the operations definitions to increase theirs readability.
+This implicit assumption is not stated in the transformation definitions to increase theirs readability.
 
 \subsection{Software Creation}
-There are defined three transformations which are able to create a software: $newClass$, $addProperty$, $addAssociation$.
+There are defined three transformations which are able to create a software: $addProperty$, $addAssociation$,  $newClass$.
 These three transformations can be used to create a software from a scratch.
-The first of them has no impact on stored data, whereas the rest of them can in some cases.
+The last of them has no impact on stored data, whereas the rest of them can in some cases.
 \begin{align*}
@@
 \end{align*}
-
-Whereas the $insertProp$ on database level can change stored data too, because a mandatory property can be added into a class with already existing instances, then a default value has to be added to all existing instances.
+The $insertProp$ on database level can change stored data too, because a mandatory property can be added into a class with already existing instances, then a default value has to be added to all existing instances.
 \begin{align*}
@@
 
+\paragraph{Add Class} The $insertClass$ operation changes the structure of the database schema only, because the $insertAssoc$ transformation needs a mapping function to work properly, which is not part of the $newClass$ definition.
+The situation reflect the modeling praxis when the structure of a class is defined and then it is associations are created.
 \begin{align*}
@@
 
-\paragraph{Add Class} The $insertClass$ operation changes the structure of the database schema only, because the $insertAssoc$ transformation needs a mapping function to work properly, which is not part of the $newClass$ definition.
-The situation reflect the modeling praxis when the structure of a class is defined and then it is associations are created.
 %\begin{align*}
@@
 \begin{align*}
-RemoveClass: Label \times Software \rightarrow State  \\
+\mathbf{RemoveClass}: Label \times Software \rightarrow State  \\
 \Psi RemoveClass(l, s) = applications(s) \circ remClass(l) \\
@@
 \begin{align*}
-RemoveProperty: Label \times Label \times State \rightarrow State \\
+\mathbf{RemoveProperty} : Label \times Label \times State \rightarrow State \\
 \Psi RemoveProperty(l, l_p, s) =  applications(s) \circ  remProp(l, l_p) \\
@@
 \begin{align*}
-RemoveAssociation : Label \times Labels \times State \rightarrow State \\ 
+\mathbf{RemoveAssociation} : Label \times Labels \times State \rightarrow State \\ 
 \Psi RemoveAssociation(l, l_a, s) = application(s) \circ remAss(l, l_a) \\
@@
 \begin{align*}
-RenameClass : Label \times Label \times State \rightarrow State \\ 
+\mathbf{RenameClass} : Label \times Label \times State \rightarrow State \\ 
 \Psi(RenameClass(l_o, l_n), s) = application(s) \circ renClass(l_o, l_n) \\
@@
 \begin{align*}
-	RenameProperty : Label \times Label \times Label \times State \rightarrow State \\ \\
+	\mathbf{RenameProperty} : Label \times Label \times Label \times State \rightarrow State \\ \\
 \Psi(RenameProperty(l_c, l_o, l_n, s)) = application(s) \circ renProp(l_c, l_o, l_n) \\
@@
 \begin{align*}
-RenameAssociation : Label \times Label \times Label \times State \rightarrow State \\
+\mathbf{RenameAssociation} : Label \times Label \times Label \times State \rightarrow State \\
 \Psi(RenameAssocaiation(l_c, l_o, l_n), s) = application(s) \circ renAssoc(l_c, l_o, l_n)\\
@@
 \paragraph{Copy Property}
-All transformations described earlier does change mostly the structure of a application only.
+All transformations described earlier does change primary the structure of an application.
 Therefore we define  the transformation for copying a property from one class to another.
-To copy property is the basic way to move data from instances of a class to instance of another class.
+Copy property is the basic way to move data among instances of different classes.
 Note the transformation does not move a single value, but it move the values for all instances in the software.
 
-The copy property can be implemented as a transformation composed by operations for adding and removing property, nevertheless this composition would not work with instances, where the $copyData$ transformation is used.
+The $CopyProperty$ function can be implemented as a transformation composed by operations for adding and removing property, nevertheless this composition would not work with instances, therefore transformation to update data are used.
 \begin{align*}
@@
 
+\section{Assigment for NII Internship / Future Work}
 \section{Conclusion}

### 2013-01-16 doc/description.tex
rewrited decomposition of SW

@@
 These sets are defined in this section.
+
 \subsection{Application Transformation}
@@
 
+\paragraph{Alter Table}
+\begin{align*}
+alterTable:  Label \times Label \times Database \rightarrow Database \\
+alterTable(l_o, l_n, d) = \inference{\exists t \in tables(d) : name(d) = l_o \\ \wedge \forall e \in tables(d) : name(e) \neq l_n}{\begin{gathered}
+t.name = l_n \\ \wedge \forall e \in tables(a) \wedge \forall (f \in fks(e) \wedge reference(f) = l_o ) \\ : f.reference = l_n 
+\end{gathered}}\\
+l_o, l_n \in Label, s \in States, a \lhd Applications, d \lhd Database
+\end{align*}
+
 \paragraph{Copy Column} The transformation creates a copy of a column in a another table in the database. The transformation changes only the structure of database - stored data are not changed.
@@
 	l \in Label, r \lhd Row
+\end{align*}
+
+\paragraph{Remove table}
+\begin{align*}
+	dropTable(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \not \exists r \in data(d) : owningTable(r) = l \\  \not \exists j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) = l }{ tables(d) \setminus \rho(c)}
+\end{align*}
+
+\paragraph{Remove column}
+\begin{align*}
+	dropColumn(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in columns(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{columns(t) \setminus c}\\
+	 l \in Label, d \lhd Database 
+\end{align*}
+
+\paragraph{Remove Foreign Key}
+\begin{align*}
+	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) \setminus c}\\
+	 l \in Label, d \lhd Database 
 \end{align*}
@@
 	\Psi RemoveClass(l, s) = applications(s) \circ remClass(l) \\
-\Phi RemoveClass(l, s) = database(s) \bullet dropClass(l) \\
-\\ l \in Label, s \in Software 
-\end{align*}
-\begin{align*}
-dropClass(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \not \exists r \in data(d) : owningTable(r) = l \\  \not \exists j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) = l }{ tables(d) \setminus \rho(c)}
+	\Phi RemoveClass(l, s) = database(s) \bullet dropTable(l) \\
+	l \in Label, s \in Software 
 \end{align*}
@@
 	\Psi RemoveProperty(l, l_p, s) =  applications(s) \circ  remProp(l, l_p) \\
-\Phi RemoveProperty(l, l_p, s) =  database(s) \bullet dropProp(l, l_p) \end{align*}
-\begin{align*}
-dropProp(p, d) = \begin{cases}
- \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \exists c \in columns(t) : name(c) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p)) }{columns(t) \setminus c
-} \\ \\
-  \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \not \exists c \in columns(t) : name(c) = name(p) \\ \wedge \exists u \in tables(d) : name(u) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p))}{ tables(d) \setminus u }
- \end{cases}
+	\Phi RemoveProperty(l, l_p, s) = \begin{cases}
+		 \inference{cardinality(propertyByName(l_p, application(s)) \leq 1}{database(s) \bullet dropColumn(l, l_p)}  \\\\
+		 \inference{cardinality(propertyByName(l_p, application(s)) >1}{database(s) \bullet dropTable(l)}
+ 	\end{cases}\\
+ 	l, l_p \in Label, s \lhd Software
 \end{align*}
@@
     \Psi RemoveAssociation(l, l_a, s) = application(s) \circ remAss(l, l_a) \\
-\Phi RemoveAssociation(l, l_a, s) = database(s) \bullet dropAssoc(l, l_a) \\
-l_c, l_a \in Labels, a \lhd Applications
-\end{align*}
-\begin{align*}
-dropAssoc(l, l_a, d) =  \begin{cases}
- \inference{\exists t \in tables(d) : name(t) = l
- \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_a
- \\ \wedge \forall r \in data(d) : owningTable(r) \neq t}{ foreignKeys(t) \setminus f }
- \\ \\
- \inference{\exists t \in tables(d) : name(t) = l_a
- \\ \wedge \exists f \in foreignKeys(t) : name(f) = l
- \\ \wedge \forall r \in data(d) : reference(r) \neq t}{ tables(d) \setminus t}
- \end{cases}
+    \Phi RemoveAssociation(l, l_a, s) = \begin{cases}
+     	\inference{a = associationByName(l, l_a, application(s)) \\ \wedge startCardinality(a) \leq 1 \\}{database(s) \bullet dropForeignKey(l_a, reference(a))} \\\\
+     	\inference{a = associationByName(l, l_a, application(s)) \\ \wedge startCardinality(a) > 1}{database(s) \bullet dropTable(l_a)} 
+     \end{cases}\\
+    l_c, l_a \in Labels, s \lhd Software
 \end{align*}
@@
 \Psi(RenameClass(l_o, l_n), s) = application(s) \circ renClass(l_o, l_n) \\
-\Phi(RenameClass(l_o, l_n), s) = database(s) \bullet alterClass(l_o, l_n) \\ \\
-\end{align*}
-\begin{align*}
-alterTable:  Label \times Label \times Database \rightarrow Database \\
-alterTable(l_o, l_n, d) = \inference{\exists t \in tables(d) : name(d) = l_o \\ \wedge \forall e \in tables(d) : name(e) \neq l_n}{\begin{gathered}
-t.name = l_n \\ \wedge \forall e \in tables(a) \wedge \forall (f \in fks(e) \wedge reference(f) = l_o ) \\ : f.reference = l_n 
-\end{gathered}}\\
-l_o, l_n \in Label, s \in States, a \lhd Applications, d \lhd Database
+\Phi(RenameClass(l_o, l_n), s) = database(s) \bullet alterTable(l_o, l_n) \\ \\
 \end{align*}
@@
  	\inference{cardinality(propertyByName(l_c, l_o, application(s))) \leq 1}{database(s) \bullet alterColumn(l_c, l_o, l_n))} \\\\
- 	\inference{cardinality(propertyByName(l_c, l_o, application(s))) > 1}{alterTable(l_o, l_n, database(s))}
+ 	\inference{cardinality(propertyByName(l_c, l_o, application(s))) > 1}{database(s) \bullet alterTable(l_o, l_n, database(s))}
  \end{cases}\\

### 2013-01-16 doc/description.tex
fixed description and definition of AddProperty

@@
 The last of them has no impact on stored data, whereas the rest of them can in some cases.
+\paragraph{Add Property} Adding a property to database can change not only structure but stored data too, because a mandatory property can be added into a class with already existing instances, then a default value has to be added to all existing instances.
 \begin{align*}
@@
 	\Psi(s, addProperty(l, p)) = application(s) \circ addProp(l,p) \\
-\Phi(s, addProperty(l, p)) = database(s) \bullet insertProp(l, p) \\
+	\Phi(s, addProperty(l, p)) = \begin{cases}
+    		\inference{cardinality(p) \leq 1}{database(s) \bullet addColumn(l, \rho(p,d), d)} \\ \\
+	    \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
+   database(s) \bullet addTable(\rho(p,d), d) \\ \bullet \forall r \in row(owningClass(p)) :\\ insertRow(row(name(p), ["value", defaultValue(p)], \\ [name(p), value(primaryKey(r))])) 
+    \end{gathered} } 
+   \end{cases}\\
 	s \lhd Software, l \in Label, p \in Properties
-\end{align*}
-The $insertProp$ on database level can change stored data too, because a mandatory property can be added into a class with already existing instances, then a default value has to be added to all existing instances.
-\begin{align*}
-insertProp(p, d) = \begin{cases}
-\inference{!\:mandatory(p) \vee !\:instantiated(owningClass(p))}{addColumn(l, \rho(p,d), d)} 
-\\ \\ 
-\inference{ \wedge mandatory(p) \vee instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered} addColumn(l, \rho(p,d), d) \\ \bullet \forall r \in data(d), reference(r) = name(t) : \\ addValueToTable(name(t), [name(p), defaultValue(p)], d)
-\end{gathered}
-} 
-\\ \\
-\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
-addTable(\rho(p,d), d) \bullet \forall r \in row(owningClass(p)) :\\ insertRow(row(name(p), ["value", defaultValue(p)], \\ [name(p), value(primaryKey(r))]),d) 
-\end{gathered}
-} 
-\end{cases}\\
-p \in Properties, d \in Dataabses
 \end{align*}

### 2013-01-16 doc/description.tex
refactored db and app constructors

@@
 	state(s) = \begin{cases}
- 		\inference{consistent(application(s)) \wedge \\ consistent(database(s)) \\ \wedge \rho(application(s)) = database(s)}{Consistent \; s}\\ \\
+ 		\inference{consistent(app(s)) \wedge \\ consistent(db(s)) \\ \wedge \rho(app(s)) = db(s)}{Consistent \; s}\\ \\
  		\inference{otherwise}{Inconsistent \; s}
@@
 	\mathbf{instantiated} : Class \times Software \rightarrow Boolean \\
-	instantiated(c, s) = \begin{cases} \inference{\exists t \in tables(database(s)) : name(t) = name(c) \\ \wedge contains(t, database(s))}{true} \\ \\
+	instantiated(c, s) = \begin{cases} \inference{\exists t \in tables(db(s)) : name(t) = name(c) \\ \wedge contains(t, db(s))}{true} \\ \\
   false
  \end{cases} \\ \\
-	 c \in classes(application(s)), s \lhd Software 
+	 c \in classes(app(s)), s \lhd Software 
 \end{align*}
@@
 	\mathbf{addProperty} : Label \times Property \times State \rightarrow State \\
-	\Psi(s, addProperty(l, p)) = application(s) \circ addProp(l,p) \\
+	\Psi(s, addProperty(l, p)) = app(s) \circ addProp(l,p) \\
 	\Phi(s, addProperty(l, p)) = \begin{cases}
-    		\inference{cardinality(p) \leq 1}{database(s) \bullet addColumn(l, \rho(p,d), d)} \\ \\
+    		\inference{cardinality(p) \leq 1}{db(s) \bullet addColumn(l, \rho(p,d), d)} \\ \\
 	    \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
-   database(s) \bullet addTable(\rho(p,d), d) \\ \bullet \forall r \in row(owningClass(p)) :\\ insertRow(row(name(p), ["value", defaultValue(p)], \\ [name(p), value(primaryKey(r))])) 
+   db(s) \bullet addTable(\rho(p,d), d) \\ \bullet \forall r \in row(owningClass(p)) :\\ insertRow(row(name(p), ["value", defaultValue(p)], \\ [name(p), value(primaryKey(r))])) 
     \end{gathered} } 
@@
 \mathbf{addAsociation} : Label \times Association \times (Row \times Row) \times State \rightarrow State \\
-\Psi(s, addAsociation(l, a, f_m)) = application(s) \circ addAssoc(l, a) \\
-\Phi(s, addAsociation(l, a, f_m)) = database(s) \bullet insertAssoc(l, a, f_m)
+\Psi(s, addAsociation(l, a, f_m)) = app(s) \circ addAssoc(l, a) \\
+\Phi(s, addAsociation(l, a, f_m)) = db(s) \bullet insertAssoc(l, a, f_m)
 \end{align*}
@@
 \mathbf{newClass} : Class \times State \rightarrow State \\
-\Psi(s, newClass(c)) = application(s) \circ addClass(c) \\
-\Phi(s, newClass(c)) = database(s)  \bullet addTable(\rho(c,d), d) \\ \bullet addColumn(name(c), \{\forall r \in singleProperties(c) | \rho(r,d)\}, d) \\ \bullet addTable(name(c), \{\forall r \in collections(c) | \rho(r,d)\}, d)
+\Psi(s, newClass(c)) = app(s) \circ addClass(c) \\
+\Phi(s, newClass(c)) = db(s)  \bullet addTable(\rho(c,d), d) \\ \bullet addColumn(name(c), \{\forall r \in singleProperties(c) | \rho(r,d)\}, d) \\ \bullet addTable(name(c), \{\forall r \in collections(c) | \rho(r,d)\}, d)
 \end{align*}
@@
 	\Psi RemoveClass(l, s) = applications(s) \circ remClass(l) \\
-	\Phi RemoveClass(l, s) = database(s) \bullet dropTable(l) \\
+	\Phi RemoveClass(l, s) = db(s) \bullet dropTable(l) \\
 	l \in Label, s \in Software 
@@
 	\Phi RemoveProperty(l, l_p, s) = \begin{cases}
-		 \inference{cardinality(propertyByName(l_p, application(s)) \leq 1}{database(s) \bullet dropColumn(l, l_p)}  \\\\
-		 \inference{cardinality(propertyByName(l_p, application(s)) >1}{database(s) \bullet dropTable(l)}
+		 \inference{cardinality(propertyByName(l_p, app(s)) \leq 1}{db(s) \bullet dropColumn(l, l_p)}  \\\\
+		 \inference{cardinality(propertyByName(l_p, app(s)) >1}{db(s) \bullet dropTable(l)}
  	\end{cases}\\
@@
     \mathbf{RemoveAssociation} : Label \times Labels \times State \rightarrow State \\ 
-    \Psi RemoveAssociation(l, l_a, s) = application(s) \circ remAss(l, l_a) \\
+    \Psi RemoveAssociation(l, l_a, s) = app(s) \circ remAss(l, l_a) \\
     \Phi RemoveAssociation(l, l_a, s) = \begin{cases}
-     	\inference{a = associationByName(l, l_a, application(s)) \\ \wedge startCardinality(a) \leq 1 \\}{database(s) \bullet dropForeignKey(l_a, reference(a))} \\\\
-     	\inference{a = associationByName(l, l_a, application(s)) \\ \wedge startCardinality(a) > 1}{database(s) \bullet dropTable(l_a)} 
+     	\inference{a = associationByName(l, l_a, app(s)) \\ \wedge startCardinality(a) \leq 1 \\}{db(s) \bullet dropForeignKey(l_a, reference(a))} \\\\
+     	\inference{a = associationByName(l, l_a, app(s)) \\ \wedge startCardinality(a) > 1}{db(s) \bullet dropTable(l_a)} 
      \end{cases}\\
@@
 \mathbf{RenameClass} : Label \times Label \times State \rightarrow State \\ 
-\Psi(RenameClass(l_o, l_n), s) = application(s) \circ renClass(l_o, l_n) \\
-\Phi(RenameClass(l_o, l_n), s) = database(s) \bullet alterTable(l_o, l_n) \\ \\
+\Psi(RenameClass(l_o, l_n), s) = app(s) \circ renClass(l_o, l_n) \\
+\Phi(RenameClass(l_o, l_n), s) = db(s) \bullet alterTable(l_o, l_n) \\ \\
 \end{align*}
@@
 	\mathbf{RenameProperty} : Label \times Label \times Label \times State \rightarrow State \\ \\
-\Psi(RenameProperty(l_c, l_o, l_n, s)) = application(s) \circ renProp(l_c, l_o, l_n) \\
+\Psi(RenameProperty(l_c, l_o, l_n, s)) = app(s) \circ renProp(l_c, l_o, l_n) \\
 \Phi(RenameProperty(l_c, l_o, l_n, s)) = \begin{cases}
- 	\inference{cardinality(propertyByName(l_c, l_o, application(s))) \leq 1}{database(s) \bullet alterColumn(l_c, l_o, l_n))} \\\\
- 	\inference{cardinality(propertyByName(l_c, l_o, application(s))) > 1}{database(s) \bullet alterTable(l_o, l_n, database(s))}
+ 	\inference{cardinality(propertyByName(l_c, l_o, app(s))) \leq 1}{db(s) \bullet alterColumn(l_c, l_o, l_n))} \\\\
+ 	\inference{cardinality(propertyByName(l_c, l_o, app(s))) > 1}{db(s) \bullet alterTable(l_o, l_n, db(s))}
  \end{cases}\\
@@
 \mathbf{RenameAssociation} : Label \times Label \times Label \times State \rightarrow State \\
-\Psi(RenameAssocaiation(l_c, l_o, l_n), s) = application(s) \circ renAssoc(l_c, l_o, l_n)\\
+\Psi(RenameAssocaiation(l_c, l_o, l_n), s) = app(s) \circ renAssoc(l_c, l_o, l_n)\\
 \Phi(RenameAssocaiation(l_c, l_o, l_n), s) = \begin{cases}
-	\inference{\exists a  \in associationByName(application(s)) : \\
-  name(a) = l_c \\ \wedge startCardinality(a) > 1 }{database(s) \bullet alterTable(l_c, l_o, l_n)}
+	\inference{\exists a  \in associationByName(app(s)) : \\
+  name(a) = l_c \\ \wedge startCardinality(a) > 1 }{db(s) \bullet alterTable(l_c, l_o, l_n)}
  \\ \\
-  \inference{\exists a  \in associationByName(application(s)) : \\
+  \inference{\exists a  \in associationByName(app(s)) : \\
   name(a) = l_c \\ \wedge startCardinality(a) \leq 1 }{\begin{gathered}
-database(s) \\ \bullet alterForeignKey(endClass(a) ,l_c, l_o, l_n)
+db(s) \\ \bullet alterForeignKey(endClass(a) ,l_c, l_o, l_n)
 \end{gathered}}
@@
 	\mathbf{CopyProperty} : Label \times Label \times Label \times (Row \times Row) \times State \rightarrow State \\
-	\Psi(CopyProperty(l_s, l_t, l_p, f_m), s) = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)} \\\\
+	\Psi(CopyProperty(l_s, l_t, l_p, f_m), s) = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{app(s) \circ addProp(l_t, p)} \\\\
 	\Phi(CopyProperty(l_s, l_t, l_p, f_m), s) = \begin{cases}
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) \leq 1 }{\begin{gathered}
-			database(s) \bullet copyColumn(l_s, l_t, l_p) \\ \bullet copyValues(l_s, l_t, l_p, f_m) 
+			db(s) \bullet copyColumn(l_s, l_t, l_p) \\ \bullet copyValues(l_s, l_t, l_p, f_m) 
 			\end{gathered}} \\ \\
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) > 1}{\begin{gathered}
-			database(s) \bullet copyTable(l_s, l_t, l_p) \\ \bullet copyRows(l_s, l_t, l_p, f_m) 
+			db(s) \bullet copyTable(l_s, l_t, l_p) \\ \bullet copyRows(l_s, l_t, l_p, f_m) 
 			\end{gathered}}

### 2013-01-16 doc/description.tex
application consistency + fixed mistake from last commit

--- 2013-01-16 doc/description.tex
@@
  		\inference{otherwise}{Inconsistent \; s}
- 	\end{cases} \\ \\
+ 	\end{cases} \\ 
+ 	s \lhd Software
 \end{align*}
@@
 \subsubsection{Application consistency}
+The application consistency is based on the structural consistency of its model.
+Violation of structural consistency are name collisions and non-existing references in associations.
+The state of an application is:
+\begin{align*}
+	state(a) = \begin{cases}
+	\inference{\forall c_1, c_2 \in classes(a) : name(c_1) \neq name(c_2) \\ \wedge \forall p_1, p_2 \in properties(c_1) : name(p_1) \neq name(p_2) \\ \wedge \forall a_1, a_2 \in associations(c_1) : name(a_1) \neq name (a_2) \\ \wedge \forall a \in associations(c_1) : \exists c \in classes(a) \wedge name(c) = reference(a) }{Consistent \; a}\\\\
+ 	Inconsistent \; a
+ 	\end{cases}\\
+	a \lhd Application
+\end{align*}
 

### 2013-01-16 doc/description.tex
database consistency

@@
 The database consistency is founded on the consistency of structure and consistency between structure and stored values.
-Each row in a database has to refer to a table in database schema and it has to have the structure corresponding to the one defined by a table, otherwise the database is inconsistent.
-The consistency of database is then:
+The conditions for structure consistency are similar as for application structure.
+The consistency between structure and data is:  each row in a database  refers to a table in database schema and it has to have the structure corresponding to the one defined by a table, moreover the column and foreign key constrains has to be fulfilled  otherwise the database is inconsistent.
+The consistency of database is:
 \begin{align*}
 	\mathbf{consistent} : Database \rightarrow Boolean \\
-	consistent(d) = \begin{cases}
- 		\inference{\forall r \in data(d), \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) :\\ \exists u \in columns(t) \cup foreignKeys(t) : \\ reference(p) = name(u)}{true}
+	state(d) = \begin{cases}
+ 		\inference{ \forall t_1, t_2 \in tables(d) : name(t_1) \neq name(t_2) \\ \wedge \forall c_1, c_2 \in columns(t_1) : name(c_1) \neq name(c_2) \\ \wedge \forall f_1, f_2 \in foreignKeys(t_1) : name(f_1) \neq name (f_2) \\ \wedge \forall f \in foreignKeys(t_1) : \exists t \in tables(d) \wedge name(t) = reference(f) \\ \wedge \forall r \in data(d): ( \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) :\\ \exists u \in columns(t) \cup foreignKeys(t) \cup primaryKey(t) : \\ reference(p) = name(u) \\ \wedge \forall e \in columns(t) \cup foreignKeys(t) \\ \wedge NOTNULL \in constrain(e) : \exists q \in pairs(r) \wedge name(e) = reference(q)) \\ \wedge \forall g \in columns(t) \cup foreignKeys(t) \\ \wedge UNIQUE \in constrain(g) : \\ \forall u_1, u_2 \in rows(t) : \\ value(pairByName(name(g), u_1) \neq value(pairByName(name(g), u_2) }{true}
  	\\ \\

### 2013-01-16 doc/description.tex
added description to functions

@@
 
-\paragraph{Get Property or Association}
+\paragraph{Get Property or Association} Functions $findProperty$ and $findAssociation$ returns a property (association) if it exists in given context - application and class of provided name.
 \begin{align*}
-	\mathbf{propertyByName} : Label \times Label \times Application \rightarrow Property \\
-	propertyByName(l_c, l_p, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{p} \\
+	\mathbf{findProperty} : Label \times Label \times Application \rightarrow Property \\
+	findProperty(l_c, l_p, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{p} \\
 	l_c, l_p \in Label, a \lhd Application
@@
 \begin{align*}
-	\mathbf{associationByName} : Label \times Label \times Application \rightarrow Association \\
-	associationByName(l_c, l_a, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists o \in associations(c) : name(o) = l_a}{o} \\
+	\mathbf{findAssociation} : Label \times Label \times Application \rightarrow Association \\
+	findAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists o \in associations(c) : name(o) = l_a}{o} \\
 	l_c, l_a \in Label, a \lhd Application
@@
 
-\paragraph{Compare Tables} The need to compare two tables during the evolution results into the comparing operator $\cong$. The operator has to be defined for columns and foreign keys too.
+\paragraph{Compare} The need to compare two tables during the evolution results into the comparing operator $\cong$. The operator has to be defined for columns and foreign keys too.
 \begin{align*}
@@
 
-\paragraph{Find pair}
+\paragraph{Find pair} The function returns a pair corresponding to given column (or key) name in concrete row.
 \begin{align*}
@@
 
-\paragraph{Alter Table}
+\paragraph{Alter Table} The $alterTable$ function renames a table in  the database.
 \begin{align*}
@@
 
-\paragraph{Copy rows}
+\paragraph{Copy rows} The function $copyRows$ creates a copy of a source table content in another table with the same structure. 
 \begin{align*}
@@
  	\inference{|f_m| = \varnothing}{d} \\\\
- 	\inference{|f_m| = 1 \\ \wedge \exists t_s, t_t \in tables(d) : name(t_s) = l_s, name(t_t) = l_t \\ \wedge t_s \cong t_t }{data(d) \cup insertRow(changeRef(l_t, first(f_m)), d)} \\\\
+ 	\inference{|f_m| = 1 \\ \wedge \exists t_s, t_t \in tables(d) : name(t_s) = l_s, name(t_t) = l_t \\ \wedge t_s \cong t_t }{data(d) \cup insertRow(row(l, pairs(first(f_m))), d)} \\\\
  	\inference{|f_m| > 1}{\begin{gathered}
@@
 
-\paragraph{Change ref}
+\paragraph{Remove table} The function removes a table if there are no stored data.
+This is because the data lost is undesirable in case of evolution.
+The consistency of foreign keys references has to be verified too.
 \begin{align*}
-	\mathbf{changeRef} : Label \times Row \rightarrow Row \\
-	changeRef(l, r) = \inference{}{reference(r) = l } \\
-	l \in Label, r \lhd Row
-\end{align*}
-
-\paragraph{Remove table}
-\begin{align*}
+	\mathbf{dropTable} : Label \times Database \rightarrow Database \\
 	dropTable(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \not \exists r \in data(d) : owningTable(r) = l \\  \not \exists j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) = l }{ tables(d) \setminus \rho(c)}
@@
 
-\paragraph{Remove column}
+\paragraph{Remove column or foreign key} The situation is similar to the $dropTable$ function - there has to be no data in the column or key so it can be safely removed.
 \begin{align*}
+	\mathbf{dropColumn} : Label \times Label \times Database \rightarrow Database \\
 	dropColumn(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in columns(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{columns(t) \setminus c}\\
-	 l \in Label, d \lhd Database 
+	 l , l_p\in Label, d \lhd Database 
 \end{align*}
-
-\paragraph{Remove Foreign Key}
 \begin{align*}
+	\mathbf{dropKey} : Label \times Label \times Database \rightarrow Database \\
 	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) \setminus c}\\
-	 l \in Label, d \lhd Database 
+	 l, l_p \in Label, d \lhd Database 
 \end{align*}
@@
 
+\paragraph{Add association} Adding an association into a database is complicated because of already existing instances.
+When an association between classes is added the instances of both classes has to be paired.
+This pairing is provided by the function $f_m$:
+$$f_m : \rightarrow Row \times Row $$
+
+\textbf{TODO: define the features of the $f_m$ function}
 \begin{align*}
@@
 \end{align*}
-
-Adding an association into a database is complicated because of already existing instances.
-When an association between classes is added the instances of both classes has to be paired.
-This pairing is provided by the function $f_m$:
-$$f_m : \rightarrow Instance \times Instance $$
-
-\textbf{TODO: define the features of the $f_m$ function}
-
 The $insertAssoc$ function is defined as:
@@
 insertAssoc : Association \times Database \times (Row \times Row) \\
-insertAssoc(a, d, f_m) = \begin{cases}
-\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
-addForeignKey(name(t_s), \rho(a,d), d) \\ \bullet \forall r \in f_m(owningClass(a), reference(a)) : \\ insertRow(row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]), d)
+insertAssoc(l, a, f_m, d) = \begin{cases}
+\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l\\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
+addForeignKey(name(t_s), \rho(a,d), d) \\ \bullet \forall r \in f_m(l, reference(a)) : \\ insertRow(row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]), d)
 \end{gathered}
@@
  \\ \\
-\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) > 1}{\begin{gathered} addTable(\rho(a,d),d) \\ \bullet \forall r \in f_m(owningClass(a), reference(a)),\\ t \in tables(d) : name(t) = name(a) : \\ insertRow(row(name(a), [name(t_s), \\ value(primaryKey(first(r)))] [name(t_t), \\ value(primaryKey(second(r)))]), d)
+\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) > 1}{\begin{gathered} addTable(\rho(a,d),d) \\ \bullet \forall r \in f_m(l, reference(a)),\\ t \in tables(d) : name(t) = name(a) : \\ insertRow(row(name(a), [name(t_s), \\ value(primaryKey(first(r)))] [name(t_t), \\ value(primaryKey(second(r)))]), d)
 \end{gathered}
@@
 \subsection{Software Decomposition}
+The removal of class, property or association is limited only by the constraints given in application or database transformations.
+The main constraint is there has to be no data, which can be lost during removal.
+\paragraph{Remove class}
 \begin{align*}
@@
 \end{align*}
-
+\paragraph{Remove property}
 \begin{align*}
@@
 	\Phi RemoveProperty(l, l_p, s) = \begin{cases}
-		 \inference{cardinality(propertyByName(l_p, application(s)) \leq 1}{database(s) \bullet dropColumn(l, l_p)}  \\\\
-		 \inference{cardinality(propertyByName(l_p, application(s)) >1}{database(s) \bullet dropTable(l)}
+		 \inference{cardinality(findProperty(l_p, application(s)) \leq 1}{database(s) \bullet dropColumn(l, l_p)}  \\\\
+		 \inference{cardinality(findProperty(l_p, application(s)) >1}{database(s) \bullet dropTable(l)}
  	\end{cases}\\
@@
 \end{align*}
-
+\paragraph{Remove association} 
 \begin{align*}
@@
     \Phi RemoveAssociation(l, l_a, s) = \begin{cases}
-     	\inference{a = associationByName(l, l_a, application(s)) \\ \wedge startCardinality(a) \leq 1 \\}{database(s) \bullet dropForeignKey(l_a, reference(a))} \\\\
-     	\inference{a = associationByName(l, l_a, application(s)) \\ \wedge startCardinality(a) > 1}{database(s) \bullet dropTable(l_a)} 
+     	\inference{a = findAssociation(l, l_a, application(s)) \\ \wedge startCardinality(a) \leq 1 \\}{database(s) \bullet dropForeignKey(l_a, reference(a))} \\\\
+     	\inference{a = findAssociation(l, l_a, application(s)) \\ \wedge startCardinality(a) > 1}{database(s) \bullet dropTable(l_a)} 
      \end{cases}\\
@@
 \subsection{Software Alternation}
+Alternation of a software is limited to renaming functions in our model.
 \paragraph{Rename Class}
@@
 \Phi(RenameProperty(l_c, l_o, l_n, s)) = \begin{cases}
- 	\inference{cardinality(propertyByName(l_c, l_o, application(s))) \leq 1}{database(s) \bullet alterColumn(l_c, l_o, l_n))} \\\\
- 	\inference{cardinality(propertyByName(l_c, l_o, application(s))) > 1}{database(s) \bullet alterTable(l_o, l_n, database(s))}
+ 	\inference{cardinality(findProperty(l_c, l_o, application(s))) \leq 1}{database(s) \bullet alterColumn(l_c, l_o, l_n))} \\\\
+ 	\inference{cardinality(findProperty(l_c, l_o, application(s))) > 1}{database(s) \bullet alterTable(l_o, l_n, database(s))}
  \end{cases}\\
@@
 \Phi(RenameAssocaiation(l_c, l_o, l_n), s) = \begin{cases}
-	\inference{\exists a  \in associationByName(application(s)) : \\
+	\inference{\exists a  \in findAssociation(application(s)) : \\
   name(a) = l_c \\ \wedge startCardinality(a) > 1 }{database(s) \bullet alterTable(l_c, l_o, l_n)}
  \\ \\
-  \inference{\exists a  \in associationByName(application(s)) : \\
+  \inference{\exists a  \in findAssociation(application(s)) : \\
   name(a) = l_c \\ \wedge startCardinality(a) \leq 1 }{\begin{gathered}

### 2013-01-16 doc/description.tex
small fix in add calss description

@@
 
-\paragraph{Add Class} Adding a class is limited to the case when the class has no associations, this is because it simplifies the database migration and because it is a natural way of modeling.
-Next the transformation $addClass$ has to verify there is no name collision
+\paragraph{Add Class} Adding a class is limited to the case when the class has no associations, this is because it simplifies the database migration.
+The transformation $addClass$ has to verify there is no name collision
 \begin{align*}
@@
 \paragraph{Add Class} The $insertClass$ operation changes the structure of the database schema only, because the $insertAssoc$ transformation needs a mapping function to work properly, which is not part of the $newClass$ definition.
-The situation reflect the modeling praxis when the structure of a class is defined and then it is associations are created.
+The situation reflect the modeling praxis when the structure of a class is defined and then its associations are created.
 \begin{align*}

### 2013-01-17 doc/description.tex
defined mapping from application to database

@@
 The mapping function itself are not able to create a database from the application as they are mentioned to be used together with evolutionary transformations introduced in section \ref{sec:transformations}.
+
+\paragraph{Mapping of Application} The goal of ORM is usually to determine how an instance is stored in the database.
+The ORM maps the structure of application classes to database schema.
+The assumption is a software is evolved continuously and evolutionary transformations produces consistent software, nevertheless to verify consistency of a software we need mapping between an existing application structure and database schema.
+\begin{align*}
+	\mathbf{\rho_{app}} : Application \rightarrow Database\\
+	\rho_{app}(a) = \inference{state(a) = Consistent a}{\begin{gathered}
+	db(\varnothing) \bullet addTable(classes(a)) \\ 
+	\bullet (\forall c \in classes(a), p \in properties(c) : addProperty(name(c), p) \\
+	\bullet (\forall c \in classes(a), e \in associations(c) : addAssociaiton(name(c), e) 
+\end{gathered}}
+\end{align*}
 

### 2013-01-17 doc/description.tex
fixed formating of alterTable

@@
 \begin{align*}
-alterTable:  Label \times Label \times Database \rightarrow Database \\
+	\mathbf{alterTable} : Label \times Label \times Database \rightarrow Database \\
 	alterTable(l_o, l_n, d) = \inference{\exists t \in tables(d) : name(d) = l_o \\ \wedge \forall e \in tables(d) : name(e) \neq l_n}{\begin{gathered}

### 2013-01-17 doc/description.tex
fixed name collisions in associaitons and properties + updated mapping + hidden find operations

@@
 When the property is mandatory the $NOTNULL$ constraint is added to the column with property value.
+
+There can occur name collision during mapping in case there are more classes with property representing collection. To avoid that the mapping changes the name of the table representing a collection.
 \begin{align*}
@@
      		}{\begin{gathered}
-     			table(name(p), primaryKey("Id"), \\ column("value", 			\rho_t(type(p)), \varnothing), \\ foreignKey(name(p), 			name(owningClass(p)), NOTNULL) \\
+     			table(name(t)+name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), \varnothing), \\ foreignKey(name(p), 			name(owningClass(p)), NOTNULL) \\
     	\end{gathered}}\\ \\
@@
      		}{\begin{gathered}  
-     			table(name(p), primaryKey("Id"), \\ 
+     			table(name(t)+name(p), primaryKey("Id"), \\ 
      			column("value", \rho_t(type(p)), NOTNULL), \\ 			foreignKey(name(p), name(owningClass(p)), NOTNULL)
@@
 \paragraph{Mapping of Associations} An association is mapped into database as a foreign key in table representing associated class or as a table in case the association cardinality is one to many or many to many.
-Thus there are two mapping cases as for properties.
+Thus there are two mapping cases as for properties and the same solution is used to avoid name collisions.
 \begin{align*}
@@
 		\inference{|a| > 1}{\rho_{a}(head(a)) \cup \rho_{a}(tail(a))}\\\\
-		\inference{  startCardinality > 1 \\ \wedge \rho(owningClass(p)) \in tables(d) \wedge \rho(reference(a)) \in tables(d)
+		\inference{  startCardinality > 1 \\ \wedge \exists t \in tables(d) : name(owningClass(p)) = name(t) \\ \wedge \exists e \in tables(d) : name(reference(a)) = name(e)
   		}{\begin{gathered}  
-		 table(name(a), pk("Id"), \varnothing, \\ foreignKey(owningClass(a),owningClass(a), NOTNULL) \\ foreignkey(reference(a), reference(a), NOTNULL)) 
+		 table(name(t)+name(a), pk("Id"), \varnothing, \\ foreignKey(owningClass(a),owningClass(a), NOTNULL) \\ foreignkey(reference(a), reference(a), NOTNULL)) 
   		\end{gathered}}  
@@
 
-\paragraph{Get Property or Association} Functions $findProperty$ and $findAssociation$ returns a property (association) if it exists in given context - application and class of provided name.
-\begin{align*}
-	\mathbf{findProperty} : Label \times Label \times Application \rightarrow Property \\
-	findProperty(l_c, l_p, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{p} \\
-	l_c, l_p \in Label, a \lhd Application
-\end{align*}
-\begin{align*}
-	\mathbf{findAssociation} : Label \times Label \times Application \rightarrow Association \\
-	findAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists o \in associations(c) : name(o) = l_a}{o} \\
-	l_c, l_a \in Label, a \lhd Application
-\end{align*}
+%\paragraph{Get Property or Association} Functions $findProperty$ and $findAssociation$ returns a property (association) if it exists in given context - application and class of provided name.
+%\begin{align*}
+%	\mathbf{findProperty} : Label \times Label \times Application \rightarrow Property \\
+%	findProperty(l_c, l_p, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{p} \\
+%	l_c, l_p \in Label, a \lhd Application
+%\end{align*}
+%\begin{align*}
+%	\mathbf{findAssociation} : Label \times Label \times Application \rightarrow Association \\
+%	findAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists o \in associations(c) : name(o) = l_a}{o} \\
+%	l_c, l_a \in Label, a \lhd Application
+%\end{align*}
 
@@
  	\mathbf{addProp} : Label \times Property \times Application \rightarrow Application \\
-	addProp(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(p) : name(q) \neq name(p)}{properties(c) \cup p } \\
+	addProp(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(c) : name(q) \neq name(p) \\ \wedge \forall e \in associations(c) : name(e) \neq name(p)}{properties(c) \cup p } \\
 	l \in Label, p \lhd Property, a \lhd Applications
@@
 	\mathbf{addAssoc} : Label \times Association \times Application \rightarrow Application \\
-	addAssoc(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(c) \cup s}\\
+	addAssoc(l, s, a) = \inference{
+	\exists c \in classes(a) : name(c) = l \\
+	\wedge \forall f \in associations(c) : name(f) \neq name(s) \\ 	\wedge \forall p \in properties(c) : name(p) \neq name(s) \\ 
+	\wedge \exists k \in classes(a) : name(k) = reference(s)}{associations(c) \cup s}\\
 	l \in Labels, s \in Associations, a \lhd Applications
@@
 
-\paragraph{Find Table} The function $findTable$ finds and returns a table in the database according to the provided label.
-\begin{align*}
-	\mathbf{findTable} : Label \times Database \rightarrow Table  \\
-	findTable(l, d) = \inference{\exists t \in tables(d) \wedge name(t) = l}{t}  \\ \\
-	l \in Label, d \lhd Database
-\end{align*}
+%\paragraph{Find Table} The function $findTable$ finds and returns a table in the database according to the provided label.
+%\begin{align*}
+%	\mathbf{findTable} : Label \times Database \rightarrow Table  \\
+%	findTable(l, d) = \inference{\exists t \in tables(d) \wedge name(t) = l}{t}  \\ \\
+%	l \in Label, d \lhd Database
+%\end{align*}
 
@@
 \end{align*}
+
 \paragraph{Remove property}
@@
 	\Phi RemoveProperty(l, l_p, s) = \begin{cases}
-		 \inference{cardinality(findProperty(l_p, application(s)) \leq 1}{database(s) \bullet dropColumn(l, l_p)}  \\\\
-		 \inference{cardinality(findProperty(l_p, application(s)) >1}{database(s) \bullet dropTable(l)}
+		 \inference{\exists c \in classes(application(s)) : name(c) = l\\ 
+		\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) \leq 1}{database(s) \bullet dropColumn(l, l_p)}  \\\\
+		 \inference{\exists c \in classes(application(s)) : name(c) = l\\ 
+		\wedge \exists p \in properties(c)  : name(p) = l \\ \wedge cardinality(p)>1}{database(s) \bullet dropTable(name(owningClass(p)+l)}
  	\end{cases}\\
@@
 \end{align*}
+
 \paragraph{Remove association} 
@@
     \Phi RemoveAssociation(l, l_a, s) = \begin{cases}
-     	\inference{a = findAssociation(l, l_a, application(s)) \\ \wedge startCardinality(a) \leq 1 \\}{database(s) \bullet dropForeignKey(l_a, reference(a))} \\\\
-     	\inference{a = findAssociation(l, l_a, application(s)) \\ \wedge startCardinality(a) > 1}{database(s) \bullet dropTable(l_a)} 
+     	\inference{\exists c \in classes(application(s)) : name(c) = l \\
+     	\wedge \exists a \in associaions(c) : name(a) = l_a  \\
+     	\wedge startCardinality(a) \leq 1 \\}{database(s) \bullet dropForeignKey(l_a, reference(a))} \\\\
+     	\inference{\exists c \in classes(application(s)) : name(c) = l \\
+     	\wedge \exists a \in associaions(c) : name(a) = l_a  \\
+     	\wedge startCardinality(a) > 1}{database(s) \bullet dropTable(name(owningClass(a))+l_a)} 
      \end{cases}\\
@@
 \Phi(RenameProperty(l_c, l_o, l_n, s)) = \begin{cases}
- 	\inference{cardinality(findProperty(l_c, l_o, application(s))) \leq 1}{database(s) \bullet alterColumn(l_c, l_o, l_n))} \\\\
- 	\inference{cardinality(findProperty(l_c, l_o, application(s))) > 1}{database(s) \bullet alterTable(l_o, l_n, database(s))}
+ 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 
+		\wedge \exists p \in properties(c) : name(p) = l_o \\
+ 	cardinality(p) \leq 1}{database(s) \bullet alterColumn(l_c, l_o, l_n))} \\\\
+ 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 
+		\wedge \exists p \in properties(c) : name(p) = l_o \\
+ 	cardinality(p) > 1}{\begin{gathered}
+		database(s)\\ \bullet alterTable(name(owningClass(p)+l_o, l_n, database(s))
+\end{gathered}}
  \end{cases}\\
@@
 \Phi(RenameAssocaiation(l_c, l_o, l_n), s) = \begin{cases}
-	\inference{\exists a  \in findAssociation(application(s)) : \\
-  name(a) = l_c \\ \wedge startCardinality(a) > 1 }{database(s) \bullet alterTable(l_c, l_o, l_n)}
- \\ \\
-  \inference{\exists a  \in findAssociation(application(s)) : \\
-  name(a) = l_c \\ \wedge startCardinality(a) \leq 1 }{\begin{gathered}
+	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 
+		\wedge \exists a \in associations(c) : name(a) = l_o 
+	\\ \wedge startCardinality(a) > 1 }{\begin{gathered}
+		database(s) \\ \bullet alterTable(name(owningClass(a)) + l_o, l_n)
+\end{gathered}}\\ \\
+  \inference{\exists c \in classes(application(s)) : name(c) = l\\ 
+		\wedge \exists a \in associations(c) : name(a) = l_o 
+	\\ \wedge startCardinality(a) \leq 1 }{\begin{gathered}
 database(s) \\ \bullet alterForeignKey(endClass(a) ,l_c, l_o, l_n)
 \end{gathered}}
- \end{cases}
-\\\\
-%	alterAssoc(l_c, l_o, l_n, d) = \begin{cases}
-%?? \inference{\exists t \in tables(a) : name(t) = l_c \\ \wedge \exists f \in fks(t) : name(c) = l_o \\ \wedge \forall e \in fks(t) : name(e) \neq l_n}{\begin{gathered}
-%f.name = l_n 
-%\end{gathered}}
-%\\ \\
-%\inference{\exists t \in tables(a) : name(t) = l_o \\ \wedge \forall e \in tables(t) : name(e) \neq l_n}{\begin{gathered}
-%t.name = l_n 
-%\end{gathered}}
-% \end{cases}
-% \\
+ \end{cases} \\
 	l_c, l_o, l_n \in Label, a \lhd Applications, s \in Sates, d \lhd Database

### 2013-01-17 doc/description.tex
created copyTable

@@
 
+\paragraph{Copy Table} The transformation creates a copy of a table in database schema with different name.
+\begin{align*}
+	\mathbf{copyTable} : Label \times Label \times Database \times Database \\
+	copyTable(l_s, l_t, d) = \inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \forall e \in tables(d) : name(e) \neq l_t}{d \bullet addTable(tab(l_t, primaryKey(t), columns(t), foreignKeys(t)))}
+\end{align*}
+
 \paragraph{Copy Column} The transformation creates a copy of a column in a another table in the database. The transformation changes only the structure of database - stored data are not changed.
@@
 	\mathbf{CopyProperty} : Label \times Label \times Label \times (Row \times Row) \times State \rightarrow State \\
-	\Psi(CopyProperty(l_s, l_t, l_p, f_m), s) = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)} \\\\
+	\Psi(CopyProperty(l_s, l_t, l_p, f_m), s) = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)} 
+	\\\\
 	\Phi(CopyProperty(l_s, l_t, l_p, f_m), s) = \begin{cases}
@@
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) > 1}{\begin{gathered}
-			database(s) \bullet copyTable(l_s, l_t, l_p) \\ \bullet copyRows(l_s, l_t, l_p, f_m) 
+			database(s) \bullet copyTable(l_s + l_p, l_t + l_p) \\ \bullet copyRows(l_s + l_p, l_t + l_p, f_m) 
 			\end{gathered}}

### 2013-01-17 doc/description.tex
addTable works with non-empty FK set

@@
 	addTable(t, d) = \begin{cases}
-		\inference{|t| > 1}{addTable(head(t),d) \bullet addTable(tail(t), d)}
+		\inference{|t| > 1}{d \bullet addTable(head(t)) \bullet addTable(tail(t))}
 \\ \\	
-		\inference{|t| = 1 \wedge \forall e \in tables(d) \wedge name(t) \neq name(e) \\ \wedge foreignKeys(t) = \varnothing }{tables(d) \cup t}
+		\inference{|t| = 1 \wedge \forall e \in tables(d) \wedge name(t) \neq name(e) \\ \wedge \forall f \in foreignKeys(t) : \\ s\exists e \in tables(d) \wedge name(e) = reference(f) }{tables(d) \cup t}
 \\ \\
@@
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) \leq 1 }{\begin{gathered}
-			database(s) \bullet copyColumn(l_s, l_t, l_p) \\ \bullet copyValues(l_s, l_t, l_p, f_m) 
+			database(s) \bullet copyColumn(l_s, l_t, l_p) \\ \bullet copyValues(l_s, l_t, l_p, l_p, f_m) 
 			\end{gathered}} \\ \\

### 2013-01-17 doc/description.tex
fixes adding association

@@
 \Psi(s, addAsociation(l, a, f_m)) = application(s) \circ addAssoc(l, a) \\
-\Phi(s, addAsociation(l, a, f_m)) = database(s) \bullet insertAssoc(l, a, f_m)
-\end{align*}
-The $insertAssoc$ function is defined as:
-\begin{align*}
-insertAssoc : Association \times Database \times (Row \times Row) \\
-insertAssoc(l, a, f_m, d) = \begin{cases}
-\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l\\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
-addForeignKey(name(t_s), \rho(a,d), d) \\ \bullet \forall r \in f_m(l, reference(a)) : \\ insertRow(row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]), d)
-\end{gathered}
- }
+\Phi(s, addAsociation(l, a, f_m)) = \begin{cases}
+	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l\\ 		\wedge name(t_t) = reference(a) \\ \wedge \forall f \in foreignKeys(t_t) \cuo columns(t_t) : name(f) \neq name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
+		database(s) \bullet addForeignKey(name(t_s), \rho(a,d), d) \\ \bullet \forall r \in f_m(l, reference(a)) : \\ insertRow(row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]), d)
+	\end{gathered}}
  \\ \\
-\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) > 1}{\begin{gathered} addTable(\rho(a,d),d) \\ \bullet \forall r \in f_m(l, reference(a)),\\ t \in tables(d) : name(t) = name(a) : \\ insertRow(row(name(a), [name(t_s), \\ value(primaryKey(first(r)))] [name(t_t), \\ value(primaryKey(second(r)))]), d)
-\end{gathered}
- } 
+	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l \\ \wedge name(t_t) = reference(a) \\ \wedge startCardinality(a) > 1}{\begin{gathered} 
+	database(s) \bullet addTable(\rho(a,d),d) \\ \bullet \forall r \in f_m(l, reference(a)) :\\ insertRow(row(name(owningClass(a)) + name(a), \\ [name(t_s), value(primaryKey(first(r)))]\\ [name(t_t), value(primaryKey(second(r)))]))
+	\end{gathered}} 
 \end{cases}\\
- a \in Associations, d \lhd Database
 \end{align*}

### 2013-01-17 doc/description.tex
small text improvement

@@
 \section{Transformation for Models Manipulation}
-The evolution affects both parts of software and as shown in section \ref{sec:evoIntro} it is mapped to transformation specific for application (set $AppT$ of transformations) and database (set $DbT$).
-These sets are defined in this section.
+The evolution affects both parts of software and as shown in section \ref{sec:evoIntro}.
+Evolution of software is mapped to transformation specific for application (set $AppT$ of transformations) and database (set $DbT$), both sets are defined in this section.
 
@@
 This section provide transformations for manipulating the application layer of a software - the $AppT$ set.
-The set contains not only transformations for model creation, alternation and decomposition, but some operations to obtain information about a model too.
+The set defines common operation for creation, alternation and decomposition of a model (or code) .
 
@@
 \subsection{Database Transformation}
-The database is composed from two separated parts - database schema and data. Transformations are defined to manipulate the database schema and to manipulate stored data.
+The database is composed from two separated parts - database schema and data.
+Transformations are defined to manipulate the database schema and to manipulate stored data.
+Some of presented operations has their origin in SQL queries (especially creation and decomposition), on the other hand some of them cannot be implemented as an SQL query, because of they complexity and conditions verified during theirs execution.
 

### 2013-01-17 doc/description.tex
mapping sets

@@
 \paragraph{Add association} Adding an association into a database is complicated because of already existing instances.
-When an association between classes is added the instances of both classes has to be paired.
-This pairing is provided by the function $f_m$:
-$$f_m : \rightarrow Row \times Row $$
+When an association between classes is added the instances of both classes has to be connected on the database level via foreign key or connection table.
+The information about pairing has to be provided by the user in form of set of instances pairs, because the model does not differ between one-to-one and (one)many-to-many relationships, however both cases are distinguished in the model.
+The set of instances pairs is in this paper called $O_{l_1, l_2, s}$ in case it represents one to one (or many) association or $M_{l_1, l_2, s}$ in case it represents many to many relationship, where $l_1, l_2$ are names of classes which are paired together and $s$ is software which creates the context of instances.
+\begin{align*}
+	O_{l_1, l_2} = <(r_1, r_2)*>,\\ 
+	r_1 \in tableData(l_1, database(s)) \Rightarrow \exists p \in O_{l_1, l_2} \wedge first(p) = r_1 \\
+	r_2 \in tableData(l_2, database(s)) \Rightarrow \exists p \in O_{l_1, l_2} \wedge second(p) = r_2\\
+	\forall p_1, p_2 \in O_{l_1, l_2} : first(p_1) = first(p_2) \Rightarrow p_1 = p_2 \\
+	l_1, l_2 \in Label, r_1, r_2 \lhd Row
+\end{align*}
+\begin{align*}
+	M_{l_1, l_2} =  <(r_1, r_2)*>,\\
+	r_1 \in tableData(l_1, database(s)) \Rightarrow \exists p \in O_{l_1, l_2} \wedge first(p) = r_1 \\
+	r_2 \in tableData(l_2, database(s)) \Rightarrow \exists p \in O_{l_1, l_2} \wedge second(p) = r_2\\
+	l_1, l_2 \in Label, r_1, r_2 \lhd Row
+\end{align*}
+%The features of both sets allows only pairs from existing instances and it assumes there is always at least one pair for each row from both tables.
 
-\textbf{TODO: define the features of the $f_m$ function}
 \begin{align*}

### 2013-01-17 doc/description.tex
updated definition of mapping between instances

--- 2013-01-17 doc/description.tex
@@
 
+\paragraph{Mapping of Instances} The set $M$ which contains all possible mappings between classes, has to be defined to allow several evolutionary operations.
+Each mapping represents an ordered set of pairs of instances.
+These pairs can represent one-to-one (or many) relationship or many-to-many relationship between instances.
+\begin{align*}
+	m_{l_1, l_2} =  <(r_1, r_2)*>,\\
+	r_1 \in tableData(l_1, database(s)) \Rightarrow \exists p \in m_{l_1, l_2} \wedge first(p) = r_1 \\
+	r_2 \in tableData(l_2, database(s)) \Rightarrow \exists p \in m_{l_1, l_2} \wedge second(p) = r_2\\
+	l_1, l_2 \in Label, r_1, r_2 \lhd Row
+\end{align*}
+The definition does not allow to add any default value in case the number of instances does not match.
+It is because the evolution transformations are designed to not affect stored data (instances) by creating a new data which does not origin from the software.
+%The features of both sets allows only pairs from existing instances and it assumes there is always at least one pair for each row from both tables.
+
 %\paragraph{Find Table} The function $findTable$ finds and returns a table in the database according to the provided label.
@@
 
-\paragraph{Copy Value} The function copy values from one column in a table to another column in the same or in another table - according to mapping function $f_m$. 
+\paragraph{Copy Value} The function copy values from one column in a table to another column in the same or in another table - according to provided mapping. 
 \begin{align*}
 	\mathbf{copyValue} : Label \times Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
-	copyValue(l_s, l_t, l_{sc}, l_{tc} f_m, d) = \begin{cases}
+	copyValue(l_s, l_t, l_{sc}, l_{tc} o_{l_s, l_t}, d) = \begin{cases}
 		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc}}{\begin{gathered}
-			insertValues(second(p), \\ [l_{tc},value(findPair(l_{sc}, first(p)))],d)
+		\forall	p \in m_{l_s, l_t} :	d \bullet insertValues(second(p), \\ [l_{tc},value(findPair(l_{sc}, first(p)))])
 		\end{gathered}}
@@
 	\mathbf{copyRows} : Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
-	copyRows(l_s, l_t, f_m, d) = \begin{cases}
- 	\inference{|f_m| = \varnothing}{d} \\\\
- 	\inference{|f_m| = 1 \\ \wedge \exists t_s, t_t \in tables(d) : name(t_s) = l_s, name(t_t) = l_t \\ \wedge t_s \cong t_t }{data(d) \cup insertRow(row(l, pairs(first(f_m))), d)} \\\\
- 	\inference{|f_m| > 1}{\begin{gathered}
-		 d \bullet copyRows(l_s, l_t, head(f_m)) \bullet copyRows(l_s, l_t, tail(f_m))
+	copyRows(l_s, l_t, m_{l_s, l_t}, d) = \begin{cases}
+ 	\inference{|m_{l_s, l_t}| = \varnothing}{d} \\\\
+ 	\inference{|m_{l_s, l_t}| = 1 \\ \wedge \exists t_s, t_t \in tables(d) : name(t_s) = l_s, name(t_t) = l_t \\ \wedge t_s \cong t_t }{d \bullet insertRow(row(l, pairs(first(m_{l_s, l_t}))), d)} \\\\
+ 	\inference{|m_{l_s, l_t}| > 1}{\begin{gathered}
+		 d \bullet copyRows(l_s, l_t, head(M_{l_s, l_t})) \\ \bullet copyRows(l_s, l_t, tail(m_{l_s, l_t}))
 	\end{gathered}}
   \end{cases} \\
-  l_s, l_t \in Label, d \lhd Database
+  l_s, l_t \in Label, d \lhd Database, m_{l_s, l_t} \in M
 \end{align*}
@@
 \paragraph{Add association} Adding an association into a database is complicated because of already existing instances.
-When an association between classes is added the instances of both classes has to be paired.
-This pairing is provided by the function $f_m$:
-$$f_m : \rightarrow Row \times Row $$
-
-\textbf{TODO: define the features of the $f_m$ function}
+When an association between classes is added the instances of both classes has to be connected on the database level via foreign key or connection table.
+The information about pairing has to be provided by the user in form of set of instances pairs, because the model does not contain enough information.
 \begin{align*}
 \mathbf{addAsociation} : Label \times Association \times (Row \times Row) \times State \rightarrow State \\
-\Psi(s, addAsociation(l, a, f_m)) = application(s) \circ addAssoc(l, a) \\
-\Phi(s, addAsociation(l, a, f_m)) = \begin{cases}
-	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l\\ 		\wedge name(t_t) = reference(a) \\ \wedge \forall f \in foreignKeys(t_t) \cuo columns(t_t) : name(f) \neq name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
-		database(s) \bullet addForeignKey(name(t_s), \rho(a,d), d) \\ \bullet \forall r \in f_m(l, reference(a)) : \\ insertRow(row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]), d)
+\Psi(s, addAsociation(l, a, m_{l, reference(a)})) = application(s) \circ addAssoc(l, a) \\
+\Phi(s, addAsociation(l, a, m_{l, reference(a)})) = \begin{cases}
+	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l\\ 		\wedge name(t_t) = reference(a) \\ \wedge \forall f \in foreignKeys(t_t) \cup columns(t_t) \\: name(f) \neq name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
+		database(s) \bullet addForeignKey(name(t_s), \rho(a,d), d) \\ \bullet \forall r \in m_{l, reference(a)} : \\ insertRow(row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]), d)
 	\end{gathered}}
@@
 	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l \\ \wedge name(t_t) = reference(a) \\ \wedge startCardinality(a) > 1}{\begin{gathered} 
-	database(s) \bullet addTable(\rho(a,d),d) \\ \bullet \forall r \in f_m(l, reference(a)) :\\ insertRow(row(name(owningClass(a)) + name(a), \\ [name(t_s), value(primaryKey(first(r)))]\\ [name(t_t), value(primaryKey(second(r)))]))
+	database(s) \bullet addTable(\rho(a,d),d) \\ \bullet \forall r \in m_{l, reference(a)} :\\ insertRow(row(name(owningClass(a)) + name(a), \\ [name(t_s), value(primaryKey(first(r)))]\\ [name(t_t), value(primaryKey(second(r)))]))
 	\end{gathered}} 
@@
 \end{align*}
-
-%\begin{align*}
-%insertClass : Class \times Database \rightarrow Database \\
-%insertClass(c, d) = \inference{ \forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
-%	%\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) \\ 
-%	assocaiations(c) = \varnothing
-%}{
-%\begin{gathered}
-% addTable(\rho(c, d),d) \\ \wedge t \rho(singleProperties(c)) \wedge \rho(collections(c)) \\ 
-%\end{gathered}
-%}
-%\end{align*}
 
@@
 	\mathbf{CopyProperty} : Label \times Label \times Label \times (Row \times Row) \times State \rightarrow State \\
-	\Psi(CopyProperty(l_s, l_t, l_p, f_m), s) = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)} 
+	\Psi(CopyProperty(l_s, l_t, l_p, m_{l_s, l_t}), s) = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)} 
 	\\\\
-	\Phi(CopyProperty(l_s, l_t, l_p, f_m), s) = \begin{cases}
+	\Phi(CopyProperty(l_s, l_t, l_p, m_{l_s, l_t}), s) = \begin{cases}
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) \leq 1 }{\begin{gathered}
-			database(s) \bullet copyColumn(l_s, l_t, l_p) \\ \bullet copyValues(l_s, l_t, l_p, l_p, f_m) 
+			database(s) \bullet copyColumn(l_s, l_t, l_p) \\ \bullet copyValues(l_s, l_t, l_p, l_p, m_{l_s, l_t}) 
 			\end{gathered}} \\ \\
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) > 1}{\begin{gathered}
-			database(s) \bullet copyTable(l_s + l_p, l_t + l_p) \\ \bullet copyRows(l_s + l_p, l_t + l_p, f_m) 
+			database(s) \bullet copyTable(l_s + l_p, l_t + l_p) \\ \bullet copyRows(l_s + l_p, l_t + l_p, m_{l_s, l_t}) 
 			\end{gathered}}
- \end{cases}
+ \end{cases}\\
+ l_s, l_t, l_p \in Label,  m_{l_s, l_t}) \in M, s \lhd Software
 \end{align*}
-
 With the copy property transformation defined we are able to define advanced evolution scenarios.

### 2013-01-17 doc/description.tex
fixed missing context in some transformations

@@
 		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c) \\ \wedge NOTNULL \not \in constraints(c) }{columns(t) \cup c}\\ \\
-		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c) \\ \wedge NOTNULL \in constraints(c) }{columns(t) \cup c}\\ \bullet addValueToTable(l, [name(c), defVal(c)]) \\ \\
+		(\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c) \\ \wedge NOTNULL \in constraints(c) }{columns(t) \cup c})\\ \bullet addValueToTable(l, [name(c), defVal(c)]) \\ \\
 		Inconsistent \;d 
@@
 	copyColumn(l_s, l_t, l_c, d) = \begin{cases}
- 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{state(d) \bullet addColumn(l_t, c)}
+ 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{d \bullet addColumn(l_t, c)}
 	\end{cases} \\
@@
 	insertRow(r, d) = \begin{cases}
-			\inference{|r|>1}{insertRow(head(r), d) \bullet insertRow(tail(r), d)} \\ \\
+			\inference{|r|>1}{d \bullet insertRow(head(r), d) \bullet insertRow(tail(r), d)} \\ \\
 		 	\inference{|r|=1 \wedge \exists t \in tables(d) : name(t) = reference(r) \\ \wedge \forall e \in tableData(t, d) : pkValue(e) \neq pkValue(r)}{data(d) \cup r} \\ \\
@@
 	insertValues(r, p, d) = \begin{cases}
- 		\inference{|p| > 1}{insertValues(r, head(p), d)\bullet insertValues(r, tail(p), d)}  \\ \\
+ 		\inference{|p| > 1}{\begin{gathered}
+			d \bullet insertValues(r, head(p), d) \\ \bullet insertValues(r, tail(p), d)
+	\end{gathered}}
+  \\ \\
  		\inference{|p| = 1 \wedge \forall e \in pairs(r) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p} \\ \\
@@
  		\inference{\exists t \in tables(d) : name(t) = l}	{\begin{gathered}
-		insertValue(head(tableData(l, d)), p, d) \\ \bullet addValueToTable(tail(tableData(l, d)), p, d)
+		d \bullet insertValue(head(tableData(l, d)), p, d) \\ \bullet addValueToTable(tail(tableData(l, d)), p, d)
 	\end{gathered}} \\ \\

### 2013-01-18 doc/description.tex
remove with data

@@
 
+\paragraph{Remove Data and Column} The approach we declare to introduce in this paper was declared to be data and information preserving, nevertheless there are cases when data can be deleted from the database without data lost as they are e.g.
+already copied in another column as the understanding of data changed in the whole software.
+\begin{align*}
+	\mathbf{RemoveColumAndData} : Label \times Label* \times Database \rightarrow Database \\
+	RemoveColumnAndData(l_c, ps, d) = \begin{cases}
+ 	\inference{ps = \varnothing}{d}\\
+ 	\inference{|ps| = 1 \\ \wedge \exists t \in tables(d) \\ \wedge \exists c \in columns(t) : name(c) = l_c}{\begin{gathered}
+	 	(\forall r, r \in tableData(d) \wedge reference(r) = l_c, \\ p \in pairs(r) \wedge name(p) = ps : pairs(r) \ ps) \\ \wedge columns(t) \ c  
+\end{gathered}}\\ \\
+	\inference{|ps| >1}{\begin{gathered}
+		d \bullet RemoveColumnAndData(l_c, head(ps), d) \\ \bullet RemoveColumnAndData(l_c, tail(ps), d) 
+\end{gathered}}
+ \end{cases}
+	\\
+	l_c \in Label, ps \lhd Label*, d \in Database
+\end{align*}
+
 \paragraph{Copy Value} The function copy values from one column in a table to another column in the same or in another table - according to provided mapping. 
@@
 	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l\\ 		\wedge name(t_t) = reference(a) \\ \wedge \forall f \in foreignKeys(t_t) \cup columns(t_t) \\: name(f) \neq name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
-		database(s) \bullet addForeignKey(name(t_s), \rho(a,d), d) \\ \bullet \forall r \in m_{l, reference(a)} : \\ insertRow(row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]), d)
+		database(s) \\ \bullet addForeignKey(name(t_s), \rho(a,d), d) \\ \bullet \forall r \in m_{l, reference(a)} : \\ insertRow(row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]), d)
 	\end{gathered}}
@@
  	l, l_p \in Label, s \lhd Software
+\end{align*}
+
+\paragraph{Remove Property and its Data} The $RemovePropertyAndData$ transformation helps to remove a property together with data stored in the database.
+The evolution can cause data and information lost, to prevent this the transformation $RemovePropertyAndData$ should be used carefully.
+\begin{align*}
+	\mathbf{RemovePropertyAndData} : Label \times Label* \times State \rightarrow State \\
+	\Phi(RemovePropertyAndData(L_c, ps), s) = application(s)\\ \circ remProperty(l_s, ps) \\
+	\Psi(RemovePropertyAndData(L_c, ps), s) = \begin{cases}
+		\inference{ps = \varnothing}{d}\\
+		\inference{|ps| > 1}{\begin{gathered}
+			database(s) \\ \bullet RemovePropertyAndData(L_c, head(ps)) \\ \bullet RemovePropertyAndData(L_c, tail(ps))
+	\end{gathered}}\\\\
+ 		\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) \leq 1}{database(s) \bullet RemoveColumnAndData(l_c, ps)}\\\\
+ 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) > 1}{database(s) \bullet RemoveTableAndData(l_c, ps)}
+ \end{cases}
+	\\
+	l_c \in Label, ps \lhd Label*, s \lhd Software
 \end{align*}
@@
     l_c, l_a \in Labels, s \lhd Software
+\end{align*}
+
+\paragraph{Remove Association and its Data} The $RemoveAssociationAndData$ transformation helps to remove a property together with data stored in the database.
+The evolution can cause data and information lost, to prevent this the transformation $RemoveAssociationAndData $ should be used carefully.
+\begin{align*}
+	\mathbf{RemoveAssociationAndData} : Label \times Label* \times State \rightarrow State \\
+	\Phi(RemoveAssociationAndData(L_c, ps), s) = application(s)\\ \circ remAssociation(l_s, ps) \\
+	\Psi(RemoveAssociationAndData(L_c, ps), s) = \begin{cases}
+		\inference{ps = \varnothing}{d}\\
+		\inference{|ps| > 1}{\begin{gathered}
+			database(s) \\ \bullet RemoveAssociationAndData(L_c, head(ps)) \\ \bullet RemoveAssociationAndData(L_c, tail(ps))
+	\end{gathered}}\\\\
+ 		\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists a \in associations(c) : name(a) = l \\ \wedge startCardinality(p) \leq 1}{database(s) \bullet RemoveForeignKeyAndData(reference(a), ps)}\\\\
+ 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists a \in associations(c) : name(a) = l \\ \wedge startCardinality(a) > 1}{database(s) \bullet RemoveAssociationAndData(l_c + reference(a), ps)}
+ \end{cases}
+	\\
+	l_c \in Label, ps \lhd Label*, s \lhd Software
 \end{align*}
@@
 
-\section{Refactoring Cases}
+\section{Complex Refactoring Cases}
+The concrete refactoring cases presented in this section were chosen according to the refactoring statistics provided by Eclipse.
+The refactoring cases, which does not affect data (such as change method signature or encapsulate field) were removed from the set of refactoring cases and form the rest the Move property and Extract/Inline class were chosen to
 \subsection{Move Property}
+The aim of the move property transformation is to change two classes - one will get a new property, the other will lost a property.
+An example of such transformation is merging information from one class to another.
+The implementation of move property transformation is based on copy property transformation:
+\begin{align*}
+	\mathbf{MoveProperty} : Label \times Label \times Label* \times (Row \times Row) \times State \rightarrow State \\
+	s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) = \begin{cases}
+ 	\inference{|ps| = \varnothing}{state(s)}\\\\
+ 	\inference{|ps| = 1}{\begin{gathered} s \oplus CopyProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemovePropertyData(l_s, ps) \\ \oplus RemoveProperty(l_s, ps) 
+\end{gathered}}\\\\
+	\inference{|ps| > 1}{\begin{gathered}
+		s \oplus MoveProperty(l_s, l_t, head(ps), m_{l_s, l_t}) \\ \oplus MoveProperty(l_s, l_t, tail(ps), m_{l_s, l_t}) 
+\end{gathered}}
+ \end{cases}
+	\\\\
+	l_s, l_t \in Label, ps \in Label*,  m_{l_s, l_t} \in M, s \lhd Software
+\end{align*}
+Although the $MoveProperty$ transformation contains the transformation $RemovePropertyData$ its execution preserve data and information as the information has been already copied.
+
 \subsection{Extract and Inline Class}
-
-\section{Assigment for NII Internship / Future Work}
+The extract and inline class are two opposite transformations. First of them extract a class from existing class, second of them merges two classes into one.
+\begin{align*}
+	\mathbf{ExtractClass} : Label \times Label* \times Label \times State \rightarrow State \\
+	state(s) \oplus ExtractClass(l_s, ps, l_n) = \begin{cases}
+	\inference{|ps| = 1}{\begin{gathered}
+		s \oplus addClass(cl(l_n, \varnothing, \varnothing)) \\ \oplus MoveProperty(l_s, l_n, ps, \mathbf{X}) \\ \oplus RemovePropertyData(l_s, ps) \\ \oplus RemoveProperty(l_s, ps)  
+\end{gathered}}\\\\
+	\inference{|ps| > 1}{\begin{gathered}
+		s \oplus ExtractClass(l_s, head(ps), l_n) \\\oplus ExtractClass(l_s, tail(ps), l_n)
+\end{gathered}}\\\\
+	\inference{|ps| = \varnothing}{state(s)}	 
+ \end{cases}
+	\\
+	l_s, l_n \in Label, ps \lhd Label* s \lhd Software 
+\end{align*}
+
+\begin{align*}
+	\mathbf{InlineClass} : Label \times Label \times State \rightarrow State \\
+	state(s) \oplus InlineClss(l_s, l_t, m_{l_s, l_t}) = s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemovePropertyData(l_s, ps) \\ \oplus RemoveProperty(l_s, ps)  
+\\
+l_s, l_t \in Label, m_{l_s, l_t} \in M, s \lhd Software, \\ \\
+ps = \inference{\exists c \in classes(applications)) \wedge ps = \varnothing}{\forall p \in properries(p) : ps \cup name(p)}
+\end{align*}
+
 \section{Conclusion}
+The evolution of software, which uses object oriented language together with relational database is complicated because of separate definition of evolution for application and database and possibility of data and information lost.
+The formal description of a framework for easier software evolution was introduced in this paper.
+
+The main contribution of the paper is definition of connection between application refactoring and database migration.
+It allows to define the evolution only once for both parts of the software and it is defined in a way it prevents data or information lost.
+The definition is based on small transformations of each part, which are deigned to preserve data and theirs concatenation into complex evolutionary transformations.
+
+Introduced formal definitions provides an important foundation for development of an evolutionary framework for concrete object-oriented language and relational database.
+
+\newpage
+\appendix
+\section{Assigment for NII Internship}
+The model of evolution in ORM software defined in this paper will be improved and validated during internship at NII. The main objectives of the internship are:
+\begin{enumerate}
+	\item \label{it:1} Verify the model by an implementation in the Maude tool. \emph{(Estimated time: 1 month)}
+	\item \label{it:2} More complex examples of refactoring cases should be examined with existing model.  \emph{(Estimated time: 0.5 month)}
+	\item \label{it:3} The inheritance relationship between classes should be implemented in the model together with inheritance-based refactoring cases. \emph{(Estimated time: 0.5 month)}
+	\item Prepare an article, which will contain topics \ref{it:1}, \ref{it:2} and \ref{it:3}. \emph{(Estimated time: 1 month)}
+\end{enumerate}
 

### 2013-01-19 doc/description.tex
textreview

@@
 
-There are three main goals of our evolutionary framework: 1) provide one source of evolution definition for both: code and database 2) provide a tool, which will be able to preserve stored data 3) minimize the manual work during data evolution in a software.
+There are three main goals of our evolutionary framework: 1) provide one source of evolution definition for both: code and database 2) provide a tool, which will be able to preserve stored data and information.
 
@@
 \end{align*}
+
 The function querying the state (it means consistency) of a software is defined as:
@@
 Next there is a set $Values$ containing all possible values, which can be used in the model.
+The notation: $Type*$ is used to define an ordered set of values of concrete type (e.g.
+$C \lhd Class*$ means c is an ordered set of classes).
+
+A lot of introduced transformations can be applied on an ordered set. The application proceeds on set consists in sequential application on set elements:
+\begin{align*}
+	f(x, y) = \begin{cases}
+ 		\inference{x = \varnothing}{y}\\
+ 		\inference{|x| > 1}{f(head(x), y) \cup f(tail(x), y}\\\\
+ 		\inference{|x| = 1}{effect}
+ \end{cases}\\
+ y \lhd Software \vee y \lhd Application \vee y \lhd Database
+\end{align*}
+where $x$ is a set of parameters. The effect of a transformation is provided only in next sections as the repetition of sequential application will be redundant.
 
@@
 	tableData(l, r) = \begin{cases}
- 		\inference{r = \varnothing}{\varnothing} \\
- 		\inference{|r| > 1}{tableData(head(r)) \cup tableData(tail(r))}\\ \\
+ 		%\inference{r = \varnothing}{\varnothing} \\
+ 		%\inference{|r| > 1}{tableData(head(r)) \cup tableData(tail(r))}\\ \\
  		\inference{|r| = 1 \wedge refTable(r) = l}{r}\\\\
@@
 	\mathbf{\rho_{app}} : Application \rightarrow Database\\
-	\rho_{app}(a) = \inference{state(a) = Consistent a}{\begin{gathered}
+	\rho_{app}(a) = \inference{state(a) = Consistent \; a}{\begin{gathered}
 	db(\varnothing) \bullet addTable(classes(a)) \\ 
@@
 	\rho_{c}(c) = \begin{cases}
-		\inference{c = \varnothing}{\varnothing}\\\\
-		\inference{|c| = 1}{table(name(c), primaryKey("Id"), \varnothing, \varnothing)}  \\ \\
-		\inference{|c| > 1}{\rho_c(head(c)) \cup \rho_c(tail(c))}
- \end{cases}\\
+		%\inference{c = \varnothing}{\varnothing}\\\\
+		%\inference{|c| > 1}{\rho_c(head(c)) \cup \rho_c(tail(c))} \\ \\
+		\inference{|c| = 1}{table(name(c), primaryKey("Id"), \varnothing, \varnothing)}  \end{cases}   \\ 	
  c \lhd Class
@@
 	\rho_{sp}(p) = \begin{cases}
-		\inference{p = \varnothing}{\varnothing}\\
-		\inference{|p| > 1}{\rho_{sp}(head(p)) \cup \rho_{sp}(tail(p))}\\\\
+		%\inference{p = \varnothing}{\varnothing}\\
+		%\inference{|p| > 1}{\rho_{sp}(head(p)) \cup \rho_{sp}(tail(p))}\\\\
  		\inference{cardinality(p) = 1 \wedge !\;mandatory(p)
@@
 	\rho_{cp}(p) = \begin{cases}
-		\inference{p = \varnothing}{\varnothing}\\
-		\inference{|p| > 1}{\rho_{cp}(head(p)) \cup \rho_{cp}(tail(p))}\\\\
+		%\inference{p = \varnothing}{\varnothing}\\
+		%\inference{|p| > 1}{\rho_{cp}(head(p)) \cup \rho_{cp}(tail(p))}\\\\
      	\inference{cardinality(p) > 1 \wedge !\;mandatory(p) \\ 
@@
 	\rho_a(a) = \begin{cases}
-		\inference{a = \varnothing}{\varnothing}\\
-		\inference{|a| > 1}{\rho_{a}(head(a)) \cup \rho_{a}(tail(a))}\\\\
+		%\inference{a = \varnothing}{\varnothing}\\
+		%\inference{|a| > 1}{\rho_{a}(head(a)) \cup \rho_{a}(tail(a))}\\\\
 		\inference{ startCardinality(a) = 0 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)
@@
 	\rho_a(a) = \begin{cases}
- 		\inference{a = \varnothing}{\varnothing}\\
-		\inference{|a| > 1}{\rho_{a}(head(a)) \cup \rho_{a}(tail(a))}\\\\
+ 		%\inference{a = \varnothing}{\varnothing}\\
+		%\inference{|a| > 1}{\rho_{a}(head(a)) \cup \rho_{a}(tail(a))}\\\\
 		\inference{  startCardinality > 1 \\ \wedge \exists t \in tables(d) : name(owningClass(p)) = name(t) \\ \wedge \exists e \in tables(d) : name(reference(a)) = name(e)
@@
 This section provide transformations for manipulating the application layer of a software - the $AppT$ set.
-The set defines common operation for creation, alternation and decomposition of a model (or code) .
+The set defines common operation for creation, alternation and decomposition of a model (or code).
 
@@
 		p.name = l_n 
-	\end{gathered}}
+	\end{gathered}}\\
+	l_c, l_o, l_n \in Label, a \lhd Application
 \end{align*}
@@
 	\end{gathered}}
+	l_c, l_o, l_n \in Label, a \lhd Application
 \end{align*}
@@
 	}{\begin{gathered}
-		c.name = l_n \wedge \forall e \in classes(a) \\ \wedge \forall (f \in associations(e) \wedge reference(f) = l_o ) :\\ f.reference = l_n 
+		name(c) = l_n \wedge \forall e \in classes(a) \\ \wedge \forall (f \in associations(e) \wedge reference(f) = l_o ) :\\ f.reference = l_n 
 	\end{gathered}}
@@
 
-\paragraph{Compare} The need to compare two tables during the evolution results into the comparing operator $\cong$. The operator has to be defined for columns and foreign keys too.
+\paragraph{Comparison of tables} The need to compare two tables during the evolution results into the comparing operator $\cong$. The operator has to be defined for columns and foreign keys too.
 \begin{align*}
@@
 		false
- \end{cases}
+ \end{cases}\\
+ f_1 \lhd ForeignKey, f_2 \lhd ForeignKey
 \end{align*}
@@
 \end{align*}
-
+\subsubsection{Database Schema Creation}
 \paragraph{Add Column} A new column can be add into a table if there are no name collisions.
@@
 	addColum(l, c, d) = \begin{cases}
-		\inference{|c| > 1}{addColumn(l, head(c), d) \bullet addColumn(l, tail(c), d)} \\ \\
+		%\inference{|c| > 1}{addColumn(l, head(c), d) \bullet addColumn(l, tail(c), d)} \\ \\
 		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c) \\ \wedge NOTNULL \not \in constraints(c) }{columns(t) \cup c}\\ \\
-		(\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c) \\ \wedge NOTNULL \in constraints(c) }{columns(t) \cup c})\\ \bullet addValueToTable(l, [name(c), defVal(c)]) \\ \\
-		Inconsistent \;d 
+		(\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c) \\ \wedge NOTNULL \in constraints(c) }{columns(t) \cup c})\\ \bullet addValueToTable(l, [name(c), defVal(c)])
+%		\\ \\ Inconsistent \;d 
 	 \end{cases}\\ 
@@
 
+\paragraph{Add Table} The transformation creates a new table in the database. The new table should not contain any foreign keys.
+\begin{align*}
+	\mathbf{addTable} : Table* \times Database \rightarrow Database \\ 
+	addTable(t, d) = \begin{cases}
+		%\inference{|t| > 1}{d \bullet addTable(head(t)) \bullet addTable(tail(t))}
+\\ \\	
+		\inference{|t| = 1 \wedge \forall e \in tables(d) \wedge name(t) \neq name(e) \\ \wedge \forall f \in foreignKeys(t) : \\ s\exists e \in tables(d) \wedge name(e) = reference(f) }{tables(d) \cup t}
+% \\ \\ 	Inconsistent \;d 
+	 \end{cases}\\ 
+	t = Table*, d \lhd Database
+\end{align*}
+
+\subsubsection{Database Schema Alternation}
 \paragraph{Alter Column Name}
@@
 l_c, l_o, l_n \in Label, d \lhd Database
-\end{align*}
-
-\paragraph{Add Table} The transformation creates a new table in the database. The new table should not contain any foreign keys.
-\begin{align*}
-	\mathbf{addTable} : Table* \times Database \rightarrow Database \\ 
-	addTable(t, d) = \begin{cases}
-		\inference{|t| > 1}{d \bullet addTable(head(t)) \bullet addTable(tail(t))}
-\\ \\	
-		\inference{|t| = 1 \wedge \forall e \in tables(d) \wedge name(t) \neq name(e) \\ \wedge \forall f \in foreignKeys(t) : \\ s\exists e \in tables(d) \wedge name(e) = reference(f) }{tables(d) \cup t}
-\\ \\
-		Inconsistent \;d 
-	 \end{cases}\\ 
-	t = Table*, d \lhd Database
 \end{align*}
@@
 
-\paragraph{Copy Table} The transformation creates a copy of a table in database schema with different name.
-\begin{align*}
-	\mathbf{copyTable} : Label \times Label \times Database \times Database \\
-	copyTable(l_s, l_t, d) = \inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \forall e \in tables(d) : name(e) \neq l_t}{d \bullet addTable(tab(l_t, primaryKey(t), columns(t), foreignKeys(t)))}
-\end{align*}
-
 \paragraph{Copy Column} The transformation creates a copy of a column in a another table in the database. The transformation changes only the structure of database - stored data are not changed.
@@
 
+\paragraph{Copy Table} The transformation creates a copy of a table in database schema with different name.
+\begin{align*}
+	\mathbf{copyTable} : Label \times Label \times Database \times Database \\
+	copyTable(l_s, l_t, d) = \inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \forall e \in tables(d) : name(e) \neq l_t}{d \bullet addTable(tab(l_t, primaryKey(t), columns(t), foreignKeys(t)))}
+\end{align*}
+
+\subsubsection{Data Manipulation}
 \paragraph{Insert row} The function allow to insert a row(s) into a database.
@@
 	insertRow(r, d) = \begin{cases}
-			\inference{|r|>1}{d \bullet insertRow(head(r), d) \bullet insertRow(tail(r), d)} \\ \\
-		 	\inference{|r|=1 \wedge \exists t \in tables(d) : name(t) = reference(r) \\ \wedge \forall e \in tableData(t, d) : pkValue(e) \neq pkValue(r)}{data(d) \cup r} \\ \\
-			 Inconsistent \; d
+			%\inference{|r|>1}{d \bullet insertRow(head(r), d) \bullet insertRow(tail(r), d)} \\ \\
+		 	\inference{|r|=1 \wedge \exists t \in tables(d) : name(t) = reference(r) \\ \wedge \forall e \in tableData(t, d) : pkValue(e) \neq pkValue(r)}{data(d) \cup r} 
+		 	%\\ \\	 Inconsistent \; d
 		 \end{cases}	\\
@@
 	insertValues(r, p, d) = \begin{cases}
- 		\inference{|p| > 1}{\begin{gathered}
-			d \bullet insertValues(r, head(p), d) \\ \bullet insertValues(r, tail(p), d)
-	\end{gathered}}
-  \\ \\
- 		\inference{|p| = 1 \wedge \forall e \in pairs(r) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p} \\ \\
-	 	 Inconsistent \; d
+ 		%\inference{|p| > 1}{\begin{gathered} 	d \bullet insertValues(r, head(p), d) \\ \bullet insertValues(r, tail(p), d)\end{gathered}}   \\ \\
+ 		\inference{|p| = 1 \wedge \forall e \in pairs(r) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p}
+ 		%\\ \\ Inconsistent \; d
  		\end{cases} \\
@@
 
-\paragraph{Remove Data and Column} The approach we declare to introduce in this paper was declared to be data and information preserving, nevertheless there are cases when data can be deleted from the database without data lost as they are e.g.
-already copied in another column as the understanding of data changed in the whole software.
-\begin{align*}
-	\mathbf{RemoveColumAndData} : Label \times Label* \times Database \rightarrow Database \\
-	RemoveColumnAndData(l_c, ps, d) = \begin{cases}
- 	\inference{ps = \varnothing}{d}\\
- 	\inference{|ps| = 1 \\ \wedge \exists t \in tables(d) \\ \wedge \exists c \in columns(t) : name(c) = l_c}{\begin{gathered}
-	 	(\forall r, r \in tableData(d) \wedge reference(r) = l_c, \\ p \in pairs(r) \wedge name(p) = ps : pairs(r) \ ps) \\ \wedge columns(t) \ c  
-\end{gathered}}\\ \\
-	\inference{|ps| >1}{\begin{gathered}
-		d \bullet RemoveColumnAndData(l_c, head(ps), d) \\ \bullet RemoveColumnAndData(l_c, tail(ps), d) 
-\end{gathered}}
- \end{cases}
-	\\
-	l_c \in Label, ps \lhd Label*, d \in Database
-\end{align*}
-
 \paragraph{Copy Value} The function copy values from one column in a table to another column in the same or in another table - according to provided mapping. 
@@
 	copyRows(l_s, l_t, m_{l_s, l_t}, d) = \begin{cases}
- 	\inference{|m_{l_s, l_t}| = \varnothing}{d} \\\\
+ 	%\inference{|m_{l_s, l_t}| = \varnothing}{d} \\\\
+ 	%\inference{|m_{l_s, l_t}| > 1}{\begin{gathered} d \bullet copyRows(l_s, l_t, head(M_{l_s, l_t})) \\ \bullet copyRows(l_s, l_t, tail(m_{l_s, l_t})) \end{gathered}}
  	\inference{|m_{l_s, l_t}| = 1 \\ \wedge \exists t_s, t_t \in tables(d) : name(t_s) = l_s, name(t_t) = l_t \\ \wedge t_s \cong t_t }{d \bullet insertRow(row(l, pairs(first(m_{l_s, l_t}))), d)} \\\\
- 	\inference{|m_{l_s, l_t}| > 1}{\begin{gathered}
-		 d \bullet copyRows(l_s, l_t, head(M_{l_s, l_t})) \\ \bullet copyRows(l_s, l_t, tail(m_{l_s, l_t}))
-	\end{gathered}}
   \end{cases} \\
@@
 
+\subsubsection{Database decomposition}
 \paragraph{Remove table} The function removes a table if there are no stored data.
@@
 \end{align*}
+
+\paragraph{Remove Data and Column} The approach we declare to introduce in this paper was declared to be data and information preserving, nevertheless there are cases when data can be deleted from the database without data lost as they are e.g.
+already copied in another column as the understanding of data changed in the whole software.
+\begin{align*}
+	\mathbf{RemoveColumAndData} : Label \times Label* \times Database \rightarrow Database \\
+	RemoveColumnAndData(l_c, ps, d) = \begin{cases}
+ 	%\inference{ps = \varnothing}{d}\\
+ 	 %\inference{|ps| >1}{\begin{gathered} 	d \bullet RemoveColumnAndData(l_c, head(ps), d) \\ \bullet RemoveColumnAndData(l_c, tail(ps), d) \end{gathered}} \\ \\
+ 	\inference{|ps| = 1 \\ \wedge \exists t \in tables(d) \\ \wedge \exists c \in columns(t) : name(c) = l_c}{\begin{gathered}
+	 	(\forall r, r \in tableData(d) \wedge reference(r) = l_c, \\ p \in pairs(r) \wedge name(p) = ps : pairs(r) \ ps) \\ \wedge columns(t) \ c  
+\end{gathered}}
+ \end{cases}
+	\\
+	l_c \in Label, ps \lhd Label*, d \in Database
+\end{align*}
+
 \begin{align*}
@@
 	\Psi(RemovePropertyAndData(L_c, ps), s) = \begin{cases}
-		\inference{ps = \varnothing}{d}\\
-		\inference{|ps| > 1}{\begin{gathered}
-			database(s) \\ \bullet RemovePropertyAndData(L_c, head(ps)) \\ \bullet RemovePropertyAndData(L_c, tail(ps))
-	\end{gathered}}\\\\
+		%\inference{ps = \varnothing}{d}\\
+		%\inference{|ps| > 1}{\begin{gathered} database(s) \\ \bullet RemovePropertyAndData(L_c, head(ps)) \\ \bullet RemovePropertyAndData(L_c, tail(ps)) 	\end{gathered}}\\\\
  		\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) \leq 1}{database(s) \bullet RemoveColumnAndData(l_c, ps)}\\\\
@@
 	\Psi(RemoveAssociationAndData(L_c, ps), s) = \begin{cases}
-		\inference{ps = \varnothing}{d}\\
-		\inference{|ps| > 1}{\begin{gathered}
-			database(s) \\ \bullet RemoveAssociationAndData(L_c, head(ps)) \\ \bullet RemoveAssociationAndData(L_c, tail(ps))
-	\end{gathered}}\\\\
+		%\inference{ps = \varnothing}{d}\\
+		%\inference{|ps| > 1}{\begin{gathered}			database(s) \\ \bullet RemoveAssociationAndData(L_c, head(ps)) \\ \bullet RemoveAssociationAndData(L_c, tail(ps)) \end{gathered}}\\\\
  		\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists a \in associations(c) : name(a) = l \\ \wedge startCardinality(p) \leq 1}{database(s) \bullet RemoveForeignKeyAndData(reference(a), ps)}\\\\
- 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists a \in associations(c) : name(a) = l \\ \wedge startCardinality(a) > 1}{database(s) \bullet RemoveAssociationAndData(l_c + reference(a), ps)}
+ 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists a \in associations(c) : name(a) = l \\ \wedge startCardinality(a) > 1}{database(s) \bullet RemoveTableAndData(l_c + reference(a), ps)}
  \end{cases}
@@
 	s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) = \begin{cases}
- 	\inference{|ps| = \varnothing}{state(s)}\\\\
- 	\inference{|ps| = 1}{\begin{gathered} s \oplus CopyProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemovePropertyData(l_s, ps) \\ \oplus RemoveProperty(l_s, ps) 
+ 	%\inference{|ps| = \varnothing}{state(s)}\\\\
+ 		%\inference{|ps| > 1}{\begin{gathered} 	s \oplus MoveProperty(l_s, l_t, head(ps), m_{l_s, l_t}) \\ \oplus MoveProperty(l_s, l_t, tail(ps), m_{l_s, l_t}) \end{gathered}}
+ 	\inference{|ps| = 1}{\begin{gathered} s \oplus CopyProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemovePropertyAndData(l_s, ps)
 \end{gathered}}\\\\
-	\inference{|ps| > 1}{\begin{gathered}
-		s \oplus MoveProperty(l_s, l_t, head(ps), m_{l_s, l_t}) \\ \oplus MoveProperty(l_s, l_t, tail(ps), m_{l_s, l_t}) 
-\end{gathered}}
  \end{cases}
-	\\\\
+	\\
 	l_s, l_t \in Label, ps \in Label*,  m_{l_s, l_t} \in M, s \lhd Software
@@
 \begin{align*}
-	\mathbf{ExtractClass} : Label \times Label* \times Label \times State \rightarrow State \\
-	state(s) \oplus ExtractClass(l_s, ps, l_n) = \begin{cases}
+	\mathbf{Extract} : Label \times Label* \times Label \times State \rightarrow State \\
+	state(s) \oplus Extract(l_s, ps, l_n) = \begin{cases}
 	\inference{|ps| = 1}{\begin{gathered}
@@
 \end{gathered}}\\\\
-	\inference{|ps| > 1}{\begin{gathered}
-		s \oplus ExtractClass(l_s, head(ps), l_n) \\\oplus ExtractClass(l_s, tail(ps), l_n)
-\end{gathered}}\\\\
-	\inference{|ps| = \varnothing}{state(s)}	 
+	%\inference{|ps| > 1}{\begin{gathered} 	s \oplus Extract(l_s, head(ps), l_n) \\\oplus Extract(l_s, tail(ps), l_n) \end{gathered}}\\\\
+	% \inference{|ps| = \varnothing}{state(s)}	 
  \end{cases}
@@
 \begin{align*}
-	\mathbf{InlineClass} : Label \times Label \times State \rightarrow State \\
-	state(s) \oplus InlineClss(l_s, l_t, m_{l_s, l_t}) = s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemovePropertyData(l_s, ps) \\ \oplus RemoveProperty(l_s, ps)  
-\\
+	\mathbf{Inline} : Label \times Label \times State \rightarrow State \\
+	state(s) \oplus Inline(l_s, l_t, m_{l_s, l_t}) = \inference{\exists c \in classes(application(s)) : name(c) = l_s \\\wedge !\:associated(l_s, s) \\ \wedge foreignKeys(c) = \varnothing  }{\begin{gathered}
+	s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemovePropertyAndData(l_s, ps) \\ \oplus RemoveClass(l_s)  
+\end{gathered}
+}\\
 l_s, l_t \in Label, m_{l_s, l_t} \in M, s \lhd Software, \\ \\

### 2013-01-19 doc/description.tex
dropX with data

@@
  		\inference{|x| = 1}{effect}
+ 		\inference{otherwise}{Inconsistent y}
  \end{cases}\\
@@
 \end{align*}
-where $x$ is a set of parameters. The effect of a transformation is provided only in next sections as the repetition of sequential application will be redundant.
+where $x$ is a set of parameters.
+The effect of a transformation is provided only in next sections as the repetition of sequential application will be redundant.
+The $otherwise$ branch is consider implicit for all introduced transformations and thus it is not part of transformation definitions too.
 
@@
 	\end{gathered}}
-	l_c, l_o, l_n \in Label, a \lhd Application
+	\\ l_c, l_o, l_n \in Label, a \lhd Application
 \end{align*}
@@
 \subsubsection{Database decomposition}
+\paragraph{dropValue}
+\begin{align*}
+	\mathbf{dropValue} : Label \times Row \times Database \rightarrow Database \\
+	dropValue(l, r, d) = \inference{\exists p \in pairs(r) : reference(p) = l}{pairs(r) \setminus p}
+	\\ l \in Label, r \lhd Row \wedge r \in data(d), d \lhd Database
+\end{align*}
+
+\begin{align*}
+	\mathbf{dropRow} :  Row*  \times Database \rightarrow Database\\
+	dropRow(r, d) = \inference{|r|=1}{data(d) \setminus r}
+	\\ l \in Label, r \lhd Row \wedge r \in data(d), d \lhd Database
+\end{align*}
+
 \paragraph{Remove table} The function removes a table if there are no stored data.
@@
 	\mathbf{dropTable} : Label \times Database \rightarrow Database \\
-	dropTable(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \not \exists r \in data(d) : owningTable(r) = l \\  \not \exists j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) = l }{ tables(d) \setminus \rho(c)}
+	dropTable(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \forall r \in data(d) : owningTable(r) \neq l \\  \wedge (\forall j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) \neq l) }{ tables(d) \setminus \rho(c)}
+\end{align*}
+
+\begin{align*}
+	\mathbf{dropTableAndData} : Labe \times Database \rightarrow Database \\
+	dropTableAndData(l, d) = \inference{\exists t \in tables(d) \\ \wedge ! \; associated(t, d)}{\begin{gathered}
+	(\forall r \in tableData(t,d) : dropRow(r, d))\\ \bullet dropTable(t,d)
+\end{gathered}}
+	\\ l \in Label, d \lhd Database
 \end{align*}
@@
 
+\begin{align*}
+	\mathbf{dropKey} : Label \times Label \times Database \rightarrow Database \\
+	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) \setminus c}\\
+	 l, l_p \in Label, d \lhd Database 
+\end{align*}
+
 \paragraph{Remove Data and Column} The approach we declare to introduce in this paper was declared to be data and information preserving, nevertheless there are cases when data can be deleted from the database without data lost as they are e.g.
@@
 \begin{align*}
-	\mathbf{RemoveColumAndData} : Label \times Label* \times Database \rightarrow Database \\
-	RemoveColumnAndData(l_c, ps, d) = \begin{cases}
+	\mathbf{DropColumnAndData} : Label \times Label* \times Database \rightarrow Database \\
+	DropColumnAndData(l_c, ps, d) = \begin{cases}
  	%\inference{ps = \varnothing}{d}\\
- 	 %\inference{|ps| >1}{\begin{gathered} 	d \bullet RemoveColumnAndData(l_c, head(ps), d) \\ \bullet RemoveColumnAndData(l_c, tail(ps), d) \end{gathered}} \\ \\
- 	\inference{|ps| = 1 \\ \wedge \exists t \in tables(d) \\ \wedge \exists c \in columns(t) : name(c) = l_c}{\begin{gathered}
-	 	(\forall r, r \in tableData(d) \wedge reference(r) = l_c, \\ p \in pairs(r) \wedge name(p) = ps : pairs(r) \ ps) \\ \wedge columns(t) \ c  
-\end{gathered}}
+ 	 %\inference{|ps| >1}{\begin{gathered} 	d \bullet DropColumnAndData(l_c, head(ps), d) \\ \bullet DropColumnAndData(l_c, tail(ps), d) \end{gathered}} \\ \\
+ 	(\inference{|ps| = 1 \\ \wedge \exists t \in tables(d) \\ \wedge \exists c \in columns(t) : name(c) = l_c}{\begin{gathered}
+	 	(\forall r, r \in tableData(d) \wedge reference(r) = l_c: \\d \bullet dropValue(ps, r)) \end{gathered}}) \\ \bullet dropColumn(l_c)  
  \end{cases}
@@
 \begin{align*}
-	\mathbf{dropKey} : Label \times Label \times Database \rightarrow Database \\
-	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) \setminus c}\\
-	 l, l_p \in Label, d \lhd Database 
+	\mathbf{DropKeyAndData} : Label \times Label* \times Database \rightarrow Database \\
+	DropKeyAndData(l_f, ps, d) = \begin{cases}
+ 	%\inference{ps = \varnothing}{d}\\
+ 	 %\inference{|ps| >1}{\begin{gathered} 	d \bullet DropColumnAndData(l_c, head(ps), d) \\ \bullet DropColumnAndData(l_c, tail(ps), d) \end{gathered}} \\ \\
+ 	(\inference{|ps| = 1 \\ \wedge \exists t \in tables(d) \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_f}{\begin{gathered}
+	 	(\forall r, r \in tableData(d) \wedge reference(r) = l_f: \\d \bullet dropValue(ps, r)) \end{gathered}}) \\ \bullet dropKey(l_f)  
+ \end{cases}
+	\\
+	l_f \in Label, ps \lhd Label*, d \in Database
 \end{align*}
@@
 		%\inference{|ps| > 1}{\begin{gathered} database(s) \\ \bullet RemovePropertyAndData(L_c, head(ps)) \\ \bullet RemovePropertyAndData(L_c, tail(ps)) 	\end{gathered}}\\\\
- 		\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) \leq 1}{database(s) \bullet RemoveColumnAndData(l_c, ps)}\\\\
+ 		\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) \leq 1}{database(s) \bullet DropColumnAndData(l_c, ps)}\\\\
  	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) > 1}{database(s) \bullet RemoveTableAndData(l_c, ps)}

### 2013-01-21 doc/description.tex
small fixes

@@
 
-The paper is organized as follows: in Section \ref{sec:evoIntro} the concept of ORM software and its evolution is discussed, then models of application and database are introduced in Section \ref{sec:appModel} and in Section \ref{sec:dbModel}, the ORM is defined in Section \ref{sec:orm} and the evolutionary transformations are defined in Section \ref{sec:eorm}.
+The paper is organized as follows: in Section \ref{sec:evoIntro} the concept of ORM software and its evolution is discussed, then models of application and database are introduced in Section \ref{sec:appModel} and in Section \ref{sec:dbModel}, the ORM is defined in Section \ref{sec:orm} and the evolutionary transformations are defined in Section \ref{sec:Evolutionaty-Transformations}.
 
@@
 \begin{align*}
- 	\mathbf{state} = Consistent\; s \: | \: Inconsistent \; s,  \\	s \lhd Software \vee s \lh Application \vee s \lhd Database
+ 	\mathbf{state} = Consistent\; s \: | \: Inconsistent \; s,  \\	s \lhd State \vee s \lh Application \vee s \lhd Database
 \end{align*}
@@
  	\end{cases} \\ 
- 	s \lhd Software
+ 	s \lhd State
 \end{align*}
@@
 \end{cases}\\ 
-s \lhd Software, t \in E
+s \lhd State, t \in E
 \end{align*}
@@
  \end{cases} \\ \\
-	 c \in classes(application(s)), s \lhd Software 
+	 c \in classes(application(s)), s \lhd State 
 \end{align*}
@@
 		d \bullet insertValue(head(tableData(l, d)), p, d) \\ \bullet addValueToTable(tail(tableData(l, d)), p, d)
-	\end{gathered}} \\ \\
- 		Inconsistent \; d
- 	\end{cases}\\ \\
+	\end{gathered}} 
+ 	\end{cases}\\ 
  	l \in Label, p \lhd Pair,  d \lhd Database
@@
 \subsubsection{Database decomposition}
-\paragraph{dropValue}
+\paragraph{Remove a value}  The approach introduced in this paper has to  preserve data and information, nevertheless there are cases when data can be deleted from the database without data lost as they are e.g.
+already copied in another column as the understanding of data changed in the whole software, therefore we defined two transformations - one removes a single value from a row, second removes whole row (instance) from the software.
 \begin{align*}
@@
 
-\paragraph{Remove table} The function removes a table if there are no stored data.
+\paragraph{Remove table, column or key} The functions removes a table, column or key if there are no stored data.
 This is because the data lost is undesirable in case of evolution.
-The consistency of foreign keys references has to be verified too.
+The consistency of foreign keys references has to be verified too in case a table is removed.
+There are defined two variants of remove transformations for each model entity - one preserves data, the other removes data too.
+
 \begin{align*}
@@
 \end{align*}
-
-\paragraph{Remove column or foreign key} The situation is similar to the $dropTable$ function - there has to be no data in the column or key so it can be safely removed.
 \begin{align*}
@@
 \end{align*}
-
-\paragraph{Remove Data and Column} The approach we declare to introduce in this paper was declared to be data and information preserving, nevertheless there are cases when data can be deleted from the database without data lost as they are e.g.
-already copied in another column as the understanding of data changed in the whole software.
 \begin{align*}
@@
 \section{Evolutionaty Transformations}
+\label{sec:Evolutionaty-Transformations}
 The set of all possible data evolution cases $E$ was defined in Section \ref{sec:evoIntro}.
@@
    \end{cases}\\
-	s \lhd Software, l \in Label, p \in Properties
+	s \lhd State, l \in Label, p \in Properties
 \end{align*}
@@
 	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l \\ \wedge name(t_t) = reference(a) \\ \wedge startCardinality(a) > 1}{\begin{gathered} 
-	database(s) \bullet addTable(\rho(a,d),d) \\ \bullet \forall r \in m_{l, reference(a)} :\\ insertRow(row(name(owningClass(a)) + name(a), \\ [name(t_s), value(primaryKey(first(r)))]\\ [name(t_t), value(primaryKey(second(r)))]))
+	database(s) \bullet addTable(\rho(a,d),d) \\ \bullet \forall r \in m_{l, reference(a)} :\\ insertRow(\\ row(name(owningClass(a)) + name(a), \\ [name(t_s), value(primaryKey(first(r)))]\\ [name(t_t), value(primaryKey(second(r)))]))
 	\end{gathered}} 
@@
  	\end{cases}\\
- 	l, l_p \in Label, s \lhd Software
+ 	l, l_p \in Label, s \lhd State
 \end{align*}
@@
 	\\
-	l_c \in Label, ps \lhd Label*, s \lhd Software
+	l_c \in Label, ps \lhd Label*, s \lhd State
 \end{align*}
@@
      \end{cases}\\
-    l_c, l_a \in Labels, s \lhd Software
+    l_c, l_a \in Labels, s \lhd State
 \end{align*}
@@
 		%\inference{|ps| > 1}{\begin{gathered}			database(s) \\ \bullet RemoveAssociationAndData(L_c, head(ps)) \\ \bullet RemoveAssociationAndData(L_c, tail(ps)) \end{gathered}}\\\\
- 		\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists a \in associations(c) : name(a) = l \\ \wedge startCardinality(p) \leq 1}{database(s) \bullet RemoveForeignKeyAndData(reference(a), ps)}\\\\
- 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists a \in associations(c) : name(a) = l \\ \wedge startCardinality(a) > 1}{database(s) \bullet RemoveTableAndData(l_c + reference(a), ps)}
+ 		\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists a \in associations(c) : name(a) = l \\ \wedge startCardinality(p) \leq 1}{\begin{gathered}
+	database(s) \\ \bullet RemoveForeignKeyAndData(reference(a), ps)
+\end{gathered}}
+ 		\\\\
+ 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists a \in associations(c) : name(a) = l \\ \wedge startCardinality(a) > 1}{\begin{gathered}
+	database(s) \\\bullet RemoveTableAndData(l_c + reference(a), ps)
+\end{gathered}}
  \end{cases}
 	\\
-	l_c \in Label, ps \lhd Label*, s \lhd Software
+	l_c \in Label, ps \lhd Label*, s \lhd State
 \end{align*}
@@
  \end{cases}\\
-l_c, l_o, l_n \in Label,  s \lhd Software
+l_c, l_o, l_n \in Label,  s \lhd State
 \end{align*}
@@
  \end{cases} \\
-	l_c, l_o, l_n \in Label, a \lhd Applications, s \in Sates, d \lhd Database
+	l_c, l_o, l_n \in Label, a \lhd Applications, s \lhd Sate
 \end{align*}
@@
  \end{cases}\\
- l_s, l_t, l_p \in Label,  m_{l_s, l_t}) \in M, s \lhd Software
+ l_s, l_t, l_p \in Label,  m_{l_s, l_t}) \in M, s \lhd State
 \end{align*}
@@
 
-\section{Complex Refactoring Cases}
+\section{Advanced Refactoring Cases}
 The concrete refactoring cases presented in this section were chosen according to the refactoring statistics provided by Eclipse.
@@
 	\\
-	l_s, l_t \in Label, ps \in Label*,  m_{l_s, l_t} \in M, s \lhd Software
+	l_s, l_t \in Label, ps \in Label*,  m_{l_s, l_t} \in M, s \lhd State
 \end{align*}
@@
 	\inference{|ps| = 1}{\begin{gathered}
-		s \oplus addClass(cl(l_n, \varnothing, \varnothing)) \\ \oplus MoveProperty(l_s, l_n, ps, \mathbf{X}) \\ \oplus RemovePropertyData(l_s, ps) \\ \oplus RemoveProperty(l_s, ps)  
+		s \oplus addClass(cl(l_n, \varnothing, \varnothing)) \\ \oplus MoveProperty(l_s, l_n, ps, \mathbf{X}) \\ \oplus RemovePropertyData(l_s, ps)   
 \end{gathered}}\\\\
@@
 	\\
-	l_s, l_n \in Label, ps \lhd Label* s \lhd Software 
+	l_s, l_n \in Label, ps \lhd Label* s \lhd State 
 \end{align*}
@@
 }\\
-l_s, l_t \in Label, m_{l_s, l_t} \in M, s \lhd Software, \\ \\
+l_s, l_t \in Label, m_{l_s, l_t} \in M, s \lhd State, \\ \\
 ps = \inference{\exists c \in classes(applications)) \wedge ps = \varnothing}{\forall p \in properries(p) : ps \cup name(p)}

### 2013-01-23 doc/description.tex
formatting

@@
 \usepackage{amssymb}
-\usepackage{amsmath}
+\usepackage[fleqn]{amsmath}
 \usepackage{semantic}
@@
 \begin{align*}
-	\mathbf{software} = < Application, Database, \rho > \\
-	sw : Applications \times Databases \times ORM \rightarrow Software \\ \\
-	\mathbf{application} : Software \rightarrow Application \\
-	application(sw(a, d, \rho)) = a \\ \\
-	\mathbf{database} : Software \rightarrow Database \\
-	database(sw(a, d, \rho)) = d \\
-	a \lhd Application, d \lhd Database
-\end{align*}
-The software definition above introduce the first type used in this paper. The notation $a \lhd Application$ indicates the variable $a$ has the type $Application$.
+& \mathbf{software} = < Application, Database, \rho >\\
+& \mathbf{sw} : Applications \times Databases \times ORM \rightarrow  Software \\ \\
+& \mathbf{application} : Software \rightarrow Application \\
+& application(sw(a, d, \rho)) = a \\ \\
+& \mathbf{database} : Software \rightarrow Database \\
+& database(sw(a, d, \rho)) = d \\\\
+& a : Application, d : Database
+\end{align*}
 
@@
 \begin{align*}
- 	\mathbf{state} = Consistent\; s \: | \: Inconsistent \; s,  \\	s \lhd State \vee s \lh Application \vee s \lhd Database
+& 	\mathbf{state} = Consistent\; s \: | \: Inconsistent \; s,  \\\\
+& 	s : State \vee s \lh Application \vee s : Database
 \end{align*}
@@
 \begin{align*}
- 	\mathbf{consistent} : State \rightarrow Boolean \\
- 	consistent(s) = \begin{cases}
+& 	\mathbf{consistent} : State \rightarrow Boolean \\
+& 	consistent(s) = \begin{cases}
  		\inference{s = Consistent(s)}{true}\\\\
@@
  \end{cases}\\
- s \lhd State
+& s : State
 \end{align*}
@@
 \begin{align*}
-	\mathbf{state} : Software \rightarrow State \\
-	state(s) = \begin{cases}
+&	\mathbf{state} : Software \rightarrow State \\
+&	state(s) = \begin{cases}
  		\inference{consistent(application(s)) \wedge \\ consistent(database(s)) \\ \wedge \rho(application(s)) = database(s)}{Consistent \; s}\\ \\
@@
  	\end{cases} \\ 
- 	s \lhd State
+& 	s : State
 \end{align*}
@@
 \begin{align*}
-\Psi : E \rightarrow AppT \\
-\Phi : E \rightarrow DbT
+&	\Psi : E \rightarrow AppT \\
+&	\Phi : E \rightarrow DbT
 \end{align*}
@@
 The application of a transformation on a software state (evolution) is marked by the $\oplus$ operator:
-$$\oplus : State \times E \rightarrow State $$
+\begin{align*}
+& \oplus : State \times E \rightarrow State
+\end{align*}
 and similarly the $\circ$ and the $\bullet$ operators for evolving application and database:
 \begin{align*}
-\circ : State \times AppT \rightarrow State \\
- \bullet : State \times DbT \rightarrow State
+& \circ : State \times AppT \rightarrow State \\
+& \bullet : State \times DbT \rightarrow State
 \end{align*}
@@
 \begin{align*}
- state(s) \oplus t = \begin{cases}
+& state(s) \oplus t = \begin{cases}
  	\inference{! consistent(state(s))}{u} \\ \\
@@
 \end{cases}\\ 
-s \lhd State, t \in E
-\end{align*}
-\begin{align*}
- state(a) \circ e = \begin{cases}
+& s : State, t \in E
+\end{align*}
+\begin{align*}
+& state(a) \circ e = \begin{cases}
  	\inference{!\;consistent(state(a))}{a} \\ \\
@@
 \end{cases}\\ 
-a \lhd Application, e \in AppT
-\end{align*}
-\begin{align*}
- state(d) \bullet e = \begin{cases}
+& a : Application, e \in AppT
+\end{align*}
+\begin{align*}
+&  state(d) \bullet e = \begin{cases}
  	\inference{!\;consistent(state(d))}{d} \\ \\
@@
 \end{cases}\\ 
-d \lhd Application, e \in DbT
+& d : Application, e \in DbT
 \end{align*}
@@
 The notation: $Type*$ is used to define an ordered set of values of concrete type (e.g.
-$C \lhd Class*$ means c is an ordered set of classes).
+$C : Class*$ means c is an ordered set of classes).
 
@@
 \begin{align*}
-	f(x, y) = \begin{cases}
+& 	f(x, y) = \begin{cases}
  		\inference{x = \varnothing}{y}\\
  		\inference{|x| > 1}{f(head(x), y) \cup f(tail(x), y}\\\\
- 		\inference{|x| = 1}{effect}
+ 		\inference{|x| = 1}{effect}\\\\
  		\inference{otherwise}{Inconsistent y}
- \end{cases}\\
- y \lhd Software \vee y \lhd Application \vee y \lhd Database
+ \end{cases}\\\\
+&  y : Software \vee y : Application \vee y : Database
 \end{align*}
@@
 \begin{align*}
-  \mathbf{Application} = < Class* > \\ 
-  \mathbf{app} : Class* \rightarrow  Application \\ \\
-  \mathbf{classes} : Application \rightarrow Class* \\
-  classes(app(cs)) = cs  \\\\
-  \mathbf{class} : Label \times Application \rightarrow Class   \\ 
-  class(l, a) = c, c \in classes(a) \wedge name(c) = l \\ \\
-  l \in Label,a \lhd Applications, cs = Class* 
+& \mathbf{Application} = < Class* > \\ 
+& \mathbf{app} : Class* \rightarrow  Application \\ \\
+& \mathbf{classes} : Application \rightarrow Class* \\
+&   classes(app(cs)) = cs  \\\\
+& \mathbf{class} : Label \times Application \rightarrow Class   \\ 
+&  class(l, a) = c, c \in classes(a) \wedge name(c) = l \\ \\
+&  l \in Label,a : Applications, cs = Class* 
 \end{align*}
@@
 \begin{align*}
-	\mathbf{Class} = <label, Property*, Association*> \\
-	\mathbf{cl} : Label \times Property* \times Association* \rightarrow Class \\\\
-	\mathbf{name} : Class \rightarrow Label \\
-	name(cl(l, props, assocs)) = l \\ \\
-	\mathbf{properties} : Class \rightarrow Property* \\
-	properties(cl(l, props, assocs)) = props \\ \\
-	\mathbf{associations} : Class \rightarrow Association* \\	associations(cl(l, props, assocs)) = assocs \\ 
-	l \in Label, props \lhd Property*, assocs \lhd Association*
+& 	\mathbf{Class} = <label, Property*, Association*> \\
+& 	\mathbf{cl} : Label \times Property* \times Association* \rightarrow Class \\\\
+& 	\mathbf{name} : Class \rightarrow Label \\
+& 	name(cl(l, props, assocs)) = l \\ \\
+& 	\mathbf{properties} : Class \rightarrow Property* \\
+& 	properties(cl(l, props, assocs)) = props \\ \\
+& 	\mathbf{associations} : Class \rightarrow Association* \\ 
+& 	associations(cl(l, props, assocs)) = assocs \\ 
+& 	l \in Label, props : Property*, assocs : Association*
 \end{align*}
@@
  	\end{cases} \\ \\
-	l \in Label, props = Property*, assocs = Association*,a \lhd Applications
+	l \in Label, props = Property*, assocs = Association*,a : Applications
 \end{align*}
@@
 	owningClass(p, a) = \inference{\exists c \in classes(a) \wedge p \in properties(c)}{c}  \\ \\
-	l \in Label, t \lhd AppType, val \in Values, n \in \mathbb{N^{*}}, m \in Boolean 
+	l \in Label, t : AppType, val \in Values, n \in \mathbb{N^{*}}, m \in Boolean 
 \end{align*}
@@
 	owningClass(as, a) = \inference{\exists c \in classes(a) \wedge as \in associations(c)}{c}  \\ \\	
-	l, ref \in Label,  n_1, n_2 \in \mathbb{N}, as \lhd Association
+	l, ref \in Label,  n_1, n_2 \in \mathbb{N}, as : Association
 \end{align*}
@@
  	\end{cases}\\
-	a \lhd Application
+	a : Application
 \end{align*}
@@
 	rows(db(tabs, rs)) = rs \\ \\
-	tabs = Table*, rs = Row*, l \in Label,d \lhd Database
+	tabs = Table*, rs = Row*, l \in Label,d : Database
 \end{align*}
@@
 	foreignKeys(tab(l, pk, cols, fks)) = fks \\ \\
-l \in Label, pk \lhd PrimaryKey, cols = Column*, fks = ForeignKey*
+l \in Label, pk : PrimaryKey, cols = Column*, fks = ForeignKey*
 \end{align*}
@@
 	owningTable(c, d) = \inference{\exists tab \in tables(d) \wedge c \in columns(tab)}{tab} \\ \\
-l \in Label, t \lhd DbType, val \in Values,\\ cons \lhd Constraint*, c \lhd Column,d \lhd Database
+l \in Label, t : DbType, val \in Values,\\ cons : Constraint*, c : Column,d : Database
 \end{align*}
@@
 	owningTable(fk, d) = \inference{\exists tab \in tables(d) \wedge fk \in foreignKeys(tab)}{tab}\\\\
-	l, tRef \in Label, cons \lhd Constraint*,d \lhd Database, fk \lhd ForeignKey
+	l, tRef \in Label, cons : Constraint*,d : Database, fk : ForeignKey
 \end{align*}
@@
  		\end{cases} \\ \\
- l \in Label, ps = Pair*, t \in tables(d), d \lhd Database
+ l \in Label, ps = Pair*, t \in tables(d), d : Database
 \end{align*}
@@
  \end{cases} \\ \\
-	r \lhd Row*, l \in Label
+	r : Row*, l \in Label
 \end{align*}
@@
  \end{cases} \\ \\
-	 c \in classes(application(s)), s \lhd State 
+	 c \in classes(application(s)), s : State 
 \end{align*}
@@
 		\inference{|c| = 1}{table(name(c), primaryKey("Id"), \varnothing, \varnothing)}  \end{cases}   \\ 	
- c \lhd Class
+ c : Class
 \end{align*}
@@
 	\end{cases}\\
-	p \lhd Property
+	p : Property
 \end{align*}
@@
  	\end{cases}\\
- 	p \lhd Property
+ 	p : Property
 \end{align*}
@@
 	 \end{cases}\\
-	 a \lhd Association
+	 a : Association
 \end{align*}
@@
  	\end{cases}\\
-	 a \lhd Association
+	 a : Association
 \end{align*}
@@
 %	findProperty(l_c, l_p, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{p} \\
-%	l_c, l_p \in Label, a \lhd Application
+%	l_c, l_p \in Label, a : Application
 %\end{align*}
@@
 %	findAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists o \in associations(c) : name(o) = l_a}{o} \\
-%	l_c, l_a \in Label, a \lhd Application
+%	l_c, l_a \in Label, a : Application
 %\end{align*}
@@
 	addProp(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(c) : name(q) \neq name(p) \\ \wedge \forall e \in associations(c) : name(e) \neq name(p)}{properties(c) \cup p } \\
-	l \in Label, p \lhd Property, a \lhd Applications
+	l \in Label, p : Property, a : Applications
 \end{align*}
@@
 	\wedge \exists k \in classes(a) : name(k) = reference(s)}{associations(c) \cup s}\\
-	l \in Labels, s \in Associations, a \lhd Applications
+	l \in Labels, s \in Associations, a : Applications
 \end{align*}
@@
 	addClass(c, a) = \inference{\forall s \in classes(a) : name(s) \neq name(c) \\ \wedge associations(c) = \varnothing }{classes(a) \cup c} \\ \\
-	c \lhd Class, a \lhd Applications
+	c : Class, a : Applications
 \end{align*}
@@
 	}{properties(c) \setminus p }\\
-	l_c, l_p \in Label, a \lhd Applications 
+	l_c, l_p \in Label, a : Applications 
 \end{align*}
@@
 	remAssoc(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) \setminus f} \\
-	l_c, l_a \in Labels, a \lhd Applications
+	l_c, l_a \in Labels, a : Applications
 \end{align*}
@@
 	} {classes(a) \setminus c }\\
-	l \in Label,a \lhd Applications
+	l \in Label,a : Applications
 \end{align*}
@@
 	\end{gathered}}\\
-	l_c, l_o, l_n \in Label, a \lhd Application
+	l_c, l_o, l_n \in Label, a : Application
 \end{align*}
@@
 	\end{gathered}}
-	\\ l_c, l_o, l_n \in Label, a \lhd Application
+	\\ l_c, l_o, l_n \in Label, a : Application
 \end{align*}
@@
 	r_2 \in tableData(l_2, database(s)) \Rightarrow \exists p \in m_{l_1, l_2} \wedge second(p) = r_2\\
-	l_1, l_2 \in Label, r_1, r_2 \lhd Row
+	l_1, l_2 \in Label, r_1, r_2 : Row
 \end{align*}
@@
 %	findTable(l, d) = \inference{\exists t \in tables(d) \wedge name(t) = l}{t}  \\ \\
-%	l \in Label, d \lhd Database
+%	l \in Label, d : Database
 %\end{align*}
@@
  \end{cases}\\
- t_1 \lhd Table, t_2 \lhd Table
+ t_1 : Table, t_2 : Table
 \end{align*}
@@
  \end{cases}\\
- c_1 \lhd Column, c_2 \lhd Column
+ c_1 : Column, c_2 : Column
 \end{align*}
@@
  \end{cases}\\
- f_1 \lhd ForeignKey, f_2 \lhd ForeignKey
+ f_1 : ForeignKey, f_2 : ForeignKey
 \end{align*}
@@
 		\end{cases}\\
-		l \in Label, r \lhd Row
+		l \in Label, r : Row
 \end{align*}
@@
 	 \end{cases}\\ 
-	 l \in Label, c \lhd Column*,d \lhd Database
+	 l \in Label, c : Column*,d : Database
 \end{align*}
@@
 	 \end{cases}\\ 
-	t = Table*, d \lhd Database
+	t = Table*, d : Database
 \end{align*}
@@
  \\
-l_c, l_o, l_n \in Label, d \lhd Database
+l_c, l_o, l_n \in Label, d : Database
 \end{align*}
@@
 	\end{gathered}}\\
-	l_o, l_n \in Label, s \in States, a \lhd Applications, d \lhd Database
+	l_o, l_n \in Label, s \in States, a : Applications, d : Database
 \end{align*}
@@
 	\end{cases} \\
-	l_s,l_t,l_c \in Label \wedge l_s \neq l_t,d \lhd Database
+	l_s,l_t,l_c \in Label \wedge l_s \neq l_t,d : Database
 \end{align*}
@@
 		 \end{cases}	\\
-		 r = Row*,d \lhd Database
+		 r = Row*,d : Database
 \end{align*}
@@
  		\end{cases} \\
-r \lhd Row, p = Pair*,d \lhd Databases
+r : Row, p = Pair*,d : Databases
 \end{align*}
@@
  	\end{cases}\\ 
- 	l \in Label, p \lhd Pair,  d \lhd Database
+ 	l \in Label, p : Pair,  d : Database
 \end{align*}
@@
  	\end{cases} \\
- 	l_s, l_t, l_{sc}, l_{tc} \in Label, d \lhd Database
+ 	l_s, l_t, l_{sc}, l_{tc} \in Label, d : Database
 \end{align*}
@@
   \end{cases} \\
-  l_s, l_t \in Label, d \lhd Database, m_{l_s, l_t} \in M
+  l_s, l_t \in Label, d : Database, m_{l_s, l_t} \in M
 \end{align*}
@@
 	dropValue(l, r, d) = \inference{\exists p \in pairs(r) : reference(p) = l}{pairs(r) \setminus p}
-	\\ l \in Label, r \lhd Row \wedge r \in data(d), d \lhd Database
+	\\ l \in Label, r : Row \wedge r \in data(d), d : Database
 \end{align*}
@@
 	dropRow(r, d) = \inference{|r|=1}{data(d) \setminus r}
-	\\ l \in Label, r \lhd Row \wedge r \in data(d), d \lhd Database
+	\\ l \in Label, r : Row \wedge r \in data(d), d : Database
 \end{align*}
@@
 \end{gathered}}
-	\\ l \in Label, d \lhd Database
+	\\ l \in Label, d : Database
 \end{align*}
@@
 	dropColumn(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in columns(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{columns(t) \setminus c}\\
-	 l , l_p\in Label, d \lhd Database 
+	 l , l_p\in Label, d : Database 
 \end{align*}
@@
 	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) \setminus c}\\
-	 l, l_p \in Label, d \lhd Database 
+	 l, l_p \in Label, d : Database 
 \end{align*}
@@
 	\\
-	l_c \in Label, ps \lhd Label*, d \in Database
+	l_c \in Label, ps : Label*, d \in Database
 \end{align*}
@@
 	\\
-	l_f \in Label, ps \lhd Label*, d \in Database
+	l_f \in Label, ps : Label*, d \in Database
 \end{align*}
@@
    \end{cases}\\
-	s \lhd State, l \in Label, p \in Properties
+	s : State, l \in Label, p \in Properties
 \end{align*}
@@
  	\end{cases}\\
- 	l, l_p \in Label, s \lhd State
+ 	l, l_p \in Label, s : State
 \end{align*}
@@
 	\\
-	l_c \in Label, ps \lhd Label*, s \lhd State
+	l_c \in Label, ps : Label*, s : State
 \end{align*}
@@
      \end{cases}\\
-    l_c, l_a \in Labels, s \lhd State
+    l_c, l_a \in Labels, s : State
 \end{align*}
@@
  \end{cases}\\
-l_c, l_o, l_n \in Label,  s \lhd State
+l_c, l_o, l_n \in Label,  s : State
 \end{align*}
@@
  \end{cases} \\
-	l_c, l_o, l_n \in Label, a \lhd Applications, s \lhd Sate
+	l_c, l_o, l_n \in Label, a : Applications, s : Sate
 \end{align*}
@@
  \end{cases}\\
- l_s, l_t, l_p \in Label,  m_{l_s, l_t}) \in M, s \lhd State
+ l_s, l_t, l_p \in Label,  m_{l_s, l_t}) \in M, s : State
 \end{align*}
@@
 	\\
-	l_s, l_t \in Label, ps \in Label*,  m_{l_s, l_t} \in M, s \lhd State
+	l_s, l_t \in Label, ps \in Label*,  m_{l_s, l_t} \in M, s : State
 \end{align*}
@@
 	\\
-	l_s, l_n \in Label, ps \lhd Label* s \lhd State 
+	l_s, l_n \in Label, ps : Label* s : State 
 \end{align*}
@@
 }\\
-l_s, l_t \in Label, m_{l_s, l_t} \in M, s \lhd State, \\ \\
+l_s, l_t \in Label, m_{l_s, l_t} \in M, s : State, \\ \\
 ps = \inference{\exists c \in classes(applications)) \wedge ps = \varnothing}{\forall p \in properries(p) : ps \cup name(p)}

### 2013-01-23 doc/related_work.tex
related work started

\documentclass[11pt]{article}

\begin{document}
\section{Related Work}
There are proposed several metrics for classification of impact of a change on the conceptual schema in \cite{Wedemeijer:ConceptualMetrics}. These metrics can help us to describe the impact of proposed transformation on a software which is represent as two models (application and database). Only some metrics could be used for describing transformations as the others describe the state of conceptual schema. The universe (domain) is the application as all changes origins on this level.

The first metric "\textbf{Justified change}" means (in our context) that for each change of database exists a change of application. In case the evolution of the software was based on transformations proposed in this paper the metric is equal to one - each change is justified.

The metric "\textbf{Proportional change}" which compares the size and severity of change between application and database level can be used to describe complexity of transformations. The terms size of change and complexity of change has to be defined in order to use the metric. The size of change is equal to the number of entities in the model and instances which are affected by a change. The complexity of change is measured by number of atomic operations on each level needed to fulfill the evolution. This metrics is used to classify every single transformation.

The metric "\textbf{Impact on Complexity}" determines how the complexity of software or its components  changed. The metric can be used to evaluate each single transformations

\paragraph{Prediction of database change impact}The analysis technique for identifying the impact of relational database change upon object oriented application is proposed in \cite{Maule:ImpactAnalysis}. The technique focus on relational database and its change which impact on application is predicted. The analysis technique is able to identify changes in application structure (new attributes) as well as changes in database queries used in application. In contrast approach proposed in this paper does not consider database queries and theirs change and it is not mentioned to predict the impact of a change, but to  
process the change in the whole software, although it can be used for simulation of a change on formal level.

\bibliographystyle{plain}
\bibliography{references}
\end{document}

### 2013-01-24 doc/description.tex
formating

@@
 \begin{align*}
-	\mathbf{primitives} : Class \rightarrow Property* \\	primitives(cl(l, props, assocs)) = props', \\ props' \subset props,  \forall p \in props' : cardinality(p) \leq 1 \\ \\
-	\mathbf{collections} : Class \rightarrow Property* \\	collections(cl(l, props, assocs)) = props', \\ props' \subset props, \forall p \in props' : cardinality(p) > 1 \\ \\
-	\mathbf{associated} :  Label \times Application \rightarrow Boolean \\
-  	associated(l,a) = \begin{cases}
+&	\mathbf{primitives} : Class \rightarrow Property* \\
+&	primitives(cl(l, props, assocs)) = props', \\ 
+& props' \subset props,  \forall p \in props' : cardinality(p) \leq 1 \\ \\
+&	\mathbf{collections} : Class \rightarrow Property* \\	
+& collections(cl(l, props, assocs)) = props',  \\
+& props' \subset props, \forall p \in props' : cardinality(p) > 1 \\ \\
+&	\mathbf{associated} :  Label \times Application \rightarrow Boolean \\
+& 	associated(l,a) = \begin{cases}
   		\inference{\exists c \in classes(a) \wedge \exists e \in associations(c) : \\ reference(e) = l
@@
  	\end{cases} \\ \\
-	l \in Label, props = Property*, assocs = Association*,a : Applications
+&	l \in Label, props = Property*, assocs = Association*,a : Applications
 \end{align*}
@@
 \begin{align*}
-	\mathbf{Property} = <label, AppType, DefaultValue, Cardinality, Mandatory> \\
-	\mathbf{prop} : Label \times AppType \times String \times \mathbb{N^{*}} \times Boolean \rightarrow Property \\ \\
-	\mathbf{name} : Property \rightarrow Label \\
-	name(prop(l, t, val, n, m)) = l \\ \\
-	\mathbf{type} : Property \times AppType \\
-	type(prop(l, t, val, n, m)) = t \\ \\
-	\mathbf{cardinality} : Property \rightarrow \mathbb{N^{*}} \\
-	cardinality(prop(l, t, val, n, m)) = n \\ \\
-	\mathbf{mandatory} : Property \rightarrow Boolean \\
-	mandatory(prop(l, t, val, n, m)) = m  \\ \\
-	\mathbf{owningClass} : Property \times Application \rightarrow Class  \\ 	
-	owningClass(p, a) = \inference{\exists c \in classes(a) \wedge p \in properties(c)}{c}  \\ \\
-	l \in Label, t : AppType, val \in Values, n \in \mathbb{N^{*}}, m \in Boolean 
+&	\mathbf{Property} = <label, AppType, DefaultValue, Cardinality, Mandatory> \\
+&	\mathbf{prop} : Label \times AppType \times String \times \mathbb{N^{*}} \times Boolean \rightarrow Property \\ \\
+&	\mathbf{name} : Property \rightarrow Label \\
+&	name(prop(l, t, val, n, m)) = l \\ \\
+&	\mathbf{type} : Property \times AppType \\
+&	type(prop(l, t, val, n, m)) = t \\ \\
+&	\mathbf{cardinality} : Property \rightarrow \mathbb{N^{*}} \\
+&	cardinality(prop(l, t, val, n, m)) = n \\ \\
+&	\mathbf{mandatory} : Property \rightarrow Boolean \\
+&	mandatory(prop(l, t, val, n, m)) = m  \\ \\
+&	\mathbf{owningClass} : Property \times Application \rightarrow Class  \\ 	
+&	owningClass(p, a) = \inference{\exists c \in classes(a) \wedge p \in properties(c)}{c}  \\ \\
+&	l \in Label, t : AppType, val \in Values, n \in \mathbb{N^{*}}, m \in Boolean 
 \end{align*}
@@
 \begin{align*}
-	\mathbf{Association} = <label, classRef, startCardinality, endCardinality> \\
-	\mathbf{assoc} : Label \times Label \times \mathbb{N} \times \mathbb{N} \rightarrow Association\\ \\
-	\mathbf{name} : Association \rightarrow Label \\
-	name(assoc(l, ref, n_1, n_2)) = l\\ \\
-	\mathbf{reference} : Association \rightarrow Label \\
-	reference(assoc(l, ref, n_1, n_2)) = ref\\ \\
-	\mathbf{startCardinality} : Association \rightarrow \mathbb{N} \\
-	startCardinality(assoc(l, ref, n_1, n_2)) = n_1\\ \\
-	\mathbf{endCardinality} : Association \rightarrow \mathbb{N} \\
-	endCardinality(assoc(l, ref, n_1, n_2)) = n_2 \\ \\
-	\mathbf{owningClass} : Association \times Application \rightarrow Class  \\
-	owningClass(as, a) = \inference{\exists c \in classes(a) \wedge as \in associations(c)}{c}  \\ \\	
-	l, ref \in Label,  n_1, n_2 \in \mathbb{N}, as : Association
+&	\mathbf{Association} = <label, classRef, startCardinality, endCardinality> \\
+&	\mathbf{assoc} : Label \times Label \times \mathbb{N} \times \mathbb{N} \rightarrow Association\\ \\
+&	\mathbf{name} : Association \rightarrow Label \\
+&	name(assoc(l, ref, n_1, n_2)) = l\\ \\
+&	\mathbf{reference} : Association \rightarrow Label \\
+&	reference(assoc(l, ref, n_1, n_2)) = ref\\ \\
+&	\mathbf{startCardinality} : Association \rightarrow \mathbb{N} \\
+&	startCardinality(assoc(l, ref, n_1, n_2)) = n_1\\ \\
+&	\mathbf{endCardinality} : Association \rightarrow \mathbb{N} \\
+&	endCardinality(assoc(l, ref, n_1, n_2)) = n_2 \\ \\
+&	\mathbf{owningClass} : Association \times Application \rightarrow Class  \\
+&	owningClass(as, a) = \inference{\exists c \in classes(a) \wedge as \in associations(c)}{c}  \\ \\	
+&	l, ref \in Label,  n_1, n_2 \in \mathbb{N}, as : Association
 \end{align*}
@@
 \begin{align*}
-\mathbf{AppType} = APPSTRING
+& \mathbf{AppType} = APPSTRING
 \end{align*}
@@
 \begin{align*}
-	state(a) = \begin{cases}
+&	state(a) = \begin{cases}
 	\inference{\forall c_1, c_2 \in classes(a) : name(c_1) \neq name(c_2) \\ \wedge \forall p_1, p_2 \in properties(c_1) : name(p_1) \neq name(p_2) \\ \wedge \forall a_1, a_2 \in associations(c_1) : name(a_1) \neq name (a_2) \\ \wedge \forall a \in associations(c_1) : \exists c \in classes(a) \wedge name(c) = reference(a) }{Consistent \; a}\\\\
@@
  	\end{cases}\\
-	a : Application
+&	a : Application
 \end{align*}
@@
 \begin{align*}
-	\mathbf{Database} = < Table*, Row*> \\
-	\mathbf{db} : Tables \times Rows \rightarrow Database \\ \\
-	\mathbf{tables} : Database \rightarrow Tables \\
-	tables(db(tabs, rs)) = tabs \\ \\
-	\mathbf{rows} : Database \rightarrow Rows \\
-	rows(db(tabs, rs)) = rs \\ \\
-	tabs = Table*, rs = Row*, l \in Label,d : Database
+&	\mathbf{Database} = < Table*, Row*> \\
+&	\mathbf{db} : Tables \times Rows \rightarrow Database \\ \\
+&	\mathbf{tables} : Database \rightarrow Tables \\
+&	tables(db(tabs, rs)) = tabs \\ \\
+&	\mathbf{rows} : Database \rightarrow Rows \\
+&	rows(db(tabs, rs)) = rs \\ \\
+&	tabs = Table*, rs = Row*, l \in Label,d : Database
 \end{align*}
@@
 \begin{align*}
-	\mathbf{Table} = <label, primaryKey, Column*, ForeignKey*>\\
-	\mathbf{tab} 	: Label \times PrimaryKeys \times Columns \times ForeignKeys \rightarrow Table \\ \\
-	\mathbf{name} : Table \rightarrow Label  \\
-	name(tab(l, pk, cols, fks)) = l \\ \\
-	\mathbf{primaryKey} : Table \rightarrow PrimaryKeys  \\
-	primaryKey(tab(l, pk, cols, fks)) = pk \\ \\
-	\mathbf{columns} : Table \rightarrow Columns  \\
-	columns(tab(l, pk, cols, fks)) = cols \\ \\
-	\mathbf{foreignKeys} : Table \rightarrow ForeignKeys  \\
-	foreignKeys(tab(l, pk, cols, fks)) = fks \\ \\
-l \in Label, pk : PrimaryKey, cols = Column*, fks = ForeignKey*
+&	\mathbf{Table} = <label, primaryKey, Column*, ForeignKey*>\\
+&	\mathbf{tab} 	: Label \times PrimaryKeys \times Columns \times ForeignKeys \rightarrow Table \\ \\
+&	\mathbf{name} : Table \rightarrow Label  \\
+&	name(tab(l, pk, cols, fks)) = l \\ \\
+&	\mathbf{primaryKey} : Table \rightarrow PrimaryKeys  \\
+&	primaryKey(tab(l, pk, cols, fks)) = pk \\ \\
+&	\mathbf{columns} : Table \rightarrow Columns  \\
+&	columns(tab(l, pk, cols, fks)) = cols \\ \\
+&	\mathbf{foreignKeys} : Table \rightarrow ForeignKeys  \\
+&	foreignKeys(tab(l, pk, cols, fks)) = fks \\ \\
+& l \in Label, pk : PrimaryKey, cols = Column*, fks = ForeignKey*
 \end{align*}
@@
 \begin{align*}
-	\mathbf{Column} = <label, DbType, DefaultValue, Constraint*> \\
-	\mathbf{col} : Label \times DbType \times String \times Constraints \rightarrow Column \\ \\
-	\mathbf{name} : Column \rightarrow Label  \\
-	name(col(l, t, val, cons)) = l  \\ \\
-	\mathbf{type} : Column \rightarrow DbType  \\
-	type(col(l, t, val, cons)) = t  \\ \\
-	\mathbf{constraints} : Column \rightarrow Constraints  \\
-	constraints(col(l, t, val, cons)) = cons  \\ \\
-	\mathbf{owningTable} : Column \times Database \rightarrow Table  \\
-	owningTable(c, d) = \inference{\exists tab \in tables(d) \wedge c \in columns(tab)}{tab} \\ \\
-l \in Label, t : DbType, val \in Values,\\ cons : Constraint*, c : Column,d : Database
+&	\mathbf{Column} = <label, DbType, DefaultValue, Constraint*> \\
+&	\mathbf{col} : Label \times DbType \times String \times Constraints \rightarrow Column \\ \\
+&	\mathbf{name} : Column \rightarrow Label  \\
+&	name(col(l, t, val, cons)) = l  \\ \\
+&	\mathbf{type} : Column \rightarrow DbType  \\
+&	type(col(l, t, val, cons)) = t  \\ \\
+&	\mathbf{constraints} : Column \rightarrow Constraints  \\
+&	constraints(col(l, t, val, cons)) = cons  \\ \\
+&	\mathbf{owningTable} : Column \times Database \rightarrow Table  \\
+&	owningTable(c, d) = \inference{\exists tab \in tables(d) \wedge c \in columns(tab)}{tab} \\ \\
+& l \in Label, t : DbType, val \in Values,\\ 
+& cons : Constraint*, c : Column,d : Database
 \end{align*}
@@
 \begin{align*}
-	\mathbf{ForeignKey} = <label, tableRef, Constraint*> \\ 
-	\mathbf{fk} : Label \times Label \times Constraints \rightarrow ForeignKey \\ \\
-	\mathbf{name} : ForeignKey \rightarrow Label \\
-	name(fk(l, tRef, cons)) = lab  \\ \\
-	\mathbf{reference} : ForeignKey \rightarrow Label  \\
-	reference(fk(l, tRef, cons)) = tRef  \\ \\
-	\mathbf{constaints} : ForeignKey \rightarrow Constraint*  \\
-	constaints(fk(l, tRef, cons)) = cons  \\ \\
-	\mathbf{owningTable} : ForeignKey \times Database \rightarrow Table  \\
-	owningTable(fk, d) = \inference{\exists tab \in tables(d) \wedge fk \in foreignKeys(tab)}{tab}\\\\
-	l, tRef \in Label, cons : Constraint*,d : Database, fk : ForeignKey
+&	\mathbf{ForeignKey} = <label, tableRef, Constraint*> \\ 
+&	\mathbf{fk} : Label \times Label \times Constraints \rightarrow ForeignKey \\ \\
+&	\mathbf{name} : ForeignKey \rightarrow Label \\
+&	name(fk(l, tRef, cons)) = lab  \\ \\
+&	\mathbf{reference} : ForeignKey \rightarrow Label  \\
+&	reference(fk(l, tRef, cons)) = tRef  \\ \\
+&	\mathbf{constaints} : ForeignKey \rightarrow Constraint*  \\
+&	constaints(fk(l, tRef, cons)) = cons  \\ \\
+&	\mathbf{owningTable} : ForeignKey \times Database \rightarrow Table  \\
+&	owningTable(fk, d) = \inference{\exists tab \in tables(d) \wedge fk \in foreignKeys(tab)}{tab}\\\\
+&	l, tRef \in Label, cons : Constraint*,d : Database, fk : ForeignKey
 \end{align*}
@@
 \begin{align*}
-	\mathbf{PrimaryKey} =  < label > 	\\
-	\mathbf{pk} : Label \rightarrow PrimaryKey
+&	\mathbf{PrimaryKey} =  < label > 	\\
+&	\mathbf{pk} : Label \rightarrow PrimaryKey
 \end{align*}
@@
 There is only one type of column values defined in the model called $DBSTRING$.
-Next the $DBN$ type is defined for the values of keys.
-\begin{align*}
-	\mathbf{DbType} = DBSTRING | DBN
+Next the $DBINT$ type is defined for the values of keys.
+\begin{align*}
+&	\mathbf{DbType} = DBSTRING | DBINT
 \end{align*}
@@
 \begin{align*}
-	\mathbf{Constraint} = NOTNULL \; | \; UNIQUE 
+&	\mathbf{Constraint} = NOTNULL \; | \; UNIQUE 
 \end{align*}
@@
 \begin{align*}
-	\mathbf{Row} = < refTable, pkValue, pair* > \\
-	\mathbf{row} : Label \times Pairs \rightarrow Row \\ \\
-	\mathbf{refTable} : Row \rightarrow Label \\
-	refTable(row(l, n, ps) = l \\ \\
-	\mathbf{pairs} : Row \rightarrow Pair* \\
-	pairs(row(l, n, ps)) = ps \\ \\
-	\mathbf{pkValue} : Row \rightarrow N \\
-	pkValue(row(l, n, ps)) = n \\ \\
-	 l \in Label, ps = Pair*, n \in N
+&	\mathbf{Row} = < refTable, pkValue, pair* > \\
+&	\mathbf{row} : Label \times Pairs \rightarrow Row \\ \\
+&	\mathbf{refTable} : Row \rightarrow Label \\
+&	refTable(row(l, n, ps) = l \\ \\
+&	\mathbf{pairs} : Row \rightarrow Pair* \\
+&	pairs(row(l, n, ps)) = ps \\ \\
+&	\mathbf{pkValue} : Row \rightarrow N \\
+&	pkValue(row(l, n, ps)) = n \\ \\
+&	 l \in Label, ps = Pair*, n \in N
 \end{align*}
@@
 \begin{align*}
-	\mathbf{Pair} = < refCol, value > \\
-	[Label \times Value] \rightarrow Pair\\ \\
-	\mathbf{reference} : Pair \rightarrow Label \\
-	reference([l,val]) = l \\ \\
-	\mathbf{value} : Pair \rightarrow Value \\ 
-	value([l,val]) = val \\ \\
-	l \in Label, val \in Value
+&	\mathbf{Pair} = < refCol, value > \\
+&	[Label \times Value] \rightarrow Pair\\ \\
+&	\mathbf{reference} : Pair \rightarrow Label \\
+&	reference([l,val]) = l \\ \\
+&	\mathbf{value} : Pair \rightarrow Value \\ 
+&	value([l,val]) = val \\ \\
+&	l \in Label, val \in Value
 \end{align*}
@@
 \begin{align*}
-	\mathbf{contains} : Table \times Database \rightarrow Boolean \\
-	contains(t, d) = \begin{cases}
+&	\mathbf{contains} : Table \times Database \rightarrow Boolean \\
+&	contains(t, d) = \begin{cases}
 			\inference{\exists r \in data(d) : refTable(r) = name(t) }{true} \\ \\
@@
  		\end{cases} \\ \\
- l \in Label, ps = Pair*, t \in tables(d), d : Database
-\end{align*}
-\begin{align*}
-	\mathbf{tableData} : Label \times Row* \rightarrow Row* \\
-	tableData(l, r) = \begin{cases}
+& l \in Label, ps = Pair*, t \in tables(d), d : Database
+\end{align*}
+\begin{align*}
+&	\mathbf{tableData} : Label \times Row* \rightarrow Row* \\
+&	tableData(l, r) = \begin{cases}
  		%\inference{r = \varnothing}{\varnothing} \\
  		%\inference{|r| > 1}{tableData(head(r)) \cup tableData(tail(r))}\\ \\
- 		\inference{|r| = 1 \wedge refTable(r) = l}{r}\\\\
- 		\inference{|r| = 1 \wedge refTable(r) \neq l}{\varnothing}
+&		\inference{|r| = 1 \wedge refTable(r) = l}{r}\\\\
+& 		\inference{|r| = 1 \wedge refTable(r) \neq l}{\varnothing}
  \end{cases} \\ \\
-	r : Row*, l \in Label
-\end{align*}
-\begin{align*}
-	\mathbf{instantiated} : Class \times Software \rightarrow Boolean \\
-	instantiated(c, s) = \begin{cases} \inference{\exists t \in tables(database(s)) : name(t) = name(c) \\ \wedge contains(t, database(s))}{true} \\ \\
+&	r : Row*, l \in Label
+\end{align*}
+\begin{align*}
+&	\mathbf{instantiated} : Class \times Software \rightarrow Boolean \\
+&	instantiated(c, s) = \begin{cases} \inference{\exists t \in tables(database(s)) : name(t) = name(c) \\ \wedge contains(t, database(s))}{true} \\ \\
   false
  \end{cases} \\ \\
-	 c \in classes(application(s)), s : State 
+&	 c \in classes(application(s)), s : State 
 \end{align*}
@@
 \begin{align*}
-	\mathbf{consistent} : Database \rightarrow Boolean \\
-	state(d) = \begin{cases}
- 		\inference{ \forall t_1, t_2 \in tables(d) : name(t_1) \neq name(t_2) \\ \wedge \forall c_1, c_2 \in columns(t_1) : name(c_1) \neq name(c_2) \\ \wedge \forall f_1, f_2 \in foreignKeys(t_1) : name(f_1) \neq name (f_2) \\ \wedge \forall f \in foreignKeys(t_1) : \exists t \in tables(d) \wedge name(t) = reference(f) \\ \wedge \forall r \in data(d): ( \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) :\\ \exists u \in columns(t) \cup foreignKeys(t) \cup primaryKey(t) : \\ reference(p) = name(u) \\ \wedge \forall e \in columns(t) \cup foreignKeys(t) \\ \wedge NOTNULL \in constrain(e) : \exists q \in pairs(r) \wedge name(e) = reference(q)) \\ \wedge \forall g \in columns(t) \cup foreignKeys(t) \\ \wedge UNIQUE \in constrain(g) : \\ \forall u_1, u_2 \in rows(t) : \\ value(pairByName(name(g), u_1) \neq value(pairByName(name(g), u_2) }{true}
+&	\mathbf{consistent} : Database \rightarrow Boolean \\
+&	state(d) = \begin{cases}
+&		\inference{ \forall t_1, t_2 \in tables(d) : name(t_1) \neq name(t_2) \\ \wedge \forall c_1, c_2 \in columns(t_1) : name(c_1) \neq name(c_2) \\ \wedge \forall f_1, f_2 \in foreignKeys(t_1) : name(f_1) \neq name (f_2) \\ \wedge \forall f \in foreignKeys(t_1) : \exists t \in tables(d) \wedge name(t) = reference(f) \\ \wedge \forall r \in data(d): ( \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) :\\ \exists u \in columns(t) \cup foreignKeys(t) \cup primaryKey(t) : \\ reference(p) = name(u) \\ \wedge \forall e \in columns(t) \cup foreignKeys(t) \\ \wedge NOTNULL \in constrain(e) : \exists q \in pairs(r) \wedge name(e) = reference(q)) \\ \wedge \forall g \in columns(t) \cup foreignKeys(t) \\ \wedge UNIQUE \in constrain(g) : \\ \forall u_1, u_2 \in rows(t) : \\ value(pairByName(name(g), u_1) \neq value(pairByName(name(g), u_2) }{true}
  	\\ \\
@@
 \begin{align*}
-	\mathbf{\rho_{app}} : Application \rightarrow Database\\
-	\rho_{app}(a) = \inference{state(a) = Consistent \; a}{\begin{gathered}
+&	\mathbf{\rho_{app}} : Application \rightarrow Database\\
+&	\rho_{app}(a) = \inference{state(a) = Consistent \; a}{\begin{gathered}
    db(\varnothing) \bullet addTable(classes(a)) \\ 
@@
 \begin{align*}
-	\mathbf{\rho_{t}} : AppType \rightarrow DbType  \\
- 	\rho_{t}(APPSTRING) = DBSTRING 
+&	\mathbf{\rho_{t}} : AppType \rightarrow DbType  \\
+& 	\rho_{t}(APPSTRING) = DBSTRING 
 \end{align*}
@@
 \begin{align*}
-	\mathbf{\rho_{c}} : Class*  \rightarrow Table* \\
-	\rho_{c}(c) = \begin{cases}
+&	\mathbf{\rho_{c}} : Class*  \rightarrow Table* \\
+&	\rho_{c}(c) = \begin{cases}
 		%\inference{c = \varnothing}{\varnothing}\\\\
@@
 		\inference{|c| = 1}{table(name(c), primaryKey("Id"), \varnothing, \varnothing)}  \end{cases}   \\ 	
- c : Class
+& c : Class
 \end{align*}
@@
 \begin{align*}
-	\mathbf{\rho_p} : Property* \rightarrow Column* \\ \\
-	\rho_{sp}(p) = \begin{cases}
+&	\mathbf{\rho_p} : Property* \rightarrow Column* \\ \\
+&	\rho_{sp}(p) = \begin{cases}
 		%\inference{p = \varnothing}{\varnothing}\\
@@
 	\end{cases}\\
-	p : Property
-\end{align*}
-\begin{align*}
-	\mathbf{\rho_p} : Property* \rightarrow Table* \\ \\
-	\rho_{cp}(p) = \begin{cases}
+&	p : Property
+\end{align*}
+\begin{align*}
+&	\mathbf{\rho_p} : Property* \rightarrow Table* \\ \\
+&	\rho_{cp}(p) = \begin{cases}
 		%\inference{p = \varnothing}{\varnothing}\\
@@
  	\end{cases}\\
- 	p : Property
+& 	p : Property
 \end{align*}
@@
 \begin{align*}
-	\mathbf{\rho_a} : Association* \rightarrow ForeignKey* \\
-	\rho_a(a) = \begin{cases}
+&	\mathbf{\rho_a} : Association* \rightarrow ForeignKey* \\
+&	\rho_a(a) = \begin{cases}
 		%\inference{a = \varnothing}{\varnothing}\\
@@
 	 \end{cases}\\
-	 a : Association
-\end{align*}
-\begin{align*} 
-	\mathbf{\rho_a} : Association* \rightarrow Table* \\
-	\rho_a(a) = \begin{cases}
+&	 a : Association
+\end{align*}
+\begin{align*} 
+&	\mathbf{\rho_a} : Association* \rightarrow Table* \\
+&	\rho_a(a) = \begin{cases}
  		%\inference{a = \varnothing}{\varnothing}\\
@@
  	\end{cases}\\
-	 a : Association
+&	 a : Association
 \end{align*}
@@
 \begin{align*}
- 	\mathbf{addProp} : Label \times Property \times Application \rightarrow Application \\
-	addProp(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(c) : name(q) \neq name(p) \\ \wedge \forall e \in associations(c) : name(e) \neq name(p)}{properties(c) \cup p } \\
-	l \in Label, p : Property, a : Applications
+& 	\mathbf{addProp} : Label \times Property \times Application \rightarrow Application \\
+&	addProp(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(c) : name(q) \neq name(p) \\ \wedge \forall e \in associations(c) : name(e) \neq name(p)}{properties(c) \cup p } \\
+&	l \in Label, p : Property, a : Applications
 \end{align*}
@@
 \begin{align*}
-	\mathbf{addAssoc} : Label \times Association \times Application \rightarrow Application \\
-	addAssoc(l, s, a) = \inference{
+&	\mathbf{addAssoc} : Label \times Association \times Application \rightarrow Application \\
+&	addAssoc(l, s, a) = \inference{
 	\exists c \in classes(a) : name(c) = l \\
@@
 	\wedge \exists k \in classes(a) : name(k) = reference(s)}{associations(c) \cup s}\\
-	l \in Labels, s \in Associations, a : Applications
+&	l \in Labels, s \in Associations, a : Applications
 \end{align*}
@@
 \begin{align*}
-	\mathbf{addClass} : Class \times Application \rightarrow Application \\ 
-	addClass(c, a) = \inference{\forall s \in classes(a) : name(s) \neq name(c) \\ \wedge associations(c) = \varnothing }{classes(a) \cup c} \\ \\
-	c : Class, a : Applications
+&	\mathbf{addClass} : Class \times Application \rightarrow Application \\ 
+&	addClass(c, a) = \inference{\forall s \in classes(a) : name(s) \neq name(c) \\ \wedge associations(c) = \varnothing }{classes(a) \cup c} \\ \\
+&	c : Class, a : Applications
 \end{align*}
@@
 \begin{align*}
- 	\mathbf{remProp} : Label \times Label \times Application \rightarrow Application \\
- 	remProp(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p
+&	\mathbf{remProp} : Label \times Label \times Application \rightarrow Application \\
+& 	remProp(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p
 	}{properties(c) \setminus p }\\
-	l_c, l_p \in Label, a : Applications 
-\end{align*}
-\begin{align*}
-	\mathbf{remAssoc} : Label \times Label \times Application \rightarrow Application \\
-	remAssoc(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) \setminus f} \\
-	l_c, l_a \in Labels, a : Applications
+&	l_c, l_p \in Label, a : Applications 
+\end{align*}
+\begin{align*}
+&	\mathbf{remAssoc} : Label \times Label \times Application \rightarrow Application \\
+&	remAssoc(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) \setminus f} \\
+&	l_c, l_a \in Labels, a : Applications
 \end{align*}
@@
 \begin{align*}
-	\mathbf{remClass} : Label \times Application \rightarrow Application \\
-	remClass(l, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge !\: associated(c)
+&	\mathbf{remClass} : Label \times Application \rightarrow Application \\
+&	remClass(l, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge !\: associated(c)
 	} {classes(a) \setminus c }\\
-	l \in Label,a : Applications
+&	l \in Label,a : Applications
 \end{align*}
@@
 \begin{align*}
-	\mathbf{renProp} : Label \times Label \times Label \times Application \rightarrow Application \\
-	renProp(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_o \\ \wedge \forall e \in properties(c) : name(e) \neq l_n
+&	\mathbf{renProp} : Label \times Label \times Label \times Application \rightarrow Application \\
+&	renProp(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_o \\ \wedge \forall e \in properties(c) : name(e) \neq l_n
 	}{\begin{gathered}
@@
 	\end{gathered}}\\
-	l_c, l_o, l_n \in Label, a : Application
-\end{align*}
-\begin{align*}
-	\mathbf{renAssoc} : Label \times Label \times Label \times Application \rightarrow Application \\
-	renAssoc(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists b \in associations(c) : name(b) = l_o \\ \wedge \forall e \in associations(c) : name(e) \neq l_n
+&	l_c, l_o, l_n \in Label, a : Application
+\end{align*}
+\begin{align*}
+&	\mathbf{renAssoc} : Label \times Label \times Label \times Application \rightarrow Application \\
+&	renAssoc(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists b \in associations(c) : name(b) = l_o \\ \wedge \forall e \in associations(c) : name(e) \neq l_n
 	}{\begin{gathered}
@@
 	\end{gathered}}
-	\\ l_c, l_o, l_n \in Label, a : Application
+	\\ & l_c, l_o, l_n \in Label, a : Application
 \end{align*}
@@
 \begin{align*}
-	\mathbf{renClass} :  Label \times Label \times Application \rightarrow Application \\
-	renClass(l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_o \\ \wedge \forall e \in classes(a) : name(e) \neq l_n
+&	\mathbf{renClass} :  Label \times Label \times Application \rightarrow Application \\
+&	renClass(l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_o \\ \wedge \forall e \in classes(a) : name(e) \neq l_n
 	}{\begin{gathered}
@@
 \begin{align*}
-	m_{l_1, l_2} =  <(r_1, r_2)*>,\\
-	r_1 \in tableData(l_1, database(s)) \Rightarrow \exists p \in m_{l_1, l_2} \wedge first(p) = r_1 \\
-	r_2 \in tableData(l_2, database(s)) \Rightarrow \exists p \in m_{l_1, l_2} \wedge second(p) = r_2\\
-	l_1, l_2 \in Label, r_1, r_2 : Row
+&	m_{l_1, l_2} =  <(r_1, r_2)*>,\\
+&	r_1 \in tableData(l_1, database(s)) \Rightarrow \exists p \in m_{l_1, l_2} \wedge first(p) = r_1 \\
+&	r_2 \in tableData(l_2, database(s)) \Rightarrow \exists p \in m_{l_1, l_2} \wedge second(p) = r_2\\
+&	l_1, l_2 \in Label, r_1, r_2 : Row
 \end{align*}
@@
 \begin{align*}
-	\mathbf{\cong} : Table \times Table \rightarrow Boolean \\
-	t_1 \cong t_2 = \begin{cases}
+&	\mathbf{\cong} : Table \times Table \rightarrow Boolean \\
+&	t_1 \cong t_2 = \begin{cases}
  		\inference{|columns(t_1)| = |columns(t_2)| \\ \wedge |foreignKeys(t_1)| = |foreignKeys(t_2)| \\ \wedge \forall c \in columns(t_1) \exists e \in columns(t_2) : c = e \\ \wedge \forall f \in foreignKeys(t_1) \exists g \in foreignKeys(t_2) : f = g }{true}\\\\
@@
  \end{cases}\\
- t_1 : Table, t_2 : Table
-\end{align*}
-\begin{align*}
-	\mathbf{\cong} : Column \times Column \rightarrow Boolean \\
-	c_1 \cong c_2 = \begin{cases}
+& t_1 : Table, t_2 : Table
+\end{align*}
+\begin{align*}
+&	\mathbf{\cong} : Column \times Column \rightarrow Boolean \\
+&	c_1 \cong c_2 = \begin{cases}
  		\inference{name(c_1) = name(c_2) \\ \wedge type(c_1) = type(c_2) \\ \wedge constraints(c_1) = comstraints(c_2)}{true}
@@
  \end{cases}\\
- c_1 : Column, c_2 : Column
-\end{align*}
-\begin{align*}
-	\mathbf{\cong} : ForeignKey \times ForeignKey \rightarrow Boolean \\
-	f_1 \cong f_2 = \begin{cases}
+& c_1 : Column, c_2 : Column
+\end{align*}
+\begin{align*}
+&	\mathbf{\cong} : ForeignKey \times ForeignKey \rightarrow Boolean \\
+&	f_1 \cong f_2 = \begin{cases}
  		\inference{name(f_1) = name(f_2) \\ \wedge reference(f_1) = reference(f_2) \\ \wedge constraints(f_1) = comstraints(f_2}{true}
@@
  \end{cases}\\
- f_1 : ForeignKey, f_2 : ForeignKey
+& f_1 : ForeignKey, f_2 : ForeignKey
 \end{align*}
@@
 \begin{align*}
-\mathbf{findPair} : Label \times Row \rightarrow Pair \\
-	findPair(l, r) = \begin{cases}
+& \mathbf{findPair} : Label \times Row \rightarrow Pair \\
+&	findPair(l, r) = \begin{cases}
 		\inference{\exists p \in pairs(r) : reference(p) = l}{p} \\ \\
@@
 		\end{cases}\\
-		l \in Label, r : Row
+&		l \in Label, r : Row
 \end{align*}
@@
 \begin{align*}
-	\mathbf{addColumn} : Label \times Column* \times Database \rightarrow Database \\ 	
-	addColum(l, c, d) = \begin{cases}
+&	\mathbf{addColumn} : Label \times Column* \times Database \rightarrow Database \\ 	
+&	addColum(l, c, d) = \begin{cases}
 		%\inference{|c| > 1}{addColumn(l, head(c), d) \bullet addColumn(l, tail(c), d)} \\ \\
@@
 	 \end{cases}\\ 
-	 l \in Label, c : Column*,d : Database
+&	 l \in Label, c : Column*,d : Database
 \end{align*}
@@
 \begin{align*}
-	\mathbf{addTable} : Table* \times Database \rightarrow Database \\ 
-	addTable(t, d) = \begin{cases}
+&	\mathbf{addTable} : Table* \times Database \rightarrow Database \\ 
+&	addTable(t, d) = \begin{cases}
 		%\inference{|t| > 1}{d \bullet addTable(head(t)) \bullet addTable(tail(t))}
@@
 	 \end{cases}\\ 
-	t = Table*, d : Database
+&	t = Table*, d : Database
 \end{align*}
@@
 \begin{align*}
-\mathbf{alterColumn} : Label \times Label \times Label \times Database \rightarrow Database \\
-	alterColumn(l_c, l_o, l_n, d) =
+& \mathbf{alterColumn} : Label \times Label \times Label \times Database \rightarrow Database \\
+&	alterColumn(l_c, l_o, l_n, d) =
  \inference{\exists t \in tables(a) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_o \\ \wedge \forall e \in columns(t) : name(e) \neq l_n}{\begin{gathered}
@@
  \\
-l_c, l_o, l_n \in Label, d : Database
+& l_c, l_o, l_n \in Label, d : Database
 \end{align*}
@@
 \begin{align*}
-	\mathbf{alterTable} : Label \times Label \times Database \rightarrow Database \\
-	alterTable(l_o, l_n, d) = \inference{\exists t \in tables(d) : name(d) = l_o \\ \wedge \forall e \in tables(d) : name(e) \neq l_n}{\begin{gathered}
+&	\mathbf{alterTable} : Label \times Label \times Database \rightarrow Database \\
+&	alterTable(l_o, l_n, d) = \inference{\exists t \in tables(d) : name(d) = l_o \\ \wedge \forall e \in tables(d) : name(e) \neq l_n}{\begin{gathered}
 		t.name = l_n \\ \wedge \forall e \in tables(a) \wedge \forall (f \in fks(e) \wedge reference(f) = l_o ) \\ : f.reference = l_n 
 	\end{gathered}}\\
-	l_o, l_n \in Label, s \in States, a : Applications, d : Database
+&	l_o, l_n \in Label, s \in States, a : Applications, d : Database
 \end{align*}
@@
 \begin{align*}
-	\mathbf{copyColumn} : Label \times Label \times Label \times Database \rightarrow Database \\
-	copyColumn(l_s, l_t, l_c, d) = \begin{cases}
+&	\mathbf{copyColumn} : Label \times Label \times Label \times Database \rightarrow Database \\
+&	copyColumn(l_s, l_t, l_c, d) = \begin{cases}
  		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{d \bullet addColumn(l_t, c)}
 	\end{cases} \\
-	l_s,l_t,l_c \in Label \wedge l_s \neq l_t,d : Database
+&	l_s,l_t,l_c \in Label \wedge l_s \neq l_t,d : Database
 \end{align*}
@@
 \begin{align*}
-	\mathbf{copyTable} : Label \times Label \times Database \times Database \\
-	copyTable(l_s, l_t, d) = \inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \forall e \in tables(d) : name(e) \neq l_t}{d \bullet addTable(tab(l_t, primaryKey(t), columns(t), foreignKeys(t)))}
+&	\mathbf{copyTable} : Label \times Label \times Database \times Database \\
+&	copyTable(l_s, l_t, d) = \inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \forall e \in tables(d) : name(e) \neq l_t}{d \bullet addTable(tab(l_t, primaryKey(t), columns(t), foreignKeys(t)))}
 \end{align*}
@@
 \begin{align*}
-	\mathbf{insertRow} : Row* \times Database \rightarrow Database \\
-	insertRow(r, d) = \begin{cases}
+&	\mathbf{insertRow} : Row* \times Database \rightarrow Database \\
+&	insertRow(r, d) = \begin{cases}
 			%\inference{|r|>1}{d \bullet insertRow(head(r), d) \bullet insertRow(tail(r), d)} \\ \\
@@
 		 \end{cases}	\\
-		 r = Row*,d : Database
+&		 r = Row*,d : Database
 \end{align*}
@@
 \begin{align*}
-	\mathbf{insertValues} : Row \times Pair* \times Database \rightarrow Daatabase \\
-	insertValues(r, p, d) = \begin{cases}
+&	\mathbf{insertValues} : Row \times Pair* \times Database \rightarrow Daatabase \\
+&	insertValues(r, p, d) = \begin{cases}
  		%\inference{|p| > 1}{\begin{gathered} 	d \bullet insertValues(r, head(p), d) \\ \bullet insertValues(r, tail(p), d)\end{gathered}}   \\ \\
@@
  		\end{cases} \\
-r : Row, p = Pair*,d : Databases
+& r : Row, p = Pair*,d : Databases
 \end{align*}
@@
 \begin{align*}
-	\mathbf{addValueToTable} : Label \times Pair \times Database \rightarrow Database \\
-	addValueToTable(l, p, d) = \begin{cases}
+&	\mathbf{addValueToTable} : Label \times Pair \times Database \rightarrow Database \\
+&	addValueToTable(l, p, d) = \begin{cases}
  		\inference{\exists t \in tables(d) : name(t) = l}	{\begin{gathered}
@@
  	\end{cases}\\ 
- 	l \in Label, p : Pair,  d : Database
+& 	l \in Label, p : Pair,  d : Database
 \end{align*}
@@
 \begin{align*}
-	\mathbf{copyValue} : Label \times Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
-	copyValue(l_s, l_t, l_{sc}, l_{tc} o_{l_s, l_t}, d) = \begin{cases}
+&	\mathbf{copyValue} : Label \times Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
+&	copyValue(l_s, l_t, l_{sc}, l_{tc} o_{l_s, l_t}, d) = \begin{cases}
 		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc}}{\begin{gathered}
@@
  	\end{cases} \\
- 	l_s, l_t, l_{sc}, l_{tc} \in Label, d : Database
+& 	l_s, l_t, l_{sc}, l_{tc} \in Label, d : Database
 \end{align*}
@@
 \begin{align*}
-	\mathbf{copyRows} : Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
-	copyRows(l_s, l_t, m_{l_s, l_t}, d) = \begin{cases}
+&	\mathbf{copyRows} : Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
+&	copyRows(l_s, l_t, m_{l_s, l_t}, d) = \begin{cases}
  	%\inference{|m_{l_s, l_t}| = \varnothing}{d} \\\\
@@
   \end{cases} \\
-  l_s, l_t \in Label, d : Database, m_{l_s, l_t} \in M
+&  l_s, l_t \in Label, d : Database, m_{l_s, l_t} \in M
 \end{align*}
@@
 \begin{align*}
-	\mathbf{dropValue} : Label \times Row \times Database \rightarrow Database \\
-	dropValue(l, r, d) = \inference{\exists p \in pairs(r) : reference(p) = l}{pairs(r) \setminus p}
-	\\ l \in Label, r : Row \wedge r \in data(d), d : Database
-\end{align*}
-\begin{align*}
-	\mathbf{dropRow} :  Row*  \times Database \rightarrow Database\\
-	dropRow(r, d) = \inference{|r|=1}{data(d) \setminus r}
-	\\ l \in Label, r : Row \wedge r \in data(d), d : Database
+&	\mathbf{dropValue} : Label \times Row \times Database \rightarrow Database \\
+&	dropValue(l, r, d) = \inference{\exists p \in pairs(r) : reference(p) = l}{pairs(r) \setminus p}\\ 
+& l \in Label, r : Row \wedge r \in data(d), d : Database
+\end{align*}
+\begin{align*}
+&	\mathbf{dropRow} :  Row*  \times Database \rightarrow Database\\
+&	dropRow(r, d) = \inference{|r|=1}{data(d) \setminus r}	\\ 
+& l \in Label, r : Row \wedge r \in data(d), d : Database
 \end{align*}
@@
 \begin{align*}
-	\mathbf{dropTable} : Label \times Database \rightarrow Database \\
-	dropTable(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \forall r \in data(d) : owningTable(r) \neq l \\  \wedge (\forall j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) \neq l) }{ tables(d) \setminus \rho(c)}
-\end{align*}
-
-\begin{align*}
-	\mathbf{dropTableAndData} : Labe \times Database \rightarrow Database \\
-	dropTableAndData(l, d) = \inference{\exists t \in tables(d) \\ \wedge ! \; associated(t, d)}{\begin{gathered}
+&	\mathbf{dropTable} : Label \times Database \rightarrow Database \\
+&	dropTable(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \forall r \in data(d) : owningTable(r) \neq l \\  \wedge (\forall j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) \neq l) }{ tables(d) \setminus \rho(c)}
+\end{align*}
+
+\begin{align*}
+&	\mathbf{dropTableAndData} : Labe \times Database \rightarrow Database \\
+&	dropTableAndData(l, d) = \inference{\exists t \in tables(d) \\ \wedge ! \; associated(t, d)}{\begin{gathered}
 	(\forall r \in tableData(t,d) : dropRow(r, d))\\ \bullet dropTable(t,d)
 \end{gathered}}
-	\\ l \in Label, d : Database
-\end{align*}
-\begin{align*}
-	\mathbf{dropColumn} : Label \times Label \times Database \rightarrow Database \\
-	dropColumn(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in columns(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{columns(t) \setminus c}\\
-	 l , l_p\in Label, d : Database 
-\end{align*}
-
-\begin{align*}
-	\mathbf{dropKey} : Label \times Label \times Database \rightarrow Database \\
-	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) \setminus c}\\
-	 l, l_p \in Label, d : Database 
-\end{align*}
-\begin{align*}
-	\mathbf{DropColumnAndData} : Label \times Label* \times Database \rightarrow Database \\
-	DropColumnAndData(l_c, ps, d) = \begin{cases}
+	\\& l \in Label, d : Database
+\end{align*}
+\begin{align*}
+&	\mathbf{dropColumn} : Label \times Label \times Database \rightarrow Database \\
+& 	dropColumn(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in columns(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{columns(t) \setminus c}\\
+&	 l , l_p\in Label, d : Database 
+\end{align*}
+
+\begin{align*}
+&	\mathbf{dropKey} : Label \times Label \times Database \rightarrow Database \\
+&	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) \setminus c}\\
+&	 l, l_p \in Label, d : Database 
+\end{align*}
+\begin{align*}
+&	\mathbf{DropColumnAndData} : Label \times Label* \times Database \rightarrow Database \\
+&	DropColumnAndData(l_c, ps, d) = \begin{cases}
  	%\inference{ps = \varnothing}{d}\\
@@
 	\\
-	l_c \in Label, ps : Label*, d \in Database
-\end{align*}
-
-\begin{align*}
-	\mathbf{DropKeyAndData} : Label \times Label* \times Database \rightarrow Database \\
-	DropKeyAndData(l_f, ps, d) = \begin{cases}
+&	l_c \in Label, ps : Label*, d \in Database
+\end{align*}
+
+\begin{align*}
+&	\mathbf{DropKeyAndData} : Label \times Label* \times Database \rightarrow Database \\
+&	DropKeyAndData(l_f, ps, d) = \begin{cases}
  	%\inference{ps = \varnothing}{d}\\
@@
 	\\
-	l_f \in Label, ps : Label*, d \in Database
+&	l_f \in Label, ps : Label*, d \in Database
 \end{align*}
@@
 \begin{align*}
-	\mathbf{addProperty} : Label \times Property \times State \rightarrow State \\
-	\Psi(s, addProperty(l, p)) = application(s) \circ addProp(l,p) \\
-	\Phi(s, addProperty(l, p)) = \begin{cases}
+&	\mathbf{addProperty} : Label \times Property \times State \rightarrow State \\
+&	\Psi(s, addProperty(l, p)) = application(s) \circ addProp(l,p) \\
+&	\Phi(s, addProperty(l, p)) = \begin{cases}
     		\inference{cardinality(p) \leq 1}{database(s) \bullet addColumn(l, \rho(p,d), d)} \\ \\
@@
    \end{cases}\\
-	s : State, l \in Label, p \in Properties
+&	s : State, l \in Label, p \in Properties
 \end{align*}
@@
 \begin{align*}
-\mathbf{addAsociation} : Label \times Association \times (Row \times Row) \times State \rightarrow State \\
-\Psi(s, addAsociation(l, a, m_{l, reference(a)})) = application(s) \circ addAssoc(l, a) \\
-\Phi(s, addAsociation(l, a, m_{l, reference(a)})) = \begin{cases}
+& \mathbf{addAsociation} : Label \times Association \times (Row \times Row) \times State \rightarrow State \\
+& \Psi(s, addAsociation(l, a, m_{l, reference(a)})) = application(s) \circ addAssoc(l, a) \\
+& \Phi(s, addAsociation(l, a, m_{l, reference(a)})) = \begin{cases}
 	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l\\ 		\wedge name(t_t) = reference(a) \\ \wedge \forall f \in foreignKeys(t_t) \cup columns(t_t) \\: name(f) \neq name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
@@
 \begin{align*}
-\mathbf{newClass} : Class \times State \rightarrow State \\
-\Psi(s, newClass(c)) = application(s) \circ addClass(c) \\
-\Phi(s, newClass(c)) = database(s)  \bullet addTable(\rho(c,d), d) \\ \bullet addColumn(name(c), \{\forall r \in singleProperties(c) | \rho(r,d)\}, d) \\ \bullet addTable(name(c), \{\forall r \in collections(c) | \rho(r,d)\}, d)
+& \mathbf{newClass} : Class \times State \rightarrow State \\
+& \Psi(s, newClass(c)) = application(s) \circ addClass(c) \\
+& \Phi(s, newClass(c)) = database(s)  \bullet addTable(\rho(c,d), d) \\ 
+& \bullet addColumn(name(c), \{\forall r \in singleProperties(c) | \rho(r,d)\}, d) \\
+& \bullet addTable(name(c), \{\forall r \in collections(c) | \rho(r,d)\}, d)  
 \end{align*}
@@
 \begin{align*}
-	\mathbf{RemoveClass}: Label \times Software \rightarrow State  \\
-	\Psi RemoveClass(l, s) = applications(s) \circ remClass(l) \\
-	\Phi RemoveClass(l, s) = database(s) \bullet dropTable(l) \\
-	l \in Label, s \in Software 
+&	\mathbf{RemoveClass}: Label \times Software \rightarrow State  \\
+&	\Psi RemoveClass(l, s) = applications(s) \circ remClass(l) \\
+&	\Phi RemoveClass(l, s) = database(s) \bullet dropTable(l) \\
+&	l \in Label, s \in Software 
 \end{align*}
@@
 \begin{align*}
-	\mathbf{RemoveProperty} : Label \times Label \times State \rightarrow State \\
-	\Psi RemoveProperty(l, l_p, s) =  applications(s) \circ  remProp(l, l_p) \\
-	\Phi RemoveProperty(l, l_p, s) = \begin{cases}
+&	\mathbf{RemoveProperty} : Label \times Label \times State \rightarrow State \\
+&	\Psi RemoveProperty(l, l_p, s) =  applications(s) \circ  remProp(l, l_p) \\
+&	\Phi RemoveProperty(l, l_p, s) = \begin{cases}
 		 \inference{\exists c \in classes(application(s)) : name(c) = l\\ 
@@
  	\end{cases}\\
- 	l, l_p \in Label, s : State
+&	l, l_p \in Label, s : State
 \end{align*}
@@
 \begin{align*}
-	\mathbf{RemovePropertyAndData} : Label \times Label* \times State \rightarrow State \\
-	\Phi(RemovePropertyAndData(L_c, ps), s) = application(s)\\ \circ remProperty(l_s, ps) \\
-	\Psi(RemovePropertyAndData(L_c, ps), s) = \begin{cases}
+&	\mathbf{RemovePropertyAndData} : Label \times Label* \times State \rightarrow State \\
+&	\Phi(RemovePropertyAndData(L_c, ps), s) = application(s)\\ 
+&  \circ remProperty(l_s, ps) \\
+&	\Psi(RemovePropertyAndData(L_c, ps), s) = \begin{cases}
 		%\inference{ps = \varnothing}{d}\\
@@
 	\\
-	l_c \in Label, ps : Label*, s : State
+&	l_c \in Label, ps : Label*, s : State
 \end{align*}
@@
 \begin{align*}
-    \mathbf{RemoveAssociation} : Label \times Labels \times State \rightarrow State \\ 
-    \Psi RemoveAssociation(l, l_a, s) = application(s) \circ remAss(l, l_a) \\
-    \Phi RemoveAssociation(l, l_a, s) = \begin{cases}
+&    \mathbf{RemoveAssociation} : Label \times Labels \times State \rightarrow State \\ 
+&    \Psi RemoveAssociation(l, l_a, s) = application(s) \circ remAss(l, l_a) \\
+&    \Phi RemoveAssociation(l, l_a, s) = \begin{cases}
      	\inference{\exists c \in classes(application(s)) : name(c) = l \\
@@
      \end{cases}\\
-    l_c, l_a \in Labels, s : State
+&    l_c, l_a \in Labels, s : State
 \end{align*}
@@
 \begin{align*}
-	\mathbf{RemoveAssociationAndData} : Label \times Label* \times State \rightarrow State \\
-	\Phi(RemoveAssociationAndData(L_c, ps), s) = application(s)\\ \circ remAssociation(l_s, ps) \\
-	\Psi(RemoveAssociationAndData(L_c, ps), s) = \begin{cases}
+&	\mathbf{RemoveAssociationAndData} : Label \times Label* \times State \rightarrow State \\
+&	\Phi(RemoveAssociationAndData(L_c, ps), s) = application(s)\\ 
+& \circ remAssociation(l_s, ps) \\
+&	\Psi(RemoveAssociationAndData(L_c, ps), s) = \begin{cases}
 		%\inference{ps = \varnothing}{d}\\
@@
 \begin{align*}
-\mathbf{RenameClass} : Label \times Label \times State \rightarrow State \\ 
-\Psi(RenameClass(l_o, l_n), s) = application(s) \circ renClass(l_o, l_n) \\
-\Phi(RenameClass(l_o, l_n), s) = database(s) \bullet alterTable(l_o, l_n) \\ \\
+& \mathbf{RenameClass} : Label \times Label \times State \rightarrow State \\ 
+& \Psi(RenameClass(l_o, l_n), s) = application(s) \circ renClass(l_o, l_n) \\
+& \Phi(RenameClass(l_o, l_n), s) = database(s) \bullet alterTable(l_o, l_n) \\ \\
 \end{align*}
@@
 \begin{align*}
-	\mathbf{RenameProperty} : Label \times Label \times Label \times State \rightarrow State \\ \\
-\Psi(RenameProperty(l_c, l_o, l_n, s)) = application(s) \circ renProp(l_c, l_o, l_n) \\
-\Phi(RenameProperty(l_c, l_o, l_n, s)) = \begin{cases}
+&	\mathbf{RenameProperty} : Label \times Label \times Label \times State \rightarrow State \\ \\
+& \Psi(RenameProperty(l_c, l_o, l_n, s)) = application(s) \circ renProp(l_c, l_o, l_n) \\
+& \Phi(RenameProperty(l_c, l_o, l_n, s)) = \begin{cases}
  	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 
@@
  \end{cases}\\
-l_c, l_o, l_n \in Label,  s : State
+& l_c, l_o, l_n \in Label,  s : State
 \end{align*}
@@
 \begin{align*}
-\mathbf{RenameAssociation} : Label \times Label \times Label \times State \rightarrow State \\
-\Psi(RenameAssocaiation(l_c, l_o, l_n), s) = application(s) \circ renAssoc(l_c, l_o, l_n)\\
-\Phi(RenameAssocaiation(l_c, l_o, l_n), s) = \begin{cases}
+& \mathbf{RenameAssociation} : Label \times Label \times Label \times State \rightarrow State \\
+& \Psi(RenameAssocaiation(l_c, l_o, l_n), s) = application(s) \circ renAssoc(l_c, l_o, l_n)\\
+& \Phi(RenameAssocaiation(l_c, l_o, l_n), s) = \begin{cases}
 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 
@@
  \end{cases} \\
-	l_c, l_o, l_n \in Label, a : Applications, s : Sate
+&	l_c, l_o, l_n \in Label, a : Applications, s : Sate
 \end{align*}
@@
 \begin{align*}
-	\mathbf{CopyProperty} : Label \times Label \times Label \times (Row \times Row) \times State \rightarrow State \\
-	\Psi(CopyProperty(l_s, l_t, l_p, m_{l_s, l_t}), s) = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)} 
+&	\mathbf{CopyProperty} : Label \times Label \times Label \times (Row \times Row) \times State \rightarrow State \\
+&	\Psi(CopyProperty(l_s, l_t, l_p, m_{l_s, l_t}), s) = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)}
 	\\\\
-	\Phi(CopyProperty(l_s, l_t, l_p, m_{l_s, l_t}), s) = \begin{cases}
+&	\Phi(CopyProperty(l_s, l_t, l_p, m_{l_s, l_t}), s) = \begin{cases}
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) \leq 1 }{\begin{gathered}
@@
  \end{cases}\\
- l_s, l_t, l_p \in Label,  m_{l_s, l_t}) \in M, s : State
+& l_s, l_t, l_p \in Label,  m_{l_s, l_t}) \in M, s : State
 \end{align*}
@@
 \begin{align*}
-	\mathbf{MoveProperty} : Label \times Label \times Label* \times (Row \times Row) \times State \rightarrow State \\
-	s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) = \begin{cases}
+&	\mathbf{MoveProperty} : Label \times Label \times Label* \times (Row \times Row) \times State \rightarrow State \\
+&	s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) = \begin{cases}
  	%\inference{|ps| = \varnothing}{state(s)}\\\\
@@
 	\\
-	l_s, l_t \in Label, ps \in Label*,  m_{l_s, l_t} \in M, s : State
+&	l_s, l_t \in Label, ps \in Label*,  m_{l_s, l_t} \in M, s : State
 \end{align*}
@@
 \begin{align*}
-	\mathbf{Extract} : Label \times Label* \times Label \times State \rightarrow State \\
-	state(s) \oplus Extract(l_s, ps, l_n) = \begin{cases}
+&	\mathbf{Extract} : Label \times Label* \times Label \times State \rightarrow State \\
+&	state(s) \oplus Extract(l_s, ps, l_n) = \begin{cases}
 	\inference{|ps| = 1}{\begin{gathered}
-		s \oplus addClass(cl(l_n, \varnothing, \varnothing)) \\ \oplus MoveProperty(l_s, l_n, ps, \mathbf{X}) \\ \oplus RemovePropertyData(l_s, ps)   
-\end{gathered}}\\\\
+		s \oplus addClass(cl(l_n, \varnothing, \varnothing)) \\ \oplus MoveProperty(l_s, l_n, ps, \mathbf{X}) \end{gathered}}\\\\
 	%\inference{|ps| > 1}{\begin{gathered} 	s \oplus Extract(l_s, head(ps), l_n) \\\oplus Extract(l_s, tail(ps), l_n) \end{gathered}}\\\\
@@
 	\\
-	l_s, l_n \in Label, ps : Label* s : State 
-\end{align*}
-
-\begin{align*}
-	\mathbf{Inline} : Label \times Label \times State \rightarrow State \\
-	state(s) \oplus Inline(l_s, l_t, m_{l_s, l_t}) = \inference{\exists c \in classes(application(s)) : name(c) = l_s \\\wedge !\:associated(l_s, s) \\ \wedge foreignKeys(c) = \varnothing  }{\begin{gathered}
-	s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemovePropertyAndData(l_s, ps) \\ \oplus RemoveClass(l_s)  
+&	l_s, l_n \in Label, ps : Label* s : State 
+\end{align*}
+
+\begin{align*}
+&	\mathbf{Inline} : Label \times Label \times State \rightarrow State \\
+&	state(s) \oplus Inline(l_s, l_t, m_{l_s, l_t}) = \inference{\exists c \in classes(application(s)) : name(c) = l_s \\\wedge !\:associated(l_s, s) \\ \wedge foreignKeys(c) = \varnothing  }{\begin{gathered}
+	s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemoveClass(l_s) 
 \end{gathered}
 }\\
-l_s, l_t \in Label, m_{l_s, l_t} \in M, s : State, \\ \\
-ps = \inference{\exists c \in classes(applications)) \wedge ps = \varnothing}{\forall p \in properries(p) : ps \cup name(p)}
+& l_s, l_t \in Label, m_{l_s, l_t} \in M, s : State, \\ \\
+& ps = \inference{\exists c \in classes(applications)) \wedge ps = \varnothing}{\forall p \in properries(p) : ps \cup name(p)}
 \end{align*}

### 2013-01-24 doc/description.tex
formating

@@
 & \mathbf{software} = < Application, Database, \rho >\\
-& \mathbf{sw} : Applications \times Databases \times ORM \rightarrow  Software \\ \\
-& \mathbf{application} : Software \rightarrow Application \\
+& \mathbf{sw} \in Applications \times Databases \times ORM \rightarrow  Software \\ \\
+& \mathbf{application} \in Software \rightarrow Application \\
 & application(sw(a, d, \rho)) = a \\ \\
-& \mathbf{database} : Software \rightarrow Database \\
+& \mathbf{database} \in Software \rightarrow Database \\
 & database(sw(a, d, \rho)) = d \\\\
-& a : Application, d : Database
+& a \in Application, d \in Database
 \end{align*}
@@
 & 	\mathbf{state} = Consistent\; s \: | \: Inconsistent \; s,  \\\\
-& 	s : State \vee s \lh Application \vee s : Database
+& 	s \in State \vee s \in Application \vee s \in Database
 \end{align*}
@@
 \begin{align*}
-& 	\mathbf{consistent} : State \rightarrow Boolean \\
+& 	\mathbf{consistent} \in State \rightarrow Boolean \\
 & 	consistent(s) = \begin{cases}
@@
  \end{cases}\\
-& s : State
+& s \in State
 \end{align*}
@@
 \begin{align*}
-&	\mathbf{state} : Software \rightarrow State \\
+&	\mathbf{state} \in Software \rightarrow State \\
 &	state(s) = \begin{cases}
@@
  	\end{cases} \\ 
-& 	s : State
+& 	s \in State
 \end{align*}
@@
 \begin{align*}
-& \oplus : State \times E \rightarrow State
+& \oplus \in State \times E \rightarrow State
 \end{align*}
@@
 \begin{align*}
-& \circ : State \times AppT \rightarrow State \\
-& \bullet : State \times DbT \rightarrow State
+& \circ \in State \times AppT \rightarrow State \\
+& \bullet \in State \times DbT \rightarrow State
 \end{align*}
@@
 & state(s) \oplus t = \begin{cases}
- 	\inference{! consistent(state(s))}{u} \\ \\
+ 	\inference{! consistent(state(s))}{Inconsistent s} \\ \\
  	\inference{consistent(state(s))}{state(sw(\Psi(s, t), \Phi(s,t))} 
 \end{cases}\\ 
-& s : State, t \in E
+& s \in State, t \in E
 \end{align*}
@@
 \end{cases}\\ 
-& a : Application, e \in AppT
+& a \in Application, e \in AppT
 \end{align*}
@@
 \end{cases}\\ 
-& d : Application, e \in DbT
+& d \in Database, e \in DbT
 \end{align*}
@@
 The notation: $Type*$ is used to define an ordered set of values of concrete type (e.g.
-$C : Class*$ means c is an ordered set of classes).
+$C \in Class*$ means c is an ordered set of classes).
 
@@
  \end{cases}\\\\
-&  y : Software \vee y : Application \vee y : Database
+&  y \in Software \vee y \in Application \vee y \in Database
 \end{align*}
@@
 & \mathbf{app} : Class* \rightarrow  Application \\ \\
-& \mathbf{classes} : Application \rightarrow Class* \\
+& \mathbf{classes} \in Application \rightarrow Class* \\
 &   classes(app(cs)) = cs  \\\\
@@
 &  class(l, a) = c, c \in classes(a) \wedge name(c) = l \\ \\
-&  l \in Label,a : Applications, cs = Class* 
+&  l \in Label,a \in Applications, cs = \{x | x \in Class\} 
 \end{align*}
@@
 & 	associations(cl(l, props, assocs)) = assocs \\ 
-& 	l \in Label, props : Property*, assocs : Association*
+& 	l \in Label, props = \{ x | x \in Property\}, assocs = \{ x | x \in Association \}
 \end{align*}
@@
 &	primitives(cl(l, props, assocs)) = props', \\ 
-& props' \subset props,  \forall p \in props' : cardinality(p) \leq 1 \\ \\
+& props' \subset props,  \forall p \in props' \in cardinality(p) \leq 1 \\ \\
 &	\mathbf{collections} : Class \rightarrow Property* \\	
 & collections(cl(l, props, assocs)) = props',  \\
-& props' \subset props, \forall p \in props' : cardinality(p) > 1 \\ \\
+& props' \subset props, \forall p \in props' \in cardinality(p) > 1 \\ \\
 &	\mathbf{associated} :  Label \times Application \rightarrow Boolean \\
@@
  	\end{cases} \\ \\
-&	l \in Label, props = Property*, assocs = Association*,a : Applications
+&	l \in Label, props = \{ x | x \in Porperty\}, assocs  = \{ x | x \in Association\}, \\
+& a \in Applications
 \end{align*}
@@
 &	owningClass(p, a) = \inference{\exists c \in classes(a) \wedge p \in properties(c)}{c}  \\ \\
-&	l \in Label, t : AppType, val \in Values, n \in \mathbb{N^{*}}, m \in Boolean 
+&	l \in Label, t \in AppType, val \in Values, n \in \mathbb{N^{*}}, m \in Boolean 
 \end{align*}
@@
 &	owningClass(as, a) = \inference{\exists c \in classes(a) \wedge as \in associations(c)}{c}  \\ \\	
-&	l, ref \in Label,  n_1, n_2 \in \mathbb{N}, as : Association
+&	l, ref \in Label,  n_1, n_2 \in \mathbb{N}, as \in Association
 \end{align*}
@@
  	\end{cases}\\
-&	a : Application
+&	a \in Application
 \end{align*}
@@
 &	rows(db(tabs, rs)) = rs \\ \\
-&	tabs = Table*, rs = Row*, l \in Label,d : Database
+&	tabs  = \{ x | x \in Table\}, rs  = \{ x | x \in Row\}, l \in Label,d \in Database
 \end{align*}
@@
 &	foreignKeys(tab(l, pk, cols, fks)) = fks \\ \\
-& l \in Label, pk : PrimaryKey, cols = Column*, fks = ForeignKey*
+& l \in Label, pk \in PrimaryKey, cols  = \{ x | x \in Column\}, \\
+& fks  = \{ x | x \in ForeignKey\}
 \end{align*}
@@
 &	owningTable(c, d) = \inference{\exists tab \in tables(d) \wedge c \in columns(tab)}{tab} \\ \\
-& l \in Label, t : DbType, val \in Values,\\ 
-& cons : Constraint*, c : Column,d : Database
+& l \in Label, t \in DbType, val \in Values,\\ 
+& cons  = \{ x | x \in Constraint\}, c \in Column,d \in Database
 \end{align*}
@@
 &	owningTable(fk, d) = \inference{\exists tab \in tables(d) \wedge fk \in foreignKeys(tab)}{tab}\\\\
-&	l, tRef \in Label, cons : Constraint*,d : Database, fk : ForeignKey
+&	l, tRef \in Label, cons  = \{ x | x \in Constraint\}, \\ 
+& d \in Database, fk = \{ x | x \in ForeignKey\}
 \end{align*}
@@
 &	pkValue(row(l, n, ps)) = n \\ \\
-&	 l \in Label, ps = Pair*, n \in N
+&	 l \in Label, ps = \{ x | x \in Pair\}, n \in N
 \end{align*}
@@
  		\end{cases} \\ \\
-& l \in Label, ps = Pair*, t \in tables(d), d : Database
+& l \in Label, ps  = \{ x | x \in Pair\}, t \in tables(d), d \in Database
 \end{align*}
@@
  \end{cases} \\ \\
-&	r : Row*, l \in Label
+&	r \in Row*, l \in Label
 \end{align*}
@@
  \end{cases} \\ \\
-&	 c \in classes(application(s)), s : State 
+&	 c \in classes(application(s)), s \in State 
 \end{align*}
@@
 \begin{align*}
-&	\mathbf{consistent} : Database \rightarrow Boolean \\
+&	\mathbf{consistent} \in Database \rightarrow Boolean \\
 &	state(d) = \begin{cases}
@@
  	false
- \end{cases}
+ \end{cases}\\
+& d \in Database
 \end{align*}
@@
 \begin{align*}
-&	\mathbf{\rho_{app}} : Application \rightarrow Database\\
+&	\mathbf{\rho_{app}} \in Application \rightarrow Database\\
 &	\rho_{app}(a) = \inference{state(a) = Consistent \; a}{\begin{gathered}
@@
 	\bullet (\forall c \in classes(a), e \in associations(c) : addAssociaiton(name(c), e) 
-\end{gathered}}
+\end{gathered}}\\
+& a \in Application
 \end{align*}
@@
 		\inference{|c| = 1}{table(name(c), primaryKey("Id"), \varnothing, \varnothing)}  \end{cases}   \\ 	
-& c : Class
+& c \in Class
 \end{align*}
@@
 	\end{cases}\\
-&	p : Property
+&	p \in Property
 \end{align*}
@@
  	\end{cases}\\
-& 	p : Property
+& 	p \in Property
 \end{align*}
@@
 	 \end{cases}\\
-&	 a : Association
+&	 a \in Association
 \end{align*}
@@
  	\end{cases}\\
-&	 a : Association
+&	 a \in Association
 \end{align*}
@@
 %	findProperty(l_c, l_p, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{p} \\
-%	l_c, l_p \in Label, a : Application
+%	l_c, l_p \in Label, a \in Application
 %\end{align*}
@@
 %	findAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a): name(c) = l_c \\ \wedge \exists o \in associations(c) : name(o) = l_a}{o} \\
-%	l_c, l_a \in Label, a : Application
+%	l_c, l_a \in Label, a \in Application
 %\end{align*}
@@
 &	addProp(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(c) : name(q) \neq name(p) \\ \wedge \forall e \in associations(c) : name(e) \neq name(p)}{properties(c) \cup p } \\
-&	l \in Label, p : Property, a : Applications
+&	l \in Label, p \in Property, a \in Applications
 \end{align*}
@@
 	\wedge \exists k \in classes(a) : name(k) = reference(s)}{associations(c) \cup s}\\
-&	l \in Labels, s \in Associations, a : Applications
+&	l \in Labels, s \in Associations, a \in Applications
 \end{align*}
@@
 &	addClass(c, a) = \inference{\forall s \in classes(a) : name(s) \neq name(c) \\ \wedge associations(c) = \varnothing }{classes(a) \cup c} \\ \\
-&	c : Class, a : Applications
+&	c \in Class, a \in Applications
 \end{align*}
@@
 	}{properties(c) \setminus p }\\
-&	l_c, l_p \in Label, a : Applications 
+&	l_c, l_p \in Label, a \in Applications 
 \end{align*}
@@
 &	remAssoc(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) \setminus f} \\
-&	l_c, l_a \in Labels, a : Applications
+&	l_c, l_a \in Labels, a \in Applications
 \end{align*}
@@
 	} {classes(a) \setminus c }\\
-&	l \in Label,a : Applications
+&	l \in Label,a \in Applications
 \end{align*}
@@
 	\end{gathered}}\\
-&	l_c, l_o, l_n \in Label, a : Application
+&	l_c, l_o, l_n \in Label, a \in Application
 \end{align*}
@@
 	\end{gathered}}
-	\\ & l_c, l_o, l_n \in Label, a : Application
+	\\ & l_c, l_o, l_n \in Label, a \in Application
 \end{align*}
@@
 		name(c) = l_n \wedge \forall e \in classes(a) \\ \wedge \forall (f \in associations(e) \wedge reference(f) = l_o ) :\\ f.reference = l_n 
-	\end{gathered}}
+	\end{gathered}}\\
+& l_o, l_n \in Label, a \in Application	
 \end{align*}
@@
 %	findTable(l, d) = \inference{\exists t \in tables(d) \wedge name(t) = l}{t}  \\ \\
-%	l \in Label, d : Database
+%	l \in Label, d \in Database
 %\end{align*}
@@
  \end{cases}\\
-& t_1 : Table, t_2 : Table
+& t_1 \in Table, t_2 \in Table
 \end{align*}
@@
  \end{cases}\\
-& c_1 : Column, c_2 : Column
+& c_1 \in Column, c_2 \in Column
 \end{align*}
@@
  \end{cases}\\
-& f_1 : ForeignKey, f_2 : ForeignKey
+& f_1 \in ForeignKey, f_2 \in ForeignKey
 \end{align*}
@@
 		\end{cases}\\
-&		l \in Label, r : Row
+&		l \in Label, r \in Row
 \end{align*}
@@
 	 \end{cases}\\ 
-&	 l \in Label, c : Column*,d : Database
+&	 l \in Label, c \in Column*,d \in Database
 \end{align*}
@@
 	 \end{cases}\\ 
-&	t = Table*, d : Database
+&	t = \{x | x \in Table\}, d \in Database
 \end{align*}
@@
  \\
-& l_c, l_o, l_n \in Label, d : Database
+& l_c, l_o, l_n \in Label, d \in Database
 \end{align*}
@@
 	\end{gathered}}\\
-&	l_o, l_n \in Label, s \in States, a : Applications, d : Database
+&	l_o, l_n \in Label, s \in States, a \in Applications, d \in Database
 \end{align*}
@@
 	\end{cases} \\
-&	l_s,l_t,l_c \in Label \wedge l_s \neq l_t,d : Database
+&	l_s,l_t,l_c \in Label \wedge l_s \neq l_t,d \in Database
 \end{align*}
@@
 &	\mathbf{copyTable} : Label \times Label \times Database \times Database \\
-&	copyTable(l_s, l_t, d) = \inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \forall e \in tables(d) : name(e) \neq l_t}{d \bullet addTable(tab(l_t, primaryKey(t), columns(t), foreignKeys(t)))}
+&	copyTable(l_s, l_t, d) = \inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \forall e \in tables(d) : name(e) \neq l_t}{\begin{gathered} d \bullet addTable(tab(l_t, primaryKey(t), \\ 
+ columns(t), foreignKeys(t)))
+\end{gathered}}\\
+& l_s, l_t \in Label,  d \in Database
 \end{align*}
@@
 		 \end{cases}	\\
-&		 r = Row*,d : Database
+&		 r = \{ x | x \in Row\},d \in Database
 \end{align*}
@@
  		\end{cases} \\
-& r : Row, p = Pair*,d : Databases
+& r = \{ x | x \in Row\}, p = \{ x | x \in Pair\},d \in Databases
 \end{align*}
@@
  	\end{cases}\\ 
-& 	l \in Label, p : Pair,  d : Database
+& 	l \in Label, p \in Pair,  d \in Database
 \end{align*}
@@
  	\end{cases} \\
-& 	l_s, l_t, l_{sc}, l_{tc} \in Label, d : Database
+& 	l_s, l_t, l_{sc}, l_{tc} \in Label, d \in Database
 \end{align*}
@@
   \end{cases} \\
-&  l_s, l_t \in Label, d : Database, m_{l_s, l_t} \in M
+&  l_s, l_t \in Label, d \in Database, m_{l_s, l_t} \in M
 \end{align*}
@@
 &	dropValue(l, r, d) = \inference{\exists p \in pairs(r) : reference(p) = l}{pairs(r) \setminus p}\\ 
-& l \in Label, r : Row \wedge r \in data(d), d : Database
+& l \in Label, r : Row \wedge r \in data(d), d \in Database
 \end{align*}
@@
 &	dropRow(r, d) = \inference{|r|=1}{data(d) \setminus r}	\\ 
-& l \in Label, r : Row \wedge r \in data(d), d : Database
+& l \in Label, r \in Row \wedge r \in data(d), d \in Database
 \end{align*}
@@
 &	\mathbf{dropTable} : Label \times Database \rightarrow Database \\
-&	dropTable(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \forall r \in data(d) : owningTable(r) \neq l \\  \wedge (\forall j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) \neq l) }{ tables(d) \setminus \rho(c)}
+&	dropTable(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \forall r \in data(d) : owningTable(r) \neq l \\  \wedge (\forall j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) \neq l) }{ tables(d) \setminus \rho(c)}\\
+& l \in Label, d \in Database
 \end{align*}
@@
 \end{gathered}}
-	\\& l \in Label, d : Database
+	\\& l \in Label, d \in Database
 \end{align*}
@@
 & 	dropColumn(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in columns(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{columns(t) \setminus c}\\
-&	 l , l_p\in Label, d : Database 
+&	 l , l_p\in Label, d \in Database 
 \end{align*}
@@
 &	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) \setminus c}\\
-&	 l, l_p \in Label, d : Database 
+&	 l, l_p \in Label, d \in Database 
 \end{align*}
@@
 	\\
-&	l_c \in Label, ps : Label*, d \in Database
+&	l_c \in Label, ps = \{ x | x \in Label\}, d \in Database
 \end{align*}
@@
 	\\
-&	l_f \in Label, ps : Label*, d \in Database
+&	l_f \in Label, ps = \{x | x \in Label \}, d \in Database
 \end{align*}
@@
    \end{cases}\\
-&	s : State, l \in Label, p \in Properties
+&	s \in State, l \in Label, p \in Properties
 \end{align*}
@@
 & \mathbf{addAsociation} : Label \times Association \times (Row \times Row) \times State \rightarrow State \\
-& \Psi(s, addAsociation(l, a, m_{l, reference(a)})) = application(s) \circ addAssoc(l, a) \\
-& \Phi(s, addAsociation(l, a, m_{l, reference(a)})) = \begin{cases}
+& \Psi(s, addAsociation(l, a, m_{l_1, l_2})) = application(s) \circ addAssoc(l, a) \\
+& \Phi(s, addAsociation(l, a, m_{l_1, l_2})) = \begin{cases}
 	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l\\ 		\wedge name(t_t) = reference(a) \\ \wedge \forall f \in foreignKeys(t_t) \cup columns(t_t) \\: name(f) \neq name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
-		database(s) \\ \bullet addForeignKey(name(t_s), \rho(a,d), d) \\ \bullet \forall r \in m_{l, reference(a)} : \\ insertRow(row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]), d)
+		database(s) \\ \bullet addForeignKey(name(t_s), \rho(a,d), d) \\ \bullet \forall r \in m_{l_1, l_2} : \\ insertRow(row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]), d)
 	\end{gathered}}
@@
 \end{cases}\\
+& s \in Software, l_1, l_2 \in Label, m_{l, reference(a)} \wedge l_2 = reference(a)
 \end{align*}
@@
 & \Psi(s, newClass(c)) = application(s) \circ addClass(c) \\
-& \Phi(s, newClass(c)) = database(s)  \bullet addTable(\rho(c,d), d) \\ 
-& \bullet addColumn(name(c), \{\forall r \in singleProperties(c) | \rho(r,d)\}, d) \\
-& \bullet addTable(name(c), \{\forall r \in collections(c) | \rho(r,d)\}, d)  
+& \Phi(s, newClass(c)) = database(s)  \bullet addTable(\rho(c,d)) \\ 
+& \bullet addColumn(name(c), \{\forall r \in singleProperties(c) | \rho(r,d)\}, database(s)) \\
+& \bullet addTable(name(c), \{\forall r \in collections(c) | \rho(r,database(s))\})  \\
+& s \in Software, c \in Class
 \end{align*}
@@
  	\end{cases}\\
-&	l, l_p \in Label, s : State
+&	l, l_p \in Label, s \in State
 \end{align*}
@@
 	\\
-&	l_c \in Label, ps : Label*, s : State
+&	l_c \in Label, ps = \{x | x \in Label\}, s \in State
 \end{align*}
@@
      \end{cases}\\
-&    l_c, l_a \in Labels, s : State
+&    l_c, l_a \in Labels, s \in State
 \end{align*}
@@
 	\\
-&	l_c \in Label, ps : Label*, s : State
+&	l_c \in Label, ps : Label*, s \in State
 \end{align*}
@@
  \end{cases}\\
-& l_c, l_o, l_n \in Label,  s : State
+& l_c, l_o, l_n \in Label,  s \in State
 \end{align*}
@@
  \end{cases} \\
-&	l_c, l_o, l_n \in Label, a : Applications, s : Sate
+&	l_c, l_o, l_n \in Label, a \in Applications, s : Sate
 \end{align*}
@@
  \end{cases}\\
-& l_s, l_t, l_p \in Label,  m_{l_s, l_t}) \in M, s : State
+& l_s, l_t, l_p \in Label,  m_{l_s, l_t}) \in M, s \in State
 \end{align*}
@@
 	\\
-&	l_s, l_t \in Label, ps \in Label*,  m_{l_s, l_t} \in M, s : State
+&	l_s, l_t \in Label, ps \in Label*,  m_{l_s, l_t} \in M, s \in State
 \end{align*}
@@
 	\\
-&	l_s, l_n \in Label, ps : Label* s : State 
+&	l_s, l_n \in Label, ps : Label* s \in State 
 \end{align*}
@@
 }\\
-& l_s, l_t \in Label, m_{l_s, l_t} \in M, s : State, \\ \\
-& ps = \inference{\exists c \in classes(applications)) \wedge ps = \varnothing}{\forall p \in properries(p) : ps \cup name(p)}
+& l_s, l_t \in Label, m_{l_s, l_t} \in M, s \in State, \\ 
+& ps = \{x | x \in \forall c \in classes(applications)) : \forall p \in properries(c) : name(p))\}
 \end{align*}

### 2013-01-24 doc/description.tex
formating

@@
 A software implemented using an object-oriented language, which uses relational database as a data storage (as illustrated in the Figure \ref{fig:appStructure}) is object of interest of this paper.
-There are three important components of such a software: application (its persistent layer concretely), database (consisting of database schema and stored data) and an ORM (we will use $\rho$ as a symbol for ORM), therefore we define a software as a triple:
+There are three important components of such a software: application (its persistent layer concretely), database (consisting of database schema and stored data) and an ORM (we will use $\rho$ as a symbol for ORM), therefore we define a software as a triple consisting of an application and a database which are connected together by an ORM:
 \begin{align*}
 & \mathbf{software} = < Application, Database, \rho >\\
-& \mathbf{sw} \in Applications \times Databases \times ORM \rightarrow  Software \\ \\
+& \mathbf{sw} : Applications \times Databases \times ORM \rightarrow  Software \\ 
+\\
 & \mathbf{application} \in Software \rightarrow Application \\
@@
 & 	\mathbf{consistent} \in State \rightarrow Boolean \\
-& 	consistent(s) = \begin{cases}
- 		\inference{s = Consistent(s)}{true}\\\\
- 		\inference{s = Inconsistent(s)}{false}\\
- \end{cases}\\
-& s \in State
+& 	consistent(state(s)) = \begin{cases}
+ 		\inference{s = Consistent \; s}{true}\\\\
+ 		\inference{s = Inconsistent \; s}{false}\\
+ \end{cases}\\
+& s \in Software
 \end{align*}
@@
 \begin{align*}
-&	\mathbf{state} \in Software \rightarrow State \\
+&	\mathbf{state} : Software \rightarrow State \\
 &	state(s) = \begin{cases}
@@
  	\end{cases} \\ 
-& 	s \in State
+& 	s \in Software
 \end{align*}
@@
 & state(s) \oplus t = \begin{cases}
- 	\inference{! consistent(state(s))}{Inconsistent s} \\ \\
+ 	\inference{! consistent(state(s))}{Inconsistent \; s} \\ \\
  	\inference{consistent(state(s))}{state(sw(\Psi(s, t), \Phi(s,t))} 
 \end{cases}\\ 
-& s \in State, t \in E
+& s \in Software, t \in E
 \end{align*}
@@
 & state(a) \circ e = \begin{cases}
- 	\inference{!\;consistent(state(a))}{a} \\ \\
+ 	\inference{!\;consistent(state(a))}{Inconsistent \; a} \\ \\
  	\inference{consistent(state(a))}{e(a)} 
@@
 &  state(d) \bullet e = \begin{cases}
- 	\inference{!\;consistent(state(d))}{d} \\ \\
+ 	\inference{!\;consistent(state(d))}{Inconsistent \; d} \\ \\
  	\inference{consistent(state(d))}{e(d)} 
@@
 Next there is a set $Values$ containing all possible values, which can be used in the model.
-The notation: $Type*$ is used to define an ordered set of values of concrete type (e.g.
-$C \in Class*$ means c is an ordered set of classes).
-
-A lot of introduced transformations can be applied on an ordered set. The application proceeds on set consists in sequential application on set elements:
+
+A lot of introduced transformations can be applied on an ordered sets. The application proceeds on set consists in sequential application on set elements:
 \begin{align*}
@@
  \end{cases}\\\\
-&  y \in Software \vee y \in Application \vee y \in Database
+& y \in Software \vee y \in Application \vee y \in Database \\
+& x \in Type* 
 \end{align*}
@@
 \begin{align*}
-&	\mathbf{ForeignKey} = <label, tableRef, Constraint*> \\ 
+&	\mathbf{ForeignKey} = < Label, Label, Constraint*> \\ 
 &	\mathbf{fk} : Label \times Label \times Constraints \rightarrow ForeignKey \\ \\
@@
 \begin{align*}
-&	\mathbf{Row} = < refTable, pkValue, pair* > \\
+&	\mathbf{Row} = < Label, N, Pair* > \\
 &	\mathbf{row} : Label \times Pairs \rightarrow Row \\ \\
@@
 \begin{align*}
-&	\mathbf{Pair} = < refCol, value > \\
+&	\mathbf{Pair} = < Label, Value > \\
 &	[Label \times Value] \rightarrow Pair\\ \\
@@
  \end{cases} \\ \\
-&	 c \in classes(application(s)), s \in State 
+&	 c \in classes(application(s)), s \in Software 
 \end{align*}
@@
 &	state(d) = \begin{cases}
-&		\inference{ \forall t_1, t_2 \in tables(d) : name(t_1) \neq name(t_2) \\ \wedge \forall c_1, c_2 \in columns(t_1) : name(c_1) \neq name(c_2) \\ \wedge \forall f_1, f_2 \in foreignKeys(t_1) : name(f_1) \neq name (f_2) \\ \wedge \forall f \in foreignKeys(t_1) : \exists t \in tables(d) \wedge name(t) = reference(f) \\ \wedge \forall r \in data(d): ( \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) :\\ \exists u \in columns(t) \cup foreignKeys(t) \cup primaryKey(t) : \\ reference(p) = name(u) \\ \wedge \forall e \in columns(t) \cup foreignKeys(t) \\ \wedge NOTNULL \in constrain(e) : \exists q \in pairs(r) \wedge name(e) = reference(q)) \\ \wedge \forall g \in columns(t) \cup foreignKeys(t) \\ \wedge UNIQUE \in constrain(g) : \\ \forall u_1, u_2 \in rows(t) : \\ value(pairByName(name(g), u_1) \neq value(pairByName(name(g), u_2) }{true}
+&		\inference{ \forall t_1, t_2 \in tables(d) : name(t_1) \neq name(t_2) \\ \wedge \forall c_1, c_2 \in columns(t_1) : name(c_1) \neq name(c_2) \\ \wedge \forall f_1, f_2 \in foreignKeys(t_1) : name(f_1) \neq name (f_2) \\ \wedge \forall f \in foreignKeys(t_1) : \exists t \in tables(d) \wedge name(t) = reference(f) \\ \wedge \forall r \in data(d): ( \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) :\\ \exists u \in columns(t) \cup foreignKeys(t) \cup primaryKey(t) : \\ reference(p) = name(u) \\ \wedge \forall e \in columns(t) \cup foreignKeys(t) \\ \wedge NOTNULL \in constrain(e) : \exists q \in pairs(r) \\ \wedge name(e) = reference(q)) \\ \wedge \forall g \in columns(t) \cup foreignKeys(t) \\ \wedge UNIQUE \in constrain(g) : \\ \forall u_1, u_2 \in rows(t) : \\ value(pairByName(name(g), u_1) \neq value(pairByName(name(g), u_2) }{true}
  	\\ \\
@@
    db(\varnothing) \bullet addTable(classes(a)) \\ 
-	\bullet (\forall c \in classes(a), p \in properties(c) : addProperty(name(c), p) \\
-	\bullet (\forall c \in classes(a), e \in associations(c) : addAssociaiton(name(c), e) 
+	\bullet (\forall c \in classes(a), p \in properties(c) : \\ addProperty(name(c), p) \\
+	\bullet (\forall c \in classes(a), e \in associations(c) : \\ addAssociaiton(name(c), e) 
 \end{gathered}}\\
@@
 &	\mathbf{\rho_{c}} : Class*  \rightarrow Table* \\
-&	\rho_{c}(c) = \begin{cases}
+&	\rho_{c}(c) = %\begin{cases}
 		%\inference{c = \varnothing}{\varnothing}\\\\
 		%\inference{|c| > 1}{\rho_c(head(c)) \cup \rho_c(tail(c))} \\ \\
-		\inference{|c| = 1}{table(name(c), primaryKey("Id"), \varnothing, \varnothing)}  \end{cases}   \\ 	
-& c \in Class
+		\inference{|c| = 1}{table(name(c), primaryKey("Id"), \varnothing, \varnothing)} % \end{cases} 
+\\ 	
+& c \in Class*
 \end{align*}
@@
 		%\inference{|p| > 1}{\rho_{sp}(head(p)) \cup \rho_{sp}(tail(p))}\\\\
- 		\inference{cardinality(p) = 1 \wedge !\;mandatory(p)
+ 		\inference{|p| = 1 \wedge cardinality(p) = 1 \wedge !\;mandatory(p)
  			\\ \wedge \exists t : t = table(name(owningClass(p)), d) 
@@
 		\end{gathered}} \\ \\
-  		\inference{cardinality(p) = 1 \wedge mandatory(p)
+  		\inference{|p| = 1 \wedge cardinality(p) = 1 \wedge mandatory(p)
 			 \\ \wedge \exists t : t = table(name(owningClass(p)), d)
@@
 		%\inference{|p| > 1}{\rho_{cp}(head(p)) \cup \rho_{cp}(tail(p))}\\\\
-     	\inference{cardinality(p) > 1 \wedge !\;mandatory(p) \\ 
+     	\inference{|p| = 1 \wedge cardinality(p) > 1 \wedge !\;mandatory(p) \\ 
      		\wedge 	\exists t : t = table(name(owningClass(p)), d)
@@
     	\end{gathered}}\\ \\
-     	\inference{ cardinality(p) > 1 \wedge mandatory(p) \\ 
+     	\inference{|p| = 1 \wedge  cardinality(p) > 1 \wedge mandatory(p) \\ 
      		\wedge \exists t : t = table(name(owningClass(p)), d) 
@@
 		%\inference{|a| > 1}{\rho_{a}(head(a)) \cup \rho_{a}(tail(a))}\\\\
-		\inference{ startCardinality(a) = 0 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)
+		\inference{|a| = 1 \wedge  startCardinality(a) = 0 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)
 		}{
@@
   \\ \\
- 	 \inference{ startCardinality(a) = 1 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)
+ 	 \inference{|a| = 1 \wedge  startCardinality(a) = 1 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)
  	 }{ 
 		foreignKey(name(a), name(t),  NOTNULL)
-	}
-	 \end{cases}\\
+	}	 \end{cases}
+\\
 &	 a \in Association
@@
 &	\mathbf{\rho_a} : Association* \rightarrow Table* \\
-&	\rho_a(a) = \begin{cases}
+&	\rho_a(a) =
+% 	  \begin{cases}
  		%\inference{a = \varnothing}{\varnothing}\\
 		%\inference{|a| > 1}{\rho_{a}(head(a)) \cup \rho_{a}(tail(a))}\\\\
-		\inference{  startCardinality > 1 \\ \wedge \exists t \in tables(d) : name(owningClass(p)) = name(t) \\ \wedge \exists e \in tables(d) : name(reference(a)) = name(e)
+		\inference{|a| = 1 \wedge   startCardinality > 1 \\ \wedge \exists t \in tables(d) : name(owningClass(p)) = name(t) \\ \wedge \exists e \in tables(d) : name(reference(a)) = name(e)
   		}{\begin{gathered}  
@@
   		\end{gathered}}  
- 	\end{cases}\\
+ 	%\end{cases}
+\\
 &	 a \in Association
@@
 &	\mathbf{addTable} : Table* \times Database \rightarrow Database \\ 
-&	addTable(t, d) = \begin{cases}
+&	addTable(t, d) =
+		% \begin{cases}
 		%\inference{|t| > 1}{d \bullet addTable(head(t)) \bullet addTable(tail(t))}
-\\ \\	
-		\inference{|t| = 1 \wedge \forall e \in tables(d) \wedge name(t) \neq name(e) \\ \wedge \forall f \in foreignKeys(t) : \\ s\exists e \in tables(d) \wedge name(e) = reference(f) }{tables(d) \cup t}
+% \\ \\	
+		\inference{|t| = 1 \wedge \forall e \in tables(d) \wedge name(t) \neq name(e) \\ \wedge \forall f \in foreignKeys(t) : \\ \exists e \in tables(d) \wedge name(e) = reference(f) }{tables(d) \cup t}
 % \\ \\ 	Inconsistent \;d 
-	 \end{cases}\\ 
+	 %\end{cases}
+	 \\ 
 &	t = \{x | x \in Table\}, d \in Database

### 2013-01-26 doc/description.tex
formating, added notDefine state and begin merge

@@
 The paper is organized as follows: in Section \ref{sec:evoIntro} the concept of ORM software and its evolution is discussed, then models of application and database are introduced in Section \ref{sec:appModel} and in Section \ref{sec:dbModel}, the ORM is defined in Section \ref{sec:orm} and the evolutionary transformations are defined in Section \ref{sec:Evolutionaty-Transformations}.
+
+\subsection{Note on Used Notation}
+Before the definition of the software is provided, the notation of function transcription has to be introduced.
+Usually the definition of a function consists of two parts - an effect of the function and conditions when the effect occurs.
+We decide to use bit different transcription to simplify the notation and to improve understanding of functions, instead of usually used:
+\begin{align*}
+& f(x) = effect \; if \; condition 
+\end{align*}
+we use following notation:
+\begin{align*}
+& f(x) =   \inference{condition}{effect}
+\end{align*}
 
@@
 \begin{align*}
-& \mathbf{software} = < Application, Database, \rho >\\
-& \mathbf{sw} : Applications \times Databases \times ORM \rightarrow  Software \\ 
+& \mathbf{software} = ( Application, Database, \rho )\\
+& \mathbf{sw} : Application \times Database \times ORM \rightarrow  Software \\ 
 \\
@@
 The software has to be in consistent state so its users could have benefit from its usage.
-The state of the software is defined to communicate the sate of the software to its user.
+The state of the software is defined to communicate the sate of the software to its user, the symbol $\perp$ is used as a notation for inconsistent state.
 The components of a software (application and database) are transformed likewise, thus the state is defined for them too.
 \begin{align*}
-& 	\mathbf{state} = Consistent\; s \: | \: Inconsistent \; s,  \\\\
+& 	\mathbf{state} = Consistent\; s \: | \: \perp,  \\\\
 & 	s \in State \vee s \in Application \vee s \in Database
 \end{align*}
-Before the definition of the software consistency is provided, the notation of function transcription has to be introduced.
-Usually the definition of a function consists of two parts - an effect of the function and conditions when the effect occurs We decide to use following transcription, to simplify the notation and to improve understanding of functions:
-\begin{align*}
-f(x) =   \inference{conditions}{effect}
-\end{align*}
-
-The function querying the state (it means consistency) of a software is defined as:
-\begin{align*}
-& 	\mathbf{consistent} \in State \rightarrow Boolean \\
-& 	consistent(state(s)) = \begin{cases}
- 		\inference{s = Consistent \; s}{true}\\\\
- 		\inference{s = Inconsistent \; s}{false}\\
- \end{cases}\\
-& s \in Software
-\end{align*}
+
+%The function querying the state (it means consistency) of a software is defined as:
+%\begin{align*}
+%& 	\mathbf{consistent} : State \rightarrow Boolean \\
+%& 	consistent(state(s)) = \begin{cases}
+% 		\inference{s = Consistent \; s}{true}\\\\
+% 		\inference{s = \perp}{false}\\
+% \end{cases}\\
+%& s \in Software
+%\end{align*}
 The consistency of application and database is discussed later in this paper. The definition of software state only is provided:
@@
 &	state(s) = \begin{cases}
- 		\inference{consistent(application(s)) \wedge \\ consistent(database(s)) \\ \wedge \rho(application(s)) = database(s)}{Consistent \; s}\\ \\
- 		\inference{otherwise}{Inconsistent \; s}
+ 		\inference{state(application(s)) \neq \perp \\
+ 		\wedge state(database(s)) \neq \perp \\
+ 		\wedge \rho(application(s)) = database(s)}{Consistent \; s}\\ \\
+ 		\perp
  	\end{cases} \\ 
@@
 \begin{align*}
-& \circ \in State \times AppT \rightarrow State \\
-& \bullet \in State \times DbT \rightarrow State
+& \circ \in Application \times AppT \rightarrow State \\
+& \bullet \in Database \times DbT \rightarrow State
 \end{align*}
@@
 & state(s) \oplus t = \begin{cases}
- 	\inference{! consistent(state(s))}{Inconsistent \; s} \\ \\
- 	\inference{consistent(state(s))}{state(sw(\Psi(s, t), \Phi(s,t))} 
+ 	\inference{state(s) = Consistent \; s}{state(sw(\Psi(s, t), \Phi(s,t))} \\ \\
+ 	\perp
 \end{cases}\\ 
@@
 \begin{align*}
-& state(a) \circ e = \begin{cases}
- 	\inference{!\;consistent(state(a))}{Inconsistent \; a} \\ \\
- 	\inference{consistent(state(a))}{e(a)} 
+& a \circ e = \begin{cases}
+ 	\inference{!\;consistent(a)}{\perp} \\ \\
+ 	\inference{consistent(a)}{e(a)} 
 \end{cases}\\ 
@@
 \begin{align*}
-&  state(d) \bullet e = \begin{cases}
- 	\inference{!\;consistent(state(d))}{Inconsistent \; d} \\ \\
- 	\inference{consistent(state(d))}{e(d)} 
+&  d \bullet e = \begin{cases}
+ 	\inference{!\;consistent(d)}{\perp} \\ \\
+ 	\inference{consistent(d)}{e(d)} 
 \end{cases}\\ 
@@
  		\inference{|x| = 1}{effect}\\\\
- 		\inference{otherwise}{Inconsistent y}
+ 		\inference{otherwise}{\perp}
  \end{cases}\\\\
 & y \in Software \vee y \in Application \vee y \in Database \\
-& x \in Type* 
+& x \in {y | y \in Type}
 \end{align*}
@@
 The instances are stored in the relational database in our case, therefore we define only the structure of classes in our model.
-\begin{align*}
-& \mathbf{Application} = < Class* > \\ 
-& \mathbf{app} : Class* \rightarrow  Application \\ \\
-& \mathbf{classes} \in Application \rightarrow Class* \\
-&   classes(app(cs)) = cs  \\\\
-& \mathbf{class} : Label \times Application \rightarrow Class   \\ 
-&  class(l, a) = c, c \in classes(a) \wedge name(c) = l \\ \\
-&  l \in Label,a \in Applications, cs = \{x | x \in Class\} 
-\end{align*}
-
 \paragraph{Class} Class represents a basic structural unit in the application model. It has a unique name, one or more properties and it can be associated to other classes in the application. 
@@
 & props' \subset props, \forall p \in props' \in cardinality(p) > 1 \\ \\
-&	\mathbf{associated} :  Label \times Application \rightarrow Boolean \\
-& 	associated(l,a) = \begin{cases}
-  		\inference{\exists c \in classes(a) \wedge \exists e \in associations(c) : \\ reference(e) = l
-  	}{ true } \\\\
- 		false
- 	\end{cases} \\ \\
 &	l \in Label, props = \{ x | x \in Porperty\}, assocs  = \{ x | x \in Association\}, \\
-& a \in Applications
+& a \in Application
 \end{align*}
@@
 The property can be mandatory, can have a default value and according to its cardinality it can represent a single value or a collection of values.
-The $owningClass$ function returns the class in the model which owns the property.
 \begin{align*}
@@
 &	mandatory(prop(l, t, val, n, m)) = m  \\ \\
-&	\mathbf{owningClass} : Property \times Application \rightarrow Class  \\ 	
-&	owningClass(p, a) = \inference{\exists c \in classes(a) \wedge p \in properties(c)}{c}  \\ \\
 &	l \in Label, t \in AppType, val \in Values, n \in \mathbb{N^{*}}, m \in Boolean 
@@
 &	endCardinality(assoc(l, ref, n_1, n_2)) = n_2 \\ \\
-&	\mathbf{owningClass} : Association \times Application \rightarrow Class  \\
-&	owningClass(as, a) = \inference{\exists c \in classes(a) \wedge as \in associations(c)}{c}  \\ \\	
 &	l, ref \in Label,  n_1, n_2 \in \mathbb{N}, as \in Association
@@
 
+\paragraph{Application} Application is defined as a sequence of classes and it creates the context for all structures used in the software persistent layer.
+\begin{align*}
+& \mathbf{Application} = (Class*) \\ 
+& \mathbf{app} : Class* \rightarrow  Application \\ \\
+& \mathbf{classes} \in Application \rightarrow Class* \\
+&   classes(app(cs)) = cs  \\\\
+& \mathbf{class} : Label \times Application \rightarrow Class   \\ 
+&  class(l, a) = c, c \in classes(a) \wedge name(c) = l \\ \\
+&  l \in Label,a \in Application, cs = \{x | x \in Class\} 
+\end{align*}
+The application creates the context for classes so it is possible to determine if a class is associated by other classes or not:
+\begin{align*}
+& \mathbf{associated} : Label \times Application \rightarrow Boolean \\
+& 	associated(l,a) = \begin{cases}
+  		\inference{\exists c \in classes(a) \wedge \exists e \in associations(c) : \\ reference(e) = l
+  	}{ true } \\\\
+ 		false
+ 	\end{cases} \\ \\
+&	l \in Label, a \in Application
+\end{align*}
+The $owningClass$ function returns the class in the model which owns the property or association and it is used during the evolutionary transformation.
+\begin{align*}
+&	\mathbf{owningClass} : Property \times Application \rightarrow Class  \\ 	
+&	owningClass(p, a) = \inference{\exists c \in classes(a) \wedge p \in properties(c)}{c}  \\ \\
+&	p \in Property, a \in Application
+\end{align*}
+\begin{align*}
+&	\mathbf{owningClass} : Association \times Application \rightarrow Class  \\
+&	owningClass(as, a) = \inference{\exists c \in classes(a) \wedge as \in associations(c)}{c}  \\ \\	
+&	as \in Association, a \in Application
+\end{align*}
+
 \subsubsection{Application consistency}
@@
 \begin{align*}
-&	state(a) = \begin{cases}
-	\inference{\forall c_1, c_2 \in classes(a) : name(c_1) \neq name(c_2) \\ \wedge \forall p_1, p_2 \in properties(c_1) : name(p_1) \neq name(p_2) \\ \wedge \forall a_1, a_2 \in associations(c_1) : name(a_1) \neq name (a_2) \\ \wedge \forall a \in associations(c_1) : \exists c \in classes(a) \wedge name(c) = reference(a) }{Consistent \; a}\\\\
- 	Inconsistent \; a
+&	\mathbf{consistent} : Application \rightarrow Boolean \\
+&	consistent(a) = \begin{cases}
+	\inference{\forall c_1, c_2 \in classes(a) : name(c_1) \neq name(c_2) \\ \wedge \forall p_1, p_2 \in properties(c_1) : name(p_1) \neq name(p_2) \\ \wedge \forall a_1, a_2 \in associations(c_1) : name(a_1) \neq name (a_2) \\ \wedge \forall a \in associations(c_1) : \exists c \in classes(a) \wedge name(c) = reference(a)\\
+	\wedge \forall p \in properties(c_1), a \in associations(c_1) : name(p) \neq name(a) }{true}\\\\
+ 	flase
  	\end{cases}\\
@@
 \begin{align*}
-&	\mathbf{Database} = < Table*, Row*> \\
+&	\mathbf{Database} = ( Table*, Row* ) \\
 &	\mathbf{db} : Tables \times Rows \rightarrow Database \\ \\
@@
 \begin{align*}
-&	\mathbf{Table} = <label, primaryKey, Column*, ForeignKey*>\\
+&	\mathbf{Table} = (label, primaryKey, Column*, ForeignKey*)\\
 &	\mathbf{tab} 	: Label \times PrimaryKeys \times Columns \times ForeignKeys \rightarrow Table \\ \\
@@
 \begin{align*}
-&	\mathbf{Column} = <label, DbType, DefaultValue, Constraint*> \\
+&	\mathbf{Column} = (label, DbType, DefaultValue, Constraint*) \\
 &	\mathbf{col} : Label \times DbType \times String \times Constraints \rightarrow Column \\ \\
@@
 \begin{align*}
-&	\mathbf{PrimaryKey} =  < label > 	\\
-&	\mathbf{pk} : Label \rightarrow PrimaryKey
-\end{align*}
+&	\mathbf{PrimaryKey} =  ( Label, Sequence ) 	\\
+&	\mathbf{pk} : Label \times Sequence \rightarrow PrimaryKey \\\\
+&	\mathbf{name} : PrimaryKey \rightarrow Label \\
+&	name(pk(l, s)) = Id \\\\
+&	\mathbf{seq} : PrimaryKey \rightarrow Sequence \\
+&   seq(pk(l,s)) = s \\\\
+& l \in Label, s \in Sequence
+\end{align*}
+The sequence generates values of primary key.
+The new value of a key is obtained by calling the function $next(s)$.
+The sequence which provides numbers incremented by one is called $s_1$ if the increment is 1, $s_{10}$ if the increment is 10 etc.
 
@@
 \begin{align*}
-&	\mathbf{DbType} = DBSTRING | DBINT
+&	\mathbf{DbType} = DBSTRING \; | \; DBINT
 \end{align*}
@@
 
-\paragraph{Stored Data} A database consists not only from schema but also from data which are represented as rows in a table.
-A table row in our model is a tuple consisting of value pairs, which represents concrete value of concrete column.
+\subsubsection{Stored Data} 
+A database consists not only from schema but also from data which are represented as rows in a table.
+A table row in our model is a tuple consisting of value pairs, which represents concrete value of concrete column or key.
 Each row contains a name of a table it belongs to.
@@
 &	pairs(row(l, n, ps)) = ps \\ \\
-&	\mathbf{pkValue} : Row \rightarrow N \\
-&	pkValue(row(l, n, ps)) = n \\ \\
+&	\mathbf{idnetifier} : Row \rightarrow N \\
+&	idnetifier(row(l, n, ps)) = n \\ \\
 &	 l \in Label, ps = \{ x | x \in Pair\}, n \in N
@@
 \begin{align*}
-&	\mathbf{tableData} : Label \times Row* \rightarrow Row* \\
-&	tableData(l, r) = \begin{cases}
+&	\mathbf{instances} : Label \times Row* \rightarrow Row* \\
+&	instances(l, r) = \begin{cases}
  		%\inference{r = \varnothing}{\varnothing} \\
- 		%\inference{|r| > 1}{tableData(head(r)) \cup tableData(tail(r))}\\ \\
+ 		%\inference{|r| > 1}{instances(head(r)) \cup instances(tail(r))}\\ \\
 &		\inference{|r| = 1 \wedge refTable(r) = l}{r}\\\\
@@
 &	\mathbf{consistent} \in Database \rightarrow Boolean \\
-&	state(d) = \begin{cases}
+&	consistent(d) = \begin{cases}
 &		\inference{ \forall t_1, t_2 \in tables(d) : name(t_1) \neq name(t_2) \\ \wedge \forall c_1, c_2 \in columns(t_1) : name(c_1) \neq name(c_2) \\ \wedge \forall f_1, f_2 \in foreignKeys(t_1) : name(f_1) \neq name (f_2) \\ \wedge \forall f \in foreignKeys(t_1) : \exists t \in tables(d) \wedge name(t) = reference(f) \\ \wedge \forall r \in data(d): ( \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) :\\ \exists u \in columns(t) \cup foreignKeys(t) \cup primaryKey(t) : \\ reference(p) = name(u) \\ \wedge \forall e \in columns(t) \cup foreignKeys(t) \\ \wedge NOTNULL \in constrain(e) : \exists q \in pairs(r) \\ \wedge name(e) = reference(q)) \\ \wedge \forall g \in columns(t) \cup foreignKeys(t) \\ \wedge UNIQUE \in constrain(g) : \\ \forall u_1, u_2 \in rows(t) : \\ value(pairByName(name(g), u_1) \neq value(pairByName(name(g), u_2) }{true}
@@
 There are defined functions for object-relational mapping used in this paper.
-The mapping function itself are not able to create a database from the application as they are mentioned to be used together with evolutionary transformations introduced in section \ref{sec:transformations}.
+The mapping function itself are not able to create a database from the application as they are mentioned to be used together with evolutionary transformations introduced in section \ref{sec:Evolutionaty-Transformations}.
 
@@
 
-\paragraph{Add Property} The name collision only has to be verified, before adding a property into an application structure:
+\paragraph{Add Property} The name collision only has to be verified, before adding a property into an application structure.
+The transformation inputs are the property to be added and the name of the class the property is going to be added to.
 \begin{align*}
@@
 &	addProp(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in  properties(c) : name(q) \neq name(p) \\ \wedge \forall e \in associations(c) : name(e) \neq name(p)}{properties(c) \cup p } \\
-&	l \in Label, p \in Property, a \in Applications
+&	l \in Label, p \in Property, a \in Application
 \end{align*}
@@
 	\wedge \exists k \in classes(a) : name(k) = reference(s)}{associations(c) \cup s}\\
-&	l \in Labels, s \in Associations, a \in Applications
+&	l \in Labels, s \in Associations, a \in Application
 \end{align*}
@@
 &	addClass(c, a) = \inference{\forall s \in classes(a) : name(s) \neq name(c) \\ \wedge associations(c) = \varnothing }{classes(a) \cup c} \\ \\
-&	c \in Class, a \in Applications
-\end{align*}
-
-\paragraph{Remove Property or Association} The property or association can be simply removed from a class:
+&	c \in Class, a \in Application
+\end{align*}
+
+\paragraph{Remove Property or Association} The property or association can be simply removed from a class.
+The transformation inputs (in order) are the name of the property's owning class and the name of property itself.
 \begin{align*}
@@
 	}{properties(c) \setminus p }\\
-&	l_c, l_p \in Label, a \in Applications 
+&	l_c, l_p \in Label, a \in Application 
 \end{align*}
@@
 &	remAssoc(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) \setminus f} \\
-&	l_c, l_a \in Labels, a \in Applications
+&	l_c, l_a \in Labels, a \in Application
 \end{align*}
@@
 	} {classes(a) \setminus c }\\
-&	l \in Label,a \in Applications
-\end{align*}
-
-\paragraph{Rename Property and Association} Both renaming operations can change a name of the property, association respectively if there are no name collisions:
+&	l \in Label,a \in Application
+\end{align*}
+
+\paragraph{Rename Property and Association} Both renaming operations can change a name of the property, association respectively if there are no name collisions.
+The name of owning class, the name of the property (or association) and the new name of the property (association) are inputs of the transformation.
 \begin{align*}
@@
 \begin{align*}
-&	m_{l_1, l_2} =  <(r_1, r_2)*>,\\
-&	r_1 \in tableData(l_1, database(s)) \Rightarrow \exists p \in m_{l_1, l_2} \wedge first(p) = r_1 \\
-&	r_2 \in tableData(l_2, database(s)) \Rightarrow \exists p \in m_{l_1, l_2} \wedge second(p) = r_2\\
-&	l_1, l_2 \in Label, r_1, r_2 : Row
-\end{align*}
-The definition does not allow to add any default value in case the number of instances does not match.
-It is because the evolution transformations are designed to not affect stored data (instances) by creating a new data which does not origin from the software.
-%The features of both sets allows only pairs from existing instances and it assumes there is always at least one pair for each row from both tables.
-
-%\paragraph{Find Table} The function $findTable$ finds and returns a table in the database according to the provided label.
-%\begin{align*}
-%	\mathbf{findTable} : Label \times Database \rightarrow Table  \\
-%	findTable(l, d) = \inference{\exists t \in tables(d) \wedge name(t) = l}{t}  \\ \\
-%	l \in Label, d \in Database
-%\end{align*}
+&	m_{l_1, l_2} =  (r_1, r_2)*,\\
+& 	r_1 \in instances(l_1, d), r_2 \in instances(l_2, d) \\
+&	\forall r_1, r_1 \in instances(l_1, d) \Rightarrow \exists p \in m_{l_1, l_2} \wedge first(p) = r_1 \\
+&	\forall r_2, r_2 \in instances(l_2, d) \Rightarrow \exists p \in m_{l_1, l_2} \wedge second(p) = r_2\\\\
+&	l_1, l_2 \in Label, r_1, r_2 \in Row, d \in Database
+\end{align*}
+The mapping $m_{l_1, l_2}$ can be used if there are instances in both tables in database $d$, because there are cases when one of the tables is empty the identity mapping is defined:
+\begin{align*}
+&	m_{l, l} = (r, r) * \\
+& 	\forall r \in instances(l, d) \Rightarrow \exists p \in i_{l} \wedge first(p) = r \\\\
+&	l \in Label, r \in Row,  d \in Database
+\end{align*}
 
@@
 &	f_1 \cong f_2 = \begin{cases}
- 		\inference{name(f_1) = name(f_2) \\ \wedge reference(f_1) = reference(f_2) \\ \wedge constraints(f_1) = comstraints(f_2}{true}
+ 		\inference{name(f_1) = name(f_2) \\ \wedge reference(f_1) = reference(f_2) \\ \wedge constraints(f_1) = comstraints(f_2)}{true}
  		\\ \\
@@
 If the column has the $NOTNULL$ constraint, then the default value has to be added to all rows in the table.
+The inputs of the transformation are name of the table where the column(s) will be added and a sequence of column to add.
 \begin{align*}
@@
 \subsubsection{Database Schema Alternation}
-\paragraph{Alter Column Name}
+\paragraph{Alter Column Name} The transformation changes the name of the column identified by its name in given table.
+The inputs of the transformation are name of the table, name of the column and the new name.
 \begin{align*}
@@
 	\end{gathered}}\\
-&	l_o, l_n \in Label, s \in States, a \in Applications, d \in Database
-\end{align*}
-
-\paragraph{Copy Column} The transformation creates a copy of a column in a another table in the database. The transformation changes only the structure of database - stored data are not changed.
+&	l_o, l_n \in Label, s \in States, a \in Application, d \in Database
+\end{align*}
+
+\paragraph{Copy Column} The transformation creates a copy of a column structure in a another table in the database.
+The transformation changes only the structure of database - stored data are not changed.
+The transformation inputs are the name of the table which is the source for the transformation, the name of a column to be copied and the name of the target table.
+It means the copy operation does not change the name of copied column during its execution.
 \begin{align*}
@@
 
-\paragraph{Copy Table} The transformation creates a copy of a table in database schema with different name.
+\paragraph{Copy Table} The transformation creates a copy of a table in database schema with a given name.
+The new name has to be different than the old one otherwise the transformation could not be processed.
 \begin{align*}
@@
 \subsubsection{Data Manipulation}
-\paragraph{Insert row} The function allow to insert a row(s) into a database.
+\paragraph{Insert row} The function allow to insert a row or rows into the database.
 \begin{align*}
@@
 			%\inference{|r|>1}{d \bullet insertRow(head(r), d) \bullet insertRow(tail(r), d)} \\ \\
-		 	\inference{|r|=1 \wedge \exists t \in tables(d) : name(t) = reference(r) \\ \wedge \forall e \in tableData(t, d) : pkValue(e) \neq pkValue(r)}{data(d) \cup r} 
+		 	\inference{|r|=1 \wedge \exists t \in tables(d) : name(t) = reference(r) \\ \wedge \forall e \in instances(t, d) : idnetifier(e) \neq idnetifier(r)}{data(d) \cup r} 
 		 	%\\ \\	 Inconsistent \; d
@@
 
-\paragraph{Insert Value} The function inserts a new pair into an existing row.
-\begin{align*}
-&	\mathbf{insertValues} : Row \times Pair* \times Database \rightarrow Daatabase \\
-&	insertValues(r, p, d) = \begin{cases}
- 		%\inference{|p| > 1}{\begin{gathered} 	d \bullet insertValues(r, head(p), d) \\ \bullet insertValues(r, tail(p), d)\end{gathered}}   \\ \\
- 		\inference{|p| = 1 \wedge \forall e \in pairs(r) : reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p}
+\paragraph{Insert Value} The function inserts a new pair or pairs into an existing row.
+\begin{align*}
+&	\mathbf{insertValue} : Row \times Pair* \times Database \rightarrow Daatabase \\
+&	insertValue(r, p, d) = \begin{cases}
+ 		%\inference{|p| > 1}{\begin{gathered} 	d \bullet insertValue(r, head(p), d) \\ \bullet insertValue(r, tail(p), d)\end{gathered}}   \\ \\
+ 		\inference{|p| = 1 \wedge \forall e \in pairs(r) :\\  reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p}
  		%\\ \\ Inconsistent \; d
@@
 
-\paragraph{Add Value To Table} The function appends a pair to all rows of a table.
+\paragraph{Add Value To Table} The function appends a pair to all rows of a table with given name.
 \begin{align*}
@@
  		\inference{\exists t \in tables(d) : name(t) = l}	{\begin{gathered}
-		d \bullet insertValue(head(tableData(l, d)), p, d) \\ \bullet addValueToTable(tail(tableData(l, d)), p, d)
+		d \bullet insertValue(head(instances(l, d)), p, d) \\ \bullet addValueToTable(tail(instances(l, d)), p, d)
 	\end{gathered}} 
@@
 
-\paragraph{Copy Value} The function copy values from one column in a table to another column in the same or in another table - according to provided mapping. 
-\begin{align*}
-&	\mathbf{copyValue} : Label \times Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
+\paragraph{Copy Value} The function copy value from one column in a table to another column in the same or in another table - according to provided mapping.
+First argument of the transformation is the name of the source table, the second one is the target table, then there are names of source column and target column, finally the mapping between rows (stored instances) has to be provided.
+\begin{align*}
+&	\mathbf{copyValue} : Label \times Label \times Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
 &	copyValue(l_s, l_t, l_{sc}, l_{tc} o_{l_s, l_t}, d) = \begin{cases}
 		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc}}{\begin{gathered}
-		\forall	p \in m_{l_s, l_t} :	d \bullet insertValues(second(p), \\ [l_{tc},value(findPair(l_{sc}, first(p)))])
+		\forall	p \in m_{l_s, l_t} :	d \bullet insertValue(second(p), \\ [l_{tc},value(findPair(l_{sc}, first(p)))])
 		\end{gathered}}
@@
 
-\paragraph{Copy rows} The function $copyRows$ creates a copy of a source table content in another table with the same structure. 
-\begin{align*}
-&	\mathbf{copyRows} : Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
-&	copyRows(l_s, l_t, m_{l_s, l_t}, d) = \begin{cases}
+\paragraph{Copy rows} The function $copyRow$ creates a copy of a source table content in another table with the same structure. 
+\begin{align*}
+&	\mathbf{copyRow} : Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
+&	copyRow(l_s, l_t, m_{l_s, l_t}, d) = \begin{cases}
  	%\inference{|m_{l_s, l_t}| = \varnothing}{d} \\\\
- 	%\inference{|m_{l_s, l_t}| > 1}{\begin{gathered} d \bullet copyRows(l_s, l_t, head(M_{l_s, l_t})) \\ \bullet copyRows(l_s, l_t, tail(m_{l_s, l_t})) \end{gathered}}
+ 	%\inference{|m_{l_s, l_t}| > 1}{\begin{gathered} d \bullet copyRow(l_s, l_t, head(M_{l_s, l_t})) \\ \bullet copyRow(l_s, l_t, tail(m_{l_s, l_t})) \end{gathered}}
  	\inference{|m_{l_s, l_t}| = 1 \\ \wedge \exists t_s, t_t \in tables(d) : name(t_s) = l_s, name(t_t) = l_t \\ \wedge t_s \cong t_t }{d \bullet insertRow(row(l, pairs(first(m_{l_s, l_t}))), d)} \\\\
@@
 There are defined two variants of remove transformations for each model entity - one preserves data, the other removes data too.
-
 \begin{align*}
@@
 &	dropTableAndData(l, d) = \inference{\exists t \in tables(d) \\ \wedge ! \; associated(t, d)}{\begin{gathered}
-	(\forall r \in tableData(t,d) : dropRow(r, d))\\ \bullet dropTable(t,d)
+	(\forall r \in instances(t,d) : dropRow(r, d))\\ \bullet dropTable(t,d)
 \end{gathered}}
@@
 \end{align*}
+
 \begin{align*}
 &	\mathbf{dropColumn} : Label \times Label \times Database \rightarrow Database \\
-& 	dropColumn(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in columns(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{columns(t) \setminus c}\\
+& 	dropColumn(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in columns(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in instances(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{columns(t) \setminus c}\\
 &	 l , l_p\in Label, d \in Database 
 \end{align*}
-
-\begin{align*}
-&	\mathbf{dropKey} : Label \times Label \times Database \rightarrow Database \\
-&	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) \setminus c}\\
-&	 l, l_p \in Label, d \in Database 
-\end{align*}
 \begin{align*}
 &	\mathbf{DropColumnAndData} : Label \times Label* \times Database \rightarrow Database \\
-&	DropColumnAndData(l_c, ps, d) = \begin{cases}
+&	DropColumnAndData(l_c, ps, d) = \\
+&= \begin{cases}
  	%\inference{ps = \varnothing}{d}\\
@@
  	(\inference{|ps| = 1 \\ \wedge \exists t \in tables(d) \\ \wedge \exists c \in columns(t) : name(c) = l_c}{\begin{gathered}
-	 	(\forall r, r \in tableData(d) \wedge reference(r) = l_c: \\d \bullet dropValue(ps, r)) \end{gathered}}) \\ \bullet dropColumn(l_c)  
+	 	(\forall r, r \in instances(d) \wedge reference(r) = l_c: \\d \bullet dropValue(ps, r)) \end{gathered}}) \\ \bullet dropColumn(l_c)  
  \end{cases}
@@
 \begin{align*}
+&	\mathbf{dropKey} : Label \times Label \times Database \rightarrow Database \\
+&	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in instances(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) \setminus c}\\
+&	 l, l_p \in Label, d \in Database 
+\end{align*}
+\begin{align*}
 &	\mathbf{DropKeyAndData} : Label \times Label* \times Database \rightarrow Database \\
-&	DropKeyAndData(l_f, ps, d) = \begin{cases}
+&	DropKeyAndData(l_f, ps, d) = \\
+& = \begin{cases}
  	%\inference{ps = \varnothing}{d}\\
@@
  	(\inference{|ps| = 1 \\ \wedge \exists t \in tables(d) \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_f}{\begin{gathered}
-	 	(\forall r, r \in tableData(d) \wedge reference(r) = l_f: \\d \bullet dropValue(ps, r)) \end{gathered}}) \\ \bullet dropKey(l_f)  
+	 	(\forall r, r \in instances(t,d) \wedge reference(r) = l_f: \\d \bullet dropValue(ps, r)) \end{gathered}}) \\ \bullet dropKey(l_f)  
  \end{cases}
@@
 
-The definition of a transformation consists of two parts - a conditions of transformation feasibility and its impact on the context (application or database).
-If the conditions of feasibility are not met, the transformation cannot change the context - the software cannot evolve.
-The second branch of the operation definition is an implicit assumption - if an operation is not feasible in a software, its state become inconsistent.
-This implicit assumption is not stated in the transformation definitions to increase theirs readability.
+It is important to notice that each instance of the software has different data in its database.
+It means that the evolution can be feasible only for some instances of the software and thus proposed transformation has to be always interpreted in context of the whole software.
 
@@
 &	\Psi(s, addProperty(l, p)) = application(s) \circ addProp(l,p) \\
-&	\Phi(s, addProperty(l, p)) = \begin{cases}
+&	\Phi(s, addProperty(l, p)) = \\
+& = \begin{cases}
     		\inference{cardinality(p) \leq 1}{database(s) \bullet addColumn(l, \rho(p,d), d)} \\ \\
@@
 & \Psi(s, addAsociation(l, a, m_{l_1, l_2})) = application(s) \circ addAssoc(l, a) \\
-& \Phi(s, addAsociation(l, a, m_{l_1, l_2})) = \begin{cases}
+& \Phi(s, addAsociation(l, a, m_{l_1, l_2})) = \\
+& =\begin{cases}
 	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l\\ 		\wedge name(t_t) = reference(a) \\ \wedge \forall f \in foreignKeys(t_t) \cup columns(t_t) \\: name(f) \neq name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
@@
 &	\Psi RemoveProperty(l, l_p, s) =  applications(s) \circ  remProp(l, l_p) \\
-&	\Phi RemoveProperty(l, l_p, s) = \begin{cases}
+&	\Phi RemoveProperty(l, l_p, s) = \\
+& \begin{cases}
 		 \inference{\exists c \in classes(application(s)) : name(c) = l\\ 
@@
 &  \circ remProperty(l_s, ps) \\
-&	\Psi(RemovePropertyAndData(L_c, ps), s) = \begin{cases}
+&	\Psi(RemovePropertyAndData(L_c, ps), s) = \\
+& = \begin{cases}
 		%\inference{ps = \varnothing}{d}\\
@@
  		\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) \leq 1}{database(s) \bullet DropColumnAndData(l_c, ps)}\\\\
- 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) > 1}{database(s) \bullet RemoveTableAndData(l_c, ps)}
+ 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) > 1}{database(s) \bullet RemoveTableAndData(name(owningClass(p))+l_c, ps)}
  \end{cases}
@@
 &    \Psi RemoveAssociation(l, l_a, s) = application(s) \circ remAss(l, l_a) \\
-&    \Phi RemoveAssociation(l, l_a, s) = \begin{cases}
+&    \Phi RemoveAssociation(l, l_a, s) = \\
+&= s\begin{cases}
      	\inference{\exists c \in classes(application(s)) : name(c) = l \\
@@
 & \circ remAssociation(l_s, ps) \\
-&	\Psi(RemoveAssociationAndData(L_c, ps), s) = \begin{cases}
+&	\Psi(RemoveAssociationAndData(L_c, ps), s) = \\
+&=  \begin{cases}
 		%\inference{ps = \varnothing}{d}\\
@@
  \end{cases} \\
-&	l_c, l_o, l_n \in Label, a \in Applications, s : Sate
+&	l_c, l_o, l_n \in Label, a \in Application, s : Sate
 \end{align*}
@@
 
-The $CopyProperty$ function can be implemented as a transformation composed by operations for adding and removing property, nevertheless this composition would not work with instances, therefore transformation to update data are used.
+There are two possible implementation of transformation for copying property, first implementation $CopyPropertyStructure$ copies only the structure of a property on the database level, the second implementation $CopyProperty $ copies both structure and data of the property.
+\begin{align*}
+&	\mathbf{CopyPropertyStructure} : Label \times Label \times Label \times State \rightarrow State \\
+&	\Psi(CopyPropertyStructure(l_s, l_t, l_p, s) = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)} 
+	\\\\
+&	\Phi(CopyPropertyStructure(l_s, l_t, l_p, s) = \begin{cases}
+    \inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) \leq 1 }{\begin{gathered}
+			database(s) \bullet copyColumn(l_s, l_t, l_p) 			\end{gathered}} \\ \\
+		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) > 1}{\begin{gathered}
+			database(s) \bullet copyTable(l_s + l_p, l_t + l_p) 			\end{gathered}}
+ \end{cases}\\
+& l_s, l_t, l_p \in Label, s \in State
+\end{align*}
+
 \begin{align*}
@@
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) \leq 1 }{\begin{gathered}
-			database(s) \bullet copyColumn(l_s, l_t, l_p) \\ \bullet copyValues(l_s, l_t, l_p, l_p, m_{l_s, l_t}) 
+			database(s) \bullet copyColumn(l_s, l_t, l_p) \\ \bullet copyValue(l_s, l_t, l_p, l_p, m_{l_s, l_t}) 
 			\end{gathered}} \\ \\
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) > 1}{\begin{gathered}
-			database(s) \bullet copyTable(l_s + l_p, l_t + l_p) \\ \bullet copyRows(l_s + l_p, l_t + l_p, m_{l_s, l_t}) 
+			database(s) \bullet copyTable(l_s + l_p, l_t + l_p) \\ \bullet copyRow(l_s + l_p, l_t + l_p, m_{l_s, l_t}) 
 			\end{gathered}}
@@
 \section{Advanced Refactoring Cases}
-The concrete refactoring cases presented in this section were chosen according to the refactoring statistics provided by Eclipse.
-The refactoring cases, which does not affect data (such as change method signature or encapsulate field) were removed from the set of refactoring cases and form the rest the Move property and Extract/Inline class were chosen to
-\subsection{Move Property}
+The refactoring cases presented in this section were chosen according to the refactoring statistics provided by Eclipse, concretely Move property and Extract/Inline class were chosen to prove the capabilities of defined transformations and next complex refactoring case - merge of two classes - is presented.
+\paragraph{Move Property}
 The aim of the move property transformation is to change two classes - one will get a new property, the other will lost a property.
 An example of such transformation is merging information from one class to another.
-The implementation of move property transformation is based on copy property transformation:
+The implementation of move property transformation is based on copy property transformation.
+Its inputs are the name of source class, the name of target class and the sequence of names of properties to be moved from source to target, finally the mapping between instances of source and target class.
 \begin{align*}
 &	\mathbf{MoveProperty} : Label \times Label \times Label* \times (Row \times Row) \times State \rightarrow State \\
-&	s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) = \begin{cases}
- 	%\inference{|ps| = \varnothing}{state(s)}\\\\
- 		%\inference{|ps| > 1}{\begin{gathered} 	s \oplus MoveProperty(l_s, l_t, head(ps), m_{l_s, l_t}) \\ \oplus MoveProperty(l_s, l_t, tail(ps), m_{l_s, l_t}) \end{gathered}}
- 	\inference{|ps| = 1}{\begin{gathered} s \oplus CopyProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemovePropertyAndData(l_s, ps)
-\end{gathered}}\\\\
- \end{cases}
+&	s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) =
+% \begin{cases}
+%	\inference{|ps| = \varnothing}{state(s)}\\\\
+%	\inference{|ps| > 1}{\begin{gathered} 	s \oplus MoveProperty(l_s, l_t, head(ps), m_{l_s, l_t}) \\ \oplus MoveProperty(l_s, l_t, tail(ps), m_{l_s, l_t}) \end{gathered}}
+ 	\inference{|ps| = 1 }{\begin{gathered} s \oplus CopyProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemovePropertyAndData(l_s, ps)
+\end{gathered}}
+% \end{cases}
 	\\
@@
 
-\subsection{Extract and Inline Class}
+\paragraph{Extract and Inline Class}
 The extract and inline class are two opposite transformations. First of them extract a class from existing class, second of them merges two classes into one.
 \begin{align*}
-&	\mathbf{Extract} : Label \times Label* \times Label \times State \rightarrow State \\
-&	state(s) \oplus Extract(l_s, ps, l_n) = \begin{cases}
+&	\mathbf{Extract} : Label \times Label \times Label \times (Row \times Row) \times State \rightarrow State \\
+&	s \oplus Extract(l_s, l_n, l_p, m_{l_s, l_s}) = \begin{cases}
 	\inference{|ps| = 1}{\begin{gathered}
-		s \oplus addClass(cl(l_n, \varnothing, \varnothing)) \\ \oplus MoveProperty(l_s, l_n, ps, \mathbf{X}) \end{gathered}}\\\\
+		s \oplus addClass(cl(l_n, \varnothing, \varnothing)) \\ \oplus MoveProperty(l_s, l_n, ps, m_{l_s, l_s}) \end{gathered}}\\\\
 	%\inference{|ps| > 1}{\begin{gathered} 	s \oplus Extract(l_s, head(ps), l_n) \\\oplus Extract(l_s, tail(ps), l_n) \end{gathered}}\\\\
@@
 	\\
-&	l_s, l_n \in Label, ps : Label* s \in State 
-\end{align*}
-
-\begin{align*}
-&	\mathbf{Inline} : Label \times Label \times State \rightarrow State \\
-&	state(s) \oplus Inline(l_s, l_t, m_{l_s, l_t}) = \inference{\exists c \in classes(application(s)) : name(c) = l_s \\\wedge !\:associated(l_s, s) \\ \wedge foreignKeys(c) = \varnothing  }{\begin{gathered}
+&	l_s, l_n, l_p \in Label,  s \in State 
+\end{align*}
+The opposite transformation $Inline$ gets the name of a class to inline and the name of target class.
+The result of the transformation is a merge of information from two classes in the target class, the source class is removed.
+The important condition of feasibility is there is no name collision among properties in both classes.
+\begin{align*}
+&	\mathbf{Inline} : Label \times Label \times (Row \times Row) \times State \rightarrow State \\
+&	s \oplus Inline(l_s, l_t, m_{l_s, l_t}) = \inference{\exists c_s \in classes(application(s)) : name(c_s) = l_s \\
+\wedge \exists c_t \in classes(application(s)) : name(c_t) = l_t \\
+\wedge \forall p_s \in properties(c_s), p_t \in properties(c_t) : name(ps) \neq name(p_t) \\
+\wedge !\:associated(l_s, s)   }{\begin{gathered}
 	s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemoveClass(l_s) 
@@
 & l_s, l_t \in Label, m_{l_s, l_t} \in M, s \in State, \\ 
-& ps = \{x | x \in \forall c \in classes(applications)) : \forall p \in properries(c) : name(p))\}
-\end{align*}
+& ps = \{x | x \in \forall c \in classes(application(s)) : \forall p \in properries(c) : name(p))\}
+\end{align*}
+
+\paragraph{Merge of Classes} There occur situations when there is a need to merge two classes and the inlining transformation cannot be used because of name collisions among properties in such a case the $merge$ transformation can be used.
+It can be used in case structure of both classes (respectively tables) corresponds, the transformation does not change the structure, it only moves data.
+\begin{align*}
+&	\mathbf{merge} : Label \times Label \times Label \times State \rightarrow State \\
+&	\Phi(s, merge(l_1, l_2, l_n) = \inference{equals(c_1, c_2)}{application(s) \circ copyClassStructure(l_1, l_n)} \\\\
+&	\Psi(s, merge(l_1, l_2, l_n) = \inference{\exists t_1, t_2 \in tables(database(s)) : \\ name(t_1) = l_1 \wedge name(t_2) = l_2 \\
+\wedge ! \: referenced(t_1, d) \wedge ! \: referenced(t_2, d) }{database(s) \bullet mergeTables(l_1, l_2)}\\
+& l_1, l_2, l_n \in Label, s \in State
+\end{align*}
+
+\section{Case Study}
+The case study demonstrates how the transformation helps with data evolution in real example of software evolution.
+There are two classes in our software $Person$ and $LegalPerson$ both of them contains information about address (street, city and zip code).
+To improve the design of code the class Address has to be created which associated with both original classes and contains the data already stored in the software.
+The initial situation:
+\begin{align*}
+&	S = sw(a, db(t,r))
+\end{align*}
+\begin{align*}
+a &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing))
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
+& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
+& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
+& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
+& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing)) 
+\end{align*}
+\begin{align*}
+r &= row("person", ["Id", 1], ["surname", "Griffin"], \\
+& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("person", ["Id", 2], ["surname", "Doe"], \\
+& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
+& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003]))
+\end{align*}
+The first evolution step is to create the address class for $Person$ and $LegalPerson$, then the temporary class can be merged together and finally the temporary class has to be renamed to $Address$:
+\begin{align*}
+& state(S) \oplus Extract("Person", "Address_tmp1", "street", "city", "zip") \\
+& \oplus Extract("LegalPerson", "Address_tmp2", "street", "city", "zip")  \\
+& \oplus Merge("Address_tmp1", "Address_tmp2") \\
+& \oplus Rename("Address_tmp2", "Address")
+\end{align*}
+The software state after second extraction is:
 
@@
 
+\newpage
+ExtactClass = addClass
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)),
+& cl("Address_tmp1", \varnothing, \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
+& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
+& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
+& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
+& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing),
+& tab("Address_tmp1", pk("Id"), \varnothing, \varnothing)) 
+\end{align*}
+\begin{align*}
+r &= row("person", ["Id", 1], ["surname", "Griffin"], \\
+& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("person", ["Id", 2], ["surname", "Doe"], \\
+& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
+& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003]))
+\end{align*}
+
+ExtactClass = copyProperty("street"
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)),
+& cl("Address_tmp1", & pr(street, APPSTRING, \varnothing) , \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
+& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
+& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
+& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
+& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing),
+& tab("Address_tmp1", pk("Id"), col(street, APPSTRING, \varnothing, \varnothing),\\
+& \varnothing)) 
+\end{align*}
+\begin{align*}
+r &= row("person", ["Id", 1], ["surname", "Griffin"], \\
+& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("person", ["Id", 2], ["surname", "Doe"], \\
+& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
+& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"]) \\
+\end{align*}
+
 \end{document}

### 2013-01-26 doc/description.tex
merge classes

@@
 \paragraph{Merge of Classes} There occur situations when there is a need to merge two classes and the inlining transformation cannot be used because of name collisions among properties in such a case the $merge$ transformation can be used.
-It can be used in case structure of both classes (respectively tables) corresponds, the transformation does not change the structure, it only moves data.
+It can be used in case structure of both classes (respectively tables) is equals.
+The transformation creates new rows in the new table - the primary keys from origin instances are replaced by a new ones.
 \begin{align*}
 &	\mathbf{merge} : Label \times Label \times Label \times State \rightarrow State \\
-&	\Phi(s, merge(l_1, l_2, l_n) = \inference{equals(c_1, c_2)}{application(s) \circ copyClassStructure(l_1, l_n)} \\\\
+&	\Phi(s, merge(l_1, l_2, l_n) = \inference{
+		\exists c_1, c_2 \in classes(application(s)) \wedge 
+		equals(c_1, c_2)}{\begin{gathered}
+			application(s) \circ addClass(cl(ln, \varnothing, \varnothing)\\ \circ (\forall p \in properties(c_1) : addProperty(l_n, p)) \\ \circ (\forall a \in associations(c_1): addAssociation(l_n, a))
+		\end{gathered}} \\\\
 &	\Psi(s, merge(l_1, l_2, l_n) = \inference{\exists t_1, t_2 \in tables(database(s)) : \\ name(t_1) = l_1 \wedge name(t_2) = l_2 \\
-\wedge ! \: referenced(t_1, d) \wedge ! \: referenced(t_2, d) }{database(s) \bullet mergeTables(l_1, l_2)}\\
+\wedge ! \: referenced(t_1, d) \wedge ! \: referenced(t_2, d) }{\begin{gathered}
+	 database(s) \bullet addTable(l_n, \mathbf{X}, \varnothing, \varnothing) \\ \bullet (\forall c \in columns(t_1) : addColumn(l_n, c)) \\ \bullet (\forall f \in foreignKeys(t_1) : addForeignKey(l_n, f)) \\ \bullet (\forall r \in instances(t_1, database(s)) \cup instances(t_2, database(s)) :  \\ insertRow(l_n, row(l_n, next(seq(key(l_n))), pairs(r))))
+\end{gathered}}\\
 & l_1, l_2, l_n \in Label, s \in State
@@
 & state(S) \oplus Extract("Person", "Address_tmp1", "street", "city", "zip") \\
+& \oplus AddAssociation("Person", assoc("address", "Address_tmp1", 1, 1),\\ 
+& \{(x,y)| x \in instances("Person", database(S)), \\
+& y \in instances("Address_tmp1", database(S)) : pkVal(x) = pkVal(y) \}) \\
 & \oplus Extract("LegalPerson", "Address_tmp2", "street", "city", "zip")  \\
-& \oplus Merge("Address_tmp1", "Address_tmp2") \\
-& \oplus Rename("Address_tmp2", "Address")
-\end{align*}
-The software state after second extraction is:
+& \oplus AddAssociation("LegalPerson", assoc("address", "Address_tmp1", 1, 1),\\ 
+& \{(x,y)| x \in instances("LegalPerson", database(S)), \\
+& y \in instances("Address_tmp2", database(S)) : pkVal(x) = pkVal(y) \}) \\
+& \oplus Merge("Address_tmp1", "Address_tmp2", "Address") \\
+& \oplus RemoveTableAndData("Address_tmp1") \\
+& \oplus RemoveTableAndData("Address_tmp2")
+\end{align*}
+The software state after adding the second association is:
 
@@
 & col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
-& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing),
+& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
 & tab("Address_tmp1", pk("Id"), \varnothing, \varnothing)) 
@@
 & pr(street, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)),
-& cl("Address_tmp1", & pr(street, APPSTRING, \varnothing) , \varnothing)
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl("Address_tmp1", pr(street, APPSTRING, \varnothing) , \varnothing)
 & )
@@
 & col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
-& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing),
+& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
 & tab("Address_tmp1", pk("Id"), col(street, APPSTRING, \varnothing, \varnothing),\\

### 2013-01-26 doc/description.tex
case study, updated insertValue

@@
 \begin{align*}
-&	\mathbf{insertValue} : Row \times Pair* \times Database \rightarrow Daatabase \\
-&	insertValue(r, p, d) = \begin{cases}
+&	\mathbf{insertValue} : Label \times N \times Pair* \times Database \rightarrow Daatabase \\
+&	insertValue(l, n, p, d) = \begin{cases}
  		%\inference{|p| > 1}{\begin{gathered} 	d \bullet insertValue(r, head(p), d) \\ \bullet insertValue(r, tail(p), d)\end{gathered}}   \\ \\
- 		\inference{|p| = 1 \wedge \forall e \in pairs(r) :\\  reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p}
+ 		\inference{|p| = 1 \wedge 
+ 		\\ \wedge \exist r \in instances(l, d) : value(findPair("Id", r)) = n
+ 		\\ \wedge \exists t \in tables(d) : name(t) = l, 
+ 		\\ \wedge \exists c \in columns(t) : reference(p) = name(c) 
+ 		\\ \wedge \forall e \in pairs(r) :\\  reference(e) \neq reference(p) }{pairs(r) \cup p} 
  		%\\ \\ Inconsistent \; d
@@
  		\inference{\exists t \in tables(d) : name(t) = l}	{\begin{gathered}
-		d \bullet insertValue(head(instances(l, d)), p, d) \\ \bullet addValueToTable(tail(instances(l, d)), p, d)
+		d \bullet (\forall r \in instances(l, d) \\ : insetValue(reference(r), keyValue(r), p) \\ \bullet addValueToTable(tail(instances(l, d)), p, d))
 	\end{gathered}} 
@@
 		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc}}{\begin{gathered}
-		\forall	p \in m_{l_s, l_t} :	d \bullet insertValue(second(p), \\ [l_{tc},value(findPair(l_{sc}, first(p)))])
+		\forall	p \in m_{l_s, l_t} :	d \bullet insertValue(name(second(p)), keyValue(second(p)), \\ [l_{tc},value(findPair(l_{sc}, first(p)))])
 		\end{gathered}}
@@
 &	\mathbf{Extract} : Label \times Label \times Label \times (Row \times Row) \times State \rightarrow State \\
-&	s \oplus Extract(l_s, l_n, l_p, m_{l_s, l_s}) = \begin{cases}
+&	s \oplus Extract(l_s, l_n, ps, m_{l_s, l_s}) = \begin{cases}
 	\inference{|ps| = 1}{\begin{gathered}
@@
 	\\
-&	l_s, l_n, l_p \in Label,  s \in State 
+&	l_s, l_n \in Label, ps ={ x | x  \in Label},  s \in State 
 \end{align*}
@@
 
+\newpage
 ExtactClass = copyProperty("street"
@@
 \begin{align*}
-r &= row("person", ["Id", 1], ["surname", "Griffin"], \\
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
 & ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
@@
 & row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"]) \\
+& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"] )\\
+\}
+\end{align*}
+
+\newpage
+ExtactClass = dropPropertyAndData("street"
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl("Address_tmp1", pr(street, APPSTRING, \varnothing) , \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
+& col(city, APPSTRING, \varnothing, \varnothing) \\
+& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
+& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
+& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("Address_tmp1", pk("Id"), col(street, APPSTRING, \varnothing, \varnothing),\\
+& \varnothing)) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+& ["city", "NY"], ["zip", 10001]), \\
+& row("person", ["Id", 2], ["surname", "Doe"], \\
+& ["city", "LA"], ["zip", 90003]), \\
+& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"]) \\
+& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"] )\\
+\}
+\end{align*}
+
+\newpage
+ExtactClass = copyProperty("city"
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl("Address_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing) , \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
+& col(city, DBSTRING, \varnothing, \varnothing) \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("Address_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing)
+& \varnothing)) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+& ["city", "NY"], ["zip", 10001]), \\
+& row("person", ["Id", 2], ["surname", "Doe"], \\
+&  ["city", "LA"], ["zip", 90003]), \\
+& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"]) \\
+& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"] )\\
+\}
+\end{align*}
+
+\newpage
+ExtactClass = dropPropertyAndData("city"
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\
+& pr(zip, APPSTRING, \varnothing)), \\
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl("Address_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing) , \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("Address_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing)
+& \varnothing)) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+& ["zip", 10001]), \\
+& row("person", ["Id", 2], ["surname", "Doe"], \\
+& ["zip", 90003]), \\
+& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"]) \\
+& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"] )\\
+\}
+\end{align*}
+
+\newpage
+ExtactClass = CopyProperty("zip"
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\
+& pr(zip, APPSTRING, \varnothing)), \\
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl("Address_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("Address_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)
+& \varnothing)) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+& ["zip", 10001]), \\
+& row("person", ["Id", 2], ["surname", "Doe"], \\
+& ["zip", 90003]), \\
+& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]) \\
+& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]) )\\
+\}
+\end{align*}
+
+\newpage
+ExtactClass = DropPopertyAndData("zip"
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing))), \\
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl("Address_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("Address_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)
+& \varnothing)) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
+& row("person", ["Id", 2], ["surname", "Doe"]), \\
+& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]) \\
+& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]) )\\
+\}
+\end{align*}
+
+\newpage
+$AddAssociation("Person", assoc("address", "Address_tmp1", 1, 1), 
+ \{(x,y)| x \in instances("Person", database(S)), y \in instances("Address_tmp1", database(S)) : pkVal(x) = pkVal(y) \})$
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
+& assoc("address", "Address_tmp1", 1, 1)), \\ 
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl("Address_tmp1", pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("Address_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)
+& fk("address", "Person", \varnothing))\\
+) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
+& row("person", ["Id", 2], ["surname", "Doe"]), \\
+& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]) \\
+& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]) )\\
+\}
 \end{align*}

### 2013-01-26 doc/description.tex
updated add association & case study

@@
 	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l\\ 		\wedge name(t_t) = reference(a) \\ \wedge \forall f \in foreignKeys(t_t) \cup columns(t_t) \\: name(f) \neq name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
-		database(s) \\ \bullet addForeignKey(name(t_s), \rho(a,d), d) \\ \bullet \forall r \in m_{l_1, l_2} : \\ insertRow(row(name(t_s), [name(a), \\ value(primaryKey(first(r)))]), d)
+		database(s) \\ \bullet addForeignKey(name(t_t), \rho(a,d), d) \\ \bullet \forall r \in m_{l_1, l_2} : \\
+		insertValue(name(t_t), value(primaryKey(second(r))), [name(a), value(primaryKey(first(r)))])
 	\end{gathered}}
@@
 	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l \\ \wedge name(t_t) = reference(a) \\ \wedge startCardinality(a) > 1}{\begin{gathered} 
-	database(s) \bullet addTable(\rho(a,d),d) \\ \bullet \forall r \in m_{l, reference(a)} :\\ insertRow(\\ row(name(owningClass(a)) + name(a), \\ [name(t_s), value(primaryKey(first(r)))]\\ [name(t_t), value(primaryKey(second(r)))]))
+	database(s) \bullet addTable(\rho(a,d),d) \\ \bullet \forall r \in m_{l, reference(a)} :
+	\\ insertRow(\\ row(name(owningClass(a)) + name(a), \\ [name(t_s), value(primaryKey(first(r)))]\\ [name(t_t), value(primaryKey(second(r)))]))
 	\end{gathered}} 
@@
 & ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
-& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
 & ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
 & ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003]))
@@
 & ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
-& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
 & ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
 & ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
@@
 & ["city", "LA"], ["zip", 90003]), \\
-& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
 & ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
 & ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
@@
 &  ["city", "LA"], ["zip", 90003]), \\
-& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
 & ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
 & ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
@@
 & ["zip", 90003]), \\
-& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
 & ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
 & ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
@@
 & row("person", ["Id", 2], ["surname", "Doe"]), \\
-& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
 & ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
 & ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
@@
 & tab("Address_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
-& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
 & fk("address", "Person", \varnothing))\\
@@
 & row("person", ["Id", 2], ["surname", "Doe"]), \\
-& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
 & ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
 & ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
-& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]) \\
-& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]) )\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001], ["Person", 1]) \\
+& row("address\_tmp1", ["Id", 2], ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003], ["Person", 2]) )\\
 \}
@@
 
+\newpage
+\begin{align*}
+& Extract("LegalPerson", "Address_tmp2", "street", "city", "zip")  \\
+& \oplus AddAssociation("LegalPerson", assoc("address", "Address_tmp1", 1, 1),\\ 
+& \{(x,y)| x \in instances("LegalPerson", database(S)), \\
+& y \in instances("Address_tmp2", database(S)) : pkVal(x) = pkVal(y) \}) \\
+\end{align*}
+\hline
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
+& assoc("address", "Address_tmp1", 1, 1)), \\ 
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing),\\
+& assoc("address", "Address_tmp2", 1, 1)), \\
+& cl("Address_tmp1", pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing) \\
+& cl("Address_tmp2", pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("Address_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
+& fk("address", "Person", \varnothing))\\
+& tab("Address_tmp2", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
+& fk("address", "LegalPerson", \varnothing))\\
+) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
+& row("person", ["Id", 2], ["surname", "Doe"]), \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."]), \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], \\ 
+& ["city", "NY"], ["zip", 10001], ["Person", 1]) \\
+& row("address\_tmp1", ["Id", 2], ["street", "Olive ave. 2225"], \\
+& ["city", "LA"], ["zip", 90003], ["Person", 2]) )\\
+& row("address\_tmp2", ["Id", 10], ["street", "Hill stret 22"], \\
+& ["city", "NY"], ["zip", 10001], ["LegalPerson", 10]) \\
+& row("address\_tmp2", ["Id", 20], ["street", "Pine ave. 2225"], \\
+& ["city", "LA"], ["zip", 90003], ["LegalPerson", 20]) )\\
+\}
+\end{align*}
+\newpage
+
 \end{document}

### 2013-01-26 doc/description.tex
updated casestudy

--- 2013-01-26 doc/description.tex
@@
 \begin{align*}
-& state(S) \oplus Extract("Person", "Address_tmp1", "street", "city", "zip") \\
-& \oplus AddAssociation("Person", assoc("address", "Address_tmp1", 1, 1),\\ 
+& state(S) \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip") \\
+& \oplus AddAssociation("Person", assoc("address", "Address\_tmp1", 1, 1),\\ 
 & \{(x,y)| x \in instances("Person", database(S)), \\
-& y \in instances("Address_tmp1", database(S)) : pkVal(x) = pkVal(y) \}) \\
-& \oplus Extract("LegalPerson", "Address_tmp2", "street", "city", "zip")  \\
-& \oplus AddAssociation("LegalPerson", assoc("address", "Address_tmp1", 1, 1),\\ 
+& y \in instances("Address\_tmp1", database(S)) : pkVal(x) = pkVal(y) \}) \\
+& \oplus Extract("LegalPerson", "Address\_tmp2", "street", "city", "zip")  \\
+& \oplus AddAssociation("LegalPerson", assoc("residentialAddress", "Address\_tmp2", 1, 1),\\ 
 & \{(x,y)| x \in instances("LegalPerson", database(S)), \\
-& y \in instances("Address_tmp2", database(S)) : pkVal(x) = pkVal(y) \}) \\
-& \oplus Merge("Address_tmp1", "Address_tmp2", "Address") \\
-& \oplus RemoveTableAndData("Address_tmp1") \\
-& \oplus RemoveTableAndData("Address_tmp2")
+& y \in instances("Address\_tmp2", database(S)) : pkVal(x) = pkVal(y) \}) \\
+%
+& \oplus AddAssociation("Person", \\
+& assoc("address", "Address\_tmp2", 1, 1), \varnothing) \\
+& \oplus AddAssociation("LegalPerson", \\
+& assoc("residentialAddress", "Address\_tmp1", 1, 1),\varnothing) \\
+%
+& \oplus Merge("Address\_tmp1", "Address\_tmp2", "Address") \\
+& \oplus RemoveTableAndData("Address\_tmp1") \\
+& \oplus RemoveTableAndData("Address\_tmp2")
 \end{align*}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%
 \newpage
+\section{Case Study}
 ExtactClass = addClass
+\hline
 \begin{align*}
@@
 & pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)),
-& cl("Address_tmp1", \varnothing, \varnothing)
+& cl("Address\_tmp1", \varnothing, \varnothing)
 & )
@@
 & col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab("Address_tmp1", pk("Id"), \varnothing, \varnothing)) 
+& tab("Address\_tmp1", pk("Id"), \varnothing, \varnothing)) 
 \end{align*}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
 ExtactClass = copyProperty("street"
+\hline
 \begin{align*}
@@
 & pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
-& cl("Address_tmp1", pr(street, APPSTRING, \varnothing) , \varnothing)
+& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) , \varnothing)
 & )
@@
 & col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab("Address_tmp1", pk("Id"), col(street, APPSTRING, \varnothing, \varnothing),\\
+& tab("Address\_tmp1", pk("Id"), col(street, APPSTRING, \varnothing, \varnothing),\\
 & \varnothing)) 
@@
 & ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
-& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
 & ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
 & ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
 & row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"]) \\
+& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"] )\\
+\}
+\end{align*}
+
+%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+ExtactClass = dropPropertyAndData("street"
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) , \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
+& col(city, APPSTRING, \varnothing, \varnothing) \\
+& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
+& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
+& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("Address\_tmp1", pk("Id"), col(street, APPSTRING, \varnothing, \varnothing),\\
+& \varnothing)) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+& ["city", "NY"], ["zip", 10001]), \\
+& row("person", ["Id", 2], ["surname", "Doe"], \\
+& ["city", "LA"], ["zip", 90003]), \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"]) \\
+& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"] )\\
+\}
+\end{align*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+ExtactClass = copyProperty("city"
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing) , \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
+& col(city, DBSTRING, \varnothing, \varnothing) \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing)
+& \varnothing)) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+& ["city", "NY"], ["zip", 10001]), \\
+& row("person", ["Id", 2], ["surname", "Doe"], \\
+&  ["city", "LA"], ["zip", 90003]), \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"]) \\
+& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"] )\\
+\}
+\end{align*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+ExtactClass = dropPropertyAndData("city"
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\
+& pr(zip, APPSTRING, \varnothing)), \\
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing) , \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing)
+& \varnothing)) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+& ["zip", 10001]), \\
+& row("person", ["Id", 2], ["surname", "Doe"], \\
+& ["zip", 90003]), \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"]) \\
+& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"] )\\
+\}
+\end{align*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+ExtactClass = CopyProperty("zip"
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\
+& pr(zip, APPSTRING, \varnothing)), \\
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)
+& \varnothing)) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+& ["zip", 10001]), \\
+& row("person", ["Id", 2], ["surname", "Doe"], \\
+& ["zip", 90003]), \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]) \\
+& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]) )\\
+\}
+\end{align*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+ExtactClass = DropPopertyAndData("zip"
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing))), \\
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)
+& \varnothing)) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
+& row("person", ["Id", 2], ["surname", "Doe"]), \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]) \\
+& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]) )\\
+\}
+\end{align*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+$AddAssociation("Person", assoc("address", "Address\_tmp1", 1, 1), 
+ \{(x,y)| x \in instances("Person", database(S)), y \in instances("Address\_tmp1", database(S)) : pkVal(x) = pkVal(y) \})$
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
+& assoc("address", "Address\_tmp1", 1, 1)), \\ 
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
+& fk("person\_address", "Person", \varnothing))\\
+) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
+& row("person", ["Id", 2], ["surname", "Doe"]), \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], \\
+& ["city", "NY"], ["zip", 10001], ["person\_address", 1]) \\
+& row("address\_tmp1", ["Id", 2], ["street", "Olive ave. 2225"], \\
+& ["city", "LA"], ["zip", 90003], ["person\_address", 2]) )\\
+\}
+\end{align*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\begin{align*}
+& Extract("LegalPerson", "Address\_tmp2", "street", "city", "zip")  \\
+& \oplus AddAssociation("LegalPerson", assoc("address", "Address\_tmp1", 1, 1),\\ 
+& \{(x,y)| x \in instances("LegalPerson", database(S)), \\
+& y \in instances("Address\_tmp2", database(S)) : pkVal(x) = pkVal(y) \}) \\
+\end{align*}
+\hline
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
+& assoc("address", "Address\_tmp1", 1, 1)), \\ 
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing),\\
+& assoc("address", "Address\_tmp2", 1, 1)), \\
+& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing) \\
+& cl("Address\_tmp2", pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
+& fk("person\_address", "Person", \varnothing))\\
+& tab("Address\_tmp2", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
+& fk("legalPerson\_address", "LegalPerson", \varnothing))\\
+) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
+& row("person", ["Id", 2], ["surname", "Doe"]), \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."]), \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], \\ 
+& ["city", "NY"], ["zip", 10001], ["Person", 1]) \\
+& row("address\_tmp1", ["Id", 2], ["street", "Olive ave. 2225"], \\
+& ["city", "LA"], ["zip", 90003], ["Person", 2]) )\\
+& row("address\_tmp2", ["Id", 10], ["street", "Hill stret 22"], \\
+& ["city", "NY"], ["zip", 10001], ["LegalPerson\_address", 10]) \\
+& row("address\_tmp2", ["Id", 20], ["street", "Pine ave. 2225"], \\
+& ["city", "LA"], ["zip", 90003], ["LegalPerson\_address", 20]) )\\
+\}
+\end{align*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\begin{align*}
+& \oplus AddAssociation("Person", \\
+& assoc("address", "Address\_tmp2", 1, 1), \varnothing) \\
+\end{align*}
+\hline
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
+& assoc("address", "Address\_tmp1", 1, 1), assoc("address", "Address\_tmp2", 1, 1)), \\ 
+& cl(LegalPerson, pr(businessName, APPSTRING, \varnothing),\\
+& assoc("address", "Address\_tmp2", 1, 1), assoc(\mathbf{"address"}, "Address\_tmp1", 1, 1)), \\
+& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing) \\
+& cl("Address\_tmp2", pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing)) \\
+& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
+& fk("address", "Person", \varnothing), fk("address", "LegalPerson", \varnothing))\\
+& tab("Address\_tmp2", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
+& fk("LegalPerson\_address", "LegalPerson", \varnothing), fk("Person\_address", "Person", \varnothing))\\
+) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
+& row("person", ["Id", 2], ["surname", "Doe"]), \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."]), \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], \\ 
+& ["city", "NY"], ["zip", 10001], ["Person", 1]) \\
+& row("address\_tmp1", ["Id", 2], ["street", "Olive ave. 2225"], \\
+& ["city", "LA"], ["zip", 90003], ["Person", 2]) )\\
+& row("address\_tmp2", ["Id", 10], ["street", "Hill stret 22"], \\
+& ["city", "NY"], ["zip", 10001], ["LegalPerson", 10]) \\
+& row("address\_tmp2", ["Id", 20], ["street", "Pine ave. 2225"], \\
+& ["city", "LA"], ["zip", 90003], ["LegalPerson", 20]) )\\
+\}
+\end{align*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\begin{align*}
+& \oplus AddAssociation("LegalPerson", \\
+& assoc("address", "Address\_tmp1", 1, 1),\varnothing) \\
+\end{align*}
+\hline
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
+& assoc("address", "Address\_tmp1", 1, 1)), \\ 
+& cl(LegalPerson, pr(businessName, APPSTRING, \varnothing),\\
+& assoc("address", "Address\_tmp2", 1, 1), assoc(\mathbf{"address"}, "Address\_tmp1", 1, 1)), \\
+& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing) \\
+& cl("Address\_tmp2", pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
+& )
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing)) \\
+& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
+& fk("Person\_address", "Person", \varnothing), fk("LegalPerson\_address", "LegalPerson", \varnothing))\\
+& tab("Address\_tmp2", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
+& fk("LegalPerson\_address", "LegalPerson", \varnothing))\\
+) 
+\end{align*}
+\begin{align*}
+r &= \{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
+& row("person", ["Id", 2], ["surname", "Doe"]), \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."]), \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"])\\
+& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], \\ 
+& ["city", "NY"], ["zip", 10001], ["Person", 1]) \\
+& row("address\_tmp1", ["Id", 2], ["street", "Olive ave. 2225"], \\
+& ["city", "LA"], ["zip", 90003], ["Person", 2]) )\\
+& row("address\_tmp2", ["Id", 10], ["street", "Hill stret 22"], \\
+& ["city", "NY"], ["zip", 10001], ["LegalPerson", 10]) \\
+& row("address\_tmp2", ["Id", 20], ["street", "Pine ave. 2225"], \\
+& ["city", "LA"], ["zip", 90003], ["LegalPerson", 20]) )\\
+\}
 \end{align*}

### 2013-01-30 doc/description.tex
updated merge class and case study

@@
 \begin{align*}
-&	\mathbf{merge} : Label \times Label \times Label \times State \rightarrow State \\
-&	\Phi(s, merge(l_1, l_2, l_n) = \inference{
+&	\mathbf{Merge} : Label \times Label \times Label \times State \rightarrow State \\
+&	\Phi(s, Merge(l_1, l_2, l_n) = \inference{
 		\exists c_1, c_2 \in classes(application(s)) \wedge 
 		equals(c_1, c_2)}{\begin{gathered}
-			application(s) \circ addClass(cl(ln, \varnothing, \varnothing)\\ \circ (\forall p \in properties(c_1) : addProperty(l_n, p)) \\ \circ (\forall a \in associations(c_1): addAssociation(l_n, a))
+			application(s) \\ \circ addClass(cl(ln, properties(c_1), associations(c_1)) \\
+			\circ (\forall c \in classes (application(s)), a \in associations(c) \\
+			reference(a) = c_1 \vee reference(a) = c_2 : \\
+			reference(a) = l_n
 		\end{gathered}} \\\\
-&	\Psi(s, merge(l_1, l_2, l_n) = \inference{\exists t_1, t_2 \in tables(database(s)) : \\ name(t_1) = l_1 \wedge name(t_2) = l_2 \\
-\wedge ! \: referenced(t_1, d) \wedge ! \: referenced(t_2, d) }{\begin{gathered}
-	 database(s) \bullet addTable(l_n, \mathbf{X}, \varnothing, \varnothing) \\ \bullet (\forall c \in columns(t_1) : addColumn(l_n, c)) \\ \bullet (\forall f \in foreignKeys(t_1) : addForeignKey(l_n, f)) \\ \bullet (\forall r \in instances(t_1, database(s)) \cup instances(t_2, database(s)) :  \\ insertRow(l_n, row(l_n, next(seq(key(l_n))), pairs(r))))
-\end{gathered}}\\
+&	\Psi(s, merge(l_1, l_2, l_n) = \inference{\exists t_1, t_2 \in tables(database(s)) : \\ name(t_1) = l_1 \wedge name(t_2) = l_2}{\begin{gathered}
+	 database(s) \\ \bullet addTable(l_n, pk("Id", s_1), columns(t_1), foreignKeys(t_1)) \\ 
+	 \bullet (\forall r \in instances(t_1, database(s)) \cup instances(t_2, database(s)) :  \\ n =  row(l_n, next(seq(key(l_n))), pairs(r)), insertRow(l_n, n) \\ 
+	\bullet (\forall k \in data(database(s)), f \in foreignKeys(k) \\ \wedge reference(f) = reference(r) \\ \wedge value(findPair(name(f), k)) = pkValue(r) : \\value(findPair(name(f), k) = pkValue(n)) \end{gathered}} \\
 	& l_1, l_2, l_n \in Label, s \in State
@@
 To improve the design of code the class Address has to be created which associated with both original classes and contains the data already stored in the software.
-The initial situation:
-\begin{align*}
-&	S = sw(a, db(t,r))
-\end{align*}
-\begin{align*}
-a &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
-& pr(street, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
-& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
-& pr(street, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing))
-\end{align*}
-\begin{align*}
-t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
-& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
-& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
-& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
-& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing)) 
-\end{align*}
-\begin{align*}
-r &= row("person", ["Id", 1], ["surname", "Griffin"], \\
-& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("person", ["Id", 2], ["surname", "Doe"], \\
-& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
-& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
-& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
-& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003]))
-\end{align*}
+This cen be done with following evolution:
 The first evolution step is to create the address class for $Person$ and $LegalPerson$, then the temporary class can be merged together and finally the temporary class has to be renamed to $Address$:
@@
 \end{align*}
+States of the evolution are shown in figures, although it could be described formally (for formal description see appendix \ref{sec:caseStudy}) the graphical illustrations are more illustrative.
+The initial situation is :
 The software state after adding the second association is:
@@
 \section{Case Study}
+\label{sec:caseStudy}
+Initial State
+\hline
+\begin{align*}
+&	S = sw(a, db(t,r))
+\end{align*}
+\begin{align*}
+a &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& pr(street, APPSTRING, \varnothing), \\
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing))
+\end{align*}
+\begin{align*}
+t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
+& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
+& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
+& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
+& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing)) 
+\end{align*}
+\begin{align*}
+r &= row("person", ["Id", 1], ["surname", "Griffin"], \\
+& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("person", ["Id", 2], ["surname", "Doe"], \\
+& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
+& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
+& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003]))
+\end{align*}
+\newpage
+
 ExtactClass = addClass

### 2013-01-30 doc/description.tex
case study text update

@@
 \usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 \usepackage{algpseudocode}
@@
 			reference(a) = c_1 \vee reference(a) = c_2 : \\
-			reference(a) = l_n
-		\end{gathered}} \\\\
-&	\Psi(s, merge(l_1, l_2, l_n) = \inference{\exists t_1, t_2 \in tables(database(s)) : \\ name(t_1) = l_1 \wedge name(t_2) = l_2}{\begin{gathered}
+			reference(a) = l_n) \\
+			\circ renClass(l_1) \circ renClass(l_2) \end{gathered}}
+ \\\\
+&	\Psi(s, merge(l_1, l_2, l_n)) = \inference{\exists t_1, t_2 \in tables(database(s)) : \\ name(t_1) = l_1 \wedge name(t_2) = l_2 \\ \wedge columns(t_1) = columns(t_2)}{\begin{gathered}
 	 database(s) \\ \bullet addTable(l_n, pk("Id", s_1), columns(t_1), foreignKeys(t_1)) \\ 
 \bullet (\forall r \in instances(t_1, database(s)) \cup instances(t_2, database(s)) :  \\ n =  row(l_n, next(seq(key(l_n))), pairs(r)), insertRow(l_n, n) \\ 
-	\bullet (\forall k \in data(database(s)), f \in foreignKeys(k) \\ \wedge reference(f) = reference(r) \\ \wedge value(findPair(name(f), k)) = pkValue(r) : \\value(findPair(name(f), k) = pkValue(n)) \end{gathered}} \\
+\bullet (\forall k \in data(database(s)), f \in foreignKeys(k) \\ \wedge reference(f) = reference(r) \\ \wedge value(findPair(name(f), k)) = pkValue(r) : \\value(findPair(name(f), k)) = pkValue(n))) \\  \bullet dropTableAndData(l_1) \bullet dropTableAndData(l_2)
+ \end{gathered}} \\
  & l_1, l_2, l_n \in Label, s \in State
@@
 There are two classes in our software $Person$ and $LegalPerson$ both of them contains information about address (street, city and zip code).
+The situation in software is in Figure \ref{fig:case1}, where the whole software is shown.
+
 To improve the design of code the class Address has to be created which associated with both original classes and contains the data already stored in the software.
-This cen be done with following evolution:
-The first evolution step is to create the address class for $Person$ and $LegalPerson$, then the temporary class can be merged together and finally the temporary class has to be renamed to $Address$:
+The evolution is defined in following text together with graphical illustrations of its processing.
+The whole evolution processing can be found in the appendix \ref{sec:caseStudy}.
+
+\begin{figure}
+\begin{subfigure}[b]{0.5\textwidth}
+	\includegraphics[width=\textwidth]{./images/case_app_1}
+\end{subfigure}
+\begin{subfigure}[b]{0.5\textwidth}
+	\includegraphics[width=\textwidth]{./images/case_db_1}
+\end{subfigure}
+\end{figure}
+
+The first  step of is to extract two temporary classes representing address of a $Person$ or of a $LegalPerson$. Temporary classes has to be connected with the origin classes by an association.
 \begin{align*}
@@
 & y \in instances("Address\_tmp2", database(S)) : pkVal(x) = pkVal(y) \}) \\
+\end{align*}
+
+\begin{figure}
+\begin{subfigure}[b]{0.5\textwidth}
+	\includegraphics[width=\textwidth]{./images/case_app_5}
+\end{subfigure}
+\begin{subfigure}[b]{0.5\textwidth}
+	\includegraphics[width=\textwidth]{./images/case_db_5}
+\end{subfigure}
+\end{figure}
+%Next two evolution steps are bit intuitive but necessary for the next evolution.  then the temporary class can be merged together and finally the temporary class has to be renamed to $Address$:
+The two temporary classes can be merged together as they have the same structure.
+All associations pointing to one of the temporary classes are redirected to the newly created class.
+Finally the associations are renamed.
+\begin{align*}
 %
-& \oplus AddAssociation("Person", \\
-& assoc("address", "Address\_tmp2", 1, 1), \varnothing) \\
-& \oplus AddAssociation("LegalPerson", \\
-& assoc("residentialAddress", "Address\_tmp1", 1, 1),\varnothing) \\
+%& \oplus AddAssociation("Person", \\
+%& assoc("address", "Address\_tmp2", 1, 1), \varnothing) \\
+%& \oplus AddAssociation("LegalPerson", \\
+%& assoc("residentialAddress", "Address\_tmp1", 1, 1),\varnothing) \\
 %
 & \oplus Merge("Address\_tmp1", "Address\_tmp2", "Address") \\
-& \oplus RemoveTableAndData("Address\_tmp1") \\
-& \oplus RemoveTableAndData("Address\_tmp2")
-\end{align*}
-States of the evolution are shown in figures, although it could be described formally (for formal description see appendix \ref{sec:caseStudy}) the graphical illustrations are more illustrative.
-The initial situation is :
-The software state after adding the second association is:
+\end{align*}
+Finally the associations are renamed and the evolution end.
+\begin{align*}
+& \oplus RenameAssociation("Person", "Address\_tmp1", "Address")\\
+& \oplus RenameAssociation("LegalPerson", "Address\_tmp2". "Address")\\
+\end{align*}
+
+\begin{figure}
+\begin{subfigure}[b]{0.5\textwidth}
+	\includegraphics[width=\textwidth]{./images/case_app_9}
+\end{subfigure}
+\begin{subfigure}[b]{0.5\textwidth}
+	\includegraphics[width=\textwidth]{./images/case_db_8}
+\end{subfigure}
+\end{figure}
 

### 2013-01-30 doc/description.tex
case study summary, merge transformation fixed

@@
  \\\\
-&	\Psi(s, merge(l_1, l_2, l_n)) = \inference{\exists t_1, t_2 \in tables(database(s)) : \\ name(t_1) = l_1 \wedge name(t_2) = l_2 \\ \wedge columns(t_1) = columns(t_2)}{\begin{gathered}
+&	\Psi(s, merge(l_1, l_2, l_n)) = \inference{\exists t_1, t_2 \in tables(database(s)) : \\ name(t_1) = l_1 \wedge name(t_2) = l_2 \\ \wedge columns(t_1) = columns(t_2)\\
+\forall f \in (foreignKeys(t_1) \cup foreignKey(t_2)) \\ \setminus (foreignKey(t_1) \cap foreignKey(t_2)) \\ : NOTNULL \not \in constraints(f)  }{\begin{gathered}
 	 database(s) \\ \bullet addTable(l_n, pk("Id", s_1), columns(t_1), foreignKeys(t_1)) \\ 
@@
 	\includegraphics[width=\textwidth]{./images/case_app_1}
+	\caption{The application model.}
 \end{subfigure}
@@
 	\includegraphics[width=\textwidth]{./images/case_db_1}
+	\caption{The database model.}
 \end{subfigure}
+	\caption{The initial state of the example used in the case study. There is a repetition of information structure in the $Person$ and $LegalPerson$ class.}
+	\label{fig:case1}
 \end{figure}
 
-The first  step of is to extract two temporary classes representing address of a $Person$ or of a $LegalPerson$. Temporary classes has to be connected with the origin classes by an association.
-\begin{align*}
-& state(S) \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip") \\
+The first  step of is to extract two temporary classes representing address of a $Person$ or of a $LegalPerson$.
+Temporary classes has to be connected with the origin classes by an association.
+The result is shown in Figure \ref{fig:case2}.
+\begin{align*}
+S_1 &= state(S) \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip") \\
 & \oplus AddAssociation("Person", assoc("address", "Address\_tmp1", 1, 1),\\ 
@@
 	\includegraphics[width=\textwidth]{./images/case_app_5}
+	\caption{The application model.}
 \end{subfigure}
@@
 	\includegraphics[width=\textwidth]{./images/case_db_5}
+	\caption{The database model.}
 \end{subfigure}
+	\caption{The state of the software after the information are extracted into separated classes and associations are created.}
+	\label{fig:case2}
 \end{figure}
 %Next two evolution steps are bit intuitive but necessary for the next evolution.  then the temporary class can be merged together and finally the temporary class has to be renamed to $Address$:
-The two temporary classes can be merged together as they have the same structure.
-All associations pointing to one of the temporary classes are redirected to the newly created class.
-Finally the associations are renamed.
+The two temporary classes can be merged together as they have the same structure. All associations pointing to one of the temporary classes are redirected to the newly created class.
 \begin{align*}
@@
 %
-& \oplus Merge("Address\_tmp1", "Address\_tmp2", "Address") \\
+S_2 &= S_1 \oplus Merge("Address\_tmp1", "Address\_tmp2", "Address") \\
 \end{align*}
@@
 \begin{align*}
-& \oplus RenameAssociation("Person", "Address\_tmp1", "Address")\\
+S_3 &= S_2  \oplus RenameAssociation("Person", "Address\_tmp1", "Address")\\
 & \oplus RenameAssociation("LegalPerson", "Address\_tmp2". "Address")\\
@@
 	\includegraphics[width=\textwidth]{./images/case_app_9}
+	\caption{The application model.}
 \end{subfigure}
@@
 	\includegraphics[width=\textwidth]{./images/case_db_8}
+	\caption{The database model.}
 \end{subfigure}
+	\caption{The final state of the case study example. There is one class containing information about address for the whole software and data were migrated.}
 \end{figure}
+
+The scenario was completed successfully, because there were no constraints defined in the database and the data was suitable for the evolution there were no complication.
+The example of evolution presented in case study shows the possibilities of defined transformation.
+The example itself is not large, nevertheless it shows advanced refactoring cases, which are very often used in real software development.
+We believe that composition of proposed transformation can be used even in case of large data evolution cases.
+It is obvious that usage of proposed set of transformations only will cause very large scripts describing data evolution, therefore a mechanism for creating templates (reusable evolution scripts).
 

### 2013-01-30 doc/description.tex
small formatting

@@
 \begin{align*}
-&	\mathbf{Property} = <label, AppType, DefaultValue, Cardinality, Mandatory> \\
+&	\mathbf{Property} = (label, AppType, DefaultValue, Cardinality, Mandatory) \\
 &	\mathbf{prop} : Label \times AppType \times String \times \mathbb{N^{*}} \times Boolean \rightarrow Property \\ \\
@@
 \begin{align*}
-&	\mathbf{Association} = <label, classRef, startCardinality, endCardinality> \\
+&	\mathbf{Association} = (label, classRef, startCardinality, endCardinality) \\
 &	\mathbf{assoc} : Label \times Label \times \mathbb{N} \times \mathbb{N} \rightarrow Association\\ \\
@@
 \begin{align*}
-&	\mathbf{ForeignKey} = < Label, Label, Constraint*> \\ 
+&	\mathbf{ForeignKey} = (Label, Label, Constraint*) \\ 
 &	\mathbf{fk} : Label \times Label \times Constraints \rightarrow ForeignKey \\ \\
@@
 \begin{align*}
-&	\mathbf{Row} = < Label, N, Pair* > \\
+&	\mathbf{Row} = (Label, N, Pair*) \\
 &	\mathbf{row} : Label \times Pairs \rightarrow Row \\ \\
@@
 \begin{align*}
-&	\mathbf{Pair} = < Label, Value > \\
+&	\mathbf{Pair} = (Label, Value) \\
 &	[Label \times Value] \rightarrow Pair\\ \\
@@
 & 	remProp(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p
-	}{properties(c) \setminus p }\\
+	}{properties(c) =- \; p }\\
 &	l_c, l_p \in Label, a \in Application 
@@
 &	\mathbf{remAssoc} : Label \times Label \times Application \rightarrow Application \\
-&	remAssoc(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) \setminus f} \\
+&	remAssoc(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) =- \; f} \\
 &	l_c, l_a \in Labels, a \in Application
@@
 &	remClass(l, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge !\: associated(c)
-	} {classes(a) \setminus c }\\
+	} {classes(a) =- \; c }\\
 &	l \in Label,a \in Application
@@
 &	\mathbf{dropValue} : Label \times Row \times Database \rightarrow Database \\
-&	dropValue(l, r, d) = \inference{\exists p \in pairs(r) : reference(p) = l}{pairs(r) \setminus p}\\ 
+&	dropValue(l, r, d) = \inference{\exists p \in pairs(r) : reference(p) = l}{pairs(r) =- \; p}\\ 
 & l \in Label, r : Row \wedge r \in data(d), d \in Database
@@
 &	\mathbf{dropRow} :  Row*  \times Database \rightarrow Database\\
-&	dropRow(r, d) = \inference{|r|=1}{data(d) \setminus r}	\\ 
+&	dropRow(r, d) = \inference{|r|=1}{data(d) =- \; r}	\\ 
 & l \in Label, r \in Row \wedge r \in data(d), d \in Database
@@
 &	\mathbf{dropTable} : Label \times Database \rightarrow Database \\
-&	dropTable(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \forall r \in data(d) : owningTable(r) \neq l \\  \wedge (\forall j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) \neq l) }{ tables(d) \setminus \rho(c)}\\
+&	dropTable(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \forall r \in data(d) : owningTable(r) \neq l \\  \wedge (\forall j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) \neq l) }{ tables(d) =- \; \rho(c)}\\
 & l \in Label, d \in Database
@@
 &	\mathbf{dropColumn} : Label \times Label \times Database \rightarrow Database \\
-& 	dropColumn(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in columns(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in instances(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{columns(t) \setminus c}\\
+& 	dropColumn(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in columns(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in instances(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{columns(t) =- \; c}\\
 &	 l , l_p\in Label, d \in Database 
@@
 &	\mathbf{dropKey} : Label \times Label \times Database \rightarrow Database \\
-&	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in instances(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) \setminus c}\\
+&	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in instances(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) =- \; c}\\
 &	 l, l_p \in Label, d \in Database 
@@
 \subsection{Software Creation}
-There are defined three transformations which are able to create a software: $addProperty$, $addAssociation$,  $newClass$.
+There are defined three transformations which are able to create a software: $addProperty$, $addAssociation$,  $addClass$.
 These three transformations can be used to create a software from a scratch.
@@
 & \Phi(s, addAsociation(l, a, m_{l_1, l_2})) = \\
-& =\begin{cases}
+& = \begin{cases}
 	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l\\ 		\wedge name(t_t) = reference(a) \\ \wedge \forall f \in foreignKeys(t_t) \cup columns(t_t) \\: name(f) \neq name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
 		database(s) \\ \bullet addForeignKey(name(t_t), \rho(a,d), d) \\ \bullet \forall r \in m_{l_1, l_2} : \\
-		insertValue(name(t_t), value(primaryKey(second(r))), [name(a), value(primaryKey(first(r)))])
+		insertValue(name(t_t), value(primaryKey(second(r))), \\ [name(a), value(primaryKey(first(r)))])
 	\end{gathered}}
@@
 
-\paragraph{Add Class} The $insertClass$ operation changes the structure of the database schema only, because the $insertAssoc$ transformation needs a mapping function to work properly, which is not part of the $newClass$ definition.
+\paragraph{Add Class} The $insertClass$ operation changes the structure of the database schema only, because the $insertAssoc$ transformation needs a mapping function to work properly, which is not part of the $addClass$ definition.
 The situation reflect the modeling praxis when the structure of a class is defined and then its associations are created.
 \begin{align*}
-& \mathbf{newClass} : Class \times State \rightarrow State \\
-& \Psi(s, newClass(c)) = application(s) \circ addClass(c) \\
-& \Phi(s, newClass(c)) = database(s)  \bullet addTable(\rho(c,d)) \\ 
+& \mathbf{addClass} : Class \times State \rightarrow State \\
+& \Psi(s, addClass(c)) = application(s) \circ addClass(c) \\
+& \Phi(s, addClass(c)) = database(s)  \bullet addTable(\rho(c,d)) \\ 
 & \bullet addColumn(name(c), \{\forall r \in singleProperties(c) | \rho(r,d)\}, database(s)) \\

### 2013-01-31 doc/description.tex
tables in case study

@@
 \end{subfigure}
+\begin{subfigure}[b]{\textwidth}
+	\centering
+	\begin{tabular}{| c | l | l | l | l | }
+	 	\hline
+		Id &  Surname & Street & City & ZIP  \\ \hline  
+		11 & Jackson & Central Park St & New York & 100 01  \\ \hline
+		12 & Clooney & S Orange Ave & Orlando & 320 24  \\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $Person$.}
+\end{subfigure}
+\begin{subfigure}[b]{\textwidth}
+	\centering
+	\begin{tabular}{| c | l | l | l | l | c |}
+	 	\hline
+		Id &  BusinessName & Street & City & ZIP \\ \hline  
+		100 & Tools \& Machines & Olive ave. & New York & 100 01 \\ \hline
+		200 & AI Robotics & Pine ave. & LA & 900 03  \\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $LegalPerson$.}
+\end{subfigure}
 	\caption{The initial state of the example used in the case study. There is a repetition of information structure in the $Person$ and $LegalPerson$ class.}
@@
 \end{subfigure}
+\begin{subfigure}[b]{\textwidth}
+	\centering
+	\begin{tabular}{| c | l |}
+	 	\hline
+		Id &  Surname \\ \hline  
+		11 & Jackson  \\ \hline
+		12 & Clooney  \\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $Person$.}
+\end{subfigure}
+\begin{subfigure}[b]{\textwidth}
+	\centering
+	\begin{tabular}{| c | l | l | l | l |}
+	 	\hline
+		Id & Street & City & ZIP & Person \\ \hline  
+		11 & Central Park St & New York & 100 01 & 11  \\ \hline
+		12 & Orange Ave & Orlando & 320 24 & 12 \\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $Address\_tmp1$.}
+\end{subfigure}
+\begin{subfigure}[b]{\textwidth}
+	\centering
+	\begin{tabular}{| c | l |}
+	 	\hline
+		Id &  BusinessName  \\ \hline  
+		100 & Tools \& Machines  \\ \hline
+		200 & AI Robotics \\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $LegalPerson$.}
+\end{subfigure}
+\begin{subfigure}[b]{\textwidth}
+	\centering
+	\begin{tabular}{| c | l | l | l | l |}
+	 	\hline
+		Id & Street & City & ZIP & LegalPerson\\ \hline  
+		100 & Olive ave. & New York & 100 01 & 100 \\ \hline
+		200 & Pine ave. & LA & 900 03 & 200  \\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $Address\_tmp2$.}
+\end{subfigure}
+
 	\caption{The state of the software after the information are extracted into separated classes and associations are created.}
@@
 \end{subfigure}
+
+\begin{subfigure}[b]{\textwidth}
+	\centering
+	\begin{tabular}{| c | l |}
+	 	\hline
+		Id &  Surname \\ \hline  
+		11 & Jackson  \\ \hline
+		12 & Clooney  \\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $Person$.}
+\end{subfigure}
+\begin{subfigure}[b]{\textwidth}
+	\centering
+	\begin{tabular}{| c | l |}
+	 	\hline
+		Id &  BusinessName  \\ \hline  
+		100 & Tools \& Machines  \\ \hline
+		200 & AI Robotics \\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $LegalPerson$.}
+\end{subfigure}
+\begin{subfigure}[b]{\textwidth}
+	\centering
+	\begin{tabular}{| c | l | l | l | l | l |}
+	 	\hline
+		Id & Street & City & ZIP & Person & LegalPerson\\ \hline  
+		1 & Olive ave. & New York & 100 01 & & 100 \\ \hline
+		2 & Pine ave. & LA & 900 03 & & 200  \\ \hline
+		3 & Central Park St & New York & 100 01 & 11 &  \\ \hline
+		4 & Orange Ave & Orlando & 320 24 & 12 &\\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $Address\_tmp2$.}
+\end{subfigure}
+
 	\caption{The final state of the case study example. There is one class containing information about address for the whole software and data were migrated.}

### 2013-01-31 doc/description.tex
formatting

@@
 &	\mathbf{consistent} : Application \rightarrow Boolean \\
-&	consistent(a) = \begin{cases}
-	\inference{\forall c_1, c_2 \in classes(a) : name(c_1) \neq name(c_2) \\ \wedge \forall p_1, p_2 \in properties(c_1) : name(p_1) \neq name(p_2) \\ \wedge \forall a_1, a_2 \in associations(c_1) : name(a_1) \neq name (a_2) \\ \wedge \forall a \in associations(c_1) : \exists c \in classes(a) \wedge name(c) = reference(a)\\
+&	consistent(a)= \\
+=&\begin{cases}
+	\inference{\forall c_1, c_2 \in classes(a) : name(c_1) \neq name(c_2) \\ \wedge \forall p_1, p_2 \in properties(c_1) : name(p_1) \neq name(p_2) \\ \wedge \forall a_1, a_2 \in associations(c_1) : name(a_1) \neq name (a_2) \\ \wedge \forall a \in associations(c_1) : \exists c \in classes(a) \\ \wedge name(c) = reference(a)\\
 	\wedge \forall p \in properties(c_1), a \in associations(c_1) : name(p) \neq name(a) }{true}\\\\
@@
 &	\mathbf{instantiated} : Class \times Software \rightarrow Boolean \\
-&	instantiated(c, s) = \begin{cases} \inference{\exists t \in tables(database(s)) : name(t) = name(c) \\ \wedge contains(t, database(s))}{true} \\ \\
+&	instantiated(c, s) =\\
+& \;\;\;\;\;\; = \begin{cases} \inference{\exists t \in tables(database(s)) : \\ name(t) = name(c)  \wedge contains(t, database(s))}{true} \\ \\
   false
@@
 &	\mathbf{consistent} \in Database \rightarrow Boolean \\
-&	consistent(d) = \begin{cases}
-&		\inference{ \forall t_1, t_2 \in tables(d) : name(t_1) \neq name(t_2) \\ \wedge \forall c_1, c_2 \in columns(t_1) : name(c_1) \neq name(c_2) \\ \wedge \forall f_1, f_2 \in foreignKeys(t_1) : name(f_1) \neq name (f_2) \\ \wedge \forall f \in foreignKeys(t_1) : \exists t \in tables(d) \wedge name(t) = reference(f) \\ \wedge \forall r \in data(d): ( \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) :\\ \exists u \in columns(t) \cup foreignKeys(t) \cup primaryKey(t) : \\ reference(p) = name(u) \\ \wedge \forall e \in columns(t) \cup foreignKeys(t) \\ \wedge NOTNULL \in constrain(e) : \exists q \in pairs(r) \\ \wedge name(e) = reference(q)) \\ \wedge \forall g \in columns(t) \cup foreignKeys(t) \\ \wedge UNIQUE \in constrain(g) : \\ \forall u_1, u_2 \in rows(t) : \\ value(pairByName(name(g), u_1) \neq value(pairByName(name(g), u_2) }{true}
+&	consistent(d) = \\
+& \;\;\;\;\;\; =\begin{cases}
+		\inference{ \forall t_1, t_2 \in tables(d) : name(t_1) \neq name(t_2) \\ \wedge \forall c_1, c_2 \in columns(t_1) : name(c_1) \neq name(c_2) \\ \wedge \forall f_1, f_2 \in foreignKeys(t_1) : name(f_1) \neq name (f_2) \\ \wedge \forall f \in foreignKeys(t_1) : \exists t \in tables(d) \wedge name(t) = reference(f) \\ \wedge \forall r \in data(d): ( \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) :\\ \exists u \in columns(t) \cup foreignKeys(t) \cup primaryKey(t) : \\ reference(p) = name(u) \\ \wedge \forall e \in columns(t) \cup foreignKeys(t) \\ \wedge NOTNULL \in constrain(e) : \exists q \in pairs(r) \\ \wedge name(e) = reference(q)) \\ \wedge \forall g \in columns(t) \cup foreignKeys(t) \\ \wedge UNIQUE \in constrain(g) : \\ \forall u_1, u_2 \in rows(t) : \\ value(pairByName(name(g), u_1) \neq value(pairByName(name(g), u_2) }{true}
  	\\ \\
@@
 
+\paragraph{Mapping of Names} There can occur situations when a consistent application could be mapped into an inconsistent database because of name collisions or the name can collide with some internal database construct (which are not right now implemented in the model).
+
 \paragraph{Mapping of Classes} The class is mapped into a table with  automatically created primary key.
@@
 
-\subsubsection{Mapping of Properties}
+\paragraph{Mapping of Properties}
 There are two kinds of properties - primitive property with cardinality equal to one is mapped to a column; second kind - collection with cardinality greater than one is mapped to table, thus there are two mapping cases.
@@
 			 \\ \wedge \exists t : t = table(name(owningClass(p)), d)
-			\\ \wedge \not instanciated(owningClass(p))
+			\\ \wedge ! instanciated(owningClass(p))
 		}{\begin{gathered}
@@
 &	\mathbf{\rho_p} : Property* \rightarrow Table* \\ \\
-&	\rho_{cp}(p) = \begin{cases}
+&	\rho_{cp}(p) = \\
+& \;\;\;\;\;\; =\begin{cases}
 		%\inference{p = \varnothing}{\varnothing}\\
@@
   		}{\begin{gathered}  
-		 table(name(t)+name(a), pk("Id"), \varnothing, \\ foreignKey(owningClass(a),owningClass(a), NOTNULL) \\ foreignkey(reference(a), reference(a), NOTNULL)) 
+		 table(name(t)+name(a), pk("Id"), \varnothing, \\ fk(name(owningClass(a)), name(owningClass(a)), NOTNULL) \\ fk(reference(a), reference(a), NOTNULL)) 
   		\end{gathered}}  
@@
 & 	remProp(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p
-	}{properties(c) =- \; p }\\
+	}{properties(c) - \; p }\\
 &	l_c, l_p \in Label, a \in Application 
@@
 &	\mathbf{remAssoc} : Label \times Label \times Application \rightarrow Application \\
-&	remAssoc(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) =- \; f} \\
+&	remAssoc(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) - \; f} \\
 &	l_c, l_a \in Labels, a \in Application
@@
 &	remClass(l, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge !\: associated(c)
-	} {classes(a) =- \; c }\\
+	} {classes(a) - \; c }\\
 &	l \in Label,a \in Application
@@
 &	alterTable(l_o, l_n, d) = \inference{\exists t \in tables(d) : name(d) = l_o \\ \wedge \forall e \in tables(d) : name(e) \neq l_n}{\begin{gathered}
-		t.name = l_n \\ \wedge \forall e \in tables(a) \wedge \forall (f \in fks(e) \wedge reference(f) = l_o ) \\ : f.reference = l_n 
+		t.name = l_n \wedge \forall e \in tables(a) \\ \wedge ( \forall f \in fks(e) \wedge reference(f) = l_o ) \\ : f.reference = l_n 
 	\end{gathered}}\\
@@
 &	\mathbf{insertRow} : Row* \times Database \rightarrow Database \\
-&	insertRow(r, d) = \begin{cases}
+&	insertRow(r, d) = \\
+& \;\;\;\;\;\; =\begin{cases}
 			%\inference{|r|>1}{d \bullet insertRow(head(r), d) \bullet insertRow(tail(r), d)} \\ \\
@@
 &	\mathbf{insertValue} : Label \times N \times Pair* \times Database \rightarrow Daatabase \\
-&	insertValue(l, n, p, d) = \begin{cases}
+&	insertValue(l, n, p, d) =\\
+& \;\;\;\;\;\; =\begin{cases}
  		%\inference{|p| > 1}{\begin{gathered} 	d \bullet insertValue(r, head(p), d) \\ \bullet insertValue(r, tail(p), d)\end{gathered}}   \\ \\
@@
 &	\mathbf{addValueToTable} : Label \times Pair \times Database \rightarrow Database \\
-&	addValueToTable(l, p, d) = \begin{cases}
+&	addValueToTable(l, p, d) = \\
+&= \;\;\;\;\;\;\begin{cases}
  		\inference{\exists t \in tables(d) : name(t) = l}	{\begin{gathered}
@@
 &	\mathbf{copyValue} : Label \times Label \times Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
-&	copyValue(l_s, l_t, l_{sc}, l_{tc} o_{l_s, l_t}, d) = \begin{cases}
-		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc}}{\begin{gathered}
+&	copyValue(l_s, l_t, l_{sc}, l_{tc} o_{l_s, l_t}, d) = \\
+& \;\;\;\;\;\; =	\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc}}{\begin{gathered}
 		\forall	p \in m_{l_s, l_t} :	d \bullet insertValue(name(second(p)), keyValue(second(p)), \\ [l_{tc},value(findPair(l_{sc}, first(p)))])
-		\end{gathered}}
- 	\end{cases} \\
+		\end{gathered}}\\
 & 	l_s, l_t, l_{sc}, l_{tc} \in Label, d \in Database
@@
 &	\mathbf{copyRow} : Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
-&	copyRow(l_s, l_t, m_{l_s, l_t}, d) = \begin{cases}
+&	copyRow(l_s, l_t, m_{l_s, l_t}, d) = \\
+& \;\;\;\;\;\; =%\begin{cases}
  	%\inference{|m_{l_s, l_t}| = \varnothing}{d} \\\\
@@
  	\inference{|m_{l_s, l_t}| = 1 \\ \wedge \exists t_s, t_t \in tables(d) : name(t_s) = l_s, name(t_t) = l_t \\ \wedge t_s \cong t_t }{d \bullet insertRow(row(l, pairs(first(m_{l_s, l_t}))), d)} \\\\
-  \end{cases} \\
+%  \end{cases} 
+\\
 &  l_s, l_t \in Label, d \in Database, m_{l_s, l_t} \in M
@@
 &	\mathbf{dropValue} : Label \times Row \times Database \rightarrow Database \\
-&	dropValue(l, r, d) = \inference{\exists p \in pairs(r) : reference(p) = l}{pairs(r) =- \; p}\\ 
+&	dropValue(l, r, d) = \inference{\exists p \in pairs(r) : reference(p) = l}{pairs(r) - \; p}\\ 
 & l \in Label, r : Row \wedge r \in data(d), d \in Database
@@
 &	\mathbf{dropRow} :  Row*  \times Database \rightarrow Database\\
-&	dropRow(r, d) = \inference{|r|=1}{data(d) =- \; r}	\\ 
+&	dropRow(r, d) = \inference{|r|=1}{data(d) - \; r}	\\ 
 & l \in Label, r \in Row \wedge r \in data(d), d \in Database
@@
 &	\mathbf{dropTable} : Label \times Database \rightarrow Database \\
-&	dropTable(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \forall r \in data(d) : owningTable(r) \neq l \\  \wedge (\forall j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) \neq l) }{ tables(d) =- \; \rho(c)}\\
+&	dropTable(l,d) = \\
+& \;\;\;\;\;\; =\inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \forall r \in data(d) : owningTable(r) \neq l \\  \wedge (\forall j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) \neq l) }{ tables(d) - \; \rho(c)}\\
 & l \in Label, d \in Database
@@
 &	\mathbf{dropColumn} : Label \times Label \times Database \rightarrow Database \\
-& 	dropColumn(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in columns(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in instances(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{columns(t) =- \; c}\\
+& 	dropColumn(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in columns(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in instances(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{columns(t) - \; c}\\
 &	 l , l_p\in Label, d \in Database 
@@
 &	DropColumnAndData(l_c, ps, d) = \\
-&= \begin{cases}
+& \;\;\;\;\;\; =%\begin{cases}
  	%\inference{ps = \varnothing}{d}\\
@@
  	(\inference{|ps| = 1 \\ \wedge \exists t \in tables(d) \\ \wedge \exists c \in columns(t) : name(c) = l_c}{\begin{gathered}
-	 	(\forall r, r \in instances(d) \wedge reference(r) = l_c: \\d \bullet dropValue(ps, r)) \end{gathered}}) \\ \bullet dropColumn(l_c)  
- \end{cases}
+	 	(\forall r, r \in instances(d) \wedge reference(r) = l_c: \\d \bullet dropValue(ps, r)) \end{gathered}})  \bullet dropColumn(l_c)  
+% \end{cases}
 	\\
@@
 &	\mathbf{dropKey} : Label \times Label \times Database \rightarrow Database \\
-&	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in instances(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) =- \; c}\\
+&	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in instances(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) - \; c}\\
 &	 l, l_p \in Label, d \in Database 
@@
 &	DropKeyAndData(l_f, ps, d) = \\
-& = \begin{cases}
+& \;\;\;\;\;\; = \begin{cases}
  	%\inference{ps = \varnothing}{d}\\
@@
 &	\Phi(s, addProperty(l, p)) = \\
-& = \begin{cases}
+& \;\;\;\;\;\; = \begin{cases}
     		\inference{cardinality(p) \leq 1}{database(s) \bullet addColumn(l, \rho(p,d), d)} \\ \\
@@
 & \Phi(s, addAsociation(l, a, m_{l_1, l_2})) = \\
-& = \begin{cases}
+& \;\;\;\;\;\; = \begin{cases}
 	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l\\ 		\wedge name(t_t) = reference(a) \\ \wedge \forall f \in foreignKeys(t_t) \cup columns(t_t) \\: name(f) \neq name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
@@
 &	\Phi RemoveProperty(l, l_p, s) = \\
-& \begin{cases}
+& \;\;\;\;\;\; =\begin{cases}
 		 \inference{\exists c \in classes(application(s)) : name(c) = l\\ 
@@
 &	\Psi(RemovePropertyAndData(L_c, ps), s) = \\
-& = \begin{cases}
+& \;\;\;\;\;\; = \begin{cases}
 		%\inference{ps = \varnothing}{d}\\
@@
 &    \Phi RemoveAssociation(l, l_a, s) = \\
-&= s\begin{cases}
+& \;\;\;\;\;\; =\begin{cases}
      	\inference{\exists c \in classes(application(s)) : name(c) = l \\
@@
 &	\Psi(RemoveAssociationAndData(L_c, ps), s) = \\
-&=  \begin{cases}
+& \;\;\;\;\;\; = \begin{cases}
 		%\inference{ps = \varnothing}{d}\\
@@
 & \Psi(RenameProperty(l_c, l_o, l_n, s)) = application(s) \circ renProp(l_c, l_o, l_n) \\
-& \Phi(RenameProperty(l_c, l_o, l_n, s)) = \begin{cases}
+& \Phi(RenameProperty(l_c, l_o, l_n, s)) = \\
+& \;\;\;\;\;\; =\begin{cases}
  	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 
@@
  	cardinality(p) > 1}{\begin{gathered}
-		database(s)\\ \bullet alterTable(name(owningClass(p)+l_o, l_n, database(s))
+		database(s)\\ \bullet alterTable(name(owningClass(p)+l_o, \\l_n, database(s))
 \end{gathered}}
@@
 & \Psi(RenameAssocaiation(l_c, l_o, l_n), s) = application(s) \circ renAssoc(l_c, l_o, l_n)\\
-& \Phi(RenameAssocaiation(l_c, l_o, l_n), s) = \begin{cases}
+& \Phi(RenameAssocaiation(l_c, l_o, l_n), s) = \\
+& \;\;\;\;\;\; =\begin{cases}
 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 
@@
 &	\mathbf{CopyPropertyStructure} : Label \times Label \times Label \times State \rightarrow State \\
-&	\Psi(CopyPropertyStructure(l_s, l_t, l_p, s) = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)} 
+&	\Psi(CopyPropertyStructure(l_s, l_t, l_p, s) = \\
+& \;\;\;\;\;\; =\inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)} 
 	\\\\
-&	\Phi(CopyPropertyStructure(l_s, l_t, l_p, s) = \begin{cases}
+&	\Phi(CopyPropertyStructure(l_s, l_t, l_p, s) = \\
+& \;\;\;\;\;\; =\begin{cases}
     \inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) \leq 1 }{\begin{gathered}
@@
 &	\mathbf{Inline} : Label \times Label \times (Row \times Row) \times State \rightarrow State \\
-&	s \oplus Inline(l_s, l_t, m_{l_s, l_t}) = \inference{\exists c_s \in classes(application(s)) : name(c_s) = l_s \\
+&	s \oplus Inline(l_s, l_t, m_{l_s, l_t}) = \\
+& \;\;\;\;\;\; =\inference{\exists c_s \in classes(application(s)) : name(c_s) = l_s \\
 \wedge \exists c_t \in classes(application(s)) : name(c_t) = l_t \\
@@
 \begin{align*}
-S_1 &= state(S) \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip") \\
+S_1 & = state(S) \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip") \\
 & \oplus AddAssociation("Person", assoc("address", "Address\_tmp1", 1, 1),\\ 
@@
 & \oplus Extract("LegalPerson", "Address\_tmp2", "street", "city", "zip")  \\
-& \oplus AddAssociation("LegalPerson", assoc("residentialAddress", "Address\_tmp2", 1, 1),\\ 
+& \oplus AddAssociation("LegalPerson", \\ 
+& assoc("residentialAddress", "Address\_tmp2", 1, 1),\\ 
 & \{(x,y)| x \in instances("LegalPerson", database(S)), \\
@@
 \end{subfigure}
-\begin{subfigure}[b]{\textwidth}
+\begin{subfigure}[b]{0.5\textwidth}
 	\centering
@@
 	\caption{Data stored in the table $Person$.}
+\end{subfigure}
+\begin{subfigure}[b]{0.5\textwidth}
+	\centering
+	\begin{tabular}{| c | l |}
+	 	\hline
+		Id &  BusinessName  \\ \hline  
+		100 & Tools \& Machines  \\ \hline
+		200 & AI Robotics \\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $LegalPerson$.}
 \end{subfigure}
@@
 	\centering
-	\begin{tabular}{| c | l |}
-	 	\hline
-		Id &  BusinessName  \\ \hline  
-		100 & Tools \& Machines  \\ \hline
-		200 & AI Robotics \\ \hline
-	\end{tabular}
-	\caption{Data stored in the table $LegalPerson$.}
-\end{subfigure}
-\begin{subfigure}[b]{\textwidth}
-	\centering
 	\begin{tabular}{| c | l | l | l | l |}
@@
 %
-S_2 &= S_1 \oplus Merge("Address\_tmp1", "Address\_tmp2", "Address") \\
+S_2 & \;\;\;\;\;\; =S_1 \oplus Merge("Address\_tmp1", "Address\_tmp2", "Address") \\
 \end{align*}
@@
 \begin{align*}
-S_3 &= S_2  \oplus RenameAssociation("Person", "Address\_tmp1", "Address")\\
+S_3 & \;\;\;\;\;\; =S_2  \oplus RenameAssociation("Person", "Address\_tmp1", "Address")\\
 & \oplus RenameAssociation("LegalPerson", "Address\_tmp2". "Address")\\
@@
 
-\begin{subfigure}[b]{\textwidth}
+\begin{subfigure}[b]{0.5\textwidth}
 	\centering
@@
 \newpage
-
-ExtactClass = addClass
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{align*}
+& state(S) \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip") = \\
+&=
+\end{align*}
 \hline
@@
 & pr(street, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)),
+& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)),\\
 & cl("Address\_tmp1", \varnothing, \varnothing)
-& )
+ )
 \end{align*}
@@
 & ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
-& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+& row("legalPerson", ["Id", 10], \\
+& ["businessName", "Tools and Machines ltd."], \\
 & ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
@@
 \begin{align*}
-a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
+a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
 & pr(street, APPSTRING, \varnothing), \\
@@
 \begin{align*}
-t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
+t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
 & col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
@@
 \begin{align*}
-r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+r & \;\;\;\;\;\; =\{ row("person", ["Id", 1], ["surname", "Griffin"], \\
 & ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
@@
 \begin{align*}
-a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
+a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
 & pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
@@
 \begin{align*}
-t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
+t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
 & col(city, APPSTRING, \varnothing, \varnothing) \\
@@
 \begin{align*}
-r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+r & \;\;\;\;\;\; =\{ row("person", ["Id", 1], ["surname", "Griffin"], \\
 & ["city", "NY"], ["zip", 10001]), \\
@@
 \begin{align*}
-t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
+t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
 & col(city, DBSTRING, \varnothing, \varnothing) \\
@@
 \begin{align*}
-a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\
+a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\
 & pr(zip, APPSTRING, \varnothing)), \\
@@
 \begin{align*}
-t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
+t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
 & col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
@@
 \begin{align*}
-r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+r & \;\;\;\;\;\; =\{ row("person", ["Id", 1], ["surname", "Griffin"], \\
 & ["zip", 10001]), \\
@@
 \begin{align*}
-a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing))), \\
+a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing))), \\
 & cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
@@
 \begin{align*}
-t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
 & tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
@@
 \begin{align*}
-r &= \{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
+r & \;\;\;\;\;\; =\{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
 & row("person", ["Id", 2], ["surname", "Doe"]), \\
@@
 \begin{align*}
-a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
+a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
 & assoc("address", "Address\_tmp1", 1, 1)), \\ 
@@
 \begin{align*}
-a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
+a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
 & assoc("address", "Address\_tmp1", 1, 1), assoc("address", "Address\_tmp2", 1, 1)), \\ 
@@
 \begin{align*}
-t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
 & tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing)) \\

### 2013-02-02 doc/description.tex
removed addValueToTable + formating

@@
 \begin{align*}
-& \circ \in Application \times AppT \rightarrow State \\
-& \bullet \in Database \times DbT \rightarrow State
+& \circ \in State \times AppT \rightarrow State \\
+& \bullet \in State \times DbT \rightarrow State
 \end{align*}
@@
 \begin{align*}
-& a \circ e = \begin{cases}
+& state(a) \circ e = \begin{cases}
  	\inference{!\;consistent(a)}{\perp} \\ \\
@@
 \begin{align*}
-&  d \bullet e = \begin{cases}
+&  state(d) \bullet e = \begin{cases}
  	\inference{!\;consistent(d)}{\perp} \\ \\
@@
  		\inference{x = \varnothing}{y}\\
- 		\inference{|x| > 1}{f(head(x), y) \cup f(tail(x), y}\\\\
+ 		\inference{|x| > 1}{f(head(x), y) \cup f(tail(x), y)}\\\\
  		\inference{|x| = 1}{effect}\\\\
@@
 & \;\;\;\;\;\; =\begin{cases}
-		\inference{ \forall t_1, t_2 \in tables(d) : name(t_1) \neq name(t_2) \\ \wedge \forall c_1, c_2 \in columns(t_1) : name(c_1) \neq name(c_2) \\ \wedge \forall f_1, f_2 \in foreignKeys(t_1) : name(f_1) \neq name (f_2) \\ \wedge \forall f \in foreignKeys(t_1) : \exists t \in tables(d) \wedge name(t) = reference(f) \\ \wedge \forall r \in data(d): ( \exists t \in tables(d) : reference(r) = name(t) \\ \wedge \forall p \in pairs(r) :\\ \exists u \in columns(t) \cup foreignKeys(t) \cup primaryKey(t) : \\ reference(p) = name(u) \\ \wedge \forall e \in columns(t) \cup foreignKeys(t) \\ \wedge NOTNULL \in constrain(e) : \exists q \in pairs(r) \\ \wedge name(e) = reference(q)) \\ \wedge \forall g \in columns(t) \cup foreignKeys(t) \\ \wedge UNIQUE \in constrain(g) : \\ \forall u_1, u_2 \in rows(t) : \\ value(pairByName(name(g), u_1) \neq value(pairByName(name(g), u_2) }{true}
+		\inference{ \forall t_1, t_2 \in tables(d) : name(t_1) \neq name(t_2) \\ \wedge \forall c_1, c_2 \in columns(t_1) : name(c_1) \neq name(c_2) \\ \wedge \forall f_1, f_2 \in foreignKeys(t_1) : name(f_1) \neq name (f_2) \\ \wedge \forall f \in foreignKeys(t_1) : \\(\exists t \in tables(d) \wedge name(t) = reference(f)) \\ \wedge \forall r \in data(d): \\( \exists t \in tables(d) : reference(r) = name(t)) \\ \wedge \forall p \in pairs(r) :\\ \exists u \in columns(t) \cup foreignKeys(t) \cup primaryKey(t) : \\ reference(p) = name(u) \\ \wedge \forall e \in columns(t) \cup foreignKeys(t) \\ \wedge NOTNULL \in constrain(e) : \exists q \in pairs(r) \\ \wedge name(e) = reference(q)) \\ \wedge \forall g \in columns(t) \cup foreignKeys(t) \\ \wedge UNIQUE \in constrain(g) : \\ \forall u_1, u_2 \in rows(t) : \\ (value(select(name(g), u_1) \neq \\ \neq value(select(name(g), u_2)) }{true}
  	\\ \\
@@
  		\inference{|p| = 1 \wedge cardinality(p) = 1 \wedge !\;mandatory(p)
- 			\\ \wedge \exists t : t = table(name(owningClass(p)), d) 
+ 			\\ \wedge \exists t = table(name(owningClass(p)), d) 
 		 }{\begin{gathered}
-	  		column(name(p), \rho_t(type(p)), \varnothing )
+	  		col(name(p), \rho_t(type(p)), \varnothing )
 		\end{gathered}} \\ \\
   		\inference{|p| = 1 \wedge cardinality(p) = 1 \wedge mandatory(p)
-			 \\ \wedge \exists t : t = table(name(owningClass(p)), d)
+			 \\ \wedge \exists t = table(name(owningClass(p)), d)
 			\\ \wedge ! instanciated(owningClass(p))
 		}{\begin{gathered}
- 	 		column(name(p), \rho_t(type(p)), NOTNULL ) \\ 
+ 	 		col(name(p), \rho_t(type(p)), NOTNULL ) \\ 
 		\end{gathered}}
@@
      	\inference{|p| = 1 \wedge cardinality(p) > 1 \wedge !\;mandatory(p) \\ 
-     		\wedge 	\exists t : t = table(name(owningClass(p)), d)
+     		\wedge 	\exists t = table(name(owningClass(p)), d)
      		}{\begin{gathered}
-     			table(name(t)+name(p), primaryKey("Id"), \\ column("value", \rho_t(type(p)), \varnothing), \\ foreignKey(name(p), 			name(owningClass(p)), NOTNULL) \\
+     			tab(name(t)+name(p), primaryKey("Id"), \\ col("value", \rho_t(type(p)), \varnothing), \\ fk(name(p), 			name(owningClass(p)), NOTNULL) \\
     	\end{gathered}}\\ \\
      	\inference{|p| = 1 \wedge  cardinality(p) > 1 \wedge mandatory(p) \\ 
-     		\wedge \exists t : t = table(name(owningClass(p)), d) 
+     		\wedge \exists t = table(name(owningClass(p)), d) 
      		}{\begin{gathered}  
-     			table(name(t)+name(p), primaryKey("Id"), \\ 
-     			column("value", \rho_t(type(p)), NOTNULL), \\ 			foreignKey(name(p), name(owningClass(p)), NOTNULL)
+     			tab(name(t)+name(p), primaryKey("Id"), \\ 
+     			col("value", \rho_t(type(p)), NOTNULL), \\ 			fk(name(p), name(owningClass(p)), NOTNULL)
     	 \end{gathered}}
@@
 		%\inference{|a| > 1}{\rho_{a}(head(a)) \cup \rho_{a}(tail(a))}\\\\
-		\inference{|a| = 1 \wedge  startCardinality(a) = 0 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)
+		\inference{|a| = 1 \wedge  startCardinality(a) = 0 \\ \wedge \exists t \in table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)
 		}{
@@
   \\ \\
- 	 \inference{|a| = 1 \wedge  startCardinality(a) = 1 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)
+ 	 \inference{|a| = 1 \wedge  startCardinality(a) = 1 \\ \wedge \exists t = table(name(owningClass(p)), d) \\ \wedge \exists u \in tables(d) : name(u = reference(a)
  	 }{ 
@@
   		}{\begin{gathered}  
-		 table(name(t)+name(a), pk("Id"), \varnothing, \\ fk(name(owningClass(a)), name(owningClass(a)), NOTNULL) \\ fk(reference(a), reference(a), NOTNULL)) 
+		 table(name(t)+name(a), pk("Id"), \varnothing, \\ fk(name(owningClass(a)), \\ name(owningClass(a)), NOTNULL) \\ fk(reference(a), reference(a), NOTNULL)) 
   		\end{gathered}}  
@@
 & 	remProp(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p
-	}{properties(c) - \; p }\\
+	}{properties(c) \setminus p }\\
 &	l_c, l_p \in Label, a \in Application 
@@
 &	\mathbf{remAssoc} : Label \times Label \times Application \rightarrow Application \\
-&	remAssoc(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) - \; f} \\
+&	remAssoc(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) \setminus f} \\
 &	l_c, l_a \in Labels, a \in Application
@@
 &	remClass(l, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge !\: associated(c)
-	} {classes(a) - \; c }\\
+	} {classes(a) \setminus c }\\
 &	l \in Label,a \in Application
@@
 	}{\begin{gathered}
-		p.name = l_n 
+		p.name := l_n 
 	\end{gathered}}\\
@@
 	}{\begin{gathered}
-		b.name = l_n 
+		b.name := l_n 
 	\end{gathered}}
@@
 	}{\begin{gathered}
-		name(c) = l_n \wedge \forall e \in classes(a) \\ \wedge \forall (f \in associations(e) \wedge reference(f) = l_o ) :\\ f.reference = l_n 
+		name(c) = l_n \wedge \forall e \in classes(a) \\ \wedge \forall (f \in associations(e), reference(f) = l_o ) :\\ f.reference := l_n 
 	\end{gathered}}\\
@@
 
-\paragraph{Mapping of Instances} The set $M$ which contains all possible mappings between classes, has to be defined to allow several evolutionary operations.
+\paragraph{Pairs of Instances} The set $M$ which contains all possible mappings between instances of classes, has to be defined to allow several evolutionary operations.
 Each mapping represents an ordered set of pairs of instances.
 These pairs can represent one-to-one (or many) relationship or many-to-many relationship between instances.
+The set $M$ contains functions: $\rightarrow Row \times Row$.
 \begin{align*}
@@
 
-\paragraph{Comparison of tables} The need to compare two tables during the evolution results into the comparing operator $\cong$. The operator has to be defined for columns and foreign keys too.
-\begin{align*}
-&	\mathbf{\cong} : Table \times Table \rightarrow Boolean \\
-&	t_1 \cong t_2 = \begin{cases}
- 		\inference{|columns(t_1)| = |columns(t_2)| \\ \wedge |foreignKeys(t_1)| = |foreignKeys(t_2)| \\ \wedge \forall c \in columns(t_1) \exists e \in columns(t_2) : c = e \\ \wedge \forall f \in foreignKeys(t_1) \exists g \in foreignKeys(t_2) : f = g }{true}\\\\
+\paragraph{Comparison of tables} The need to compare two tables during the evolution results into the comparing function equals. The operator has to be defined for columns and foreign keys too.
+\begin{align*}
+& \mathbf{equals} : Table \times Table \rightarrow Boolean \\
+&	equals(t_1, t_2) = \begin{cases}
+ 		\inference{|columns(t_1)| = |columns(t_2)| \\ \wedge |foreignKeys(t_1)| = |foreignKeys(t_2)| \\ \wedge \forall c \in columns(t_1) \exists e \in columns(t_2) : c = e \\ \wedge \forall f \in foreignKeys(t_1) : \\ \exists g \in foreignKeys(t_2) : f = g }{true}\\\\
  false
@@
 \begin{align*}
-&	\mathbf{\cong} : Column \times Column \rightarrow Boolean \\
-&	c_1 \cong c_2 = \begin{cases}
+& \mathbf{equals} : Column \times Column \rightarrow Boolean \\
+&	equals(c_1, c_2) = \begin{cases}
  		\inference{name(c_1) = name(c_2) \\ \wedge type(c_1) = type(c_2) \\ \wedge constraints(c_1) = comstraints(c_2)}{true}
@@
 \begin{align*}
-&	\mathbf{\cong} : ForeignKey \times ForeignKey \rightarrow Boolean \\
-&	f_1 \cong f_2 = \begin{cases}
+&	\mathbf{equals} : ForeignKey \times ForeignKey \rightarrow Boolean \\
+&	equals(f_1, f_2) = \begin{cases}
  		\inference{name(f_1) = name(f_2) \\ \wedge reference(f_1) = reference(f_2) \\ \wedge constraints(f_1) = comstraints(f_2)}{true}
@@
 
-\paragraph{Find pair} The function returns a pair corresponding to given column (or key) name in concrete row.
-\begin{align*}
-& \mathbf{findPair} : Label \times Row \rightarrow Pair \\
-&	findPair(l, r) = \begin{cases}
+\paragraph{Select a pair} The function returns a pair corresponding to given column (or key) name in concrete row.
+\begin{align*}
+& \mathbf{select} : Label \times Row \rightarrow Pair \\
+&	select(l, r) = \begin{cases}
 		\inference{\exists p \in pairs(r) : reference(p) = l}{p} \\ \\
@@
 		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c) \\ \wedge NOTNULL \not \in constraints(c) }{columns(t) \cup c}\\ \\
-		(\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c) \\ \wedge NOTNULL \in constraints(c) }{columns(t) \cup c})\\ \bullet addValueToTable(l, [name(c), defVal(c)])
+		\inference{|c| = 1 \wedge \exists t \in tables(d) \wedge name(t) = l \\ \wedge \forall f \in columns(t) : name(f) \neq name(c) \\ \wedge NOTNULL \in constraints(c)}{\begin{gathered}
+	 (columns(t) \cup c) \\ \bullet (\forall r \in instances(l, d): \\ insetValue(reference(r), keyValue(r), \\ [name(c), defVal(c)]) 
+	\end{gathered}}
 %		\\ \\ Inconsistent \;d 
@@
  		\inference{|p| = 1 \wedge 
- 		\\ \wedge \exist r \in instances(l, d) : value(findPair("Id", r)) = n
+ 		\\ \wedge \exist r \in instances(l, d) : value(select("Id", r)) = n
  		\\ \wedge \exists t \in tables(d) : name(t) = l, 
@@
 
-\paragraph{Add Value To Table} The function appends a pair to all rows of a table with given name.
-\begin{align*}
-&	\mathbf{addValueToTable} : Label \times Pair \times Database \rightarrow Database \\
-&	addValueToTable(l, p, d) = \\
-&= \;\;\;\;\;\;\begin{cases}
- 		\inference{\exists t \in tables(d) : name(t) = l}	{\begin{gathered}
-		d \bullet (\forall r \in instances(l, d) \\ : insetValue(reference(r), keyValue(r), p) \\ \bullet addValueToTable(tail(instances(l, d)), p, d))
-	\end{gathered}} 
- 	\end{cases}\\ 
-& 	l \in Label, p \in Pair,  d \in Database
-\end{align*}
-
 \paragraph{Copy Value} The function copy value from one column in a table to another column in the same or in another table - according to provided mapping.
@@
 & \;\;\;\;\;\; =	\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc}}{\begin{gathered}
-		\forall	p \in m_{l_s, l_t} :	d \bullet insertValue(name(second(p)), keyValue(second(p)), \\ [l_{tc},value(findPair(l_{sc}, first(p)))])
+		\forall	p \in m_{l_s, l_t} :	d \bullet insertValue(name(second(p)), keyValue(second(p)), \\ [l_{tc},value(select(l_{sc}, first(p)))])
 		\end{gathered}}\\
@@
  	%\inference{|m_{l_s, l_t}| > 1}{\begin{gathered} d \bullet copyRow(l_s, l_t, head(M_{l_s, l_t})) \\ \bullet copyRow(l_s, l_t, tail(m_{l_s, l_t})) \end{gathered}}
- 	\inference{|m_{l_s, l_t}| = 1 \\ \wedge \exists t_s, t_t \in tables(d) : name(t_s) = l_s, name(t_t) = l_t \\ \wedge t_s \cong t_t }{d \bullet insertRow(row(l, pairs(first(m_{l_s, l_t}))), d)} \\\\
+ 	\inference{|m_{l_s, l_t}| = 1 \\ \wedge \exists t_s, t_t \in tables(d) : name(t_s) = l_s, name(t_t) = l_t \\ \wedge equals(t_s, t_t) }{d \bullet insertRow(row(l, pairs(first(m_{l_s, l_t}))), d)} \\\\
 %  \end{cases} 
@@
 \bullet (\forall r \in instances(t_1, database(s)) \cup instances(t_2, database(s)) :  \\ n =  row(l_n, next(seq(key(l_n))), pairs(r)), insertRow(l_n, n) \\ 
-\bullet (\forall k \in data(database(s)), f \in foreignKeys(k) \\ \wedge reference(f) = reference(r) \\ \wedge value(findPair(name(f), k)) = pkValue(r) : \\value(findPair(name(f), k)) = pkValue(n))) \\  \bullet dropTableAndData(l_1) \bullet dropTableAndData(l_2)
+\bullet (\forall k \in data(database(s)), f \in foreignKeys(k) \\ \wedge reference(f) = reference(r) \\ \wedge value(select(name(f), k)) = pkValue(r) : \\value(select(name(f), k)) = pkValue(n))) \\  \bullet dropTableAndData(l_1) \bullet dropTableAndData(l_2)
  \end{gathered}} \\

### 2013-02-02 doc/description.tex
formating

@@
 \begin{align*}
-&	\mathbf{copyValue} : Label \times Label \times Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
-&	copyValue(l_s, l_t, l_{sc}, l_{tc} o_{l_s, l_t}, d) = \\
+&	\mathbf{copyValue} : Label \times Label \times Label \times Label \times M \times State \rightarrow \\ & \;\;\;\;\;\; \rightarrow Database \\
+&	copyValue(l_s, l_t, l_{sc}, l_{tc}, m_{l_s, l_t}, d) = \\
 & \;\;\;\;\;\; =	\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc}}{\begin{gathered}
-		\forall	p \in m_{l_s, l_t} :	d \bullet insertValue(name(second(p)), keyValue(second(p)), \\ [l_{tc},value(select(l_{sc}, first(p)))])
+		\forall	p \in m_{l_s, l_t} :	\\ state(d) \bullet insertValue(name(second(p)), \\ pkValue(second(p)), [l_{tc},value(select(l_{sc}, first(p)))])
 		\end{gathered}}\\
-& 	l_s, l_t, l_{sc}, l_{tc} \in Label, d \in Database
+& 	l_s, l_t, l_{sc}, l_{tc} \in Label, d \in Database, m_{l_s, l_t} \in M
 \end{align*}
@@
     		\inference{cardinality(p) \leq 1}{database(s) \bullet addColumn(l, \rho(p,d), d)} \\ \\
-	    \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
-   database(s) \bullet addTable(\rho(p,d), d) \\ \bullet \forall r \in row(owningClass(p)) :\\ insertRow(row(name(p), ["value", defaultValue(p)], \\ [name(p), value(primaryKey(r))])) 
+	    \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) > 1}{\begin{gathered}
+   database(s) \bullet addTable(\rho(p,d), d) \\ \bullet \forall r \in row(owningClass(p)) :\\ insertRow(row(name(p), ["value", defaultValue(p)], \\ [name(p), pkValue(r)])) 
     \end{gathered} } 
@@
 \begin{align*}
-& \mathbf{addAsociation} : Label \times Association \times (Row \times Row) \times State \rightarrow State \\
+& \mathbf{addAsociation} : Label \times Association \times M \times State \rightarrow State \\
 & \Psi(s, addAsociation(l, a, m_{l_1, l_2})) = application(s) \circ addAssoc(l, a) \\
@@
 	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l\\ 		\wedge name(t_t) = reference(a) \\ \wedge \forall f \in foreignKeys(t_t) \cup columns(t_t) \\: name(f) \neq name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
-		database(s) \\ \bullet addForeignKey(name(t_t), \rho(a,d), d) \\ \bullet \forall r \in m_{l_1, l_2} : \\
-		insertValue(name(t_t), value(primaryKey(second(r))), \\ [name(a), value(primaryKey(first(r)))])
+		database(s) \\ \bullet addForeignKey(name(t_t), \rho(a,d), d) \\ \bullet (\forall r \in m_{l_1, l_2} : 
+		insertValue(\\ name(t_t), pkValue(second(r)), \\ [name(a), pkValue(first(r))]))
 	\end{gathered}}
@@
 	\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = l \\ \wedge name(t_t) = reference(a) \\ \wedge startCardinality(a) > 1}{\begin{gathered} 
-	database(s) \bullet addTable(\rho(a,d),d) \\ \bullet \forall r \in m_{l, reference(a)} :
-	\\ insertRow(\\ row(name(owningClass(a)) + name(a), \\ [name(t_s), value(primaryKey(first(r)))]\\ [name(t_t), value(primaryKey(second(r)))]))
+	database(s) \bullet addTable(\rho(a,d),d) \\ \bullet (\forall r \in m_{l, l_2} :
+	 insertRow(\\ row(name(owningClass(a)) + name(a), \\ [name(t_s), pkValue(first(r))]\\ [name(t_t), pkValue(second(r))])))
 	\end{gathered}} 
 \end{cases}\\
-& s \in Software, l_1, l_2 \in Label, m_{l, reference(a)} \wedge l_2 = reference(a)
+& s \in Software, l, l_2 \in Label, m_{l, l_2} \in M \wedge l_2 = reference(a)
 \end{align*}
@@
 & \mathbf{addClass} : Class \times State \rightarrow State \\
-& \Psi(s, addClass(c)) = application(s) \circ addClass(c) \\
-& \Phi(s, addClass(c)) = database(s)  \bullet addTable(\rho(c,d)) \\ 
+& \Psi(s, addClass(c)) = state(application(s)) \circ addClass(c) \\
+& \Phi(s, addClass(c)) = state(database(s))  \bullet addTable(\rho(c,d)) \\ 
 & \bullet addColumn(name(c), \{\forall r \in singleProperties(c) | \rho(r,d)\}, database(s)) \\
-& \bullet addTable(name(c), \{\forall r \in collections(c) | \rho(r,database(s))\})  \\
+& \bullet addTable(name(c), \{\forall r \in collections(c) | \rho(r,database(s))\})  \\\\
 & s \in Software, c \in Class
@@
 \begin{align*}
-&	\mathbf{RemoveClass}: Label \times Software \rightarrow State  \\
-&	\Psi RemoveClass(l, s) = applications(s) \circ remClass(l) \\
-&	\Phi RemoveClass(l, s) = database(s) \bullet dropTable(l) \\
+&	\mathbf{RemoveClass}: Label \times State \rightarrow State  \\
+&	\Psi RemoveClass(l, s) = state(applications(s)) \circ remClass(l) \\
+&	\Phi RemoveClass(l, s) = state(database(s)) \bullet dropTable(l) \\\\
 &	l \in Label, s \in Software 
@@
 &	\mathbf{RemoveProperty} : Label \times Label \times State \rightarrow State \\
-&	\Psi RemoveProperty(l, l_p, s) =  applications(s) \circ  remProp(l, l_p) \\
+&	\Psi RemoveProperty(l, l_p, s) =  state(applications(s)) \circ  remProp(l, l_p) \\
 &	\Phi RemoveProperty(l, l_p, s) = \\
@@
 		 \inference{\exists c \in classes(application(s)) : name(c) = l\\ 
-		\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) \leq 1}{database(s) \bullet dropColumn(l, l_p)}  \\\\
+		\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) \leq 1}{state(database(s)) \bullet dropColumn(l, l_p)}  \\\\
 		 \inference{\exists c \in classes(application(s)) : name(c) = l\\ 
-		\wedge \exists p \in properties(c)  : name(p) = l \\ \wedge cardinality(p)>1}{database(s) \bullet dropTable(name(owningClass(p)+l)}
- 	\end{cases}\\
-&	l, l_p \in Label, s \in State
+		\wedge \exists p \in properties(c)  : name(p) = l \\ \wedge cardinality(p)>1}{state(database(s)) \bullet dropTable(name(owningClass(p)+l)}
+ 	\end{cases}\\\\
+&	l, l_p \in Label, s \in Software
 \end{align*}
@@
 &	\mathbf{RemovePropertyAndData} : Label \times Label* \times State \rightarrow State \\
-&	\Phi(RemovePropertyAndData(L_c, ps), s) = application(s)\\ 
+&	\Phi(RemovePropertyAndData(L_c, ps), s) = state(application(s)) \\ 
 &  \circ remProperty(l_s, ps) \\
@@
 		%\inference{|ps| > 1}{\begin{gathered} database(s) \\ \bullet RemovePropertyAndData(L_c, head(ps)) \\ \bullet RemovePropertyAndData(L_c, tail(ps)) 	\end{gathered}}\\\\
- 		\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) \leq 1}{database(s) \bullet DropColumnAndData(l_c, ps)}\\\\
- 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) > 1}{database(s) \bullet RemoveTableAndData(name(owningClass(p))+l_c, ps)}
+ 		\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) \leq 1}{state(database(s)) \bullet DropColumnAndData(l_c, ps)}\\\\
+ 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) > 1}{state(database(s)) \bullet RemoveTableAndData(name(owningClass(p))+l_c, ps)}
  \end{cases}
@@
 &    \mathbf{RemoveAssociation} : Label \times Labels \times State \rightarrow State \\ 
-&    \Psi RemoveAssociation(l, l_a, s) = application(s) \circ remAss(l, l_a) \\
+&    \Psi RemoveAssociation(l, l_a, s) = state(application(s)) \circ remAss(l, l_a) \\
 &    \Phi RemoveAssociation(l, l_a, s) = \\
@@
      	\wedge \exists a \in associaions(c) : name(a) = l_a  \\
-     	\wedge startCardinality(a) \leq 1 \\}{database(s) \bullet dropForeignKey(l_a, reference(a))} \\\\
+     	\wedge startCardinality(a) \leq 1 \\}{state(database(s)) \bullet dropForeignKey(l_a, reference(a))} \\\\
      	\inference{\exists c \in classes(application(s)) : name(c) = l \\
      	\wedge \exists a \in associaions(c) : name(a) = l_a  \\
-     	\wedge startCardinality(a) > 1}{database(s) \bullet dropTable(name(owningClass(a))+l_a)} 
+     	\wedge startCardinality(a) > 1}{state(database(s)) \bullet dropTable(name(owningClass(a))+l_a)} 
      \end{cases}\\
@@
 &	\mathbf{RemoveAssociationAndData} : Label \times Label* \times State \rightarrow State \\
-&	\Phi(RemoveAssociationAndData(L_c, ps), s) = application(s)\\ 
+&	\Phi(RemoveAssociationAndData(L_c, ps), s) = state(application(s))\\ 
 & \circ remAssociation(l_s, ps) \\
@@
  		\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists a \in associations(c) : name(a) = l \\ \wedge startCardinality(p) \leq 1}{\begin{gathered}
-	database(s) \\ \bullet RemoveForeignKeyAndData(reference(a), ps)
+	state(database(s)) \\ \bullet RemoveForeignKeyAndData(reference(a), ps)
 \end{gathered}}
@@
  	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists a \in associations(c) : name(a) = l \\ \wedge startCardinality(a) > 1}{\begin{gathered}
-	database(s) \\\bullet RemoveTableAndData(l_c + reference(a), ps)
+	state(database(s)) \\\bullet RemoveTableAndData(l_c + reference(a), ps)
 \end{gathered}}
@@
 & \mathbf{RenameClass} : Label \times Label \times State \rightarrow State \\ 
-& \Psi(RenameClass(l_o, l_n), s) = application(s) \circ renClass(l_o, l_n) \\
-& \Phi(RenameClass(l_o, l_n), s) = database(s) \bullet alterTable(l_o, l_n) \\ \\
+& \Psi(RenameClass(l_o, l_n), s) = state(application(s)) \circ renClass(l_o, l_n) \\
+& \Phi(RenameClass(l_o, l_n), s) = state(database(s)) \bullet alterTable(l_o, l_n) \\ \\
 \end{align*}
@@
 &	\mathbf{RenameProperty} : Label \times Label \times Label \times State \rightarrow State \\ \\
-& \Psi(RenameProperty(l_c, l_o, l_n, s)) = application(s) \circ renProp(l_c, l_o, l_n) \\
+& \Psi(RenameProperty(l_c, l_o, l_n, s)) = state(application(s)) \circ renProp(l_c, l_o, l_n) \\
 & \Phi(RenameProperty(l_c, l_o, l_n, s)) = \\
@@
 		\wedge \exists p \in properties(c) : name(p) = l_o \\
- 	cardinality(p) \leq 1}{database(s) \bullet alterColumn(l_c, l_o, l_n))} \\\\
+ 	cardinality(p) \leq 1}{state(database(s)) \bullet alterColumn(l_c, l_o, l_n))} \\\\
  	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 
@@
  	cardinality(p) > 1}{\begin{gathered}
-		database(s)\\ \bullet alterTable(name(owningClass(p)+l_o, \\l_n, database(s))
+		state(database(s)) \\ \bullet alterTable(name(owningClass(p)+l_o, \\l_n, database(s))
 \end{gathered}}
@@
 & \mathbf{RenameAssociation} : Label \times Label \times Label \times State \rightarrow State \\
-& \Psi(RenameAssocaiation(l_c, l_o, l_n), s) = application(s) \circ renAssoc(l_c, l_o, l_n)\\
+& \Psi(RenameAssocaiation(l_c, l_o, l_n), s) = state(application(s)) \\ &\;\;\;\;\;\;  \circ renAssoc(l_c, l_o, l_n)\\
 & \Phi(RenameAssocaiation(l_c, l_o, l_n), s) = \\
@@
 		\wedge \exists a \in associations(c) : name(a) = l_o 
-	\\ \wedge startCardinality(a) > 1 }{\begin{gathered}
-		database(s) \\ \bullet alterTable(name(owningClass(a)) + l_o, l_n)
+	\\ \wedge startCardinality(a) \leq 1 }{\begin{gathered}
+state(database(s)) \\ \bullet alterForeignKey(endClass(a) ,l_c, l_o, l_n)
 \end{gathered}}\\ \\
@@
 		\wedge \exists a \in associations(c) : name(a) = l_o 
-	\\ \wedge startCardinality(a) \leq 1 }{\begin{gathered}
-database(s) \\ \bullet alterForeignKey(endClass(a) ,l_c, l_o, l_n)
-\end{gathered}}
+	\\ \wedge startCardinality(a) > 1 }{\begin{gathered}
+		state(database(s)) \\ \bullet alterTable(name(owningClass(a)) + l_o, l_n)
+\end{gathered}}\\ 
  \end{cases} \\
@@
 &	\Psi(CopyPropertyStructure(l_s, l_t, l_p, s) = \\
-& \;\;\;\;\;\; =\inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)} 
+& \;\;\;\;\;\; =\inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{state(application(s)) \circ addProp(l_t, p)} 
 	\\\\
@@
     \inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) \leq 1 }{\begin{gathered}
-			database(s) \bullet copyColumn(l_s, l_t, l_p) 			\end{gathered}} \\ \\
+			state(database(s)) \bullet copyColumn(l_s, l_t, l_p) 			\end{gathered}} \\ \\
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) > 1}{\begin{gathered}
-			database(s) \bullet copyTable(l_s + l_p, l_t + l_p) 			\end{gathered}}
+			state(database(s)) \bullet copyTable(l_s + l_p, l_t + l_p) 			\end{gathered}}
  \end{cases}\\
@@
 &	\mathbf{CopyProperty} : Label \times Label \times Label \times (Row \times Row) \times State \rightarrow State \\
-&	\Psi(CopyProperty(l_s, l_t, l_p, m_{l_s, l_t}), s) = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{application(s) \circ addProp(l_t, p)}
+&	\Psi(CopyProperty(l_s, l_t, l_p, m_{l_s, l_t}), s) = \\
+& \;\;\;\;\;\; = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{state(application(s)) \circ addProp(l_t, p)} 
 	\\\\
-&	\Phi(CopyProperty(l_s, l_t, l_p, m_{l_s, l_t}), s) = \begin{cases}
+&	\Phi(CopyProperty(l_s, l_t, l_p, m_{l_s, l_t}), s) = \\
+& \;\;\;\;\;\; = \begin{cases}
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) \leq 1 }{\begin{gathered}
-			database(s) \bullet copyColumn(l_s, l_t, l_p) \\ \bullet copyValue(l_s, l_t, l_p, l_p, m_{l_s, l_t}) 
+			state(database(s)) \bullet copyColumn(l_s, l_t, l_p) \\ \bullet copyValue(l_s, l_t, l_p, l_p, m_{l_s, l_t}) 
 			\end{gathered}} \\ \\
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) > 1}{\begin{gathered}
-			database(s) \bullet copyTable(l_s + l_p, l_t + l_p) \\ \bullet copyRow(l_s + l_p, l_t + l_p, m_{l_s, l_t}) 
+			state(database(s)) \bullet copyTable(l_s + l_p, l_t + l_p) \\ \bullet copyRow(l_s + l_p, l_t + l_p, m_{l_s, l_t}) 
 			\end{gathered}}
- \end{cases}\\
+ \end{cases}\\\\
 & l_s, l_t, l_p \in Label,  m_{l_s, l_t}) \in M, s \in State
@@
 \begin{align*}
-&	\mathbf{MoveProperty} : Label \times Label \times Label* \times (Row \times Row) \times State \rightarrow State \\
-&	s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) =
+&	\mathbf{MoveProperty} : Label \times Label \times Label* \times M \times State \rightarrow State \\
+&	s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) = \\
 % \begin{cases}
@@
 %	\inference{|ps| > 1}{\begin{gathered} 	s \oplus MoveProperty(l_s, l_t, head(ps), m_{l_s, l_t}) \\ \oplus MoveProperty(l_s, l_t, tail(ps), m_{l_s, l_t}) \end{gathered}}
- 	\inference{|ps| = 1 }{\begin{gathered} s \oplus CopyProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemovePropertyAndData(l_s, ps)
+& \;\;\;\;\;\; = 	\inference{|ps| = 1 }{\begin{gathered} s \oplus CopyProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemovePropertyAndData(l_s, ps)
 \end{gathered}}
 % \end{cases}
-	\\
+	\\ \\
 &	l_s, l_t \in Label, ps \in Label*,  m_{l_s, l_t} \in M, s \in State
@@
 \begin{align*}
-&	\mathbf{Extract} : Label \times Label \times Label \times (Row \times Row) \times State \rightarrow State \\
-&	s \oplus Extract(l_s, l_n, ps, m_{l_s, l_s}) = \begin{cases}
+&	\mathbf{Extract} : Label \times Label \times Label \times M \times State \rightarrow State \\
+&	s \oplus Extract(l_s, l_n, ps, m_{l_s, l_s}) = %\begin{cases}
 	\inference{|ps| = 1}{\begin{gathered}
@@
 	% \inference{|ps| = \varnothing}{state(s)}	 
- \end{cases}
-	\\
-&	l_s, l_n \in Label, ps ={ x | x  \in Label},  s \in State 
+ %\end{cases}
+\\\\
+&	l_s, l_n \in Label, ps ={ x | x  \in Label},  s \in State, m_{l_s, l_s} \in M
 \end{align*}
@@
 \begin{align*}
-&	\mathbf{Inline} : Label \times Label \times (Row \times Row) \times State \rightarrow State \\
+&	\mathbf{Inline} : Label \times Label \times M \times State \rightarrow State \\
 &	s \oplus Inline(l_s, l_t, m_{l_s, l_t}) = \\
@@
 \wedge \exists c_t \in classes(application(s)) : name(c_t) = l_t \\
-\wedge \forall p_s \in properties(c_s), p_t \in properties(c_t) : name(ps) \neq name(p_t) \\
+\wedge \forall p_s \in properties(c_s), p_t \in properties(c_t) :\\  name(ps) \neq name(p_t) \\
 \wedge !\:associated(l_s, s)   }{\begin{gathered}
@@
 \end{gathered}
-}\\
+}\\\\
 & l_s, l_t \in Label, m_{l_s, l_t} \in M, s \in State, \\ 
-& ps = \{x | x \in \forall c \in classes(application(s)) : \forall p \in properries(c) : name(p))\}
+& ps = \{x | x \in (\forall c \in classes(application(s)) : \\ 
+& \;\;\;\;\;\; \forall p \in properries(c) : name(p)))\}
 \end{align*}
@@
 &	\mathbf{Merge} : Label \times Label \times Label \times State \rightarrow State \\
-&	\Phi(s, Merge(l_1, l_2, l_n) = \inference{
+&	\Phi(s, Merge(l_1, l_2, l_n) = \\
+& \;\;\;\;\;\; = \inference{
 		\exists c_1, c_2 \in classes(application(s)) \wedge 
 		equals(c_1, c_2)}{\begin{gathered}
-			application(s) \\ \circ addClass(cl(ln, properties(c_1), associations(c_1)) \\
+			state(application(s)) \\ \circ addClass(cl(ln, properties(c_1), associations(c_1)) \\
 			\circ (\forall c \in classes (application(s)), a \in associations(c) \\
@@
  \\\\
-&	\Psi(s, merge(l_1, l_2, l_n)) = \inference{\exists t_1, t_2 \in tables(database(s)) : \\ name(t_1) = l_1 \wedge name(t_2) = l_2 \\ \wedge columns(t_1) = columns(t_2)\\
+&	\Psi(s, merge(l_1, l_2, l_n)) = \\
+& \;\;\;\;\;\; = \inference{\exists t_1, t_2 \in tables(database(s)) : \\ name(t_1) = l_1 \wedge name(t_2) = l_2 \\ \wedge columns(t_1) = columns(t_2)\\
 \forall f \in (foreignKeys(t_1) \cup foreignKey(t_2)) \\ \setminus (foreignKey(t_1) \cap foreignKey(t_2)) \\ : NOTNULL \not \in constraints(f)  }{\begin{gathered}
-	 database(s) \\ \bullet addTable(l_n, pk("Id", s_1), columns(t_1), foreignKeys(t_1)) \\ 
+	state(database(s)) \\ \bullet addTable(l_n, pk("Id", s_1), columns(t_1), foreignKeys(t_1)) \\ 
 \bullet (\forall r \in instances(t_1, database(s)) \cup instances(t_2, database(s)) :  \\ n =  row(l_n, next(seq(key(l_n))), pairs(r)), insertRow(l_n, n) \\ 
 \bullet (\forall k \in data(database(s)), f \in foreignKeys(k) \\ \wedge reference(f) = reference(r) \\ \wedge value(select(name(f), k)) = pkValue(r) : \\value(select(name(f), k)) = pkValue(n))) \\  \bullet dropTableAndData(l_1) \bullet dropTableAndData(l_2)
- \end{gathered}} \\
+ \end{gathered}} \\\\
  & l_1, l_2, l_n \in Label, s \in State

### 2013-02-02 doc/description.tex
new images

@@
 \begin{figure}
-\begin{subfigure}[b]{0.5\textwidth}
+\begin{subfigure}[b]{0.45\textwidth}
 	\includegraphics[width=\textwidth]{./images/case_app_1}
@@
 \end{subfigure}
-\begin{subfigure}[b]{0.5\textwidth}
+\quad
+\begin{subfigure}[b]{0.45\textwidth}
 	\includegraphics[width=\textwidth]{./images/case_db_1}
@@
 \begin{align*}
-S_1 & = state(S) \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip") \\
-& \oplus AddAssociation("Person", assoc("address", "Address\_tmp1", 1, 1),\\ 
-& \{(x,y)| x \in instances("Person", database(S)), \\
-& y \in instances("Address\_tmp1", database(S)) : pkVal(x) = pkVal(y) \}) \\
-& \oplus Extract("LegalPerson", "Address\_tmp2", "street", "city", "zip")  \\
-& \oplus AddAssociation("LegalPerson", \\ 
-& assoc("residentialAddress", "Address\_tmp2", 1, 1),\\ 
-& \{(x,y)| x \in instances("LegalPerson", database(S)), \\
-& y \in instances("Address\_tmp2", database(S)) : pkVal(x) = pkVal(y) \}) \\
+S_1 & = state(S) \\
+& \;\;\;\;\;\; \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip") \\
+& \;\;\;\;\;\; \oplus AddAssociation("Person", \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp1", 1, 1),\\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("Person", database(S)), \\
+& \;\;\;\;\;\; \;\;\;\;\;\; y \in instances("Address\_tmp1", database(S)) : \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \}) \\
+&\;\;\;\;\;\; \oplus Extract("LegalPerson", "Address\_tmp2", "street", \\
+& \;\;\;\;\;\;\;\;\;\;\;\; "city", "zip")  \\
+& \;\;\;\;\;\; \oplus AddAssociation("LegalPerson", \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; assoc("residentialAddress", "Address\_tmp2", 1, 1),\\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("LegalPerson", database(S)), \\
+& \;\;\;\;\;\;\;\;\;\;\;\; y \in instances("Address\_tmp2", database(S)) : \\
+& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \}) \\
 \end{align*}
@@
 \begin{figure}
-\begin{subfigure}[b]{0.5\textwidth}
+\begin{subfigure}[b]{0.45\textwidth}
 	\includegraphics[width=\textwidth]{./images/case_app_5}
@@
 \end{subfigure}
-\begin{subfigure}[b]{0.5\textwidth}
+\begin{subfigure}[b]{0.45\textwidth}
 	\includegraphics[width=\textwidth]{./images/case_db_5}
@@
 %
-S_2 & \;\;\;\;\;\; =S_1 \oplus Merge("Address\_tmp1", "Address\_tmp2", "Address") \\
+S_2 &= S_1 \oplus Merge("Address\_tmp1", "Address\_tmp2", "Address") \\
 \end{align*}
@@
 \begin{align*}
-S_3 & \;\;\;\;\;\; =S_2  \oplus RenameAssociation("Person", "Address\_tmp1", "Address")\\
+S_3 &= S_2  \oplus RenameAssociation("Person", "Address\_tmp1", "Address")\\
 & \oplus RenameAssociation("LegalPerson", "Address\_tmp2". "Address")\\
@@
 \begin{figure}
-\begin{subfigure}[b]{0.5\textwidth}
+\begin{subfigure}[b]{0.45\textwidth}
+	\centering
 	\includegraphics[width=\textwidth]{./images/case_app_9}
@@
 \end{subfigure}
-\begin{subfigure}[b]{0.5\textwidth}
+\quad
+\begin{subfigure}[b]{0.45\textwidth}
+	\centering
 	\includegraphics[width=\textwidth]{./images/case_db_8}

### 2013-02-02 doc/description.tex
equals for classes, associations and properties

@@
 
+\paragraph{Comparison of Structures} During the evolution there occur situations when the structure of two classes has to be compared.
+\begin{align*}
+& \mathbf{equals} : Class \times Class \rightarrow Boolean \\
+& equals(c_1, c_2) = \begin{cases}
+ 	\inference{|properties(c_1)| = |properties(c_2)| \\ \wedge
+ 	\forall p \in properties(c_1) \exists r \in properties(c_2): \\ equals(p, r) \\
+ 	|associations(c_1)| = |associations(c_2)| \\
+ 	\wedge \forall a \in associations(c_1) \exists b \in asccosiations(c_2): \\ equals(a,b) \\ }{true}\\\\
+ false
+ \end{cases}\\\\
+& c_1, c_2 \in Class
+\end{align*}
+\begin{align*}
+& \mathbf{equals} : Property \times Property \rightarrow Boolean \\
+& equals(p_1, p_2) = \begin{cases}
+	\inference{name(p_1) = name(p_2) \\
+	\wedge mandatory(p_1) = mandatory(p_2) \\
+	\wedge cardinality(p_1) = cardinality(p_2)}{true}
+ \\\\ false
+ \end{cases}\\\\
+& p_1, p_2 \in Property
+\end{align*}
+\begin{align*}
+& \mathbf{equals} : Association \times Association \rightarrow Boolean \\
+& equals(a_1, a_2) = \begin{cases}
+	\inference{name(a_1) = name(a_2)\\
+	\wedge reference(a_1) = reference(a_2) \\
+	\wedge startCardinality(a_1) = startCardinality(a_2) \\
+	\wedge endCardinaility(a_1) = endCardinality(a_2)
+	}{true}
+ \\\\ false
+ \end{cases}\\\\
+& a_1, a_2 \in Association
+\end{align*}
+
 \paragraph{Add Property} The name collision only has to be verified, before adding a property into an application structure.
@@
 \paragraph{Merge of Classes} There occur situations when there is a need to merge two classes and the inlining transformation cannot be used because of name collisions among properties in such a case the $merge$ transformation can be used.
-It can be used in case structure of both classes (respectively tables) is equals.
+It can be used in case structure of both classes (respectively tables) are equal.
 The transformation creates new rows in the new table - the primary keys from origin instances are replaced by a new ones.

### 2013-02-02 doc/description.tex
mapping of names fixed

@@
 \paragraph{Mapping of Names} There can occur situations when a consistent application could be mapped into an inconsistent database because of name collisions or the name can collide with some internal database construct (which are not right now implemented in the model).
+Very simple mapping is provided - name of classes are mapped directly, whereas names of properties and associations are mapped as concatenation of name of theirs owning class and the name of the property (association), this prevents name collisions in the database schema.
+The mapping of names is implicit and it is not mentioned in this paper to increase readability.
 
@@
      		}{\begin{gathered}
-     			tab(name(t)+name(p), primaryKey("Id"), \\ col("value", \rho_t(type(p)), \varnothing), \\ fk(name(p), 			name(owningClass(p)), NOTNULL) \\
+     			tab(name(p), primaryKey("Id"), \\ col("value", \rho_t(type(p)), \varnothing), \\ fk(name(p), 			name(owningClass(p)), NOTNULL) \\
     	\end{gathered}}\\ \\
@@
      		}{\begin{gathered}  
-     			tab(name(t)+name(p), primaryKey("Id"), \\ 
+     			tab(name(p), primaryKey("Id"), \\ 
      			col("value", \rho_t(type(p)), NOTNULL), \\ 			fk(name(p), name(owningClass(p)), NOTNULL)
@@
 		}{
-			foreignKey(name(a), name(t),  \varnothing) 
+			fk(name(a), name(t),  \varnothing) 
 	 	}
   \\ \\
- 	 \inference{|a| = 1 \wedge  startCardinality(a) = 1 \\ \wedge \exists t = table(name(owningClass(p)), d) \\ \wedge \exists u \in tables(d) : name(u = reference(a)
+ 	 \inference{|a| = 1 \wedge  startCardinality(a) = 1 \\ \wedge \exists t = tab(name(owningClass(p)), d) \\ \wedge \exists u \in tables(d) : name(u = reference(a)
  	 }{ 
-		foreignKey(name(a), name(t),  NOTNULL)
+		fk(name(a), name(t),  NOTNULL)
 	}	 \end{cases}
@@
   		}{\begin{gathered}  
-		 table(name(t)+name(a), pk("Id"), \varnothing, \\ fk(name(owningClass(a)), \\ name(owningClass(a)), NOTNULL) \\ fk(reference(a), reference(a), NOTNULL)) 
+		 table(name(a), pk("Id"), \varnothing, \\ fk(name(owningClass(a)), \\ name(owningClass(a)), NOTNULL) \\ fk(reference(a), reference(a), NOTNULL)) 
   		\end{gathered}}  
@@
 	database(s) \bullet addTable(\rho(a,d),d) \\ \bullet (\forall r \in m_{l, l_2} :
-	 insertRow(\\ row(name(owningClass(a)) + name(a), \\ [name(t_s), pkValue(first(r))]\\ [name(t_t), pkValue(second(r))])))
+	 insertRow(\\ row(name(a), \\ [name(t_s), pkValue(first(r))]\\ [name(t_t), pkValue(second(r))])))
 	\end{gathered}} 
@@
 		 \inference{\exists c \in classes(application(s)) : name(c) = l\\ 
-		\wedge \exists p \in properties(c)  : name(p) = l \\ \wedge cardinality(p)>1}{state(database(s)) \bullet dropTable(name(owningClass(p)+l)}
+		\wedge \exists p \in properties(c)  : name(p) = l \\ \wedge cardinality(p)>1}{state(database(s)) \bullet dropTable(l)}
  	\end{cases}\\\\
@@
  		\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) \leq 1}{state(database(s)) \bullet DropColumnAndData(l_c, ps)}\\\\
- 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) > 1}{state(database(s)) \bullet RemoveTableAndData(name(owningClass(p))+l_c, ps)}
+ 	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists p \in properties(c) : name(p) = l \\ \wedge cardinality(p) > 1}{state(database(s)) \bullet DropTableAndData(l_c, ps)}
  \end{cases}
@@
      	\wedge \exists a \in associaions(c) : name(a) = l_a  \\
-     	\wedge startCardinality(a) > 1}{state(database(s)) \bullet dropTable(name(owningClass(a))+l_a)} 
+     	\wedge startCardinality(a) > 1}{state(database(s)) \bullet dropTable(l, l_a)} 
      \end{cases}\\
@@
  	\inference{\exists c \in classes(application(s)) : name(c) = l\\ 	\wedge \exists a \in associations(c) : name(a) = l \\ \wedge startCardinality(a) > 1}{\begin{gathered}
-	state(database(s)) \\\bullet RemoveTableAndData(l_c + reference(a), ps)
+	state(database(s)) \\\bullet RemoveTableAndData(l_c, ps)
 \end{gathered}}
@@
  	cardinality(p) > 1}{\begin{gathered}
-		state(database(s)) \\ \bullet alterTable(name(owningClass(p)+l_o, \\l_n, database(s))
+		state(database(s)) \\ \bullet alterTable(l_o, \\l_n, database(s))
 \end{gathered}}
@@
 \begin{align*}
-&	\mathbf{CopyPropertyStructure} : Label \times Label \times Label \times State \rightarrow State \\
+&	\mathbf{CopyPropertyStructure} : Label \times Label \times Label  \times Label \times State \rightarrow \\ &  \;\;\;\;\;\;  \;\;\;\;\;\;  \rightarrow State \\
 &	\Psi(CopyPropertyStructure(l_s, l_t, l_p, s) = \\
@@
 	\\\\
-&	\Phi(CopyPropertyStructure(l_s, l_t, l_p, s) = \\
+&	\Phi(CopyPropertyStructure(l_s, l_t, l_p, l_n, s) = \\
 & \;\;\;\;\;\; =\begin{cases}
     \inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) \leq 1 }{\begin{gathered}
-			state(database(s)) \bullet copyColumn(l_s, l_t, l_p) 			\end{gathered}} \\ \\
+			state(database(s)) \bullet copyColumn(l_s, l_t, l_p, l_n) 			\end{gathered}} \\ \\
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) > 1}{\begin{gathered}
-			state(database(s)) \bullet copyTable(l_s + l_p, l_t + l_p) 			\end{gathered}}
- \end{cases}\\
+			state(database(s)) \bullet copyTable(l_p, l_n) 			\end{gathered}}
+ \end{cases}\\\\
 & l_s, l_t, l_p \in Label, s \in State
@@
 \begin{align*}
-&	\mathbf{CopyProperty} : Label \times Label \times Label \times (Row \times Row) \times State \rightarrow State \\
+&	\mathbf{CopyProperty} : Label \times Label \times Label \times Label \times M \times State \rightarrow \\ &\;\;\;\;\;\; \;\;\;\;\;\;  \rightarrow State \\
 &	\Psi(CopyProperty(l_s, l_t, l_p, m_{l_s, l_t}), s) = \\
@@
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) > 1}{\begin{gathered}
-			state(database(s)) \bullet copyTable(l_s + l_p, l_t + l_p) \\ \bullet copyRow(l_s + l_p, l_t + l_p, m_{l_s, l_t}) 
+			state(database(s)) \bullet copyTable(l_s + l_p, l_t + l_p) \\ \bullet copyRow(l_p, l_n, m_{l_s, l_t}) 
 			\end{gathered}}

### 2013-02-02 doc/description.tex
copy property with names

@@
 \begin{align*}
-&	\mathbf{copyColumn} : Label \times Label \times Label \times Database \rightarrow Database \\
-&	copyColumn(l_s, l_t, l_c, d) = \begin{cases}
- 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_c}{d \bullet addColumn(l_t, c)}
-	\end{cases} \\
+&	\mathbf{copyColumn} : Label \times Label \times Label \times Label \times Database \rightarrow Database \\
+&	copyColumn(l_s, l_t, l_c, l_n, d) = \\
+& \;\;\;\;\;\;=% \begin{cases}
+ 		\inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \exists c \in columns(t) : name(c) = l_c \\ \wedge \exists e \in tables(d) : name(e) = l_t  \\ \wedge \forall q \in properties(e) : name(q) \neq l_n}{\begin{gathered}
+d \bullet addColumn(l_t, col(l_n, type(c), constraints(c))
+\end{gathered}} 
+	%\end{cases} 
+	\\\\
 &	l_s,l_t,l_c \in Label \wedge l_s \neq l_t,d \in Database
@@
 
-\paragraph{Copy Table} The transformation creates a copy of a table in database schema with a given name.
+\paragraph{Copy Table Structure} The transformation creates a copy of a table in database schema with a given name.
 The new name has to be different than the old one otherwise the transformation could not be processed.
@@
 &	\mathbf{copyTable} : Label \times Label \times Database \times Database \\
-&	copyTable(l_s, l_t, d) = \inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \forall e \in tables(d) : name(e) \neq l_t}{\begin{gathered} d \bullet addTable(tab(l_t, primaryKey(t), \\ 
+&	copyTable(l_s, l_n, d) = \inference{\exists t \in tables(d) : name(t) = l_s \\ \wedge \forall e \in tables(d) : name(e) \neq l_n}{\begin{gathered} d \bullet addTable(tab(l_n, primaryKey(t), \\ 
  columns(t), foreignKeys(t)))
@@
 There are two possible implementation of transformation for copying property, first implementation $CopyPropertyStructure$ copies only the structure of a property on the database level, the second implementation $CopyProperty $ copies both structure and data of the property.
+The copy is created with name ($l_n$), which can be the same or different than original name ($l_p$) possible name collision in database schema is eliminated due to the mapping of names between application and database schema.
 \begin{align*}
 &	\mathbf{CopyPropertyStructure} : Label \times Label \times Label  \times Label \times State \rightarrow \\ &  \;\;\;\;\;\;  \;\;\;\;\;\;  \rightarrow State \\
-&	\Psi(CopyPropertyStructure(l_s, l_t, l_p, s) = \\
+&	\Psi(CopyPropertyStructure(l_s, l_t, l_p, l_n, s) = \\
 & \;\;\;\;\;\; =\inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{state(application(s)) \circ addProp(l_t, p)} 
@@
 &	\mathbf{CopyProperty} : Label \times Label \times Label \times Label \times M \times State \rightarrow \\ &\;\;\;\;\;\; \;\;\;\;\;\;  \rightarrow State \\
-&	\Psi(CopyProperty(l_s, l_t, l_p, m_{l_s, l_t}), s) = \\
-& \;\;\;\;\;\; = \inference{\exists c \in classes(s) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p }{state(application(s)) \circ addProp(l_t, p)} 
+&	\Psi(CopyProperty(l_s, l_t, l_p, l_n, m_{l_s, l_t}), s) = \\
+& \;\;\;\;\;\; = \inference{\exists c \in classes(application(s)) : name(c) = l_s \\ \wedge \exists p \in properties(c) : name(p) = l_p \\ \wedge \exists c_2 \in classes(application(s)): name(c_2) = l_t \\
+\wedge \forall r \in properties(c_2): name(r) \neq l_n}{\begin{gathered}
+ state(application(s)) \circ addProp(l_t, \\ prop(l_n, type(p), defaultValue(p), \\ cardinality(p), mandatory(p)))
+\end{gathered}
+} 
 	\\\\
-&	\Phi(CopyProperty(l_s, l_t, l_p, m_{l_s, l_t}), s) = \\
+&	\Phi(CopyProperty(l_s, l_t, l_p, l_n, m_{l_s, l_t}), s) = \\
 & \;\;\;\;\;\; = \begin{cases}
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) \leq 1 }{\begin{gathered}
-			state(database(s)) \bullet copyColumn(l_s, l_t, l_p) \\ \bullet copyValue(l_s, l_t, l_p, l_p, m_{l_s, l_t}) 
+			state(database(s)) \bullet copyColumn(l_s, l_t, l_p, l_n) \\ \bullet copyValue(l_s, l_t, l_p, l_p, m_{l_s, l_t}) 
 			\end{gathered}} \\ \\
 		\inference{\exists p \in properties(c), c \in classes(s) : \\ name(p) = l_p \wedge cardinality(p) > 1}{\begin{gathered}
-			state(database(s)) \bullet copyTable(l_s + l_p, l_t + l_p) \\ \bullet copyRow(l_p, l_n, m_{l_s, l_t}) 
+			state(database(s)) \bullet copyTable(l_p, l_n) \\ \bullet copyRow(l_p, l_n, m_{l_s, l_t}) 
 			\end{gathered}}

### 2013-02-02 doc/description.tex
default value function added

@@
 &	type(prop(l, t, val, n, m)) = t \\ \\
+&	\mathbf{defaultValue} : Property \rightarrow Value \\
+&	defaultValue(prop(l, t, val, n, m)) = val \\ \\
 &	\mathbf{cardinality} : Property \rightarrow \mathbb{N^{*}} \\

### 2013-02-02 doc/description.tex
case study

@@
 			reference(a) = l_n) \\
-			\circ renClass(l_1) \circ renClass(l_2) \end{gathered}}
+			\circ remClass(l_1) \circ remClass(l_2) \end{gathered}}
  \\\\
@@
 The case study demonstrates how the transformation helps with data evolution in real example of software evolution.
-There are two classes in our software $Person$ and $LegalPerson$ both of them contains information about address (street, city and zip code).
+There are two classes in our software $Person$ and $LegalParty$ both of them contains information about address (street, city and zip code).
 The situation in software is in Figure \ref{fig:case1}, where the whole software is shown.
@@
 	\end{tabular}
-	\caption{Data stored in the table $LegalPerson$.}
+	\caption{Data stored in the table $LegalParty$.}
 \end{subfigure}
-	\caption{The initial state of the example used in the case study. There is a repetition of information structure in the $Person$ and $LegalPerson$ class.}
+	\caption{The initial state of the example used in the case study. There is a repetition of information structure in the $Person$ and $LegalParty$ class.}
 	\label{fig:case1}
@@
 
-The first  step of is to extract two temporary classes representing address of a $Person$ or of a $LegalPerson$.
+The first  step of is to extract two temporary classes representing address of a $Person$ or of a $LegalParty$.
 Temporary classes has to be connected with the origin classes by an association.
@@
 S_1 & = state(S) \\
-& \;\;\;\;\;\; \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip") \\
+& \;\;\;\;\;\; \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip",\\ 
+& \;\;\;\;\;\; \;\;\;\;\;\;  \{(x,x) | x \in instances("Person", database(S))\}) \\
 & \;\;\;\;\;\; \oplus AddAssociation("Person", \\ 
@@
 & \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \}) \\
-&\;\;\;\;\;\; \oplus Extract("LegalPerson", "Address\_tmp2", "street", \\
+&\;\;\;\;\;\; \oplus Extract("LegalParty", "Address\_tmp2", "street", \\
 & \;\;\;\;\;\;\;\;\;\;\;\; "city", "zip")  \\
-& \;\;\;\;\;\; \oplus AddAssociation("LegalPerson", \\ 
-& \;\;\;\;\;\;\;\;\;\;\;\; assoc("residentialAddress", "Address\_tmp2", 1, 1),\\ 
-& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("LegalPerson", database(S)), \\
+& \;\;\;\;\;\; \oplus AddAssociation("LegalParty", \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp2", 1, 1),\\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("LegalParty", database(S)), \\
 & \;\;\;\;\;\;\;\;\;\;\;\; y \in instances("Address\_tmp2", database(S)) : \\
@@
 	\end{tabular}
-	\caption{Data stored in the table $LegalPerson$.}
+	\caption{Data stored in the table $LegalParty$.}
 \end{subfigure}
@@
 	 	\hline
-		Id & Street & City & ZIP & Person \\ \hline  
+		Id & Street & City & ZIP & Person\_Address \\ \hline  
 		11 & Central Park St & New York & 100 01 & 11  \\ \hline
@@
 	 	\hline
-		Id & Street & City & ZIP & LegalPerson\\ \hline  
+		Id & Street & City & ZIP & LegalParty\_Address\\ \hline  
 		100 & Olive ave. & New York & 100 01 & 100 \\ \hline
@@
 %& assoc("address", "Address\_tmp2", 1, 1), \varnothing) \\
-%& \oplus AddAssociation("LegalPerson", \\
+%& \oplus AddAssociation("LegalParty", \\
 %& assoc("residentialAddress", "Address\_tmp1", 1, 1),\varnothing) \\
@@
 \end{align*}
-Finally the associations are renamed and the evolution end.
-\begin{align*}
-S_3 &= S_2  \oplus RenameAssociation("Person", "Address\_tmp1", "Address")\\
-& \oplus RenameAssociation("LegalPerson", "Address\_tmp2". "Address")\\
-\end{align*}
-
 \begin{figure}
@@
 	 	\hline
-		Id & Street & City & ZIP & Person & LegalPerson\\ \hline  
+		Id & Street & City & ZIP & Person\_Address & LegalParty\_Address\\ \hline  
 		1 & Olive ave. & New York & 100 01 & & 100 \\ \hline
@@
 \end{figure}
-
 The scenario was completed successfully, because there were no constraints defined in the database and the data was suitable for the evolution there were no complication.
@@
 \begin{enumerate}
-	\item \label{it:1} Verify the model by an implementation in the Maude tool. \emph{(Estimated time: 1 month)}
-	\item \label{it:2} More complex examples of refactoring cases should be examined with existing model.  \emph{(Estimated time: 0.5 month)}
-	\item \label{it:3} The inheritance relationship between classes should be implemented in the model together with inheritance-based refactoring cases. \emph{(Estimated time: 0.5 month)}
+	\item \label{it:1} The inheritance relationship between classes should be implemented in the model together with inheritance-based refactoring cases. \emph{(Estimated time: 0.5 month)}
+	\item \label{it:2} Verify the model by an implementation in the Maude tool. \emph{(Estimated time: 1 month)}
+	\item \label{it:3} More complex examples of refactoring cases should be examined with existing model.  \emph{(Estimated time: 0.5 month)}
 	\item Prepare an article, which will contain topics \ref{it:1}, \ref{it:2} and \ref{it:3}. \emph{(Estimated time: 1 month)}
@@
 \label{sec:caseStudy}
-Initial State
+The first part of the evolution case study extract a new temporary class and then creates an association.
+\begin{align*}
+S_1 &= state(S) \\
+& \;\;\;\;\;\; \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip",\\ 
+& \;\;\;\;\;\; \;\;\;\;\;\;  \{(x,x) | x \in instances("Person", database(S)\}) \\
+& \;\;\;\;\;\; \oplus AddAssociation("Person", \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp1", 1, 1),\\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("Person", database(S)), \\
+& \;\;\;\;\;\; \;\;\;\;\;\; y \in instances("Address\_tmp1", database(S)) : \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \}) = \\
+& \;\;\;\;\;\; =  state(S) \oplus addClass(cl("Address\_tmp1", \varnothing, \varnothing)) \\
+& \oplus MoveProperty("Person", "Address\_tmp1", "street", "city", "zip", \\
+& \;\;\;\;\;\; \;\;\;\;\;\;    \{(x,x) | x \in instances("Person", database(S))\}) \\
+& \;\;\;\;\;\; \oplus AddAssociation("Person", \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp1", 1, 1),\\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("Person", database(S)), \\
+& \;\;\;\;\;\; \;\;\;\;\;\; y \in instances("Address\_tmp1", database(S)) : \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \})
+\end{align*}
+
+\newpage
+Initial state of the software:
+\medskip
 \hline
@@
 & pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
-& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& cl(LegalParty, (pr(businessName, APPSTRING, \varnothing), \\
 & pr(street, APPSTRING, \varnothing), \\
@@
 & col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
+& tab(LegalParty, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
 & col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
@@
 & ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
-& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+& row("LegalParty", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
 & ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
+& row("LegalParty", ["Id", 20], ["businessName", "Doe"], \\
 & ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003]))
 \end{align*}
+
 \newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{align*}
-& state(S) \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip") = \\
-&=
+\begin{align*}
+S_1 &= state(S) \oplus \mathbf{addClass}(cl("Address\_tmp1", \varnothing, \varnothing)) \\
+& \oplus MoveProperty("Person", "Address\_tmp1", "street", "city", "zip", \\
+& \;\;\;\;\;\; \;\;\;\;\;\;    \{(x,x) | x \in instances("Person", database(S))\}) \\
+& \;\;\;\;\;\; \oplus AddAssociation("Person", \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp1", 1, 1),\\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("Person", database(S)), \\
+& \;\;\;\;\;\; \;\;\;\;\;\; y \in instances("Address\_tmp1", database(S)) : \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \})
 \end{align*}
@@
 & pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
-& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& cl(LegalParty, (pr(businessName, APPSTRING, \varnothing), \\
 & pr(street, APPSTRING, \varnothing), \\
 & pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)),\\
-& cl("Address\_tmp1", \varnothing, \varnothing)
+& \mathbf{cl("Address\_tmp1", \varnothing, \varnothing)}
  )
@@
 & col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
+& tab(LegalParty, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
 & col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
 & col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab("Address\_tmp1", pk("Id"), \varnothing, \varnothing)) 
-\end{align*}
-\begin{align*}
-r &= row("person", ["Id", 1], ["surname", "Griffin"], \\
-& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("person", ["Id", 2], ["surname", "Doe"], \\
-& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
-& row("legalPerson", ["Id", 10], \\
-& ["businessName", "Tools and Machines ltd."], \\
-& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
-& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003]))
-\end{align*}
+& \mathbf{tab("Address\_tmp1", pk("Id"), \varnothing, \varnothing)}) 
+\end{align*}
+%\begin{align*}
+%r &= row("person", ["Id", 1], ["surname", "Griffin"], \\
+%& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
+%& row("person", ["Id", 2], ["surname", "Doe"], \\
+%& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
+%& row("LegalParty", ["Id", 10], \\
+%& ["businessName", "Tools and Machines ltd."], \\
+%& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+%& row("LegalParty", ["Id", 20], ["businessName", "Doe"], \\
+%& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003]))
+%\end{align*}
 
@@
 \newpage
-ExtactClass = copyProperty("street"
+The $MoveProperty$ transforms the software in three steps (one for each property to move).
+First each property and its data are copied from the instances of an old class to the new one, then the copied property and its data are removed.
+\begin{align*}
+	S_2 &= state(S_1) \oplus \mathbf{MoveProperty}("Person", "Address\_tmp1", \\ 
+& \;\;\;\;\;\; \;\;\;\;\;\; "street", "city", "zip", \\
+& \;\;\;\;\;\; \;\;\;\;\;\;    \{(x,x) | x \in instances("Person", database(S))\}) \\
+&  \;\;\;\;\;\; \oplus AddAssociation("Person", \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp1", 1, 1),\\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("Person", database(S)), \\
+& \;\;\;\;\;\; \;\;\;\;\;\; y \in instances("Address\_tmp1", database(S)) : \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \}) = \\
+& = state(S_1) \oplus \mathbf{CopyProperty}("Person", "Address\_tmp1" , "street", \\
+& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,x) | x \in instances("Person", database(S))\}) \\
+& \;\;\;\;\;\; \oplus RemovePropertyAndData("Person", "street") \\
+& \;\;\;\;\;\; \oplus MoveProperty("Person", "Address\_tmp1", \\ 
+& \;\;\;\;\;\; \;\;\;\;\;\; "street", "city", "zip", \\
+& \;\;\;\;\;\; \;\;\;\;\;\;    \{(x,x) | x \in instances("Person", database(S))\}) \\
+&  \;\;\;\;\;\; \oplus AddAssociation("Person", \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp1", 1, 1),\\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("Person", database(S)), \\
+& \;\;\;\;\;\; \;\;\;\;\;\; y \in instances("Address\_tmp1", database(S)) : \\ 
+\end{align*}
+
 \hline
 \begin{align*}
-a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
+a' &=app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
 & pr(street, APPSTRING, \varnothing), \\
 & pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
-& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+& cl(LegalParty, (pr(businessName, APPSTRING, \varnothing), \\
 & pr(street, APPSTRING, \varnothing), \\
 & pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
-& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) , \varnothing)
+& cl("Address\_tmp1", \mathbf{pr(street, APPSTRING, \varnothing)} , \varnothing)
 & )
@@
 \begin{align*}
-t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
+t &=db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
 & col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
 & col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
+& tab(LegalParty, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
 & col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
 & col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab("Address\_tmp1", pk("Id"), col(street, APPSTRING, \varnothing, \varnothing),\\
+& tab("Address\_tmp1", pk("Id"), \mathbf{col(street, APPSTRING, \varnothing, \varnothing)},\\
 & \varnothing)) 
 \end{align*}
-\begin{align*}
-r & \;\;\;\;\;\; =\{ row("person", ["Id", 1], ["surname", "Griffin"], \\
-& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("person", ["Id", 2], ["surname", "Doe"], \\
-& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
-& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
-& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
-& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
-& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"]) \\
-& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"] )\\
-\}
-\end{align*}
+%\begin{align*}
+%r &=\{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+%& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
+%& row("person", ["Id", 2], ["surname", "Doe"], \\
+%& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
+%& row("LegalParty", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+%& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+%& row("LegalParty", ["Id", 20], ["businessName", "Doe"], \\
+%& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+%& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"]) \\
+%& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"] )\\
+%\}
+%\end{align*}
 
 %%%%%%%%%%%%%%%%%%%%%%%
+%\newpage
+%ExtactClass = dropPropertyAndData("street"
+%\begin{align*}
+%a' &=app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
+%& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+%& cl(LegalParty, (pr(businessName, APPSTRING, \varnothing), \\
+%& pr(street, APPSTRING, \varnothing), \\
+%& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+%& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) , \varnothing)
+%& )
+%\end{align*}
+%\begin{align*}
+%t &=db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
+%& col(city, APPSTRING, \varnothing, \varnothing) \\
+%& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
+%& tab(LegalParty, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
+%& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
+%& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
+%& tab("Address\_tmp1", pk("Id"), col(street, APPSTRING, \varnothing, \varnothing),\\
+%& \varnothing)) 
+%\end{align*}
+%\begin{align*}
+%r &=\{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+%& ["city", "NY"], ["zip", 10001]), \\
+%& row("person", ["Id", 2], ["surname", "Doe"], \\
+%& ["city", "LA"], ["zip", 90003]), \\
+%& row("LegalParty", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+%& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+%& row("LegalParty", ["Id", 20], ["businessName", "Doe"], \\
+%& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+%& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"]) \\
+%& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"] )\\
+%\}
+%\end{align*}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\newpage
+%ExtactClass = copyProperty("city"
+%\begin{align*}
+%a' &=app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
+%& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+%& cl(LegalParty, (pr(businessName, APPSTRING, \varnothing), \\
+%& pr(street, APPSTRING, \varnothing), \\
+%& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+%& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing) , \varnothing)
+%& )
+%\end{align*}
+%\begin{align*}
+%t &=db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
+%& col(city, DBSTRING, \varnothing, \varnothing) \\
+%& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+%& tab(LegalParty, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+%& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+%& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+%& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+%& col(city, DBSTRING, \varnothing, \varnothing)
+%& \varnothing)) 
+%\end{align*}
+%\begin{align*}
+%r &=\{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+%& ["city", "NY"], ["zip", 10001]), \\
+%& row("person", ["Id", 2], ["surname", "Doe"], \\
+%&  ["city", "LA"], ["zip", 90003]), \\
+%& row("LegalParty", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+%& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+%& row("LegalParty", ["Id", 20], ["businessName", "Doe"], \\
+%& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+%& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"]) \\
+%& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"] )\\
+%\}
+%\end{align*}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\newpage
+%ExtactClass = dropPropertyAndData("city"
+%\begin{align*}
+%a' &=app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\
+%& pr(zip, APPSTRING, \varnothing)), \\
+%& cl(LegalParty, (pr(businessName, APPSTRING, \varnothing), \\
+%& pr(street, APPSTRING, \varnothing), \\
+%& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+%& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing) , \varnothing)
+%& )
+%\end{align*}
+%\begin{align*}
+%t &=db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
+%& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+%& tab(LegalParty, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+%& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+%& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+%& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+%& col(city, DBSTRING, \varnothing, \varnothing)
+%& \varnothing)) 
+%\end{align*}
+%\begin{align*}
+%r &=\{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+%& ["zip", 10001]), \\
+%& row("person", ["Id", 2], ["surname", "Doe"], \\
+%& ["zip", 90003]), \\
+%& row("LegalParty", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+%& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+%& row("LegalParty", ["Id", 20], ["businessName", "Doe"], \\
+%& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+%& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"]) \\
+%& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"] )\\
+%\}
+%\end{align*}
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\newpage
+%ExtactClass = CopyProperty("zip"
+%\begin{align*}
+%a' &=app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\
+%& pr(zip, APPSTRING, \varnothing)), \\
+%& cl(LegalParty, (pr(businessName, APPSTRING, \varnothing), \\
+%& pr(street, APPSTRING, \varnothing), \\
+%& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+%& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
+%& )
+%\end{align*}
+%\begin{align*}
+%t &=db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
+%& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+%& tab(LegalParty, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+%& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
+%& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+%& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+%& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)
+%& \varnothing)) 
+%\end{align*}
+%\begin{align*}
+%r &=\{ row("person", ["Id", 1], ["surname", "Griffin"], \\
+%& ["zip", 10001]), \\
+%& row("person", ["Id", 2], ["surname", "Doe"], \\
+%& ["zip", 90003]), \\
+%& row("LegalParty", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+%& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
+%& row("LegalParty", ["Id", 20], ["businessName", "Doe"], \\
+%& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
+%& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]) \\
+%& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]) )\\
+%\}
+%\end{align*}
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \newpage
-ExtactClass = dropPropertyAndData("street"
-\begin{align*}
-a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
-& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+The software after moving all properties from $Person$ to $Address\_tmp1$.
+\medskip
+\hline
+\begin{align*}
+a' &=app(cl(Person, (pr(surname, APPSTRING, \varnothing))), \\
+& cl(LegalParty, (pr(businessName, APPSTRING, \varnothing), \\
 & pr(street, APPSTRING, \varnothing), \\
 & pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
-& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) , \varnothing)
+& \mathbf{cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) } \\
+& \mathbf{pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)}
 & )
@@
 \begin{align*}
-t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
-& col(city, APPSTRING, \varnothing, \varnothing) \\
-& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
-& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
-& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab("Address\_tmp1", pk("Id"), col(street, APPSTRING, \varnothing, \varnothing),\\
-& \varnothing)) 
-\end{align*}
-\begin{align*}
-r & \;\;\;\;\;\; =\{ row("person", ["Id", 1], ["surname", "Griffin"], \\
-& ["city", "NY"], ["zip", 10001]), \\
-& row("person", ["Id", 2], ["surname", "Doe"], \\
-& ["city", "LA"], ["zip", 90003]), \\
-& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
-& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
-& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
-& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"]) \\
-& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"] )\\
-\}
-\end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-ExtactClass = copyProperty("city"
-\begin{align*}
-a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
-& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
-& pr(street, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
-& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing) , \varnothing)
-& )
-\end{align*}
-\begin{align*}
-t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
-& col(city, DBSTRING, \varnothing, \varnothing) \\
-& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+t &=db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalParty, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
 & col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
 & col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
-& col(city, DBSTRING, \varnothing, \varnothing)
+& \mathbf{tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),}\\
+& \mathbf{col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)}
 & \varnothing)) 
@@
 \begin{align*}
-r & \;\;\;\;\;\; =\{ row("person", ["Id", 1], ["surname", "Griffin"], \\
-& ["city", "NY"], ["zip", 10001]), \\
-& row("person", ["Id", 2], ["surname", "Doe"], \\
-&  ["city", "LA"], ["zip", 90003]), \\
-& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+r &=\{ \mathbf{row("person", ["Id", 1], ["surname", "Griffin"])}, \\
+& \mathbf{row("person", ["Id", 2], ["surname", "Doe"])}, \\
+& row("LegalParty", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
 & ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
+& row("LegalParty", ["Id", 20], ["businessName", "Doe"], \\
 & ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
-& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"]) \\
-& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"] )\\
-\}
-\end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-ExtactClass = dropPropertyAndData("city"
-\begin{align*}
-a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\
-& pr(zip, APPSTRING, \varnothing)), \\
-& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
-& pr(street, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
-& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing) , \varnothing)
-& )
-\end{align*}
-\begin{align*}
-t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
-& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
-& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
-& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
-& col(city, DBSTRING, \varnothing, \varnothing)
-& \varnothing)) 
-\end{align*}
-\begin{align*}
-r & \;\;\;\;\;\; =\{ row("person", ["Id", 1], ["surname", "Griffin"], \\
-& ["zip", 10001]), \\
-& row("person", ["Id", 2], ["surname", "Doe"], \\
-& ["zip", 90003]), \\
-& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
-& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
-& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
-& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"]) \\
-& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"] )\\
-\}
-\end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-ExtactClass = CopyProperty("zip"
-\begin{align*}
-a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\
-& pr(zip, APPSTRING, \varnothing)), \\
-& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
-& pr(street, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
-& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
-& )
-\end{align*}
-\begin{align*}
-t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
-& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
-& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
-& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
-& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)
-& \varnothing)) 
-\end{align*}
-\begin{align*}
-r & \;\;\;\;\;\; =\{ row("person", ["Id", 1], ["surname", "Griffin"], \\
-& ["zip", 10001]), \\
-& row("person", ["Id", 2], ["surname", "Doe"], \\
-& ["zip", 90003]), \\
-& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
-& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
-& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
-& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]) \\
-& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]) )\\
-\}
-\end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-ExtactClass = DropPopertyAndData("zip"
-\begin{align*}
-a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing))), \\
-& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
-& pr(street, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
-& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
-& )
-\end{align*}
-\begin{align*}
-t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
-& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
-& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
-& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)
-& \varnothing)) 
-\end{align*}
-\begin{align*}
-r & \;\;\;\;\;\; =\{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
-& row("person", ["Id", 2], ["surname", "Doe"]), \\
-& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
-& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
-& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
-& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]) \\
-& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]) )\\
+& \mathbf{row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"],} \\
+& \mathbf{["city", "NY"], ["zip", 10001])} \\
+& \mathbf{row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"],} \\
+& \mathbf{ ["city", "LA"], ["zip", 90003])} )\\
 \}
@@
 \newpage
-$AddAssociation("Person", assoc("address", "Address\_tmp1", 1, 1), 
- \{(x,y)| x \in instances("Person", database(S)), y \in instances("Address\_tmp1", database(S)) : pkVal(x) = pkVal(y) \})$
-\begin{align*}
-a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
-& assoc("address", "Address\_tmp1", 1, 1)), \\ 
-& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
+\begin{align*}
+	S_3 &= state(S_2) \oplus AddAssociation("Person", \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp1", 1, 1),\\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("Person", database(S)), \\
+& \;\;\;\;\;\; \;\;\;\;\;\; y \in instances("Address\_tmp1", database(S)) : \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \}) = \\
+\end{align*}
+
+\hline
+\begin{align*}
+a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
+& \mathbf{assoc("address", "Address\_tmp1", 1, 1)}), \\ 
+& cl(LegalParty, pr(businessName, APPSTRING, \varnothing), \\
 & pr(street, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
+& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing)), \\
 & cl("Address\_tmp1", pr(street, APPSTRING, \varnothing), \\
@@
 \begin{align*}
-t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+t' &=db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalParty, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
 & col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
@@
 & col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
-& fk("person\_address", "Person", \varnothing))\\
+& \mathbf{fk("person\_address", "Person", \varnothing)})\\
 ) 
@@
 \begin{align*}
-r & \;\;\;\;\;\; =\{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
+r &=\{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
 & row("person", ["Id", 2], ["surname", "Doe"]), \\
-& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
+& row("LegalParty", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
 & ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
-& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
+& row("LegalParty", ["Id", 20], ["businessName", "Doe"], \\
 & ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
 & row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], \\
-& ["city", "NY"], ["zip", 10001], ["person\_address", 1]) \\
+& ["city", "NY"], ["zip", 10001], \mathbf{["person\_address", 1]}) \\
 & row("address\_tmp1", ["Id", 2], ["street", "Olive ave. 2225"], \\
-& ["city", "LA"], ["zip", 90003], ["person\_address", 2]) )\\
+& ["city", "LA"], ["zip", 90003], \mathbf{["person\_address", 2]}) )\\
 \}
@@
 \newpage
-\begin{align*}
-& Extract("LegalPerson", "Address\_tmp2", "street", "city", "zip")  \\
-& \oplus AddAssociation("LegalPerson", assoc("address", "Address\_tmp1", 1, 1),\\ 
-& \{(x,y)| x \in instances("LegalPerson", database(S)), \\
-& y \in instances("Address\_tmp2", database(S)) : pkVal(x) = pkVal(y) \}) \\
-\end{align*}
+The next steps are very similar to the previous ones - extracting a temporary class and creates an association.
+\begin{align*}
+S_4 &= S_3 \oplus Extract("LegalParty", "Address\_tmp2", "street", "city", "zip",\\ 
+& \;\;\;\;\;\; \;\;\;\;\;\;  \{(x,x) | x \in instances("LegalParty", database(S)\}) \\
+& \;\;\;\;\;\; \oplus AddAssociation("LegalParty", \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp1", 1, 1),\\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("LegalParty", database(S)), \\
+& \;\;\;\;\;\; \;\;\;\;\;\; y \in instances("Address\_tmp2", database(S)) : \\ 
+& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \})\end{align*}
 \hline
 \begin{align*}
-a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
+a' &=app(cl(Person, pr(surname, APPSTRING, \varnothing), \\ 
 & assoc("address", "Address\_tmp1", 1, 1)), \\ 
-& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing),\\
-& assoc("address", "Address\_tmp2", 1, 1)), \\
+& \mathbf{cl(LegalParty, pr(businessName, APPSTRING, \varnothing)},\\
+& \mathbf{assoc("address", "Address\_tmp2", 1, 1))}, \\
 & cl("Address\_tmp1", pr(street, APPSTRING, \varnothing), \\
 & pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing) \\
-& cl("Address\_tmp2", pr(street, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
+& \mathbf{cl("Address\_tmp2", pr(street, APPSTRING, \varnothing)}, \\
+& \mathbf{pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)}
 & )
@@
 \begin{align*}
-t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
-& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
-& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+t &=db(tab("Person", pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& \mathbf{tab("LegalParty", pk("Id"), } \\
+& \mathbf{col(businessName, DBSTRING, \varnothing, \varnothing), \varnothing),} \\
 & tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
@@
 & fk("person\_address", "Person", \varnothing))\\
-& tab("Address\_tmp2", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
-& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
-& fk("legalPerson\_address", "LegalPerson", \varnothing))\\
-) 
-\end{align*}
-\begin{align*}
-r & \;\;\;\;\;\; =\{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
+& \mathbf{tab("Address\_tmp2", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing)},\\
+& \mathbf{col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)} \\
+& \mathbf{fk("LegalParty\_address", "LegalParty", \varnothing))}) 
+\end{align*}
+\begin{align*}
+r &=\{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
 & row("person", ["Id", 2], ["surname", "Doe"]), \\
-& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."]), \\
-& row("legalPerson", ["Id", 20], ["businessName", "Doe"])\\
+& \mathbf{row("LegalParty", ["Id", 10],} \\
+& \mathbf{["businessName", "Tools and Machines ltd."])}, \\
+& \mathbf{row("LegalParty", ["Id", 20], ["businessName", "Doe"])}\\
 & row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], \\ 
@@
 & ["city", "LA"], ["zip", 90003], ["Person", 2]) )\\
-& row("address\_tmp2", ["Id", 10], ["street", "Hill stret 22"], \\
-& ["city", "NY"], ["zip", 10001], ["LegalPerson\_address", 10]) \\
-& row("address\_tmp2", ["Id", 20], ["street", "Pine ave. 2225"], \\
-& ["city", "LA"], ["zip", 90003], ["LegalPerson\_address", 20]) )\\
+& \mathbf{row("address\_tmp2", ["Id", 10], ["street", "Hill stret 22"], }\\
+& \mathbf{["city", "NY"], ["zip", 10001], ["LegalParty\_address", 10]}) \\
+& \mathbf{row("address\_tmp2", ["Id", 20], ["street", "Pine ave. 2225"],} \\
+& \mathbf{["city", "LA"], ["zip", 90003], ["LegalParty\_address", 20])} )\\
 \}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \newpage
 \begin{align*}
-& \oplus AddAssociation("Person", \\
-& assoc("address", "Address\_tmp2", 1, 1), \varnothing) \\
+	S_5 &= S_4 \oplus Merge("Address\_tmp1", "Address\_tmp2", "Address") \\
+& 
 \end{align*}
@@
 \begin{align*}
-a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
-& assoc("address", "Address\_tmp1", 1, 1), assoc("address", "Address\_tmp2", 1, 1)), \\ 
-& cl(LegalPerson, pr(businessName, APPSTRING, \varnothing),\\
-& assoc("address", "Address\_tmp2", 1, 1), assoc(\mathbf{"address"}, "Address\_tmp1", 1, 1)), \\
-& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing) \\
-& cl("Address\_tmp2", pr(street, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
-& )
-\end{align*}
-\begin{align*}
-t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing)) \\
-& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
-& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
-& fk("address", "Person", \varnothing), fk("address", "LegalPerson", \varnothing))\\
-& tab("Address\_tmp2", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
-& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
-& fk("LegalPerson\_address", "LegalPerson", \varnothing), fk("Person\_address", "Person", \varnothing))\\
-) 
-\end{align*}
-\begin{align*}
-r & \;\;\;\;\;\; =\{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
+a' &=app(cl(Person, pr(surname, APPSTRING, \varnothing), \\ 
+& \mathbf{assoc("address", "Address", 1, 1)}), \\ 
+& cl(LegalParty, pr(businessName, APPSTRING, \varnothing),\\
+& \mathbf{assoc("address", "Address", 1, 1)}), \\
+& \mathbf{cl("Address", pr(street, APPSTRING, \varnothing),} \\
+& \mathbf{pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing)}
+\end{align*}
+\begin{align*}
+t &=db(tab("Person", pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("LegalParty", pk("Id"),  \\
+& col(businessName, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& \mathbf{tab("Address", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing)},\\
+& \mathbf{col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)} \\
+& \mathbf{fk("person\_address", "Person", \varnothing)},\\
+& \mathbf{fk("LegalParty\_address", "LegalParty", \varnothing)}) 
+\end{align*}
+\begin{align*}
+r &= (row("person", ["Id", 1], ["surname", "Griffin"]), \\
 & row("person", ["Id", 2], ["surname", "Doe"]), \\
-& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."]), \\
-& row("legalPerson", ["Id", 20], ["businessName", "Doe"])\\
-& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], \\ 
-& ["city", "NY"], ["zip", 10001], ["Person", 1]) \\
-& row("address\_tmp1", ["Id", 2], ["street", "Olive ave. 2225"], \\
-& ["city", "LA"], ["zip", 90003], ["Person", 2]) )\\
-& row("address\_tmp2", ["Id", 10], ["street", "Hill stret 22"], \\
-& ["city", "NY"], ["zip", 10001], ["LegalPerson", 10]) \\
-& row("address\_tmp2", ["Id", 20], ["street", "Pine ave. 2225"], \\
-& ["city", "LA"], ["zip", 90003], ["LegalPerson", 20]) )\\
-\}
-\end{align*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\begin{align*}
-& \oplus AddAssociation("LegalPerson", \\
-& assoc("address", "Address\_tmp1", 1, 1),\varnothing) \\
-\end{align*}
-\hline
-\begin{align*}
-a' & \;\;\;\;\;\; =app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
-& assoc("address", "Address\_tmp1", 1, 1)), \\ 
-& cl(LegalPerson, pr(businessName, APPSTRING, \varnothing),\\
-& assoc("address", "Address\_tmp2", 1, 1), assoc(\mathbf{"address"}, "Address\_tmp1", 1, 1)), \\
-& cl("Address\_tmp1", pr(street, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing) \\
-& cl("Address\_tmp2", pr(street, APPSTRING, \varnothing), \\
-& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
-& )
-\end{align*}
-\begin{align*}
-t & \;\;\;\;\;\; =db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing)) \\
-& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
-& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
-& fk("Person\_address", "Person", \varnothing), fk("LegalPerson\_address", "LegalPerson", \varnothing))\\
-& tab("Address\_tmp2", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
-& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
-& fk("LegalPerson\_address", "LegalPerson", \varnothing))\\
-) 
-\end{align*}
-\begin{align*}
-r & \;\;\;\;\;\; =\{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
-& row("person", ["Id", 2], ["surname", "Doe"]), \\
-& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."]), \\
-& row("legalPerson", ["Id", 20], ["businessName", "Doe"])\\
-& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], \\ 
-& ["city", "NY"], ["zip", 10001], ["Person", 1]) \\
-& row("address\_tmp1", ["Id", 2], ["street", "Olive ave. 2225"], \\
-& ["city", "LA"], ["zip", 90003], ["Person", 2]) )\\
-& row("address\_tmp2", ["Id", 10], ["street", "Hill stret 22"], \\
-& ["city", "NY"], ["zip", 10001], ["LegalPerson", 10]) \\
-& row("address\_tmp2", ["Id", 20], ["street", "Pine ave. 2225"], \\
-& ["city", "LA"], ["zip", 90003], ["LegalPerson", 20]) )\\
-\}
+& row("LegalParty", ["Id", 10], \\
+& ["businessName", "Tools and Machines ltd."]), \\
+& row("LegalParty", ["Id", 20], ["businessName", "Doe"])\\
+& \mathbf{row("address", ["Id", 1], ["street", "S. Mary Evenue 22"]}, \\ 
+& \mathbf{["city", "NY"], ["zip", 10001], ["Person", 1])} \\
+& \mathbf{row("address", ["Id", 2], ["street", "Olive ave. 2225"],} \\
+& \mathbf{["city", "LA"], ["zip", 90003], ["Person", 2])} )\\
+& \mathbf{row("address", ["Id", 3], ["street", "Hill stret 22"], }\\
+& \mathbf{["city", "NY"], ["zip", 10001], ["LegalParty\_address", 10]}) \\
+& \mathbf{row("address", ["Id", 4], ["street", "Pine ave. 2225"],} \\
+& \mathbf{["city", "LA"], ["zip", 90003], ["LegalParty\_address", 20])} ))
 \end{align*}

### 2013-02-02 doc/description.tex
small fix in merge class

@@
 \forall f \in (foreignKeys(t_1) \cup foreignKey(t_2)) \\ \setminus (foreignKey(t_1) \cap foreignKey(t_2)) \\ : NOTNULL \not \in constraints(f)  }{\begin{gathered}
-	state(database(s)) \\ \bullet addTable(l_n, pk("Id", s_1), columns(t_1), foreignKeys(t_1)) \\ 
+	state(database(s)) \\ \bullet addTable(l_n, pk("Id", s_1), columns(t_1), foreignKeys(t_1) \cup foreignKeys(t_2)  ) \\ 
 \bullet (\forall r \in instances(t_1, database(s)) \cup instances(t_2, database(s)) :  \\ n =  row(l_n, next(seq(key(l_n))), pairs(r)), insertRow(l_n, n) \\ 

### 2013-02-02 doc/description.tex
primary key has a sequence

@@
 		%\inference{|c| > 1}{\rho_c(head(c)) \cup \rho_c(tail(c))} \\ \\
-		\inference{|c| = 1}{table(name(c), primaryKey("Id"), \varnothing, \varnothing)} % \end{cases} 
+		\inference{|c| = 1}{table(name(c), primaryKey("Id", s_1), \varnothing, \varnothing)} % \end{cases} 
 \\ 	
@@
      		}{\begin{gathered}
-     			tab(name(p), primaryKey("Id"), \\ col("value", \rho_t(type(p)), \varnothing), \\ fk(name(p), 			name(owningClass(p)), NOTNULL) \\
+     			tab(name(p), primaryKey("Id", s_1), \\ col("value", \rho_t(type(p)), \varnothing), \\ fk(name(p), 			name(owningClass(p)), NOTNULL) \\
     	\end{gathered}}\\ \\
@@
      		}{\begin{gathered}  
-     			tab(name(p), primaryKey("Id"), \\ 
+     			tab(name(p), primaryKey("Id", s_1), \\ 
      			col("value", \rho_t(type(p)), NOTNULL), \\ 			fk(name(p), name(owningClass(p)), NOTNULL)
@@
   		}{\begin{gathered}  
-		 table(name(a), pk("Id"), \varnothing, \\ fk(name(owningClass(a)), \\ name(owningClass(a)), NOTNULL) \\ fk(reference(a), reference(a), NOTNULL)) 
+		 table(name(a), pk("Id", s_1), \varnothing, \\ fk(name(owningClass(a)), \\ name(owningClass(a)), NOTNULL) \\ fk(reference(a), reference(a), NOTNULL)) 
   		\end{gathered}}  
@@
 \begin{align*}
-t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
+t &= db(tab(Person, pk("Id", s_1), col(surname, APPSTRING, \varnothing, \varnothing), \\
 & col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
 & col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalParty, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
+& tab(LegalParty, pk("Id", s_1),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
 & col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
@@
 \begin{align*}
-t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
+t &= db(tab(Person, pk("Id", s_1), col(surname, APPSTRING, \varnothing, \varnothing), \\
 & col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
 & col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalParty, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
+& tab(LegalParty, pk("Id", s_1),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
 & col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
 & col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& \mathbf{tab("Address\_tmp1", pk("Id"), \varnothing, \varnothing)}) 
+& \mathbf{tab("Address\_tmp1", pk("Id", s_1), \varnothing, \varnothing)}) 
 \end{align*}
@@
 \begin{align*}
-t &=db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
+t &=db(tab(Person, pk("Id", s_1), col(surname, APPSTRING, \varnothing, \varnothing), \\
 & col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
 & col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalParty, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
+& tab(LegalParty, pk("Id", s_1),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
 & col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
 & col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab("Address\_tmp1", pk("Id"), \mathbf{col(street, APPSTRING, \varnothing, \varnothing)},\\
+& tab("Address\_tmp1", pk("Id", s_1), \mathbf{col(street, APPSTRING, \varnothing, \varnothing)},\\
 & \varnothing)) 
@@
 %\begin{align*}
-%t &=db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
+%t &=db(tab(Person, pk("Id", s_1), col(surname, APPSTRING, \varnothing, \varnothing), \\
 %& col(city, APPSTRING, \varnothing, \varnothing) \\
 %& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-%& tab(LegalParty, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
+%& tab(LegalParty, pk("Id", s_1),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
 %& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
 %& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
-%& tab("Address\_tmp1", pk("Id"), col(street, APPSTRING, \varnothing, \varnothing),\\
+%& tab("Address\_tmp1", pk("Id", s_1), col(street, APPSTRING, \varnothing, \varnothing),\\
 %& \varnothing)) 
@@
 %\begin{align*}
-%t &=db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
+%t &=db(tab(Person, pk("Id", s_1), col(surname, DBSTRING, \varnothing, \varnothing), \\
 %& col(city, DBSTRING, \varnothing, \varnothing) \\
 %& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-%& tab(LegalParty, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+%& tab(LegalParty, pk("Id", s_1),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
 %& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
 %& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-%& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+%& tab("Address\_tmp1", pk("Id", s_1), col(street, DBSTRING, \varnothing, \varnothing),\\
 %& col(city, DBSTRING, \varnothing, \varnothing)
@@
 %\begin{align*}
-%t &=db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
+%t &=db(tab(Person, pk("Id", s_1), col(surname, DBSTRING, \varnothing, \varnothing), \\
 %& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-%& tab(LegalParty, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+%& tab(LegalParty, pk("Id", s_1),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
 %& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
 %& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-%& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+%& tab("Address\_tmp1", pk("Id", s_1), col(street, DBSTRING, \varnothing, \varnothing),\\
 %& col(city, DBSTRING, \varnothing, \varnothing)
@@
 %\begin{align*}
-%t &=db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
+%t &=db(tab(Person, pk("Id", s_1), col(surname, DBSTRING, \varnothing, \varnothing), \\
 %& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-%& tab(LegalParty, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+%& tab(LegalParty, pk("Id", s_1),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
 %& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
 %& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-%& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+%& tab("Address\_tmp1", pk("Id", s_1), col(street, DBSTRING, \varnothing, \varnothing),\\
 %& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)
@@
 \begin{align*}
-t &=db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalParty, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+t &=db(tab(Person, pk("Id", s_1), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalParty, pk("Id", s_1),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
 & col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
 & col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& \mathbf{tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),}\\
+& \mathbf{tab("Address\_tmp1", pk("Id", s_1), col(street, DBSTRING, \varnothing, \varnothing),}\\
 & \mathbf{col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)}
@@
 \begin{align*}
-t' &=db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab(LegalParty, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
+t' &=db(tab(Person, pk("Id", s_1), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab(LegalParty, pk("Id", s_1),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
 & col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
 & col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& tab("Address\_tmp1", pk("Id", s_1), col(street, DBSTRING, \varnothing, \varnothing),\\
 & col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
@@
 \begin{align*}
-t &=db(tab("Person", pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& \mathbf{tab("LegalParty", pk("Id"), } \\
+t &=db(tab("Person", pk("Id", s_1), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& \mathbf{tab("LegalParty", pk("Id", s_1), } \\
 & \mathbf{col(businessName, DBSTRING, \varnothing, \varnothing), \varnothing),} \\
-& tab("Address\_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
+& tab("Address\_tmp1", pk("Id", s_1), col(street, DBSTRING, \varnothing, \varnothing),\\
 & col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing) \\
 & fk("person\_address", "Person", \varnothing))\\
-& \mathbf{tab("Address\_tmp2", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing)},\\
+& \mathbf{tab("Address\_tmp2", pk("Id", s_1), col(street, DBSTRING, \varnothing, \varnothing)},\\
 & \mathbf{col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)} \\
@@
 \begin{align*}
-t &=db(tab("Person", pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& tab("LegalParty", pk("Id"),  \\
+t &=db(tab("Person", pk("Id", s_1), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
+& tab("LegalParty", pk("Id", s_1),  \\
 & col(businessName, DBSTRING, \varnothing, \varnothing), \varnothing), \\
-& \mathbf{tab("Address", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing)},\\
+& \mathbf{tab("Address", pk("Id", s_1), col(street, DBSTRING, \varnothing, \varnothing)},\\
 & \mathbf{col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)} \\

### 2013-02-02 doc/description.tex
small fixes

@@
 \begin{align*}
-& 	\mathbf{Class} = <label, Property*, Association*> \\
+& 	\mathbf{Class} = (label, Property*, Association*) \\
 & 	\mathbf{cl} : Label \times Property* \times Association* \rightarrow Class \\\\

### 2013-02-02 doc/paper.tex
work on paper begun

\documentclass[11pt]{article}
\usepackage{amssymb}
\usepackage[fleqn]{amsmath}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{semantic}

\title{Refactoring of Software Using ORM: A Formal Description}
\author{Ondrej Macek}

\begin{document}
\maketitle
\section{Introduction}
\label{sex:intro}
The evolution (change) of a software is a common issue during software development. It occurs from many reasons in all phases of software lifecycle. The evolution severity usually depends on number of changes which has to be made and on the number of affected software components. The focus of this paper is in data evolution context of software implemented in object oriented language using relational database. It means that the data evolution affect at least two software components: persistent objects (entities) and the database. The evolution is based on developer's point of view in this paper as it origins in the need to evolve code. This kind of evolution is straightforward in case there are no data stored in the database, because the database schema can be generated according to an evolved code, moreover some object-relational mapping (ORM) frameworks are able to process basic evolutionary transformation on database with stored data. The evolution of an software with ORM become more difficult when there are stored data and when the evolutionary scenario is more complex. The data migration to the new database schema is usually processed manually, which is expensive and error prone.

This paper shows how the code refactoring can be used as the source for data migration. The approach based on changes in code is useful in case when there is no model of database as it is derived directly from code. Next it can help developers to understand impact  of code changes on database. The formal description of code refactoring impact on database and stored data can help with construction of a standalone migration framework or evolutionary extension for existing ORM frameworks.

The paper is organized as follows: first the evolution of software is described in detail in Sec. \ref{sec:problem}, then the related work in presented in Sec. \ref{sec:related-work}. In Sec. \ref{sec:models} the models of software and database is provided together with a set of basic evolutionary transformations. The impact of code refactoring on database is discussed in Sec. \ref{sec:evolution}. Finally the applicability and limitations of the proposed solution is discussed in Sec. \ref{sec:contribution}.

\section{Problem Definition} 
\label{sec:problem}
A software implemented using an object-oriented language, which uses relational database as a data storage (as illustrated in the Figure \ref{fig:evolution}) is object of interest of this paper. There are three important components of such a software: application (its persistent layer concretely), database (consisting of database schema and stored data) and an ORM, therefore we define a software as a triple consisting of an application and a database which are connected together by an ORM:
\begin{align*}
& \mathbf{software} = ( Application, Database, ORM )
\end{align*}
The software has to be in consistent state so its users could have benefit from its usage. The state of the software is defined to communicate the sate of the software to its user, the symbol $\perp$ is used as a notation for inconsistent state. 
\begin{align*}
& \mathbf{state} = Consistent\; s \: | \: \perp, s \in State 
\end{align*}
The software is in consistent state if application and database are consistent and the database structure corresponds to the application structure according to the ORM.

The evolution of a software is a transformation from one consistent state to another one. These states are called generations of the software and the function which changed the state of a software are called transformations. The evolution consists of change in application and in database as shown in the Figure \ref{fig:evolution}. 
\begin{figure}
\centering
	\includegraphics[scale=0.4]{./images/evolution_simple}
	\caption{The evolution of data changes the system on all levels. The figure shows all the components of the evolution process.}
	\label{fig:evolution}
\end{figure}
The usual approach is to refactor code (which is quite fast due to help of IDE) and then re-generation or manual migration of the database. Another approach is to evolve a database (e.g. due to database optimizations) and then update the code. This approach is very common too, nevertheless this paper focus on evolution which origin from the application point of view.

The refactoring itself does not provide enough information for the complex migration of database and data, therefore a level defining the evolution is added - the set of all possible transformations $E$. The evolution definition from this level is propagated into both involved levels - application and database. The evolution is defined as:
\begin{align*}
& t(state(s)) = \begin{cases}
state(sw(\Psi(s, t), \Phi(s,t)) \; \mathbf{if} \; state(s) = Consistent \; s\\\\
 	\perp
\end{cases}\\ 
& s \in Software, t \in E
\end{align*}
Where $\Psi$ maps the software evolution cases to the code refactoring and $\Phi$ to maps to the evolutionary transformation of database.

The main issue during the software evolution is the preservation of data and information consistency. The data lost is a situation when data are damaged or lost during the migration. The information consistency means the data was transformed according to the evolution of the application and they correspond with developer's intention.

\section{Related Work}
\label{sec:related-work}

\section{Model of Software}
\label{sec:models}
The aim of this paper is to provide an definition of data evolution, which can be applied in case of various programming language, therefore the platform independent models of application and database are introduced. These models also provide simplification, which should increase the understanding of the proposed evolutions.

\subsection{Application Model}
%TODO popis
\paragraph{Application} Application is defined as a sequence of classes and it creates the context for all structures used in the software persistent layer.
\begin{align*}
& \mathbf{Application} = (Class*)
\end{align*}

\paragraph{Class} Class represents a basic structural unit in the application model. It has a unique name, one or more properties and it can be associated to other classes in the application. 
\begin{align*}
& 	\mathbf{Class} = (label, Property*, Association*)
\end{align*}
The properties of a class can be of two kinds - values or collections - according of theirs cardinality.

\paragraph{Property} Property represents a feature of a class which is represented as a primitive type. The property can be mandatory, can have a default value and according to its cardinality it can represent a single value or a collection of values. 
\begin{align*}
&	\mathbf{Property} = (label, AppType, DefaultValue, Cardinality, Mandatory)
\end{align*}

\paragraph {Association} Association represents a connection between two classes. It has a unique name and the reference is represented by the label of referenced class. The class which owns the association is consider to be the starting class of an association, referenced class is consider to be the ending class of an association. The cardinalities defines the multiplicity of both association ends.
\begin{align*}
&	\mathbf{Association} = (label, classRef, startCardinality, endCardinality) 
\end{align*}

\paragraph{Application type} Application type ($AppType$) represents primitive type in the application. There are usually defined types such as String, Integer, Boolean etc. in contrast there is only one type in our model, because we focus on structural and data changes and type casting operations are not important for us. The only $AppType$ type is called $APPSTRING$.
\begin{align*}
& \mathbf{AppType} = APPSTRING
\end{align*}

%TODO consistency

\subsection{Database Model}
he relational database consists of database schema which defines the structure of the database and data, which in our ORM system represents stored instances. The database is defined as a tuple:
\begin{align*}
&	\mathbf{Database} = ( Table*, Row* )
\end{align*}

\paragraph{Table} Table represents a basic concept of database schema. It has a unique name, one or more columns and it can be related to other tables in the schema by foreign keys. Rows in the table represents stored data.
\begin{align*}
&	\mathbf{Table} = (label, primaryKey, Column*, ForeignKey*)
\end{align*}

\paragraph{Column} Column defines data values and types which can be part of a table record.
\begin{align*}
&	\mathbf{Column} = (label, DbType, DefaultValue, Constraint*) 
\end{align*}

\paragraph{Foreign key} Foreign key is a reference to another table's primary key, it has an unique name and it can be constrained. The value of a foreign key is $\varnothing$ or a non-zero natural number.
\begin{align*}
&	\mathbf{ForeignKey} = (Label, Label, Constraint*) 
\end{align*}

\paragraph{Primary key} Primary key is unambiguous identifier of a record in a table. The primary key is always provided (automatically generated) by the database as a non-zero natural number. Primary key is always defined with constraints $NOTNULL$ and $UNIQUE$. 
\begin{align*}
&	\mathbf{PrimaryKey} =  ( Label, Sequence ) \end{align*}
The sequence generates values of primary key. The new value of a key is obtained by calling the function $next(s)$. The sequence which provides numbers incremented by one is called $s_1$ if the increment is 1, $s_{10}$ if the increment is 10 etc.

\paragraph{Data Types} Database data types $DbType$ represents primitive types in the database. There are usually defined types such as Varchar, Integer, Boolean etc. There is only one type of column values defined in the model called $DBSTRING$. Next the $DBINT$ type is defined for the values of keys.
\begin{align*}
&	\mathbf{DbType} = DBSTRING \; | \; DBINT
\end{align*}

\paragraph{Constraints} There are two types of constraints defined in the model. Both constraints are column constraints - first constraint defines non-empty columns, second constraint defines there has to be unique records in a column or foreign key.
\begin{align*}
&	\mathbf{Constraint} = NOTNULL \; | \; UNIQUE 
\end{align*}

A database consists not only from schema but also from data which are represented as rows in a table. A table row in our model is a tuple consisting of value pairs, which represents concrete value of concrete column or key. Each row contains a name of a table it belongs to.
\begin{align*}
&	\mathbf{Row} = (Label, N, Pair*) \\
&	\mathbf{Pair} = (Label, Value) 
\end{align*}

%TODO consistency

\subsection{Object Relational Mapping}
The ORM can differ from application to application, its impact on the evolution is not crucial but it has to be present to provide full information about the method. The mapping is similar to the Hibernate
mapping, thus a lot of developers should be familiar with it. 
% add mapping

\section{Data Evolution of Software}
\label{sec:evolution}

\section{Contribution}
\label{sec:contribution}

\section{Conclusion}

\end{document}

### 2013-02-03 doc/paper.tex
chapter application evolution

@@
 \section{Introduction}
-\label{sex:intro}
+\label{sec:intro}
 The evolution (change) of a software is a common issue during software development.
@@
 \label{sec:evolution}
+The evolution of the software is mapped to the layer specific transformation, which are defined in Sec.
+\ref{sec:app-evolution} and \ref{sec:db-evolution}.
+Finally the evolution cases and their mapping on application and database components are presented in \ref{sec:sw-evolution}.
+The transformations are divided into groups according to their impact on structure and data.
+
+%TODO evalutation of trnasformations
+
+\subsection{Application Evolution}
+\label{sec:app-evolution}
+The application model defines only the structure of an application thus there are defined transformation for its creating, altering and disassembling.
+Therefore the conditions for successful transformations are simple 1) when creating an object in the model or altering its name collisions has to be prevent, 2) when renaming a class the references have to be updated 3) when creating or altering an association the existence of referenced class has to be verified.
+The only issue during disassembling is the class can be removed only in case it is not associated by other classes.
+If a transformation cannot be processed the state of the application become inconsistent.
+The transformations are in table \ref{tab:app-evolution}.
+The list of altering operations is not complete as there are missing transformation for changing the obligation or cardinality of properties and associations.
+It is because these transformations are not so important for the next text when the complex refactorings are considered.
+\begin{table}
+\centering
+	\begin{tabular}{|l|}
+	\hline
+	\textbf{Application Creation} \\
+	newProperty(className, property, application) \\
+	newAssociation(className, association, application) \\
+	newClass(class, application) \\
+	\textbf{Application Alternation} \\
+	renameProperty(className, oldName, newName, application) \\
+	renameAssociation(className, oldName, newName, application) \\
+	renameClass(oldName, newName, application)\\
+	\textbf{Application Disassembly} \\
+	removeProperty(className, propertyName, application) \\
+	removeAssociation(className, associationName, application) \\
+	removeClass(className, application)\\
+	\hline
+	\end{tabular}
+	\caption{The transformations for application evolution}
+	\label{tab:app-evolution}
+\end{table}
+
+\subsection{Database Evolution}
+\label{sec:db-evolution}
+
+\subsection{Software Evolution}
+\label{sec:sw-evolution}
 

### 2013-02-03 doc/paper.tex
chapter about database evolution

@@
 \label{sec:problem}
+% víc rozepsat o změnách a jak se jim bránit a kam tím směřujeme
 A software implemented using an object-oriented language, which uses relational database as a data storage (as illustrated in the Figure \ref{fig:evolution}) is object of interest of this paper.
@@
 	\end{tabular}
-	\caption{The transformations for application evolution}
+	\caption{The transformations for application evolution.}
 	\label{tab:app-evolution}
@@
 \label{sec:db-evolution}
+The database consists of two parts - from database schema which defines the structure and from stored data - hence the evolutionary transformation has to consider both parts.
+The transformation for evolution of structure has similar conditions to the evolution of application application.
+The transformations for data manipulation are inspired by the SQL language.
+The basic transformations are in Table \ref{tab:db-basic-evolution}.
+\begin{table}
+\centering
+	\begin{tabular}{|l|}
+	\hline
+	\textbf{Database Creation} \\
+	addColumn(tableName, column, database) \\
+	addForeignKey(tableName, key, database) \\
+	addTable(table, database)\\
+	\textbf{Database Alternation} \\
+	alterColumnName(tableName, oldName, newName, dataabse) \\
+	alterForeignKeyName(tableName, oldName, newName, dataabse) \\
+	alterTableName(oldName, newName, dataabse) \\
+	\textbf{Database Disasembling} \\
+	dropColumn(tableName, columnName, database) \\
+	dropForeignKey(tableName, keyName, database) \\
+	dropTable(tableName, database) \\
+	\textbf{Data Manipulation} \\
+	select(tableName, rowId, database) \\
+	select(tableName, database) \\
+	insert(row, database) \\
+	insert(rowId, value, database) \\
+	delete(tableName, rowId, database) \\
+	delete(tableName, rowId, columnName, database) \\
+	\hline
+	\end{tabular}
+	\caption{The transformations for database evolution.}
+	\label{tab:db-basic-evolution}
+\end{table}
+\begin{table}
+\centering
+	\begin{tabular}{|l|}
+	\hline
+	\textbf{Copy Structure} \\
+	copyColumnStructure(sourceTableName, targetTableName, \\  \hspace{1in} sourceColumnName, targetColumnName, database) \\
+	copyTableStructure(sourceTable, newTableName, database) \\
+	\textbf{Copy Values} \\
+	copyColumn(sourceTable, targetTable, \\
+	\hspace{1in} sourceColumnName, targetColumnName, \\ 
+	\hspace{1in} mapping, database) \\
+	copyTabele(sourceTable, targetTable, mapping,  database) \\
+	\textbf{Data Secure Database Disassembling} \\
+	dropEmptyColumn(tableName, columnName, database) \\
+	dropEmptyForeignKey(tableName, keyName, database) \\
+	dropEmptyTable(tableName, database) \\
+	\hline
+	\end{tabular}
+	\caption{The transformations for database evolution which are created as SQL functions.}
+	\label{tab:db-adv-evolution}
+\end{table}
+
+The transformation for disassembly the database has fatal impact on the data preservation, therefore the advanced set of transformations is created.
+These transformations are not part of SQL standard, although they can be implemented as database functions.
+These transformations creates a structural copy of database structures or data or creates a secure way to disassembly the database.
+These transformations are in table \ref{tab:db-adv-evolution}.
+
+The copying of structural elements has the same limits as database creation.
+The column copying transformation needs one more information to succeed - there has to be provided mapping between the source and target table to assure the data information consistency.
+Moreover the source and target should have the same structure, type and constraints.
+
+The data secure transformations for disassembling the database can succeed (e.g. drop a table) when there are no data contained within the dropped element.
 
@@
 \label{sec:sw-evolution}
+The evolution of the whole software is based on the basic transformations specific for each software part.
+The evolution is described from the developer point of view, therefore the transformations use the names and elements from application context, these names and elements has to be transformed into a database context - this is assured by the ORM.
 

### 2013-02-03 doc/paper.tex
advances transformation started

@@
 
+The defined evolution are divided into two groups - first group contains the basic transformations for creating, altering and disassembling the software, second group contains the advanced refactoring cases.
+
+\subsection{Basic Evolutionary Transformations}
+The basic evolutionary transformations of the software are based on basic evolution of an application defined in Sec.
+\ref{sec:app-evolution}, nevertheless the evolution of the database is not so straightforward because properties and associations can be mapped as a column (foreign key respectively) or as a table, this has to be reflected during the evolution as shown on the example of creating a new property:
+\begin{align*}
+& newProperty(className, property, application) \Rightarrow \\
+& \;\;\; \Rightarrow \begin{cases}
+  addColumn(ORM(className), ORM(property), database) \\\mathbf{if} \; cardinality(property \leq 1)  \\\\
+  addTable(ORM(property), database) \\
+  \mathbf{if} \; cardinality(property > 1)  
+ \end{cases}
+\end{align*}
+Next transformation are provided in the shortened way in table \ref{tab:sw-basic-evolution}, there is the basic application transformation and possible mappings on database level according to to the cardinality.
+\begin{table}
+\centering
+	\begin{tabular}{|p{0.4\textwidth} p{0.4\textwidth}|}
+	\hline
+	Application Transformation & Database Transformation \\
+	\multicolumn{2}{|c|}{\textbf{Software Creation}} \\
+	newProperty & addColumn \\
+	& addTable \\ 
+	& \\
+	newAssociation & addColumn \\
+	& addTable \\
+	& \\
+	newClass & addTable \\
+	\multicolumn{2}{|c|}{\textbf{Application Alternation}} \\
+	renameProperty & renameColumn \\
+	& renameTable \\
+	& \\
+	renameAssociation & renameForeignKey \\
+	& renameTable \\
+	& \\
+	renameClass & renameTable \\
+	\multicolumn{2}{|c|}{\textbf{Application Disassembly}} \\
+	removeProperty & removeColumn \\
+	& removeTable \\
+	& \\
+	removeAssociation & removeForeignKey \\
+	& removeTable \\
+	& \\
+	removeClass & removeTable\\
+	\hline
+	\end{tabular}
+	\caption{The mapping between evolution of the application to the database transformations.}
+	\label{tab:sw-basic-evolution}
+\end{table}
+
+\subsection{Advanced Evolutionary Transformations}
+Advanced evolutionary transformations are based on the basic ones, they can be obtained as a concatenation of transformations.
+The concatenation of transformation is possible because each transformation produces a state of a software (or application or database) which can be used as an input for the next transformation.
+
+All transformations presented so far has the same input information for both application and database, whereas the advanced transformations usually needs a mapping between stored data (instances) as theirs input.
+First two of them - $copyProperty$ and $mergeClasses$ completes the set of basic transformations.
+
+The $copyProperty$ creates a duplicate of a property in a given class.
+It can not rename the property so it can not be used for creating duplicates of a property in the property owning class.
+If the mapping is not provided the transformation creates only a structural copy of the property otherwise it copies the values of the given column too.
+\begin{align*}
+& \mathbf{copyProperty}(sourceClass, propertyName, targetClass, \\
+& \; \; \; mapping, software) 
+\end{align*}
+The transformation is interpreted for both software components - in case of the application a new property is added:
+\begin{align*}
+& copyProperty(sourceClass, property, targetClass, mapping, \\ 
+&  \; \; \; software) \rightarrow newProperty(targetClass, property, application) 
+\end{align*}
+in case of database the copy of a column or table is created according to property's cardinality and then values are copied:
+\begin{align*}
+& copyPropertysourceClass, property, targetClass, mapping, \\ 
+&  \; \; \; software) \rightarrow 
+\end{align*}
+
+The mapping provides information which value from the source table rows will be copied to which row in target table. 
+\begin{align*}
+&	mapping =  (r_1, r_2)*,\\
+& 	r_1 \in select(ORM(sourceClassName), database), \\
+& 	r_2 \in select(ORM(targetClassName), database) 
+\end{align*}
+
+The mapping features determine if the $moveProperty$ is data inconsistent or not.
+
+\begin{align*}
+&	\forall r_1, r_1 \in instances(l_1, d) \Rightarrow \exists p \in m_{l_1, l_2} \wedge first(p) = r_1 \\
+&	\forall r_2, r_2 \in instances(l_2, d) \Rightarrow \exists p \in m_{l_1, l_2} \wedge second(p) = r_2
+\end{align*}
+
 \section{Contribution}

### 2013-02-06 doc/paper.tex
reorganization of the paper

--- 2013-02-03 doc/paper.tex
@@
 \usepackage{semantic}
+\usepackage{caption}
+\usepackage{subcaption}
 
-\title{Refactoring of Software Using ORM: A Formal Description}
+\title{Refactoring of Software Using Relational Database}
 \author{Ondrej Macek}
@@
 Finally the applicability and limitations of the proposed solution is discussed in Sec.
-\ref{sec:contribution}.
+\ref{sec:case}.
 
@@
 \label{sec:problem}
+% víc rozepsat o změnách a jak se jim bránit a kam tím směřujeme
+\textbf{TODO: write an introduction when and why the software changes. and possible solutions.}
 A software implemented using an object-oriented language, which uses relational database as a data storage (as illustrated in the Figure \ref{fig:evolution}) is object of interest of this paper.
@@
 
+\textbf{TODO}
+
 \section{Model of Software}
@@
 %TODO consistency
+
+\subsubsection{Application Manipulation}
+\label{sec:app-evolution}
+The application model defines only the structure of an application thus there are defined transformation for its creating, altering and disassembling.
+Therefore the conditions for successful transformations are simple 1) when creating an object in the model or altering its name collisions has to be prevent, 2) when renaming a class the references have to be updated 3) when creating or altering an association the existence of referenced class has to be verified.
+The only issue during disassembling is the class can be removed only in case it is not associated by other classes.
+If a transformation cannot be processed the state of the application become inconsistent.
+The transformations are in table \ref{tab:app-evolution}.
+The list of altering operations is not complete as there are missing transformation for changing the obligation or cardinality of properties and associations.
+It is because these transformations are not so important for the next text when the complex refactorings are considered.
+\begin{table}
+\centering
+	\begin{tabular}{|l|}
+	\hline
+	\textbf{Application Creation} \\
+	newProperty(className, property, application) \\
+	newAssociation(className, association, mapping, application) \\
+	newClass(class, application) \\
+	\textbf{Application Alternation} \\
+	renameProperty(className, oldName, newName, application) \\
+	renameAssociation(className, oldName, newName, application) \\
+	renameClass(oldName, newName, application)\\
+	changeReference(referncedClass, newReferencedClass) \\
+	\textbf{Application Disassembly} \\
+	removeProperty(className, propertyName, application) \\
+	removeAssociation(className, associationName, application) \\
+	removeClass(className, application)\\
+	\textbf{Copy Structure}\\
+	copyProperty(sourceClass, property, targetClass, newName, application)  \\
+	copyClass(class, newName, application) \\
+	\hline
+	\end{tabular}
+	\caption{The transformations for application evolution.}
+	\label{tab:app-evolution}
+\end{table}
 
@@
 
+\subsubsection{Database Manipulation}
+\label{sec:db-evolution}
+The database consists of two parts - from database schema which defines the structure and from stored data - hence the transformations have to consider both parts.
+The transformation for manipualtion of structure has similar conditions to the evolution of application application.
+The transformations for data manipulation are inspired by the SQL language.
+The basic transformations are in Table \ref{tab:db-basic-evolution}.
+\begin{table}
+\centering
+	\begin{tabular}{|l|}
+	\hline
+	\textbf{Database Creation} \\
+	addColumn(tableName, column, database) \\
+	addForeignKey(tableName, key, database) \\
+	addTable(table, database)\\
+	\textbf{Database Alternation} \\
+	alterColumnName(tableName, oldName, newName, dataabse) \\
+	alterForeignKeyName(tableName, oldName, newName, dataabse) \\
+	alterTableName(oldName, newName, dataabse) \\
+	\textbf{Database Disasembling} \\
+	dropColumn(tableName, columnName, database) \\
+	dropForeignKey(tableName, keyName, database) \\
+	dropTable(tableName, database) \\
+	\textbf{Data Manipulation} \\
+	select(tableName, rowId, database) \\
+	select(tableName, database) \\
+	insert(row, database) \\
+	insert(rowId, value, database) \\
+	delete(tableName, rowId, database) \\
+	delete(tableName, rowId, columnName, database) \\
+	\textbf{Copy Structure} \\
+	copyColumnStructure(sourceTableName, targetTableName, \\  \hspace{0.5in} sourceColumnName, newName, database) \\
+	copyTableStructure(sourceTable, newTableName, database) \\
+	\textbf{Copy Strucutre and Values} \\
+	copyColumn(sourceTable, targetTable, \\
+	\hspace{0.5in} sourceColumnName, targetColumnName, \\ 
+	\hspace{0.5in} mapping, database) \\
+	copyTable(sourceTable, newName, database) \\
+	\hline
+	\end{tabular}
+	\caption{The transformations for database evolution.}
+	\label{tab:db-basic-evolution}
+\end{table}
+The set of transformations for database manipulation is limited in contrast with the SQL language it is because mentioned transformation are necessary for application of data evolution on database level.
+
+\paragraph{Copying of Database Elements} The transformations for copying the structure of a database elements serves more as helpers in advanced evolution cases as they adds a structural duplicate of an existing element only, nevertheless they are valid part of the transformation for database manipulation.
+The copying of structural elements has the same limits as database creation - the transformation can not cause name collisions.
+
+The mapping between instances of the classes (stored data) to provides information which value from the source table rows will be copied to which row in target table.
+The value of a column is then inserted for each row into the copied column during $copyColumn$ transformation.
+The mapping is usually defined as: 
+\begin{align*}
+&	mapping =  (r_1, r_2)*,\\
+& 	r_1 \in select(ORM(sourceName), database) \cup \varnothing, \\
+& 	r_2 \in select(ORM(targetName), database) \cup \varnothing
+\end{align*}
+The mapping is not limited to the bijection and it is possible to not include a row into a mapping - this has to be verified because of possible $NOTNULL$ constraints on a target column, as well as the $UNIQUE$ constraint has to be verified because the mapping can violate it in general.
+On the other hand the benevolent mapping help with creating various kinds on advanced evolutionary transformations.
+The mapping can cause data losses in some cases if used as a part of advanced refactoring.
+This situation is discussed later on.
+
+\paragraph{Data Secure Disasembling of a Software} The transformation for disassembly the database can have fatal impact on the data preservation, therefore the set of transformations is extended by a data secure transformations for database disassembling.
+These transformations are not part of the SQL standard, although they can be implemented as database functions.
+These transformations creates a secure way to disassembly the database as they drops empty structural elements only.
+These transformations are in table \ref{tab:db-adv-evolution}.
+\begin{table}
+\centering
+	\begin{tabular}{|l|}
+	\hline
+	\textbf{Data Secure Database Disassembling} \\
+	dropEmptyColumn(tableName, columnName, database) \\
+	dropEmptyForeignKey(tableName, keyName, database) \\
+	dropEmptyTable(tableName, database) \\
+	\hline
+	\end{tabular}
+	\caption{The data secure transformations for database evolution which can be created as SQL functions.}
+	\label{tab:db-adv-evolution}
+\end{table}
+
 \subsection{Object Relational Mapping}
@@
 
-\section{Data Evolution of Software}
-\label{sec:evolution}
-The evolution of the software is mapped to the layer specific transformation, which are defined in Sec.
-\ref{sec:app-evolution} and \ref{sec:db-evolution}.
-Finally the evolution cases and their mapping on application and database components are presented in \ref{sec:sw-evolution}.
-The transformations are divided into groups according to their impact on structure and data.
+\textbf{TODO: Describe ORM}
+
+\subsection{Software Evolution}
+\label{sec:sw-evolution}
+The evolution of the whole software consider its both components application and database containing data.
+The way how evolutionary transformation are created based on the set of component specific transformations is introduced in this chapter.
+
+The evolution is described from the developer point of view, therefore the transformations use the names and elements from application context.
+The names of application elements and elements themselves have to be transformed into a database context - this is assured by the ORM.
+
+%Finally the evolution cases and their mappings on application and database components are presented in \ref{sec:sw-evolution}.
 
@@
 
-\subsection{Application Evolution}
-\label{sec:app-evolution}
-The application model defines only the structure of an application thus there are defined transformation for its creating, altering and disassembling.
-Therefore the conditions for successful transformations are simple 1) when creating an object in the model or altering its name collisions has to be prevent, 2) when renaming a class the references have to be updated 3) when creating or altering an association the existence of referenced class has to be verified.
-The only issue during disassembling is the class can be removed only in case it is not associated by other classes.
-If a transformation cannot be processed the state of the application become inconsistent.
-The transformations are in table \ref{tab:app-evolution}.
-The list of altering operations is not complete as there are missing transformation for changing the obligation or cardinality of properties and associations.
-It is because these transformations are not so important for the next text when the complex refactorings are considered.
+\subsection{Basic Evolutionary Transformations}
+\label{sec:sw-basic-evolution}
+The evolution of the whole software is based on the atomic transformations specific for each software part, which are defined in Sec.
+\ref{sec:app-evolution} and \ref{sec:db-evolution}.
+This section introduces how these primitives can be used to manipulate the software elements.
+
+\paragraph{Basic Software Evolution} The basic evolutionary transformations of the software are based on basic evolution of an application.
+These transformation cannot be  mapped directly on the evolutionary transformations of the database, because properties and associations can be mapped as a column (foreign key respectively) or as a table, this has to be reflected during the evolution as shown on the example of creating a new property:
+\begin{align*}
+& \Phi(newProperty(className, property, application)) = \\
+& \;\;\; = \begin{cases}
+  addColumn(ORM(className), ORM(property), database) \\\mathbf{if} \; cardinality(property \leq 1)  \\\\
+  addTable(ORM(property), database) \\
+  \mathbf{if} \; cardinality(property > 1)  
+ \end{cases}
+\end{align*}
+
+Mapping between basic evolutionary transformations is provided in the shortened way in table \ref{tab:sw-basic-evolution}, there is the basic application transformation and possible mappings on database level according to to the cardinality.
+
 \begin{table}
 \centering
-	\begin{tabular}{|l|}
+	\begin{tabular}{|p{0.4\textwidth} p{0.5\textwidth}|}
 	\hline
-	\textbf{Application Creation} \\
-	newProperty(className, property, application) \\
-	newAssociation(className, association, application) \\
-	newClass(class, application) \\
-	\textbf{Application Alternation} \\
-	renameProperty(className, oldName, newName, application) \\
-	renameAssociation(className, oldName, newName, application) \\
-	renameClass(oldName, newName, application)\\
-	\textbf{Application Disassembly} \\
-	removeProperty(className, propertyName, application) \\
-	removeAssociation(className, associationName, application) \\
-	removeClass(className, application)\\
+	Application Transformation & Database Transformation \\
+	\multicolumn{2}{|c|}{\textbf{Software Creation}} \\
+	newProperty & addColumn  ($cardinality \leq 1$)\\
+	& addTable \\ 
+	& \\
+	newAssociation & addColumn ($cardinality \leq 1$)\\
+	& addTable \\
+	& \\
+	newClass & addTable \\
+	\multicolumn{2}{|c|}{\textbf{Application Alternation}} \\
+	renameProperty & renameColumn ($cardinality \leq 1$)\\
+	& renameTable \\
+	& \\
+	renameAssociation & renameForeignKey ($cardinality \leq 1$)\\
+	& renameTable \\
+	& \\
+	renameClass & renameTable \\
+	\multicolumn{2}{|c|}{\textbf{Application Disassembly}} \\
+	removeProperty & removeColumn ($cardinality \leq 1$)\\
+	& removeTable \\
+	& \\
+	removeAssociation & removeForeignKey ($cardinality \leq 1$)\\
+	& removeTable \\
+	& \\
+	removeClass & removeTable\\
 	\hline
 	\end{tabular}
-	\caption{The transformations for application evolution}
-	\label{tab:app-evolution}
+	\caption{The mapping between evolution of the application to the database transformations.}
+	\label{tab:sw-basic-evolution}
 \end{table}
 
-\subsection{Database Evolution}
-\label{sec:db-evolution}
+The data secure transformations for disassembling the database can succeed (e.g. drop a table) when there are no data contained within the dropped element.
 
-\subsection{Software Evolution}
-\label{sec:sw-evolution}
+\subsection{Advanced Evolutionary Transformations}
+\label{sec:sw-adv-evolution}
+Advanced evolutionary transformations are based on the basic ones, they can be obtained as a concatenation of transformations.
+The concatenation of transformation is possible because each transformation produces a state of a software (or application or database) which can be used as an input for the next transformation.
 
-\section{Contribution}
-\label{sec:contribution}
+All transformations presented so far has the same input information for both application and database, whereas the advanced transformations usually needs a mapping between stored data (instances) as theirs input.
+
+\paragraph{Copy Property}
+The $copyProperty$ creates a duplicate of a property in a given class.
+It can rename the property so it can be used for creating duplicates of a property in the property owning class.
+If the mapping is not provided the transformation creates only a structural copy of the property otherwise it copies the values too.
+Therefore the  $copyProperty$ transformation is the simplest way to manipulate stored data.
+
+The  $copyProperty$ is the first transformation when an additional information has to be added to the usual code refactoring.
+It is because the $copyColumn$ transformation needs one more information to succeed - there has to be provided mapping between the source and target table to assure the data information consistency.
+%Moreover the source and target should have the same structure, type and constraints.
+\begin{align*}
+& \mathbf{copyProperty}(sourceClass, propertyName, targetClass, \\
+& \; \; \; mapping, software) 
+\end{align*}
+The transformation is interpreted for both software components - in case of the application a new property is added:
+\begin{align*}
+& \Psi(copyProperty(sourceClass, property, targetClass, mapping, \\ 
+&  \; \; \; software)) = newProperty(targetClass, property, application) 
+\end{align*}
+in case of database the copy of a column or table is created according to property's cardinality and then values are copied:
+\begin{align*}
+& \Phi(copyProperty(sourceClass, property, targetClass, newName,  \\ 
+&  \; \; \; mapping, software)) = \begin{cases}
+ coypColumn(ORM(sourceClass),  \\ \;\;\; ORM(targetClass),  \\ \;\;\;alterColumnName(ORM(property), \\ \;\;\; ORM(newName)), mapping, database) \\ \; \mathbf{if} \; cardinality(property) \leq 1 \\\\
+ copyTable(ORM(Property), \\ \;\;\; ORM(newName))
+ \end{cases}
+\end{align*}
+
+If there are no data in the target table an identity can be used as a mapping function instead of mapping above.
+
+\paragraph{Move Property}
+The $moveProperty$ transformation is based on the $copy$ $Property$ transformation followed by the $removeProperty$ so its construction is easy.
+On the other hand special attention has to be paid to the provided mapping of instances, because it can cause data loses.
+The ideal case is when the mapping is a bijection among instances, then the transformation can not cause the lost of data.
+On the other hand the situation when the mapping does not provides pairs for all rows in a source table can lead to data loss.
+\begin{align*}
+& moveProperty(sourceClass, property, targetClass, newName, \\
+& \;\;\; mapping, software) = removeProperty(sourceClass, property, \\
+& \;\;\; CopyProperty(sourceClass, property, targetClass, \\ 
+& \;\;\;newName, mapping, software))
+\end{align*}
+
+\paragraph{Merge Classes}
+There occur situations when there is a need to merge two classes and the inlining transformation cannot be used because of name collisions among properties in such a case the $merge$ transformation can be used.
+It can be used in case structure of both classes (respectively tables) are equal.
+The transformation creates new rows in the new table - the primary keys from origin instances are replaced by a new ones.
+\begin{align*}
+& merge(firstClass, secondClass, newClassName, software) = \\
+& \; \; \; = removeClass(secondClass, removeClass(firstClass, \\
+& \; \; \; changeReference(secondClass, newClassName, \\
+& \; \; \; changeReference(firstClass, newClassName, \\
+& \; \; \; copyClass(firstClass, newClassName, software))))))
+\end{align*}
+
+\paragraph{Extract and Inline Class}
+The extract and inline class are two opposite transformations. First of them extract a class from an existing class, second of them moves data from one class into another.
+\begin{align*}
+& extract(sourceClass, newClass, property*, software) = \\ 
+& \; \; \; = moveProperty(sourceClass, newClass, property*, \\ 
+& \; \; \; addClass(newClass, software))
+\\\\
+& inline(targetClass, classToInline, software)) = \\
+& \; \; \; = removeClass(classToInline, moveProperty(classToInline, \\
+& \; \; \; targetClass, properties(classToInline), software))
+\end{align*}
+
+The advanced refactoring cases are limited by the transformations they are created from.
+
+\section{Case Study}
+\label{sec:case}
+It is obvious that the set of defined transformation is limited because of focus on data preservation or transformation concatenation.
+On the other hand defined transformations are strong enough to handle a lot of refactoring cases.
+
+Before one complex case study example is shown a set of small refactoring cases is discussed.
+These cases are chosen based on the Eclipse foundation refactoring statistics (the transformation from the top ten influencing data chosen only) and on Martin Fowler's book: Refactoring - improving design of existing code, where one chapter is dedicated to the data organization in the software.
+
+Selected refactorings are:
+\begin{enumerate}
+	\item \textbf{Rename} is the most used refactoring used according to Eclipse statistic and it is one of the basic refactoring cases introduced in the Sec. \ref{sec:sw-basic-evolution}.
+	\item  \textbf{Move} refactoring is used in Eclipse to move properties from a class to another class within an inheritance hierarchy or to move classes between packages.
+In contrast our model does not consider packages, on the other hand it is able to move property from one class to another according given mapping, which contains not only move in hierarchy but a lot of other cases too.
+	\item \textbf{Extract Class} is often used refactoring in Eclipse and it is mentioned in Fowler's book too.
+It is introduced as an example of advanced refactorings in Sec.
+\ref{sec:sw-adv-evolution} together with its opposite transformation $inline$.
+	\item \textbf{Move field} from the Fowler's book is introduced as $moveProperty$ in Sec. \ref{sec:sw-basic-evolution}.
+	\item \textbf{Replace Data Value with Object} is a refactoring case which is not mentioned among the evolutionary transformations, on the other hand it can be composed from already defined transformations:
+	\begin{align*}
+& replaceDataWithObject(sourceClass, property, newObject, \\
+& \; \; \; software) = removeProperty(sourceClass, property, \\
+& \; \; \; newAssociation(sourceClass, newObject,  mapping, \\
+& \; \; \; copyProperty(sourceClass, property, newObject, \\ & \; \; \;  software)
+	\end{align*}
+Where the mapping use the values copied in the first transformation as a matching point for creating connection between the old and the new class.
+\end{enumerate}
+The transformations proposed in this paper are able to implement all data refactorings used on Eclipse platform.
+There are some refactorings from Fowler which are not considered in the proposal: replace array with object, change unidirectional association to bidirectional (and reverse) and operations which replace types with polymorphism.
+
+\subsection{Example of Usage}
+The example demonstrates how the transformation helps with data evolution in real example of software evolution.
+There are two classes in our software $Person$ and $LegalParty$ both of them contains information about address (street, city and zip code).
+The situation in software is in Figure \ref{fig:case1}, where the whole software is shown.
+
+To improve the design of code the class Address has to be created which associated with both original classes and contains the data already stored in the software. 
+\begin{figure}
+\begin{subfigure}[b]{0.45\textwidth}
+	\includegraphics[width=\textwidth]{./images/case_app_1}
+	\caption{The application model.}
+\end{subfigure}
+\quad
+\begin{subfigure}[b]{0.45\textwidth}
+	\includegraphics[width=\textwidth]{./images/case_db_1}
+	\caption{The database model.}
+\end{subfigure}
+\begin{subfigure}[b]{\textwidth}
+	\centering
+	\begin{tabular}{| c | l | l | l | l | }
+	 	\hline
+		Id &  Surname & Street & City & ZIP  \\ \hline  
+		11 & Jackson & Central Park St & New York & 100 01  \\ \hline
+		12 & Clooney & S Orange Ave & Orlando & 320 24  \\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $Person$.}
+\end{subfigure}
+\begin{subfigure}[b]{\textwidth}
+	\centering
+	\begin{tabular}{| c | l | l | l | l | c |}
+	 	\hline
+		Id &  BusinessName & Street & City & ZIP \\ \hline  
+		100 & Tools \& Machines & Olive ave. & New York & 100 01 \\ \hline
+		200 & AI Robotics & Pine ave. & LA & 900 03  \\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $LegalParty$.}
+\end{subfigure}
+	\caption{The initial state of the example used in the case study. There is a repetition of information structure in the $Person$ and $LegalParty$ class.}
+	\label{fig:case1}
+\end{figure}
+The first  step of is to extract two temporary classes representing address of a $Person$ or of a $LegalParty$. Temporary classes has to be connected with the origin classes by an association.
+\begin{align*}
+ s' &= merge("Address\_tmp1", "Address\_tmp2", "Address", \\
+& \;\;\;\;  newAssociation("LegalParty", association("address", \\
+& \;\;\;\; "Address\_tmp2", 1, 1), mapping_2, extract("LegalParty", \\
+& \;\;\;\;  "Address\_tmp2", "street", "city", "zip", \varnothing  \\
+& \;\;\;\; newAssociation("Person", association("address", \\
+& \;\;\;\; "Address\_tmp1", 1, 1), mapping_1, extract("Person", \\
+& \;\;\;\; "Address\_tmp1", "street", "city", "zip", \varnothing, software)))
+\end{align*}
+Both $mapping_1$ and $mapping_2$ are identities so the creation of the association is straightforward. The result of the transformations is in the Fig. \ref{fig:case2}. 
+\begin{figure}
+\begin{subfigure}[b]{0.45\textwidth}
+	\centering
+	\includegraphics[width=\textwidth]{./images/case_app_9}
+	\caption{The application model.}
+\end{subfigure}
+\quad
+\begin{subfigure}[b]{0.45\textwidth}
+	\centering
+	\includegraphics[width=\textwidth]{./images/case_db_8}
+	\caption{The database model.}
+\end{subfigure}
+
+\begin{subfigure}[b]{0.5\textwidth}
+	\centering
+	\begin{tabular}{| c | l |}
+	 	\hline
+		Id &  Surname \\ \hline  
+		11 & Jackson  \\ \hline
+		12 & Clooney  \\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $Person$.}
+\end{subfigure}
+\begin{subfigure}[b]{0.5\textwidth}
+	\centering
+	\begin{tabular}{| c | l |}
+	 	\hline
+		Id &  BusinessName  \\ \hline  
+		100 & Tools \& Machines  \\ \hline
+		200 & AI Robotics \\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $LegalParty$.}
+\end{subfigure}
+\begin{subfigure}[b]{\textwidth}
+	\centering
+	\begin{tabular}{| c | l | l | l | l | l |}
+	 	\hline
+		Id & Street & City & ZIP & Person\_Address & LegalParty\_Address\\ \hline  
+		1 & Olive ave. & New York & 100 01 & & 100 \\ \hline
+		2 & Pine ave. & LA & 900 03 & & 200  \\ \hline
+		3 & Central Park St & New York & 100 01 & 11 &  \\ \hline
+		4 & Orange Ave & Orlando & 320 24 & 12 &\\ \hline
+	\end{tabular}
+	\caption{Data stored in the table $Address\_tmp2$.}
+\end{subfigure}
+	\caption{The initial state of the example used in the case study. There is a repetition of information structure in the $Person$ and $LegalParty$ class.}
+	\label{fig:case2}
+\end{figure}
+
+The example shows that defined transformations are able to perform regular data evolutions.
+It shows that usage of the transformations are sometimes not intuitive for user (e.g.
+creating temporary classes before merge).
 

### 2013-02-06 doc/related_work.tex
extended related work - evolution in OODB

@@
 \section{Related Work}
+
+\subsection{Database Evolution}
+\subsubsection{Evolution in Objec-Oriented Databases}
+The problem of database schema evolution is well known issue which is discussed for a long time.
+There is proposed an approach for  schemes of object-oriented databases in \cite{Banerjee:SemanticsOfSchemaEvolutionInOODB} which based on a set of schema invariants (representing the properties of the schema) and a set of rules for invariant preserving schema changes.
+The impact of evolution on instances is considered too.
+The approach is presented informally, while our approach is based on set of formal definitions.
+
+The formal definition of evolution of object-oriented databases is presented in \cite{Peters:AxiomaticModelOfDynamicEvolutionInOODB}.
+The approach is based on type system representing the database schema, on axioms which defines the schema structural constraints and on changes which evolves the system.
+There are three change operators: add, remove, modify defined in the paper.
+The model satisfies the properties of soundness and completeness.
+The approach in this paper is based on small transformations (changes) too, but its aim is to combine them in a way advanced refactoring cases can be easily applied on a software.
+
+The SERF framework \cite{Claypool:SERF} provides a similar functionality to the user of an object-oriented database as we do for the user of a software based on ORM.
+The approach is based on templates a user can create and maintain.
+Limitation and condition of some evolution cases is discussed in \cite{Claypool:SERF-formal}.
+
+All mentioned approaches address similar  issue as this paper, in contrast it focusses on the homogenous environment of object-oriented databases, whereas our approach aims the heterogenous world of ORM.
+
+The formal approach based on epistemic logic (logic of knowledge) is introduced in \cite{Chang-LogicFrameworkForDbRefactoring}.
+The approach presents a database as a set of beliefs (knowledge), which is revision can be seen as refactoring.
+The approach consider not only database structure but database queries too, on the other hand data are not considered in this approach.
+
+\subsection{Data Evolution Metrics}
 There are proposed several metrics for classification of impact of a change on the conceptual schema in \cite{Wedemeijer:ConceptualMetrics}.
@@
 
-\paragraph{Prediction of database change impact}The analysis technique for identifying the impact of relational database change upon object oriented application is proposed in \cite{Maule:ImpactAnalysis}.
+\subsection{Impact of Database Evolution on Application}
+\paragraph{Prediction of database change impact on application}The analysis technique for identifying the impact of relational database change upon object oriented application is proposed in \cite{Maule:ImpactAnalysis}.
 The technique focus on relational database and its change which impact on application is predicted.
-The analysis technique is able to identify changes in application structure (new attributes) as well as changes in database queries used in application.
+The analysis technique is able to identify changes in application structure as well as changes in database queries used in application.
 In contrast approach proposed in this paper does not consider database queries and theirs change and it is not mentioned to predict the impact of a change, but to

### 2013-02-07 doc/related_work.tex
finished OODB and RDB databases

@@
 \subsection{Database Evolution}
-\subsubsection{Evolution in Objec-Oriented Databases}
-The problem of database schema evolution is well known issue which is discussed for a long time.
+The problem of database schema evolution is well known issue which is discussed for a long time, nevertheless the term "database refactoring" is quite new.
+This section provides an overview of significant related work in area of database evolution and refactoring.
+
+\subsubsection{Evolution of Relational Database}
+\paragraph{Informal definitions} The taxonomy of relational database evolution based on the entity-relationship model  is proposed in \cite{Roddick:TaxonomyOnERM}.
+The evolution is described as a change in entity-relationship model and change in relational database.
+The semantics change patterns in context of conceptual schema is described in \cite{Wedemeijer:SemanticChangePatternsInConceptualSchema}, although its impact on database schema or data is not described.
+The main cases of data evolution are defined in both publications, however the description is informal.
+The extensive set of possible database refactorings is provided in \cite{Ambler:DbRefactoringBook}, where both schema and data evolution is discussed.
+The refactorings are intended to be used by database administrators, thus it assume database-first approach to evolution whereas this proposal is application-first.
+
+\paragraph{Formal framework} A general formal framework for database evolution is defined in \cite{McBrien:formal-framework-transformation}.
+The framework is based on a set of basic graph transformations which are then extended to transformations of the entity-relationship model.
+The framework and the defined transformations can be used for comparison with our framework, where our framework consider stored data.
+The contribution of the formal framework is definition of equivalent structures in schemas.
+
+\paragraph{Tools for Database Evolution} The MeDEA project \cite{Meda:Main}, \cite{Meda:MDDViewBasedEv} offers a tool for evolution of both database schema and stored data based on model-driven approach.
+Both MeDEA and proposed framework allows automatic database evolution, the difference is in different target audience.
+The framework is aimed to be used by database administrators, therefore it is aimed on database structures (such as views) which are not considered in our approach as it is aimed to be used by developers.
+Proposed framework can use MeDEA to enrich its ability on database model, whereas MeDEA can use the code-centric point of view to design developer friendly transformations.
+
+The PRISM is a research project for data management under schema evolution \cite{PRISM} in contrast with our proposal and with MeDEA it uses it extends the SQL command set by so called schema modification operators which implements the schema evolution.
+The project is aimed to  be used by database administrators (as well as MeDEA).
+The PRISM's set of modification operators could be used for database transformation in our proposal, however we prefer to work with a standard SQL set of commands.
+
+\paragraph{ORM Frameworks} here are many object-relational mapping frameworks available for developers, and some of them provide tools for database migration.
+Hibernate \cite{Hibernate} is one of the most popular ORM frameworks in the Java community.
+It provides customizable ORM for a wide range of databases, however it does not support complex database migrations.
+It is capable only of creating a new table or adding a new column, hence it is not possible to, for example, drop a table or copy values from one column to another.
+Active Record \cite{Active_Record} is an ORM framework in the Ruby on Rails environment.
+Since its first version it has contained support for database evolution according to the create-update-delete principle, in the form of so-called migrations \cite{Rails:Migrations} which can be extended by adding user SQL commands.
+Entity Framework \cite{Entity_Framework} is Microsoft's ORM solution for the .NET platform.
+Its capabilities of data evolution support are similar to those of Active Record.
+
+Each of presented ORM frameworks provides support for creating data evolution, as they are all capable of adding a new class as an entity and creating a corresponding table in the database.
+On the other hand, only two of them (Active Record and Entity Framework) are capable of updating and deleting evolution.
+In the case of evolutions which need to manipulate data, e.g.
+moving the property, none of the frameworks mentioned provides a built-in solution.
+Active Record and Entity Framework allow developers to describe the data move manually.
+In contrast to these ORMFs, our proposal  provides a higher level of user comfort because of its capability to handle complex data evolutions.
+
+\subsubsection{Evolution of Object-Oriented Databases}
+The issue of database evolution is discussed not only for relational databases, but for object-oriented databases too.
 There is proposed an approach for  schemes of object-oriented databases in \cite{Banerjee:SemanticsOfSchemaEvolutionInOODB} which based on a set of schema invariants (representing the properties of the schema) and a set of rules for invariant preserving schema changes.

### 2013-02-12 doc/paper.tex
added todos
related work

@@
 \label{sec:related-work}
+Research of data and database evolution is not limited to relational ones only, there are also works in the areas of object databases and XML databases.
+These works provide solutions specific to concrete types of databases using various ranges of solutions - domain specific languages \cite{SERF}, extensions of existing standards or MDD \cite{Evolution_vs_Reorganization} or formal specification \cite{Tresch:1991vi}.
+These solutions are inspiring, however the domain of the ORM has its specific issues, so a solution from another domain has to be adapted carefully.
 
-\textbf{TODO}
+\paragraph{Informal definitions} The taxonomy of relational database evolution based on the entity-relationship model  is proposed in \cite{Roddick:TaxonomyOnERM}.
+The evolution is described as a change in entity-relationship model and change in relational database.
+The semantics change patterns in context of conceptual schema is described in \cite{Wedemeijer:SemanticChangePatternsInConceptualSchema}, although its impact on database schema or data is not described.
+The main cases of data evolution are defined in both publications, however the description is informal.
+The extensive set of possible database refactorings is provided in \cite{Ambler:DbRefactoringBook}, where both schema and data evolution is discussed.
+The refactorings are intended to be used by database administrators, thus it assume database-first approach to evolution whereas this proposal is application-first.
+
+\paragraph{Formal framework} A general formal framework for database evolution is defined in \cite{McBrien:formal-framework-transformation}.
+The framework is based on a set of basic graph transformations which are then extended to transformations of the entity-relationship model.
+The framework and the defined transformations can be used for comparison with our framework, where our framework consider stored data.
+The contribution of the formal framework is definition of equivalent structures in schemas.
+
+A categorical framework for migration of object-oriented systems is proposed in \cite{Schulz:CategoricalModel}.
+This framework defines the refactoring of objects, data and methods, which are the main objectives of the framework.
+The impact of the object change on a relational database is not considered in the paper as it is aimed at object-oriented systems only.
+
+\paragraph{Tools for Database Evolution} The MeDEA project \cite{Meda:Main} offers a tool for evolution of both database schema and stored data based on model-driven approach.
+Both MeDEA and proposed framework allows automatic database evolution, the difference is in different target audience.
+The framework is aimed to be used by database administrators, therefore it is aimed on database structures (such as views) which are not considered in our approach as it is aimed to be used by developers.
+Proposed framework can use MeDEA to enrich its ability on database model, whereas MeDEA can use the code-centric point of view to design developer friendly transformations.
+
+The PRISM is a research project for data management under schema evolution \cite{PRISM} in contrast with our proposal, and with MeDEA it extends the SQL command set by so-called schema modification operators which implements the schema evolution.
+The project is aimed to  be used by database administrators (as well as MeDEA).
+The PRISM's set of modification operators could be used for database transformation in our proposal, however we prefer to work with a standard SQL set of commands.
+
+The project DB-MAIN \cite{Hick:DBMAIN} provides a MDD approach to data evolution on all the levels of software we do.
+The project is well documented formally.
+The difference between DB-MAIN and our framework is that each has a different focus.
+DB-MAIN starts with database modification and database migration, and entities are created accordingly; whereas our focus is on entities evolution, which is then propagated to a database with an emphasis on automatization.
+Our goal is to hide the entire database level from our users.
+
+\paragraph{ORM Frameworks} here are many object-relational mapping frameworks available for developers, and some of them provide tools for database migration.
+Hibernate \cite{Hibernate} is one of the most popular ORM frameworks in the Java community.
+It provides customizable ORM for a wide range of databases, however it does not support complex database migrations.
+It is capable only of creating a new table or adding a new column, hence it is not possible to, for example, drop a table or copy values from one column to another.
+Active Record \cite{Active_Record} is an ORM framework in the Ruby on Rails environment.
+Since its first version it has contained support for database evolution according to the create-update-delete principle, in the form of so-called migrations \cite{Rails:Migrations} which can be extended by adding user SQL commands.
+Entity Framework \cite{Entity_Framework} is Microsoft's ORM solution for the .NET platform.
+Its capabilities of data evolution support are similar to those of Active Record.
+
+Each of presented ORM frameworks provides support for creating data evolution, as they are all capable of adding a new class as an entity and creating a corresponding table in the database.
+On the other hand, only two of them (Active Record and Entity Framework) are capable of updating and deleting evolution.
+In the case of evolutions which need to manipulate data, e.g.
+moving the property, none of the frameworks mentioned provides a built-in solution.
+Active Record and Entity Framework allow developers to describe the data move manually.
+In contrast to these ORMFs, our proposal  provides a higher level of user comfort because of its capability to handle complex data evolutions.
+
+\paragraph{Language Specific Solution} An evolution approach for COBOL language is described in \cite{Cleve:2006ug}, which allows a database schema change to be propagated into stored data and entities in the COBOL language.
+The evolution described in this paper are for create, update and delete.
+In contrast, our framework is platform independent and covers a wider range of evolution cases including data manipulation; it is aimed at being used by software developers, and thus it starts on the entities level.
+
+Very promising solution is Liquibase \cite{Liquibase} a tool for database refactoring and evolution.
+It is capable to migrate both database schema and stored data.
+The evolution is described in a form of a XML document which can be interpreted on various databases.
+The difference is we try not to bother the developer with database-specific details and make all changes according to a change in application code.
+
+\paragraph{Forward Round-Trip Engineering in Data Evolution} A round-trip approach for data evolution was described in \cite{VanPaesschen:2005to}, which was implemented in the SELF language.
+This approach proposed a forward-oriented evolution on all application levels in the software application.
+In contrast with our framework, the round-trip approach does not care about stored data because it is focused only on create, update and delete types of evolution.
+
+A meta-model based approach to data evolution is proposed in\cite{Aboulsamh:MetaModelBasedApproachToIsDataEvolution} and very similar solution for data evolution is proposed, which is based on extended UML meta-model.
+The solution provides similar capabilities of change of application and database as does our framework.
+In contrast our work is created with respect to ORM domain and therefore we extended the application meta-model with constructs typical for this domain.
 
@@
 
+\bibliographystyle{plain}      % basic style, author-year citations
+\bibliography{references}   % name your BibTeX data base
+
 \end{document}

### 2013-02-12 doc/paper.tex
shortened related work, compact models

@@
 \paragraph{Tools for Database Evolution} The MeDEA project \cite{Meda:Main} offers a tool for evolution of both database schema and stored data based on model-driven approach.
-Both MeDEA and proposed framework allows automatic database evolution, the difference is in different target audience.
-The framework is aimed to be used by database administrators, therefore it is aimed on database structures (such as views) which are not considered in our approach as it is aimed to be used by developers.
-Proposed framework can use MeDEA to enrich its ability on database model, whereas MeDEA can use the code-centric point of view to design developer friendly transformations.
-
-The PRISM is a research project for data management under schema evolution \cite{PRISM} in contrast with our proposal, and with MeDEA it extends the SQL command set by so-called schema modification operators which implements the schema evolution.
-The project is aimed to  be used by database administrators (as well as MeDEA).
-The PRISM's set of modification operators could be used for database transformation in our proposal, however we prefer to work with a standard SQL set of commands.
+The PRISM is a research project for data management under schema evolution \cite{PRISM} in contrast with our proposal, and with MeDEA, it extends the SQL command set by so-called schema modification operators which implements the schema evolution.
 
@@
 Our goal is to hide the entire database level from our users.
+
+All mentioned project is aimed to  be used by database administrators
 
@@
 Its capabilities of data evolution support are similar to those of Active Record.
-
-Each of presented ORM frameworks provides support for creating data evolution, as they are all capable of adding a new class as an entity and creating a corresponding table in the database.
-On the other hand, only two of them (Active Record and Entity Framework) are capable of updating and deleting evolution.
-In the case of evolutions which need to manipulate data, e.g.
-moving the property, none of the frameworks mentioned provides a built-in solution.
-Active Record and Entity Framework allow developers to describe the data move manually.
-In contrast to these ORMFs, our proposal  provides a higher level of user comfort because of its capability to handle complex data evolutions.
-
-\paragraph{Language Specific Solution} An evolution approach for COBOL language is described in \cite{Cleve:2006ug}, which allows a database schema change to be propagated into stored data and entities in the COBOL language.
-The evolution described in this paper are for create, update and delete.
-In contrast, our framework is platform independent and covers a wider range of evolution cases including data manipulation; it is aimed at being used by software developers, and thus it starts on the entities level.
 
@@
 \begin{align*}
-& \mathbf{Application} = (Class*)
+& \mathbf{Application} = (Class*) \\
+& 	\mathbf{Class} = (label, Property*, Association*) \\
+& \mathbf{Property} = (label, AppType, DefaultValue, Cardinality, Mandatory) \\
+&	\mathbf{Association} = (label, classRef, startCardinality, endCardinality)  \\
+& \mathbf{AppType} = APPSTRING
 \end{align*}
 
-\paragraph{Class} Class represents a basic structural unit in the application model. It has a unique name, one or more properties and it can be associated to other classes in the application. 
-\begin{align*}
-& 	\mathbf{Class} = (label, Property*, Association*)
-\end{align*}
+\paragraph{Class} Class represents a basic structural unit in the application model.
+It has a unique name, one or more properties and it can be associated to other classes in the application.
 The properties of a class can be of two kinds - values or collections - according of theirs cardinality.
@@
 The property can be mandatory, can have a default value and according to its cardinality it can represent a single value or a collection of values.
-\begin{align*}
-&	\mathbf{Property} = (label, AppType, DefaultValue, Cardinality, Mandatory)
-\end{align*}
 
@@
 The cardinalities defines the multiplicity of both association ends.
-\begin{align*}
-&	\mathbf{Association} = (label, classRef, startCardinality, endCardinality) 
-\end{align*}
 
@@
 The only $AppType$ type is called $APPSTRING$.
-\begin{align*}
-& \mathbf{AppType} = APPSTRING
-\end{align*}
 
@@
 he relational database consists of database schema which defines the structure of the database and data, which in our ORM system represents stored instances. The database is defined as a tuple:
+
 \begin{align*}
-&	\mathbf{Database} = ( Table*, Row* )
+&	\mathbf{Database} = ( Table*, Row* )\\
+&	\mathbf{Table} = (label, primaryKey, Column*, ForeignKey*)\\
+&	\mathbf{Column} = (label, DbType, DefaultValue, Constraint*) \\
+&	\mathbf{ForeignKey} = (Label, Label, Constraint*) \\
+&	\mathbf{PrimaryKey} =  ( Label, Sequence ) \\
+&	\mathbf{DbType} = DBSTRING \; | \; DBINT\\
+&	\mathbf{Constraint} = NOTNULL \; | \; UNIQUE 
 \end{align*}
@@
 Rows in the table represents stored data.
-\begin{align*}
-&	\mathbf{Table} = (label, primaryKey, Column*, ForeignKey*)
-\end{align*}
 
 \paragraph{Column} Column defines data values and types which can be part of a table record.
-\begin{align*}
-&	\mathbf{Column} = (label, DbType, DefaultValue, Constraint*) 
-\end{align*}
 
@@
 The value of a foreign key is $\varnothing$ or a non-zero natural number.
-\begin{align*}
-&	\mathbf{ForeignKey} = (Label, Label, Constraint*) 
-\end{align*}
 
@@
 Primary key is always defined with constraints $NOTNULL$ and $UNIQUE$.
-\begin{align*}
-&	\mathbf{PrimaryKey} =  ( Label, Sequence ) \end{align*}
+
 The sequence generates values of primary key.
@@
 Next the $DBINT$ type is defined for the values of keys.
-\begin{align*}
-&	\mathbf{DbType} = DBSTRING \; | \; DBINT
-\end{align*}
 
@@
 Both constraints are column constraints - first constraint defines non-empty columns, second constraint defines there has to be unique records in a column or foreign key.
-\begin{align*}
-&	\mathbf{Constraint} = NOTNULL \; | \; UNIQUE 
-\end{align*}
 

### 2013-02-13 doc/paper.tex
new style, formatting, changes in related work

@@
-\documentclass[11pt]{article}
+\documentclass[runningheads]{comsis}
+
+\def\journalissue{ComSIS Vol.~0, No.~0, December 2009}
+\def\paperidnum{UDC, DOI: N/A}
+\setcounter{page}{1}
+
+\usepackage[dvipdfm]{hyperref}
+
 \usepackage{amssymb}
@@
 \title{Refactoring of Software Using Relational Database}
-\author{Ondrej Macek}
+\author{Ondrej Macek\inst{1}}
+\authorrunning{Ondrej Macek}
+\institute{Czech Technical University in Prague\\
+  Karlovo namesti 13 \\
+121 35 Praha 2, Czech Republic\\
+  \email{ondrej.macek@fel.cvut.cz}}
 
@@
 \maketitle
+
+\begin{abstract}
+
+\vspace{6pt}\textbf{Keywords:}
+\end{abstract}
+
 \section{Introduction}
@@
 There are three important components of such a software: application (its persistent layer concretely), database (consisting of database schema and stored data) and an ORM, therefore we define a software as a triple consisting of an application and a database which are connected together by an ORM:
-\begin{align*}
+\begin{align}
 & \mathbf{software} = ( Application, Database, ORM )
-\end{align*}
+\end{align}
 The software has to be in consistent state so its users could have benefit from its usage.
 The state of the software is defined to communicate the sate of the software to its user, the symbol $\perp$ is used as a notation for inconsistent state.
-\begin{align*}
+\begin{align}
 & \mathbf{state} = Consistent\; s \: | \: \perp, s \in State 
-\end{align*}
+\end{align}
 The software is in consistent state if application and database are consistent and the database structure corresponds to the application structure according to the ORM.
@@
 \centering
-	\includegraphics[scale=0.4]{./images/evolution_simple}
+	\includegraphics[width=.7\textwidth]{./images/evolution_simple}
 	\caption{The evolution of data changes the system on all levels. The figure shows all the components of the evolution process.}
@@
 The evolution is defined as:
-\begin{align*}
+\begin{align}
 & t(state(s)) = \begin{cases}
@@
 \end{cases}\\ 
-& s \in Software, t \in E
-\end{align*}
+& s \in Software, t \in E \nonumber
+\end{align}
 Where $\Psi$ maps the software evolution cases to the code refactoring and $\Phi$ to maps to the evolutionary transformation of database.
@@
 The information consistency means the data was transformed according to the evolution of the application and they correspond with developer's intention.
-
-\section{Related Work}
-\label{sec:related-work}
-Research of data and database evolution is not limited to relational ones only, there are also works in the areas of object databases and XML databases.
-These works provide solutions specific to concrete types of databases using various ranges of solutions - domain specific languages \cite{SERF}, extensions of existing standards or MDD \cite{Evolution_vs_Reorganization} or formal specification \cite{Tresch:1991vi}.
-These solutions are inspiring, however the domain of the ORM has its specific issues, so a solution from another domain has to be adapted carefully.
-
-\paragraph{Informal definitions} The taxonomy of relational database evolution based on the entity-relationship model  is proposed in \cite{Roddick:TaxonomyOnERM}.
-The evolution is described as a change in entity-relationship model and change in relational database.
-The semantics change patterns in context of conceptual schema is described in \cite{Wedemeijer:SemanticChangePatternsInConceptualSchema}, although its impact on database schema or data is not described.
-The main cases of data evolution are defined in both publications, however the description is informal.
-The extensive set of possible database refactorings is provided in \cite{Ambler:DbRefactoringBook}, where both schema and data evolution is discussed.
-The refactorings are intended to be used by database administrators, thus it assume database-first approach to evolution whereas this proposal is application-first.
-
-\paragraph{Formal framework} A general formal framework for database evolution is defined in \cite{McBrien:formal-framework-transformation}.
-The framework is based on a set of basic graph transformations which are then extended to transformations of the entity-relationship model.
-The framework and the defined transformations can be used for comparison with our framework, where our framework consider stored data.
-The contribution of the formal framework is definition of equivalent structures in schemas.
-
-A categorical framework for migration of object-oriented systems is proposed in \cite{Schulz:CategoricalModel}.
-This framework defines the refactoring of objects, data and methods, which are the main objectives of the framework.
-The impact of the object change on a relational database is not considered in the paper as it is aimed at object-oriented systems only.
-
-\paragraph{Tools for Database Evolution} The MeDEA project \cite{Meda:Main} offers a tool for evolution of both database schema and stored data based on model-driven approach.
-The PRISM is a research project for data management under schema evolution \cite{PRISM} in contrast with our proposal, and with MeDEA, it extends the SQL command set by so-called schema modification operators which implements the schema evolution.
-
-The project DB-MAIN \cite{Hick:DBMAIN} provides a MDD approach to data evolution on all the levels of software we do.
-The project is well documented formally.
-The difference between DB-MAIN and our framework is that each has a different focus.
-DB-MAIN starts with database modification and database migration, and entities are created accordingly; whereas our focus is on entities evolution, which is then propagated to a database with an emphasis on automatization.
-Our goal is to hide the entire database level from our users.
-
-All mentioned project is aimed to  be used by database administrators
-
-\paragraph{ORM Frameworks} here are many object-relational mapping frameworks available for developers, and some of them provide tools for database migration.
-Hibernate \cite{Hibernate} is one of the most popular ORM frameworks in the Java community.
-It provides customizable ORM for a wide range of databases, however it does not support complex database migrations.
-It is capable only of creating a new table or adding a new column, hence it is not possible to, for example, drop a table or copy values from one column to another.
-Active Record \cite{Active_Record} is an ORM framework in the Ruby on Rails environment.
-Since its first version it has contained support for database evolution according to the create-update-delete principle, in the form of so-called migrations \cite{Rails:Migrations} which can be extended by adding user SQL commands.
-Entity Framework \cite{Entity_Framework} is Microsoft's ORM solution for the .NET platform.
-Its capabilities of data evolution support are similar to those of Active Record.
-
-Very promising solution is Liquibase \cite{Liquibase} a tool for database refactoring and evolution.
-It is capable to migrate both database schema and stored data.
-The evolution is described in a form of a XML document which can be interpreted on various databases.
-The difference is we try not to bother the developer with database-specific details and make all changes according to a change in application code.
-
-\paragraph{Forward Round-Trip Engineering in Data Evolution} A round-trip approach for data evolution was described in \cite{VanPaesschen:2005to}, which was implemented in the SELF language.
-This approach proposed a forward-oriented evolution on all application levels in the software application.
-In contrast with our framework, the round-trip approach does not care about stored data because it is focused only on create, update and delete types of evolution.
-
-A meta-model based approach to data evolution is proposed in\cite{Aboulsamh:MetaModelBasedApproachToIsDataEvolution} and very similar solution for data evolution is proposed, which is based on extended UML meta-model.
-The solution provides similar capabilities of change of application and database as does our framework.
-In contrast our work is created with respect to ORM domain and therefore we extended the application meta-model with constructs typical for this domain.
 
@@
 \paragraph{Application} Application is defined as a sequence of classes and it creates the context for all structures used in the software persistent layer.
-\begin{align*}
+\begin{align}
 & \mathbf{Application} = (Class*) \\
 & 	\mathbf{Class} = (label, Property*, Association*) \\
-& \mathbf{Property} = (label, AppType, DefaultValue, Cardinality, Mandatory) \\
+& \mathbf{Property} = (label, AppType, DefaultValue, \nonumber \\ & \;\;\;\;  Cardinality, Mandatory) \\
 &	\mathbf{Association} = (label, classRef, startCardinality, endCardinality)  \\
 & \mathbf{AppType} = APPSTRING
-\end{align*}
+\end{align}
 
@@
 \begin{table}
+\caption{The transformations for application evolution.}
+	\label{tab:app-evolution}
 \centering
-	\begin{tabular}{|l|}
+	\begin{tabular}{ll}
 	\hline
-	\textbf{Application Creation} \\
-	newProperty(className, property, application) \\
-	newAssociation(className, association, mapping, application) \\
-	newClass(class, application) \\
-	\textbf{Application Alternation} \\
-	renameProperty(className, oldName, newName, application) \\
-	renameAssociation(className, oldName, newName, application) \\
-	renameClass(oldName, newName, application)\\
-	changeReference(referncedClass, newReferencedClass) \\
-	\textbf{Application Disassembly} \\
-	removeProperty(className, propertyName, application) \\
-	removeAssociation(className, associationName, application) \\
-	removeClass(className, application)\\
-	\textbf{Copy Structure}\\
-	copyProperty(sourceClass, property, targetClass, newName, application)  \\
-	copyClass(class, newName, application) \\
+	Type of transformation & Defined transformations \\
+	\hline
+	\\[-2ex] 
+	\textbf{Application Creation} 
+	& newProperty(className, property, application) \\
+	& newAssociation(className, association, mapping, \\ &  \hspace{0.5in}application) \\
+	&  newClass(class, application) \\
+	\textbf{Application Alternation} 
+		& renameProperty(className, oldName, newName, \\ &  \hspace{0.5in} application) \\
+		& renameAssociation(className, oldName, newName, \\ &  \hspace{0.5in}application) \\
+		& renameClass(oldName, newName, application)\\
+		& changeReference(referncedClass, newReferencedClass) \\
+	\textbf{Application Disassembly} 
+		& removeProperty(className, propertyName,\\ &  \hspace{0.5in} application) \\
+		& removeAssociation(className, associationName, \\ &  \hspace{0.5in}application) \\
+		& removeClass(className, application)\\
+	\textbf{Copy Structure} 	 
+		& copyProperty(sourceClass, property, targetClass,\\ &  \hspace{0.5in} newName, application)  \\
+		& copyClass(class, newName, application) \\
 	\hline
 	\end{tabular}
-	\caption{The transformations for application evolution.}
-	\label{tab:app-evolution}
+
 \end{table}
@@
 
-\begin{align*}
+\begin{align}
 &	\mathbf{Database} = ( Table*, Row* )\\
@@
 &	\mathbf{Constraint} = NOTNULL \; | \; UNIQUE 
-\end{align*}
+\end{align}
 
@@
 Each row contains a name of a table it belongs to.
-\begin{align*}
+\begin{align}
 &	\mathbf{Row} = (Label, N, Pair*) \\
 &	\mathbf{Pair} = (Label, Value) 
-\end{align*}
+\end{align}
 
@@
 \begin{table}
+\caption{The transformations for database evolution.}
+	\label{tab:db-basic-evolution}
 \centering
-	\begin{tabular}{|l|}
+\setlength\tabcolsep{0.5em}
+	\begin{tabular}{ll}
 	\hline
-	\textbf{Database Creation} \\
-	addColumn(tableName, column, database) \\
-	addForeignKey(tableName, key, database) \\
-	addTable(table, database)\\
-	\textbf{Database Alternation} \\
-	alterColumnName(tableName, oldName, newName, dataabse) \\
-	alterForeignKeyName(tableName, oldName, newName, dataabse) \\
-	alterTableName(oldName, newName, dataabse) \\
-	\textbf{Database Disasembling} \\
-	dropColumn(tableName, columnName, database) \\
-	dropForeignKey(tableName, keyName, database) \\
-	dropTable(tableName, database) \\
-	\textbf{Data Manipulation} \\
-	select(tableName, rowId, database) \\
-	select(tableName, database) \\
-	insert(row, database) \\
-	insert(rowId, value, database) \\
-	delete(tableName, rowId, database) \\
-	delete(tableName, rowId, columnName, database) \\
-	\textbf{Copy Structure} \\
-	copyColumnStructure(sourceTableName, targetTableName, \\  \hspace{0.5in} sourceColumnName, newName, database) \\
-	copyTableStructure(sourceTable, newTableName, database) \\
-	\textbf{Copy Strucutre and Values} \\
-	copyColumn(sourceTable, targetTable, \\
-	\hspace{0.5in} sourceColumnName, targetColumnName, \\ 
-	\hspace{0.5in} mapping, database) \\
-	copyTable(sourceTable, newName, database) \\
+	Type of transformation & Defined transformations \\
+	\hline
+	\\[-2ex] Database Creation
+	& addColumn(tableName, column, database) \\
+	& addForeignKey(tableName, key, database) \\
+	& addTable(table, database)\\
+	Database Alternation
+	& alterColumnName(tableName, oldName, \\ & \hspace{0.5in}newName, dataabse) \\
+	& alterForeignKeyName(tableName, oldName, \\ & \hspace{0.5in}newName, dataabse) \\
+	& alterTableName(oldName, newName, dataabse) \\
+	Database Disasembling
+	& dropColumn(tableName, columnName, database) \\
+	& dropForeignKey(tableName, keyName, database) \\
+	& dropTable(tableName, database) \\
+	Data Manipulation
+	& select(tableName, rowId, database) \\
+	& select(tableName, database) \\
+	& insert(row, database) \\
+	& insert(rowId, value, database) \\
+	& delete(tableName, rowId, database) \\
+	& delete(tableName, rowId, columnName, database) \\
+	Copy Structure
+	& copyColumnStructure(sourceTableName,\\ &  \hspace{0.5in} targetTableName,  sourceColumnName, \\ &  \hspace{0.5in} newName, database) \\
+	& copyTableStructure(sourceTable, newTableName, \\ &  \hspace{0.5in} database) \\
+	Copy Strucutre and Values
+	& copyColumn(sourceTable, targetTable, \\
+	& \hspace{0.5in} sourceColumnName, targetColumnName, \\ 
+	& \hspace{0.5in} mapping, database) \\
+	& copyTable(sourceTable, newName, database) \\
 	\hline
 	\end{tabular}
-	\caption{The transformations for database evolution.}
-	\label{tab:db-basic-evolution}
 \end{table}
@@
 The mapping is usually defined as: 
-\begin{align*}
+\begin{align}
 &	mapping =  (r_1, r_2)*,\\
@@
 & 	r_2 \in select(ORM(targetName), database) \cup \varnothing
-\end{align*}
+\end{align}
 The mapping is not limited to the bijection and it is possible to not include a row into a mapping - this has to be verified because of possible $NOTNULL$ constraints on a target column, as well as the $UNIQUE$ constraint has to be verified because the mapping can violate it in general.
@@
 \centering
-	\begin{tabular}{|l|}
+\caption{The data secure transformations for database evolution which can be created as SQL functions.}
+	\label{tab:db-adv-evolution}
+	\begin{tabular}{ll}
 	\hline
-	\textbf{Data Secure Database Disassembling} \\
-	dropEmptyColumn(tableName, columnName, database) \\
-	dropEmptyForeignKey(tableName, keyName, database) \\
-	dropEmptyTable(tableName, database) \\
+	Type of transformation & Defined transformations \\
+	\hline
+	\\[-2ex] 
+	Data Secure Database Disassembling &
+	dropEmptyColumn(tableName, columnName, \\ & \hspace{0.25in} database) \\
+	& dropEmptyForeignKey(tableName, keyName, \\ & \hspace{0.25in} database) \\
+	& dropEmptyTable(tableName, database) \\
 	\hline
 	\end{tabular}
-	\caption{The data secure transformations for database evolution which can be created as SQL functions.}
-	\label{tab:db-adv-evolution}
 \end{table}
@@
 These transformation cannot be  mapped directly on the evolutionary transformations of the database, because properties and associations can be mapped as a column (foreign key respectively) or as a table, this has to be reflected during the evolution as shown on the example of creating a new property:
-\begin{align*}
-& \Phi(newProperty(className, property, application)) = \\
+\begin{align}
+& \Phi(newProperty(className, property, application)) = \nonumber \\
 & \;\;\; =  \begin{cases}
@@
  \end{cases}
-\end{align*}
+\end{align}
 
@@
 \begin{table}
+	\caption{The mapping between evolution of the application to the database transformations.}
+	\label{tab:sw-basic-evolution}
 \centering
-	\begin{tabular}{|p{0.4\textwidth} p{0.5\textwidth}|}
+	\begin{tabular}{ll}
 		\hline
 	Application Transformation & Database Transformation \\
-	\multicolumn{2}{|c|}{\textbf{Software Creation}} \\
-	newProperty & addColumn  ($cardinality \leq 1$)\\
-	& addTable \\ 
-	& \\
-	newAssociation & addColumn ($cardinality \leq 1$)\\
-	& addTable \\
-	& \\
+	\hline
+	\\[-2ex] 
+	newProperty & \textbf{if} $cardinality \leq 1$ \textbf{then} addColumn \textbf{else} addTable \\
+
+	newAssociation & \textbf{if} $cardinality \leq 1$ \textbf{then} addColumn  \textbf{else} addTable \\
+
 	newClass & addTable \\
-	\multicolumn{2}{|c|}{\textbf{Application Alternation}} \\
-	renameProperty & renameColumn ($cardinality \leq 1$)\\
-	& renameTable \\
-	& \\
-	renameAssociation & renameForeignKey ($cardinality \leq 1$)\\
-	& renameTable \\
-	& \\
+	renameProperty & \textbf{if} $cardinality \leq 1$ \textbf{then}  renameColumn  \textbf{else} renameTable \\
+
+	renameAssociation & \textbf{if} $cardinality \leq 1$ \textbf{then} renameForeignKey \\ & \hspace{1in} \textbf{else} renameTable \\
+
 	renameClass & renameTable \\
-	\multicolumn{2}{|c|}{\textbf{Application Disassembly}} \\
-	removeProperty & removeColumn ($cardinality \leq 1$)\\
-	& removeTable \\
-	& \\
-	removeAssociation & removeForeignKey ($cardinality \leq 1$)\\
-	& removeTable \\
-	& \\
+	removeProperty & \textbf{if} $cardinality \leq 1$ \textbf{then} removeColumn  \\ & \hspace{1in}\textbf{else} removeTable \\
+
+	removeAssociation & \textbf{if} $cardinality \leq 1$ \textbf{then}  removeForeignKey \\ & \hspace{1in} \textbf{else} removeTable \\
+
 	removeClass & removeTable\\
@@
 	\end{tabular}
-	\caption{The mapping between evolution of the application to the database transformations.}
-	\label{tab:sw-basic-evolution}
 \end{table}
@@
 %Moreover the source and target should have the same structure, type and constraints.
-\begin{align*}
-& \mathbf{copyProperty}(sourceClass, propertyName, targetClass, \\
+\begin{align}
+& \mathbf{copyProperty}(sourceClass, propertyName, targetClass, \nonumber  \\
 & \; \; \; mapping, software) 
-\end{align*}
+\end{align}
 The transformation is interpreted for both software components - in case of the application a new property is added:
-\begin{align*}
-& \Psi(copyProperty(sourceClass, property, targetClass, mapping, \\ 
+\begin{align}
+& \Psi(copyProperty(sourceClass, property, targetClass, mapping, \nonumber  \\ 
 &  \; \; \; software)) = newProperty(targetClass, property, application) 
-\end{align*}
+\end{align}
 in case of database the copy of a column or table is created according to property's cardinality and then values are copied:
-\begin{align*}
-& \Phi(copyProperty(sourceClass, property, targetClass, newName,  \\ 
+\begin{align}
+& \Phi(copyProperty(sourceClass, property, targetClass, newName, \nonumber \\ 
 &  \; \; \; mapping, software)) = \begin{cases}
@@
 On the other hand the situation when the mapping does not provides pairs for all rows in a source table can lead to data loss.
-\begin{align*}
-& moveProperty(sourceClass, property, targetClass, newName, \\
-& \;\;\; mapping, software) = removeProperty(sourceClass, property, \\
-& \;\;\; CopyProperty(sourceClass, property, targetClass, \\ 
+\begin{align}
+& moveProperty(sourceClass, property, targetClass, newName,\nonumber  \\
+& \;\;\; mapping, software) = removeProperty(sourceClass, property, \nonumber \\
+& \;\;\; CopyProperty(sourceClass, property, targetClass, \nonumber \\ 
 & \;\;\;newName, mapping, software))
-\end{align*}
+\end{align}
 
@@
 The transformation creates new rows in the new table - the primary keys from origin instances are replaced by a new ones.
-\begin{align*}
-& merge(firstClass, secondClass, newClassName, software) = \\
-& \; \; \; = removeClass(secondClass, removeClass(firstClass, \\
-& \; \; \; changeReference(secondClass, newClassName, \\
-& \; \; \; changeReference(firstClass, newClassName, \\
+\begin{align}
+& merge(firstClass, secondClass, newClassName, software) = \nonumber \\
+& \; \; \; = removeClass(secondClass, removeClass(firstClass,\nonumber  \\
+& \; \; \; changeReference(secondClass, newClassName, \nonumber \\
+& \; \; \; changeReference(firstClass, newClassName,\nonumber  \\
 & \; \; \; copyClass(firstClass, newClassName, software))))))
-\end{align*}
+\end{align}
 
@@
 The extract and inline class are two opposite transformations. First of them extract a class from an existing class, second of them moves data from one class into another.
-\begin{align*}
-& extract(sourceClass, newClass, property*, software) = \\ 
-& \; \; \; = moveProperty(sourceClass, newClass, property*, \\ 
+\begin{align}
+& extract(sourceClass, newClass, property*, software) = \nonumber  \\ 
+& \; \; \; = moveProperty(sourceClass, newClass, property*, \nonumber \\ 
 & \; \; \; addClass(newClass, software))
-\\\\
-& inline(targetClass, classToInline, software)) = \\
-& \; \; \; = removeClass(classToInline, moveProperty(classToInline, \\
+\\ \nonumber  \\
+& inline(targetClass, classToInline, software)) =  \nonumber \\
+& \; \; \; = removeClass(classToInline, moveProperty(classToInline, \nonumber \\
 & \; \; \; targetClass, properties(classToInline), software))
-\end{align*}
+\end{align}
 
@@
 	\item \textbf{Replace Data Value with Object} is a refactoring case which is not mentioned among the evolutionary transformations, on the other hand it can be composed from already defined transformations:
-	\begin{align*}
-& replaceDataWithObject(sourceClass, property, newObject, \\
-& \; \; \; software) = removeProperty(sourceClass, property, \\
-& \; \; \; newAssociation(sourceClass, newObject,  mapping, \\
-& \; \; \; copyProperty(sourceClass, property, newObject, \\ & \; \; \;  software)
-	\end{align*}
+	\begin{align}
+& replaceDataWithObject(sourceClass, property, newObject, \nonumber \\
+& \; \; \; software) = removeProperty(sourceClass, property, \nonumber \\
+& \; \; \; newAssociation(sourceClass, newObject,  mapping,\nonumber  \\
+& \; \; \; copyProperty(sourceClass, property, newObject, \nonumber \\ & \; \; \;  software)
+	\end{align}
 Where the mapping use the values copied in the first transformation as a matching point for creating connection between the old and the new class.
@@
 The first  step of is to extract two temporary classes representing address of a $Person$ or of a $LegalParty$. Temporary classes has to be connected with the origin classes by an association.
-\begin{align*}
- s' &= merge("Address\_tmp1", "Address\_tmp2", "Address", \\
-& \;\;\;\;  newAssociation("LegalParty", association("address", \\
-& \;\;\;\; "Address\_tmp2", 1, 1), mapping_2, extract("LegalParty", \\
-& \;\;\;\;  "Address\_tmp2", "street", "city", "zip", \varnothing  \\
-& \;\;\;\; newAssociation("Person", association("address", \\
-& \;\;\;\; "Address\_tmp1", 1, 1), mapping_1, extract("Person", \\
+\begin{align}
+ s' &= merge("Address\_tmp1", "Address\_tmp2", "Address", \nonumber \\
+& \;\;\;\;  newAssociation("LegalParty", association("address", \nonumber \\
+& \;\;\;\; "Address\_tmp2", 1, 1), mapping_2, extract("LegalParty", \nonumber \\
+& \;\;\;\;  "Address\_tmp2", "street", "city", "zip", \varnothing \nonumber  \\
+& \;\;\;\; newAssociation("Person", association("address",\nonumber  \\
+& \;\;\;\; "Address\_tmp1", 1, 1), mapping_1, extract("Person", \nonumber \\
 & \;\;\;\; "Address\_tmp1", "street", "city", "zip", \varnothing, software)))
-\end{align*}
+\end{align}
 Both $mapping_1$ and $mapping_2$ are identities so the creation of the association is straightforward. The result of the transformations is in the Fig. \ref{fig:case2}. 
@@
 
+\section{Related Work}
+\label{sec:related-work}
+Research of data and database evolution is not limited to relational ones only, there are also works in the areas of object databases and XML databases.
+These works provide solutions specific to concrete types of databases using various ranges of solutions - domain specific languages \cite{SERF}, extensions of existing standards or MDD \cite{Evolution_vs_Reorganization} or formal specification \cite{Tresch:1991vi}.
+These solutions are inspiring, however the domain of the ORM has its specific issues, so a solution from another domain has to be adapted carefully.
+
+\paragraph{Informal definitions} The taxonomy of relational database evolution based on the entity-relationship model  is proposed in \cite{Roddick:TaxonomyOnERM}.
+The evolution is described as a change in entity-relationship model and change in relational database.
+The semantics change patterns in context of conceptual schema is described in \cite{Wedemeijer:SemanticChangePatternsInConceptualSchema}, although its impact on database schema or data is not described.
+The main cases of data evolution are defined in both publications, however the description is informal.
+The extensive set of possible database refactorings is provided in \cite{Ambler:DbRefactoringBook}, where both schema and data evolution is discussed.
+The refactorings are intended to be used by database administrators, thus it assume database-first approach to evolution whereas this proposal is application-first.
+
+\paragraph{Formal framework} A general formal framework for database evolution is defined in \cite{McBrien:formal-framework-transformation}.
+The framework is based on a set of basic graph transformations which are then extended to transformations of the entity-relationship model.
+The framework and the defined transformations can be used for comparison with our framework, where our framework consider stored data.
+The contribution of the formal framework is definition of equivalent structures in schemas.
+
+A categorical framework for migration of object-oriented systems is proposed in \cite{Schulz:CategoricalModel}.
+This framework defines the refactoring of objects, data and methods, which are the main objectives of the framework.
+The impact of the object change on a relational database is not considered in the paper as it is aimed at object-oriented systems only.
+
+\paragraph{ORM Frameworks} There are many object-relational mapping frameworks available for developers, and some of them provide tools for database migration.
+Hibernate \cite{Hibernate} is one of the most popular ORM frameworks in the Java community which is capable of creating a new table or adding a new column according to a change in the application.
+Active Record \cite{Active_Record} is an ORM framework in the Ruby on Rails environment.
+Since its first version it has contained support for database evolution according to the create-update-delete principle, in the form of so-called migrations \cite{Rails:Migrations} which can be extended by adding user SQL commands.
+Entity Framework \cite{Entity_Framework} is Microsoft's ORM solution for the .NET platform.
+Its capabilities of data evolution support are similar to those of Active Record.
+Neither of the frameworks is capable of automatization of complex refactoring cases.
+
+\paragraph{Forward Round-Trip Engineering in Data Evolution} A round-trip approach for data evolution was described in \cite{VanPaesschen:2005to}, which was implemented in the SELF language.
+This approach proposed a forward-oriented evolution on all application levels in the software application.
+In contrast with our framework, the round-trip approach does not care about stored data because it is focused only on create, update and delete types of evolution.
+
+A meta-model based approach to data evolution is proposed in\cite{Aboulsamh:MetaModelBasedApproachToIsDataEvolution} and very similar solution for data evolution is proposed, which is based on extended UML meta-model.
+The solution provides similar capabilities of change of application and database as does our framework.
+In contrast our work is created with respect to ORM domain and therefore we extended the application meta-model with constructs typical for this domain.
+
+\paragraph{Tools for Database Evolution} The MeDEA project \cite{Meda:Main} offers a tool for evolution of both database schema and stored data based on model-driven approach.
+The project DB-MAIN \cite{Hick:DBMAIN} provides a MDD approach to data evolution on all the levels of software we do.
+The project DB-MAIN is well documented formally.
+The PRISM is a research project for data management under schema evolution \cite{PRISM} in contrast with our proposal, and with MeDEA, it extends the SQL command set by so-called schema modification operators which implements the schema evolution.
+Very promising solution is Liquibase \cite{Liquibase} a tool for database refactoring and evolution.
+It is capable to migrate both database schema and stored data.
+The evolution is described in a form of a XML document which can be interpreted on various databases.
+
+The difference between these frameworks and our proposal is that each has a different focus.
+All mentioned projects are aimed to  be used by database administrators; whereas our focus is on entities evolution, which is then propagated to a database with an emphasis on automatization.
+Our goal is to hide the entire database level from our users.
+
 \section{Conclusion}
 
-\bibliographystyle{plain}      % basic style, author-year citations
+\bibliographystyle{splncs03}      % basic style, author-year citations
 \bibliography{references}   % name your BibTeX data base

### 2013-02-13 doc/paper.tex
updated formatting

@@
 	\\[-2ex] 
-	\textbf{Application Creation} 
+	Application Creation
 	& newProperty(className, property, application) \\
@@
 	&  newClass(class, application) \\
-	\textbf{Application Alternation} 
+	Application Alternation
 		& renameProperty(className, oldName, newName, \\ &  \hspace{0.5in} application) \\
@@
 		& changeReference(referncedClass, newReferencedClass) \\
-	\textbf{Application Disassembly} 
+	Application Disassembly 
 		& removeProperty(className, propertyName,\\ &  \hspace{0.5in} application) \\
@@
 		& removeClass(className, application)\\
-	\textbf{Copy Structure} 	 
+	Copy Structure
 		& copyProperty(sourceClass, property, targetClass,\\ &  \hspace{0.5in} newName, application)  \\
@@
 The primary key is always provided (automatically generated) by the database as a non-zero natural number.
-Primary key is always defined with constraints $NOTNULL$ and $UNIQUE$.
+Primary key is always defined with constraints\\ $NOTNULL$ and $UNIQUE$.
 
@@
 There are usually defined types such as Varchar, Integer, Boolean etc.
-There is only one type of column values defined in the model called $DBSTRING$.
+There is only one type of column values defined in the model called \\ $DBSTRING$.
 Next the $DBINT$ type is defined for the values of keys.
@@
 \begin{align}
-& \mathbf{copyProperty}(sourceClass, propertyName, targetClass, \nonumber  \\
+& copyProperty(sourceClass, propertyName, targetClass, \nonumber  \\
 & \; \; \; mapping, software) 
@@
 The framework is based on a set of basic graph transformations which are then extended to transformations of the entity-relationship model.
-The framework and the defined transformations can be used for comparison with our framework, where our framework consider stored data.
-The contribution of the formal framework is definition of equivalent structures in schemas.
+The framework and the defined transformations can be implemented in our proposal too.
+The contribution of the formal framework is definition of equivalent structures in relational database schemas.
+Our proposal is aimed to be used in domain of object-oriented languages where the class model is more common than entity-relationship model which is respected in our proposal.
 
@@
 
-A meta-model based approach to data evolution is proposed in\cite{Aboulsamh:MetaModelBasedApproachToIsDataEvolution} and very similar solution for data evolution is proposed, which is based on extended UML meta-model.
-The solution provides similar capabilities of change of application and database as does our framework.
+A meta-model based approach to data evolution is proposed in \cite{Aboulsamh:MetaModelBasedApproachToIsDataEvolution} and very similar solution for data evolution is proposed, which is based on extended UML meta-model.
+The solution provides similar capabilities of change of application and database as does our proposal.
 In contrast our work is created with respect to ORM domain and therefore we extended the application meta-model with constructs typical for this domain.

### 2013-02-13 doc/paper.tex
intro updated

@@
 
-\title{Refactoring of Software Using Relational Database}
+\title{Refactoring of Application Using Relational Database}
 \author{Ondrej Macek\inst{1}}
@@
 
-\vspace{6pt}\textbf{Keywords:}
+\vspace{6pt}\textbf{Keywords:} code refactoring, relational database, formal model
 \end{abstract}
@@
 The evolution severity usually depends on number of changes which has to be made and on the number of affected software components.
-The focus of this paper is in data evolution context of software implemented in object oriented language using relational database.
+Refactoring \cite{Fowler:Refactoring} is very popular practice in object-oriented environment for evolving the source code and software architecture.
+The evolution of database schema and stored data is implemented separately from  source code refactoring, although the change of application affects the database too.
+There exist object-relational mapping frameworks, which can help with propagation of evolution from application to database.
+However these framework are usually not capable to solve complex refactoring cases nor migrate data properly as shown in Sect.
+\ref{sec:problem}.
+
+%The focus of this paper is in data evolution context of software implemented in object oriented language using relational database.
 It means that the data evolution affect at least two software components: persistent objects (entities) and the database.
@@
 
-This paper shows how the code refactoring can be used as the source for data migration.
+The problem of application and database evolution is discussed from developers point of view and the formal model of application refactoring and its impact is shown in Sect.
+\ref{sec:models}.
+There are introduced basic refactoring cases as well as the complex ones which are created as a sequences of the basic ones.
+The capabilities of proposed formal models are illustrated on the common refactoring issues in Sect.
+\ref{sec:case}.
+
+%This paper shows how the application refactoring can be used as the source for database migration.
 The approach based on changes in code is useful in case when there is no model of database as it is derived directly from code.
@@
 
-The paper is organized as follows: first the evolution of software is described in detail in Sec.
+%The paper is organized as follows: first the evolution of software is described in detail in Sec.
 \ref{sec:problem}, then the related work in presented in Sec.
@@
 \textbf{TODO: write an introduction when and why the software changes. and possible solutions.}
+
 A software implemented using an object-oriented language, which uses relational database as a data storage (as illustrated in the Figure \ref{fig:evolution}) is object of interest of this paper.
@@
 
-\paragraph{Formal framework} A general formal framework for database evolution is defined in \cite{McBrien:formal-framework-transformation}.
+\paragraph{Formal frameworks} A general formal framework for database evolution is defined in \cite{McBrien:formal-framework-transformation}.
 The framework is based on a set of basic graph transformations which are then extended to transformations of the entity-relationship model.
@@
 Our proposal is aimed to be used in domain of object-oriented languages where the class model is more common than entity-relationship model which is respected in our proposal.
+
+The formal definition of MDD approach to database schema evolution is proposed in \cite{Aboulsamh:FormalModelingEvolution}, where a changes of database conceptual schema is interpreted on both physical schema and data.
+In contrast our proposal is aimed to solve the problem of code refactoring and we propose more complex transformations such as \emph{merge}.
 
@@
 
+\paragraph{Data Evolution of Software} The project IMIS \cite{Bordbar:ModelBasedMigration} follows the same idea of applying MDD into evolution of a whole software, but does not provide a formal model or an overview of capabilities (defined transformations).
+
 \section{Conclusion}

### 2013-02-13 doc/paper.tex
text

@@
 The evolution of database schema and stored data is implemented separately from  source code refactoring, although the change of application affects the database too.
-There exist object-relational mapping frameworks, which can help with propagation of evolution from application to database.
+There exist object-relational mapping (ORM) frameworks, which can help with propagation of evolution from application to database.
 However these framework are usually not capable to solve complex refactoring cases nor migrate data properly as shown in Sect.
@@
 
-\section{Problem Definition} 
+\section{Refactoring in Context of ORM} 
 \label{sec:problem}
-% víc rozepsat o změnách a jak se jim bránit a kam tím směřujeme
-\textbf{TODO: write an introduction when and why the software changes. and possible solutions.}
+A software implemented using an object-oriented language, which uses relational database as a data storage consists of four main components.
+There is the application itself, the database schema, stored data and the object-relational mapping.
+The refactoring affects not only the application but the database too as the database has to be evolved when the persistent layer of the application changed to fit the object-relational mapping used in the software.
+The common current solution is based on capabilities of object-relational mapping frameworks which are capable to create a database schema according to the given source code or model.
+The process of evolution then proceed in following steps:
+\begin{enumerate}
+	\item The code is refactored (usually using the developers IDE).
+	\item The ORM framework generates a new database schema.
+	\item The data are migrated manually (if needed) from the old to the new database.
+\end{enumerate}
+The last step is error-prone as it is processed manually and the prone to error increases with complexity of refactoring.
+The evolution process require cooperation of a developer and a database administrator or it requires the developer has a knowledge of the database used by the software.
+The knowledge of the ORM is needed in both cases.
+Remarkable is the feasibility of data evolution has to be verified for each deployed software instance, because the data can differ.
+The data and information preservation is crucial issue of database evolution.
+Next observation is the evolution of the software is defined twice for one software - first for the application then for the database.
 
-A software implemented using an object-oriented language, which uses relational database as a data storage (as illustrated in the Figure \ref{fig:evolution}) is object of interest of this paper.
-There are three important components of such a software: application (its persistent layer concretely), database (consisting of database schema and stored data) and an ORM, therefore we define a software as a triple consisting of an application and a database which are connected together by an ORM:
-\begin{align}
-& \mathbf{software} = ( Application, Database, ORM )
-\end{align}
-The software has to be in consistent state so its users could have benefit from its usage.
-The state of the software is defined to communicate the sate of the software to its user, the symbol $\perp$ is used as a notation for inconsistent state.
-\begin{align}
-& \mathbf{state} = Consistent\; s \: | \: \perp, s \in State 
-\end{align}
-The software is in consistent state if application and database are consistent and the database structure corresponds to the application structure according to the ORM.
+\paragraph{Example 1} Let us have only two classes \emph{A} and \emph{B} in the application which are not connected by an association and there are corresponding tables \emph{tab\_a} and \emph{tab\_b} in the database, which contain data.
+We decide to inline \emph{B} into \emph{A}  during the development.
+The class \emph{A} contains all properties of \emph{B} and \emph{B} is removed from the application, the database schema is generated by the ORM framework and it contains only the table \emph{tab\_a}'.
+The data migration has to be created manually - the developer has to define the evolution twice.
+The mapping between data in \emph{tab\_a} and \emph{tab\_b} has to be provided to merge stored data correctly and impact of this mapping the database has to be verified: is there data which can be lost during inlining and is this lost intentional?
 
-The evolution of a software is a transformation from one consistent state to another one.
-These states are called generations of the software and the function which changed the state of a software are called transformations.
-The evolution consists of change in application and in database as shown in the Figure \ref{fig:evolution}.
+\subsection{Solution Proposal}
+We believe there exists better solution which automatize the process of database evolution according to the code refactoring.
+The solution is illustrated in the Fig.
+\ref{fig:evolution}.
+It is based on a change in the evolution process which assumes the ORM does not change during the evolution:
+\begin{enumerate}
+	\item The evolution of the whole software is defined independently on application or database.
+	\item The evolution is interpreted for application and database.
+	\item The evolution is executed.
+\end{enumerate}
+The process can decrease the human mistakes in the evolution process, because there is only one source for the evolution and the evolution is automatically interpreted for the application (as a refactoring) and database (as schema and data migration).
+The existence of the set of all possible software evolutions $E$ is based on a evolutionary transformations specific for application and database.
+Each transformation contains conditions of transformation feasibility, thus the feasibility of the evolution can be verified.
+
+% The evolution consists of change in application and in database as shown in the Figure \ref{fig:evolution}. 
 \begin{figure}
@@
 \end{figure}
-The usual approach is to refactor code (which is quite fast due to help of IDE) and then re-generation or manual migration of the database.
+
+The set $E$ is created with respect to the needs of a developer, therefore the evolutionary cases are similar to the refactoring cases.
+Nevertheless the refactoring itself does not provide enough information for the complex migration of database and data, therefore the inputs of transformation in $E$ are extended beyond normal refactoring inputs.
+
+\paragraph{Example 2} The same situation from Example 1 can be solved more effectively if the \emph{inline} evolution is defined in $E$.
+All information needed for software evolution are provided as the input of the \emph{inline} transformation which is then automatically interpreted as inline of application classes and merge of tables on database level.
+The mapping is provided as the input of \emph{inline} too, thus the script for data migration can be generated.
+The structural feasibility is verified during the interpretation of the \emph{inline} and the migration script contains conditions which verifies the feasibility on level of data.
+
+%The usual approach is to refactor code (which is quite fast due to help of IDE) and then re-generation or manual migration of the database.
 Another approach is to evolve a database (e.g.
@@
 
-The refactoring itself does not provide enough information for the complex migration of database and data, therefore a level defining the evolution is added - the set of all possible transformations $E$.
+%The refactoring itself does not provide enough information for the complex migration of database and data, therefore a level defining the evolution is added - the set of all possible transformations $E$.
 The evolution definition from this level is propagated into both involved levels - application and database.
-The evolution is defined as:
+
+\section{Model of Software Evolution}
+\label{sec:models}
+Three sets of transformations has to be defined to provide the capabilities described in Example 2: the set $E$ of all possible software transformations:
+\begin{equation}
+E = \{t | t: Software \rightarrow Software\}\:,
+\end{equation}
+the set of application refactorings:
+\begin{equation}
+	A = \{t | t: Application \rightarrow Application\}
+\end{equation}
+and the set of database evolutionary transformations
+\begin{equation}
+ D = \{t | t: Database \rightarrow Database\}\:.
+\end{equation}
+
+The evolution of the software is defined as the interpretation of the transformation for each component of the software:
 \begin{align}
-& t(state(s)) = \begin{cases}
-state(sw(\Psi(s, t), \Phi(s,t)) \; \mathbf{if} \; state(s) = Consistent \; s\\\\
+& t(s) = \begin{cases}
+software(\Psi(s, t), \Phi(s,t)) \; \mathbf{if} \; s \neq \perp \\\\
+ 	\perp
+\end{cases}
+s \in Software, t \in E \nonumber
+\end{align}
+Where $\Psi : E \rightarrow A $ interprets the software evolution cases to the code refactoring and $\Phi : E \rightarrow D $ interprets to the evolutionary transformation of database.
+The $\perp$ symbol denotes an inconsistent state of a software or its component.
+The definition of evolution allows to concatenate evolutionary transformations easily.
+
+\subsection{Software Model}
+The software consists of three important components (as seen in Fig.
+\ref{fig:evolution}: application (its persistent layer concretely), database (consisting of database schema and stored data) and an ORM, therefore we define a software as a triple consisting of an application and a database which are connected together by an ORM:
+\begin{align}
+& software = ( Application, Database, ORM )
+\end{align}
+The software has to be in consistent state so its users could have benefit from its usage.
+The software is in consistent state if application and database are consistent and the database structure corresponds to the application structure according to the ORM, therefore the constructor of the software is defined as:
+\begin{align}
+& software(a, d, \rho) = \begin{cases}
+ software(a, d, \rho) \; \textbf{if} \; a \neq \perp \; \wedge \; b \neq \perp \\ \wedge \; \rho(a) = schema(d) 
+ \\\\
  \perp
  \end{cases}\\
-& s \in Software, t \in E \nonumber
+& a \in Application, d \in Database, \rho \in ORM \nonumber
 \end{align}
-Where $\Psi$ maps the software evolution cases to the code refactoring and $\Phi$ to maps to the evolutionary transformation of database.
+The evolution of a software is a transformation from one consistent state to another one.
+These states are called generations of the software and the function which changed the state of a software are called transformations.
 
-The main issue during the software evolution is the preservation of data and information consistency.
+%The main issue during the software evolution is the preservation of data and information consistency.
 The data lost is a situation when data are damaged or lost during the migration.
@@
 
-\section{Model of Software}
-\label{sec:models}
-The aim of this paper is to provide an definition of data evolution, which can be applied in case of various programming language, therefore the platform independent models of application and database are introduced.
+%The aim of this paper is to provide an definition of data evolution, which can be applied in case of various programming language, therefore the platform independent models of application and database are introduced.
 These models also provide simplification, which should increase the understanding of the proposed evolutions.
@@
 \subsection{Application Model}
-%TODO popis
-\paragraph{Application} Application is defined as a sequence of classes and it creates the context for all structures used in the software persistent layer.
+Application is defined as a sequence of classes and it creates the context for all structures used in the software persistent layer:
 \begin{align}
@@
 
-\subsubsection{Application Manipulation}
+\subsection{Application Manipulation}
 \label{sec:app-evolution}
 The application model defines only the structure of an application thus there are defined transformation for its creating, altering and disassembling.
-Therefore the conditions for successful transformations are simple 1) when creating an object in the model or altering its name collisions has to be prevent, 2) when renaming a class the references have to be updated 3) when creating or altering an association the existence of referenced class has to be verified.
-The only issue during disassembling is the class can be removed only in case it is not associated by other classes.
-If a transformation cannot be processed the state of the application become inconsistent.
-The transformations are in table \ref{tab:app-evolution}.
+Therefore the conditions for successful transformations are simple in case of creating or altering:
+\begin{enumerate}
+	\item The name collisions has to be prevent when creating an class in the model or altering its name.
+	\item All references have to be updated when renaming a class.
+	\item The existence of referenced class has to be verified when creating or altering an association.  
+\end{enumerate}
+The only issue during deleting classes etc.
+from an application is the class can be removed only in case it is not associated by other classes.
+The transformations from $A$ are in table \ref{tab:app-evolution}.
 The list of altering operations is not complete as there are missing transformation for changing the obligation or cardinality of properties and associations.
-It is because these transformations are not so important for the next text when the complex refactorings are considered.
+It is because these transformations are not so important for complex refactorings.
 \begin{table}
@@
 \subsection{Database Model}
-he relational database consists of database schema which defines the structure of the database and data, which in our ORM system represents stored instances. The database is defined as a tuple:
-
+The relational database consists of database schema which defines the structure of the database and data, which in the software represents stored instances. The database is defined as:
 \begin{align}
@@
 Primary key is always defined with constraints\\ $NOTNULL$ and $UNIQUE$.
-
 The sequence generates values of primary key.
@@
 
-A database consists not only from schema but also from data which are represented as rows in a table.
+\paragraph{Data} A database consists not only from schema but also from data which are represented as rows in a table.
 A table row in our model is a tuple consisting of value pairs, which represents concrete value of concrete column or key.
@@
 \begin{align}
-&	\mathbf{Row} = (Label, N, Pair*) \\
+&	\mathbf{Row} = (Label, Pair*) \\
 &	\mathbf{Pair} = (Label, Value) 
@@
 
-\subsubsection{Database Manipulation}
+\subsection{Database Manipulation}
 \label{sec:db-evolution}
 The database consists of two parts - from database schema which defines the structure and from stored data - hence the transformations have to consider both parts.
-The transformation for manipualtion of structure has similar conditions to the evolution of application application.
+The transformation for manipulation of structure has similar conditions to the evolution of application application.
 The transformations for data manipulation are inspired by the SQL language.
 The basic transformations are in Table \ref{tab:db-basic-evolution}.
+The set of transformations for database manipulation is limited in contrast with the SQL language as we mentioned only transformation necessary for data evolution on database level.
+
 \begin{table}
@@
 	& copyTable(sourceTable, newName, database) \\
-	\hline
-	\end{tabular}
-\end{table}
-The set of transformations for database manipulation is limited in contrast with the SQL language it is because mentioned transformation are necessary for application of data evolution on database level.
-
-\paragraph{Copying of Database Elements} The transformations for copying the structure of a database elements serves more as helpers in advanced evolution cases as they adds a structural duplicate of an existing element only, nevertheless they are valid part of the transformation for database manipulation.
-The copying of structural elements has the same limits as database creation - the transformation can not cause name collisions.
-
-The mapping between instances of the classes (stored data) to provides information which value from the source table rows will be copied to which row in target table.
-The value of a column is then inserted for each row into the copied column during $copyColumn$ transformation.
-The mapping is usually defined as: 
-\begin{align}
-&	mapping =  (r_1, r_2)*,\\
-& 	r_1 \in select(ORM(sourceName), database) \cup \varnothing, \\
-& 	r_2 \in select(ORM(targetName), database) \cup \varnothing
-\end{align}
-The mapping is not limited to the bijection and it is possible to not include a row into a mapping - this has to be verified because of possible $NOTNULL$ constraints on a target column, as well as the $UNIQUE$ constraint has to be verified because the mapping can violate it in general.
-On the other hand the benevolent mapping help with creating various kinds on advanced evolutionary transformations.
-The mapping can cause data losses in some cases if used as a part of advanced refactoring.
-This situation is discussed later on.
-
-\paragraph{Data Secure Disasembling of a Software} The transformation for disassembly the database can have fatal impact on the data preservation, therefore the set of transformations is extended by a data secure transformations for database disassembling.
-These transformations are not part of the SQL standard, although they can be implemented as database functions.
-These transformations creates a secure way to disassembly the database as they drops empty structural elements only.
-These transformations are in table \ref{tab:db-adv-evolution}.
-\begin{table}
-\centering
-\caption{The data secure transformations for database evolution which can be created as SQL functions.}
-	\label{tab:db-adv-evolution}
-	\begin{tabular}{ll}
-	\hline
-	Type of transformation & Defined transformations \\
-	\hline
-	\\[-2ex] 
-	Data Secure Database Disassembling &
-	dropEmptyColumn(tableName, columnName, \\ & \hspace{0.25in} database) \\
+	Data-secure Database Elements &
+	dropEmptyColumn(tableName, columnName, \\
+	Removal & \hspace{0.25in} database) \\
 	& dropEmptyForeignKey(tableName, keyName, \\ & \hspace{0.25in} database) \\
@@
 
+\paragraph{Copying Database Elements}The transformations for copying the structure of a column or table serves more as helpers for advanced evolution cases.
+The \emph{copyColumn} transformation contains a mapping between instances (stored data).
+The mapping to provides information which value from the source table rows is copied to which row in target table.
+The mapping is usually defined as:
+\begin{align}
+&	mapping : select(ORM(sourceName), database) \nonumber \\ 
+& \;\;\;\;\; \rightarrow  (select(ORM(targetName), database) 
+\end{align}
+The character of transformation depend on the feature of provided mapping.
+There occurs no data lost if the mapping is surjective.
+The lost of target data can be caused by non-surjective function.
+The result of the mapping has to respect the constraints on column (e.g.
+$NULL$ cannot be inserted into a column with $NOTNULL$ constraint defined)
+
+\paragraph{Data-secure Database Elements Removal} The transformation for remove elements from the database can have fatal impact on the data preservation, therefore the set of transformations is extended by a data secure transformations for removing database elements.
+These transformations are not part of the SQL standard, although they can be implemented as database functions.
+These transformations creates a secure way to remove elements from the database as they drops empty structural elements only.
+
 \subsection{Object Relational Mapping}
-The ORM can differ from application to application, its impact on the evolution is not crucial but it has to be present to provide full information about the method.
-The mapping is similar to the Hibernate
-mapping, thus a lot of developers should be familiar with it. 
-% add mapping
-
-\textbf{TODO: Describe ORM}
+The ORM is the only fixed point in the software model we use.
+The mapping is similar to the Hibernate mapping \cite{Hibernate} , thus a lot of developers should be familiar with it.
+The main ideas are:
+\begin{itemize}
+	\item classes are mapped to tables,
+	\item single properties are mapped to column or if the property is collection then the property is mapped to a table,
+	\item associations are mapped to foreign keys or to tables if the association represents many-to-many  relationship,
+	\item the primary keys are created automatically for each table,
+	\item names used on application are mapped into the database schema (e.g. because possible name collision of application classes and names of database schema elements) 
+\end{itemize}
+The ORM mapping is a function $ORM : Application \rightarrow Database $.
 
@@
 \label{sec:sw-evolution}
-The evolution of the whole software consider its both components application and database containing data.
-The way how evolutionary transformation are created based on the set of component specific transformations is introduced in this chapter.
-
-The evolution is described from the developer point of view, therefore the transformations use the names and elements from application context.
+The evolution of the whole software is described from the developer point of view, therefore the transformations use the names and elements from application context.
 The names of application elements and elements themselves have to be transformed into a database context - this is assured by the ORM.
-
-%Finally the evolution cases and their mappings on application and database components are presented in \ref{sec:sw-evolution}.
-
-%TODO evalutation of trnasformations
 
@@
 
-\paragraph{Basic Software Evolution} The basic evolutionary transformations of the software are based on basic evolution of an application.
-These transformation cannot be  mapped directly on the evolutionary transformations of the database, because properties and associations can be mapped as a column (foreign key respectively) or as a table, this has to be reflected during the evolution as shown on the example of creating a new property:
+The basic evolutionary transformations of the software are based on basic evolution of an application.
+These transformations has to respect the ORM, because properties and associations can be mapped as a column (foreign key respectively) as shown on the example of creating a new property:
 \begin{align}
@@
 Mapping between basic evolutionary transformations is provided in the shortened way in table \ref{tab:sw-basic-evolution}, there is the basic application transformation and possible mappings on database level according to to the cardinality.
-
 \begin{table}
@@
 	renameClass & renameTable \\
-	removeProperty & \textbf{if} $cardinality \leq 1$ \textbf{then} removeColumn  \\ & \hspace{1in}\textbf{else} removeTable \\
+	removeProperty & \textbf{if} $cardinality \leq 1$ \textbf{then} dropEmptyColumn  \\ & \hspace{1in}\textbf{else} dropEmptyTable \\
 
-	removeAssociation & \textbf{if} $cardinality \leq 1$ \textbf{then}  removeForeignKey \\ & \hspace{1in} \textbf{else} removeTable \\
+	removeAssociation & \textbf{if} $cardinality \leq 1$ \textbf{then}  dropEmptyForeignKey \\ & \hspace{1in} \textbf{else} dropEmptyTable \\
 
-	removeClass & removeTable\\
+	removeClass & dropEmptyTable\\
 	\hline
@@
 \end{table}
-
-The data secure transformations for disassembling the database can succeed (e.g. drop a table) when there are no data contained within the dropped element.
+The transformations for data-secure removal are used as default removing transformations, although the classic drop- transformations can be used.
+This should lead to the more careful usage of removing trnasformations.
 
@@
 Advanced evolutionary transformations are based on the basic ones, they can be obtained as a concatenation of transformations.
-The concatenation of transformation is possible because each transformation produces a state of a software (or application or database) which can be used as an input for the next transformation.
-
-All transformations presented so far has the same input information for both application and database, whereas the advanced transformations usually needs a mapping between stored data (instances) as theirs input.
+This means the complex transformations are limited in the same way as their basic components.
+All transformations from the set $E$ presented so far has the same input information for both application and database, whereas the advanced transformations usually needs a mapping between stored data (instances) as theirs input.
 
@@
 Therefore the  $copyProperty$ transformation is the simplest way to manipulate stored data.
-
-The  $copyProperty$ is the first transformation when an additional information has to be added to the usual code refactoring.
-It is because the $copyColumn$ transformation needs one more information to succeed - there has to be provided mapping between the source and target table to assure the data information consistency.
-%Moreover the source and target should have the same structure, type and constraints.
 \begin{align}
@@
 \end{align}
+The  $copyProperty$ is the first transformation when an additional information has to be added to the usual code refactoring.
+It is because the $copyColumn$ transformation needs one more information to succeed - there has to be provided mapping between the source and target table to assure the data information consistency.
+%Moreover the source and target should have the same structure, type and constraints.
 The transformation is interpreted for both software components - in case of the application a new property is added:
@@
 
-If there are no data in the target table an identity can be used as a mapping function instead of mapping above.
-
 \paragraph{Move Property}
-The $moveProperty$ transformation is based on the $copy$ $Property$ transformation followed by the $removeProperty$ so its construction is easy.
+The $moveProperty$ transformation is based on the $copyProperty$ transformation followed by the $removeProperty$ so its construction is easy.
 On the other hand special attention has to be paid to the provided mapping of instances, because it can cause data loses.
-The ideal case is when the mapping is a bijection among instances, then the transformation can not cause the lost of data.
-On the other hand the situation when the mapping does not provides pairs for all rows in a source table can lead to data loss.
+The ideal case is when the mapping is surjective, then the transformation can not cause the lost of data.
+%On the other hand the situation when the mapping does not provides pairs for all rows in a source table can lead to data loss.
 \begin{align}
@@
 & \;\;\; mapping, software) = removeProperty(sourceClass, property, \nonumber \\
-& \;\;\; CopyProperty(sourceClass, property, targetClass, \nonumber \\ 
+& \;\;\; copyProperty(sourceClass, property, targetClass, \nonumber \\ 
 & \;\;\;newName, mapping, software))
-\end{align}
-
-\paragraph{Merge Classes}
-There occur situations when there is a need to merge two classes and the inlining transformation cannot be used because of name collisions among properties in such a case the $merge$ transformation can be used.
-It can be used in case structure of both classes (respectively tables) are equal.
-The transformation creates new rows in the new table - the primary keys from origin instances are replaced by a new ones.
-\begin{align}
-& merge(firstClass, secondClass, newClassName, software) = \nonumber \\
-& \; \; \; = removeClass(secondClass, removeClass(firstClass,\nonumber  \\
-& \; \; \; changeReference(secondClass, newClassName, \nonumber \\
-& \; \; \; changeReference(firstClass, newClassName,\nonumber  \\
-& \; \; \; copyClass(firstClass, newClassName, software))))))
 \end{align}
@@
 
-The advanced refactoring cases are limited by the transformations they are created from.
+\paragraph{Merge Classes}
+There occur situations when there is a need to merge two classes and the inlining transformation cannot be used because of name collisions among properties in such a case the $merge$ transformation can be used.
+It can be used in case structure of classes and corresponding tables is the same.
+This is because the redundant columns or foreign keys can be constrained with $NOTNULL$ constraint and during merge we are not able to fill missing values.
 
-\section{Case Study}
-\label{sec:case}
-It is obvious that the set of defined transformation is limited because of focus on data preservation or transformation concatenation.
-On the other hand defined transformations are strong enough to handle a lot of refactoring cases.
+The transformation creates new rows in the new table - the primary keys from origin instances are replaced by a new ones.
+\begin{align}
+& merge(firstClass, secondClass, newClassName, software) = \nonumber \\
+& \; \; \; = removeClass(secondClass, removeClass(firstClass,\nonumber  \\
+& \; \; \; changeReference(secondClass, newClassName, \nonumber \\
+& \; \; \; changeReference(firstClass, newClassName,\nonumber  \\
+& \; \; \; copyClass(firstClass, newClassName, software))))))
+\end{align}
 
-Before one complex case study example is shown a set of small refactoring cases is discussed.
-These cases are chosen based on the Eclipse foundation refactoring statistics (the transformation from the top ten influencing data chosen only) and on Martin Fowler's book: Refactoring - improving design of existing code, where one chapter is dedicated to the data organization in the software.
-
-Selected refactorings are:
-\begin{enumerate}
-	\item \textbf{Rename} is the most used refactoring used according to Eclipse statistic and it is one of the basic refactoring cases introduced in the Sec. \ref{sec:sw-basic-evolution}.
-	\item  \textbf{Move} refactoring is used in Eclipse to move properties from a class to another class within an inheritance hierarchy or to move classes between packages.
-In contrast our model does not consider packages, on the other hand it is able to move property from one class to another according given mapping, which contains not only move in hierarchy but a lot of other cases too.
-	\item \textbf{Extract Class} is often used refactoring in Eclipse and it is mentioned in Fowler's book too.
-It is introduced as an example of advanced refactorings in Sec.
-\ref{sec:sw-adv-evolution} together with its opposite transformation $inline$.
-	\item \textbf{Move field} from the Fowler's book is introduced as $moveProperty$ in Sec. \ref{sec:sw-basic-evolution}.
-	\item \textbf{Replace Data Value with Object} is a refactoring case which is not mentioned among the evolutionary transformations, on the other hand it can be composed from already defined transformations:
-	\begin{align}
-& replaceDataWithObject(sourceClass, property, newObject, \nonumber \\
-& \; \; \; software) = removeProperty(sourceClass, property, \nonumber \\
-& \; \; \; newAssociation(sourceClass, newObject,  mapping,\nonumber  \\
-& \; \; \; copyProperty(sourceClass, property, newObject, \nonumber \\ & \; \; \;  software)
-	\end{align}
-Where the mapping use the values copied in the first transformation as a matching point for creating connection between the old and the new class.
-\end{enumerate}
-The transformations proposed in this paper are able to implement all data refactorings used on Eclipse platform.
-There are some refactorings from Fowler which are not considered in the proposal: replace array with object, change unidirectional association to bidirectional (and reverse) and operations which replace types with polymorphism.
-
-\subsection{Example of Usage}
+\section{Example of Usage}
 The example demonstrates how the transformation helps with data evolution in real example of software evolution.
@@
 \end{figure}
-The first  step of is to extract two temporary classes representing address of a $Person$ or of a $LegalParty$. Temporary classes has to be connected with the origin classes by an association.
+The first  step of is to extract two temporary classes representing address of a $Person$ or of a $LegalParty$:
 \begin{align}
- s' &= merge("Address\_tmp1", "Address\_tmp2", "Address", \nonumber \\
-& \;\;\;\;  newAssociation("LegalParty", association("address", \nonumber \\
-& \;\;\;\; "Address\_tmp2", 1, 1), mapping_2, extract("LegalParty", \nonumber \\
-& \;\;\;\;  "Address\_tmp2", "street", "city", "zip", \varnothing \nonumber  \\
-& \;\;\;\; newAssociation("Person", association("address",\nonumber  \\
-& \;\;\;\; "Address\_tmp1", 1, 1), mapping_1, extract("Person", \nonumber \\
-& \;\;\;\; "Address\_tmp1", "street", "city", "zip", \varnothing, software)))
+s' &= extract("LegalParty",  "Address\_tmp2", "street", "city", "zip", \varnothing, \nonumber  \\ 
+& \;\;\;\; extract("Person", "Address\_tmp1", "street", "city", "zip", \nonumber \\ & \;\;\;\;  \varnothing, software)))
 \end{align}
+Temporary classes has to be connected with the origin classes by an association:
+\begin{align}
+ s'' &= newAssociation("LegalParty", association("address", \nonumber \\
+& \;\;\;\; "Address\_tmp2", 1, 1), mapping_2, \nonumber \\ & \;\;\;\; newAssociation("Person", association("address",\nonumber  \\
+& \;\;\;\; "Address\_tmp1", 1, 1), mapping_1, s')))
+\end{align}
+Finally the temporary class are merged into one:
+\begin{align}
+ s''' &= merge("Address\_tmp1", "Address\_tmp2", "Address", s'')
+\end{align}
+
 Both $mapping_1$ and $mapping_2$ are identities so the creation of the association is straightforward. The result of the transformations is in the Fig. \ref{fig:case2}. 
@@
 \end{subfigure}
-
-\begin{subfigure}[b]{0.5\textwidth}
-	\centering
-	\begin{tabular}{| c | l |}
-	 	\hline
-		Id &  Surname \\ \hline  
-		11 & Jackson  \\ \hline
-		12 & Clooney  \\ \hline
-	\end{tabular}
-	\caption{Data stored in the table $Person$.}
-\end{subfigure}
-\begin{subfigure}[b]{0.5\textwidth}
-	\centering
-	\begin{tabular}{| c | l |}
-	 	\hline
-		Id &  BusinessName  \\ \hline  
-		100 & Tools \& Machines  \\ \hline
-		200 & AI Robotics \\ \hline
-	\end{tabular}
-	\caption{Data stored in the table $LegalParty$.}
-\end{subfigure}
 \begin{subfigure}[b]{\textwidth}
@@
 	\end{tabular}
-	\caption{Data stored in the table $Address\_tmp2$.}
+	\caption{Data stored in the table $Address$.}
 \end{subfigure}
-	\caption{The initial state of the example used in the case study. There is a repetition of information structure in the $Person$ and $LegalParty$ class.}
+	\caption{The final state of the example used in the case study. There is only one table containing all addresses in the system, which reference the $Party$ and $LegalParty$.}
 	\label{fig:case2}
@@
 creating temporary classes before merge).
+
+\section{Capabilities of the Proposal}
+\label{sec:case}
+It is obvious that the set of defined transformation is limited because of focus on data preservation or transformation concatenation.
+On the other hand defined transformations are strong enough to handle a lot of refactoring cases.
+To verify this we use refactoring cases based on the Eclipse foundation refactoring statistics \cite{Eclipse:Refactoring} (the transformation from the top ten influencing data chosen only) and on Fowler's book \cite{Fowler:Refactoring}.
+Selected refactorings are:
+\begin{enumerate}
+	\item \textbf{Rename} is the most used refactoring used according to Eclipse statistic and it is one of the basic refactoring cases introduced in the Sec. \ref{sec:sw-basic-evolution}.
+	\item  \textbf{Move} refactoring is used in Eclipse to move properties from a class to another class within an inheritance hierarchy or to move classes between packages.
+In contrast our model does not consider packages, on the other hand it is able to move property from one class to another according given mapping, which contains not only move in hierarchy but a lot of other cases too.
+	\item \textbf{Extract Class} is often used refactoring in Eclipse and it is mentioned in Fowler's book too.
+It is introduced as an example of advanced refactorings in Sec.
+\ref{sec:sw-adv-evolution} together with its opposite transformation $inline$.
+	\item \textbf{Move field} from the Fowler's book is introduced as $moveProperty$ in Sec. \ref{sec:sw-basic-evolution}.
+	\item \textbf{Replace Data Value with Object} is a refactoring case which is not mentioned among the evolutionary transformations, on the other hand it can be composed from already defined transformations:
+	\begin{align}
+& replaceDataWithObject(sourceClass, property, newObject, \nonumber \\
+& \; \; \; software) = removeProperty(sourceClass, property, \nonumber \\
+& \; \; \; newAssociation(sourceClass, newObject,  mapping,\nonumber  \\
+& \; \; \; copyProperty(sourceClass, property, newObject, \nonumber \\ & \; \; \;  software)
+	\end{align}
+Where the mapping use the values copied in the first transformation as a matching point for creating connection between the old and the new class.
+\end{enumerate}
+The transformations proposed in this paper are able to implement all data refactorings used on Eclipse platform.
+There are some refactorings from Fowler which are not considered in the proposal: replace array with object, change unidirectional association to bidirectional (and reverse) and operations which replace types with polymorphism.
+However the proposal solves the most used refactorings.
 

### 2013-02-14 doc/paper.tex
completing

@@
 121 35 Praha 2, Czech Republic\\
-  \email{ondrej.macek@fel.cvut.cz}}
+  \email{macekond@fel.cvut.cz}}
 
@@
 \begin{abstract}
+The change of application source code caused by refactoring affected not only the application, but it affect the stored data too if the persistent objects change.
+The change is usually processed in two steps: refactoring and data migration.
+We provide a solution capable to migrate database according to a refactoring in the application code.
+The feasibility of change and its data-secure processing is addressed too.
 
@@
 \label{sec:models}
+The model-driven development (MDD) is a good approach to data evolution of various software components \cite{Moravec:PracticalApproach}.
+The software is represented by a set of models - application model and database model concretely.
+The evolution and ORM is represented by a set of model-to-model transformations and the interpretation of transformations from $E$ is a model-to-model transformation again.
+
 Three sets of transformations has to be defined to provide the capabilities described in Example 2: the set $E$ of all possible software transformations:
@@
 \end{equation}
-
 The evolution of the software is defined as the interpretation of the transformation for each component of the software:
@@
 The software consists of three important components (as seen in Fig.
-\ref{fig:evolution}: application (its persistent layer concretely), database (consisting of database schema and stored data) and an ORM, therefore we define a software as a triple consisting of an application and a database which are connected together by an ORM:
-\begin{align}
-& software = ( Application, Database, ORM )
-\end{align}
+\ref{fig:evolution}: application (its persistent layer concretely), database (consisting of database schema and stored data) and an ORM, therefore we define a software as a triple consisting of an application and a database which are connected together by an ORM.
 The software has to be in consistent state so its users could have benefit from its usage.
-The software is in consistent state if application and database are consistent and the database structure corresponds to the application structure according to the ORM, therefore the constructor of the software is defined as:
+The software is in consistent state if application and database are consistent and the database structure corresponds to the application structure according to the ORM, therefore the the software is defined as:
 \begin{align}
@@
 \label{sec:app-evolution}
-The application model defines only the structure of an application thus there are defined transformation for its creating, altering and disassembling.
-Therefore the conditions for successful transformations are simple in case of creating or altering:
+The application model defines only the structure of an application thus there are defined transformation for its creating, altering and removing parts of the model (the set $A$).
+The conditions for successful transformations are simple in case of creating or altering:
 \begin{enumerate}
@@
 \label{sec:db-evolution}
-The database consists of two parts - from database schema which defines the structure and from stored data - hence the transformations have to consider both parts.
+The database consists of two parts - from database schema which defines the structure and from stored data - hence the transformations from set $D$ have to consider both parts.
 The transformation for manipulation of structure has similar conditions to the evolution of application application.
@@
 \begin{table}
-\caption{The transformations for database evolution.}
+\caption{The transformations for database evolution. The transformations can be mapped to SQL intuitevly.}
 	\label{tab:db-basic-evolution}
@@
 The mapping is usually defined as:
-\begin{align}
+\begin{align}\label{eq:mapping}
 &	mapping : select(ORM(sourceName), database) \nonumber \\ 
@@
 \end{align}
-The character of transformation depend on the feature of provided mapping.
+The result of the mapping has to respect the constraints on column (e.g.
+$NULL$ cannot be inserted into a column with $NOTNULL$ constraint defined) The character of transformation depend on the feature of provided mapping.
 There occurs no data lost if the mapping is surjective.
-The lost of target data can be caused by non-surjective function.
-The result of the mapping has to respect the constraints on column (e.g.
-$NULL$ cannot be inserted into a column with $NOTNULL$ constraint defined)
 
@@
 \end{itemize}
-The ORM mapping is a function $ORM : Application \rightarrow Database $.
+%The ORM mapping is a function $ORM : Application \rightarrow Database $.
 
@@
 \paragraph{Extract and Inline Class}
-The extract and inline class are two opposite transformations. First of them extract a class from an existing class, second of them moves data from one class into another.
+Extract and inline class are two opposite transformations.
+First of them extract a new class from an existing class, second of them moves data from one class into another.
+Both transformations can be composed from already mentioned basic trasnformations:
 \begin{align}
 & extract(sourceClass, newClass, property*, software) = \nonumber  \\ 
-& \; \; \; = moveProperty(sourceClass, newClass, property*, \nonumber \\ 
+& \; \; \; = moveProperty(sourceClass, newClass, property*, mapping \nonumber \\ 
 & \; \; \; addClass(newClass, software))
-\\ \nonumber  \\
-& inline(targetClass, classToInline, software)) =  \nonumber \\
+\end{align}
+The \emph{mapping} cannot be created according to the definition \ref{eq:mapping}, because there are no stored data in the table representing the \emph{newClass}, therefore the \emph{mapping} is defined as identity in case of \emph{extract}, therefore the rows in the new table has same primaryKey value as in their source class.
+The \emph{extract} transformations assumes the \emph{sourceClass} and \emph{newClass} has the same type of primary keys sequence.
+\begin{align}
+& inline(targetClass, classToInline, mapping, software)) =  \nonumber \\
 & \; \; \; = removeClass(classToInline, moveProperty(classToInline, \nonumber \\
-& \; \; \; targetClass, properties(classToInline), software))
+& \; \; \; targetClass, properties(classToInline), mapping, software))
 \end{align}
@@
 There occur situations when there is a need to merge two classes and the inlining transformation cannot be used because of name collisions among properties in such a case the $merge$ transformation can be used.
-It can be used in case structure of classes and corresponding tables is the same.
-This is because the redundant columns or foreign keys can be constrained with $NOTNULL$ constraint and during merge we are not able to fill missing values.
-
-The transformation creates new rows in the new table - the primary keys from origin instances are replaced by a new ones.
 \begin{align}
-& merge(firstClass, secondClass, newClassName, software) = \nonumber \\
+& \Psi(merge(firstClass, secondClass, newClassName, software)) = \nonumber \\
 & \; \; \; = removeClass(secondClass, removeClass(firstClass,\nonumber  \\
@@
 & \; \; \; changeReference(firstClass, newClassName,\nonumber  \\
-& \; \; \; copyClass(firstClass, newClassName, software))))))
+& \; \; \; addClass(class(newClassName, \nonumber  \\
+& \; \; \; \; \;properties(firstClass) \cup properties(secondClass), \nonumber  \\
+& \; \; \;\; \;associations(firstClass) \cup associations(secondClass), \nonumber  \\
+& \; \; \;application)))))) 
 \end{align}
+The interpretation for the database level is more complicated as the transformation copies rows from different sources with different primary key sequences in the new (empty) table.
+Therefore the primary keys from origin instances has to be replaced by a new ones and all references (foreign keys) has to be updated accordingly.
+Next the \emph{merge} can be used only if the structure of classes and corresponding tables is the same or if the extra columns (foreign keys) are not constrained, because we are not able to automatically solve the  duplicity of values or missing values.
 
 \section{Example of Usage}
-The example demonstrates how the transformation helps with data evolution in real example of software evolution.
-There are two classes in our software $Person$ and $LegalParty$ both of them contains information about address (street, city and zip code).
-The situation in software is in Figure \ref{fig:case1}, where the whole software is shown.
+The examples demonstrate how the transformations help with data evolution in real example of software evolution.
+There are two classes in our software $Person$ and $LegalParty$ both of them contains information about address (street, city and zip code) as shown in the Fig.
+\ref{fig:case1} where the database tables are shown.
 
@@
 \begin{figure}
-\begin{subfigure}[b]{0.45\textwidth}
-	\includegraphics[width=\textwidth]{./images/case_app_1}
-	\caption{The application model.}
-\end{subfigure}
-\quad
-\begin{subfigure}[b]{0.45\textwidth}
-	\includegraphics[width=\textwidth]{./images/case_db_1}
-	\caption{The database model.}
-\end{subfigure}
 \begin{subfigure}[b]{\textwidth}
@@
 \end{align}
-Temporary classes has to be connected with the origin classes by an association:
+Temporary classes has to be connected with the origin classes by an association.
+The associations are creates as optional because foreign key constrained with $NOTNULL$ can cause problems in the next phase of evolution.
 \begin{align}
  s'' &= newAssociation("LegalParty", association("address", \nonumber \\
-& \;\;\;\; "Address\_tmp2", 1, 1), mapping_2, \nonumber \\ & \;\;\;\; newAssociation("Person", association("address",\nonumber  \\
-& \;\;\;\; "Address\_tmp1", 1, 1), mapping_1, s')))
+& \;\;\;\; "Address\_tmp2", 0, 1), mapping_2, \nonumber \\ & \;\;\;\; newAssociation("Person", association("address",\nonumber  \\
+& \;\;\;\; "Address\_tmp1", 0, 1), mapping_1, s')))
 \end{align}
+The $mapping_1$ and $mapping_2$ can be defined using the equality of primary keys values, because the temporary classes were extracted from the $Person$ and $LegalParty$.
 Finally the temporary class are merged into one:
@@
 \end{align}
-
-Both $mapping_1$ and $mapping_2$ are identities so the creation of the association is straightforward. The result of the transformations is in the Fig. \ref{fig:case2}. 
+The merge of classes is possible because the associations (and corresponding foreign keys) created in the previous step are not constrained.
+The result of the transformations is in the Fig.
+\ref{fig:case2}.
 \begin{figure}
@@
 \end{figure}
+The change of the code can continue e.g. by adding more addresses to the $Person$ class, creating an enumeration for cities etc.
 
@@
 It shows that usage of the transformations are sometimes not intuitive for user (e.g.
-creating temporary classes before merge).
+creating temporary classes before merge), therefore next complex transformations has to be define based on user experience.
+The examples shows that the column constraint s often decide the feasibility of a transformation.
 
@@
 On the other hand defined transformations are strong enough to handle a lot of refactoring cases.
-To verify this we use refactoring cases based on the Eclipse foundation refactoring statistics \cite{Eclipse:Refactoring} (the transformation from the top ten influencing data chosen only) and on Fowler's book \cite{Fowler:Refactoring}.
+To verify this we use refactoring cases based on the Eclipse foundation refactoring statistics \cite{Eclipse:Refactoring} (the transformation influencing data chosen only) and on Fowler's book \cite{Fowler:Refactoring}.
 Selected refactorings are:
@@
 The formal definition of MDD approach to database schema evolution is proposed in \cite{Aboulsamh:FormalModelingEvolution}, where a changes of database conceptual schema is interpreted on both physical schema and data.
-In contrast our proposal is aimed to solve the problem of code refactoring and we propose more complex transformations such as \emph{merge}.
+In contrast our proposal is aimed to solve the problem of code refactoring and its impact on relational database, next we propose more complex transformations and examine the impact of platform specific constructs (such as foreign keys or constraints) on the evolution.
 
@@
 The impact of the object change on a relational database is not considered in the paper as it is aimed at object-oriented systems only.
+
+\paragraph{Forward Round-Trip Engineering in Data Evolution} A round-trip approach for data evolution was described in \cite{VanPaesschen:2005to}, which was implemented in the SELF language.
+This approach proposed a forward-oriented evolution on all application levels in the software application.
+In contrast with our framework, the round-trip approach does not care about stored data because it is focused only on create, update and delete types of evolution.
+
+A meta-model based approach to data evolution is proposed in \cite{Aboulsamh:MetaModelBasedApproachToIsDataEvolution} and very similar solution for data evolution is proposed, which is based on extended UML meta-model.
+The solution provides similar capabilities of change of application and database as does our proposal.
+In contrast our work is created with respect to ORM domain and therefore we extended the application meta-model with constructs typical for this domain.
 
@@
 Neither of the frameworks is capable of automatization of complex refactoring cases.
-
-\paragraph{Forward Round-Trip Engineering in Data Evolution} A round-trip approach for data evolution was described in \cite{VanPaesschen:2005to}, which was implemented in the SELF language.
-This approach proposed a forward-oriented evolution on all application levels in the software application.
-In contrast with our framework, the round-trip approach does not care about stored data because it is focused only on create, update and delete types of evolution.
-
-A meta-model based approach to data evolution is proposed in \cite{Aboulsamh:MetaModelBasedApproachToIsDataEvolution} and very similar solution for data evolution is proposed, which is based on extended UML meta-model.
-The solution provides similar capabilities of change of application and database as does our proposal.
-In contrast our work is created with respect to ORM domain and therefore we extended the application meta-model with constructs typical for this domain.
 
@@
 
-\paragraph{Data Evolution of Software} The project IMIS \cite{Bordbar:ModelBasedMigration} follows the same idea of applying MDD into evolution of a whole software, but does not provide a formal model or an overview of capabilities (defined transformations).
+The project IMIS \cite{Bordbar:ModelBasedMigration} follows the same idea of applying MDD into evolution of a whole software, but does not provide a formal model or an overview of capabilities (defined transformations).
 
 \section{Conclusion}
+We discuss the impact of application refactoring on relational database schema and stored data.
+We introduce basic transformations of application and database, next show how the complex refactorings can be constructed based on basic refactorings.
+The transformations presented are capable to solve the main refactoring cases and provide both verification of transformations structural feasibility and data-secure migration for database schema and data.
 
-\bibliographystyle{splncs03}      % basic style, author-year citations
-\bibliography{references}   % name your BibTeX data base
+The proposal is based on the idea of MDD which is implemented by a set of models and transformations rules.
+This allows to simulate the behavior of an evolution in the platform independent environment.
+
+The main contribution is the new point of view on the application refactoring.
+The application code and relational database co-evolution can improve the capabilities of IDEs and speed up the work of developers.
+Next contributions are definitions of impact of advanced refactoring cases on relational database and stored data.
+
+\bibliographystyle{splncs03}     
+\bibliography{references}
 

### 2013-02-15 doc/paper.tex
added inheritance

@@
 & \mathbf{Application} = (Class*) \\
-& 	\mathbf{Class} = (label, Property*, Association*) \\
+& 	\mathbf{Class} = (label, Property*, Association*, Inheritance) \\
 & \mathbf{Property} = (label, AppType, DefaultValue, \nonumber \\ & \;\;\;\;  Cardinality, Mandatory) \\
 &	\mathbf{Association} = (label, classRef, startCardinality, endCardinality)  \\
+& \mathbf{Inheritance} = (class, InheritanceType) \\¨
+& \mathbf{InheritanceType} = SINGLETABLE \; | \; TABLE-PER-CLASS \\
 & \mathbf{AppType} = APPSTRING
@@
 		& copyClass(class, newName, application) \\
+	Inheritance manipulation   & setParent(class, parentClass) \\
+		& removeParent(childClass) \\
+		& pushDown(class, property)\\
+		& pullUp(class, property)\\
 	\hline
 	\end{tabular}
-
 \end{table}
@@
 	removeClass & dropEmptyTable\\
+	setParent & \\
+	removeParent & \\
+	pushDown &\\
+	pullUp &\\
 	\hline
@@
 The transformations for data-secure removal are used as default removing transformations, although the classic drop- transformations can be used.
-This should lead to the more careful usage of removing trnasformations.
+This should lead to the more careful usage of removing transformations.
 
@@
 Next the \emph{merge} can be used only if the structure of classes and corresponding tables is the same or if the extra columns (foreign keys) are not constrained, because we are not able to automatically solve the  duplicity of values or missing values.
+
+\subsection{Inheritance Manipulation}
+\begin{align}
+& extractSuperClass(class, parentName, property*) \\
+  & extractCommonSuperClass(firstClass, secondClass, parentName,  property* ) \\
+\end{align}
 

### 2013-02-15 doc/paper.tex
inheritance-related transformations

@@
 
-%TODO consistency
+ There is not allowed multiple ancestors and of course a class can not be its own ancestor cyclically. We also assume there is only one type of inheritance type in the class hierarchy.
 
@@
 		& removeClass(className, application)\\
-	Copy Structure
-		& copyProperty(sourceClass, property, targetClass,\\ &  \hspace{0.5in} newName, application)  \\
-		& copyClass(class, newName, application) \\
+%	Copy Structure
+%		& copyProperty(sourceClass, property, targetClass,\\ &  \hspace{0.5in} newName, application)  \\
+%		& copyClass(class, newName, application) \\
 	Inheritance manipulation   & setParent(class, parentClass) \\
@@
 
-%TODO consistency
-
 \subsection{Database Manipulation}
@@
 	& copyTableStructure(sourceTable, newTableName, \\ &  \hspace{0.5in} database) \\
-	Copy Strucutre and Values
+	Copy Structure and Values
 	& copyColumn(sourceTable, targetTable, \\
@@
 \begin{table}
-	\caption{The mapping between evolution of the application to the database transformations.}
+	\caption{The mapping between evolution of the application and database transformations. The  transformation connected with inheritance are in Sect.\ref{sec:inheritance}. }
 	\label{tab:sw-basic-evolution}
@@
 	removeAssociation & \textbf{if} $cardinality \leq 1$ \textbf{then}  dropEmptyForeignKey \\ & \hspace{1in} \textbf{else} dropEmptyTable \\
-
 	removeClass & dropEmptyTable\\
-	setParent & \\
-	removeParent & \\
-	pushDown &\\
-	pullUp &\\
 	\hline
@@
 \subsection{Inheritance Manipulation}
+\label{sec:inheritance}
+The inheritance is important part of object-oriented world.
+The impact of change of this relationship to the database depends on the type of inheritance mapping similarly as in the case of \emph{moveProperty}.
+The \emph{setParent} transformation is described as example:
 \begin{align}
-& extractSuperClass(class, parentName, property*) \\
-  & extractCommonSuperClass(firstClass, secondClass, parentName,  property* ) \\
+&\Phi(setParent(childClass, parentClass)) = \nonumber \\
+&= \begin{cases}
+mergeTables(ORM(childClass), ORM(parentClass),\\
+addColumn(column("inheritanceType", DBSTRING, "parentClass", \\ \; \; NOTNULL) ORM(parentClass), \\
+     addColumn(column("inheritanceType", DBSTRING, "childClass", \\ NOTNULL) ORM(childClass), database))) \\
+     \mathbf{if} \; inheritanceType = SINGLETABLE 
+     \\\\
+     addForeignKey(foreignKey("parent", ORM(parentClass), \\
+     \; \; (NOTNULL, UNIQUE)), ORM(childClass), database) \\
+     \mathbf{if} \; inheritanceType = TABLE-PER-CLASS \\ \wedge  
+     isEmpty(ORM(childClass)) \wedge isEmpty(ORM(parentClass))
+ \end{cases}
 \end{align}
+The inverse transformation \emph{removeParent} moves all columns from the parent table to the child table if there are data in the parent table, then the parent can be removed securely.
+This does not cover the case when the information from parent are not needed in its child, such as transformation has to be defined in several steps.
+
+Next transformations connected with inheritance are \emph{pushDown} and \emph{pullUp}.
+Pull up moves a property from child to parent so the \emph{moveProperty} can be used, the mapping of instances is based on the parent-child relationship.
+It means the column cannot be constrained with $NOTNULL$ constraint if there are more siblings in the hierarchy.
+The \emph{pushDown} transformation works in opposite direction.
+The easiest situation is where there no sibling in the hierarchy, then the \emph{moveProperty} can be used, otherwise we assume there are no stored data in the siblings.
+It is because the change of parent changes affects the instances of its children, when a property is moved only into one children the information consistency is violated, therefore we forbid such transformation, because we cannot anticipate developers intents.
+
+The next two inheritance-related transformations are \emph{extractParent} and \emph{extractCommonParent} serves for extracting a parent class in based on given set of property from a class or from a couple of classes.
+The transformation is based on the \emph{extractClass} (and \emph{mergeClasses} in the second case).
 
@@
 The formal definition of MDD approach to database schema evolution is proposed in \cite{Aboulsamh:FormalModelingEvolution}, where a changes of database conceptual schema is interpreted on both physical schema and data.
-In contrast our proposal is aimed to solve the problem of code refactoring and its impact on relational database, next we propose more complex transformations and examine the impact of platform specific constructs (such as foreign keys or constraints) on the evolution.
+In contrast our proposal is aimed to solve the problem of code refactoring and its impact on relational database, next we propose more complex transformations (such as \emph{copyProperty} or inheritance-related transformations) and examine the impact of platform specific constructs (such as foreign keys or constraints) on the evolution.
 

### 2013-02-19 doc/paper.tex
added after acceptance info (as comment)

@@
 \usepackage[dvipdfm]{hyperref}
-
 \usepackage{amssymb}
@@
 \title{Refactoring of Application Using Relational Database}
+
 \author{Ondrej Macek\inst{1}}
@@
 \label{sec:intro}
-The evolution (change) of a software is a common issue during software development.
-It occurs from many reasons in all phases of software lifecycle.
-The evolution severity usually depends on number of changes which has to be made and on the number of affected software components.
+Evolution (change) of a software is a common issue during the software development.
+It occurs from many reasons in all phases of the software lifecycle.
+The evolution severity usually depends on number of changes which have to be made and on number of affected software components.
 Refactoring \cite{Fowler:Refactoring} is very popular practice in object-oriented environment for evolving the source code and software architecture.
-The evolution of database schema and stored data is implemented separately from  source code refactoring, although the change of application affects the database too.
-There exist object-relational mapping (ORM) frameworks, which can help with propagation of evolution from application to database.
-However these framework are usually not capable to solve complex refactoring cases nor migrate data properly as shown in Sect.
+Evolution of database schema and stored data is implemented separately from  source code refactoring, although the change of application affects the database too.
+Object-relational mapping (ORM) frameworks can help with propagation of the evolution from an application to a database.
+However these frameworks are usually not capable to solve complex refactoring cases nor migrate data properly as shown in Sect.
 \ref{sec:problem}.
@@
 \ref{sec:models}.
-There are introduced basic refactoring cases as well as the complex ones which are created as a sequences of the basic ones.
-The capabilities of proposed formal models are illustrated on the common refactoring issues in Sect.
+Introduced are basic refactoring cases as well as the complex ones which are created as sequences of the basic ones.
+Capabilities of the proposed formal models are illustrated in the common refactoring issues in Sect.
 \ref{sec:case}.
@@
 \label{sec:problem}
-A software implemented using an object-oriented language, which uses relational database as a data storage consists of four main components.
+A software implemented by using an object-oriented language, which uses relational database as a data storage, consists of four main components.
 There is the application itself, the database schema, stored data and the object-relational mapping.
-The refactoring affects not only the application but the database too as the database has to be evolved when the persistent layer of the application changed to fit the object-relational mapping used in the software.
+The refactoring affects not only the application but the database as well, as the database has to be evolved when the persistent layer of the application changed to fit the object-relational mapping used in the software.
 The common current solution is based on capabilities of object-relational mapping frameworks which are capable to create a database schema according to the given source code or model.
-The process of evolution then proceed in following steps:
+The process of evolution then proceeds in following steps:
 \begin{enumerate}
-	\item The code is refactored (usually using the developers IDE).
+	\item The code is refactored (usually by using the developers IDE).
 	\item The ORM framework generates a new database schema.
@@
 The last step is error-prone as it is processed manually and the prone to error increases with complexity of refactoring.
-The evolution process require cooperation of a developer and a database administrator or it requires the developer has a knowledge of the database used by the software.
+The evolution process requires cooperation of a developer and a database administrator or it requires the developer to have a knowledge of the database used in the software.
 The knowledge of the ORM is needed in both cases.
-Remarkable is the feasibility of data evolution has to be verified for each deployed software instance, because the data can differ.
-The data and information preservation is crucial issue of database evolution.
-Next observation is the evolution of the software is defined twice for one software - first for the application then for the database.
+Remarkable is the fact that the feasibility of data evolution has to be verified for each deployed software instance, because the data can differ.
+The data and information preservation is a crucial issue of database evolution.
+The next observation is the evolution of the software is defined twice for one software - first for the application then for the database.
 
-\paragraph{Example 1} Let us have only two classes \emph{A} and \emph{B} in the application which are not connected by an association and there are corresponding tables \emph{tab\_a} and \emph{tab\_b} in the database, which contain data.
+\paragraph{Example 1} Let us have only two classes \emph{A} and \emph{B} in the application which are not connected by an association and there are corresponding tables \emph{tab\_a} and \emph{tab\_b} in the database, which contain some data.
 We decide to inline \emph{B} into \emph{A}  during the development.
@@
 The data migration has to be created manually - the developer has to define the evolution twice.
-The mapping between data in \emph{tab\_a} and \emph{tab\_b} has to be provided to merge stored data correctly and impact of this mapping the database has to be verified: is there data which can be lost during inlining and is this lost intentional?
+The mapping between the data in \emph{tab\_a} and \emph{tab\_b} has to be provided to merge the stored data correctly and the impact of this mapping on the database has to be verified: are there any data which can be lost during inlining and is this lost intentional?
 
 \subsection{Solution Proposal}
-We believe there exists better solution which automatize the process of database evolution according to the code refactoring.
+We believe there exists a better solution which automatizes the process of database evolution according to the code refactoring.
 The solution is illustrated in the Fig.
 \ref{fig:evolution}.
-It is based on a change in the evolution process which assumes the ORM does not change during the evolution:
+It is based on a change in the evolution process which assumes that the ORM does not change during the evolution:
 \begin{enumerate}
-	\item The evolution of the whole software is defined independently on application or database.
-	\item The evolution is interpreted for application and database.
+	\item The evolution of the whole software is defined independently on the application or the database.
+	\item The evolution is interpreted for the application and the database.
 	\item The evolution is executed.
@@
 
-\paragraph{Copying Database Elements}The transformations for copying the structure of a column or table serves more as helpers for advanced evolution cases.
+\paragraph{Copying Database Elements} The transformations for copying the structure an values of a column or table serve more as helpers for advanced evolution cases, where they are discussed in detail.
+
+%\paragraph{Copying Database Elements}The transformations for copying the structure of a column or table serves more as helpers for advanced evolution cases.
 The \emph{copyColumn} transformation contains a mapping between instances (stored data).
@@
 The mapping is usually defined as:
-\begin{align}\label{eq:mapping}
-&	mapping : select(ORM(sourceName), database) \nonumber \\ 
-& \;\;\;\;\; \rightarrow  (select(ORM(targetName), database) 
-\end{align}
-The result of the mapping has to respect the constraints on column (e.g.
+%\begin{align}\label{eq:mapping}
+%&	mapping : select(ORM(sourceName), database) \nonumber \\ 
+%& \;\;\;\;\; \rightarrow  (select(ORM(targetName), database) 
+%\end{align}
+%The result of the mapping has to respect the constraints on column (e.g.
 $NULL$ cannot be inserted into a column with $NOTNULL$ constraint defined) The character of transformation depend on the feature of provided mapping.
@@
 \end{align}
+The \emph{copyProperty} transformation gets a mapping between instances (stored data) on imput.
+The mapping to provides information which value from the source table rows is copied to which row in target table.
+The mapping is usually defined as:
+\begin{align}\label{eq:mapping}
+&	mapping : select(ORM(sourceName), database) \nonumber \\ 
+& \;\;\;\;\; \rightarrow  (select(ORM(targetName), database) 
+\end{align}
+However different definitions can be used the only condition is the result of the mapping has to respect the structure of the instances and constraints on column (e.g.
+$NULL$ cannot be inserted into a column with $NOTNULL$ constraint defined) The character of transformation depend on the feature of provided mapping.
+There occurs no data lost if the mapping is surjective.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% For the final version of the paper: %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Author information
+%\vspace{4ex}\noindent
+%\textbf{Ondrej Macek} is\dots
+
+%% Reception and acceptance information
+%\bigskip
+%\paragraph{Received: Month DD, 20YY; Accepted: Month DD, 20YY.}
+
+%% Insert a blank page at the end if necessary, to keep the page count even
+%\newpage\mbox{}
+%\thispagestyle{empty}
+
 \end{document}

### 2013-02-22 doc/paper.tex
review from David

@@
 We believe there exists a better solution which automatizes the process of database evolution according to the code refactoring.
-The solution is illustrated in the Fig.
-\ref{fig:evolution}.
+The solution is illustrated in the Fig.~\ref{fig:evolution}.
 It is based on a change in the evolution process which assumes that the ORM does not change during the evolution:
@@
 The software has to be in consistent state so its users could have benefit from its usage.
-The software is in consistent state if application and database are consistent and the database structure corresponds to the application structure according to the ORM, therefore the the software is defined as:
+The software is in consistent state if application and database are consistent and the database structure corresponds to the application structure according to the ORM, therefore the software is defined as:
 \begin{align}
@@
 It has a unique name, one or more properties and it can be associated to other classes in the application.
-The properties of a class can be of two kinds - values or collections - according of theirs cardinality.
+The properties of a class can be of two kinds - values or collections - according to theirs cardinality.
 
@@
 \label{sec:app-evolution}
-The application model defines only the structure of an application thus there are defined transformation for its creating, altering and removing parts of the model (the set $A$).
+The application model defines only the structure of an application thus there are defined transformations for its creating, altering and removing parts of the model (the set $A$).
 The conditions for successful transformations are simple in case of creating or altering:
 \begin{enumerate}
-	\item The name collisions has to be prevent when creating an class in the model or altering its name.
+	\item The name collisions has to be prevented when creating an class in the model or altering its name.
 	\item All references have to be updated when renaming a class.
@@
 The database consists of two parts - from database schema which defines the structure and from stored data - hence the transformations from set $D$ have to consider both parts.
-The transformation for manipulation of structure has similar conditions to the evolution of application application.
+The transformation for manipulation of structure has similar conditions to the evolution of application.
 The transformations for data manipulation are inspired by the SQL language.
@@
 \end{align}
-The \emph{copyProperty} transformation gets a mapping between instances (stored data) on imput.
+The \emph{copyProperty} transformation gets a mapping between instances (stored data) on input.
 The mapping to provides information which value from the source table rows is copied to which row in target table.
@@
 However different definitions can be used the only condition is the result of the mapping has to respect the structure of the instances and constraints on column (e.g.
-$NULL$ cannot be inserted into a column with $NOTNULL$ constraint defined) The character of transformation depend on the feature of provided mapping.
+$NULL$ cannot be inserted into a column with $NOTNULL$ constraint defined) The character of transformation depends on the feature of provided mapping.
 There occurs no data lost if the mapping is surjective.
@@
 The $moveProperty$ transformation is based on the $copyProperty$ transformation followed by the $removeProperty$ so its construction is easy.
-On the other hand special attention has to be paid to the provided mapping of instances, because it can cause data loses.
+On the other hand special attention has to be paid to the provided mapping of instances, because it can cause lost of data.
 The ideal case is when the mapping is surjective, then the transformation can not cause the lost of data.
@@
 The \emph{pushDown} transformation works in opposite direction.
-The easiest situation is where there no sibling in the hierarchy, then the \emph{moveProperty} can be used, otherwise we assume there are no stored data in the siblings.
-It is because the change of parent changes affects the instances of its children, when a property is moved only into one children the information consistency is violated, therefore we forbid such transformation, because we cannot anticipate developers intents.
+The easiest situation is where there is no sibling in the hierarchy, then the \emph{moveProperty} can be used, otherwise we assume there are no stored data in the siblings.
+It is because the change of parent affects the instances of its children, when a property is moved only into one children the information consistency is violated, therefore we forbid such transformation, because we cannot anticipate developers intents.
 
-The next two inheritance-related transformations are \emph{extractParent} and \emph{extractCommonParent} serves for extracting a parent class in based on given set of property from a class or from a couple of classes.
+The next two inheritance-related transformations are \emph{extractParent} and \emph{extractCommonParent} which serve for extracting a parent class based on given set of property from a class or from a couple of classes.
 The transformation is based on the \emph{extractClass} (and \emph{mergeClasses} in the second case).
@@
 \end{figure}
-The first  step of is to extract two temporary classes representing address of a $Person$ or of a $LegalParty$:
+The first  step is to extract two temporary classes representing address of a $Person$ or of a $LegalParty$:
 \begin{align}
@@
 Temporary classes has to be connected with the origin classes by an association.
-The associations are creates as optional because foreign key constrained with $NOTNULL$ can cause problems in the next phase of evolution.
+The associations are created as optional because foreign key constrained with $NOTNULL$ can cause problems in the next phase of evolution.
 \begin{align}
@@
 It shows that usage of the transformations are sometimes not intuitive for user (e.g.
-creating temporary classes before merge), therefore next complex transformations has to be define based on user experience.
+creating temporary classes before merge), therefore next complex transformations has to be defined based on user experience.
 The examples shows that the column constraint s often decide the feasibility of a transformation.
@@
 
-The formal definition of MDD approach to database schema evolution is proposed in \cite{Aboulsamh:FormalModelingEvolution}, where a changes of database conceptual schema is interpreted on both physical schema and data.
+The formal definition of MDD approach to database schema evolution is proposed in \cite{Aboulsamh:FormalModelingEvolution}, where changes of database conceptual schema is interpreted on both physical schema and data.
 In contrast our proposal is aimed to solve the problem of code refactoring and its impact on relational database, next we propose more complex transformations (such as \emph{copyProperty} or inheritance-related transformations) and examine the impact of platform specific constructs (such as foreign keys or constraints) on the evolution.

### 2013-03-04 doc/ComSIS-typeinst.tex
text improvements

\documentclass[runningheads]{comsis}

\def\journalissue{ComSIS Vol.~0, No.~0, December 2009}
\def\paperidnum{UDC, DOI: N/A}
\setcounter{page}{1}

\usepackage[dvipdfm]{hyperref}
\usepackage{verbatim}
\usepackage{graphicx}

\title{Authors' Instructions for the Preparation\\of Camera-Ready Contributions with \LaTeX}
\titlerunning{Authors' Instructions}
\author{Vladan Deved\v{z}i\'c\inst{1} \and Marijana Despotovi\'c\inst{1} \and Ivan Lukovi\'c\inst{2} \and
        Violeta Damjanovi\'c\inst{3} \and Milo\v{s} Radovanovi\'c\inst{4}}
\authorrunning{Vladan Deved\v{z}i\'c et al.}
\institute{Faculty of Organizational Sciences, POB 52\\
  11000 Belgrade, Serbia\\
  \email{\{devedzic,maja\}@fon.rs}
  \and
  Faculty of Technical Sciences, Trg D.\ Obradovi\'ca 6\\
  21000 Novi Sad, Serbia\\
  \email{ivan@uns.ac.rs}
  \and
  Postal Savings Bank, 27.\ marta 71\\
  11000 Belgrade, Serbia\\
  \email{vdamjanovic@posted.co.rs}
  \and
  Dept.\ of Mathematics and Informatics, Trg D.\ Obradovi\'ca 4\\
  21000 Novi Sad, Serbia\\
  \email{radacha@dmi.uns.ac.rs}}

\begin{document}

\maketitle

\begin{abstract}
The abstract should summarize the contents of the paper and should contain at
least 70 and at most 150 words. It should be set using the \texttt{abstract}
environment.

\vspace{6pt}\textbf{Keywords:} computer science, information systems, \LaTeX,
typesetting.
\end{abstract}

\section{Introduction}

These instructions and the corresponding \LaTeX\ style are based
on the corresponding Springer instructions and \LaTeX\ document class for
preparing camera ready papers to be published in the Springer series Lecture
Notes in Computer Science.

The preparation of manuscripts which are to be reproduced by photo-offset
requires special care. Papers submitted in a technically unsuitable form will
be returned for retyping, or canceled if the volume cannot otherwise be
finished on~time.

\section{Manuscript Preparation}

You are encouraged to use \LaTeX\ for the preparation of your camera-ready
manuscript together with the corresponding document class comsis.cls (see
Sect.~\ref{sec:latex}), and send all necessary source files. Please follow
these instructions closely in order to make the volume look as uniform as
possible. Please also send a corresponding pdf file.

We would like to stress that the class/style files should not be manipulated
and that the guidelines regarding font sizes and format should be adhered to.
This is to ensure that the end product is as homogeneous as possible.

\subsection{Printing Area}

The printing area is 122mm $\times$ 193mm. The text should be justified to
occupy the full line width, so that the right margin is not ragged, with words
hyphenated as appropriate. The length of the text on a page should not be less
than~180mm.

\subsection{Layout, Typeface, Font Sizes, and Numbering}

The default font for all text formatting is Helvetica. Use 10-point type for
the name(s) of the author(s) and 9-point type for the address(es) and the
abstract. For the main text, please use 10-point type and single-line spacing.
Italic type may be used to emphasize words in running text. Bold type and
underlining should be avoided. With these sizes, the interline distance should
be such that some 45 lines occur on a full-text page.

\subsubsection{Headings.} Headings should be capitalized (i.e., nouns, verbs, and
all other words except articles, prepositions, and conjunctions should be set
with an initial capital) and should, with the exception of the title, be
aligned to the left. Words joined by a hyphen are subject to a special rule.
If the first word can stand alone, the second word should be capitalized.
Headings and appropriate \LaTeX\ commands are enumerated in
Table~\ref{tbl:headings}.

\begin{table}
\caption{Font sizes of headings. Table captions should always be positioned
\emph{above} the tables. Tables should be horizontally centered. The final
sentence of a table caption should end without a period}\label{tbl:headings}
\centering
\setlength\tabcolsep{0.5em}
\begin{tabular}{lll}
\hline
Heading level      & Example                                    & \LaTeX\ command \\ \hline \\[-2ex]
Title (centered)   & \textbf{\Large Authors' Instructions\dots} & \texttt{$\backslash$title} \\
1st-level heading  & \textbf{\large 1.\ Introduction}           & \texttt{$\backslash$section} \\
2nd-level heading  & \textbf{2.1.\ Printing Area}               & \texttt{$\backslash$subsection} \\
3rd-level heading  & \textbf{Headings.} Text follows\dots       & \texttt{$\backslash$subsubsection} \\
4th-level heading  & \textit{Remark.} Text follows\dots         & \texttt{$\backslash$paragraph} \\ \hline
\end{tabular}
\end{table}

Here are some examples of headings: ``Criteria to Disprove Context-Freeness of
Collage Languages'', ``On Correcting the Intrusion of Tracing
Non-deterministic Programs by Software'', ``A User-Friendly and Extendable
Data Distribution System'', ``Multi-flip Networks: Parallelizing GenSAT'',
``Self-determinations of Man''.

\subsubsection{Lemmas, Propositions, and Theorems.} The numbers accorded to lemmas,
propositions, and theorems etc.\ should appear in consecutive order, starting
with the number 1, and not, for example, with the number 11.

\subsection{Figures and Photographs}

Please produce your figures electronically, if possible, and refer to them
from your source file using, for example, the
\texttt{$\backslash$includegraphics} command from package \texttt{graphicx}.
Format the figures using the standard \texttt{figure} environment, and center
them horizontally.

Check that in line drawings, lines are not interrupted and have constant
width. Grids and details within the figures must be clearly readable and may
not be written one on top of the other. Line drawings should have a resolution
of at least 800dpi (preferably 1200dpi). For digital halftones 300dpi is
usually sufficient. The lettering in figures should have a height of 2mm
(10-point type). Figures should be scaled up or down accordingly. Please do
not use any absolute coordinates in figures. If possible, the files of figures
(e.g.\ eps files) should not contain binary data, but be saved in ASCII
format.

Figures should be numbered and should have a caption which should always be
positioned under the figures, in contrast to the caption belonging to a table,
which should always appear above the table. The final sentence of a caption,
be it for a table or a figure, should end without a period. The
\texttt{$\backslash$caption} command will center the captions between the
margins and set them in 9-point type (Fig.~\ref{fig:afigure} shows an
example). The distance between text and figure should be about~8mm, the
distance between figure and caption about~5mm.

\begin{figure}
\centering
\includegraphics[width=.7\textwidth]{figure}
\caption{One kernel at~$x_s$ (\emph{dotted kernel}) or two kernels at $x_i$ and~$x_j$
(\emph{left and right}) lead to the same summed estimate at~$x_s$. This shows
a figure consisting of different types of lines. Elements of the figure described
in the caption should be set in italics, in parentheses, as shown in this sample
caption. The last sentence of a figure caption should generally end without a period}\label{fig:afigure}
\end{figure}

\paragraph{Remark 1.} In the printed volumes, illustrations are generally black and
white (halftones), and only in exceptional cases, and if the author is
prepared to cover the extra cost for color reproduction, are color pictures
accepted. If color illustrations are necessary, please send us color-separated
files if possible. Color pictures are welcome in the electronic version at no
additional cost.

\paragraph{Remark 2.} To ensure that the reproduction of your illustrations is of
reasonable quality we advise against the use of shading. The contrast should
be as pronounced as possible. This particularly applies for screenshots.

\subsection{Formulas}

Displayed equations or formulas are centered and set on a separate line (with
an extra line or halfline space above and below). Displayed expressions should
be numbered for reference. The numbers should be consecutive within each
section or within the contribution, with numbers enclosed in parentheses and
set on the right margin. For example,
\begin{equation}
x + y = z\,.
\end{equation}

Please punctuate a displayed equation in the same way as ordinary text but
with a small space before the end punctuation.

\subsection{Program Code}

Program listings or program commands in the text are normally set in
typewriter font (Courier).

Example of a Computer Program from Jensen K., Wirth N. (1991) Pascal user
manual and report. Springer, New York:

\begin{verbatim}
  program Inflation (Output)
    {Assuming annual inflation rates of 7%, 8%, and
    10%,...   years};
    const  MaxYears = 10;
    var    Year: 0..MaxYears;
           Factor1, Factor2, Factor3: Real;
    begin
      Year := 0;
      Factor1 := 1.0; Factor2 := 1.0; Factor3 := 1.0;
      WriteLn('Year 7% 8% 10%'); WriteLn;
      repeat
        Year := Year + 1;
        Factor1 := Factor1 * 1.07;
        Factor2 := Factor2 * 1.08;
        Factor3 := Factor3 * 1.10;
        WriteLn(Year:5,Factor1:7:3,Factor2:7:3,
          Factor3:7:3)
      until Year = MaxYears
  end.
\end{verbatim}

\subsection{Footnotes}

The superscript numeral used to refer to a footnote appears in the text either
directly after the word to be discussed or -- in relation to a phrase or a
sentence -- following the punctuation sign (comma, semicolon, or period).
Footnotes are set at the bottom of the normal text area, with a line of
about~5cm immediately above them.\footnote{The footnote numeral is set flush
left and the text follows with the usual word spacing. Second and subsequent
lines are indented. Footnotes should end with a period.}

\subsection{Citations}

The list of references is headed ``References'' and is not assigned a number
in the decimal system of headings. The list should be set in small print and
placed at the end of your contribution, in front of the appendix, if one
exists. Please do not insert a pagebreak before the list of references if the
page is not completely filled. An example is given at the end of this
information sheet. For citations in the text please use the
\texttt{$\backslash$cite} command, and organize your references using Bib\TeX,
with \texttt{$\backslash$bibliographystyle\{splncs03\}}, producing numbered
citations in square brackets, such as: \cite{Agrawal94VLDB}, \cite{Enc97LNCS},
\cite{DatabaseBook02}, \dots

\subsection{Page Numbering and Running Heads}

Page numbers and running heads are automatically generated by \LaTeX. If the
title of your paper is too long for the header please specify a running title
using the \texttt{$\backslash$titlerunning} command. Other parameters will be
allocated by the publisher.

\subsection{Printing Quality}

You are normally not expected to send a printed copy of your manuscript if you
provide the source and pdf files matching exactly each other when printed.
Only in exceptional cases, you will be asked to send sheets which are printed
on one side only. In such a case, please use a high-resolution printer,
preferably a laser printer with at least 300dpi. We prefer the text to be
centered on the pages (i.e., equal margins left and right and top and bottom).
The format of the paper (A4, Letter, etc.) is irrelevant.

\section{Using \LaTeX}\label{sec:latex}

We encourage using \LaTeX\ for preparing final manuscripts. We provide the
document class comsis.cls to help \LaTeX\ users prepare their camera-ready
manuscript and to enable us to use their source files. The necessary style
files and documentation can be downloaded from the ComSIS Web page at
\url{http://www.comsis.org/information.php}.

\section{Supplementary Material}

If you wish to include color illustrations in the electronic version in place
of or in addition to any black and white illustrations in the printed version,
please provide the managing editor and the editorial assistant with the
appropriate files.

If you have supplementary material, e.g., executable files, video clips, or
audio recordings, on your server, simply send the managing editor and the
editorial assistant a short description of the supplementary material and
inform them of the URL at which it can be found. We will add the description
of the supplementary material to the online version of the corresponding
ComSIS volume and create a link to your server. Alternatively, if this
supplementary material is not to be updated at any stage, then it can be sent
directly to the managing editor and the editorial assistant, together with all
the other files.

\section{Copyright Form}

For details regarding the transfer of copyright, please see the ``Copyright
and Use Agreement'' section under ``Information for Contributors'' at the
ComSIS Web page. Generally, we ask contributing authors to complete and sign
the ``Transfer of Copyright'' agreement before the paper may be published. (It
is sufficient if one author from each contribution signs the form on behalf of
all the other authors.) The copyright form is located on our Web page at
\url{http://www.comsis.org/download/ComSISCopyright.doc}. The printed form should be
completed and signed and sent to the editorial assistant by normal mail.
Alternatively, a scanned copy of the form may be sent by e-mail.

\section{Checklist}

When submitting your camera-ready manuscript to the managing editor and the
editorial assistant, please make sure you include the following:
\begin{itemize}
  \item your source (input) files, including the figures;
  \item a single-sided printout (not a photocopy) of the final version of your contribution, if requested by the managing editor;
  \item any style files, templates, and special fonts you may have used;
  \item the completed and signed copyright form.
\end{itemize}

If supplementary material is available, please provide the editorial
assistant~with:
\begin{itemize}
  \item a short description of the supplementary material;
  \item the supplementary material or the URL at which it can be found;
  \item the files of color figures for the electronic version.
\end{itemize}

\nocite{Temporal97TDS,CS95,Ontology02}

\bibliographystyle{splncs03}
\bibliography{example}

\end{document}

### 2013-03-04 doc/paper.tex
text improvements

@@
 
-\title{Refactoring of Application Using Relational Database}
+\title{Application and Relational Database Co-Refactoring}
 
@@
 \begin{abstract}
-The change of application source code caused by refactoring affected not only the application, but it affect the stored data too if the persistent objects change.
+A change of application source code caused by refactoring affects not only the application, but it affects the stored data too if the persistent objects change.
 The change is usually processed in two steps: refactoring and data migration.
-We provide a solution capable to migrate database according to a refactoring in the application code.
-The feasibility of change and its data-secure processing is addressed too.
+We provide a solution capable to migrate database according to refactoring in the application code.
+The feasibility of the change and its data-secure processing is addressed too.
 
@@
 \subsection{Solution Proposal}
-We believe there exists a better solution which automatizes the process of database evolution according to the code refactoring.
+We believe that there exists a better solution which automatizes the process of database evolution according to the code refactoring.
 The solution is illustrated in the Fig.~\ref{fig:evolution}.
@@
 \end{enumerate}
-The process can decrease the human mistakes in the evolution process, because there is only one source for the evolution and the evolution is automatically interpreted for the application (as a refactoring) and database (as schema and data migration).
+The process can decrease human mistakes in the evolution process, because there is only one source for the evolution and the evolution is automatically interpreted for the application (as a refactoring) and database (as schema and data migration).
 The existence of the set of all possible software evolutions $E$ is based on a evolutionary transformations specific for application and database.
@@
 
-\paragraph{Example 2} The same situation from Example 1 can be solved more effectively if the \emph{inline} evolution is defined in $E$.
-All information needed for software evolution are provided as the input of the \emph{inline} transformation which is then automatically interpreted as inline of application classes and merge of tables on database level.
+\paragraph{Example 2} The same situation from Example 1 can be solved more effectively when the \emph{inline} evolution is defined in $E$.
+All information needed for software evolution is provided as the input of the \emph{inline} transformation which is then automatically interpreted as inline of application classes and merge of tables on database level.
 The mapping is provided as the input of \emph{inline} too, thus the script for data migration can be generated.
-The structural feasibility is verified during the interpretation of the \emph{inline} and the migration script contains conditions which verifies the feasibility on level of data.
+The structural feasibility is verified during the interpretation of the \emph{inline} and the migration script contains conditions which verifies the feasibility on the level of data.
 
@@
 \label{sec:models}
-The model-driven development (MDD) is a good approach to data evolution of various software components \cite{Moravec:PracticalApproach}.
-The software is represented by a set of models - application model and database model concretely.
+Model-driven development (MDD) is a good approach to data evolution of various software components \cite{Moravec:PracticalApproach}.
+The software is represented by a set of models -- an application model and a database model concretely.
 The evolution and ORM is represented by a set of model-to-model transformations and the interpretation of transformations from $E$ is a model-to-model transformation again.
@@
 \end{equation}
-The evolution of the software is defined as the interpretation of the transformation for each component of the software:
+Evolution of the software is defined as an interpretation of the transformation for each component of the software:
 \begin{align}
@@
 The software consists of three important components (as seen in Fig.
-\ref{fig:evolution}: application (its persistent layer concretely), database (consisting of database schema and stored data) and an ORM, therefore we define a software as a triple consisting of an application and a database which are connected together by an ORM.
-The software has to be in consistent state so its users could have benefit from its usage.
-The software is in consistent state if application and database are consistent and the database structure corresponds to the application structure according to the ORM, therefore the software is defined as:
+\ref{fig:evolution}: an application (its persistent layer concretely), a database (consisting of a database schema and stored data) and an ORM, therefore we define a software as a triple consisting of an application and a database which are connected together by an ORM.
+The software has to be in consistent state so its users could benefit from its usage.
+The software is in a consistent state if the application and the database are consistent and the database structure corresponds to the application structure according to the ORM, therefore the software is defined as:
 \begin{align}
@@
 \end{align}
-The evolution of a software is a transformation from one consistent state to another one.
-These states are called generations of the software and the function which changed the state of a software are called transformations.
+Evolution of a software is a transformation from one consistent state to another one.
+These states are called generations of the software and the function which changed the state of the software are called transformations.
 
@@
 \subsection{Application Model}
-Application is defined as a sequence of classes and it creates the context for all structures used in the software persistent layer:
+An application is defined as a sequence of classes and it creates the context for all structures used in the software persistent layer:
 \begin{align}
@@
 It has a unique name, one or more properties and it can be associated to other classes in the application.
-The properties of a class can be of two kinds - values or collections - according to theirs cardinality.
+The properties of a class can be of two kinds - values or collections - according to their cardinality.
 
 \paragraph{Property} Property represents a feature of a class which is represented as a primitive type.
-The property can be mandatory, can have a default value and according to its cardinality it can represent a single value or a collection of values.
+A property can be mandatory, can have a default value and according to its cardinality it can represent a single value or a collection of values.
 
-\paragraph {Association} Association represents a connection between two classes.
+\paragraph{Association} Association represents a connection between two classes.
 It has a unique name and the reference is represented by the label of referenced class.
@@
 The cardinalities defines the multiplicity of both association ends.
+
+\paragraph{Inheritance} 
+%TODO Inheritance
+Multiple ancestors are not allowed and of course a class can not be its own ancestor cyclically. We also assume there is only one type of inheritance type in the class hierarchy.
 
@@
 
- There is not allowed multiple ancestors and of course a class can not be its own ancestor cyclically. We also assume there is only one type of inheritance type in the class hierarchy.
-
 \subsection{Application Manipulation}
@@
 \begin{enumerate}
-	\item The name collisions has to be prevented when creating an class in the model or altering its name.
+	\item The name collisions have to be prevented when creating a class in the model or altering its name.
 	\item All references have to be updated when renaming a class.
@@
 \end{enumerate}
-The only issue during deleting classes etc.
-from an application is the class can be removed only in case it is not associated by other classes.
-The transformations from $A$ are in table \ref{tab:app-evolution}.
-The list of altering operations is not complete as there are missing transformation for changing the obligation or cardinality of properties and associations.
+The only problem during deleting classes etc.
+from an application is  that the class can be removed only in case it is not associated by other classes.
+The transformations from $A$ are in Table \ref{tab:app-evolution}.
+The list of altering operations is not complete as transformation for changing the obligation or cardinality of properties and associations are missing.
 It is because these transformations are not so important for complex refactorings.
@@
 %		& copyClass(class, newName, application) \\
-	Inheritance manipulation   & setParent(class, parentClass) \\
+	Inheritance manipulation   & addParent(class, parentClass) \\
 		& removeParent(childClass) \\
@@
 \subsection{Database Model}
-The relational database consists of database schema which defines the structure of the database and data, which in the software represents stored instances. The database is defined as:
+A relational database consists of a database schema which defines the structure of the database and data, which represents stored instances in the software. A database is defined as:
 \begin{align}
@@
 
-\paragraph{Table} Table represents a basic concept of database schema.
+\paragraph{Table} Table represents a basic concept of a database schema.
 It has a unique name, one or more columns and it can be related to other tables in the schema by foreign keys.
@@
 The sequence generates values of primary key.
-The new value of a key is obtained by calling the function $next(s)$.
-The sequence which provides numbers incremented by one is called $s_1$ if the increment is 1, $s_{10}$ if the increment is 10 etc.
+A new value of a key is obtained by calling the function $next(s)$.
+The sequence which provides numbers incremented by one is called $s_1$ when the increment is 1, $s_{10}$ when the increment is 10 etc.
 
@@
 
-\paragraph{Data} A database consists not only from schema but also from data which are represented as rows in a table.
-A table row in our model is a tuple consisting of value pairs, which represents concrete value of concrete column or key.
+\paragraph{Data} A database consists not only of a schema but also of data which are represented as rows in a table.
+A table row in our model is a tuple consisting of value pairs, which represent concrete value of a concrete column or key.
 Each row contains a name of a table it belongs to.
@@
 \label{sec:db-evolution}
-The database consists of two parts - from database schema which defines the structure and from stored data - hence the transformations from set $D$ have to consider both parts.
+A database consists of two parts - of a database schema, which defines the structure, and of stored data - hence the transformations from set $D$ have to consider both parts.
 The transformation for manipulation of structure has similar conditions to the evolution of application.
@@
 The basic transformations are in Table \ref{tab:db-basic-evolution}.
-The set of transformations for database manipulation is limited in contrast with the SQL language as we mentioned only transformation necessary for data evolution on database level.
+The set of transformations for database manipulation is limited in contrast to the SQL language.
+Transformations only necessary for data evolution on database level are introduced.
 
 \begin{table}
-\caption{The transformations for database evolution. The transformations can be mapped to SQL intuitevly.}
+\caption{The transformations for database evolution. Transformations can be mapped to SQL intuitively.}
 	\label{tab:db-basic-evolution}
@@
 
-\paragraph{Copying Database Elements} The transformations for copying the structure an values of a column or table serve more as helpers for advanced evolution cases, where they are discussed in detail.
+\paragraph{Copying Database Elements} The transformations for copying the structure and values of a column or table serve more as helpers for advanced evolution cases, where they are discussed in detail.
 
@@
 
-\paragraph{Data-secure Database Elements Removal} The transformation for remove elements from the database can have fatal impact on the data preservation, therefore the set of transformations is extended by a data secure transformations for removing database elements.
-These transformations are not part of the SQL standard, although they can be implemented as database functions.
-These transformations creates a secure way to remove elements from the database as they drops empty structural elements only.
+\paragraph{Data-secure Database Elements Removal} The transformation for remove elements from the database can have fatal impact on the data preservation, therefore the set of transformations is extended by data secure transformations for removing database elements.
+These transformations are not a part of the SQL standard, although they can be implemented as database functions.
+These transformations create a secure way to remove elements from the database as they drop empty structural elements only.
 
 \subsection{Object Relational Mapping}
-The ORM is the only fixed point in the software model we use.
-The mapping is similar to the Hibernate mapping \cite{Hibernate} , thus a lot of developers should be familiar with it.
-The main ideas are:
+ORM is the only fixed point in the software model we use. The mapping is similar to the Hibernate mapping \cite{Hibernate}, thus a lot of developers should be familiar with it. The main ideas are: 
 \begin{itemize}
 	\item classes are mapped to tables,
-	\item single properties are mapped to column or if the property is collection then the property is mapped to a table,
+	\item single properties are mapped to a column or if the property is a collection then the property is mapped to a table,
 	\item associations are mapped to foreign keys or to tables if the association represents many-to-many  relationship,
-	\item the primary keys are created automatically for each table,
-	\item names used on application are mapped into the database schema (e.g. because possible name collision of application classes and names of database schema elements) 
+	\item a primary keys are created automatically for each table,
+	\item names used in application are mapped into the database schema (e.g. because of possible name collision of application classes and names of database schema elements) 
 \end{itemize}
@@
 \ref{sec:app-evolution} and \ref{sec:db-evolution}.
-This section introduces how these primitives can be used to manipulate the software elements.
+This section introduces how those primitives can be used to manipulate the software elements.
 
 The basic evolutionary transformations of the software are based on basic evolution of an application.
-These transformations has to respect the ORM, because properties and associations can be mapped as a column (foreign key respectively) as shown on the example of creating a new property:
+These transformations has to respect the ORM, because properties and associations can be mapped as a column (foreign key respectively) as shown in the example of creating a new property:
 \begin{align}
@@
 
-Mapping between basic evolutionary transformations is provided in the shortened way in table \ref{tab:sw-basic-evolution}, there is the basic application transformation and possible mappings on database level according to to the cardinality.
+Mapping between basic evolutionary transformations is provided in the shortened way in Table \ref{tab:sw-basic-evolution}.
+There is the basic application transformation and possible mappings on database level according to the cardinality.
 \begin{table}
@@
 This means the complex transformations are limited in the same way as their basic components.
-All transformations from the set $E$ presented so far has the same input information for both application and database, whereas the advanced transformations usually needs a mapping between stored data (instances) as theirs input.
+All transformations from the set $E$ presented so far have the same input information for both an application and a database, whereas the advanced transformations usually need a mapping between stored data (instances) as their input.
 
@@
 The $copyProperty$ creates a duplicate of a property in a given class.
-It can rename the property so it can be used for creating duplicates of a property in the property owning class.
-If the mapping is not provided the transformation creates only a structural copy of the property otherwise it copies the values too.
-Therefore the  $copyProperty$ transformation is the simplest way to manipulate stored data.
+It can rename the property so it can be used for creating of duplicates of a property in the property owning class.
+If the mapping is not provided the transformation creates only a structural copy of the property, otherwise it copies the values too.
+Therefore the  $copyProperty$ transformation is the simplest way to manipulate the stored data.
 \begin{align}
@@
 The  $copyProperty$ is the first transformation when an additional information has to be added to the usual code refactoring.
-It is because the $copyColumn$ transformation needs one more information to succeed - there has to be provided mapping between the source and target table to assure the data information consistency.
+It is because the $copyColumn$ transformation needs one more information to succeed - there has to be provided mapping between the source and the target table to assure the data information consistency.
 %Moreover the source and target should have the same structure, type and constraints.
@@
 \end{align}
-in case of database the copy of a column or table is created according to property's cardinality and then values are copied:
+In case of database the copy of a column or table is created according to property's cardinality and then values are copied:
 \begin{align}
@@
 The \emph{copyProperty} transformation gets a mapping between instances (stored data) on input.
-The mapping to provides information which value from the source table rows is copied to which row in target table.
+The mapping provides information which value from the source table rows is copied to which row in the target table.
 The mapping is usually defined as:
@@
 \end{align}
-However different definitions can be used the only condition is the result of the mapping has to respect the structure of the instances and constraints on column (e.g.
-$NULL$ cannot be inserted into a column with $NOTNULL$ constraint defined) The character of transformation depends on the feature of provided mapping.
+However, different definitions can be used, the only condition is the result of the mapping has to respect the structure of the instances and constraints on column (e.g.
+$NULL$ cannot be inserted into a column with $NOTNULL$ constraint defined).
+The character of transformation depends on the feature of provided mapping.
 There occurs no data lost if the mapping is surjective.
@@
 The $moveProperty$ transformation is based on the $copyProperty$ transformation followed by the $removeProperty$ so its construction is easy.
-On the other hand special attention has to be paid to the provided mapping of instances, because it can cause lost of data.
-The ideal case is when the mapping is surjective, then the transformation can not cause the lost of data.
+On the other hand, special attention has to be paid to the provided mapping of instances, because it can cause lost of data.
+The ideal case is when the mapping is surjective, then the transformation can not cause loss of data.
 %On the other hand the situation when the mapping does not provides pairs for all rows in a source table can lead to data loss.
@@
 Extract and inline class are two opposite transformations.
-First of them extract a new class from an existing class, second of them moves data from one class into another.
-Both transformations can be composed from already mentioned basic trasnformations:
+First of them extracts a new class from an existing class, second of them moves data from one class into another.
+Both transformations can be composed from already mentioned basic transformations:
 \begin{align}
@@
 \end{align}
-The \emph{mapping} cannot be created according to the definition \ref{eq:mapping}, because there are no stored data in the table representing the \emph{newClass}, therefore the \emph{mapping} is defined as identity in case of \emph{extract}, therefore the rows in the new table has same primaryKey value as in their source class.
-The \emph{extract} transformations assumes the \emph{sourceClass} and \emph{newClass} has the same type of primary keys sequence.
+The \emph{mapping} cannot be created according to the Definition \ref{eq:mapping}, because there are no stored data in the table representing the \emph{newClass}, therefore the \emph{mapping} is defined as identity in case of \emph{extract}, therefore the rows in the new table has same primaryKey value as in their source class.
+The \emph{extract} transformations assumes the \emph{sourceClass} and \emph{newClass} have the same type of primary keys sequence.
 \begin{align}
